### 2021-02-01 README.Rmd
set up repo with rrtools

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# DeadwaterPaper

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/BiomarkABS/DeadwaterPaper/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): R Markdown source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered version,
    `paper.docx`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: data](/analysis/data): Data used in the analysis.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder:
    supplementary-materials](/analysis/supplementary-materials):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.

## How to run in your broswer or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

The simplest way to explore the text, code and data is to click on
[binder](https://mybinder.org/v2/gh/BiomarkABS/DeadwaterPaper/master?urlpath=rstudio)
to open an instance of RStudio in your browser, which will have the
compendium files ready to work with. Binder uses rocker-project.org
Docker images to ensure a consistent and reproducible computational
environment. These Docker images can also be used locally.

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
[DESCRIPTION](/DESCRIPTION) file).
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.Rmd")` in the R console

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2021-02-01 README.Rmd
updated README.Rmd

@@
 # Please put your title here to include it in the file below.
-Title <- "Title of your paper goes here"
+Title <- "Assessment of Northern Pikeminnow Abundance in Deadwater Slough and Potential Impacts to Local Chinook Salmon"
+Authors <- "See, Kevin E., Nick Porter, Richard A. Carmichael, and Michael W. Ackerman"
+yr <- 2021
+Journal <- "TBD"
 ```
@@
 
-> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>
+> `r Authors` (`r yr`). _`r Title`_. `r Journal` <https://doi.org/xxx/xxx>
 
@@
 
-> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+> `r Authors` (`r yr`). _`r Title`_. `r Journal`, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
 
@@
 
-> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+> `r Authors`, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
 

### 2021-02-01 analysis/paper/paper.Rmd
added Richie's outline

---
title: "Title Goes Here"
author:
  - Myfirstname Mylastname:
      email: fl@oneplace.edu
      institute: [UofO]
      correspondence: true
  - Yourfirstname Yourlastname:
      email: fl@another.edu
      institute: [UofA]
      correspondence: false
institute:
  - UofO: University of One Place
  - UofA: University of Another Place
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Although well documented through the Columbia and Snake River hydropower system, little is known about movement and survival of Salmon River juvenile Chinook salmon through their migratory corridors above the hydropower system. With increases in water temperatures, predator abundance, and predator distribution, this knowledge is evermore necessary to characterize limiting factors and recovery potential of this imperiled species. To further understanding of predation impacts on juvenile Chinook Salmon in the Salmon River, specifically, the anthropogenically altered Deadwater Reach, we estimated the population of predatory species, calculated the bioenergetic requirement of predator species to maintain size, and estimated the number of juveniles predated upon within said reach. We then used previously established smolt-to-adult return ratios to understand the impact that predation within the Deadwater Reach has on adult returns to the Lemhi River of Idaho, with broader implications for anadromous populations outside of the Lemhi River specifically. 
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

<!-- The actual document text starts here: -->

# Introduction

* Lemhi River supports ESA listed Chinook salmon, steelhead and bull trout, making it a prime candidate for protection/recovery.
    * ESA listed species helps motivate the research, why it is important to study not only the Lemhi fish, but also the corridor that impacts all fish upstream of the Deadwater Reach.
    * Heavy anthropogenic influence, but little to no hatchery influence or supplementation.
* Little is known about the movements and survival of juveniles as they exit the Lemhi River.
    * Why is this? No PIT tag infrastructure, remote hard to access areas, large river system.
* A lot is known about movement and survival through the hydropower system, however little is known about survival through sections of the salmon river (Axel et al Sockeye Study, Copeland et al NRR DRR).
    * Discuss the two life histories NRR vs DRR?
*	A previous study showed that the lowest transitional probabilities of Red Fish Lake Sockeye occurred in the Dead Water reach.
    * Information about the dead water reach. Where is it? What is it? Any publications that mention it?
* Paragraph on pike minnow and small mouth bass? populations in the Columbia and Salmon River and implications of predation. 
    * Are there any studies that have assess predation, movement, populations of predatory fish (piscivorous) within the Columbia, Snake, or Salmon River?
* Any other background information to help motivate the study that I might be missing?
* Paragraph on what we did to address the knowledge gap.
    * Pike minnow population estimate. Bioenergetics model. Utilized previously established SARs to answer the questions about how this anthropogenically altered reach is impacting the recovery of adults back to the Lemhi and overall on the system.

Here is a citation [@Marwick2017]

# Methods

* Briefly touch on radio telemetry methods and cite the RT publication if it is out at this point? (might not be) or that goes in intro and discussion.
* Hook and line sampling mark recapture to estimate predator population in November 2019, April 2020, September/October 2020 (these dates may be wrong).
    * Provide calculations for total pop abundance estimate with uncertainty.
* Conducted gastric lavage sampling at multiple occasions to try and estimate the percent of the diet that is juvenile chinook salmon. 
    * Wet weight after dividing by prey type 
    *	Are there any reference in the literature about this? Or any other studies? 
*	Water Temperature measurements.
    *	Hobo tidbit placed at site. 
    *	Calculated daily average water temperature for 365 days of the year.
*	Developed a bioenergetics model for the average sized Pike Minnow observed to maintain body weight through the 365 days modelled. 
    *	Incorporated daily temperature time step.
    * Calculated total biomass consumed to maintain weight for an individual fish across 365 days.
    *	Multiplied by the total population estimates. (this may vary with time, but not sure yet).
*	Adult estimates.
    *	Calculated total mass consumed based on above. Multiplied fraction of diet that is chinook by the total mass consumed. 
    *	Then we have to divide by average mass of a fish from the Lemhi screw trap? Smolt, pre-smolt, average?
*	Used SARs (published or ours?) to then estimate the number of adults that may return if predation is reduced. 
    *	Do we want to say predation is reduced to 0, or maybe a range of reduction?
*	Anything else?

# Results
*	Total population estimates of Pike Minnow. Chart or table, but likely a chart.
*	Gastric lavage sampling results. What percent of the diet is chinook salmon from each sampling event. Pie chart? Mike loves pie charts.
*	Water temperature plot for the 365 average.
*	Bioenergetics results. Total mass consumed over a 365 day period. Plot of daily consumption with day of year on x and mass on y.
*	SAR results. Maybe a plot of multiple scenarios depending on gastric lavage results? And pop estimates if they vary in time.
*	Anything else?

<!-- Here's some example analysis code: -->

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r demo-plot, fig.cap="A plot of random numbers"}
plot(rnorm(10))
```

Figure \@ref(fig:demo-plot) shows how we can have a caption and cross-reference for a plot

```{r demo-inline-code}
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion
*	Here we will need to discuss for each result above: 
    *	How do our results compare to other sources within the literature?
    *	What are the broader implications of each result?
*	Management implications and how these results, analysis, modelling framework could be applied elsewhere?

## Conclusion
*	What are the most important results?
    *	Adult returns
    *	Pop estimate of pike minnow
    *	Total predation on juveniles
*	We have also presented a novel modeling framework for estimating predation on native, critically endangered anadromous species which can be applied to other areas of interest. John day? Others?
*	Also, this is an anthropogenically created dead water area, which could be a candidate for restoration (removing the impoundment). Is there anything in the literature that discusses habitat preferences for pike minnow? If we speed up velocities and add some cover potentially pike minnow predation success will be lowered. Will these fish just move elsewhere? 
*	The end.

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

<div id="refs"></div>

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2021-06-11 README.Rmd
updates to README.Rmd

@@
 # Please put your title here to include it in the file below.
-Title <- "Assessment of Northern Pikeminnow Abundance in Deadwater Slough and Potential Impacts to Local Chinook Salmon"
-Authors <- "See, Kevin E., Nick Porter, Richard A. Carmichael, and Michael W. Ackerman"
+Title <- "Northern Pikeminnow Abundance in Deadwater Slough and Potential Impacts to Local Chinook Salmon Populations"
+Authors <- "Carmichael, Richard A., Kevin E. See, Nick Porter, and Michael W. Ackerman"
 yr <- 2021
 Journal <- "TBD"
+
 ```
@@
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/BiomarkABS/DeadwaterPaper/master?urlpath=rstudio)
+[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/BiomarkABS/DeadwaterPaper/main?urlpath=rstudio)
 
@@
 
-## How to run in your broswer or download and run locally
+## How to run in your browser or download and run locally
 

### 2021-06-11 README.Rmd
master -> main in README.Rmd

@@
 The simplest way to explore the text, code and data is to click on
-[binder](https://mybinder.org/v2/gh/BiomarkABS/DeadwaterPaper/master?urlpath=rstudio)
+[binder](https://mybinder.org/v2/gh/BiomarkABS/DeadwaterPaper/main?urlpath=rstudio)
 to open an instance of RStudio in your browser, which will have the
@@
 You can download the compendium as a zip from from this URL:
-[master.zip](/archive/master.zip). After unzipping:
+[main.zip](/archive/main.zip). After unzipping:
 - open the `.Rproj` file in RStudio

### 2021-06-11 analysis/paper/paper.Rmd
cleaning up paper.Rmd

@@
 ---
-title: "Title Goes Here"
+title: "Northern Pikeminnow Abundance in Deadwater Slough and Potential Impacts to Local Chinook Salmon Populations"
 author:
-  - Myfirstname Mylastname:
-      email: fl@oneplace.edu
-      institute: [UofO]
+  - Richard A. Carmichael:
+      email: Richard.Carmichael@merck.com
+      institute: [biomark]
       correspondence: true
-  - Yourfirstname Yourlastname:
-      email: fl@another.edu
-      institute: [UofA]
+  - Kevin E. See:
+      email: Kevin.See@merck.com
+      institute: [biomark]
+      correspondence: false
+  - Nicholas J. Porter:
+      email: Nicholas.Porter@merck.com
+      institute: [biomark]
+      correspondence: false
+  - Michael Hall:
+      email: Mike.Hall1@merck.com
+      institute: [biomark]
+      correspondence: false
+  - Michael W. Ackerman:
+      email: Mike.Ackerman@merck.com
+      institute: [biomark]
       correspondence: false
 institute:
-  - UofO: University of One Place
-  - UofA: University of Another Place
+  - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
       reference_docx: "../templates/template.docx" # Insert path for the DOCX file
+      number_sections: FALSE
+      toc: no
+      always_allow_html: true
+      fig_width: 6
       pandoc_args:
@@
       - --lua-filter=../templates/pagebreak.lua
-bibliography: references.bib
-csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
+fontsize: 12pt
+mainfont: Times New Roman
+bibliography: 
+- references.bib
+- packages.bib
+csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
 abstract: |
@@
 keywords: |
-  keyword 1; keyword 2; keyword 3
+  northern pikeminnow; Chinook salmon; predation; mark-recapture; bioenergetics
 highlights: |
@@
 
-<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->
-
-<!-- With the following code you can access and display values from the yml header above. -->
-
 Keywords: `r rmarkdown::metadata$keywords`
-
-Highlights: `r rmarkdown::metadata$highlights`
-
-<!-- The following code chunk defines some general settings how code chunks should behave. -->
 
@@
 ```
-
-<!-- The actual document text starts here: -->
 
@@
 Utilized previously established SARs to answer the questions about how this anthropogenically altered reach is impacting the recovery of adults back to the Lemhi and overall on the system.
-
-Here is a citation [@Marwick2017]
 
@@
 # Results
+
 *	Total population estimates of Pike Minnow. Chart or table, but likely a chart.
@@
 
-<!-- Here's some example analysis code: -->
+# Discussion
 
-```{r get-data, eval = FALSE}
-# Note the path that we need to use to access our data files when rendering this document
-my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
-```
-
-```{r demo-plot, fig.cap="A plot of random numbers"}
-plot(rnorm(10))
-```
-
-Figure \@ref(fig:demo-plot) shows how we can have a caption and cross-reference for a plot
-
-```{r demo-inline-code}
-x <- round(pi, 2)
-```
-
-Here is an example of inline code `r x` in the middle of a sentence.
-
-# Discussion
 *	Here we will need to discuss for each result above: 
@@
 ## Conclusion
+
 *	What are the most important results?
@@
 # Acknowledgements
-
-<!-- The following line inserts a page break  -->
 

### 2021-06-11 analysis/paper/paper.Rmd
hashing out the paper outline

@@
 
-* Lemhi River supports ESA listed Chinook salmon, steelhead and bull trout, making it a prime candidate for protection/recovery.
-    * ESA listed species helps motivate the research, why it is important to study not only the Lemhi fish, but also the corridor that impacts all fish upstream of the Deadwater Reach.
-    * Heavy anthropogenic influence, but little to no hatchery influence or supplementation.
-* Little is known about the movements and survival of juveniles as they exit the Lemhi River.
-    * Why is this? No PIT tag infrastructure, remote hard to access areas, large river system.
-* A lot is known about movement and survival through the hydropower system, however little is known about survival through sections of the salmon river (Axel et al Sockeye Study, Copeland et al NRR DRR).
-    * Discuss the two life histories NRR vs DRR?
-*	A previous study showed that the lowest transitional probabilities of Red Fish Lake Sockeye occurred in the Dead Water reach.
-    * Information about the dead water reach. Where is it? What is it? Any publications that mention it?
-* Paragraph on pike minnow and small mouth bass? populations in the Columbia and Salmon River and implications of predation. 
-    * Are there any studies that have assess predation, movement, populations of predatory fish (piscivorous) within the Columbia, Snake, or Salmon River?
-* Any other background information to help motivate the study that I might be missing?
-* Paragraph on what we did to address the knowledge gap.
-    * Pike minnow population estimate.
-Bioenergetics model.
-Utilized previously established SARs to answer the questions about how this anthropogenically altered reach is impacting the recovery of adults back to the Lemhi and overall on the system.
+***Paragraph 1***
+
+* The Lemhi River supports an ESA-listed Chinook salmon population with substantial cultural and recreational value and is a candidate for protection and/or recovery.
+* The population has become depleted with lower adult returns in recent years.
+* Decreased returns and/or increased mortality of the population have often been attributed to (in chronological order by life stage):
+  - Degraded natal tributary habitat (e.g., high amount of agriculture)
+  - Reduced survival in the mainstem corridor e.g., Snake and Columbia rivers
+  - Poor ocean conditions
+  - Others worth mentioning? Perhaps adult harvest in ocean and/or in-river returns?
+
+***Paragraph 2***
+
+* One potentially under-appreciated source of mortality on juvenile Chinook salmon is predation from piscivorous fishes.
+  - At this point, I'm thinking leave avian predation out of the equation, but instead include in Discussion
+  - and keep focus on the freshwater environment
+* What is known about downriver (Columbia/Snake) piscivorous fish populations and potential predation on juvenile Chinook salmon?
+  - literature review of northern pikeminnow, smallmouth bass, other? predation on juvenile Chinook salmon
+  - still stay away from avian predation
+
+***Paragraph 3***
+
+* Summary of findings from Axel et al.
+and early RT reports and papers (Ackerman et al.
+2019, Porter et al.
+2020, 2021, See et al.
+RT manuscript) in Deadwater slough.
+Recent studies have demonstrated decreased movement and survival of juvenile salmonids emigrating through the Deadwater Slough.
+Axel et al.
+(2015) demonstrated decreased rates of emigration and survival for sockeye salmon *Oncorhynchus nerka* from Redfish Lake emigrating during the spring and recent winter telemetry studies have indicated decreased transition probabilities (approximately 10% less than nearby reaches) of juvenile Chinook salmon *Oncorhynchus tschawytscha* through Deadwater Slough during fall and winter months (Ackerman et al.
+2018; Porter et al.
+2019).
+* Axel et al.
+(2015) found increased mortality in a particular reach spanning location x to location y in the mainstem Salmon River on a study of juvenile (hatchery?) sockeye salmon released near Redfish Lake and actively migrating downriver.
+  - in their case, these were actively migrating juveniles, and so transitional probabilities can likely be interpreted (almost) directly as survival
+  - A previous study showed that the lowest transitional probabilities of Red Fish Lake Sockeye occurred in the Dead Water reach.
+* Ackerman et al. and Porter et al. found reduced (apparent) survival in a reach spanning near the town of North Fork to just downstream of Deadwater Slough.
+  - in these cases, there's a chance that juvenile Chinook could exhibit concealment behavior prior to winter and so transition probability is interpreted as the probability of a fish surviving AND moving
+* Anecdotally, locals (and others? maybe NOAA?) informed us that northern pikeminno and perhaps other species are prevalent in Deadwater Slough.
+
+***Paragraph 4***
+
+* Chinook salmon in the Lemhi River exhibit 2 juvenile migration tactics (site Copeland and Venditti [2014] and See et al. *in prep*).
+  - downstream rearing (DSR)
+  - natal reach rearing (NRR)
+* Define these juvenile migration tactics for stream-type Chinook salmon
+
+***Paragraph 5***
+
+* Some concern arose that predation at or near Deadwater Slough was *not* inconsequential to the Lemhi River Chinook salmon (and adjacent) population(s), impacting both DSR and NRR emigrants
+* We hypothesize that decreases in survival (or apparent survival) of juvenile Chinook salmon (and sockeye?) are largely, or at least partially, due to increased densities of piscivorous predators in the Deadwater Slough area resulting from anthropogenically impacted habitat after the Dump Creek alluvial fan was formed.
+* Beginning in 2019, we set out to evaluate fish predator populations in the Deadwater Slough and their potential predation upon and impacts to juvenile emigrants from the Lemhi River population, both DSR and NRR.
+* Our objectives are four-fold:
+  1. Estimate the abundance of fish predators in Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations from the Lemhi River.
+  2. Document (and perhaps quantify) predation on juvenile Chinook salmon during the sampling/emigration periods.
+  3. Using a bioenergetics approach, attempt to estimate the number of juvenile Chinook salmon consumed during some defined fall and/or spring sampling/emigration period.
+  4. After making some assumptions (e.g., on SARs) attempt to estimate how results from objective #3 could translate into impacts on adult returns (via a thought experiment or simulation approach)
+
+* Additionally, Deadwater Slough occurs in a reach of the Salmon River believed to be important for overwinter rearing of presmolts attempting to exhibit concealment behavior to survive winter prior to spring emigration to the ocean.
+Deadwater Slough further contains little to no hydrological or structure features (i.e., fish cover) allowing juvenile salmonids to be predated upon without being able to seek adequate refuge.
+This is also an extremely important reach and relevant study to the entirety of the upper Salmon River anadromous population(s), as each emigrating fish must pass through the Deadwater area.
 
@@
 
-* Briefly touch on radio telemetry methods and cite the RT publication if it is out at this point? (might not be) or that goes in intro and discussion.
-* Hook and line sampling mark recapture to estimate predator population in November 2019, April 2020, September/October 2020 (these dates may be wrong).
-    * Provide calculations for total pop abundance estimate with uncertainty.
-* Conducted gastric lavage sampling at multiple occasions to try and estimate the percent of the diet that is juvenile chinook salmon. 
-    * Wet weight after dividing by prey type 
-    *	Are there any reference in the literature about this? Or any other studies? 
+## Study Site
+
+<!-- The following was pulled directly from a technical report, but probably could/should be improved for the manuscript -->
+
+The Deadwater Slough is an approximately two kilometer (km) section of the mainstem Salmon River located rough 5.8 river kms downstream from North Fork, Idaho (publication quality map referenced here).
+The downstream end of the Deadwater Slough occurs at the confluence of Dump Creek and the Salmon River.
+Around 1897, a small mining diversion reservoir failed in the Dump Creek drainage, creating upslope instability, causing a large erosion event that deposited large amounts of sediment at the confluence of the Salmon River and Dump Creek (Emerson 1973 ref here).
+The event created a large alluvial fan which backed up the Salmon River creating an approximately 25.3 acre (maybe convert to square meters or kilometers?) reservoir-like area that is unnaturally slow and deep (map reference or image here).
+The river in this section has slower water velocities, a deepened channel, and (anecdotally) warmer water temperatures which can favor piscivorous fishes (e.g., northern pikeminnow *Ptychocheilus oregonensis*, smallmouth bass *Micropterus dolomieu*).
+
+## Abundance of Piscivorous Fishes (Objective 1)
+
+**Data = C/M/R and Effort**
+
+* Capture methods used - We initially attempted electrofishing, snorkeling, netting, others?, but eventually settled on angling as best method for mark-recapture.
+* Mark-recapture - We kept the mark and recapture events close together in an attempt to use a closed population model
+* What are the assumptions of a closed population model?
+* How was sampling performed? What relevant information was captured for each fish? How/where were fish released? How was effort recorded to calculate CPUE?
+* Description of sampling time frame:
+  - Fall 2019: Goal was to estimate abundance during the peak of DSR emigration i.e., timing was done to coincide to be shortly after peak fall (DSR) emigration at the lower Lemhi River screw trap with the hopes of documenting predation.
+DSR are the more abundant juvenile emigration tactic.
+  - Fall 2020: Intent was to repeat the above, but instead during peak NRR emigration, but effort canceled due to COVID-19.
+Instead, sampling was re-schedule to fall when social distancing, etc.
+could be put in place.
+  - Spring 2021: Did a relative abundance effort during spring 2021.
+Had to reduce to a CPUE (single-week) effort due to funding constraints.
+However, goal was to 1) document presence during NRR outmigration and 2) estimate CPUE to provide a relative comparison to the fall efforts.
+
+## Predation Upon Juvenile Chinook Salmon i.e., Gastric Lavage (Objective 2)
+
+**Data = Gastric Lavage Data**
+
+The goal was to document predation upon juvenile Chinook salmon (or other targets), and ideally, estimate the proportion of their diet that consisted of juvenile Chinook salmon at the time of sampling?
+
+* Which individuals was gastric lavage performed upon?
+* How was gastric lavage performed?
+* Intent was to quantify predation on juvenile Chinook salmon
+  - or incidentally, juvenile steelhead or sockeye salmon
+
+## Bioenergetics (Objective 3)
+
+**Data = Temperature, Bioenergetics Inputs, Others?**
+
+Pull material from technical report
+
+## Impacts to Adult Returns (Objective 4)
+
+**Data = SARs, Adult Escapements, Others?**
+
+Pull material from technical report
+
+# Results
+
+## Abundance of Piscivorous Fishes (Objective 1)
+
+* Total population estimates of Pike Minnow. Chart or table, but likely a chart.
+* What is that equate to as a density? e.g., how many pikeminnow per 10 square meters? Is that a reasonable number?
+
+## Predation Upon Juvenile Chinook Salmon i.e., Gastric Lavage (Objective 2)
+
+*	Gastric lavage sampling results. What percent of the diet is chinook salmon from each sampling event. Pie chart? Mike loves pie charts.
+
+## Bioenergetics (Objective 3)
+
 * Water Temperature measurements.
@@
     *	Calculated daily average water temperature for 365 days of the year.
-*	Developed a bioenergetics model for the average sized Pike Minnow observed to maintain body weight through the 365 days modelled. 
-    *	Incorporated daily temperature time step.
-    * Calculated total biomass consumed to maintain weight for an individual fish across 365 days.
-    *	Multiplied by the total population estimates. (this may vary with time, but not sure yet).
-*	Adult estimates.
-    *	Calculated total mass consumed based on above. Multiplied fraction of diet that is chinook by the total mass consumed. 
-    *	Then we have to divide by average mass of a fish from the Lemhi screw trap? Smolt, pre-smolt, average?
-*	Used SARs (published or ours?) to then estimate the number of adults that may return if predation is reduced. 
-    *	Do we want to say predation is reduced to 0, or maybe a range of reduction?
-*	Anything else?
-
-# Results
-
-*	Total population estimates of Pike Minnow. Chart or table, but likely a chart.
-*	Gastric lavage sampling results. What percent of the diet is chinook salmon from each sampling event. Pie chart? Mike loves pie charts.
 *	Water temperature plot for the 365 average.
 *	Bioenergetics results. Total mass consumed over a 365 day period. Plot of daily consumption with day of year on x and mass on y.
+
+## Impacts to Adult Returns (Objective 4)
+
 *	SAR results. Maybe a plot of multiple scenarios depending on gastric lavage results? And pop estimates if they vary in time.
-*	Anything else?
 
@@
 
-*	Here we will need to discuss for each result above: 
-    *	How do our results compare to other sources within the literature?
-    *	What are the broader implications of each result?
-*	Management implications and how these results, analysis, modelling framework could be applied elsewhere?
+Start with a summary of the most important results! How many pikeminnow are there? Did we document/quantify predation? What's the potential impacts to adult returns?
 
-## Conclusion
+## Mark-Recapture Model
 
-*	What are the most important results?
-    *	Adult returns
-    *	Pop estimate of pike minnow
-    *	Total predation on juveniles
-*	We have also presented a novel modeling framework for estimating predation on native, critically endangered anadromous species which can be applied to other areas of interest. John day? Others?
+Re-hash out our assumptions?
+What assumptions were most likely violated e.g., the closed population assumption?
+If we assumed an open population with varying immigration/emigration rates how might that affect our estimate?
+In the end, how good do we think our estimate is and is the bottom-line that we belive there still to be a shit-ton of pikeminnow even if our assumptions were violated?
+
+## Gastric Lavage
+
+## Bioenergetics
+
+## Impacts to Adult Returns
+
+## Avian Predation
+
+In the discussion, let's also acknowledge the potential impacts from avian predation.
+Outside of the scope of the study, but what did we find in the first year?
+Other large studies that estimate impacts of avian predaton (herons in particular) on juvenile Chinook salmon, especially in altered habitats?
+Anecdotally, what did we see i.e., we noticed lots of herons traveling to/from the site especially in the anastomosizing section above Deadwater Slough?
+
+* In addition to fish predators the Deadwater Slough is also recognized as an important bird watching and nesting area due to the riparian and backwater habitats created by the feature [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
+
+## Management Implications
+
+Discuss the potential for various management actions:
+* Removing the Dump Creek alluvial fan and restoring flow. What would that look like?
+* A local bounty program? What would that look like? Have they been successful elsewhere e.g., in the Columbia River?
+
 *	Also, this is an anthropogenically created dead water area, which could be a candidate for restoration (removing the impoundment).
@@
 Will these fish just move elsewhere?
+
+## Conclusions
+
+*	We have also presented a novel modeling framework for estimating predation on native, critically endangered anadromous species which can be applied to other areas of interest. John day? Others?
 *	The end.

### 2021-06-11 analysis/paper/paper.Rmd
final paper edits of the day

@@
 
+<!-- below pulled from technical report -->
+
+We collected stomach contents from most captured individuals using gastric lavage (Foster 1977) and examined contents for the presence or absence of juvenile Chinook salmon and other incidentals (e.g., steelhead, sockeye salmon) juveniles and the proportion of stomach contents containing targets versus non-targets (e.g., macroinvertebrates).
+Stomach contents were stored in whirl-paks, preserved with 99% isopropyl alcohol, and analyzed one week later in a controlled environment.
+Each sample was uniquely identified to match up with the appropriate fish record, contents were identified down to its unique composition, total weight of all content was measured in grams, and total weight of fish content, if found, was measured in grams.
+Throughout the sampling period a proportion of the captured individuals were sacrificed after gastric lavage to validate that the gastric lavage was successful at flushing all or most of the stomach contents from the northern pikeminnow.
+
 ## Bioenergetics (Objective 3)
@@
 ## Gastric Lavage
+
+Concerns with gastric lavage.
+We only observed juvenile Chinook salmon (and/or other fishes) in a very small number of stomach contents.
+However, we don't necessarily believe that to mean that pikeminnow in Deadwater Slough rarely consume juvenile Chinook salmon.
+Are there other cases where gastric lavage failed to document predation?
+How soon after consumption do you need to take a sample?
+So instead, we made some assumptions about the proportion of a pikeminnows diet consisting of Chinook salmon.
 

### 2021-06-15 analysis/paper/paper.Rmd
update to ms outline

@@
       correspondence: false
-  - Michael Hall:
+  - Michael S. Hall:
       email: Mike.Hall1@merck.com
@@
 bibliography: 
-- references.bib
-- packages.bib
+- references.bib      # Kevin's library
+- AckermanLibrary.bib # Mike's library
 csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
 abstract: |
-  Although well documented through the Columbia and Snake River hydropower system, little is known about movement and survival of Salmon River juvenile Chinook salmon through their migratory corridors above the hydropower system.
-With increases in water temperatures, predator abundance, and predator distribution, this knowledge is evermore necessary to characterize limiting factors and recovery potential of this imperiled species.
-To further understanding of predation impacts on juvenile Chinook Salmon in the Salmon River, specifically, the anthropogenically altered Deadwater Reach, we estimated the population of predatory species, calculated the bioenergetic requirement of predator species to maintain size, and estimated the number of juveniles predated upon within said reach.
-We then used previously established smolt-to-adult return ratios to understand the impact that predation within the Deadwater Reach has on adult returns to the Lemhi River of Idaho, with broader implications for anadromous populations outside of the Lemhi River specifically.
+  Abstract text here. 
 keywords: |
@@
 
+```{r load-libraries}
+library(tidyverse)
+library(janitor)
+
+```
+
+```{r set-prefix}
+# set NAS prefix, depending on operating system
+if(.Platform$OS.type != 'unix') {
+  nas_prefix = "S:"
+} else if(.Platform$OS.type == 'unix') {
+  nas_prefix = "~/../../Volumes/ABS"
+}
+
+```
+
+\newpage
+
 # Introduction
@@
 
-* One potentially under-appreciated source of mortality on juvenile Chinook salmon is predation from piscivorous fishes.
+* One potentially under-appreciated source of mortality on juvenile Chinook salmon is predation from piscivorous fishes, native and/or non-native
   - At this point, I'm thinking leave avian predation out of the equation, but instead include in Discussion
@@
 * What is known about downriver (Columbia/Snake) piscivorous fish populations and potential predation on juvenile Chinook salmon?
-  - literature review of northern pikeminnow, smallmouth bass, other? predation on juvenile Chinook salmon
+  - literature review of northern pikeminnow, smallmouth bass, other? predation on juvenile Chinook salmon, particulary from the Snake and Columbia rivers
   - still stay away from avian predation
@@
 
-* Summary of findings from Axel et al.
-and early RT reports and papers (Ackerman et al.
-2019, Porter et al.
-2020, 2021, See et al.
-RT manuscript) in Deadwater slough.
+* Summary of findings from @Axel2015 and early RT reports and papers (@Ackerman2018a, @Porter2019, @Porter2020) in Deadwater Slough.
 Recent studies have demonstrated decreased movement and survival of juvenile salmonids emigrating through the Deadwater Slough.
-Axel et al.
-(2015) demonstrated decreased rates of emigration and survival for sockeye salmon *Oncorhynchus nerka* from Redfish Lake emigrating during the spring and recent winter telemetry studies have indicated decreased transition probabilities (approximately 10% less than nearby reaches) of juvenile Chinook salmon *Oncorhynchus tschawytscha* through Deadwater Slough during fall and winter months (Ackerman et al.
-2018; Porter et al.
-2019).
-* Axel et al.
-(2015) found increased mortality in a particular reach spanning location x to location y in the mainstem Salmon River on a study of juvenile (hatchery?) sockeye salmon released near Redfish Lake and actively migrating downriver.
+@Axel2015 demonstrated decreased rates of emigration and survival for sockeye salmon *Oncorhynchus nerka* from Redfish Lake emigrating during the spring and recent winter telemetry studies have indicated decreased transition probabilities (approximately 10% less than nearby reaches) of juvenile Chinook salmon *Oncorhynchus tschawytscha* through Deadwater Slough during fall and winter months (@Ackerman2018a, @Porter2019, @Porter2020).
+* @Axel2015 found increased mortality in a particular reach spanning location x to location y in the mainstem Salmon River on a study of juvenile (hatchery?) sockeye salmon released near Redfish Lake and actively migrating downriver.
   - in their case, these were actively migrating juveniles, and so transitional probabilities can likely be interpreted (almost) directly as survival
   - A previous study showed that the lowest transitional probabilities of Red Fish Lake Sockeye occurred in the Dead Water reach.
-* Ackerman et al. and Porter et al. found reduced (apparent) survival in a reach spanning near the town of North Fork to just downstream of Deadwater Slough.
+* @Ackerman2018a and @Porter2019, @Porter2020 found reduced (apparent) survival in a reach spanning near the town of North Fork to just downstream of Deadwater Slough.
   - in these cases, there's a chance that juvenile Chinook could exhibit concealment behavior prior to winter and so transition probability is interpreted as the probability of a fish surviving AND moving
-* Anecdotally, locals (and others? maybe NOAA?) informed us that northern pikeminno and perhaps other species are prevalent in Deadwater Slough.
+* Anecdotally, locals (and others? maybe NOAA?) informed us that northern pikeminnow and perhaps other species are prevalent in Deadwater Slough.
 
@@
 
-* Chinook salmon in the Lemhi River exhibit 2 juvenile migration tactics (site Copeland and Venditti [2014] and See et al. *in prep*).
+* Chinook salmon in the Lemhi River exhibit two juvenile migration tactics (@Copeland2014c and See et al. *in prep*).
   - downstream rearing (DSR)
@@
 The downstream end of the Deadwater Slough occurs at the confluence of Dump Creek and the Salmon River.
-Around 1897, a small mining diversion reservoir failed in the Dump Creek drainage, creating upslope instability, causing a large erosion event that deposited large amounts of sediment at the confluence of the Salmon River and Dump Creek (Emerson 1973 ref here).
+Around 1897, a small mining diversion reservoir failed in the Dump Creek drainage, creating upslope instability, causing a large erosion event that deposited large amounts of sediment at the confluence of the Salmon River and Dump Creek (@Emerson1973).
 The event created a large alluvial fan which backed up the Salmon River creating an approximately 25.3 acre (maybe convert to square meters or kilometers?) reservoir-like area that is unnaturally slow and deep (map reference or image here).
@@
 ## Abundance of Piscivorous Fishes (Objective 1)
+
+```{r cmr-effort-data}
+cmr_df = read_csv(paste0(nas_prefix, "/data/fish/deadwater/deadwater_cmr_effort_20210614.csv")) %>%
+  clean_names()
+
+```
 
@@
 
+The Chapman-modified Lincoln-Peterson estimator is below, where $M$ is the number of fish marked and returned to the population, $n$ is the number of fish caught in the second/recapture event and $m$ is the number of marked fish in the second sample.
+$$
+\hat{N} = \frac{(M + 1)(n + 1)}{(m + 1)} - 1
+$$
+The Schnabel estimator is shown below, where the $M$, $n$ and $m$ are indexed by the sampling occasion, $i$.
+This estimator does not have an associate standard error, but 95% confidence intervals can be calculated.
+$$
+\hat{N} = \frac{\sum\limits_{i = 1}^k n_i M_i}{\left(\sum\limits_{i = 1}^k m_i \right) + 1}
+$$
+
 ## Predation Upon Juvenile Chinook Salmon i.e., Gastric Lavage (Objective 2)
@@
 
-We collected stomach contents from most captured individuals using gastric lavage (Foster 1977) and examined contents for the presence or absence of juvenile Chinook salmon and other incidentals (e.g., steelhead, sockeye salmon) juveniles and the proportion of stomach contents containing targets versus non-targets (e.g., macroinvertebrates).
+We collected stomach contents from most captured individuals using gastric lavage (@Foster1977) and examined contents for the presence or absence of juvenile Chinook salmon and other incidentals (e.g., steelhead, sockeye salmon) juveniles and the proportion of stomach contents containing targets versus non-targets (e.g., macroinvertebrates).
 Stomach contents were stored in whirl-paks, preserved with 99% isopropyl alcohol, and analyzed one week later in a controlled environment.
@@
 ## Bioenergetics (Objective 3)
+
+```{r temp-data}
+temp_df = read_delim(paste0(nas_prefix,
+                            "/data/temperature/deadwater/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
+                     delim = "\t",
+                     comment = "#") %>%
+  rename(discharge_cfs = `47768_00060`,
+         discharge_cfs_cd = `47768_00060_cd`,
+         gage_ht_ft = `47769_00065`,
+         gage_ht_ft_cd = `47769_00065_cd`,
+         water_temp_c = `47773_00010`,
+         water_temp_c_cd = `47773_00010_cd`)
+
+```
 
@@
 
+* Deadwater is an opportunity to benefit multiple (6 or 7) local populations with a single management or restoration action.
+
 *	Also, this is an anthropogenically created dead water area, which could be a candidate for restoration (removing the impoundment).

### 2021-11-03 analysis/R scripts/Nick Script 11.2.21 cleaning fish data.R
Updating Bioenergetics and adding files
Updating bioenergetics and adding script to investigate data to repository

@@ -1,0 +1,77 @@
+# working to clean up and provide summaries for Deadwater Data
+# Nick Porter
+# 11/2/20201
+rm(list = ls()) # clear the working environment
+
+
+
+require(tidyverse)
+require(lubridate)
+require(boot)
+require(bootstrap)
+require(lme4)
+require(Matrix)
+require(numDeriv)
+require(FSA)
+require(FSAdata)
+require(magrittr)
+require(car)
+require(dplyr)
+require(plotrix)
+require(nnet)
+require(Hmisc)
+require(Matching)
+require(Rcapture)
+require(nlstools)
+require(stringi)
+require(MASS)
+
+deadwater <- read_csv("analysis/data/raw data/deadwater_cmr_effort_20210614.csv") %>%
+  mutate(Date = as.POSIXct(Date, format = "%m/%d/%y"))
+
+NPM = deadwater %>%
+  filter(Species == "Northern Pikeminnow")
+
+Clean_NPM = NPM %>%
+  mutate(Fish_or_Parts = case_when(stri_detect_fixed(StomachContents, "Fish") ~ "Fish or Fish Parts",
+                                   stri_detect_fixed(StomachContents, "Empty") ~ "Empty", TRUE ~ "Other")) %>%
+  drop_na(Length)
+
+
+
+#Histogram Ogle
+
+Clean_NPM$NPM_50=lencat(Clean_NPM$Length, w = 50)
+
+esquisse::esquisser(Clean_NPM)
+
+
+##group_by(Species) %>% 165,640,25
+#custbins=seq(150,675,25)
+#hist(~Length,data=Clean_NPM,xlab="TL (mm)", ylim=c(0,500))
+##xtick<-seq(150,650,20)####creating minor tick marks
+#axis(side=1, at=xtick, labels = FALSE)####setting minor tick marks
+
+
+ggplot(data=Clean_NPM, aes(Length)) +
+  geom_histogram(breaks=seq(150, 650, by = 50),
+                 col="#DB740E",
+                 fill="#DB740E",
+                 alpha = .2) +
+  labs(x="TL (mm)", y="Frequency") +
+  xlim(c(100,700)) +
+  ylim(c(0,500)) +
+  theme_classic()
+
+
+Clean_NPM %>%
+  filter(Date >= "2020-05-17 23:00:00" & Date <= "2020-11-21 00:00:00") %>%
+  filter(Method %in%
[…]

### 2021-11-04 analysis/paper/paper.Rmd
changed paths to data

@@
       correspondence: true
+  - Nicholas J. Porter:
+      email: Nicholas.Porter@merck.com
+      institute: [biomark]
+      correspondence: false
   - Kevin E. See:
       email: Kevin.See@merck.com
-      institute: [biomark]
-      correspondence: false
-  - Nicholas J. Porter:
-      email: Nicholas.Porter@merck.com
       institute: [biomark]
@@
 library(tidyverse)
+library(magrittr)
 library(janitor)
-
-```
-
-```{r set-prefix}
-# set NAS prefix, depending on operating system
-if(.Platform$OS.type != 'unix') {
-  nas_prefix = "S:"
-} else if(.Platform$OS.type == 'unix') {
-  nas_prefix = "~/../../Volumes/ABS"
-}
 
@@
 ```{r cmr-effort-data}
-cmr_df = read_csv(paste0(nas_prefix, "/data/fish/deadwater/deadwater_cmr_effort_20210614.csv")) %>%
+cmr_df = read_csv("../data/raw data/deadwater_cmr_effort_20210614.csv") %>%
   clean_names()
@@
 
-```{r temp-data}
-temp_df = read_delim(paste0(nas_prefix,
-                            "/data/temperature/deadwater/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
-                     delim = "\t",
+```{r shoup-temp-data}
+temp_df = read_tsv("../data/raw data/USGS_13307000_Salmon_River_Nr_Shoup.txt",
                    comment = "#") %>%

### 2021-11-04 analysis/paper/paper.Rmd
additions to Introduction

@@
       email: Kevin.See@merck.com
-      institute: [biomark]
+      institute: [biomark, wdfw]
       correspondence: false
@@
       email: Mike.Ackerman@merck.com
-      institute: [biomark]
+      institute: [biomark, mhe]
       correspondence: false
@@
   - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
+  - wdfw: Washington Department of Fish and Wildife, *Address Here*
+  - mhe: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-***Paragraph 1***
-
-* The Lemhi River supports an ESA-listed Chinook salmon population with substantial cultural and recreational value and is a candidate for protection and/or recovery.
-* The population has become depleted with lower adult returns in recent years.
-* Decreased returns and/or increased mortality of the population have often been attributed to (in chronological order by life stage):
-  - Degraded natal tributary habitat (e.g., high amount of agriculture)
-  - Reduced survival in the mainstem corridor e.g., Snake and Columbia rivers
-  - Poor ocean conditions
-  - Others worth mentioning? Perhaps adult harvest in ocean and/or in-river returns?
+The Upper Salmon River major population group (MPG) supports eight extant spring/summer Chinook salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
+For the MPG to be considered viable, at least five of these eight populations should meet viability criteria set forth by @McElhany2000 and @ICTRT2007, and is necessary for recovery of the Snake River Evolutionary Significant Unit (ESU).
+Populations within the ESU have substantial cultural value, support downriver mainstem Snake and Columbia River commercial and subsistence fisheries, and in year with sufficient abundance, support local fisheries that contribute to the local economies.
+All Upper Salmon River MPG populations have become depleted in recent decades.
+Declines in survival of juvenile Chinook salmon have been attributed to the removal of beavers from the landscape (fur trade), mining activities, simplification of rivers resulting from the construction of railroads and agriculture, water withdrawals, logging activities, urbanization, avian predation, proliferation of non-native species (e.g., non-native coastal rainbow trout *O.
+mykiss irideus* and brook trout *Salvelinus fontinalis*), warming streams and rivers, and modifications to downriver migration corridors (e.g., from hydropower projects).
+Moreover, the abundance of adult returns are further impacted by ocean and downriver harvests, altered ocean conditions, and changes to the spawning migration corridor.
+Each of these factors have contributed, to varying and unknown extents, to reduced adult escapement, the primary metric used to assess population viability.
+In response to the above human activities and reduced abundances, action agencies have invested substantial funds in recent years to rehabilitate tributary ecosystems in an attempt to improve juvenile survival and adult spawning conditions.
 
@@
 
-* One potentially under-appreciated source of mortality on juvenile Chinook salmon is predation from piscivorous fishes, native and/or non-native
+One potentially important, but perhaps under-appreciated source of mortality on Chinook salmon is predation on emigrating juveniles by piscivorious fishes, including both native and non-native species.
+As an example,...
+
   - At this point, I'm thinking leave avian predation out of the equation, but instead include in Discussion
@@
 
-* Chinook salmon in the Lemhi River exhibit two juvenile migration tactics (@Copeland2014c and See et al. *in prep*).
-  - downstream rearing (DSR)
-  - natal reach rearing (NRR)
-* Define these juvenile migration tactics for stream-type Chinook salmon
+Spring/summer Chinook salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics, termed downstream rearing (DSR) and natal reach rearing (NRR) by [@Copeland2014].
+The DSR migrants leave the natal spawning area as subyearlings during June through November and typically winter in downstream, mainstem habitats prior to emigration to the ocean the following spring as smolts.
+Alternatively, NRR migrants remain in their natal areas for approximately one year after emergence prior to emigration to the ocean as smolts.
+The diversity of emigration tactics provides for increases population resiliency due to changes in the freshwater rearing and migration environments and buffers against catastrophic events.
 

### 2021-11-05 analysis/paper/paper.Rmd
additional intro and discussion content

@@
 ---
-title: "Northern Pikeminnow Abundance in Deadwater Slough and Potential Impacts to Local Chinook Salmon Populations"
+title: "Predation of Northern Pikeminnow on Juvenile Chinook Salmon and Potential Impacts to Local Populations"
 author:
@@
   - Kevin E. See:
-      email: Kevin.See@merck.com
+      email: Kevin.See@dfw.wa.gov
       institute: [biomark, wdfw]
@@
   - Michael W. Ackerman:
-      email: Mike.Ackerman@merck.com
+      email: mike.ackerman@mthoodenvironmental.com
       institute: [biomark, mhe]
@@
   - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
-  - wdfw: Washington Department of Fish and Wildife, *Address Here*
+  - wdfw: Washington Department of Fish and Wildife, Under A Bridge, Seattle, Washington, 00000, USA
   - mhe: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
@@
 
-* Summary of findings from @Axel2015 and early RT reports and papers (@Ackerman2018a, @Porter2019, @Porter2020) in Deadwater Slough.
+* Summary of findings from @Axel2015 and early RT reports and papers (@Ackerman2018, @Porter2019, @Porter2020) in Deadwater Slough.
 Recent studies have demonstrated decreased movement and survival of juvenile salmonids emigrating through the Deadwater Slough.
-@Axel2015 demonstrated decreased rates of emigration and survival for sockeye salmon *Oncorhynchus nerka* from Redfish Lake emigrating during the spring and recent winter telemetry studies have indicated decreased transition probabilities (approximately 10% less than nearby reaches) of juvenile Chinook salmon *Oncorhynchus tschawytscha* through Deadwater Slough during fall and winter months (@Ackerman2018a, @Porter2019, @Porter2020).
+@Axel2015 demonstrated decreased rates of emigration and survival for sockeye salmon *Oncorhynchus nerka* from Redfish Lake emigrating during the spring and recent winter telemetry studies have indicated decreased transition probabilities (approximately 10% less than nearby reaches) of juvenile Chinook salmon *Oncorhynchus tschawytscha* through Deadwater Slough during fall and winter months (@Ackerman2018, @Porter2019, @Porter2020).
 * @Axel2015 found increased mortality in a particular reach spanning location x to location y in the mainstem Salmon River on a study of juvenile (hatchery?) sockeye salmon released near Redfish Lake and actively migrating downriver.
@@
   - A previous study showed that the lowest transitional probabilities of Red Fish Lake Sockeye occurred in the Dead Water reach.
-* @Ackerman2018a and @Porter2019, @Porter2020 found reduced (apparent) survival in a reach spanning near the town of North Fork to just downstream of Deadwater Slough.
+* @Ackerman2018 and @Porter2019, @Porter2020 found reduced (apparent) survival in a reach spanning near the town of North Fork to just downstream of Deadwater Slough.
   - in these cases, there's a chance that juvenile Chinook could exhibit concealment behavior prior to winter and so transition probability is interpreted as the probability of a fish surviving AND moving
 * Anecdotally, locals (and others? maybe NOAA?) informed us that northern pikeminnow and perhaps other species are prevalent in Deadwater Slough.
-
-***Paragraph 4***
 
@@
 
-***Paragraph 5***
+The Deadwater Slough is the result of human activities and resulted from an alluvial fan that formed at the bottom of Deadwater following a storm event and resulting failure of a mining dam on Dump Creek.
+Additionally, the Deadwater Slough occurs in a reach of the Salmon River believed to be an important overwinter rearing area for DSR emigrants attempting to exhibit concealment behavior to survive winter prior to emigration to the ocean the following spring.
+Deadwater Slough contains little to no hydrological or structure features (i.e., little fish cover) and fine sediment or mud substrate; all conditions making concealment behavior difficult and providing little refuge from predation.
+The Deadwater reach of the Salmon River is also important in that juvenile DSR and NRR emigrant from all eight extant populations must pass through the area during ocean migration.
 
-* Some concern arose that predation at or near Deadwater Slough was *not* inconsequential to the Lemhi River Chinook salmon (and adjacent) population(s), impacting both DSR and NRR emigrants
-* We hypothesize that decreases in survival (or apparent survival) of juvenile Chinook salmon (and sockeye?) are largely, or at least partially, due to increased densities of piscivorous predators in the Deadwater Slough area resulting from anthropogenically impacted habitat after the Dump Creek alluvial fan was formed.
-* Beginning in 2019, we set out to evaluate fish predator populations in the Deadwater Slough and their potential predation upon and impacts to juvenile emigrants from the Lemhi River population, both DSR and NRR.
-* Our objectives are four-fold:
-  1. Estimate the abundance of fish predators in Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations from the Lemhi River.
-  2. Document (and perhaps quantify) predation on juvenile Chinook salmon during the sampling/emigration periods.
-  3. Using a bioenergetics approach, attempt to estimate the number of juvenile Chinook salmon consumed during some defined fall and/or spring sampling/emigration period.
-  4. After making some assumptions (e.g., on SARs) attempt to estimate how results from objective #3 could translate into impacts on adult returns (via a thought experiment or simulation approach)
+In recent years, concerns arose that predation on juvenile Chinook salmon at or near the Deadwater Slough was not inconsequential to Chinook salmon populations in the Upper Salmon River MPG, impacting both DSR migrants in the fall and NRR emigrants during the following spring.
+We hypothesize that decreases in survival (or apparent survival) of juvenile Chinook salmon observed by @Ackerman2018, and similar for sockeye salmon by @Axel2015, are largely, or at least partially, due to increased densities of piscivorous predators in the Deadwater Slough area.
+=Beginning in 2019, we set out to evaluate potential piscivorous fish predator populations in the Deadwater Slough and evaluate their potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook salmon.
+Our objectives for the study were four-fold:
 
-* Additionally, Deadwater Slough occurs in a reach of the Salmon River believed to be important for overwinter rearing of presmolts attempting to exhibit concealment behavior to survive winter prior to spring emigration to the ocean.
-Deadwater Slough further contains little to no hydrological or structure features (i.e., fish cover) allowing juvenile salmonids to be predated upon without being able to seek adequate refuge.
-This is also an extremely important reach and relevant study to the entirety of the upper Salmon River anadromous population(s), as each emigrating fish must pass through the Deadwater area.
+1. Estimate the abundance (or relative abundance) of potential predators in the Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations.
+2. Document predation on juvenile Chinook salmon during the emigration periods using gastric lavage.
+3. Using a bioenergetics approach, attempt to estimate the number of juvenile Chinook salmon consumed during defined emigration periods for DSR and NRR juveniles.
+4. Estimate potential impacts to Chinook salmon adult returns based on the estimate of juvenile Chinook salmon consumed during the defined emigration periods.
 
@@
 
+  4. After making some assumptions (e.g., on SARs) attempt to estimate how results from objective #3 could translate into impacts on adult returns (via a thought experiment or simulation approach)
+
 # Results
@@
 
-Start with a summary of the most important results! How many pikeminnow are there? Did we document/quantify predation? What's the potential impacts to adult returns?
+We estimated the population size of northern pikeminnow in the Deadwater Slough and during the fall emigration period to be greater than ***xx,xxx***.
+And although we did not attempt to estimate population size during the spring emigration period, the relative abundance measured as CPUE was comparable to that during the fall sampling periods (update statement later).
+That fall population size estimate translates to ***xxx*** northern pikeminnow per 100 m or ***xxx*** pikeminnow per 100 m^2^.
+For context, that is similar/more/less than estimates from the Columbia River where predation impacts have been demonstrated as substantial and bounty programs have been implemented to reduce northern pikeminnow abundance.
+Given that estimate of abundance and some assumptions, we estimated that northern pikeminnow potentially consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook salmon during the *x* sampling periods resulting in a reduction of adult returns between ***xxx*** and ***x,xxx***.
+We suggest that modifications to the habitat due to human activities in the Deadwater Reach has resulted in favorable conditions for northern pikeminnow, including improving conditions for predation upon juvenile Chinook salmon, which has resulted in a consequential impact to ESA-listed Chinook salmon populations in the Upper Salmon River MPG.
 
@@
 
+What assumptions did we make during the bioenergetics assessment?
+And how might violations of those assumptions change our estimate of the number of juvenile Chinook salmon consumed and resulting impacts to adult returns?
+
 ## Impacts to Adult Returns
+
+Again, what assumptions did we make here and how might violations of those assumptions change our estimate of impacts to adult returns.
 
@@
 
-In the discussion, let's also acknowledge the potential impacts from avian predation.
-Outside of the scope of the study, but what did we find in the first year?
-Other large studies that estimate impacts of avian predaton (herons in particular) on juvenile Chinook salmon, especially in altered habitats?
-Anecdotally, what did we see i.e., we noticed lots of herons traveling to/from the site especially in the anastomosizing section above Deadwater Slough?
-
-* In addition to fish predators the Deadwater Slough is also recognized as an important bird watching and nesting area due to the riparian and backwater habitats created by the feature [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
+Although not formally assessed here, another potential source of predation upon juvenile salmon in the Deadwater Slough is from avian predators including great blue herons *Ardea herodias* and bald eagles *Haliaeetus leucocephalus*.
+The Deadwater Slough is recognized as an important bird watching and nesting area due to the riparian and backwater habitats created by the feature [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
+During the initial sampling event in 2019, a two-person crew equipped with a mobile passive integrated transponder (PIT) tag antenna walked the entire reach and surrounding and upstream areas scanning for PIT tags.
+During the informal survey, we did identify nine PIT tags on surfaces surrounding Deadwater Slough, many of which were near nests or in an upstream anastomizing reach where herons and eagles are prevalent.
+Investigating their tag histories in [PTAGIS](https://www.ptagis.org/), these tags were implanted into a combination of juvenile Chinook Salmon (3), Sockeye Salmon (3), and steelhead (3) suggesting avian predation does occur at and adjacent to the Deadwater Reach, although to an unknown degree.
+***Follow up with a brief lit review on avian predation on juvenile Chinook salmon elsewhere***
 

### 2021-11-08 analysis/R scripts/Nick Script 11.2.21 cleaning fish data.R
Updated Bioenergetics with date

@@ -25,9 +25,13 @@ require(Rcapture)
 require(nlstools)
 require(stringi)
 require(MASS)
-
+require(readr)
 deadwater <- read_csv("analysis/data/raw data/deadwater_cmr_effort_20210614.csv") %>%
   mutate(Date = as.POSIXct(Date, format = "%m/%d/%y"))
+
+bioenergetics<- read_csv("analysis/data/raw data/2020 Bioenergetics 1 year.csv") #%>%
+   mutate(day = as.POSIXct(day))
+  mutate(day = as.Date(day, origin=as.Date("1960-01-01")))
 
 NPM = deadwater %>%
   filter(Species == "Northern Pikeminnow")
@@ -43,7 +47,7 @@ Clean_NPM = NPM %>%
 
 Clean_NPM$NPM_50=lencat(Clean_NPM$Length, w = 50)
 
-esquisse::esquisser(Clean_NPM)
+#esquisse::esquisser(Clean_NPM)
 
 
 ##group_by(Species) %>% 165,640,25

### 2021-11-08 analysis/paper/paper.Rmd
done for the week; added concluding statement to intro
incorporated TBs edits to Intro

@@
 ---
-title: "Predation of Northern Pikeminnow on Juvenile Chinook Salmon and Potential Impacts to Local Populations"
+title: "Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations"
 author:
@@
 
-The Upper Salmon River major population group (MPG) supports eight extant spring/summer Chinook salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
-For the MPG to be considered viable, at least five of these eight populations should meet viability criteria set forth by @McElhany2000 and @ICTRT2007, and is necessary for recovery of the Snake River Evolutionary Significant Unit (ESU).
-Populations within the ESU have substantial cultural value, support downriver mainstem Snake and Columbia River commercial and subsistence fisheries, and in year with sufficient abundance, support local fisheries that contribute to the local economies.
-All Upper Salmon River MPG populations have become depleted in recent decades.
-Declines in survival of juvenile Chinook salmon have been attributed to the removal of beavers from the landscape (fur trade), mining activities, simplification of rivers resulting from the construction of railroads and agriculture, water withdrawals, logging activities, urbanization, avian predation, proliferation of non-native species (e.g., non-native coastal rainbow trout *O.
+The Upper Salmon River major population group (MPG) supports eight extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
+At least five of these eight populations must meet criteria set forth by @McElhany2000 and @ICTRT2007 for the MPG to be considered viable and for the recovery of the Snake River Evolutionary Significant Unit (ESU).
+Populations within the ESU have substantial cultural value, support downriver mainstem Snake and Columbia River commercial and subsistence fisheries, and support local fisheries and economies in years with sufficient abundance.
+All populations within the Upper Salmon River MPG have become depleted in recent decades.
+Declines in survival of juvenile Chinook Salmon have been attributed to the removal of beavers from the landscape (fur trade), mining activities, river simplification, water withdrawals, logging activities, urbanization, avian predation, proliferation of non-native species (e.g., non-native coastal rainbow trout *O.
 mykiss irideus* and brook trout *Salvelinus fontinalis*), warming streams and rivers, and modifications to downriver migration corridors (e.g., from hydropower projects).
-Moreover, the abundance of adult returns are further impacted by ocean and downriver harvests, altered ocean conditions, and changes to the spawning migration corridor.
+Moreover, the abundance of returning adults are further impacted by ocean and downriver harvests, poor ocean conditions, and changes to the spawning migration corridor.
 Each of these factors have contributed, to varying and unknown extents, to reduced adult escapement, the primary metric used to assess population viability.
-In response to the above human activities and reduced abundances, action agencies have invested substantial funds in recent years to rehabilitate tributary ecosystems in an attempt to improve juvenile survival and adult spawning conditions.
+In response to the decline in Chinook Salmon abundance from the myriad human activities and associated habitat degradation, action agencies have attempted to improve juvenile survival and adult spawning conditions by investing in the rehabilitation of tributary ecosystems.
 
@@
 
-One potentially important, but perhaps under-appreciated source of mortality on Chinook salmon is predation on emigrating juveniles by piscivorious fishes, including both native and non-native species.
+One potentially important, but perhaps under-appreciated source of mortality on Chinook salmon is predation on emigrating juveniles by piscivorous fishes, including both native and non-native species.
 As an example,...
@@
 
-Spring/summer Chinook salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics, termed downstream rearing (DSR) and natal reach rearing (NRR) by [@Copeland2014].
-The DSR migrants leave the natal spawning area as subyearlings during June through November and typically winter in downstream, mainstem habitats prior to emigration to the ocean the following spring as smolts.
-Alternatively, NRR migrants remain in their natal areas for approximately one year after emergence prior to emigration to the ocean as smolts.
-The diversity of emigration tactics provides for increases population resiliency due to changes in the freshwater rearing and migration environments and buffers against catastrophic events.
+Spring/summer Chinook Salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014].
+The DSR migrants leave the natal spawning area as subyearlings between June and November and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts.
+Alternatively, NRR migrants remain in their natal spawning areas for approximately one year after emergence until emigration to the ocean as smolts.
+Diversity of migratory tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency.
 
-The Deadwater Slough is the result of human activities and resulted from an alluvial fan that formed at the bottom of Deadwater following a storm event and resulting failure of a mining dam on Dump Creek.
-Additionally, the Deadwater Slough occurs in a reach of the Salmon River believed to be an important overwinter rearing area for DSR emigrants attempting to exhibit concealment behavior to survive winter prior to emigration to the ocean the following spring.
-Deadwater Slough contains little to no hydrological or structure features (i.e., little fish cover) and fine sediment or mud substrate; all conditions making concealment behavior difficult and providing little refuge from predation.
-The Deadwater reach of the Salmon River is also important in that juvenile DSR and NRR emigrant from all eight extant populations must pass through the area during ocean migration.
+The Deadwater Slough is located in a reach of the Salmon River that is believed to be a historically important overwinter rearing area for DSR emigrants.
+Moreover, this reach is part of the migratory pathway for juvenile DSR and NRR emigrants from all eight extant populations.
+The slough lacks hydrological and structural features (i.e., homogeneous channel with fine substrate) that can provide essential refuge from predation.
+As a result, predation on juvenile Chinook salmon proximal to the Deadwater Slough has been cited as a concern for the Upper Salmon River MPG, impacting DSR migrants in the fall and NRR emigrants during the spring.
 
-In recent years, concerns arose that predation on juvenile Chinook salmon at or near the Deadwater Slough was not inconsequential to Chinook salmon populations in the Upper Salmon River MPG, impacting both DSR migrants in the fall and NRR emigrants during the following spring.
-We hypothesize that decreases in survival (or apparent survival) of juvenile Chinook salmon observed by @Ackerman2018, and similar for sockeye salmon by @Axel2015, are largely, or at least partially, due to increased densities of piscivorous predators in the Deadwater Slough area.
-=Beginning in 2019, we set out to evaluate potential piscivorous fish predator populations in the Deadwater Slough and evaluate their potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook salmon.
+We hypothesize that increased densitied of piscivorous predators in the Deadwater Slough area may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon [@Ackerman2018] and Sockeye Salmon [@Axel2015].
+In this study, we estimated the abundance of a piscivorous fish predator population in the Deadwater Slough and their potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
 Our objectives for the study were four-fold:
@@
 1. Estimate the abundance (or relative abundance) of potential predators in the Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations.
-2. Document predation on juvenile Chinook salmon during the emigration periods using gastric lavage.
-3. Using a bioenergetics approach, attempt to estimate the number of juvenile Chinook salmon consumed during defined emigration periods for DSR and NRR juveniles.
-4. Estimate potential impacts to Chinook salmon adult returns based on the estimate of juvenile Chinook salmon consumed during the defined emigration periods.
+2. Document predation on juvenile Chinook Salmon during the emigration periods using gastric lavage.
+3. Use a bioenergetics approach to estimate the total consumption of juvenile DSR and NRR Chinook salmon during defined emigration periods.
+4. Quantify the potential impacts of estimated predation on juvenile to Chinook Salmon adult returns.
+
+We follow with a discussion of the various assumptions that went into the mark-recapture and bioenergetics models and assessment of impacts to adult returns and how violations of some assumptions may affect overall results and inferences from the study.
 

### 2021-11-08 analysis/R scripts/Nick Script 11.2.21 cleaning fish data.R
Added new Bioenergetics model with 10% growth over a year

@@ -26,12 +26,10 @@ require(nlstools)
 require(stringi)
 require(MASS)
 require(readr)
+
+#Read in the Deadwater dataset
 deadwater <- read_csv("analysis/data/raw data/deadwater_cmr_effort_20210614.csv") %>%
   mutate(Date = as.POSIXct(Date, format = "%m/%d/%y"))
-
-bioenergetics<- read_csv("analysis/data/raw data/2020 Bioenergetics 1 year.csv") #%>%
-   mutate(day = as.POSIXct(day))
-  mutate(day = as.Date(day, origin=as.Date("1960-01-01")))
 
 NPM = deadwater %>%
   filter(Species == "Northern Pikeminnow")
@@ -40,6 +38,32 @@ Clean_NPM = NPM %>%
   mutate(Fish_or_Parts = case_when(stri_detect_fixed(StomachContents, "Fish") ~ "Fish or Fish Parts",
                                    stri_detect_fixed(StomachContents, "Empty") ~ "Empty", TRUE ~ "Other")) %>%
   drop_na(Length)
+
+#Read in teh Bioenergetics
+
+bioenergetics<- read_csv("analysis/data/raw data/2020 Bioenergetics 1 year.csv") %>%
+  mutate(Date = as.POSIXct(Date, format = "%m/%d/%y")) %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g))
+
+#Fall bioenergetics looking at 78 days
+Fall_bioenergetics = bioenergetics %>%
+  filter(Date > "2020-08-31") %>%
+  filter(Date < "2020-11-18") %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g))
+
+#Plot of fish eatten
+Fall_bioenergetics %>%
+  filter(Date >= "2020-09-01 06:00:00" & Date <= "2020-11-17 07:00:00") %>%
+  ggplot() +
+  aes(x = Date, y = Cumu_fish_eatten, colour = Temperature_C) +
+  geom_line(size = 2.5) +
+  scale_color_viridis_c(option = "viridis", direction = 1) +
+  labs(y = "Cumulative fish eatten g") +
+  theme_classic()
+
+
+
+esquisse::esquisser(Fall_bioenergetics)
 
 
 

### 2021-11-09 analysis/R scripts/Nick Script 11.2.21 cleaning fish data.R
Added figures folder in the paper folder
Added figures for the bioenergetics and a length frequency plot. There are 3 different plots one for a full year and no growth, one for a full year and NPM 10% growth, and one for no growth and only the 78 day fall season.

@@ -3,8 +3,6 @@
 # 11/2/20201
 rm(list = ls()) # clear the working environment
 
-
-
 require(tidyverse)
 require(lubridate)
 require(boot)
@@ -19,8 +17,8 @@ require(car)
 require(dplyr)
 require(plotrix)
 require(nnet)
-require(Hmisc)
-require(Matching)
+#require(Hmisc) # does not work
+#require(Matching)
 require(Rcapture)
 require(nlstools)
 require(stringi)
@@ -39,33 +37,82 @@ Clean_NPM = NPM %>%
                                    stri_detect_fixed(StomachContents, "Empty") ~ "Empty", TRUE ~ "Other")) %>%
   drop_na(Length)
 
-#Read in teh Bioenergetics
+#Read in the Bioenergetics
 
 bioenergetics<- read_csv("analysis/data/raw data/2020 Bioenergetics 1 year.csv") %>%
   mutate(Date = as.POSIXct(Date, format = "%m/%d/%y")) %>%
-  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g))
+  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+  mutate(Date = as.Date(Date)) %>%
+  mutate(month = month(Date))
+
+bioenergetics_10percet <- read_csv("analysis/data/raw data/2020 Bioenergetics 1 year 10% growth.csv") %>%
+mutate(Date = as.POSIXct(Date, format = "%m/%d/%y")) %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+  mutate(Date = as.Date(Date)) %>%
+  mutate(month = month(Date))
+
 
 #Fall bioenergetics looking at 78 days
 Fall_bioenergetics = bioenergetics %>%
   filter(Date > "2020-08-31") %>%
   filter(Date < "2020-11-18") %>%
-  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g))
+  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+  mutate(Date = as.Date(Date))
 
 #Plot of fish eatten
-Fall_bioenergetics %>%
-  filter(Date >= "2020-09-01 06:00:00" & Date <= "2020-11-17 07:00:00") %>%
+fall_bioenergetics = Fall_bioenergetics %>%
[…]

### 2021-11-09 analysis/paper/paper.Rmd

@@
 
-The Deadwater Slough is an approximately two kilometer (km) section of the mainstem Salmon River located rough 5.8 river kms downstream from North Fork, Idaho (publication quality map referenced here).
-The downstream end of the Deadwater Slough occurs at the confluence of Dump Creek and the Salmon River.
-Around 1897, a small mining diversion reservoir failed in the Dump Creek drainage, creating upslope instability, causing a large erosion event that deposited large amounts of sediment at the confluence of the Salmon River and Dump Creek (@Emerson1973).
-The event created a large alluvial fan which backed up the Salmon River creating an approximately 25.3 acre (maybe convert to square meters or kilometers?) reservoir-like area that is unnaturally slow and deep (map reference or image here).
-The river in this section has slower water velocities, a deepened channel, and (anecdotally) warmer water temperatures which can favor piscivorous fishes (e.g., northern pikeminnow *Ptychocheilus oregonensis*, smallmouth bass *Micropterus dolomieu*).
+The Deadwater Slough is an approximately two-kilometer (km) section of the mainstem Salmon River locations roughly 5.8 river kms downstream from the town of North Fork, Idaho ***(publication quality map referenced here)***.
+The downstream end of the Deadwater Slough is located at the confluence of Dump Creek and the Salmon River.
+Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River and Dump Creek [@Emerson1973].
+The influx of sediment created a delta that impeded flow of the Salmon River, thereby creating an unnaturally slow and deep section in the river, spanning approximately 25.3 acres.
+The slower water velocities, deepened channel, and (anecdotally) warmer water temperatures favor piscivorous fishes known to occupy the reach (e.g., northern pikeminnow *Ptychocheilus oregonensis*, smallmouth bass *Micropterus dolomieu*).
 

### 2021-11-09 analysis/R scripts/Nick Script 11.2.21 cleaning fish data.R
updates

@@ -35,7 +35,13 @@ NPM = deadwater %>%
 Clean_NPM = NPM %>%
   mutate(Fish_or_Parts = case_when(stri_detect_fixed(StomachContents, "Fish") ~ "Fish or Fish Parts",
                                    stri_detect_fixed(StomachContents, "Empty") ~ "Empty", TRUE ~ "Other")) %>%
-  drop_na(Length)
+  drop_na(Length) %>%
+  mutate(Non_fish_wt = StomachContentsWeight - FishContentWeight)
+
+Stomach_content_DF = Clean_NPM %>%
+  dplyr::select(Date, Length, LavageID, StomachContents, Comments, Fish_or_Parts, StomachContentsWeight, Non_fish_wt, FishContentWeight)
+
+unique(Clean_NPM)
 
 #Read in the Bioenergetics
 
@@ -114,22 +120,32 @@ Bioenergetics_plot_10percent
 
 #esquisse::esquisser(Fall_bioenergetics)
 
-#Histogram Ogle
+# Stomach contents and length frequency ############################################################
 
 Clean_NPM$NPM_50=lencat(Clean_NPM$Length, w = 50)
 
+Stomach_catagories_sum = Stomach_content_DF %>%
+  drop_na(StomachContentsWeight) %>%
+  group_by(Fish_or_Parts) %>%
+  count()
+
+Stomach_wt_sum = Stomach_content_DF %>%
+  drop_na(StomachContentsWeight) %>%
+  group_by(Fish_or_Parts) %>%
+  summarise(sum(StomachContentsWeight))
+
+Non_fishwt_sum = Stomach_content_DF %>%
+  drop_na(StomachContentsWeight) %>%
+  group_by(Fish_or_Parts) %>%
+  summarise(sum(Non_fish_wt))
+
+Fish_weights_sum = Stomach_content_DF %>%
+  drop_na(StomachContentsWeight) %>%
+  group_by(Fish_or_Parts) %>%
+  summarise(sum(FishContentWeight))
+
+
 #esquisse::esquisser(Clean_NPM)
-
-
-##group_by(Species) %>% 165,640,25
-#custbins=seq(150,675,25)
-#hist(~Length,data=Clean_NPM,xlab="TL (mm)", ylim=c(0,500))
-##xtick<-seq(150,650,20)####creating minor tick marks
-#axis(side=1, at=xtick, labels = FALSE)####setting minor tick marks
-
-
-
-
 
[…]

### 2021-11-09 analysis/paper/paper.Rmd
added map and updated site descriptions

@@
 Moreover, this reach is part of the migratory pathway for juvenile DSR and NRR emigrants from all eight extant populations.
-The slough lacks hydrological and structural features (i.e., homogeneous channel with fine substrate) that can provide essential refuge from predation.
+The slough lacks hydrological and structural features (i.e., a homogenous channel with fine substrate and little cover) that can provide essential refuge from predation.
 As a result, predation on juvenile Chinook salmon proximal to the Deadwater Slough has been cited as a concern for the Upper Salmon River MPG, impacting DSR migrants in the fall and NRR emigrants during the spring.
@@
 
-<!-- The following was pulled directly from a technical report, but probably could/should be improved for the manuscript -->
-
-The Deadwater Slough is an approximately two-kilometer (km) section of the mainstem Salmon River locations roughly 5.8 river kms downstream from the town of North Fork, Idaho ***(publication quality map referenced here)***.
-The downstream end of the Deadwater Slough is located at the confluence of Dump Creek and the Salmon River.
-Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River and Dump Creek [@Emerson1973].
-The influx of sediment created a delta that impeded flow of the Salmon River, thereby creating an unnaturally slow and deep section in the river, spanning approximately 25.3 acres.
-The slower water velocities, deepened channel, and (anecdotally) warmer water temperatures favor piscivorous fishes known to occupy the reach (e.g., northern pikeminnow *Ptychocheilus oregonensis*, smallmouth bass *Micropterus dolomieu*).
+The Deadwater Slough is an approximately 1.5 kilometer section of the mainstem Salmon River located roughly 5.8 river kilometers downstream from the town of North Fork, Idaho ***(publication quality map referenced here)***.
+The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
+Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River, thereby creating an unnaturally slow and deep section in the river, spanning approximately 30 acres and averaging 68 m width.
+Both northern pikeminnow *Ptychocheilus oregonensis* and smallmouth bass *Micropterus dolomieu* inhabit the slough which likely provides favorable conditions for their feeding and growth (e.g., reduced water velocity, deep channel, warmer water temperature).
 
@@
 
-# References
+# Literature Cited
 
@@
 <div id="refs"></div>
+
+\newpage
+
+# Tables
+
+<div id="tables"></div>
+
+\newpage
+
+# Figures
+
+<div id="figures"></div>
+
+```{r map, out.width = "85%", fig.cap = "Map of the Deadwater Slough study area.
+The high-resolution orthoimage portion directly surrounding the Deadwater Slough was generated from aerial images taken by a drone.
+The red polygon indicates the reach characterized by unnaturally slow water velocities and a deepened channel.
+The location of the Dump Creek delta is indicated."}
+knitr::include_graphics("../figures/deadwater_map.png")
+
+```
 

### 2021-11-09 analysis/paper/paper.Rmd
updated data filepaths in paper for "raw_data" change

@@
 ```{r cmr-effort-data}
-cmr_df = read_csv("../data/raw data/deadwater_cmr_effort_20210614.csv") %>%
+cmr_df = read_csv("../data/raw_data/deadwater_cmr_effort_20210614.csv") %>%
   clean_names()
@@
 ```{r shoup-temp-data}
-temp_df = read_tsv("../data/raw data/USGS_13307000_Salmon_River_Nr_Shoup.txt",
+temp_df = read_tsv("../data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt",
                    comment = "#") %>%
@@
 The location of the Dump Creek delta is indicated."}
-knitr::include_graphics("../figures/deadwater_map.png")
+knitr::include_graphics("../figures/deadwater_map_low_res.png")
 

### 2021-11-09 analysis/paper/paper.Rmd
added inline figure reference

@@
 
-The Deadwater Slough is an approximately 1.5 kilometer section of the mainstem Salmon River located roughly 5.8 river kilometers downstream from the town of North Fork, Idaho ***(publication quality map referenced here)***.
+The Deadwater Slough is an approximately 1.5 kilometer section of the mainstem Salmon River located roughly 5.8 river kilometers downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
 The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.

### 2021-11-09 analysis/R scripts/Nick Script 11.2.21 cleaning fish data.R
Updates for Bioenergetics

@@ -65,6 +65,15 @@ Fall_bioenergetics = bioenergetics %>%
   mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
   mutate(Date = as.Date(Date))
 
+Alternative_bioenergetics <- read_csv("analysis/data/raw data/Bioenergetics 32 fish 68 inverts.csv") %>%
+  mutate(Date = as.POSIXct(Date, format = "%m/%d/%y")) %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+  mutate(Date = as.Date(Date)) %>%
+  mutate(month = month(Date))
+
+
+
+
 #Plot of fish eatten
 fall_bioenergetics = Fall_bioenergetics %>%
   filter(Date >= "2020-09-01" & Date <= "2020-11-17") %>%
@@ -116,7 +125,21 @@ Bioenergetics_plot_10percent <- bioenergetics_10percet %>%
 
 Bioenergetics_plot_10percent
 
+Alt_Bioenergetics_plot <- Alternative_bioenergetics %>%
+  ggplot() +
+  aes(x = Date, y = Cumu_fish_eatten, colour = Temperature_C) +
+  geom_line(size = 2.5) +
+  scale_color_distiller(palette = "Spectral", direction = -1) +
+  #labs(y = "Cumulative fish eatten (g)") +
+  #scale_x_date(breaks = "1 month") +
+  scale_y_continuous(name = "Cumulative fish eatten (g)", breaks = seq(0,600, by = 50),
+                     sec.axis = sec_axis(~. /12, name = "Potential number of Chinook smolts", breaks = seq(0,30, by = 2))) +
+  #theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
+  theme_classic() +
+  theme(legend.position = "top") +
+  labs(color = "Temperature (C)", x = "Month")
 
+Alt_Bioenergetics_plot
 
 #esquisse::esquisser(Fall_bioenergetics)
 

### 2021-11-09 analysis/R scripts/Nick Script 11.2.21 cleaning fish data.R
I hope last update
Adjusted bioenergetics inputs for more accurate graph.

@@ -34,7 +34,9 @@ NPM = deadwater %>%
 
 Clean_NPM = NPM %>%
   mutate(Fish_or_Parts = case_when(stri_detect_fixed(StomachContents, "Fish") ~ "Fish or Fish Parts",
-                                   stri_detect_fixed(StomachContents, "Empty") ~ "Empty", TRUE ~ "Other")) %>%
+                                   stri_detect_fixed(StomachContents, "Empty") ~ "Empty",
+                                   stri_detect_fixed(Comments, "Whole scuplin") ~ "Fish or Fish Parts",
+                                   TRUE ~ "Other")) %>%
   drop_na(Length) %>%
   mutate(Non_fish_wt = StomachContentsWeight - FishContentWeight)
 
@@ -57,21 +59,18 @@ mutate(Date = as.POSIXct(Date, format = "%m/%d/%y")) %>%
   mutate(Date = as.Date(Date)) %>%
   mutate(month = month(Date))
 
-
-#Fall bioenergetics looking at 78 days
-Fall_bioenergetics = bioenergetics %>%
-  filter(Date > "2020-08-31") %>%
-  filter(Date < "2020-11-18") %>%
-  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
-  mutate(Date = as.Date(Date))
-
 Alternative_bioenergetics <- read_csv("analysis/data/raw data/Bioenergetics 32 fish 68 inverts.csv") %>%
   mutate(Date = as.POSIXct(Date, format = "%m/%d/%y")) %>%
   mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
   mutate(Date = as.Date(Date)) %>%
   mutate(month = month(Date))
 
-
+#Fall bioenergetics looking at 78 days
+Fall_bioenergetics = Alternative_bioenergetics %>%
+  filter(Date > "2020-08-31") %>%
+  filter(Date < "2020-11-18") %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+  mutate(Date = as.Date(Date))
 
 
 #Plot of fish eatten
@@ -202,6 +201,8 @@ Bioenergetics_plot
 Bioenergetics_plot_10percent
 fall_bioenergetics
 
+Alt_Bioenergetics_plot
+
 # Save fish passing plot
 ggsave('analysis/paper/figures/length_freq.jpg',
        length_freq,
[…]

### 2021-11-09 analysis/paper/paper.Rmd
typo

@@
 
-We hypothesize that increased densitied of piscivorous predators in the Deadwater Slough area may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon [@Ackerman2018] and Sockeye Salmon [@Axel2015].
+We hypothesize that increased densities of piscivorous predators in the Deadwater Slough area may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon [@Ackerman2018] and Sockeye Salmon [@Axel2015].
 In this study, we estimated the abundance of a piscivorous fish predator population in the Deadwater Slough and their potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.

### 2021-11-10 analysis/R scripts/Nick Script 11.2.21 cleaning fish data.R
updating script

@@ -26,7 +26,7 @@ require(MASS)
 require(readr)
 
 #Read in the Deadwater dataset
-deadwater <- read_csv("analysis/data/raw data/deadwater_cmr_effort_20210614.csv") %>%
+deadwater <- read_csv("analysis/data/raw data/deadwater_cmr_effort_20210614.csv") #%>%
   mutate(Date = as.POSIXct(Date, format = "%m/%d/%y"))
 
 NPM = deadwater %>%

### 2021-11-10 analysis/R scripts/Nick Script 11.10.21 cleaning fish data.R
Nicks Script for bioenergetics and length freq

@@ -1,0 +1,233 @@
+# working to clean up and provide summaries for Deadwater Data
+# Nick Porter
+# 11/2/20201
+rm(list = ls()) # clear the working environment
+
+require(tidyverse)
+require(lubridate)
+require(boot)
+require(bootstrap)
+require(lme4)
+require(Matrix)
+require(numDeriv)
+require(FSA)
+require(FSAdata)
+require(magrittr)
+require(car)
+require(dplyr)
+require(plotrix)
+require(nnet)
+#require(Hmisc) # does not work
+#require(Matching)
+require(Rcapture)
+require(nlstools)
+require(stringi)
+require(MASS)
+require(readr)
+
+#Read in the Deadwater dataset
+deadwater <- read_csv("analysis/data/raw_data/deadwater_cmr_effort_20210614.csv") %>%
+  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
+
+
+NPM = deadwater %>%
+  filter(Species == "Northern Pikeminnow")
+
+Clean_NPM = NPM %>%
+  mutate(Fish_or_Parts = case_when(stri_detect_fixed(StomachContents, "Fish") ~ "Fish or Fish Parts",
+                                   stri_detect_fixed(StomachContents, "Empty") ~ "Empty",
+                                   stri_detect_fixed(Comments, "Whole scuplin") ~ "Fish or Fish Parts",
+                                   TRUE ~ "Other")) %>%
+  drop_na(Length) %>%
+  mutate(Non_fish_wt = StomachContentsWeight - FishContentWeight)
+
+Stomach_content_DF = Clean_NPM %>%
+  dplyr::select(Date, Length, LavageID, StomachContents, Comments,
+                Fish_or_Parts, StomachContentsWeight, Non_fish_wt, FishContentWeight) %>%
+  filter(!StomachContents %in% c("Lost sample CG27-001","Did Not Lavage")) %>%
+  drop_na(StomachContents)
+
+
+#Read in the Bioenergetics
+
+bioenergetics<- read_csv("analysis/data/raw_data/2020 Bioenergetics 1 year.csv") %>%
+  mutate(Date = as.Date(Date, format = "%m/%d/%y")) %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+  mutate(Date = as.Date(Date)) %>%
+  mutate(month = month(Date))
+
[…]

### 2021-11-10 analysis/R scripts/Nick Script 11.10.21 cleaning fish data.R
Working on Code and updating stomach contents per size class

@@ -37,6 +37,7 @@ Clean_NPM = NPM %>%
   mutate(Fish_or_Parts = case_when(stri_detect_fixed(StomachContents, "Fish") ~ "Fish or Fish Parts",
                                    stri_detect_fixed(StomachContents, "Empty") ~ "Empty",
                                    stri_detect_fixed(Comments, "Whole scuplin") ~ "Fish or Fish Parts",
+                                   stri_detect_fixed(StomachContents, "Whitefish") ~ "Fish or Fish Parts",
                                    TRUE ~ "Other")) %>%
   drop_na(Length) %>%
   mutate(Non_fish_wt = StomachContentsWeight - FishContentWeight)
@@ -167,6 +168,29 @@ Fish_weights_sum = Stomach_content_DF %>%
   drop_na(StomachContentsWeight) %>%
   group_by(Fish_or_Parts) %>%
   summarise(sum(FishContentWeight))
+
+#determining fish spp vs unknown
+
+Num_Hungry_fish = Stomach_content_DF %>%
+  filter(!Fish_or_Parts == "Empty") %>%
+  group_by(Fish_or_Parts) %>%
+  count()
+
+Num_Hungry_fish_YR = Stomach_content_DF %>%
+  filter(!Fish_or_Parts == "Empty") %>%
+  mutate(Year = year(Date)) %>%
+  group_by(Year, Fish_or_Parts) %>%
+  count()
+
+Hungry_fish_DF = Stomach_content_DF %>%
+  filter(Fish_or_Parts == "Fish or Fish Parts") %>%
+  mutate(Fish = case_when(stri_detect_fixed(StomachContents, "Shiner") ~ "Redside Shiner",
+                          stri_detect_fixed(StomachContents, "Whitefish") ~ "Mountin Whitefish",
+                          stri_detect_fixed(StomachContents, "scuplin") ~ "Sculpin",
+                          stri_detect_fixed(StomachContents, "Chinook") ~ "Chinook",
+                          stri_detect_fixed(Comments, "whitefish") ~ "Mountin Whitefish",
+                          stri_detect_fixed(Comments, "scuplin") ~ "Sculpin",
+                          stri_detect_fixed(Comments, "chinook") ~ "Chinook", TRUE ~ "Other"))
 
 
 #esquisse::esquisser(Clean_NPM)

### 2021-11-11 analysis/R scripts/abund_est.R
added updated data file from Mike Hall. Started some chunks to estimate abundance and effort (CPUE) in the manuscript Rmd file.

@@ -43,22 +43,6 @@ raw_df %>%
   # tabyl(date, capture_type)
   tabyl(fish_status)
 
-# calculate effort
-raw_df %>%
-  select(event_name,
-         site,
-         date,
-         crew,
-         method,
-         equipment,
-         start_time:total_effort,
-         total_day_effort) %>%
-  distinct() %>%
-  group_by(event_name,
-           date) %>%
-  summarise(across(total_effort,
-                   sum),
-            .groups = "drop")
 
 # using Schnabel model
 N_mods = raw_df %>%
@@ -215,3 +199,60 @@ LP_mod %>%
   mutate(range = Uci - Lci,
          prop_range = range / N) %>%
   arrange(event_name, model)
+
+#-----------------------------------------------------------------
+# CPUE
+#-----------------------------------------------------------------
+
+# calculate effort
+cpue_df = raw_df %>%
+  select(event_name,
+         site,
+         date,
+         crew,
+         method,
+         equipment,
+         start_time:total_effort,
+         total_day_effort) %>%
+  distinct() %>%
+  group_by(event_name,
+           date) %>%
+  summarise(across(total_effort,
+                   sum),
+            .groups = "drop") %>%
+  # add how many pikeminnow were caught
+  left_join(raw_df %>%
+              filter(species == "Northern Pikeminnow") %>%
+              filter(capture_type != "Recapture") %>%
+              group_by(event_name, species, date) %>%
+              summarise(n_fish = n(),
+                        .groups = "drop")) %>%
+  mutate(cpue = n_fish / total_effort)
+
+cpue_df %>%
+  ggplot(aes(x = date,
+             y = cpue,
+             color = event_name)) +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~ event_name,
+             scales = 'free_x')
+
+
+cpue_df %>%
+  group_by(species, event_name) %>%
+  summarise(across(c(total_effort, n_fish),
+                   sum),
+            .groups = "drop") %>%
+  mutate(across(total_effort,
+                ~ . / 60)) %>%
+  mutate(cpue = n_fish / total_effort) %>%
+  ggplot(aes(x = event_name,
+             y = cpue,
+             fill = event_name)) +
+  geom_col() +
+  scale_fill_brewer(palette = "Set1",
+                    name = "Event") +
+  labs(x = "Event",
+       y = "CPUE (Pikeminnow caught per hour)") +
+  theme(legend.position = "none")

### 2021-11-11 analysis/R scripts/Nick Script 11.10.21 cleaning fish data.R
updates to bioenergetics

@@ -69,16 +69,25 @@ Alternative_bioenergetics <- read_csv("analysis/data/raw_data/Bioenergetics 32 f
   mutate(Date = as.Date(Date)) %>%
   mutate(month = month(Date))
 
+sevendyeight_bioenergetics <- read_csv("analysis/data/raw_data/78_day_bioenergetics.csv") %>%
+  mutate(Date = as.Date(Date, format = "%m/%d/%y")) %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+  mutate(Date = as.Date(Date)) %>%
+  mutate(month = month(Date))
+
 #Fall bioenergetics looking at 78 days
 Fall_bioenergetics = Alternative_bioenergetics %>%
   filter(Date >= "2020-09-01" & Date <= "2020-11-17") %>%
   mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
   mutate(Date = as.Date(Date))
 
+#Fall bioenergetics looking at 78 days
+Fall_bioenergetics_2 = sevendyeight_bioenergetics %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+  mutate(Date = as.Date(Date))
 
 #Plot of fish eatten
-fall_bioenergetics = Fall_bioenergetics %>%
-  filter(Date >= "2020-09-01" & Date <= "2020-11-17") %>%
+fall_bioenergetics = Fall_bioenergetics_2 %>%
   ggplot() +
   aes(x = Date, y = Cumu_fish_eatten, colour = Temperature_C) +
   geom_line(size = 2.5) +

### 2021-11-11 analysis/paper/paper.Rmd

@@
 ```{r cmr-effort-data}
-cmr_df = read_csv("../data/raw_data/deadwater_cmr_effort_20210614.csv") %>%
-  clean_names()
+# cmr_df = read_csv(here("analysis/data/raw_data/deadwater_cmr_effort_20210614.csv")) %>%
+cmr_df = readxl::read_excel(here("analysis/data/raw_data/Deadwater_predation_mark_recapture_11.09.21.xlsx")) %>%
+  clean_names() %>%
+  filter(count > 0) %>%
+  # pull out only Northern Pikeminnow
+  filter(species == "Northern Pikeminnow") %>%
+  # mutate(across(date,
+  #               mdy)) %>%
+  mutate(across(date,
+                ymd)) %>%
+  mutate(across(c(start_time, end_time),
+                ~ ymd_hms(paste(year(date), month(date), day(date), hour(.), minute(.), second(.))))) %>%
+  # split into sampling periods
+  mutate(event_name = if_else(month(date) > 8,
+                              paste("Fall", year(date), sep = "_"),
+                              paste("Spring", year(date), sep = "_")))
 
+effort_df = cmr_df %>%
+  select(event_name,
+         site,
+         date,
+         crew,
+         method,
+         equipment,
+         start_time:total_effort,
+         total_day_effort) %>%
+  distinct() %>%
+  group_by(event_name,
+           date) %>%
+  summarise(across(total_effort,
+                   sum),
+            .groups = "drop")
+
+cpue_df = effort_df %>%
+  # add how many pikeminnow were caught
+  left_join(cmr_df %>%
+              filter(capture_type != "Recapture") %>%
+              group_by(event_name, species, date) %>%
+              summarise(n_fish = n(),
+                        .groups = "drop")) %>%
+  mutate(cpue = n_fish / total_effort)
+
+```
+
+```{r lincoln-petersen-estimator}
+# using Lincoln-Petersen model, combining first 4 and last four days
+# define mark and recapture periods
+cap_periods = tibble(event_name = "Fall_2019",
+                     cap_period = "Mark",
+                     int = interval(start = ymd('20191112'),
+                                    end = ymd('20191114'))) %>%
+  bind_rows(tibble(event_name = "Fall_2019",
+                   cap_period = "Recapture",
+                   int = interval(start = ymd('20191119'),
+                                  end = ymd('20191121')))) %>%
+  bind_rows(tibble(event_name = "Fall_2020",
+                   cap_period = "Mark",
+                   int = interval(start = ymd('20201020'),
+                                  end = ymd('20201023')))) %>%
+  bind_rows(tibble(event_name = "Fall_2020",
+                   cap_period = "Recapture",
+                   int = interval(start = ymd('20201027'),
+                                  end = ymd('20201030'))))
+
+cmr_df$cap_period = NA_character_
+for(i in 1:nrow(cap_periods)) {
+  my_int = cap_periods$int[i]
+  index = which(cmr_df$date %within% my_int)
+  cmr_df$cap_period[index] = cap_periods$cap_period[i]
+}
+
+LP_mod = cmr_df %>%
+  filter(species == "Northern Pikeminnow") %>%
+  # filter out spring 2021 (no recapture)
+  filter(event_name != "Spring_2021") %>%
+  mutate(mr_type = if_else(cap_period == "Mark" & fish_status == "Live",
+                           "M",
+                           if_else(cap_period == "Recapture",
+                                   if_else(capture_type == "Mark",
+                                           "C",
+                                           if_else(capture_type == "Recapture",
+                                                   "R",
+                                                   NA_character_)),
+                                   NA_character_))) %>%
+  filter(!is.na(mr_type)) %>%
+  group_by(event_name, species) %>%
+  summarise(M = sum(mr_type == "M"),
+            n = sum(mr_type %in% c("C", "R")),
+            m = sum(mr_type == 'R'),
+            .groups = "drop") %>%
+  nest(data = -c(event_name, species)) %>%
+  crossing(model = c("Petersen",
+                     "Chapman")) %>%
+  mutate(mr_mod = map2(data,
+                       model,
+                      .f = function(x, y) {
+                        with(x,
+                             mrClosed(M = M,
+                                      n = n,
+                                      m = m,
+                                      method = y))
+                      }),
+         N_summ = map(mr_mod,
+                      .f = summary,
+                      incl.SE = T),
+         CI = map(mr_mod,
+                  .f = confint)) %>%
+  mutate(N = map_dbl(N_summ,
+                     .f = function(x) x[1]),
+         SE = map_dbl(N_summ,
+                      .f = function(x) x[2]),
+         N.lci = map_dbl(CI,
+                         .f = function(x) x[1]),
+         N.uci = map_dbl(CI,
+                         .f = function(x) x[2])) %>%
+  select(-N_summ, -CI)
+```
+
+```{r schnabel-estimator}
+# using Schnabel model
+N_mods = cmr_df %>%
+  filter(species == "Northern Pikeminnow") %>%
+  # filter out spring 2021 (no recapture)
+  filter(event_name != "Spring_2021") %>%
+  group_by(event_name, species, date, fish_status, capture_type) %>%
+  summarise(n_fish = n(),
+            .groups = "drop") %>%
+  group_by(event_name, species, date) %>%
+  summarise(n = sum(n_fish),
+            m = sum(n_fish[capture_type == "Recapture"]),
+            R = sum(n_fish[fish_status == "Live"]),
+            dead_fish = sum(n_fish[fish_status == "Dead"]),
+            .groups = "drop") %>%
+  group_by(event_name, species) %>%
+  nest() %>%
+  mutate(mod = map(data,
+                   .f = function(x) {
+                     with(x,
+                          mrClosed(n = n,
+                                   m = m,
+                                   R = R,
+                                   method = "Schnabel",
+                                   chapman.mod = TRUE))
+                   }),
+         N = map_dbl(mod,
+                     .f = summary),
+         CI = map(mod,
+                  .f = confint))
+
+N_mods %>%
+  select(species, N, CI) %>%
+  mutate(N.lci = map_dbl(CI,
+                         .f = function(x) x[1]),
+         N.uci = map_dbl(CI,
+                         .f = function(x) x[2])) %>%
+  select(-CI)
 ```

### 2021-11-12 analysis/paper/paper.Rmd
edits to Discussion; added 5 packages to load-libraries

@@
 library(magrittr)
+library(lubridate)
+library(tidytable)
 library(janitor)
+library(purrr)
+library(here)
+library(FSA)
 
@@
 
-The Upper Salmon River major population group (MPG) supports eight extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
+The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
 At least five of these eight populations must meet criteria set forth by @McElhany2000 and @ICTRT2007 for the MPG to be considered viable and for the recovery of the Snake River Evolutionary Significant Unit (ESU).
@@
 
-```{r schnabel-estimator}
+```{r schnabel-estimator, eval = F}
 # using Schnabel model
@@
 
-We estimated the population size of northern pikeminnow in the Deadwater Slough and during the fall emigration period to be greater than ***xx,xxx***.
-And although we did not attempt to estimate population size during the spring emigration period, the relative abundance measured as CPUE was comparable to that during the fall sampling periods (update statement later).
-That fall population size estimate translates to ***xxx*** northern pikeminnow per 100 m or ***xxx*** pikeminnow per 100 m^2^.
-For context, that is similar/more/less than estimates from the Columbia River where predation impacts have been demonstrated as substantial and bounty programs have been implemented to reduce northern pikeminnow abundance.
-Given that estimate of abundance and some assumptions, we estimated that northern pikeminnow potentially consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook salmon during the *x* sampling periods resulting in a reduction of adult returns between ***xxx*** and ***x,xxx***.
-We suggest that modifications to the habitat due to human activities in the Deadwater Reach has resulted in favorable conditions for northern pikeminnow, including improving conditions for predation upon juvenile Chinook salmon, which has resulted in a consequential impact to ESA-listed Chinook salmon populations in the Upper Salmon River MPG.
+We estimated the population size of Northern Pikeminnow in the Deadwater Slough to be greater than ***xx,xxx*** during the fall emigration period for DSR Chinook Salmon.
+That estimate translates to a density of ***xxx*** Northern Pikeminnow per 100 m or ***xxx*** per 100 m^2^ which is ***similar/more/less*** than estimates from elsewhere in the Columbia River ***(citation)*** where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminno abundance.
+The population size of Northern Pikeminnow was not directly estimated during the spring NRR Chinook salmon emigration period however, the relative abundance measured at CPUE was comparable to the fall sampling periods ***(update statement later)***.
+The population of Northern Pikeminnow in Deadwater Slough was estimated to consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook Salmon during the ***x*** sampling periods and result in an estimated reduction of returning adults between ***xxx*** and ***x,xxx***.
+We suggest that the habitat modifications that created the Deadwater Slough reach have resulted in favorable conditions for Northern Pikeminnow, including improved conditions for predation upon juvenile Chinook Salmon ***(add detail here)***.
+Therefore, predation by Northern Pikeminnow in the Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG.
 
@@
 
-Although not formally assessed here, another potential source of predation upon juvenile salmon in the Deadwater Slough is from avian predators including great blue herons *Ardea herodias* and bald eagles *Haliaeetus leucocephalus*.
+Although not formally assessed in this study, avian predators including Great Blue Herons *Ardea herodias* and Bald Eagles *Haliaeetus leucocephalus* are another potential source of mortality for juvenile salmon in the Deadwater Slough.
 The Deadwater Slough is recognized as an important bird watching and nesting area due to the riparian and backwater habitats created by the feature [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
-During the initial sampling event in 2019, a two-person crew equipped with a mobile passive integrated transponder (PIT) tag antenna walked the entire reach and surrounding and upstream areas scanning for PIT tags.
-During the informal survey, we did identify nine PIT tags on surfaces surrounding Deadwater Slough, many of which were near nests or in an upstream anastomizing reach where herons and eagles are prevalent.
-Investigating their tag histories in [PTAGIS](https://www.ptagis.org/), these tags were implanted into a combination of juvenile Chinook Salmon (3), Sockeye Salmon (3), and steelhead (3) suggesting avian predation does occur at and adjacent to the Deadwater Reach, although to an unknown degree.
-***Follow up with a brief lit review on avian predation on juvenile Chinook salmon elsewhere***
+Several piscivorous bird species have been documented using the Deadwater Slough that include the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
+During the intial sampling event in 2019, a two-person crew walked the entire reach and surrounding and upstream areas scanning for passive integrated transponder (PIT) tags.
+During that informal survey, nine PIT tags were recovered near active bird nests or in an upstream anastomizing reach where herons and eagles are prevalent, suggesting that mortality may have been a result of avian predation.
+The PIT tag histories in [PTAGIS](https://www.ptagis.org/) indicate these tags were implanted into a combination of juvenile Chinook Salmon (3), Sockeye Salmon (3), and steelhead (3).
+Avian predation contributes a major component of the total mortality for yearling Chinook Salmon in some locations in the lower Snake River and Columbia River, particularly at hydroelectric dams and within reservoirs [@Evans2012, @Evans2016]; however, we did not observe large colonies of piscivorous birds within the study area.
+Although there is documentation of individual Double-crested Cormorants [@eBird2021] at the Deadwater Slough, the site is not within their breeding range, rather, it is part of a migration corridor.
+Given the current avian species known to occupy Deadwater Slough, it is unlikely that avian predation on juvenile salmonids is comparable to elsewhere in the Columbia River basin with large piscivorous bird colonies.
+Nevertheless, we hypothesize that the reservoir-like conditions at the Deadwater Slough may increase the probability of avian predation on juvenile Chinook Salmon from the many piscivorous birds known to use the site.
+Future estimates of predation would benefit from consideration of the contribution of piscivorous avian predators.
 

### 2021-11-12 analysis/paper/paper.Rmd
added TB as author
added citations for discussion; updated MA bibliography

@@
       institute: [biomark]
+      correspondence: false
+  - Tara E. Blackman:
+      email: tara.blackman@mthoodenvironmental.com
+      institute: [mhe]
       correspondence: false
@@
 The PIT tag histories in [PTAGIS](https://www.ptagis.org/) indicate these tags were implanted into a combination of juvenile Chinook Salmon (3), Sockeye Salmon (3), and steelhead (3).
-Avian predation contributes a major component of the total mortality for yearling Chinook Salmon in some locations in the lower Snake River and Columbia River, particularly at hydroelectric dams and within reservoirs [@Evans2012, @Evans2016]; however, we did not observe large colonies of piscivorous birds within the study area.
+Avian predation contributes a major component of the total mortality for yearling Chinook Salmon in some locations in the lower Snake River and Columbia River, particularly at hydroelectric dams and within reservoirs [@Evans2012; -@Evans2016]; however, we did not observe large colonies of piscivorous birds within the study area.
 Although there is documentation of individual Double-crested Cormorants [@eBird2021] at the Deadwater Slough, the site is not within their breeding range, rather, it is part of a migration corridor.

### 2021-11-12 analysis/paper/paper.Rmd
updated chunks to estimate abundance and CPUE, added some tables and figures showing results.

@@
       - --lua-filter=../templates/pagebreak.lua
+    bookdown::html_document2:
+      fig_caption: yes
+      number_sections: FALSE
+      toc: no
+      fig_width: 6
+      pandoc_args:
+      - --lua-filter=../templates/scholarly-metadata.lua
+      - --lua-filter=../templates/author-info-blocks.lua
+      - --lua-filter=../templates/pagebreak.lua      
 fontsize: 12pt
@@
 library(FSA)
+library(knitr)
+library(kableExtra)
+
+# format NA values in tables
+options(knitr.kable.NA = '-')
+
+# set default theme for figures
+theme_set(theme_bw())
 
@@
 ```{r cmr-effort-data}
-# cmr_df = read_csv(here("analysis/data/raw_data/deadwater_cmr_effort_20210614.csv")) %>%
-cmr_df = readxl::read_excel(here("analysis/data/raw_data/Deadwater_predation_mark_recapture_11.09.21.xlsx")) %>%
+raw_df = readxl::read_excel(here("analysis/data/raw_data/Deadwater_predation_mark_recapture_11.09.21.xlsx")) %>%
   clean_names() %>%
-  filter(count > 0) %>%
-  # pull out only Northern Pikeminnow
-  filter(species == "Northern Pikeminnow") %>%
-  # mutate(across(date,
-  #               mdy)) %>%
   mutate(across(date,
@@
 
-effort_df = cmr_df %>%
+# focus on pikeminnow
+cmr_df = raw_df %>%
+  filter(count > 0) %>%
+  # pull out only Northern Pikeminnow
+  filter(species == "Northern Pikeminnow") %>%
+  mutate(mark_release = if_else(fish_status == "Live" &
+                                  capture_type == "Mark",
+                                T, F),
+         recap = if_else(capture_type == "Recapture",
+                         T, F))
+
+effort_df = raw_df %>%
   select(event_name,
@@
          crew,
+         crew_count,
          method,
@@
   distinct() %>%
+  mutate(across(total_effort,
+                ~ . * crew_count)) %>%
   group_by(event_name,
@@
   left_join(cmr_df %>%
-              filter(capture_type != "Recapture") %>%
               group_by(event_name, species, date) %>%
@@
 LP_mod = cmr_df %>%
-  filter(species == "Northern Pikeminnow") %>%
   # filter out spring 2021 (no recapture)
   filter(event_name != "Spring_2021") %>%
-  mutate(mr_type = if_else(cap_period == "Mark" & fish_status == "Live",
-                           "M",
-                           if_else(cap_period == "Recapture",
-                                   if_else(capture_type == "Mark",
-                                           "C",
-                                           if_else(capture_type == "Recapture",
-                                                   "R",
-                                                   NA_character_)),
-                                   NA_character_))) %>%
-  filter(!is.na(mr_type)) %>%
   group_by(event_name, species) %>%
-  summarise(M = sum(mr_type == "M"),
-            n = sum(mr_type %in% c("C", "R")),
-            m = sum(mr_type == 'R'),
+  summarise(M = sum(cap_period == "Mark" & fish_status == "Live", na.rm = T),
+            n = sum(cap_period == "Recapture"),
+            m = sum(cap_period == "Recapture" & capture_type == "Recapture", na.rm = T),
             .groups = "drop") %>%
@@
                      "Chapman")) %>%
-  mutate(mr_mod = map2(data,
+  mutate(mr_model = map2(data,
                        model,
@@
                       }),
-         N_summ = map(mr_mod,
+         N_summ = map(mr_model,
                       .f = summary,
                       incl.SE = T),
-         CI = map(mr_mod,
+         CI = map(mr_model,
                   .f = confint)) %>%
@@
                       .f = function(x) x[2]),
-         N.lci = map_dbl(CI,
+         Lci = map_dbl(CI,
                          .f = function(x) x[1]),
-         N.uci = map_dbl(CI,
+         Uci = map_dbl(CI,
                          .f = function(x) x[2])) %>%
   select(-N_summ, -CI)
+
 ```
@@
 N_mods = cmr_df %>%
-  filter(species == "Northern Pikeminnow") %>%
   # filter out spring 2021 (no recapture)
   filter(event_name != "Spring_2021") %>%
-  group_by(event_name, species, date, fish_status, capture_type) %>%
-  summarise(n_fish = n(),
-            .groups = "drop") %>%
   group_by(event_name, species, date) %>%
-  summarise(n = sum(n_fish),
-            m = sum(n_fish[capture_type == "Recapture"]),
-            R = sum(n_fish[fish_status == "Live"]),
-            dead_fish = sum(n_fish[fish_status == "Dead"]),
+  summarise(n = n(),
+            m = sum(capture_type == "Recapture", na.rm = T),
+            u = sum(capture_type == "Mark" | is.na(capture_type)),
+            R = sum(fish_status == "Live", na.rm = T),
             .groups = "drop") %>%
@@
   nest() %>%
-  mutate(mod = map(data,
-                   .f = function(x) {
+  crossing(model = c("Schnabel",
+                     "SchumacherEschmeyer")) %>%
+  mutate(mr_model = map2(data,
+                         model,
+                         .f = function(x, y) {
                            with(x,
@@
                                          R = R,
-                                   method = "Schnabel",
+                                         method = y,
                                          chapman.mod = TRUE))
                          }),
-         N = map_dbl(mod,
+         N = map_dbl(mr_model,
                      .f = summary),
-         CI = map(mod,
-                  .f = confint))
+         CI = map(mr_model,
+                  .f = confint),
+         Lci = map_dbl(CI,
+                       .f = function(x) x[1]), 
+         Uci = map_dbl(CI,
+                       .f = function(x) x[2])) %>%
+  ungroup() %>%
+  select(-CI)
 
-N_mods %>%
-  select(species, N, CI) %>%
-  mutate(N.lci = map_dbl(CI,
-                         .f = function(x) x[1]),
-         N.uci = map_dbl(CI,
-                         .f = function(x) x[2])) %>%
-  select(-CI)
 ```
@@
 ```{r shoup-temp-data}
-temp_df = read_tsv("../data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt",
+temp_df = read_tsv(here("analysis/data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
                    comment = "#") %>%
@@
 ## Abundance of Piscivorous Fishes (Objective 1)
+
+```{r abund-est}
+all_est = LP_mod %>%
+  select(event_name, species, model, N:Uci) %>%
+  bind_rows(N_mods %>%
+             select(event_name, species, model, N:Uci)) %>%
+  mutate(range = Uci - Lci,
+         prop_range = range / N) %>%
+  mutate(across(event_name,
+                ~ str_replace(., "_", " "))) %>%
+  mutate(model = fct_recode(model,
+                            "Schumacher-Eschmeyer" = "SchumacherEschmeyer")) %>%
+  arrange(event_name, model)
+```
+
+```{r abund-tab}
+all_est %>%
+  select(-c(species, range:prop_range)) %>%
+  rename(`Sampling Event` = event_name,
+         Estimator = model) %>%
+  kbl(booktabs = T,
+      format.args = list(big.mark = ",")) %>%
+  kable_styling()
+```
+
+```{r abund-fig, fig.cap = "Estimates of abundance of northern pikeminnow using different estimators."}
+all_est %>%
+  mutate(across(c(N, Lci, Uci),
+                 ~ . / 1000)) %>%
+  ggplot(aes(x = model,
+             y = N,
+             color = model)) +
+  geom_errorbar(aes(ymin = Lci,
+                    ymax = Uci),
+                width = 0.1) +
+  geom_point(size = 4) +
+  facet_wrap(~ event_name) +
+  scale_color_brewer(palette = "Set1") +
+  coord_cartesian(ylim = c(0, 100)) +
+  labs(x = "Estimator",
+       y = "Pikeminnow Abundance (x 1,000)") +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45,
+                                   hjust = 1))
+
+```
+
+```{r cpue-fig, fig.cap = "Total catch per unit effort across entire sampling event."}
+cpue_df %>%
+  mutate(across(event_name,
+                ~ str_replace(., "_", " "))) %>%
+  group_by(species, event_name) %>%
+  summarise(across(c(total_effort, n_fish),
+                   sum),
+            .groups = "drop") %>%
+  mutate(across(total_effort,
+                ~ . / 60)) %>%
+  mutate(cpue = n_fish / total_effort) %>%
+  ggplot(aes(x = event_name,
+             y = cpue,
+             fill = event_name)) +
+  geom_col() +
+  scale_fill_brewer(palette = "Set1",
+                    name = "Event") +
+  labs(x = "Event",
+       y = "CPUE (Pikeminnow caught per hour)") +
+  theme(legend.position = "none")
+
+```
 

### 2021-11-12 analysis/paper/paper.Rmd
knited paper.Rmd into an html doc

@@
 output: 
+    bookdown::html_document2:
+      fig_caption: yes
+      number_sections: FALSE
+      toc: no
+      fig_width: 6
+      pandoc_args:
+      - --lua-filter=../templates/scholarly-metadata.lua
+      - --lua-filter=../templates/author-info-blocks.lua
+      - --lua-filter=../templates/pagebreak.lua      
     bookdown::word_document2:
@@
       always_allow_html: true
-      fig_width: 6
-      pandoc_args:
-      - --lua-filter=../templates/scholarly-metadata.lua
-      - --lua-filter=../templates/author-info-blocks.lua
-      - --lua-filter=../templates/pagebreak.lua
-    bookdown::html_document2:
-      fig_caption: yes
-      number_sections: FALSE
-      toc: no
       fig_width: 6
@@
 
-```{r schnabel-estimator, eval = F}
+```{r schnabel-estimator, eval = T}
 # using Schnabel model

### 2021-11-12 analysis/paper/paper.Rmd
added some equations for several estimators

@@
 ```{r schnabel-estimator, eval = T}
-# using Schnabel model
+# using Schnabel and Schumacher-Eschmeyer estimators
 N_mods = cmr_df %>%
@@
 
-The Chapman-modified Lincoln-Peterson estimator is below, where $M$ is the number of fish marked and returned to the population, $n$ is the number of fish caught in the second/recapture event and $m$ is the number of marked fish in the second sample.
+The Lincoln-Petersen estimator is below, where $M$ is the number of fish marked and returned to the population, $n$ is the number of fish caught in the second/recapture event and $m$ is the number of marked fish in the second sample.
+
+$$
+\hat{N} = \frac{(M)(n)}{(m)}
+$$
+
+The Lincoln-Petersen estimator can be biased with small samples, so we also investigated the Chapman-modified Lincoln-Petersen estimator which is shown below.
 $$
@@
 The Schnabel estimator is shown below, where the $M$, $n$ and $m$ are indexed by the sampling occasion, $i$.
+In our example, the sampling occasions are defined as each day of sampling.
 This estimator does not have an associate standard error, but 95% confidence intervals can be calculated.
@@
 \hat{N} = \frac{\sum\limits_{i = 1}^k n_i M_i}{\left(\sum\limits_{i = 1}^k m_i \right) + 1}
+$$
+There is another estimator for this type of multiple census surveys, called the Schumacher-Eschmeyer estimator, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
+It is shown below.
+
+$$
+\hat{N} = \frac{\sum\limits_{i = 1}^k n_i M^2_i}{\sum\limits_{i = 1}^k m_i M_i}
 $$

### 2021-11-15 analysis/paper/paper.Rmd
added tables showing abundance data used in estimators.

@@
 
+```{r abund-data-lp}
+LP_mod %>%
+  select(event_name, species, data) %>%
+  distinct() %>%
+  unnest(data) %>%
+  mutate(across(event_name,
+                ~ str_replace(., "_", " "))) %>%
+  rename(`Sampling Event` = event_name,
+         Species = species) %>%
+  kbl(booktabs = T,
+      format.args = list(big.mark = ","),
+      caption = "Data summary for single census estimators.
+M is the number of fish caught in the first sample, marked and returned to the population.
+n is the number of fish caught in the second sample, and m is the number of marked fish caught in the second sample.") %>%
+  kable_styling()
+```
+
+```{r abund-data-schn}
+N_mods %>%
+  select(event_name, species, data) %>%
+  distinct() %>%
+  unnest(data) %>%
+  mutate(across(event_name,
+                ~ str_replace(., "_", " "))) %>%
+  rename(`Sampling Event` = event_name,
+         Species = species,
+         Date = date) %>%
+  kbl(booktabs = T,
+      format.args = list(big.mark = ","),
+      caption = "Data summary for multiple census estimators.
+n is number of fish caught, m is the number of marked fish caught, u is the number of unmarked fish caught, and R is the number of marked fish returned to the population.") %>%
+  kable_styling()
+```
+
 ```{r abund-est}

### 2021-11-16 analysis/R scripts/Nick Script 11.10.21 cleaning fish data.R
updated code for Nicks script

@@ -158,6 +158,11 @@ Alt_Bioenergetics_plot
 
 Clean_NPM$NPM_50=lencat(Clean_NPM$Length, w = 50)
 
+PSD_DF = Clean_NPM %>%
+  group_by(NPM_50) %>%
+  summarise(n = n())
+
+
 Stomach_catagories_sum = Stomach_content_DF %>%
   drop_na(StomachContentsWeight) %>%
   group_by(Fish_or_Parts) %>%

### 2021-11-16 analysis/paper/paper.Rmd

@@
 
-The Deadwater Slough is located in a reach of the Salmon River that is believed to be a historically important overwinter rearing area for DSR emigrants.
+The Deadwater Slough is in a reach of the Salmon River that is believed to be a historically important overwinter rearing area for DSR emigrants.
 Moreover, this reach is part of the migratory pathway for juvenile DSR and NRR emigrants from all eight extant populations.
@@
 
-We hypothesize that increased densities of piscivorous predators in the Deadwater Slough area may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon [@Ackerman2018] and Sockeye Salmon [@Axel2015].
+We hypothesize that increased densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon [@Ackerman2018] and Sockeye Salmon [@Axel2015].
 In this study, we estimated the abundance of a piscivorous fish predator population in the Deadwater Slough and their potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
@@
 The population of Northern Pikeminnow in Deadwater Slough was estimated to consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook Salmon during the ***x*** sampling periods and result in an estimated reduction of returning adults between ***xxx*** and ***x,xxx***.
-We suggest that the habitat modifications that created the Deadwater Slough reach have resulted in favorable conditions for Northern Pikeminnow, including improved conditions for predation upon juvenile Chinook Salmon ***(add detail here)***.
+We suggest that the habitat modifications that created the Deadwater Slough have resulted in favorable conditions for Northern Pikeminnow, including improved conditions for predation upon juvenile Chinook Salmon ***(add detail here)***.
 Therefore, predation by Northern Pikeminnow in the Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG.
@@
 
+<!--Please feel free to add others that I missed or improve*** -->
+
+The authors extent much appreciation to the many volunteers who assisted with field efforts including collaborators from Bureau of Reclamation, Idaho Department of Fish and Game, and Lemhi Regional Land Trust, among others.
+This manuscript benefited from reviews and contributions from colleagues at the Idaho Governor's Office of Species Conservation, Rio Applied Science and Engineering, and from Sean Gibbs and Ben Briscoe at Mount Hood Environmental.
+Funding for this study was provided by the Bureau of Reclamation, Pacific Northwest Regional Office (contract No.
+140R1021F0018).
+Special thanks to Caitlin Alcott and Inter-Fluve for their administrative support and guidance.
+
 \newpage

### 2021-11-16 analysis/paper/paper.Rmd
added Nick's content to paper.Rmd

@@
 
-***Paragraph 3***
+An estimated 200 million juvenile salmonids emigrate through the lower Snake and Columbia rivers, annually.
+Of those, approximately 16.4 million (8%) are consumed by northern pikeminnow [Beamesderfer1996].
+@Peterson1994 estimated that the annual loss of juvenile salmonids to northern pikeminnow to be 7.3% of all juvenile salmonids entering the John Day Reservoir.
+Additionally, 78% of these salmonids consumed near a dam were consumed when alive [@Peterson1994].
+Dams are known to be areas of predation, but @Ward1995 estimated that 48% of predation is occurring away from dams in mid-reservoir areas.
+Most of the research has been targeted around dams and reservoirs, but there are areas functioning similar to a reservoir in the rivers upstream.
+One of these areas is the Deadwater Slough in the Salmon River.
+The Deadwater Slough is a 25.3-acre area that is unnaturally slow and deep within the Salmon River and resembles a small reservoir (Figure \@ref(fig:map)).
+This unnatural area has little to no cover, higher depth, and fine substrate which favor piscivorous fish predators (e.g., northern pikeminnow *Ptychocheilus oregonensis*, smallmouth bass *Micropterus dolomieu*) [@Watkins2015].
+@Axel2015 demonstrated decreased rates of emigration and apparent survival for Sockeye Salmon *Oncorhynchus nerka* from Redfish Lake emigrating during the spring.
+Further, recent winter telemetry studies have indicated decreased transition probabilities (approximately 10% less than surrounding reaches) of juvenile Chinook salmon through Deadwater Slough during fall and winter months [@Ackerman2018; @Porter2019].
 
-* Summary of findings from @Axel2015 and early RT reports and papers (@Ackerman2018, @Porter2019, @Porter2020) in Deadwater Slough.
-Recent studies have demonstrated decreased movement and survival of juvenile salmonids emigrating through the Deadwater Slough.
-@Axel2015 demonstrated decreased rates of emigration and survival for sockeye salmon *Oncorhynchus nerka* from Redfish Lake emigrating during the spring and recent winter telemetry studies have indicated decreased transition probabilities (approximately 10% less than nearby reaches) of juvenile Chinook salmon *Oncorhynchus tschawytscha* through Deadwater Slough during fall and winter months (@Ackerman2018, @Porter2019, @Porter2020).
+<!--
 * @Axel2015 found increased mortality in a particular reach spanning location x to location y in the mainstem Salmon River on a study of juvenile (hatchery?) sockeye salmon released near Redfish Lake and actively migrating downriver.
@@
 * Anecdotally, locals (and others? maybe NOAA?) informed us that northern pikeminnow and perhaps other species are prevalent in Deadwater Slough.
+-->
 
@@
 
-## Abundance of Piscivorous Fishes (Objective 1)
+## Abundance of Piscivorous Fishes
 
@@
 $$
+We additionally calculated the proportional stock density (PSD) of northern pikeminnow with 300 mm total length (TL) for stock and 400 mm TL for quality.
+$$
+PSD_{i} = 100 * \frac{FQ_{i}}{FS_{i}}
+$$
+where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
 
-## Predation Upon Juvenile Chinook Salmon i.e., Gastric Lavage (Objective 2)
+## Predation On Juvenile Chinook Salmon
 
-**Data = Gastric Lavage Data**
+We were interested in documenting potential consumption of juvenile Chinook Salmon by northern pikeminnow, and ideally, estimate the proportion of their diet consisting of juvenile Chinook Salmon at the time of sampling.
+To accomplish this, we collected stomach contents using gastric lavage as described in @Foster1977 and examined contents for the presence or absence of juvenile Chinook Salmon or other incidentals (e.g., steelhead, Sockeye Salmon, Redside Shiner, etc.), and the proportion of any stomach contents containing fish or fish parts versus non-targets (e.g., macroinvertebrates, organic matter, etc.).
+Stomach contents were stored in whirl-paks, preserved with 99% isopropyl alcohol, and analyzed within the following week in a controlled environment.
+Each sample was uniquely identified to match up with the appropriate fish record, contents were identified down to their unique composition, total weight of all content was measured as wet weight in grams, and total weight of fish content, if found, was measured in grams.
+Fish and fish parts were identified down to species, if possible, or categorized as unknown if unidentifiable.
+Throughout the sampling periods a subset of the capture individuals were sacrificed after gastric lavage to validate whether gastric lavage was successful at flushing stomach contents from the northern pikeminnow.
 
-The goal was to document predation upon juvenile Chinook salmon (or other targets), and ideally, estimate the proportion of their diet that consisted of juvenile Chinook salmon at the time of sampling?
-
-* Which individuals was gastric lavage performed upon?
-* How was gastric lavage performed?
-* Intent was to quantify predation on juvenile Chinook salmon
-  - or incidentally, juvenile steelhead or sockeye salmon
-
-<!-- below pulled from technical report -->
-
-We collected stomach contents from most captured individuals using gastric lavage (@Foster1977) and examined contents for the presence or absence of juvenile Chinook salmon and other incidentals (e.g., steelhead, sockeye salmon) juveniles and the proportion of stomach contents containing targets versus non-targets (e.g., macroinvertebrates).
-Stomach contents were stored in whirl-paks, preserved with 99% isopropyl alcohol, and analyzed one week later in a controlled environment.
-Each sample was uniquely identified to match up with the appropriate fish record, contents were identified down to its unique composition, total weight of all content was measured in grams, and total weight of fish content, if found, was measured in grams.
-Throughout the sampling period a proportion of the captured individuals were sacrificed after gastric lavage to validate that the gastric lavage was successful at flushing all or most of the stomach contents from the northern pikeminnow.
-
-## Bioenergetics (Objective 3)
+## Bioenergetics
 
@@
 
-**Data = Temperature, Bioenergetics Inputs, Others?**
+To estimate the impacts from northern pikeminnow on juvenile Chinook Salmon outmigrants in Deadwater Slough, we used an R-based application of Fish Bioenergetics v4.0 developed by @Deslauriers2017.
+We estimated a daily rate of consumption for an individual northern pikeminnow based on predator and prey energy densities, water temperatures, and predator start and end weights.
+We also ran two alternative models one with a year of temperature inputs from 2019, and another with the same temperature inputs but a 10% growth in weight of the average pikeminnow at the end of the model.
+The average length of northern pikeminnow captured at the Deadwater Slough during our study was 370.6 mm which calculates to an average starting weight of 847 g using the FSA package (@Ogle2019) in R which provides a weight-length formula for northern pikeminnow sourced from @Parker1995.
+Predator energy density for northern pikeminnow was available in the Bioenergetics v4.0 application and was fixed at 6,703 Joules(J)/g.
+Prey energy densities was taken from @Moss2016 where they estimated juvenile Chinook Salmon at 21,500 J/g.
+Using these temperatures, weights, and energy inputs, we estimated a total consumption in grams for an individual norther pikeminnow within Deadwater Slough.
 
-Pull material from technical report
+We ran three bioenergetic models for this study.
+First, we chose a 78-day period from September 1 through November 17 where we know Chinook Salmon DSR emigrants begin to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin to migrate downstream, but when temperatures are still high enough that juveniles are not yet exhibiting concealment behavior or torpor.
+The second model was still assuming no growth but taking the entire year into account for the model rather than the 78-day fall period.
+The assumption of no growth for pikeminnow was included to show what may be occurring if the population of northern pikeminnow is stable.
+If growther of the population is occurring, the alternative model shows how an increase of 10% body weight changes the estimates of consumption over a full year period.
+To estimate the number of possible Chinook Salmon that could be consumed, we assumed a weight of 12 g per individual chinook Salmon.
+This weight was an average from juvenile Chinook salmon emigrating past six rotary screw traps operating in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.
 
-## Impacts to Adult Returns (Objective 4)
+## Impacts to Adult Returns
 
-**Data = SARs, Adult Escapements, Others?**
+Text here...
 
-Pull material from technical report
-
-  4. After making some assumptions (e.g., on SARs) attempt to estimate how results from objective #3 could translate into impacts on adult returns (via a thought experiment or simulation approach)
+<!-- After making some assumptions (e.g., on SARs) attempt to estimate how results from objective #3 could translate into impacts on adult returns (via a thought experiment or simulation approach) -->
 
@@
 
+<!--
 * Total population estimates of Pike Minnow. Chart or table, but likely a chart.
 * What is that equate to as a density? e.g., how many pikeminnow per 10 square meters? Is that a reasonable number?
+-->
 
-## Predation Upon Juvenile Chinook Salmon i.e., Gastric Lavage (Objective 2)
+PSD for northern pikeminnow across all three years was 42.26% or 42.26% of fish were over the quality size classification 400 mm TL we denoted in the Methods, demonstrating a high percentage of above average size classs present within Deadwater Slough.
 
-*	Gastric lavage sampling results. What percent of the diet is chinook salmon from each sampling event. Pie chart? Mike loves pie charts.
+## Predation On Juvenile Chinook Salmon
 
-## Bioenergetics (Objective 3)
+During 2019, 660 northern pikeminnow were gastric lavaged.
+Of those fish, 603 had empty stomachs, 57 had stomach contents, 12 had fish parts, and of those 12 with fish parts 1 MWF and 2 shiners.
+During the 2020 season 805 northern pikeminnow were gastric lavaged with 613 empty stomachs, 188 with stomach contents and 23 with fish parts.
+Out of the 23 stomachs with fish parts present 10 were unknown, 6 red sided shiner, 4 suckers, 2 sculpin, and 1 chinook.
+In 2021, we had 105 stomach samples with 98 being non-fish and 7 having fish or fish parts.
 
+<!--- Need to add Nick's gastric lavage table here -->
+
+## Bioenergetics
+
+<!-- Need to add Nick's figures here -->
+
+Over the fall time period a single northern pikeminnow is modeled to eat 61 g of fish.
+If we extrapolate this out to the potential effect the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook salmon having the potential to consume ***xxx** juvenile Chinook as a discrete population.
+
+Over the point of a whole year, a single northern pikeminnow is modeled to eat 283 g of fish and 613 g of other food types.
+
+<!--
 * Water Temperature measurements.
@@
 *	Bioenergetics results. Total mass consumed over a 365 day period. Plot of daily consumption with day of year on x and mass on y.
+-->
 
-## Impacts to Adult Returns (Objective 4)
+## Impacts to Adult Returns
 
@@
 
+<!--
 Re-hash out our assumptions?
@@
 In the end, how good do we think our estimate is and is the bottom-line that we belive there still to be a shit-ton of pikeminnow even if our assumptions were violated?
+-->
 
@@
 
+We were able to gastric lavage the majority of fish collected including some non-predatory species.
+We euthanized nine northern pikeminnow for dissection to confirm that the gastric lavage was effective in removing stomach contents.
+Our results support the findings the previous year that dissected individuals that the dissected gastric lavage was effective in removing stomach contents (@Lott2020).
+Additionally , we found that the fish captured in trap-nets had a similar proportion of stomach content samples.
+Of those, 1,214 (76.2%) were completely empty, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish parts.
+We were able to identify juvenile shiners, a largescale sucker, sculpin, mountain whitefish, and one Chinook Salmon within stomach contents, but most samples were too digested to identify to species.
+
+<!--
 Concerns with gastric lavage.
@@
 So instead, we made some assumptions about the proportion of a pikeminnows diet consisting of Chinook salmon.
+-->
 
@@
 
-Discuss the potential for various management actions:
-* Removing the Dump Creek alluvial fan and restoring flow. What would that look like?
-* A local bounty program? What would that look like? Have they been successful elsewhere e.g., in the Columbia River?
+We estimated that consumption of juvenile Chinook Salmon by northern pikeminnow in the Deadwater Slough potentially reduces annual adult returns by ***xxx - x,xxx*** to upriver populations.
+Presumably, that reduction in adult returns impacts both the ESA-listed natural populations in the Upper Salmon River MPG plus the two hatchery populations in the Upper Salmon, Pahsimeroi and Sawtooth hatcheries, which provide for recreational fishing opportunities.
+Fisheries managers desire increased adult returns to both aide in the recovery of natural populations as well as to provide additional harvest in recreational fisheries to boost local economies.
+A reduction in predation mortality at Deadwater Slough on juvenile Chinook Salmon has potential to benefit multiple upriver natural and hatchery populations, in contrast to typical tributary habitat rehabilition actions which typically benefit a single population.
+Moreover, the deepened, slack water conditions that favor northern pikeminnow at Deadwater Slough are indirectly the result of manmade activities i.e., the failure of a manmade mining reservoir dam.
+Given these reasons, it seems that Deadwater Slough could be a candidate for management or restoration actions to benefit local Chinook Salmon populations.
 
-* Deadwater is an opportunity to benefit multiple (6 or 7) local populations with a single management or restoration action.
+We see two potential management actions: 1) removing the Dump Creek delta to restore flow and 2) a local northern pikeminnow bounty program to encourage harvest of northern pikeminnow in Deadwater Slough.
 
-*	Also, this is an anthropogenically created dead water area, which could be a candidate for restoration (removing the impoundment).
 Is there anything in the literature that discusses habitat preferences for pike minnow?
@@
 
-<!--Please feel free to add others that I missed or improve*** -->
+<!-- Please feel free to add others that I missed or improve -->
 

### 2021-11-16 analysis/paper/paper.Rmd
added always_allow_html to YAML header, knit .html and .docx

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
+always_allow_html: true
 output: 
-    bookdown::html_document2:
-      fig_caption: yes
-      number_sections: FALSE
-      toc: no
-      fig_width: 6
-      pandoc_args:
-      - --lua-filter=../templates/scholarly-metadata.lua
-      - --lua-filter=../templates/author-info-blocks.lua
-      - --lua-filter=../templates/pagebreak.lua      
     bookdown::word_document2:
@@
       always_allow_html: true
+      fig_width: 6
+      pandoc_args:
+      - --lua-filter=../templates/scholarly-metadata.lua
+      - --lua-filter=../templates/author-info-blocks.lua
+      - --lua-filter=../templates/pagebreak.lua
+    bookdown::html_document2:
+      fig_caption: yes
+      number_sections: FALSE
+      toc: no
       fig_width: 6

### 2021-11-19 analysis/paper/paper.Rmd
added citations and re-knit .docx

@@
 Of those, approximately 16.4 million (8%) are consumed by northern pikeminnow [Beamesderfer1996].
-@Peterson1994 estimated that the annual loss of juvenile salmonids to northern pikeminnow to be 7.3% of all juvenile salmonids entering the John Day Reservoir.
-Additionally, 78% of these salmonids consumed near a dam were consumed when alive [@Peterson1994].
+@Petersen1994 estimated that the annual loss of juvenile salmonids to northern pikeminnow to be 7.3% of all juvenile salmonids entering the John Day Reservoir.
+Additionally, 78% of these salmonids consumed near a dam were consumed when alive [@Petersen1994].
 Dams are known to be areas of predation, but @Ward1995 estimated that 48% of predation is occurring away from dams in mid-reservoir areas.
@@
 We also ran two alternative models one with a year of temperature inputs from 2019, and another with the same temperature inputs but a 10% growth in weight of the average pikeminnow at the end of the model.
-The average length of northern pikeminnow captured at the Deadwater Slough during our study was 370.6 mm which calculates to an average starting weight of 847 g using the FSA package (@Ogle2019) in R which provides a weight-length formula for northern pikeminnow sourced from @Parker1995.
+The average length of northern pikeminnow captured at the Deadwater Slough during our study was 370.6 mm which calculates to an average starting weight of 847 g using the FSA package (@Ogle2021) in R which provides a weight-length formula for northern pikeminnow sourced from @Parker1995.
 Predator energy density for northern pikeminnow was available in the Bioenergetics v4.0 application and was fixed at 6,703 Joules(J)/g.

### 2021-11-19 analysis/paper/paper.Rmd
moved tables back to Table section

@@
 theme_set(theme_bw())
+
+# when knitting to Word, use this
+doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
+
+if(doc.type == 'docx') {
+  options(knitr.table.format = "pandoc")
+}
 
@@
 
-```{r abund-data-lp}
-LP_mod %>%
-  select(event_name, species, data) %>%
-  distinct() %>%
-  unnest(data) %>%
-  mutate(across(event_name,
-                ~ str_replace(., "_", " "))) %>%
-  rename(`Sampling Event` = event_name,
-         Species = species) %>%
-  kbl(booktabs = T,
-      format.args = list(big.mark = ","),
-      caption = "Data summary for single census estimators.
-M is the number of fish caught in the first sample, marked and returned to the population.
-n is the number of fish caught in the second sample, and m is the number of marked fish caught in the second sample.") %>%
-  kable_styling()
-```
-
-```{r abund-data-schn}
-N_mods %>%
-  select(event_name, species, data) %>%
-  distinct() %>%
-  unnest(data) %>%
-  mutate(across(event_name,
-                ~ str_replace(., "_", " "))) %>%
-  rename(`Sampling Event` = event_name,
-         Species = species,
-         Date = date) %>%
-  kbl(booktabs = T,
-      format.args = list(big.mark = ","),
-      caption = "Data summary for multiple census estimators.
-n is number of fish caught, m is the number of marked fish caught, u is the number of unmarked fish caught, and R is the number of marked fish returned to the population.") %>%
-  kable_styling()
-```
-
 ```{r abund-est}
@@
   arrange(event_name, model)
-```
-
-```{r abund-tab}
-all_est %>%
-  select(-c(species, range:prop_range)) %>%
-  rename(`Sampling Event` = event_name,
-         Estimator = model) %>%
-  kbl(booktabs = T,
-      format.args = list(big.mark = ",")) %>%
-  kable_styling()
 ```
@@
 
+```{r abund-data-lp}
+LP_mod %>%
+  select(event_name, species, data) %>%
+  distinct() %>%
+  unnest(data) %>%
+  mutate(across(event_name,
+                ~ str_replace(., "_", " "))) %>%
+  rename(`Sampling Event` = event_name,
+         Species = species) %>%
+  kable(booktabs = T,
+        format.args = list(big.mark = ","),
+        caption = "Data summary for single census estimators.
+M is the number of fish caught in the first sample, marked and returned to the population.
+n is the number of fish caught in the second sample, and m is the number of marked fish caught in the second sample.") %>%
+  kable_styling()
+
+```
+
+\newpage
+
+```{r abund-data-schn}
+N_mods %>%
+  select(event_name, species, data) %>%
+  distinct() %>%
+  unnest(data) %>%
+  mutate(across(event_name,
+                ~ str_replace(., "_", " "))) %>%
+  rename(`Sampling Event` = event_name,
+         Species = species,
+         Date = date) %>%
+  kable(booktabs = T,
+        format.args = list(big.mark = ","),
+        caption = "Data summary for multiple census estimators.
+n is number of fish caught, m is the number of marked fish caught, u is the number of unmarked fish caught, and R is the number of marked fish returned to the population.") %>%
+  kable_styling()
+
+```
+
+\newpage
+
+```{r abund-tab}
+all_est %>%
+  select(-c(species, range:prop_range)) %>%
+  rename(`Sampling Event` = event_name,
+         Estimator = model) %>%
+  kable(booktabs = T,
+        format.args = list(big.mark = ","),
+        caption = "Temporary caption.") %>%
+  kable_styling()
+
+```
+
 \newpage

### 2021-11-19 analysis/paper/paper.Rmd
and moved figures to Figures
edits to Predation on Juvenile Chinook Salmon methods

@@
 
-We were interested in documenting potential consumption of juvenile Chinook Salmon by northern pikeminnow, and ideally, estimate the proportion of their diet consisting of juvenile Chinook Salmon at the time of sampling.
-To accomplish this, we collected stomach contents using gastric lavage as described in @Foster1977 and examined contents for the presence or absence of juvenile Chinook Salmon or other incidentals (e.g., steelhead, Sockeye Salmon, Redside Shiner, etc.), and the proportion of any stomach contents containing fish or fish parts versus non-targets (e.g., macroinvertebrates, organic matter, etc.).
-Stomach contents were stored in whirl-paks, preserved with 99% isopropyl alcohol, and analyzed within the following week in a controlled environment.
-Each sample was uniquely identified to match up with the appropriate fish record, contents were identified down to their unique composition, total weight of all content was measured as wet weight in grams, and total weight of fish content, if found, was measured in grams.
-Fish and fish parts were identified down to species, if possible, or categorized as unknown if unidentifiable.
-Throughout the sampling periods a subset of the capture individuals were sacrificed after gastric lavage to validate whether gastric lavage was successful at flushing stomach contents from the northern pikeminnow.
+Gastric lavage [@Foster1977] was used to examine the stomach contents of Northern Pikeminnow for the presence of juvenile Chinook Salmon, other fishes (e.g., juvenile steelhead, juvenile Sockeye Salmon, Redside Shiner, etc.), and non-targets (e.g., macroinvertebrates, organic matter, etc.).
+Stomach contents were preserved with 99% isopropyl alcohol in whirl-paks and analyzed in a laboratory.
+For each sample, wet weight (grams) was recorded for the combined stomach content and for the fish content.
+Fish and fish remnants were identified to the lowest taxonomic unit, when possible, or were categorized as unknown.
+A subset of Northern Pikeminnow captures were euthanized for dissection after gastric lavage to validate the efficacy of the methodology.
 
@@
   arrange(event_name, model)
-```
-
-```{r abund-fig, fig.cap = "Estimates of abundance of northern pikeminnow using different estimators."}
-all_est %>%
-  mutate(across(c(N, Lci, Uci),
-                 ~ . / 1000)) %>%
-  ggplot(aes(x = model,
-             y = N,
-             color = model)) +
-  geom_errorbar(aes(ymin = Lci,
-                    ymax = Uci),
-                width = 0.1) +
-  geom_point(size = 4) +
-  facet_wrap(~ event_name) +
-  scale_color_brewer(palette = "Set1") +
-  coord_cartesian(ylim = c(0, 100)) +
-  labs(x = "Estimator",
-       y = "Pikeminnow Abundance (x 1,000)") +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45,
-                                   hjust = 1))
-
-```
-
-```{r cpue-fig, fig.cap = "Total catch per unit effort across entire sampling event."}
-cpue_df %>%
-  mutate(across(event_name,
-                ~ str_replace(., "_", " "))) %>%
-  group_by(species, event_name) %>%
-  summarise(across(c(total_effort, n_fish),
-                   sum),
-            .groups = "drop") %>%
-  mutate(across(total_effort,
-                ~ . / 60)) %>%
-  mutate(cpue = n_fish / total_effort) %>%
-  ggplot(aes(x = event_name,
-             y = cpue,
-             fill = event_name)) +
-  geom_col() +
-  scale_fill_brewer(palette = "Set1",
-                    name = "Event") +
-  labs(x = "Event",
-       y = "CPUE (Pikeminnow caught per hour)") +
-  theme(legend.position = "none")
-
 ```
@@
 
+```{r abund-fig, fig.cap = "Estimates of abundance of northern pikeminnow using different estimators."}
+all_est %>%
+  mutate(across(c(N, Lci, Uci),
+                 ~ . / 1000)) %>%
+  ggplot(aes(x = model,
+             y = N,
+             color = model)) +
+  geom_errorbar(aes(ymin = Lci,
+                    ymax = Uci),
+                width = 0.1) +
+  geom_point(size = 4) +
+  facet_wrap(~ event_name) +
+  scale_color_brewer(palette = "Set1") +
+  coord_cartesian(ylim = c(0, 100)) +
+  labs(x = "Estimator",
+       y = "Pikeminnow Abundance (x 1,000)") +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45,
+                                   hjust = 1))
+
+```
+
+\newpage
+
+```{r cpue-fig, fig.cap = "Total catch per unit effort across entire sampling event."}
+cpue_df %>%
+  mutate(across(event_name,
+                ~ str_replace(., "_", " "))) %>%
+  group_by(species, event_name) %>%
+  summarise(across(c(total_effort, n_fish),
+                   sum),
+            .groups = "drop") %>%
+  mutate(across(total_effort,
+                ~ . / 60)) %>%
+  mutate(cpue = n_fish / total_effort) %>%
+  ggplot(aes(x = event_name,
+             y = cpue,
+             fill = event_name)) +
+  geom_col() +
+  scale_fill_brewer(palette = "Set1",
+                    name = "Event") +
+  labs(x = "Event",
+       y = "CPUE (Pikeminnow caught per hour)") +
+  theme(legend.position = "none")
+
+```
+
+\newpage
+
 ### Colophon
@@
 
+\newpage
+
 The current Git commit details are:

### 2021-11-19 analysis/paper/paper.Rmd
minor edits to bioenergetics

@@
 
-An estimated 200 million juvenile salmonids emigrate through the lower Snake and Columbia rivers, annually.
-Of those, approximately 16.4 million (8%) are consumed by northern pikeminnow [Beamesderfer1996].
-@Petersen1994 estimated that the annual loss of juvenile salmonids to northern pikeminnow to be 7.3% of all juvenile salmonids entering the John Day Reservoir.
-Additionally, 78% of these salmonids consumed near a dam were consumed when alive [@Petersen1994].
-Dams are known to be areas of predation, but @Ward1995 estimated that 48% of predation is occurring away from dams in mid-reservoir areas.
-Most of the research has been targeted around dams and reservoirs, but there are areas functioning similar to a reservoir in the rivers upstream.
+Northern Pikeminnow are estimated to consume 8% (16.4 million) of the estimated 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996].
+Dams and reservoirs in the Columbia River are associated with high rates of predation on salmonids [@Petersen1994; @Ward1995].
+Because migration rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], migrating smolts may become more vulnerable to predation in areas with slower velocities.
+Most of the research has been targeted around dams and reservoirs, but there are areas functioning similar to a reservoir in the ruvers upstream.
 One of these areas is the Deadwater Slough in the Salmon River.
@@
 $$
-We additionally calculated the proportional stock density (PSD) of northern pikeminnow with 300 mm total length (TL) for stock and 400 mm TL for quality.
+We additionally calculated the proportional stock density (PSD) of Northern Pikeminnow with 300 mm total length (TL) for stock and 400 mm TL for quality.
 $$
@@
 
+To estimate the consumption of juvenile Chinook Salmon outmigrants by Northern Pikeminnow in the Deadwater Slough, we used the Fish Bioenergetics v4.0 applicatioon developed by @Deslauriers2017.
+The daily rate of consumption in grams for an individual Northern Pikeminnow was based on predator and prey energy densities, predator start and end weights, and water temperatures.
+Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017].
+Prey energy densities were fixed at 21,500 J/g for juvenile Chinook salmon based on estimate from @Moss2016.
+To calculate the average predator start weight, we converted the average length of Northern Pikeminnow captured during our study to a weight using a weight-length formula for Northern Pikeminnow [@Parker1995] from the FSA package [@Ogle2021].
+Continuous water temperature data from 2019 were used in two alternative models, on using the baseline start and end weights and the second with a 10% increase in average end weight of Northern Pikeminnow.
+
+We ran three bioenergetic models for this study.
+For the first model, we chose a 78-day period from September 1 through November 17 when DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration.
+During this time, water temperatures exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon.
+This model assumed no growth for Northern Pikeminnow.
+The second model was run for one-year and assumed no growth for Northern Pikeminnow.
+The assumption of no growth for pikeminnow was included to show what may be occurring if the populaiton of Northern Pikeminnow is stable.
+If growth of the population is occurring, the alternative model shows how an increase of 10% body weight changes the estimtes of consumption over a full year period.
+To estimate the number of possible Chinook Salmon that could be consumed, we assumed a weight of 12 g per individual Chinook Salmon.
+This weight was an average from juvenile Chinook Salmon emigrating past six rotary screw traps operating in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.
+
+<!--
 To estimate the impacts from northern pikeminnow on juvenile Chinook Salmon outmigrants in Deadwater Slough, we used an R-based application of Fish Bioenergetics v4.0 developed by @Deslauriers2017.
@@
 This weight was an average from juvenile Chinook salmon emigrating past six rotary screw traps operating in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.
+-->
 
@@
 
-We were able to gastric lavage the majority of fish collected including some non-predatory species.
-We euthanized nine northern pikeminnow for dissection to confirm that the gastric lavage was effective in removing stomach contents.
-Our results support the findings the previous year that dissected individuals that the dissected gastric lavage was effective in removing stomach contents (@Lott2020).
-Additionally , we found that the fish captured in trap-nets had a similar proportion of stomach content samples.
-Of those, 1,214 (76.2%) were completely empty, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish parts.
-We were able to identify juvenile shiners, a largescale sucker, sculpin, mountain whitefish, and one Chinook Salmon within stomach contents, but most samples were too digested to identify to species.
+The majority of fish collected during this study received gastric lavage, including some non-predatory species.
+To validate the efficacy of this method, we euthanized nine Northern Pikeminnow after gastric lavage was completed and removed the remaining stomach contents via dissection.
+Our results support previous findings that gastric lavage effectively removes stomach contents (@Lott2020).
+Additionally, we found that the fish captured in net-traps had a similar proportion of stomach content samples.
+of those, 1,214 (76.2%) were completely emply, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
+We were able to identify juvenile Red Shiners, a Largescale Sucker, Sculpin, Mountain Whitefish, and one Chinook Salmon within stomach contents, but most samples were too digested to identify to species.
 
@@
 Although not formally assessed in this study, avian predators including Great Blue Herons *Ardea herodias* and Bald Eagles *Haliaeetus leucocephalus* are another potential source of mortality for juvenile salmon in the Deadwater Slough.
-The Deadwater Slough is recognized as an important bird watching and nesting area due to the riparian and backwater habitats created by the feature [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
+The Deadwater Slough is recognized as an important bird watching and nesting area due to the associated riparian and backwater habitats [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
 Several piscivorous bird species have been documented using the Deadwater Slough that include the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
@@
 
+<!--
 Is there anything in the literature that discusses habitat preferences for pike minnow?
@@
 Will these fish just move elsewhere?
+-->
 

### 2021-11-20 analysis/paper/paper.Rmd
added Nick's gastric lavage code

@@
 library(kableExtra)
+library(stringi)
 
@@
 ## Predation On Juvenile Chinook Salmon
+
+```{r gastric-data}
+gastric_df = raw_df %>%
+  filter(species == "Northern Pikeminnow") %>%
+  mutate(fish_or_parts = case_when(stri_detect_fixed(stomach_contents, "Fish") ~ "Fish or Fish Parts",
+                                   stri_detect_fixed(stomach_contents, "Empty") ~"Empty",
+                                   stri_detect_fixed(comments, "Whole scuplin") ~ "Fish or Fish Parts",
+                                   stri_detect_fixed(stomach_contents, "Whitefish") ~ "Fish or Fish Parts",
+                                   TRUE ~ "Other")) %>%
+  drop_na(length) %>%
+  mutate(non_fish_wt = stomach_contents_weight - fish_content_weight) %>%
+  dplyr::select(date, length, lavage_id, stomach_contents, comments, fish_or_parts, stomach_contents_weight, non_fish_wt,                  fish_content_weight) %>%
+  filter(!stomach_contents %in% c("Lost sample CG27-001", "Did Not Lavage")) %>%
+  drop_na(stomach_contents)
+
+```
 

### 2021-11-23 analysis/paper/paper.Rmd
additions to Intro

@@
 
-***Paragraph 2***
+One potentially important, but perhaps under-appreciated source of mortality for Chinook Salmon is predation on emigrating juveniles by piscivorous fishes, including both native and non-native species.
+Dams and reservoirs in the Columbia River basin are the primary locations associated with high rates of predation on salmonids [@Petersen1994; @Ward1995].
+There are generally two maechanisms that explain these high predation zones.
+First, migration rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation.
+Second, reservoirs and downstream tailraces created by dams on the Columbia and Snake rivers have created favorable slow-water habitat for several species of piscivorous fish known to consume juvenile Chinook Salmon including Walleye *Sander vitreus*, Smallmouth Bass *Micropterus dolomieu*, and Northern Pikeminnow *Ptychocheilus oregonensis*.
+In the Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of non-native Smallmouth Bass during some seasons [@Erhardt2018].
+Walleye also prefer slow water habitats and are considered a potentially important source for Juvenile Chinook Salmon mortality [@NOAA2020], particularly as their abundance has significantly increased in Snake River dam facilities between 2008 and 2020.
 
-One potentially important, but perhaps under-appreciated source of mortality on Chinook salmon is predation on emigrating juveniles by piscivorous fishes, including both native and non-native species.
-As an example,...
+In addition to non-native species, Northern Pikeminnow are a native, piscivorous fish preferring slow water habitats.
+Consequently, they too have benefited from dams on the Columbia River and become abundance predators on salmonid outmigrants [@Knutsen1999].
+Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the estimated 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996].
+Predation by piscivorous fish may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@NOAA2020], however there is a dearth of current data on piscivorous predator populations in the Columbia River.
 
-  - At this point, I'm thinking leave avian predation out of the equation, but instead include in Discussion
-  - and keep focus on the freshwater environment
-* What is known about downriver (Columbia/Snake) piscivorous fish populations and potential predation on juvenile Chinook salmon?
-  - literature review of northern pikeminnow, smallmouth bass, other? predation on juvenile Chinook salmon, particulary from the Snake and Columbia rivers
-  - still stay away from avian predation
-
-Northern Pikeminnow are estimated to consume 8% (16.4 million) of the estimated 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996].
-Dams and reservoirs in the Columbia River are associated with high rates of predation on salmonids [@Petersen1994; @Ward1995].
-Because migration rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], migrating smolts may become more vulnerable to predation in areas with slower velocities.
-Most of the research has been targeted around dams and reservoirs, but there are areas functioning similar to a reservoir in the ruvers upstream.
-One of these areas is the Deadwater Slough in the Salmon River.
-The Deadwater Slough is a 25.3-acre area that is unnaturally slow and deep within the Salmon River and resembles a small reservoir (Figure \@ref(fig:map)).
-This unnatural area has little to no cover, higher depth, and fine substrate which favor piscivorous fish predators (e.g., northern pikeminnow *Ptychocheilus oregonensis*, smallmouth bass *Micropterus dolomieu*) [@Watkins2015].
+Habitat containing slow-velocities and other attributes that support piscivorous predators are not limited to dams and reservoirs.
+One such area is the Deadwater Slough in the Salmon River, which occupies a 25.3-acre reach of unnaturally slow and deep water (Figure \@ref(fig:map)).
+The Deadwater Slough contains little to no cover, increased depth, and fine substrate which favor piscivorous fish predators (e.g., Northern Pikeminnow, Smallmouth Bass) [@Watkins2015].
 @Axel2015 demonstrated decreased rates of emigration and apparent survival for Sockeye Salmon *Oncorhynchus nerka* from Redfish Lake emigrating during the spring.

### 2021-11-23 analysis/paper/paper.Rmd
round of TB edits added

@@
 $$
-We additionally calculated the proportional stock density (PSD) of Northern Pikeminnow with 300 mm total length (TL) for stock and 400 mm TL for quality.
+Additionally, we calculated the proportional stock density (PSD) of Northern Pikeminnow with 300 mm total length (TL) for stock and 400 mm TL for quality.
 $$
@@
 
-During 2019, 660 northern pikeminnow were gastric lavaged.
-Of those fish, 603 had empty stomachs, 57 had stomach contents, 12 had fish parts, and of those 12 with fish parts 1 MWF and 2 shiners.
-During the 2020 season 805 northern pikeminnow were gastric lavaged with 613 empty stomachs, 188 with stomach contents and 23 with fish parts.
-Out of the 23 stomachs with fish parts present 10 were unknown, 6 red sided shiner, 4 suckers, 2 sculpin, and 1 chinook.
-In 2021, we had 105 stomach samples with 98 being non-fish and 7 having fish or fish parts.
+During 2019, 660 Northern Pikeminnow were gastric lavaged.
+Of those fish, we observed 603 individuals with empty stomachs, and 57 individuals with stomach contents.
+Twelve of the samples contained fish remnants including a single Mountain Whitefish and two Redside Shiners.
+During the 2020 season, 805 Northern Pikeminnow were gastric lavaged comprised of 613 individuals with empty stomachs and 188 individuals with stomach contents.
+Twenty-three individuals had fish remnants in their stomach contents including ten with unidentifiable remnants, six Redside Shiners, four suckers, two sculpin, and one juvenile Chinook Salmon.
+In 2021, we collected stomach contents from 105 Northern Pikeminnow with 98 containing non-fish prey items and observed seven containing fish remnants.
 
@@
 
-Over the fall time period a single northern pikeminnow is modeled to eat 61 g of fish.
-If we extrapolate this out to the potential effect the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook salmon having the potential to consume ***xxx** juvenile Chinook as a discrete population.
+Over the fall time period a single Northern Pikeminnow is modeled to eat 61 g of fish.
+For the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook Salmon.
+Therefore, Northern Pikeminnow in the Deadwater Slough can potentially consume ***xx,xxx*** juvenile Chinook Salmon as a discrete population.
 
-Over the point of a whole year, a single northern pikeminnow is modeled to eat 283 g of fish and 613 g of other food types.
+Over an entire year, we estimate that a single Northern Pikeminnow will eat 283 g of fish and 613 g of other food types.
 
@@
 Additionally, we found that the fish captured in net-traps had a similar proportion of stomach content samples.
-of those, 1,214 (76.2%) were completely emply, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
-We were able to identify juvenile Red Shiners, a Largescale Sucker, Sculpin, Mountain Whitefish, and one Chinook Salmon within stomach contents, but most samples were too digested to identify to species.
+of those, 1,214 (76.2%) were completely empty, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
+We were successfully able to identify Redside Shiners, a Largescale Sucker, sculpin, Mountain Whitefish, and one juvenile Chinook Salmon in the stomach contents.
+However, decomposition from digestion rendered most stomach contents unidentifiable.
 
@@
 Several piscivorous bird species have been documented using the Deadwater Slough that include the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
-During the intial sampling event in 2019, a two-person crew walked the entire reach and surrounding and upstream areas scanning for passive integrated transponder (PIT) tags.
+During the initial sampling event in 2019, a two-person crew walked the entire reach and surrounding and upstream areas scanning for passive integrated transponder (PIT) tags.
 During that informal survey, nine PIT tags were recovered near active bird nests or in an upstream anastomizing reach where herons and eagles are prevalent, suggesting that mortality may have been a result of avian predation.
@@
 
-We estimated that consumption of juvenile Chinook Salmon by northern pikeminnow in the Deadwater Slough potentially reduces annual adult returns by ***xxx - x,xxx*** to upriver populations.
-Presumably, that reduction in adult returns impacts both the ESA-listed natural populations in the Upper Salmon River MPG plus the two hatchery populations in the Upper Salmon, Pahsimeroi and Sawtooth hatcheries, which provide for recreational fishing opportunities.
-Fisheries managers desire increased adult returns to both aide in the recovery of natural populations as well as to provide additional harvest in recreational fisheries to boost local economies.
-A reduction in predation mortality at Deadwater Slough on juvenile Chinook Salmon has potential to benefit multiple upriver natural and hatchery populations, in contrast to typical tributary habitat rehabilition actions which typically benefit a single population.
+We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces annual adult returns by ***xxx - x,xxx*** (perhaps express as percent, too) to upriver populations.
+Presumably, that reduction in adult returns impacts both the ESA-listed natural populations in the Upper Salmon River MPG and two hatchery populations in the Upper Salmon, Pahsimeroi and Sawtooth hatcheries, which provide for recrational fishing opportunities.
+Consequently, reducing juvenile Chinook Salmon predation mortalities at Deadwater Slough can potentially benefit multiple upriver natural and hatchery populations upriver.
+Because the Deadwater Slough is part of the migratory pathway for emigrating salmonids, including for multiple species and populations, the impact of that benefit would likely be higher than tributary habitat rehabilitation actions, which typically benefit a single population.
 Moreover, the deepened, slack water conditions that favor northern pikeminnow at Deadwater Slough are indirectly the result of manmade activities i.e., the failure of a manmade mining reservoir dam.

### 2021-11-23 analysis/paper/paper.Rmd
all bioenergetics plots added

@@
 )
+
 ```
@@
 ```{r shoup-temp-data}
-temp_df = read_tsv(here("analysis/data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
+shoup_temp_df = read_tsv(here("analysis/data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
                    comment = "#") %>%
@@
          water_temp_c_cd = `47773_00010_cd`)
+
+```
+
+```{r read-bio-outputs}
+# full year, no growth
+full_yr_no_growth = read_csv(here("analysis/data/raw_data/2020 Bioenergetics 1 year.csv")) %>%
+  clean_names() %>%
+  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
+  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
+  mutate(month = month(date))
+
+# full year, 10% growth
+full_yr_10p_growth = read_csv(here("analysis/data/raw_data/2020 Bioenergetics 1 year 10% growth.csv")) %>%
+  clean_names() %>%
+  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
+  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
+  mutate(month = month(date))
+
+# an alternative model
+alternative_mod = read_csv(here("analysis/data/raw_data/Bioenergetics 32 fish 68 inverts.csv")) %>%
+  clean_names() %>%
+  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
+  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
+  mutate(month = month(date))
+
+# Fall emigration period (9/1 - 11/17), no growth
+fall_only_no_growth = read_csv(here("analysis/data/raw_data/78_day_bioenergetics.csv")) %>%
+  clean_names() %>%
+  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
+  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
+  mutate(month = month(date))
 
@@
   arrange(event_name, model)
+
 ```
@@
 
+```{r fall-bio-fig, fig.cap = "First bioenergetics plot."}
+fall_bio_p = fall_only_no_growth %>%
+  ggplot() +
+  aes(x = date,
+      y = cum_cons_fish_g,
+      colour = temperature_c) +
+  geom_line(size = 2.5) +
+  scale_color_distiller(palette = "Spectral",
+                        direction = -1) +
+  scale_x_date(breaks = "2 week") +
+  scale_y_continuous(name = "Scumulative fish eaten (g)",
+                     breaks = seq(0, 200, by = 10),
+                     sec.axis = sec_axis(~. / 12, 
+                                         name = "Potential number of Chinook Salmon smolts",
+                                         breaks = seq(0, 20, by = 1))) +
+  theme_classic() +
+  theme(legend.position = "top") +
+  labs(color = "Temperature (C)")
+
+fall_bio_p
+
+```
+
+\newpage
+
+```{r full-yr-ng-fig, fig.cap = "Next bioenergetics plot."}
+full_yr_ng_p = full_yr_no_growth %>%
+  ggplot() +
+  aes(x = date,
+      y = cum_cons_fish_g,
+      colour = temperature_c) +
+  geom_line(size = 2.5) +
+  scale_color_distiller(palette = "Spectral",
+                        direction = -1) +
+  #scale_x_date(breaks = "2 week") +
+  scale_y_continuous(name = "Cumulative fish eaten (g)",
+                     breaks = seq(0, 600, by = 50),
+                     sec.axis = sec_axis(~. / 12,
+                                         name = "Potential number of Chinook Salmon smolts",
+                                         breaks = seq(0, 30, by = 2))) +
+  theme_classic() +
+  theme(legend.position = "top") +
+  labs(color = "Temperature (C)",
+       x = "Month")
+
+full_yr_ng_p
+
+```
+
+```{r full-yr-10p-fig, eval = F, fig.cap = "Another bioenergetics plot."}
+full_yr_10p_p = full_yr_10p_growth %>%
+  ggplot() +
+  aes(x = date,
+      y = cum_cons_fish_g,
+      colour = temperature_c) +
+  geom_line(size = 2.5) +
+  scale_color_distiller(palette = "Spectral",
+                        direction = -1) +
+  #scale_x_date(breaks = "2 week") +
+  scale_y_continuous(name = "Cumulative fish eaten (g)",
+                     breaks = seq(0, 600, by = 50),
+                     sec.axis = sec_axis(~. / 12,
+                                         name = "Potential number of Chinook Salmon smolts",
+                                         breaks = seq(0, 40, by = 2))) +
+  theme_classic() +
+  theme(legend.position = "top") +
+  labs(color = "Temperature (C)",
+       x = "Month")
+
+full_yr_10p_p
+
+```
+
+```{r alt-mod-fig, eval = F, fig.cap = "Another bioenergetics plot, again."}
+alt_mod_p = alternative_mod %>%
+  ggplot() +
+  aes(x = date,
+      y = cum_cons_fish_g,
+      colour = temperature_c) +
+  geom_line(size = 2.5) +
+  scale_color_distiller(palette = "Spectral",
+                        direction = -1) +
+  #scale_x_date(breaks = "2 week") +
+  scale_y_continuous(name = "Cumulative fish eaten (g)",
+                     breaks = seq(0, 600, by = 50),
+                     sec.axis = sec_axis(~. / 12,
+                                         name = "Potential number of Chinook Salmon smolts",
+                                         breaks = seq(0, 30, by = 2))) +
+  theme_classic() +
+  theme(legend.position = "top") +
+  labs(color = "Temperature (C)",
+       x = "Month")
+
+alt_mod_p
+
+```
+
+\newpage
+
 ### Colophon
@@
 if ("devtools" %in% installed.packages()) devtools::session_info()
+
 ```
@@
 if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())
+
 ```

### 2021-11-23 analysis/paper/paper.Rmd
added Nick's remaining plots to ms

@@
 In the Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of non-native Smallmouth Bass during some seasons [@Erhardt2018].
-Walleye also prefer slow water habitats and are considered a potentially important source for Juvenile Chinook Salmon mortality [@NOAA2020], particularly as their abundance has significantly increased in Snake River dam facilities between 2008 and 2020.
+Walleye also prefer slow water habitats and are considered a potentially important source for Juvenile Chinook Salmon mortality [@Widener2021], particularly as their abundance has significantly increased in Snake River dam facilities between 2008 and 2020.
 
@@
 Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the estimated 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996].
-Predation by piscivorous fish may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@NOAA2020], however there is a dearth of current data on piscivorous predator populations in the Columbia River.
+Predation by piscivorous fish may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@Widener2021], however there is a dearth of current data on piscivorous predator populations in the Columbia River.
 
@@
   mutate(non_fish_wt = stomach_contents_weight - fish_content_weight) %>%
-  dplyr::select(date, length, lavage_id, stomach_contents, comments, fish_or_parts, stomach_contents_weight, non_fish_wt,                  fish_content_weight) %>%
+  dplyr::select(date, method, length, lavage_id, stomach_contents, comments, fish_or_parts, stomach_contents_weight, non_fish_wt, fish_content_weight) %>%
   filter(!stomach_contents %in% c("Lost sample CG27-001", "Did Not Lavage")) %>%
   drop_na(stomach_contents)
+
+# hungry_fish_df = gastric_df %>%
+#   filter(fish_or_parts == "Fish or Fish Parts") %>%
+#   mutate(fish = case_when(stri_detect_fixed(stomach_contents, "Shiner") ~ "Redside Shiner",
+#                           stri_detect_fixed(stomach_contents, "Whitefish") ~ "Mountain Whitefish",
+#                           stri_detect_fixed(stomach_contents, "scuplin") ~ "Sculpin",
+#                           stri_detect_fixed(stomach_contents, "Chinook") ~ "Chinook",
+#                           stri_detect_fixed(comments, "whitefish") ~ "Mountin Whitefish",
+#                           stri_detect_fixed(comments, "scuplin") ~ "Sculpin",
+#                           stri_detect_fixed(comments, "chinook") ~ "Chinook",
+#                           TRUE ~ "Other"))
+
+```
+
+```{r gastric-summ}
+gastric_df$npm_50 = lencat(gastric_df$length, w = 50)
+
+psd_df = gastric_df %>%
+  group_by(npm_50) %>%
+  summarize(n = n())
+
+gastric_df2 = gastric_df %>%
+  drop_na(stomach_contents_weight) %>%
+  group_by(fish_or_parts)
+
+stmch_ctgry_summ = gastric_df2 %>%
+  count()
+
+stmch_wt_summ = gastric_df2 %>%
+  summarise(sum(stomach_contents_weight))
+
+non_fish_wt_summ = gastric_df2 %>%
+  summarise(sum(non_fish_wt))
+
+fish_wt_summ = gastric_df2 %>%
+  summarise(sum(fish_content_weight))
 
@@
 
+```{r stmch-cont-p, eval = F, fig.cap = "A stomach contents plot."}
+stmch_cont_per_sz_p = gastric_df %>%
+  filter(method %in% "Angling") %>%
+  ggplot() +
+  aes(x = fish_or_parts) +
+  geom_bar(fill = "#DB740E") +
+  labs(x = "Stomach Contents",
+       y = "Frequency") +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 90,
+                                   vjust = 0.5,
+                                   hjust = 1)) +
+  facet_wrap(vars(npm_50),
+             nrow = 2L)
+
+stmch_cont_per_sz_p
+
+```
+
+\newpage
+
+```{r length-freq-p, eval = T, fig.cap = "A length frequency plot."}
+length_freq_p = gastric_df %>%
+  ggplot(aes(length)) +
+  geom_histogram(breaks = seq(150, 650, by = 50),
+                 col = "#DB740E",
+                 fill = "#DB740E",
+                 alpha = .2) +
+  labs(x = "TL (mm)",
+       y = "Frequency") +
+  xlim(c(100, 700)) +
+  ylim(c(0, 800)) +
+  scale_x_continuous(breaks = seq(0, 650, by = 50)) +
+  scale_y_continuous(breaks = seq(0, 550, by = 50)) +
+  theme_classic()
+
+length_freq_p
+
+```
+
 \newpage

### 2021-11-23 analysis/paper/paper.Rmd
minor intro/discussion edits

@@
 The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
-At least five of these eight populations must meet criteria set forth by @McElhany2000 and @ICTRT2007 for the MPG to be considered viable and for the recovery of the Snake River Evolutionary Significant Unit (ESU).
-Populations within the ESU have substantial cultural value, support downriver mainstem Snake and Columbia River commercial and subsistence fisheries, and support local fisheries and economies in years with sufficient abundance.
+All populations within the Upper Salmon River MPG have become depleted in recent decades.
+Declines in survival of juvenile Chinook Salmon have been associated with hydrosystem passage (@Schaller2014), ocean conditions and regional climate (@Welch2021; @Gosselin2021), and freshwater habitat modification.
+
 All populations within the Upper Salmon River MPG have become depleted in recent decades.
@@
 
+At least five of the eight populations in the Upper Salmon MPG must meet criteria set forth by @McElhany2000 and @ICTRT2007 for the MPG to be considered viable and for the recovery of the Snake River Evolutionary Significant Unit.
+Populations with the ESU have substantial cultural value, support downriver mainstem Snake and Columbia River commercial and subsistence fisheries, and support local fisheries and economies in years with sufficient abundance.
+
 We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces annual adult returns by ***xxx - x,xxx*** (perhaps express as percent, too) to upriver populations.

### 2021-11-23 analysis/paper/paper.Rmd
added more about mark-recapture assumptions and more description of the various estimators we used.

--- 2021-11-23 analysis/paper/paper.Rmd
@@
 )
-
 ```
@@
   group_by(event_name, species) %>%
+  mutate(M = lag(R - m),
+         M = replace_na(M, 0),
+         M = cumsum(M)) %>%
   nest() %>%
@@
 
+All of our estimators rely on meeting several assumptions:
+1. The population in closed (no immigration, emigration, births or deaths during the sampling period).
+1. Marking a fish does not affect its chances of being caught
+1. No loss of marks
+1. No marks are missed or mistaken
+1. All fish have the same chance of being caught in the second (and subsequent) sample(s)
+
+We applied several abundance estimators, which fell into two broad categories: single census and multiple census.
+For the single census estimators, we treated the first week of sampling as the mark event, and the second week as the recapture event, pooling data within each of those weeks.
+The multiple census estimators treat each day as a survey, and use information about the total marked fish from all previous surveys to infer the total abundance.
+
 The Lincoln-Petersen estimator is below, where $M$ is the number of fish marked and returned to the population, $n$ is the number of fish caught in the second/recapture event and $m$ is the number of marked fish in the second sample.
@@
 The Lincoln-Petersen estimator can be biased with small samples, so we also investigated the Chapman-modified Lincoln-Petersen estimator which is shown below.
+
 $$
@@
 $$
+
 The Schnabel estimator is shown below, where the $M$, $n$ and $m$ are indexed by the sampling occasion, $i$.
@@
 This estimator does not have an associate standard error, but 95% confidence intervals can be calculated.
+The Schnabel estimator is essentially a weighted average of a series of Lincoln-Petersen estimators (with a Chapman modification).
+
 $$
@@
 $$
+
 There is another estimator for this type of multiple census surveys, called the Schumacher-Eschmeyer estimator, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
@@
 $$
-Additionally, we calculated the proportional stock density (PSD) of Northern Pikeminnow with 300 mm total length (TL) for stock and 400 mm TL for quality.
+We additionally calculated the proportional stock density (PSD) of Northern Pikeminnow with 300 mm total length (TL) for stock and 400 mm TL for quality.
 $$
@@
 ```{r shoup-temp-data}
-shoup_temp_df = read_tsv(here("analysis/data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
+temp_df = read_tsv(here("analysis/data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
                    comment = "#") %>%
@@
          water_temp_c_cd = `47773_00010_cd`)
-
-```
-
-```{r read-bio-outputs}
-# full year, no growth
-full_yr_no_growth = read_csv(here("analysis/data/raw_data/2020 Bioenergetics 1 year.csv")) %>%
-  clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
-  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
-  mutate(month = month(date))
-
-# full year, 10% growth
-full_yr_10p_growth = read_csv(here("analysis/data/raw_data/2020 Bioenergetics 1 year 10% growth.csv")) %>%
-  clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
-  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
-  mutate(month = month(date))
-
-# an alternative model
-alternative_mod = read_csv(here("analysis/data/raw_data/Bioenergetics 32 fish 68 inverts.csv")) %>%
-  clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
-  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
-  mutate(month = month(date))
-
-# Fall emigration period (9/1 - 11/17), no growth
-fall_only_no_growth = read_csv(here("analysis/data/raw_data/78_day_bioenergetics.csv")) %>%
-  clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
-  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
-  mutate(month = month(date))
 
@@
   arrange(event_name, model)
-
 ```
@@
 
-During 2019, 660 Northern Pikeminnow were gastric lavaged.
-Of those fish, we observed 603 individuals with empty stomachs, and 57 individuals with stomach contents.
-Twelve of the samples contained fish remnants including a single Mountain Whitefish and two Redside Shiners.
-During the 2020 season, 805 Northern Pikeminnow were gastric lavaged comprised of 613 individuals with empty stomachs and 188 individuals with stomach contents.
-Twenty-three individuals had fish remnants in their stomach contents including ten with unidentifiable remnants, six Redside Shiners, four suckers, two sculpin, and one juvenile Chinook Salmon.
-In 2021, we collected stomach contents from 105 Northern Pikeminnow with 98 containing non-fish prey items and observed seven containing fish remnants.
+During 2019, 660 northern pikeminnow were gastric lavaged.
+Of those fish, 603 had empty stomachs, 57 had stomach contents, 12 had fish parts, and of those 12 with fish parts 1 MWF and 2 shiners.
+During the 2020 season 805 northern pikeminnow were gastric lavaged with 613 empty stomachs, 188 with stomach contents and 23 with fish parts.
+Out of the 23 stomachs with fish parts present 10 were unknown, 6 red sided shiner, 4 suckers, 2 sculpin, and 1 chinook.
+In 2021, we had 105 stomach samples with 98 being non-fish and 7 having fish or fish parts.
 
@@
 
-Over the fall time period a single Northern Pikeminnow is modeled to eat 61 g of fish.
-For the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook Salmon.
-Therefore, Northern Pikeminnow in the Deadwater Slough can potentially consume ***xx,xxx*** juvenile Chinook Salmon as a discrete population.
+Over the fall time period a single northern pikeminnow is modeled to eat 61 g of fish.
+If we extrapolate this out to the potential effect the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook salmon having the potential to consume ***xxx** juvenile Chinook as a discrete population.
 
-Over an entire year, we estimate that a single Northern Pikeminnow will eat 283 g of fish and 613 g of other food types.
+Over the point of a whole year, a single northern pikeminnow is modeled to eat 283 g of fish and 613 g of other food types.
 
@@
 Additionally, we found that the fish captured in net-traps had a similar proportion of stomach content samples.
-of those, 1,214 (76.2%) were completely empty, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
-We were successfully able to identify Redside Shiners, a Largescale Sucker, sculpin, Mountain Whitefish, and one juvenile Chinook Salmon in the stomach contents.
-However, decomposition from digestion rendered most stomach contents unidentifiable.
+of those, 1,214 (76.2%) were completely emply, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
+We were able to identify juvenile Red Shiners, a Largescale Sucker, Sculpin, Mountain Whitefish, and one Chinook Salmon within stomach contents, but most samples were too digested to identify to species.
 
@@
 Several piscivorous bird species have been documented using the Deadwater Slough that include the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
-During the initial sampling event in 2019, a two-person crew walked the entire reach and surrounding and upstream areas scanning for passive integrated transponder (PIT) tags.
+During the intial sampling event in 2019, a two-person crew walked the entire reach and surrounding and upstream areas scanning for passive integrated transponder (PIT) tags.
 During that informal survey, nine PIT tags were recovered near active bird nests or in an upstream anastomizing reach where herons and eagles are prevalent, suggesting that mortality may have been a result of avian predation.
@@
 
-We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces annual adult returns by ***xxx - x,xxx*** (perhaps express as percent, too) to upriver populations.
-Presumably, that reduction in adult returns impacts both the ESA-listed natural populations in the Upper Salmon River MPG and two hatchery populations in the Upper Salmon, Pahsimeroi and Sawtooth hatcheries, which provide for recrational fishing opportunities.
-Consequently, reducing juvenile Chinook Salmon predation mortalities at Deadwater Slough can potentially benefit multiple upriver natural and hatchery populations upriver.
-Because the Deadwater Slough is part of the migratory pathway for emigrating salmonids, including for multiple species and populations, the impact of that benefit would likely be higher than tributary habitat rehabilitation actions, which typically benefit a single population.
+We estimated that consumption of juvenile Chinook Salmon by northern pikeminnow in the Deadwater Slough potentially reduces annual adult returns by ***xxx - x,xxx*** to upriver populations.
+Presumably, that reduction in adult returns impacts both the ESA-listed natural populations in the Upper Salmon River MPG plus the two hatchery populations in the Upper Salmon, Pahsimeroi and Sawtooth hatcheries, which provide for recreational fishing opportunities.
+Fisheries managers desire increased adult returns to both aide in the recovery of natural populations as well as to provide additional harvest in recreational fisheries to boost local economies.
+A reduction in predation mortality at Deadwater Slough on juvenile Chinook Salmon has potential to benefit multiple upriver natural and hatchery populations, in contrast to typical tributary habitat rehabilition actions which typically benefit a single population.
 Moreover, the deepened, slack water conditions that favor northern pikeminnow at Deadwater Slough are indirectly the result of manmade activities i.e., the failure of a manmade mining reservoir dam.
@@
 
-```{r fall-bio-fig, fig.cap = "First bioenergetics plot."}
-fall_bio_p = fall_only_no_growth %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Scumulative fish eaten (g)",
-                     breaks = seq(0, 200, by = 10),
-                     sec.axis = sec_axis(~. / 12, 
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 20, by = 1))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)")
-
-fall_bio_p
-
-```
-
-\newpage
-
-```{r full-yr-ng-fig, fig.cap = "Next bioenergetics plot."}
-full_yr_ng_p = full_yr_no_growth %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  #scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Cumulative fish eaten (g)",
-                     breaks = seq(0, 600, by = 50),
-                     sec.axis = sec_axis(~. / 12,
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 30, by = 2))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)",
-       x = "Month")
-
-full_yr_ng_p
-
-```
-
-```{r full-yr-10p-fig, eval = F, fig.cap = "Another bioenergetics plot."}
-full_yr_10p_p = full_yr_10p_growth %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  #scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Cumulative fish eaten (g)",
-                     breaks = seq(0, 600, by = 50),
-                     sec.axis = sec_axis(~. / 12,
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 40, by = 2))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)",
-       x = "Month")
-
-full_yr_10p_p
-
-```
-
-```{r alt-mod-fig, eval = F, fig.cap = "Another bioenergetics plot, again."}
-alt_mod_p = alternative_mod %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  #scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Cumulative fish eaten (g)",
-                     breaks = seq(0, 600, by = 50),
-                     sec.axis = sec_axis(~. / 12,
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 30, by = 2))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)",
-       x = "Month")
-
-alt_mod_p
-
-```
-
-\newpage
-
 ### Colophon
@@
 if ("devtools" %in% installed.packages()) devtools::session_info()
-
 ```
@@
 if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
-
 ```

### 2021-11-24 analysis/paper/paper.Rmd
completed draft of Intro

--- 2021-11-23 analysis/paper/paper.Rmd
@@
 
-The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
-All populations within the Upper Salmon River MPG have become depleted in recent decades.
-Declines in survival of juvenile Chinook Salmon have been associated with hydrosystem passage (@Schaller2014), ocean conditions and regional climate (@Welch2021; @Gosselin2021), and freshwater habitat modification.
-
-All populations within the Upper Salmon River MPG have become depleted in recent decades.
-Declines in survival of juvenile Chinook Salmon have been attributed to the removal of beavers from the landscape (fur trade), mining activities, river simplification, water withdrawals, logging activities, urbanization, avian predation, proliferation of non-native species (e.g., non-native coastal rainbow trout *O.
-mykiss irideus* and brook trout *Salvelinus fontinalis*), warming streams and rivers, and modifications to downriver migration corridors (e.g., from hydropower projects).
-Moreover, the abundance of returning adults are further impacted by ocean and downriver harvests, poor ocean conditions, and changes to the spawning migration corridor.
+The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017]; all of which have become depleted in recent decades.
+Declines in survival of local juvenile Chinook Salmon have been associated with hydrosystem passage [@Schaller2014], ocean conditions and regional climate [@Welch2021; @Gosselin2021], and changes to the freshwater habitat including removal of beavers, river channel simplification, water withdrawals, and the proliferation of non-native species (e.g., non-native coastal rainbow trout *O.
+mykiss irideus* and brook trout *Salvelinus fontinalis*) among others [@IdahoOSCTeam2019].
+Moreoever, the abundance of returning adults are further impacted by ocean and downriver harvests, poor ocean conditions, and changes to the spawning migration corridor.
 Each of these factors have contributed, to varying and unknown extents, to reduced adult escapement, the primary metric used to assess population viability.
@@
 Dams and reservoirs in the Columbia River basin are the primary locations associated with high rates of predation on salmonids [@Petersen1994; @Ward1995].
-There are generally two maechanisms that explain these high predation zones.
+There are generally two mechanisms that explain these high predation zones.
 First, migration rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation.
-Second, reservoirs and downstream tailraces created by dams on the Columbia and Snake rivers have created favorable slow-water habitat for several species of piscivorous fish known to consume juvenile Chinook Salmon including Walleye *Sander vitreus*, Smallmouth Bass *Micropterus dolomieu*, and Northern Pikeminnow *Ptychocheilus oregonensis*.
+Second, reservoirs and downstream tailraces created by dams on the Columbia and Snake rivers have created favorable slow-water habitat for several species of piscivorous fish known to consume juvenile Chinook Salmon including both non-native (e.g., Smallmouth Bass *Micropterus dolomieu*, Walleye *Sander vitreus*) and native (e.g., Northern Pikeminnow *Ptychocheilus oregonensis*) species.
 In the Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of non-native Smallmouth Bass during some seasons [@Erhardt2018].
-Walleye also prefer slow water habitats and are considered a potentially important source for Juvenile Chinook Salmon mortality [@Widener2021], particularly as their abundance has significantly increased in Snake River dam facilities between 2008 and 2020.
+And non-native Walleye are considered a potentially important source for juvenile Chinook Salmon mortality in the Snake River[@Widener2021], particularly as their abundance has significantly increased surrounding Snake River dam facilities between 2008 and 2020 related to their preference for slow water habitats.
 
-In addition to non-native species, Northern Pikeminnow are a native, piscivorous fish preferring slow water habitats.
-Consequently, they too have benefited from dams on the Columbia River and become abundance predators on salmonid outmigrants [@Knutsen1999].
+In addition to non-native species, Northern Pikeminnow are a native, piscivorous fish preferring warm, slow water habitats.
+Consequently, they too have benefited from dams on the Columbia River and become abundant predators on salmonid outmigrants [@Knutsen1999].
 Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the estimated 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996].
-Predation by piscivorous fish may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@Widener2021], however there is a dearth of current data on piscivorous predator populations in the Columbia River.
+Predation by piscivorous fishes, including Northern Pikeminnow, may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@Widener2021].
+However, there is a dearth of current data on piscivorous predator populations in the Columbia River.
 
-Habitat containing slow-velocities and other attributes that support piscivorous predators are not limited to dams and reservoirs.
-One such area is the Deadwater Slough in the Salmon River, which occupies a 25.3-acre reach of unnaturally slow and deep water (Figure \@ref(fig:map)).
-The Deadwater Slough contains little to no cover, increased depth, and fine substrate which favor piscivorous fish predators (e.g., Northern Pikeminnow, Smallmouth Bass) [@Watkins2015].
-@Axel2015 demonstrated decreased rates of emigration and apparent survival for Sockeye Salmon *Oncorhynchus nerka* from Redfish Lake emigrating during the spring.
-Further, recent winter telemetry studies have indicated decreased transition probabilities (approximately 10% less than surrounding reaches) of juvenile Chinook salmon through Deadwater Slough during fall and winter months [@Ackerman2018; @Porter2019].
-
-<!--
-* @Axel2015 found increased mortality in a particular reach spanning location x to location y in the mainstem Salmon River on a study of juvenile (hatchery?) sockeye salmon released near Redfish Lake and actively migrating downriver.
-  - in their case, these were actively migrating juveniles, and so transitional probabilities can likely be interpreted (almost) directly as survival
-  - A previous study showed that the lowest transitional probabilities of Red Fish Lake Sockeye occurred in the Dead Water reach.
-* @Ackerman2018 and @Porter2019, @Porter2020 found reduced (apparent) survival in a reach spanning near the town of North Fork to just downstream of Deadwater Slough.
-  - in these cases, there's a chance that juvenile Chinook could exhibit concealment behavior prior to winter and so transition probability is interpreted as the probability of a fish surviving AND moving
-* Anecdotally, locals (and others? maybe NOAA?) informed us that northern pikeminnow and perhaps other species are prevalent in Deadwater Slough.
--->
+Habitats containing slower water velocities and other attributes that support piscivorous predators are not limited to dams and reservoirs.
+One such area is the Deadwater Slough in the Salmon River, Idaho which is an unnaturally slow and deep reach of the mainstem river approximately 1.5 kilometers in length and occupying 25.3 acres (Figure \@ref(fig:map)).
+The slough contains "reservoir-like" conditions which typically favor piscivorous fish predators like Northern Pikeminnow or Smallmouth Bass [@Watkins2015], and further, lacks hydrological and structural features and fish cover (e.g., large wood or substrate, woody debris, bank undercuts) that can provide essential refuge from predation for juvenile fishes.
+As a result, predation of juvenile Chinook Salmon and other local ESA-listed salmonids (Sockeye Salmon *O.
+nerka*, redband steelhead *O.
+mykiss gairdneri*) has been cited as a concern for populations in the Upper Salmon River
 
 Spring/summer Chinook Salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014].
-The DSR migrants leave the natal spawning area as subyearlings between June and November and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts.
+The DSR migrants leave the natal spawning area as subyearlings between June and november and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts.
 Alternatively, NRR migrants remain in their natal spawning areas for approximately one year after emergence until emigration to the ocean as smolts.
@@
 
-The Deadwater Slough is in a reach of the Salmon River that is believed to be a historically important overwinter rearing area for DSR emigrants.
-Moreover, this reach is part of the migratory pathway for juvenile DSR and NRR emigrants from all eight extant populations.
-The slough lacks hydrological and structural features (i.e., a homogenous channel with fine substrate and little cover) that can provide essential refuge from predation.
-As a result, predation on juvenile Chinook salmon proximal to the Deadwater Slough has been cited as a concern for the Upper Salmon River MPG, impacting DSR migrants in the fall and NRR emigrants during the spring.
+The Deadwater Slough is in a reach of the Salmon River believed to be a historically important overwinter rearing area for DSR emigrants.
+Moreover, the slough is part of the migratory pathway for not only juvenile Chinook Salmon (both DSR and NRR emigrants), but also for the endangered Snake River Sockeye Salmon population and several population of threatened Snake River steelhead.
+As such, the main Salmon River has been the target of recent radio telemetry studies to examine the downstream movement, distribution, and (apparent survival) of juvenile salmonid emigrants.
+@Axel2015 demonstrated decreased rates of emigration and survival for Sockeye Salmon from Redfish Lake during the spring emigration.
+Further, recent telemetry studies of Chinook salmon emigration during fall and early winter months indicated a decreased transition probability in the reach spanning Deadwater Slough (approximately 10% less than surrounding reaches).
 

### 2021-11-24 analysis/paper/paper.Rmd
knit docs w/ Kevin's recent edits

@@
 output: 
-    bookdown::html_document2:
-      fig_caption: yes
-      number_sections: FALSE
-      toc: no
-      fig_width: 6
-      pandoc_args:
-      - --lua-filter=../templates/scholarly-metadata.lua
-      - --lua-filter=../templates/author-info-blocks.lua
-      - --lua-filter=../templates/pagebreak.lua      
     bookdown::word_document2:

### 2021-11-24 analysis/paper/paper.Rmd
added SAR data, and reference, with pdf. added some preliminary impacts to adult returns to the manuscript

--- 2021-11-24 analysis/paper/paper.Rmd
@@
 
-The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017]; all of which have become depleted in recent decades.
-Declines in survival of local juvenile Chinook Salmon have been associated with hydrosystem passage [@Schaller2014], ocean conditions and regional climate [@Welch2021; @Gosselin2021], and changes to the freshwater habitat including removal of beavers, river channel simplification, water withdrawals, and the proliferation of non-native species (e.g., non-native coastal rainbow trout *O.
-mykiss irideus* and brook trout *Salvelinus fontinalis*) among others [@IdahoOSCTeam2019].
-Moreoever, the abundance of returning adults are further impacted by ocean and downriver harvests, poor ocean conditions, and changes to the spawning migration corridor.
+The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
+All populations within the Upper Salmon River MPG have become depleted in recent decades.
+Declines in survival of juvenile Chinook Salmon have been associated with hydrosystem passage (@Schaller2014), ocean conditions and regional climate (@Welch2021; @Gosselin2021), and freshwater habitat modification.
+
+All populations within the Upper Salmon River MPG have become depleted in recent decades.
+Declines in survival of juvenile Chinook Salmon have been attributed to the removal of beavers from the landscape (fur trade), mining activities, river simplification, water withdrawals, logging activities, urbanization, avian predation, proliferation of non-native species (e.g., non-native coastal rainbow trout *O.
+mykiss irideus* and brook trout *Salvelinus fontinalis*), warming streams and rivers, and modifications to downriver migration corridors (e.g., from hydropower projects).
+Moreover, the abundance of returning adults are further impacted by ocean and downriver harvests, poor ocean conditions, and changes to the spawning migration corridor.
 Each of these factors have contributed, to varying and unknown extents, to reduced adult escapement, the primary metric used to assess population viability.
@@
 Dams and reservoirs in the Columbia River basin are the primary locations associated with high rates of predation on salmonids [@Petersen1994; @Ward1995].
-There are generally two mechanisms that explain these high predation zones.
+There are generally two maechanisms that explain these high predation zones.
 First, migration rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation.
-Second, reservoirs and downstream tailraces created by dams on the Columbia and Snake rivers have created favorable slow-water habitat for several species of piscivorous fish known to consume juvenile Chinook Salmon including both non-native (e.g., Smallmouth Bass *Micropterus dolomieu*, Walleye *Sander vitreus*) and native (e.g., Northern Pikeminnow *Ptychocheilus oregonensis*) species.
+Second, reservoirs and downstream tailraces created by dams on the Columbia and Snake rivers have created favorable slow-water habitat for several species of piscivorous fish known to consume juvenile Chinook Salmon including Walleye *Sander vitreus*, Smallmouth Bass *Micropterus dolomieu*, and Northern Pikeminnow *Ptychocheilus oregonensis*.
 In the Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of non-native Smallmouth Bass during some seasons [@Erhardt2018].
-And non-native Walleye are considered a potentially important source for juvenile Chinook Salmon mortality in the Snake River[@Widener2021], particularly as their abundance has significantly increased surrounding Snake River dam facilities between 2008 and 2020 related to their preference for slow water habitats.
-
-In addition to non-native species, Northern Pikeminnow are a native, piscivorous fish preferring warm, slow water habitats.
-Consequently, they too have benefited from dams on the Columbia River and become abundant predators on salmonid outmigrants [@Knutsen1999].
+Walleye also prefer slow water habitats and are considered a potentially important source for Juvenile Chinook Salmon mortality [@Widener2021], particularly as their abundance has significantly increased in Snake River dam facilities between 2008 and 2020.
+
+In addition to non-native species, Northern Pikeminnow are a native, piscivorous fish preferring slow water habitats.
+Consequently, they too have benefited from dams on the Columbia River and become abundance predators on salmonid outmigrants [@Knutsen1999].
 Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the estimated 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996].
-Predation by piscivorous fishes, including Northern Pikeminnow, may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@Widener2021].
-However, there is a dearth of current data on piscivorous predator populations in the Columbia River.
-
-Habitats containing slower water velocities and other attributes that support piscivorous predators are not limited to dams and reservoirs.
-One such area is the Deadwater Slough in the Salmon River, Idaho which is an unnaturally slow and deep reach of the mainstem river approximately 1.5 kilometers in length and occupying 25.3 acres (Figure \@ref(fig:map)).
-The slough contains "reservoir-like" conditions which typically favor piscivorous fish predators like Northern Pikeminnow or Smallmouth Bass [@Watkins2015], and further, lacks hydrological and structural features and fish cover (e.g., large wood or substrate, woody debris, bank undercuts) that can provide essential refuge from predation for juvenile fishes.
-As a result, predation of juvenile Chinook Salmon and other local ESA-listed salmonids (Sockeye Salmon *O.
-nerka*, redband steelhead *O.
-mykiss gairdneri*) has been cited as a concern for populations in the Upper Salmon River
+Predation by piscivorous fish may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@Widener2021], however there is a dearth of current data on piscivorous predator populations in the Columbia River.
+
+Habitat containing slow-velocities and other attributes that support piscivorous predators are not limited to dams and reservoirs.
+One such area is the Deadwater Slough in the Salmon River, which occupies a 25.3-acre reach of unnaturally slow and deep water (Figure \@ref(fig:map)).
+The Deadwater Slough contains little to no cover, increased depth, and fine substrate which favor piscivorous fish predators (e.g., Northern Pikeminnow, Smallmouth Bass) [@Watkins2015].
+@Axel2015 demonstrated decreased rates of emigration and apparent survival for Sockeye Salmon *Oncorhynchus nerka* from Redfish Lake emigrating during the spring.
+Further, recent winter telemetry studies have indicated decreased transition probabilities (approximately 10% less than surrounding reaches) of juvenile Chinook salmon through Deadwater Slough during fall and winter months [@Ackerman2018; @Porter2019].
+
+<!--
+* @Axel2015 found increased mortality in a particular reach spanning location x to location y in the mainstem Salmon River on a study of juvenile (hatchery?) sockeye salmon released near Redfish Lake and actively migrating downriver.
+  - in their case, these were actively migrating juveniles, and so transitional probabilities can likely be interpreted (almost) directly as survival
+  - A previous study showed that the lowest transitional probabilities of Red Fish Lake Sockeye occurred in the Dead Water reach.
+* @Ackerman2018 and @Porter2019, @Porter2020 found reduced (apparent) survival in a reach spanning near the town of North Fork to just downstream of Deadwater Slough.
+  - in these cases, there's a chance that juvenile Chinook could exhibit concealment behavior prior to winter and so transition probability is interpreted as the probability of a fish surviving AND moving
+* Anecdotally, locals (and others? maybe NOAA?) informed us that northern pikeminnow and perhaps other species are prevalent in Deadwater Slough.
+-->
 
 Spring/summer Chinook Salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014].
-The DSR migrants leave the natal spawning area as subyearlings between June and november and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts.
+The DSR migrants leave the natal spawning area as subyearlings between June and November and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts.
 Alternatively, NRR migrants remain in their natal spawning areas for approximately one year after emergence until emigration to the ocean as smolts.
@@
 
-The Deadwater Slough is in a reach of the Salmon River believed to be a historically important overwinter rearing area for DSR emigrants.
-Moreover, the slough is part of the migratory pathway for not only juvenile Chinook Salmon (both DSR and NRR emigrants), but also for the endangered Snake River Sockeye Salmon population and several population of threatened Snake River steelhead.
-As such, the main Salmon River has been the target of recent radio telemetry studies to examine the downstream movement, distribution, and (apparent survival) of juvenile salmonid emigrants.
-@Axel2015 demonstrated decreased rates of emigration and survival for Sockeye Salmon from Redfish Lake during the spring emigration.
-Further, recent telemetry studies of Chinook salmon emigration during fall and early winter months indicated a decreased transition probability in the reach spanning Deadwater Slough (approximately 10% less than surrounding reaches).
+The Deadwater Slough is in a reach of the Salmon River that is believed to be a historically important overwinter rearing area for DSR emigrants.
+Moreover, this reach is part of the migratory pathway for juvenile DSR and NRR emigrants from all eight extant populations.
+The slough lacks hydrological and structural features (i.e., a homogenous channel with fine substrate and little cover) that can provide essential refuge from predation.
+As a result, predation on juvenile Chinook salmon proximal to the Deadwater Slough has been cited as a concern for the Upper Salmon River MPG, impacting DSR migrants in the fall and NRR emigrants during the spring.
 
@@
 
-Text here...
+```{r}
+# pull in SAR numbers
+sar = read_csv(here("analysis/data/raw_data", 
+                    "WSSCh_sar_4See.csv")) %>%
+  summarise(across(c(medsar, sdsar),
+                   median)) %>%
+  summarise(across(c(medsar, sdsar),
+               ~ . / 100))
+
+adult_impact = fall_only_no_growth %>%
+  filter(cum_cons_fish_g == max(cum_cons_fish_g)) %>%
+  select(cum_cons_fish_g) %>%
+  mutate(n_chnk_smolts = cum_cons_fish_g / 12) %>%
+  bind_cols(all_est %>%
+              select(event_name, 
+                     model,
+                     N,
+                     Lci,
+                     Uci)) %>%
+  mutate(across(c(smolt_N = N, 
+                  smolt_Lci = Lci, 
+                  smolt_Uci = Uci),
+                ~ . * n_chnk_smolts)) %>%
+  bind_cols(sar) %>%
+  mutate(adult_N = smolt_N * medsar,
+         adult_Lci = smolt_Lci * qnorm(0.025, medsar, sdsar),
+         adult_Uci = smolt_Uci * qnorm(0.975, medsar, sdsar))
+```
+
+To estimate the number of possible Chinook salmon that could be consumed, we divided the cumulative fish consumption in grams for an individual pikeminnow by the average weight of a Chinook emigrant to come up with a presmolt equivalency and then multiplied this by our abundance estimate of northern pikeminnow in Deadwater Slough, resulting in an estimate of the equivalent total number of Chinook salmon presmolts consumed.
+Finally, we estimated the number of additional adult Chinook salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough reach was reduced or eliminated.
+We multiplied the total estimated consumed juvenile Chinook salmon by an estimate of Granite-to-Granite SARs [@McCann2019].
 
 <!-- After making some assumptions (e.g., on SARs) attempt to estimate how results from objective #3 could translate into impacts on adult returns (via a thought experiment or simulation approach) -->
+
+<!-- Currently only using one bioenergetics model run.
+Could incorporate more bioenergetics models, but if so then maybe consider focusing on a single abundance estimator (e.g.
+Schnabel) to avoid too many combinations in the results -->
 
@@
 * What is that equate to as a density? e.g., how many pikeminnow per 10 square meters? Is that a reasonable number?
+* To calculate density, need the area of the extent surveyed each year. Was that area consistent year-to-year?
 -->
 
-PSD for northern pikeminnow across all three years was 42.26% or 42.26% of fish were over the quality size classification 400 mm TL we denoted in the Methods, demonstrating a high percentage of above average size classs present within Deadwater Slough.
+Abundance estimates for northern pikeminnow during the fall sampling events ranged from `r prettyNum(min(all_est$N), big.mark = ",")` to `r prettyNum(max(all_est$N), big.mark = ",")`.
+The multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied.
+
+PSD for northern pikeminnow across all three years was 42.26% or 42.26% of fish were over the quality size classification 400 mm TL we denoted in the Methods, demonstrating a high percentage of above average size classes present within Deadwater Slough.
 
@@
 
+```{r adult-imp-p}
+adult_impact %>%
+  mutate(stage = "Smolt") %>%
+  select(event_name, model,
+         stage,
+         starts_with("smolt")) %>%
+  rename_with(.fn = ~ str_remove(., "smolt_"),
+              .cols = starts_with("smolt")) %>%
+  bind_rows(adult_impact %>%
+              mutate(stage = "Adult Equivalent") %>%
+              select(event_name, model,
+                     stage,
+                     starts_with("adult")) %>%
+              rename_with(.fn = ~ str_remove(., "adult_"),
+                          .cols = starts_with("adult"))) %>%
+  filter(str_detect(model, "Schuma", negate = T)) %>%
+  ggplot(aes(x = model,
+             y = N,
+             color = model)) +
+  geom_errorbar(aes(ymin = Lci,
+                    ymax = Uci),
+                width = 0.5) +
+  geom_point(size = 4) +
+  facet_grid(stage ~ event_name,
+             scales = "free_y") +
+  scale_color_brewer(palette = "Set1",
+                     name = "Model") +
+  theme(axis.text.x = element_blank(),
+        legend.position = "bottom") +
+  labs(x = "N. Pikeminnow Abundance Model")
+```
+
+\newpage
+
 ### Colophon

### 2021-11-24 analysis/paper/paper.Rmd
moved a chunk of code

@@
 
+```{r abund-est}
+all_est = LP_mod %>%
+  select(event_name, species, model, N:Uci) %>%
+  bind_rows(N_mods %>%
+             select(event_name, species, model, N:Uci)) %>%
+  mutate(range = Uci - Lci,
+         prop_range = range / N) %>%
+  mutate(across(event_name,
+                ~ str_replace(., "_", " "))) %>%
+  mutate(model = fct_recode(model,
+                            "Schumacher-Eschmeyer" = "SchumacherEschmeyer")) %>%
+  arrange(event_name, model)
+
+```
+
 ```{r}
@@
          adult_Uci = smolt_Uci * qnorm(0.975, medsar, sdsar))
+
 ```
@@
 ## Abundance of Piscivorous Fishes
-
-```{r abund-est}
-all_est = LP_mod %>%
-  select(event_name, species, model, N:Uci) %>%
-  bind_rows(N_mods %>%
-             select(event_name, species, model, N:Uci)) %>%
-  mutate(range = Uci - Lci,
-         prop_range = range / N) %>%
-  mutate(across(event_name,
-                ~ str_replace(., "_", " "))) %>%
-  mutate(model = fct_recode(model,
-                            "Schumacher-Eschmeyer" = "SchumacherEschmeyer")) %>%
-  arrange(event_name, model)
-
-```
 

### 2021-11-24 analysis/paper/paper.Rmd
updated Study Site

@@
 Habitats containing slower water velocities and other attributes that support piscivorous predators are not limited to dams and reservoirs.
-One such area is the Deadwater Slough in the Salmon River, Idaho which is an unnaturally slow and deep reach of the mainstem river approximately 1.5 kilometers in length and occupying 25.3 acres (Figure \@ref(fig:map)).
+One such area is the Deadwater Slough in the Salmon River, Idaho which is an unnaturally slow and deep reach of the mainstem river approximately 1.5 kilometers in length and occupying 29.5 acres (Figure \@ref(fig:map)).
 The slough contains "reservoir-like" conditions which typically favor piscivorous fish predators like Northern Pikeminnow or Smallmouth Bass [@Watkins2015], and further, lacks hydrological and structural features and fish cover (e.g., large wood or substrate, woody debris, bank undercuts) that can provide essential refuge from predation for juvenile fishes.
@@
 
-The Deadwater Slough is an approximately 1.5 kilometer section of the mainstem Salmon River located roughly 5.8 river kilometers downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
+The Deadwater Slough is located approximately 6 kms downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
 The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
-Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River, thereby creating an unnaturally slow and deep section in the river, spanning approximately 30 acres and averaging 68 m width.
-Both northern pikeminnow *Ptychocheilus oregonensis* and smallmouth bass *Micropterus dolomieu* inhabit the slough which likely provides favorable conditions for their feeding and growth (e.g., reduced water velocity, deep channel, warmer water temperature).
+Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River, forming an alluvial fan and creating an unnaturally slow and deep section of the river resembling a small reservoir.
 

### 2021-11-24 analysis/paper/paper.Rmd
added methods for fish capture, data collection

@@
 
+Our goal was to estimate the abundance of potentially piscivorous predators in the Deadwater Slough using a mark-recapture approach both during the fall and spring near the peaks of the juvenile Chinook Salmon DSR and NRR emigrations, respectively.
+The initial intent was to sample during fall 2019 and spring 2020 to estimate predator abundances during the emigration of the 2018 brood year.
+However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 sampling event was delayed until fall 2020.
+Although this provided the benefit of an additional abundance estimates during the DSR emigration, estimates during the NRR emigration were still lacking, and so an additional sampling event was added for spring 2021.
+The study design was such that both the fall 2019 and fall 2020 sampling efforts included both a "mark week" and "recapture week" intended for mark-recapture models.
+However, due to budget constraints the spring 2021 effort only included a single "mark week"; instead, we calculated the catch-per-unit-effort (CPUE; fish per hour) for the spring 2021 effort for comparison to the fall efforts.
+In the end, sampling dates were as follows: fall 2019 (mark = November 12-14; recapture = November 19-21), fall 2020 (mark = October 20-23; recapture = October 27-30); spring 2021 (mark = May 18-21).
+The recapture events were held just one week following the mark events to reduce opportunity for immigration/emigration of predators between events i.e., to avoid violation of a closed population for mark-recapture models.
+
+We attempted multiple capture methods during the first year including raft electrofishing, fyke netting, snorkeling, and angling to reduce selectivity, both for species and size classes.
+Unfortunately, some of the methods proved ineffective.
+For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) we lacked the appropriate motor to effectively “chase” fish given the lack of flow.
+Snorkeling was ineffective due to the high turbidity of the mainstem Salmon River and Deadwater Slough.
+And due to debris floating in the water column, there were concerns with fyke nets being pulled from their anchored locations and becoming entangled.
+Additionally, there were concerns with raft electrofishers and fyke nets incidentally capturing adult steelhead that are present in the mainstem Salmon River during fall and spring.
+In the end, angling proved to be the most effective method for capturing piscivorous predators while minimizing potential impacts to steelhead but recognizing that angling can have both species and size selectivity (bias).
+From here on, our focus was on fish captured by angling, unless otherwise noted.
+
+Our study relied heavily on volunteer anglers who were allowed to fish anywhere within Deadwater Slough; periodically or upon filling a livewell anglers would boat or hike their catch to a processing station located at the boat ramp approximately 500 m downstream from the top of the slough.
+For each fish, processors would record the date of capture, species, total length (TL; mm), and whether the fish was previously marked.
+If not previously marked, each fish was provided a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) and marks were unique for each marking day.
+We also recorded whether marked individuals were released back to be captured again.
+For release, processors attempted to distribute released individuals throughout Deadwater Slough to facilitate mixing back into the population.
+We also recorded the crew (person or combination of persons) that captured each fish and the start and end times for each crew were recorded to estimate total effort.
+
 ```{r cmr-effort-data}
@@
 
-**Data = C/M/R and Effort**
-
-* Capture methods used - We initially attempted electrofishing, snorkeling, netting, others?, but eventually settled on angling as best method for mark-recapture.
-* Mark-recapture - We kept the mark and recapture events close together in an attempt to use a closed population model
-* What are the assumptions of a closed population model?
-* How was sampling performed? What relevant information was captured for each fish? How/where were fish released? How was effort recorded to calculate CPUE?
-* Description of sampling time frame:
-  - Fall 2019: Goal was to estimate abundance during the peak of DSR emigration i.e., timing was done to coincide to be shortly after peak fall (DSR) emigration at the lower Lemhi River screw trap with the hopes of documenting predation.
-DSR are the more abundant juvenile emigration tactic.
-  - Fall 2020: Intent was to repeat the above, but instead during peak NRR emigration, but effort canceled due to COVID-19.
-Instead, sampling was re-schedule to fall when social distancing, etc.
-could be put in place.
-  - Spring 2021: Did a relative abundance effort during spring 2021.
-Had to reduce to a CPUE (single-week) effort due to funding constraints.
-However, goal was to 1) document presence during NRR outmigration and 2) estimate CPUE to provide a relative comparison to the fall efforts.
-
 All of our estimators rely on meeting several assumptions:

### 2021-11-24 analysis/paper/paper.Rmd
minor edits to mark-recap meth

@@
 
-All of our estimators rely on meeting several assumptions:
-1. The population in closed (no immigration, emigration, births or deaths during the sampling period).
+We used several mark-recapture abundance estimators that each rely on meeting several assumptions:
+
+1. The population in closed (no immigration, emigration, births or deaths during the sampling period)
 1. Marking a fish does not affect its chances of being caught
@@
 
-We applied several abundance estimators, which fell into two broad categories: single census and multiple census.
-For the single census estimators, we treated the first week of sampling as the mark event, and the second week as the recapture event, pooling data within each of those weeks.
-The multiple census estimators treat each day as a survey, and use information about the total marked fish from all previous surveys to infer the total abundance.
-
-The Lincoln-Petersen estimator is below, where $M$ is the number of fish marked and returned to the population, $n$ is the number of fish caught in the second/recapture event and $m$ is the number of marked fish in the second sample.
+The estimators fell into two broad categories: single census and multiple census.
+For the single census estimators, we treated the first week of sampling as the mark event, and the following (second) week as the recapture event, pooling data within each of those two weeks.
+Alternatively, the multiple census estimators treat each day as a survey, and use information about the total marked fish from all previous surveys to infer the total abundance.
+
+The Lincoln-Peterson estimator, a single-census estimator, is below, where $M$ is the number of fish marked and returned to the population, the $n$ is the number of fish caught in the second/recapture event, and $m$ is the number of marked fish in the second sample.
 
@@
 
-The Lincoln-Petersen estimator can be biased with small samples, so we also investigated the Chapman-modified Lincoln-Petersen estimator which is shown below.
+The Lincoln-Petersen estimator can be biased with small samples, so we also investigated the Chapman-modified Lincoln-Petersen estimator:
 

### 2021-11-24 analysis/paper/paper.Rmd
added comment to do some additional work

@@
 
+<!-- use ratio of CPUEs between fall and spring, and assume same capture probability in both periods to estimate abundance in the spring.
+Recognize that capture probability is probably smaller in the spring, so our estimate is a conservative one (the truth is probably higher).
+Then re-do the bioenergetics model to look at how many smolts are consumed in the spring.
+-->
+
 ## Predation On Juvenile Chinook Salmon

### 2021-11-24 analysis/paper/paper.Rmd
calculated NP abundance in the spring using ratios of CPUE
added some notes to the Bioenergetics methods section about some additional things to add.

@@
 
+```{r abund-est}
+all_est = LP_mod %>%
+  select(event_name, species, model, N:Uci) %>%
+  bind_rows(N_mods %>%
+             select(event_name, species, model, N:Uci)) %>%
+  mutate(range = Uci - Lci,
+         prop_range = range / N) %>%
+  mutate(across(event_name,
+                ~ str_replace(., "_", " "))) %>%
+  mutate(model = fct_recode(model,
+                            "Schumacher-Eschmeyer" = "SchumacherEschmeyer")) %>%
+  arrange(event_name, model)
+
+```
+
+```{r spring-abund}
+total_cpue = cpue_df %>%
+              mutate(across(event_name,
+                            ~ str_replace(., "_", " "))) %>%
+              group_by(species, event_name) %>%
+              summarise(across(c(total_effort, n_fish),
+                               sum),
+                        .groups = "drop") %>%
+              mutate(across(total_effort,
+                            ~ . / 60)) %>%
+              mutate(cpue = n_fish / total_effort)
+
+spring_abund = all_est %>%
+  select(-range, -prop_range) %>%
+  inner_join(total_cpue,
+             by = c("event_name", "species")) %>%
+  select(species, model, 
+         model_event = event_name,
+         N_F = N, 
+         Lci_F = Lci,
+         Uci_F = Uci,
+         cpue_F = cpue) %>%
+  full_join(total_cpue %>%
+              filter(str_detect(event_name, "Spring")) %>%
+              select(species,
+                     cpue_S = cpue),
+            by = "species") %>%
+  mutate(N_S = N_F * (cpue_S / cpue_F),
+         Lci_S = Lci_F * (cpue_S / cpue_F),
+         Uci_S = Uci_F * (cpue_S / cpue_F))
+
+```
+
 We used several mark-recapture abundance estimators that each rely on meeting several assumptions:
@@
 -->
+
+To estimate the abundance of Northern Pikeminnow in the spring, we used the ratio of total CPUE in a fall sampling event (total Northern Pikeminnow caught / effort) to the total CPUE in the spring, and multiply that ratio by the abundance of Northern Pikeminnow in the fall.
+This assumes the same capture probability in the fall as in the spring.
+If that does differ, we believe the capture probability to be lower in the spring due to higher flows, which would lead to our spring abundance estimate to be biased low.
+Therefore, we consider this a conservative estimate of the number of Northern Pikeminnow present in the Deadwater area during the spring Chinook smolt out migration.
 
@@
 
+A few things that need to be clarified in this section:
+* Why were the dates Sept 1 - Nov 17 chosen? They make sense, but we should include our rationale
+* Where did the temperature data come from? Hobo? Stream gauge? The methods say temperature data from 2019 was used, but the results seem to use 2020? 
+* Why does it look like we only ran it for dates in 2020? Is that the only year we had temperature data for? Would the results look different if we used temperature data from a different year?
+* The lengths of pikeminnow: did we only use pikeminnow we caught in the fall to get average length? Is there a difference in lengths between 2019 and 2020? 
+* There needs to be more explanation of the inverts vs. fish. What proportion went to each and why? Did it change through time? Based on literature at all? 
+* Where is the data that was used to get a weight of 12 g for Chinook emigrant?
+Was that for Chinook presmolts, or a combination of smolts and presmolts?
+If the latter, can we filter it to look at the average weight for presmolts only?
+And maybe smolts only for a different analysis?
+* Could we run a similar model for the spring emigration period?
+Figure out the dates, look at lengths of pikeminnow from the spring, convert to average weight, use average size of Chinook smolts (not presmolts) from all the Upper Salmon traps?
+Kevin can work on converting the spring catch into some kind of estimate of pikeminnow at Deadwater in the spring
+
 ```{r shoup-temp-data}
@@
 The second model was run for one-year and assumed no growth for Northern Pikeminnow.
-The assumption of no growth for pikeminnow was included to show what may be occurring if the populaiton of Northern Pikeminnow is stable.
-If growth of the population is occurring, the alternative model shows how an increase of 10% body weight changes the estimtes of consumption over a full year period.
+The assumption of no growth for pikeminnow was included to show what may be occurring if the population of Northern Pikeminnow is stable.
+If growth of the population is occurring, the alternative model shows how an increase of 10% body weight changes the estimates of consumption over a full year period.
 To estimate the number of possible Chinook Salmon that could be consumed, we assumed a weight of 12 g per individual Chinook Salmon.
@@
 ## Impacts to Adult Returns
-
-```{r abund-est}
-all_est = LP_mod %>%
-  select(event_name, species, model, N:Uci) %>%
-  bind_rows(N_mods %>%
-             select(event_name, species, model, N:Uci)) %>%
-  mutate(range = Uci - Lci,
-         prop_range = range / N) %>%
-  mutate(across(event_name,
-                ~ str_replace(., "_", " "))) %>%
-  mutate(model = fct_recode(model,
-                            "Schumacher-Eschmeyer" = "SchumacherEschmeyer")) %>%
-  arrange(event_name, model)
-
-```
 

### 2021-11-24 analysis/paper/paper.Rmd
moved Impacts to Adult Returns to Discussion

@@
 
+<!--
 A few things that need to be clarified in this section:
@@
 Kevin can work on converting the spring catch into some kind of estimate of pikeminnow at Deadwater in the spring
+-->
 
@@
 -->
+
+# Results
+
+## Abundance of Piscivorous Fishes
+
+<!--
+* Total population estimates of Pike Minnow. Chart or table, but likely a chart.
+* What is that equate to as a density? e.g., how many pikeminnow per 10 square meters? Is that a reasonable number?
+* To calculate density, need the area of the extent surveyed each year. Was that area consistent year-to-year?
+-->
+
+Abundance estimates for northern pikeminnow during the fall sampling events ranged from `r prettyNum(min(all_est$N), big.mark = ",")` to `r prettyNum(max(all_est$N), big.mark = ",")`.
+The multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied.
+
+PSD for northern pikeminnow across all three years was 42.26% or 42.26% of fish were over the quality size classification 400 mm TL we denoted in the Methods, demonstrating a high percentage of above average size classes present within Deadwater Slough.
+
+## Predation On Juvenile Chinook Salmon
+
+During 2019, 660 Northern Pikeminnow were gastric lavaged.
+Of those fish, we observed 603 individuals with empty stomachs, and 57 individuals with stomach contents.
+Twelve of the samples contained fish remnants including a single Mountain Whitefish and two Redside Shiners.
+During the 2020 season, 805 Northern Pikeminnow were gastric lavaged comprised of 613 individuals with empty stomachs and 188 individuals with stomach contents.
+Twenty-three individuals had fish remnants in their stomach contents including ten with unidentifiable remnants, six Redside Shiners, four suckers, two sculpin, and one juvenile Chinook Salmon.
+In 2021, we collected stomach contents from 105 Northern Pikeminnow with 98 containing non-fish prey items and observed seven containing fish remnants.
+
+<!--- Need to add Nick's gastric lavage table here -->
+
+## Bioenergetics
+
+Over the fall time period a single Northern Pikeminnow is modeled to eat 61 g of fish.
+For the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook Salmon.
+Therefore, Northern Pikeminnow in the Deadwater Slough can potentially consume ***xx,xxx*** juvenile Chinook Salmon as a discrete population.
+
+Over an entire year, we estimate that a single Northern Pikeminnow will eat 283 g of fish and 613 g of other food types.
+
+<!--
+* Water Temperature measurements.
+    *	Hobo tidbit placed at site. 
+    *	Calculated daily average water temperature for 365 days of the year.
+*	Water temperature plot for the 365 average.
+*	Bioenergetics results. Total mass consumed over a 365 day period. Plot of daily consumption with day of year on x and mass on y.
+-->
+
+# Discussion
+
+We estimated the population size of Northern Pikeminnow in the Deadwater Slough to be greater than ***xx,xxx*** during the fall emigration period for DSR Chinook Salmon.
+That estimate translates to a density of ***xxx*** Northern Pikeminnow per 100 m or ***xxx*** per 100 m^2^ which is ***similar/more/less*** than estimates from elsewhere in the Columbia River ***(citation)*** where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminno abundance.
+The population size of Northern Pikeminnow was not directly estimated during the spring NRR Chinook salmon emigration period however, the relative abundance measured at CPUE was comparable to the fall sampling periods ***(update statement later)***.
+The population of Northern Pikeminnow in Deadwater Slough was estimated to consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook Salmon during the ***x*** sampling periods and result in an estimated reduction of returning adults between ***xxx*** and ***x,xxx***.
+We suggest that the habitat modifications that created the Deadwater Slough have resulted in favorable conditions for Northern Pikeminnow, including improved conditions for predation upon juvenile Chinook Salmon ***(add detail here)***.
+Therefore, predation by Northern Pikeminnow in the Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG.
+
+## Mark-Recapture Model
+
+<!--
+Re-hash out our assumptions?
+What assumptions were most likely violated e.g., the closed population assumption?
+If we assumed an open population with varying immigration/emigration rates how might that affect our estimate?
+In the end, how good do we think our estimate is and is the bottom-line that we belive there still to be a shit-ton of pikeminnow even if our assumptions were violated?
+-->
+
+## Gastric Lavage
+
+The majority of fish collected during this study received gastric lavage, including some non-predatory species.
+To validate the efficacy of this method, we euthanized nine Northern Pikeminnow after gastric lavage was completed and removed the remaining stomach contents via dissection.
+Our results support previous findings that gastric lavage effectively removes stomach contents (@Lott2020).
+Additionally, we found that the fish captured in net-traps had a similar proportion of stomach content samples.
+of those, 1,214 (76.2%) were completely empty, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
+We were successfully able to identify Redside Shiners, a Largescale Sucker, sculpin, Mountain Whitefish, and one juvenile Chinook Salmon in the stomach contents.
+However, decomposition from digestion rendered most stomach contents unidentifiable.
+
+<!--
+Concerns with gastric lavage.
+We only observed juvenile Chinook salmon (and/or other fishes) in a very small number of stomach contents.
+However, we don't necessarily believe that to mean that pikeminnow in Deadwater Slough rarely consume juvenile Chinook salmon.
+Are there other cases where gastric lavage failed to document predation?
+How soon after consumption do you need to take a sample?
+So instead, we made some assumptions about the proportion of a pikeminnows diet consisting of Chinook salmon.
+-->
+
+## Bioenergetics
+
+What assumptions did we make during the bioenergetics assessment?
+And how might violations of those assumptions change our estimate of the number of juvenile Chinook salmon consumed and resulting impacts to adult returns?
 
@@
 
-<!-- After making some assumptions (e.g., on SARs) attempt to estimate how results from objective #3 could translate into impacts on adult returns (via a thought experiment or simulation approach) -->
-
-<!-- Currently only using one bioenergetics model run.
+<!-- 
+After making some assumptions (e.g., on SARs) attempt to estimate how results from objective #3 could translate into impacts on adult returns (via a thought experiment or simulation approach)
+
+Currently only using one bioenergetics model run.
 Could incorporate more bioenergetics models, but if so then maybe consider focusing on a single abundance estimator (e.g.
-Schnabel) to avoid too many combinations in the results -->
-
-# Results
-
-## Abundance of Piscivorous Fishes
-
-<!--
-* Total population estimates of Pike Minnow. Chart or table, but likely a chart.
-* What is that equate to as a density? e.g., how many pikeminnow per 10 square meters? Is that a reasonable number?
-* To calculate density, need the area of the extent surveyed each year. Was that area consistent year-to-year?
+Schnabel) to avoid too many combinations in the results
 -->
 
-Abundance estimates for northern pikeminnow during the fall sampling events ranged from `r prettyNum(min(all_est$N), big.mark = ",")` to `r prettyNum(max(all_est$N), big.mark = ",")`.
-The multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied.
-
-PSD for northern pikeminnow across all three years was 42.26% or 42.26% of fish were over the quality size classification 400 mm TL we denoted in the Methods, demonstrating a high percentage of above average size classes present within Deadwater Slough.
-
-## Predation On Juvenile Chinook Salmon
-
-During 2019, 660 Northern Pikeminnow were gastric lavaged.
-Of those fish, we observed 603 individuals with empty stomachs, and 57 individuals with stomach contents.
-Twelve of the samples contained fish remnants including a single Mountain Whitefish and two Redside Shiners.
-During the 2020 season, 805 Northern Pikeminnow were gastric lavaged comprised of 613 individuals with empty stomachs and 188 individuals with stomach contents.
-Twenty-three individuals had fish remnants in their stomach contents including ten with unidentifiable remnants, six Redside Shiners, four suckers, two sculpin, and one juvenile Chinook Salmon.
-In 2021, we collected stomach contents from 105 Northern Pikeminnow with 98 containing non-fish prey items and observed seven containing fish remnants.
-
-<!--- Need to add Nick's gastric lavage table here -->
-
-## Bioenergetics
-
-<!-- Need to add Nick's figures here -->
-
-Over the fall time period a single Northern Pikeminnow is modeled to eat 61 g of fish.
-For the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook Salmon.
-Therefore, Northern Pikeminnow in the Deadwater Slough can potentially consume ***xx,xxx*** juvenile Chinook Salmon as a discrete population.
-
-Over an entire year, we estimate that a single Northern Pikeminnow will eat 283 g of fish and 613 g of other food types.
-
-<!--
-* Water Temperature measurements.
-    *	Hobo tidbit placed at site. 
-    *	Calculated daily average water temperature for 365 days of the year.
-*	Water temperature plot for the 365 average.
-*	Bioenergetics results. Total mass consumed over a 365 day period. Plot of daily consumption with day of year on x and mass on y.
--->
-
-## Impacts to Adult Returns
+Again, what assumptions did we make here and how might violations of those assumptions change our estimate of impacts to adult returns.
 
 *	SAR results. Maybe a plot of multiple scenarios depending on gastric lavage results? And pop estimates if they vary in time.
-
-# Discussion
-
-We estimated the population size of Northern Pikeminnow in the Deadwater Slough to be greater than ***xx,xxx*** during the fall emigration period for DSR Chinook Salmon.
-That estimate translates to a density of ***xxx*** Northern Pikeminnow per 100 m or ***xxx*** per 100 m^2^ which is ***similar/more/less*** than estimates from elsewhere in the Columbia River ***(citation)*** where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminno abundance.
-The population size of Northern Pikeminnow was not directly estimated during the spring NRR Chinook salmon emigration period however, the relative abundance measured at CPUE was comparable to the fall sampling periods ***(update statement later)***.
-The population of Northern Pikeminnow in Deadwater Slough was estimated to consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook Salmon during the ***x*** sampling periods and result in an estimated reduction of returning adults between ***xxx*** and ***x,xxx***.
-We suggest that the habitat modifications that created the Deadwater Slough have resulted in favorable conditions for Northern Pikeminnow, including improved conditions for predation upon juvenile Chinook Salmon ***(add detail here)***.
-Therefore, predation by Northern Pikeminnow in the Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG.
-
-## Mark-Recapture Model
-
-<!--
-Re-hash out our assumptions?
-What assumptions were most likely violated e.g., the closed population assumption?
-If we assumed an open population with varying immigration/emigration rates how might that affect our estimate?
-In the end, how good do we think our estimate is and is the bottom-line that we belive there still to be a shit-ton of pikeminnow even if our assumptions were violated?
--->
-
-## Gastric Lavage
-
-The majority of fish collected during this study received gastric lavage, including some non-predatory species.
-To validate the efficacy of this method, we euthanized nine Northern Pikeminnow after gastric lavage was completed and removed the remaining stomach contents via dissection.
-Our results support previous findings that gastric lavage effectively removes stomach contents (@Lott2020).
-Additionally, we found that the fish captured in net-traps had a similar proportion of stomach content samples.
-of those, 1,214 (76.2%) were completely empty, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
-We were successfully able to identify Redside Shiners, a Largescale Sucker, sculpin, Mountain Whitefish, and one juvenile Chinook Salmon in the stomach contents.
-However, decomposition from digestion rendered most stomach contents unidentifiable.
-
-<!--
-Concerns with gastric lavage.
-We only observed juvenile Chinook salmon (and/or other fishes) in a very small number of stomach contents.
-However, we don't necessarily believe that to mean that pikeminnow in Deadwater Slough rarely consume juvenile Chinook salmon.
-Are there other cases where gastric lavage failed to document predation?
-How soon after consumption do you need to take a sample?
-So instead, we made some assumptions about the proportion of a pikeminnows diet consisting of Chinook salmon.
--->
-
-## Bioenergetics
-
-What assumptions did we make during the bioenergetics assessment?
-And how might violations of those assumptions change our estimate of the number of juvenile Chinook salmon consumed and resulting impacts to adult returns?
-
-## Impacts to Adult Returns
-
-Again, what assumptions did we make here and how might violations of those assumptions change our estimate of impacts to adult returns.
 

### 2021-11-24 analysis/paper/paper.Rmd
updated objectives to reflect moving 'Adult Impacts'

@@
 In this study, we estimated the abundance of a piscivorous fish predator population in the Deadwater Slough and their potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
-Our objectives for the study were four-fold:
+Our objectives for the study were three-fold:
 
@@
 3. Use a bioenergetics approach to estimate the total consumption of juvenile DSR and NRR Chinook salmon during defined emigration periods.
-4. Quantify the potential impacts of estimated predation on juvenile to Chinook Salmon adult returns.
-
-We follow with a discussion of the various assumptions that went into the mark-recapture and bioenergetics models and assessment of impacts to adult returns and how violations of some assumptions may affect overall results and inferences from the study.
+
+We follow with a discussion of the various assumptions that went into the mark-recapture and bioenergetics models and how violations of some assumptions may affect overall results and inferences from the study.
+We also consider how consumption of juvenile Chinook Salmon by predators at Deadwater Slough might impact adult returns, hindering recovery of populations in the Upper Salmon River MPG.
 
@@
 
+<!-- START HERE -->
+
 The Schnabel estimator is shown below, where the $M$, $n$ and $m$ are indexed by the sampling occasion, $i$.

### 2021-12-02 analysis/paper/paper.Rmd
updated CPUE to also include a metric of fish / angler day, and large fish (>= 228mm) / angler day, to compare with other reports.

@@
   group_by(species, event_name) %>%
-              summarise(across(c(total_effort, n_fish),
+  summarise(across(c(total_effort, 
+                     n_fish,
+                     n_large_fish,
+                     angler_days = crew_count),
                    sum),
@@
                 ~ . / 60)) %>%
-              mutate(cpue = n_fish / total_effort)
+  mutate(cpue = n_fish / total_effort,
+         cpue_per_day = n_fish / angler_days,
+         cpue_large_per_day = n_large_fish / angler_days)
 
@@
 ```{r cpue-fig, fig.cap = "Total catch per unit effort across entire sampling event."}
-cpue_df %>%
-  mutate(across(event_name,
-                ~ str_replace(., "_", " "))) %>%
-  group_by(species, event_name) %>%
-  summarise(across(c(total_effort, n_fish),
-                   sum),
-            .groups = "drop") %>%
-  mutate(across(total_effort,
-                ~ . / 60)) %>%
-  mutate(cpue = n_fish / total_effort) %>%
+total_cpue %>%
   ggplot(aes(x = event_name,
-             y = cpue,
+             y = cpue_large_per_day,
              fill = event_name)) +
@@
   labs(x = "Event",
-       y = "CPUE (Pikeminnow caught per hour)") +
+       y = "CPUE (Large Pikeminnow caught per angler day)") +
   theme(legend.position = "none")

### 2021-12-02 analysis/R scripts/abund_est.R

@@ -32,7 +32,13 @@ raw_df = readxl::read_excel(here("analysis/data/raw_data/Deadwater_predation_mar
 cmr_df = raw_df %>%
   filter(count > 0) %>%
   # pull out only Northern Pikeminnow
-  filter(species == "Northern Pikeminnow")
+  filter(species == "Northern Pikeminnow") %>%
+  mutate(mark_release = if_else(fish_status == "Live" &
+                                  capture_type == "Mark",
+                                T, F),
+         recap = if_else(capture_type == "Recapture",
+                         T, F))
+
 
 #------------------------------------------------------------
 # CPUE
@@ -54,23 +60,29 @@ effort_df = raw_df %>%
                 ~ . * crew_count)) %>%
   group_by(event_name,
            date) %>%
-  summarise(across(total_effort,
+  summarise(across(c(total_effort,
+                     crew_count),
                    sum),
             .groups = "drop")
 
-# calculate CPUE
 cpue_df = effort_df %>%
   # add how many pikeminnow were caught
   left_join(cmr_df %>%
               group_by(event_name, species, date) %>%
               summarise(n_fish = n(),
+                        n_large_fish = sum(length>=228,
+                                           na.rm = T),
                         .groups = "drop")) %>%
-  mutate(cpue = n_fish / total_effort)
+  mutate(cpue = n_fish / total_effort,
+         cpue_per_day = n_fish / crew_count,
+         cpue_large_per_day = n_large_fish / crew_count)
+
+
 
 #------------------------------------------------------------
 # Abundance
 #------------------------------------------------------------
-# using Schnabel and Schumacher-Eschmeyer models
+# using Schnabel and Schumacher-Eschmeyer estimators
 N_mods = cmr_df %>%
   # filter out spring 2021 (no recapture)
   filter(event_name != "Spring_2021") %>%
@@ -81,6 +93,9 @@ N_mods = cmr_df %>%
             R = sum(fish_status == "Live", na.rm = T),
             .groups = "drop") %>%
   group_by(event_name, species) %>%
+  mutate(M = lag(R - m),
+         M = replace_na(M, 0),
+         M = cumsum(M)) %>%
   nest() %>%
   crossing(model = c("Schnabel",
                      "SchumacherEschmeyer")) %>%
@@ -167,28 +182,28 @@ LP_mod = cmr_df %>%
   nest(data = -c(event_name, species)) %>%
   crossing(model = c("Petersen",
                      "Chapman")) %>%
[…]

### 2021-12-02 analysis/paper/paper.Rmd
knitted into a html

@@
            date) %>%
-  summarise(across(total_effort,
+  summarise(across(c(total_effort,
+                     crew_count),
                    sum),
@@
               summarise(n_fish = n(),
+                        n_large_fish = sum(length>=228,
+                                           na.rm = T),
                         .groups = "drop")) %>%
-  mutate(cpue = n_fish / total_effort)
+  # calculate CPUE a few different ways
+  mutate(cpue = n_fish / total_effort,
+         cpue_per_day = n_fish / crew_count,
+         cpue_large_per_day = n_large_fish / crew_count)
 

### 2021-12-02 analysis/paper/paper.Rmd
draft of abundance methods

@@
 
-<!-- START HERE -->
-
-The Schnabel estimator is shown below, where the $M$, $n$ and $m$ are indexed by the sampling occasion, $i$.
-In our example, the sampling occasions are defined as each day of sampling.
-This estimator does not have an associate standard error, but 95% confidence intervals can be calculated.
-The Schnabel estimator is essentially a weighted average of a series of Lincoln-Petersen estimators (with a Chapman modification).
+For a multiple census estimator we first considered the Schnabel estimator where the $M$, $n$ and $m$ are indexed by the sampling occassion, $i$.
+In our case, the sampling occassions are defined as each day of sampling.
+The Schnabel estimator does not have an associated standard error; however, the 90% confidence intervals can be calculated.
+The Schnabel estimator is essentially a weighted average of a series of Lincoln-Peterson estimators (with a Chapman modfication).
+The Schnabel estimator is shown here:
 
@@
 
-There is another estimator for this type of multiple census surveys, called the Schumacher-Eschmeyer estimator, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
-It is shown below.
+Finally, we explored one additional estimator for multiple census surveys, called the Schumacher-Eschmeyer estimator, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
+The Schumacher-Eschmeyer estimator is as follows:
 
@@
 $$
-Additionally, we calculated the proportional stock density (PSD) of Northern Pikeminnow with 300 mm total length (TL) for stock and 400 mm TL for quality.
+
+<!-- Here, let's attempt to use the ratio of CPUEs between the fall and spring to estimate abundance in the spring.
+We'll assume equal capture probability in both periods.
+We recognize that capture probability is probably smaller in the spring (higher flows, more turbidity, perhaps cooler?), so we think our estimate is a conservative one (i.e., the truth is probably higher).
+Then we'll re-do the bioenergetics model to explore how many NFF smolts might be consumed in the spring -->
+
+The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and fall 2020 sampling periods.
+However, due to logistical constraints, we were only able to sample during a single week in spring 2021 and released fish were released unmarked.
+Instead, to estimate the abundance of Northern Pikeminnow in the spring, we used the ratio of total CPUE in the fall sampling events to the total CPUE in the spring sampling event, and multiplied that ratio by the abundance of Northern Pikeminnow in the fall.
+This assumes the same capture probability in the fall as in the spring.
+If that assumption is not true and capture probability did differ, we believe that the capture probability to be lower in the spring due to higher flows, which would lead to our spring abundance estimate to be biased low.
+Therefore, we consider this a conservative estimate of the number of Northern Pikeminnow present in the Deadwater Slough during the spring emigration of NRR Chinook Salmon.
+
+Lastly, we calculated the proportional stock density (PSD) for Northern Pikeminnow captured during field efforts, which is a measure of species size structure.
+PSD is a ratio, typically expressed as a percentage, between the number of "quality-sized" individuals or larger individuals and "stock-sized" individuals.
+Stock and quality size definitions vary by species.
+PSD is expressed as:
+
 $$
@@
 $$
+
 where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
-
-<!-- use ratio of CPUEs between fall and spring, and assume same capture probability in both periods to estimate abundance in the spring.
-Recognize that capture probability is probably smaller in the spring, so our estimate is a conservative one (the truth is probably higher).
-Then re-do the bioenergetics model to look at how many smolts are consumed in the spring.
--->
-
-To estimate the abundance of Northern Pikeminnow in the spring, we used the ratio of total CPUE in a fall sampling event (total Northern Pikeminnow caught / effort) to the total CPUE in the spring, and multiply that ratio by the abundance of Northern Pikeminnow in the fall.
-This assumes the same capture probability in the fall as in the spring.
-If that does differ, we believe the capture probability to be lower in the spring due to higher flows, which would lead to our spring abundance estimate to be biased low.
-Therefore, we consider this a conservative estimate of the number of Northern Pikeminnow present in the Deadwater area during the spring Chinook smolt out migration.
+We calculated the PSD for Northern Pikeminnow in our study using 300 mm TL for stock-size and 400 mm TL for quality-size.
 
@@
 
-```{r adult-imp-p}
+```{r adult-imp-p, fig.cap = "The adult impacts plot."}
 adult_impact %>%

### 2021-12-03 analysis/paper/paper.Rmd
minor edits to results

@@
 
+Gastric lavage [@Foster1977] was used to examine the stomach contents of Northern Pikeminnow for the presence of juvenile Chinook Salmon, other fishes (e.g., juvenile steelhead, juvenile Sockeye Salmon, Redside Shiner, etc.), and non-targets (e.g., macroinvertebrates, organic matter, etc.).
+Stomach contents were preserved with 99% isopropyl alcohol in whirl-paks and analyzed in a laboratory.
+For each sample, wet weight (grams) was recorded for the combined stomach content and for the fish content.
+Fish and fish remnants were identified to the lowest taxonomic unit, when possible, or were categorized as unknown.
+A subset of Northern Pikeminnow captures were euthanized for dissection after gastric lavage to validate the efficacy of the methodology.
+
 ```{r gastric-data}
@@
 
-Gastric lavage [@Foster1977] was used to examine the stomach contents of Northern Pikeminnow for the presence of juvenile Chinook Salmon, other fishes (e.g., juvenile steelhead, juvenile Sockeye Salmon, Redside Shiner, etc.), and non-targets (e.g., macroinvertebrates, organic matter, etc.).
-Stomach contents were preserved with 99% isopropyl alcohol in whirl-paks and analyzed in a laboratory.
-For each sample, wet weight (grams) was recorded for the combined stomach content and for the fish content.
-Fish and fish remnants were identified to the lowest taxonomic unit, when possible, or were categorized as unknown.
-A subset of Northern Pikeminnow captures were euthanized for dissection after gastric lavage to validate the efficacy of the methodology.
-
 ## Bioenergetics
+
+```{r shoup-temp-data}
+shoup_temp_df = read_tsv(here("analysis/data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
+                   comment = "#") %>%
+  rename(discharge_cfs = `47768_00060`,
+         discharge_cfs_cd = `47768_00060_cd`,
+         gage_ht_ft = `47769_00065`,
+         gage_ht_ft_cd = `47769_00065_cd`,
+         water_temp_c = `47773_00010`,
+         water_temp_c_cd = `47773_00010_cd`)
+
+```
+
+```{r read-bio-outputs}
+# full year, no growth
+full_yr_no_growth = read_csv(here("analysis/data/raw_data/2020 Bioenergetics 1 year.csv")) %>%
+  clean_names() %>%
+  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
+  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
+  mutate(month = month(date))
+
+# full year, 10% growth
+full_yr_10p_growth = read_csv(here("analysis/data/raw_data/2020 Bioenergetics 1 year 10% growth.csv")) %>%
+  clean_names() %>%
+  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
+  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
+  mutate(month = month(date))
+
+# an alternative model
+alternative_mod = read_csv(here("analysis/data/raw_data/Bioenergetics 32 fish 68 inverts.csv")) %>%
+  clean_names() %>%
+  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
+  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
+  mutate(month = month(date))
+
+# Fall emigration period (9/1 - 11/17), no growth
+fall_only_no_growth = read_csv(here("analysis/data/raw_data/78_day_bioenergetics.csv")) %>%
+  clean_names() %>%
+  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
+  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
+  mutate(month = month(date))
+
+```
+
+To estimate the consumption of juvenile Chinook Salmon outmigrants by Northern Pikeminnow in the Deadwater Slough, we used the Fish Bioenergetics v4.0 applicatioon developed by @Deslauriers2017.
+The daily rate of consumption in grams for an individual Northern Pikeminnow was based on predator and prey energy densities, predator start and end weights, and water temperatures.
+Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017].
+Prey energy densities were fixed at 21,500 J/g for juvenile Chinook salmon based on estimate from @Moss2016.
+To calculate the average predator start weight, we converted the average length of Northern Pikeminnow captured during our study to a weight using a weight-length formula for Northern Pikeminnow [@Parker1995] from the FSA package [@Ogle2021].
+Continuous water temperature data from 2019 were used in two alternative models, on using the baseline start and end weights and the second with a 10% increase in average end weight of Northern Pikeminnow.
+
+We ran three bioenergetic models for this study.
+For the first model, we chose a 78-day period from September 1 through November 17 when DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration.
+During this time, water temperatures exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon.
+This model assumed no growth for Northern Pikeminnow.
+The second model was run for one-year and assumed no growth for Northern Pikeminnow.
+The assumption of no growth for pikeminnow was included to show what may be occurring if the population of Northern Pikeminnow is stable.
+If growth of the population is occurring, the alternative model shows how an increase of 10% body weight changes the estimates of consumption over a full year period.
+To estimate the number of possible Chinook Salmon that could be consumed, we assumed a weight of 12 g per individual Chinook Salmon.
+This weight was an average from juvenile Chinook Salmon emigrating past six rotary screw traps operating in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.
 
@@
 
-```{r shoup-temp-data}
-shoup_temp_df = read_tsv(here("analysis/data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
-                   comment = "#") %>%
-  rename(discharge_cfs = `47768_00060`,
-         discharge_cfs_cd = `47768_00060_cd`,
-         gage_ht_ft = `47769_00065`,
-         gage_ht_ft_cd = `47769_00065_cd`,
-         water_temp_c = `47773_00010`,
-         water_temp_c_cd = `47773_00010_cd`)
-
-```
-
-```{r read-bio-outputs}
-# full year, no growth
-full_yr_no_growth = read_csv(here("analysis/data/raw_data/2020 Bioenergetics 1 year.csv")) %>%
-  clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
-  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
-  mutate(month = month(date))
-
-# full year, 10% growth
-full_yr_10p_growth = read_csv(here("analysis/data/raw_data/2020 Bioenergetics 1 year 10% growth.csv")) %>%
-  clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
-  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
-  mutate(month = month(date))
-
-# an alternative model
-alternative_mod = read_csv(here("analysis/data/raw_data/Bioenergetics 32 fish 68 inverts.csv")) %>%
-  clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
-  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
-  mutate(month = month(date))
-
-# Fall emigration period (9/1 - 11/17), no growth
-fall_only_no_growth = read_csv(here("analysis/data/raw_data/78_day_bioenergetics.csv")) %>%
-  clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
-  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
-  mutate(month = month(date))
-
-```
-
-To estimate the consumption of juvenile Chinook Salmon outmigrants by Northern Pikeminnow in the Deadwater Slough, we used the Fish Bioenergetics v4.0 applicatioon developed by @Deslauriers2017.
-The daily rate of consumption in grams for an individual Northern Pikeminnow was based on predator and prey energy densities, predator start and end weights, and water temperatures.
-Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017].
-Prey energy densities were fixed at 21,500 J/g for juvenile Chinook salmon based on estimate from @Moss2016.
-To calculate the average predator start weight, we converted the average length of Northern Pikeminnow captured during our study to a weight using a weight-length formula for Northern Pikeminnow [@Parker1995] from the FSA package [@Ogle2021].
-Continuous water temperature data from 2019 were used in two alternative models, on using the baseline start and end weights and the second with a 10% increase in average end weight of Northern Pikeminnow.
-
-We ran three bioenergetic models for this study.
-For the first model, we chose a 78-day period from September 1 through November 17 when DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration.
-During this time, water temperatures exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon.
-This model assumed no growth for Northern Pikeminnow.
-The second model was run for one-year and assumed no growth for Northern Pikeminnow.
-The assumption of no growth for pikeminnow was included to show what may be occurring if the population of Northern Pikeminnow is stable.
-If growth of the population is occurring, the alternative model shows how an increase of 10% body weight changes the estimates of consumption over a full year period.
-To estimate the number of possible Chinook Salmon that could be consumed, we assumed a weight of 12 g per individual Chinook Salmon.
-This weight was an average from juvenile Chinook Salmon emigrating past six rotary screw traps operating in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.
-
-<!--
-To estimate the impacts from northern pikeminnow on juvenile Chinook Salmon outmigrants in Deadwater Slough, we used an R-based application of Fish Bioenergetics v4.0 developed by @Deslauriers2017.
-We estimated a daily rate of consumption for an individual northern pikeminnow based on predator and prey energy densities, water temperatures, and predator start and end weights.
-We also ran two alternative models one with a year of temperature inputs from 2019, and another with the same temperature inputs but a 10% growth in weight of the average pikeminnow at the end of the model.
-The average length of northern pikeminnow captured at the Deadwater Slough during our study was 370.6 mm which calculates to an average starting weight of 847 g using the FSA package (@Ogle2021) in R which provides a weight-length formula for northern pikeminnow sourced from @Parker1995.
-Predator energy density for northern pikeminnow was available in the Bioenergetics v4.0 application and was fixed at 6,703 Joules(J)/g.
-Prey energy densities was taken from @Moss2016 where they estimated juvenile Chinook Salmon at 21,500 J/g.
-Using these temperatures, weights, and energy inputs, we estimated a total consumption in grams for an individual norther pikeminnow within Deadwater Slough.
-
-We ran three bioenergetic models for this study.
-First, we chose a 78-day period from September 1 through November 17 where we know Chinook Salmon DSR emigrants begin to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin to migrate downstream, but when temperatures are still high enough that juveniles are not yet exhibiting concealment behavior or torpor.
-The second model was still assuming no growth but taking the entire year into account for the model rather than the 78-day fall period.
-The assumption of no growth for pikeminnow was included to show what may be occurring if the population of northern pikeminnow is stable.
-If growther of the population is occurring, the alternative model shows how an increase of 10% body weight changes the estimates of consumption over a full year period.
-To estimate the number of possible Chinook Salmon that could be consumed, we assumed a weight of 12 g per individual chinook Salmon.
-This weight was an average from juvenile Chinook salmon emigrating past six rotary screw traps operating in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.
+# Results
+
+## Abundance of Piscivorous Fishes
+
+Abundance estimates for Northern Pikeminnow in the Deadwater Slough during the fall 2019 and 2020 sampling events ranges from `r prettyNum(min(all_est$N), big.mark = ",")` to `r prettyNum(max(all_est$N), big.mark = ",")`.
+The multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied (Figure \@ref(fig:abund-fig)).
+<!-- Need to fill in these fall results more -->
+
+<!-- Add spring abundance results here -->
+
+The PSD for Northern Pikeminnow in our study and across all three years was 42.3%.
+In other words, 42.3% of captured fish were over the quality size classification of 400 mm TL, demonstrating that a high percentage of Northern Pikeminnow within Deadwater Slough are above average size classes defined for the species <!-- Is this useful?
+And if useful, does it make sense to calculate PSD separately for the fall and spring periods?
+I'd expect PSD to vary between seasons.
 -->
 
-# Results
-
-## Abundance of Piscivorous Fishes
-
-<!--
-* Total population estimates of Pike Minnow. Chart or table, but likely a chart.
-* What is that equate to as a density? e.g., how many pikeminnow per 10 square meters? Is that a reasonable number?
-* To calculate density, need the area of the extent surveyed each year. Was that area consistent year-to-year?
--->
-
-Abundance estimates for northern pikeminnow during the fall sampling events ranged from `r prettyNum(min(all_est$N), big.mark = ",")` to `r prettyNum(max(all_est$N), big.mark = ",")`.
-The multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied.
-
-PSD for northern pikeminnow across all three years was 42.26% or 42.26% of fish were over the quality size classification 400 mm TL we denoted in the Methods, demonstrating a high percentage of above average size classes present within Deadwater Slough.
+<!-- Include some results for density e.g., pikeminnow per 100 linear meters or pikeminnow per acre. -->
 

### 2021-12-03 analysis/paper/paper.Rmd
added psd calculation

@@
 
+```{r psd}
+s_size = 300
+q_size = 400
+
+psd_df = cmr_df %>%
+  select(species, count, length, recap, event_name) %>%
+  filter(recap  != TRUE)
+
+fs = psd_df %>%
+  filter(length >= s_size) %>%
+  nrow()
+
+fq = psd_df %>%
+  filter(length >= q_size) %>%
+  nrow()
+
+psd = 100 * (fq/fs)
+
+```
+
 Lastly, we calculated the proportional stock density (PSD) for Northern Pikeminnow captured during field efforts, which is a measure of species size structure.
@@
 where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
-We calculated the PSD for Northern Pikeminnow in our study using 300 mm TL for stock-size and 400 mm TL for quality-size.
+We calculated the PSD for Northern Pikeminnow in our study using `r prettyNum(s_size, digits = 0)` mm TL for stock-size and `r prettyNum(q_size, digits = 0)` mm TL for quality-size.
 
@@
 
-The PSD for Northern Pikeminnow in our study and across all three years was 42.3%.
-In other words, 42.3% of captured fish were over the quality size classification of 400 mm TL, demonstrating that a high percentage of Northern Pikeminnow within Deadwater Slough are above average size classes defined for the species <!-- Is this useful?
+The PSD for Northern Pikeminnow in our study and across all three years was `r prettyNum(psd, digits = 1)`%.
+In other words, `r prettyNum(psd, digits = 1)`% of captured fish were over the quality size classification of `r prettyNum(q_size, digits = 0)` mm TL, demonstrating that a high percentage of Northern Pikeminnow within Deadwater Slough are above average size classes defined for the species <!-- Is this useful?
 And if useful, does it make sense to calculate PSD separately for the fall and spring periods?

### 2021-12-03 analysis/paper/paper.Rmd
done for day

@@
 
-<!-- Here, let's attempt to use the ratio of CPUEs between the fall and spring to estimate abundance in the spring.
-We'll assume equal capture probability in both periods.
-We recognize that capture probability is probably smaller in the spring (higher flows, more turbidity, perhaps cooler?), so we think our estimate is a conservative one (i.e., the truth is probably higher).
-Then we'll re-do the bioenergetics model to explore how many NFF smolts might be consumed in the spring -->
-
 The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and fall 2020 sampling periods.

### 2021-12-07 analysis/paper/paper.Rmd
incorporated RC edits

@@
 
-The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017]; all of which have become depleted in recent decades.
-Declines in survival of local juvenile Chinook Salmon have been associated with hydrosystem passage [@Schaller2014], ocean conditions and regional climate [@Welch2021; @Gosselin2021], and changes to the freshwater habitat including removal of beavers, river channel simplification, water withdrawals, and the proliferation of non-native species (e.g., non-native coastal rainbow trout *O.
-mykiss irideus* and brook trout *Salvelinus fontinalis*) among others [@IdahoOSCTeam2019].
-Moreoever, the abundance of returning adults are further impacted by ocean and downriver harvests, poor ocean conditions, and changes to the spawning migration corridor.
-Each of these factors have contributed, to varying and unknown extents, to reduced adult escapement, the primary metric used to assess population viability.
-In response to the decline in Chinook Salmon abundance from the myriad human activities and associated habitat degradation, action agencies have attempted to improve juvenile survival and adult spawning conditions by investing in the rehabilitation of tributary ecosystems.
-
-One potentially important, but perhaps under-appreciated source of mortality for Chinook Salmon is predation on emigrating juveniles by piscivorous fishes, including both native and non-native species.
+The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* each providing important ecological, social, and economic benefits to the region.
+The distinct populations include Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017]; all of which have become depleted in recent decades.
+Reduction in overall population size, and more specifically, declines in survival of juvenile Chinook Salmon have been associated with hydrosystem passage [@Schaller2014], ocean conditions, regional climate [@Welch2021; @Gosselin2021], and downstream predation by native and non-native piscivorous fish [@Poe1991; @Friesen1999; @Harvey2004].
+Anthropogenic changes to freshwater habitat including removal of beavers, river channel simplification, water withdrawals, and the proliferation of non-native species (e.g., non-native coastal rainbow trout *O.
+mykiss irideus* and brook trout *Salvelinus fontinalis*) among others [@IdahoOSCTeam2019] have also been attributed to the reductions in the population.
+Moreover, the abundance of returning adults is futher impacted by ocean and mainstem harvests, poor ocean conditions, and changes to the spawning migration corridor [@FCRPSBiOp2020].
+Each of these factors have contributed, to varying extents, to reduced adult escapement, the primary metric used to assess population viability.
+In response to the decline in Upper Salmon Chinook Salmon abundance from the myriad of human activities and associated habitat degradation, action agencies have attempted to improve juvenile survival and adult spawning conditions by investing in rehabilitation of tributary ecosystems as well as investing in the study and evaluation of many factors attributed to the species decline [@Roni].
+
+One potentially important, but perhaps under-appreciated source of mortality for Upper Salmon Chinook Salmon is predation on emigrating juveniles by piscine fishes.
+Research in the lower mainstem Snake and Columbia rivers has estimated that during peak emigration, predation on out-migrating salmonids has a significant negative impact (most notably in the presence on non-native piscivores) on the overall population and success of population recovery [@Fresh2003; @Winther2019].
+However, these studies have focused on larger, mainstem and reservoir habitats and direct predation, but very little is known about the interaction between upstream, river dwelling piscine fishes, and their impact on salmonid recovery above reservoir systems of the Columbia and Snake Rivers [@Rubenson2020].
+Adding to this, there is little consensus within the literature related to the magnitude of predation on salmonids in less altered, free flowing systems, with some suggesting lower levels of predation [@Brown1981; @Buchana1981] and others suggesting increased levels of predation can occur [@Tabor1993; @Ward1995; @Shively1996; @Zimmerman1999].
+
 Dams and reservoirs in the Columbia River basin are the primary locations associated with high rates of predation on salmonids [@Petersen1994; @Ward1995].
@@
 First, migration rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation.
-Second, reservoirs and downstream tailraces created by dams on the Columbia and Snake rivers have created favorable slow-water habitat for several species of piscivorous fish known to consume juvenile Chinook Salmon including both non-native (e.g., Smallmouth Bass *Micropterus dolomieu*, Walleye *Sander vitreus*) and native (e.g., Northern Pikeminnow *Ptychocheilus oregonensis*) species.
+Second, reservoirs and downstream tailraces resulting from dams on the Columbia and Snake rivers have created favorable slow-water habitat for several species of piscivorous fish known to consume juvenile Chinook Salmon including both non-native (e.g., Smallmouth Bass *Micropterus dolomieu*, Walleye *Sander vitreus*) and native (e.g., Northern Pikeminnow *Ptychocheilus oregonensis*) species.
 In the Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of non-native Smallmouth Bass during some seasons [@Erhardt2018].
 And non-native Walleye are considered a potentially important source for juvenile Chinook Salmon mortality in the Snake River[@Widener2021], particularly as their abundance has significantly increased surrounding Snake River dam facilities between 2008 and 2020 related to their preference for slow water habitats.
-
-In addition to non-native species, Northern Pikeminnow are a native, piscivorous fish preferring warm, slow water habitats.
+Much effort has been devoted to understanding predation related mortality throughout the hydropower system of the Lower Snake and Columbia rivers, but little is known about native and non-native predation on salmonids throughout free-flowing sections of the Columbia Basin such as the Salmon River [@Rubenson2020].
+
+Along with non-native species, the Salmon River supports native Northern Pikeminnow, a piscine fish tolerating and thriving in relatively warm, slow water habitats.
 Consequently, they too have benefited from dams on the Columbia River and become abundant predators on salmonid outmigrants [@Knutsen1999].
@@
 Predation by piscivorous fishes, including Northern Pikeminnow, may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@Widener2021].
-However, there is a dearth of current data on piscivorous predator populations in the Columbia River.
-
-Habitats containing slower water velocities and other attributes that support piscivorous predators are not limited to dams and reservoirs.
-One such area is the Deadwater Slough in the Salmon River, Idaho which is an unnaturally slow and deep reach of the mainstem river approximately 1.5 kilometers in length and occupying 29.5 acres (Figure \@ref(fig:map)).
-The slough contains "reservoir-like" conditions which typically favor piscivorous fish predators like Northern Pikeminnow or Smallmouth Bass [@Watkins2015], and further, lacks hydrological and structural features and fish cover (e.g., large wood or substrate, woody debris, bank undercuts) that can provide essential refuge from predation for juvenile fishes.
+However, there is a dearth of current data on piscivorous predator populations in the Columbia River with little known about their movement, life history, or impacts upstream of the mainstem habitats [@Gadomski2001].
+
+Habitats containing slower water velocities and other attributes that support piscivorous predators are not limited to areas near dams and reservoirs.
+One such example is the Deadwater Slough on the mainstem Salmon River, Idaho which is an unnaturally slow and deep reach approximately 1.5 kilometers in length and occupying 29.5 acres (Figure \@ref(fig:map)).
+The slough is located approximately 6 kms downstream from the town of North Fork, Idaho and contains anthropogenically created "reservoir-like" conditions which typically favor piscivorous fish predators like Northern Pikeminnow or Smallmouth Bass [@Watkins2015], and further, lacks hydrological and structural features with little to no fish cover (e.g., large wood or substrate, woody debris, bank undercuts) that can provide essential refuge from predation for juvenile fishes.
+The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
+Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River, forming an alluvial fan and creating an unnaturally slow and deep section of the river resembling a small reservoir.
 As a result, predation of juvenile Chinook Salmon and other local ESA-listed salmonids (Sockeye Salmon *O.
 nerka*, redband steelhead *O.
-mykiss gairdneri*) has been cited as a concern for populations in the Upper Salmon River
-
+mykiss gairdneri*) has been cited as a concern for populations in the Upper Salmon River.
+
+The Deadwater Slough is in a reach of the Salmon River believed to be a historically important overwinter rearing and spring emigration area for Chinook Salmon juveniles.
 Spring/summer Chinook Salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014].
@@
 Diversity of migratory tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency.
-
-The Deadwater Slough is in a reach of the Salmon River believed to be a historically important overwinter rearing area for DSR emigrants.
-Moreover, the slough is part of the migratory pathway for not only juvenile Chinook Salmon (both DSR and NRR emigrants), but also for the endangered Snake River Sockeye Salmon population and several population of threatened Snake River steelhead.
+Adding to its importance, the slough is part of the migratory pathway for not only juvenile Chinook Salmon (both DSR and NRR emigrants), but also for the endangered Snake River Sockeye Salmon population and several population of threatened Snake River steelhead.
 As such, the main Salmon River has been the target of recent radio telemetry studies to examine the downstream movement, distribution, and (apparent survival) of juvenile salmonid emigrants.
@@
 We hypothesize that increased densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon [@Ackerman2018] and Sockeye Salmon [@Axel2015].
-In this study, we estimated the abundance of a piscivorous fish predator population in the Deadwater Slough and their potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
+To test this, we estimated the abundance of a piscivorous fish predator population (Northern Pikeminnow) in the Deadwater Slough and their potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
 Our objectives for the study were three-fold:
@@
 2. Document predation on juvenile Chinook Salmon during the emigration periods using gastric lavage.
-3. Use a bioenergetics approach to estimate the total consumption of juvenile DSR and NRR Chinook salmon during defined emigration periods.
-
+3.
+Use a bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population at Deadwater Slough and the percent of consumption consisting of juvenile DSR and NRR Chinook salmon during defined emigration periods.
+
+This research is an essential first step in characterizing conditions through the Deadwater Reach and providing data necessary to discuss potential mitigation strategies (if needed), where Northern Pikeminnow removal has proven fruitful in increasing survival through the reservoir systems [@Winther2019].
 We follow with a discussion of the various assumptions that went into the mark-recapture and bioenergetics models and how violations of some assumptions may affect overall results and inferences from the study.
@@
 # Methods
-
-## Study Site
-
-The Deadwater Slough is located approximately 6 kms downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
-The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
-Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River, forming an alluvial fan and creating an unnaturally slow and deep section of the river resembling a small reservoir.
 
@@
 However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 sampling event was delayed until fall 2020.
-Although this provided the benefit of an additional abundance estimates during the DSR emigration, estimates during the NRR emigration were still lacking, and so an additional sampling event was added for spring 2021.
-The study design was such that both the fall 2019 and fall 2020 sampling efforts included both a "mark week" and "recapture week" intended for mark-recapture models.
-However, due to budget constraints the spring 2021 effort only included a single "mark week"; instead, we calculated the catch-per-unit-effort (CPUE; fish per hour) for the spring 2021 effort for comparison to the fall efforts.
+Although this provided the benefit of an additional abundance estimate during the DSR emigration, estimates during the NRR emigration were still lacking, and so an additional sampling event was added for spring 2021.
+The study design was such that the fall 2019 and fall 2020 sampling efforts included both a "mark week" and "recapture week" intended for mark-recapture models.
+However, due to budget constraints the spring 2021 effort only included a single "mark week"; instead, we calculated the catch-per-unit-effort (CPUE; fish per hour) for the spring 2021 sampling event for comparison to the fall efforts.
 In the end, sampling dates were as follows: fall 2019 (mark = November 12-14; recapture = November 19-21), fall 2020 (mark = October 20-23; recapture = October 27-30); spring 2021 (mark = May 18-21).
@@
 
-We attempted multiple capture methods during the first year including raft electrofishing, fyke netting, snorkeling, and angling to reduce selectivity, both for species and size classes.
+We attempted a multiple of capture methods during the first year including raft electrofishing, fyke netting, snorkeling, and angling to reduce selectivity, both for species and size classes.
 Unfortunately, some of the methods proved ineffective.
-For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) we lacked the appropriate motor to effectively “chase” fish given the lack of flow.
-Snorkeling was ineffective due to the high turbidity of the mainstem Salmon River and Deadwater Slough.
+For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) our inability to "chase" fish downstream as they evaded the electricity.
+Snorkeling was ineffective due to the high turbidity of the mainstem Salmon River and Deadwater Slough area.
 And due to debris floating in the water column, there were concerns with fyke nets being pulled from their anchored locations and becoming entangled.
 Additionally, there were concerns with raft electrofishers and fyke nets incidentally capturing adult steelhead that are present in the mainstem Salmon River during fall and spring.
-In the end, angling proved to be the most effective method for capturing piscivorous predators while minimizing potential impacts to steelhead but recognizing that angling can have both species and size selectivity (bias).
-From here on, our focus was on fish captured by angling, unless otherwise noted.
+Eventually, angling proved to be the most effective method for capturing piscivorous predators while minimizing potential impacts to steelhead but recognizing that angling can have both species and size selectivity (bias).
+From here on, we focus on fish captured by angling, unless otherwise noted.
 
@@
 For each fish, processors would record the date of capture, species, total length (TL; mm), and whether the fish was previously marked.
-If not previously marked, each fish was provided a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) and marks were unique for each marking day.
+If not previously marked, each fish was provided a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) unique for each marking day.
 We also recorded whether marked individuals were released back to be captured again.
@@
 
-To estimate the consumption of juvenile Chinook Salmon outmigrants by Northern Pikeminnow in the Deadwater Slough, we used the Fish Bioenergetics v4.0 applicatioon developed by @Deslauriers2017.
+To estimate the consumption of juvenile Chinook Salmon outmigrants by Northern Pikeminnow in the Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 and applied in the R statistical software [@RcitationFromNick].
 The daily rate of consumption in grams for an individual Northern Pikeminnow was based on predator and prey energy densities, predator start and end weights, and water temperatures.
 Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017].
-Prey energy densities were fixed at 21,500 J/g for juvenile Chinook salmon based on estimate from @Moss2016.
-To calculate the average predator start weight, we converted the average length of Northern Pikeminnow captured during our study to a weight using a weight-length formula for Northern Pikeminnow [@Parker1995] from the FSA package [@Ogle2021].
-Continuous water temperature data from 2019 were used in two alternative models, on using the baseline start and end weights and the second with a 10% increase in average end weight of Northern Pikeminnow.
+Prey energy densities were fixed at 21,500 J/g for juvenile Chinook salmon based on an estimate from @Moss2016.
+We used a weight-length formula [@Parker1995] to calculate the average predator start weight, converted from the average length of Northern Pikeminnow captured during our study within the FSA package [@Ogle2021].
+Continuous water temperature data from 2019 was used in two alternative models, one using the baseline start and end weights and the second with a 10% increase in average end weight of Northern Pikeminnow.
 
 We ran three bioenergetic models for this study.
-For the first model, we chose a 78-day period from September 1 through November 17 when DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration.
+For the first model, we chose a 78-day period from September 1 through November 17 where DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration.
 During this time, water temperatures exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon.
 This model assumed no growth for Northern Pikeminnow.
-The second model was run for one-year and assumed no growth for Northern Pikeminnow.
+The second model was run for one full year and assumed no growth for Northern Pikeminnow.
 The assumption of no growth for pikeminnow was included to show what may be occurring if the population of Northern Pikeminnow is stable.
@@
 
-Over the fall time period a single Northern Pikeminnow is modeled to eat 61 g of fish.
+Over the fall time period a single average-sized Northern Pikeminnow is modeled to eat 61 g of fish.
 For the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook Salmon.

### 2021-12-07 analysis/paper/paper.Rmd
updated KS address

@@
       email: tara.blackman@mthoodenvironmental.com
-      institute: [mhe]
+      institute: [mhe_sandy]
       correspondence: false
@@
       email: mike.ackerman@mthoodenvironmental.com
-      institute: [biomark, mhe]
+      institute: [biomark, mhe_mccall]
       correspondence: false
@@
   - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
-  - wdfw: Washington Department of Fish and Wildife, Under A Bridge, Seattle, Washington, 00000, USA
-  - mhe: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
+  - wdfw: Washington Department of Fish and Wildife, Fish Program, Science Division, 1111 Washington Street NE, Olympia, Washington, 98501, USA
+  - mhe_mccall: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
+  - mhe_sandy: Mount Hood Environmental, 39085 Pioneer Boulevard \#100 Mezzanine, Sandy, Oregon, 97055, USA 
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* each providing important ecological, social, and economic benefits to the region.
-The distinct populations include Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017]; all of which have become depleted in recent decades.
+The distinct populations include Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017], all of which have become depleted in recent decades.
 Reduction in overall population size, and more specifically, declines in survival of juvenile Chinook Salmon have been associated with hydrosystem passage [@Schaller2014], ocean conditions, regional climate [@Welch2021; @Gosselin2021], and downstream predation by native and non-native piscivorous fish [@Poe1991; @Friesen1999; @Harvey2004].
@@
 Each of these factors have contributed, to varying extents, to reduced adult escapement, the primary metric used to assess population viability.
-In response to the decline in Upper Salmon Chinook Salmon abundance from the myriad of human activities and associated habitat degradation, action agencies have attempted to improve juvenile survival and adult spawning conditions by investing in rehabilitation of tributary ecosystems as well as investing in the study and evaluation of many factors attributed to the species decline [@Roni].
+In response to the decline in Upper Salmon Chinook Salmon abundance from the myriad of human activities and associated habitat degradation, action agencies have attempted to improve juvenile survival and adult spawning conditions by investing in rehabilitation of tributary ecosystems as well as investing in the study and evaluation of many factors attributed to the species decline [@Roni2018].
 
@@
 However, these studies have focused on larger, mainstem and reservoir habitats and direct predation, but very little is known about the interaction between upstream, river dwelling piscine fishes, and their impact on salmonid recovery above reservoir systems of the Columbia and Snake Rivers [@Rubenson2020].
-Adding to this, there is little consensus within the literature related to the magnitude of predation on salmonids in less altered, free flowing systems, with some suggesting lower levels of predation [@Brown1981; @Buchana1981] and others suggesting increased levels of predation can occur [@Tabor1993; @Ward1995; @Shively1996; @Zimmerman1999].
+Adding to this, there is little consensus within the literature related to the magnitude of predation on salmonids in less altered, free flowing systems, with some suggesting lower levels of predation [@Brown1981; @Buchanan1981] and others suggesting increased levels of predation can occur [@Tabor1993; @Ward1995; @Shively1996; @Zimmerman1999].
 

### 2021-12-07 analysis/paper/paper.Rmd
first draft methods complete

@@
 
-Our goal was to estimate the abundance of potentially piscivorous predators in the Deadwater Slough using a mark-recapture approach both during the fall and spring near the peaks of the juvenile Chinook Salmon DSR and NRR emigrations, respectively.
-The initial intent was to sample during fall 2019 and spring 2020 to estimate predator abundances during the emigration of the 2018 brood year.
+The first objective was to estimate the population size of piscine predators in the Deadwater Slough to begin to their potential impacts on juvenile Chinook Salmon.
+Our initial intent was to estimate predator abundances using mark-recapture methods during fall 2019 and spring 2020 during the emigration of the 2018 brood year.
+We attempted to sample near the peaks of the juvenile DSR (fall) and NRR (spring) emigrations.
 However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 sampling event was delayed until fall 2020.
@@
 We attempted a multiple of capture methods during the first year including raft electrofishing, fyke netting, snorkeling, and angling to reduce selectivity, both for species and size classes.
-Unfortunately, some of the methods proved ineffective.
+Unfortunately, some methods proved ineffective during initial efforts.
 For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) our inability to "chase" fish downstream as they evaded the electricity.
@@
 And due to debris floating in the water column, there were concerns with fyke nets being pulled from their anchored locations and becoming entangled.
-Additionally, there were concerns with raft electrofishers and fyke nets incidentally capturing adult steelhead that are present in the mainstem Salmon River during fall and spring.
-Eventually, angling proved to be the most effective method for capturing piscivorous predators while minimizing potential impacts to steelhead but recognizing that angling can have both species and size selectivity (bias).
+Additionally, there were concerns with raft electrofishers and fyke nets incidentally capturing ESA-listed adult steelhead that are present in the mainstem Salmon River during fall and spring.
+Eventually, angling proved to be the most effective method for capturing piscine predators while minimizing potential impacts to steelhead but recognizing that angling can have both species and size selectivity (bias).
 From here on, we focus on fish captured by angling, unless otherwise noted.
@@
 For release, processors attempted to distribute released individuals throughout Deadwater Slough to facilitate mixing back into the population.
-We also recorded the crew (person or combination of persons) that captured each fish and the start and end times for each crew were recorded to estimate total effort.
+We also recorded the crew (person or combination of persons) that captured each fish and the start and end times for each crew to estimate total effort.
 
@@
 
-1. The population in closed (no immigration, emigration, births or deaths during the sampling period)
+1. The population is closed (no immigration, emigration, births or deaths during the sampling period)
 1. Marking a fish does not affect its chances of being caught
@@
 
-The estimators fell into two broad categories: single census and multiple census.
+Estimators fell into two broad categories: single census and multiple census.
 For the single census estimators, we treated the first week of sampling as the mark event, and the following (second) week as the recapture event, pooling data within each of those two weeks.
@@
 The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and fall 2020 sampling periods.
-However, due to logistical constraints, we were only able to sample during a single week in spring 2021 and released fish were released unmarked.
-Instead, to estimate the abundance of Northern Pikeminnow in the spring, we used the ratio of total CPUE in the fall sampling events to the total CPUE in the spring sampling event, and multiplied that ratio by the abundance of Northern Pikeminnow in the fall.
-This assumes the same capture probability in the fall as in the spring.
-If that assumption is not true and capture probability did differ, we believe that the capture probability to be lower in the spring due to higher flows, which would lead to our spring abundance estimate to be biased low.
-Therefore, we consider this a conservative estimate of the number of Northern Pikeminnow present in the Deadwater Slough during the spring emigration of NRR Chinook Salmon.
+However, due to logistical constraints, we were only able to sample during a single week in spring 2021 and released fish were left unmarked.
+Instead, to estimate the abundance of Northern Pikeminnow during spring 2021, we used the ratio of total CPUE in the fall sampling events to the total CPUE in the spring sampling event, and multiplied that ratio by the average abundance of Northern Pikeminnow from the fall sampling events.<!-- Kevin, is this previous statement accurate?
+--> This approach assumes equal capture probabilities between the fall and spring capture events.
+If that assumption is not true and capture probability did differ, we believe that the capture probability  in the spring would be lower due to higher flows, which would lead to our spring abundance estimate to be biased low.
+Therefore, we presume our spring abundance estimate during the NRR Chinook Salmon emigration to be a conservative estimate of the number of Northern Pikeminnow present in the Deadwater Slough.
 
@@
 
-Lastly, we calculated the proportional stock density (PSD) for Northern Pikeminnow captured during field efforts, which is a measure of species size structure.
+Lastly, we calculated the proportional stock density (PSD) for Northern Pikeminnow captured during all field efforts, which is a measure of species size structure.
 PSD is a ratio, typically expressed as a percentage, between the number of "quality-sized" individuals or larger individuals and "stock-sized" individuals.
@@
 
-Gastric lavage [@Foster1977] was used to examine the stomach contents of Northern Pikeminnow for the presence of juvenile Chinook Salmon, other fishes (e.g., juvenile steelhead, juvenile Sockeye Salmon, Redside Shiner, etc.), and non-targets (e.g., macroinvertebrates, organic matter, etc.).
-Stomach contents were preserved with 99% isopropyl alcohol in whirl-paks and analyzed in a laboratory.
-For each sample, wet weight (grams) was recorded for the combined stomach content and for the fish content.
+Gastric lavage [@Foster1977] was used to examine the stomach contents of Northern Pikeminnow for the presence of juvenile Chinook Salmon, other fishes (e.g., juvenile steelhead, juvenile Sockeye Salmon, Redside Shiner *Richardsonius balteatus*, etc.), and non-targets (e.g., macroinvertebrates, organic matter, etc.).
+Stomach contents of individuals were preserved with 99% isopropyl alcohol in whirl-paks immediately following lavage for later analysis in a controlled environment.
+For each sample, wet weight (grams) was recorded for the total combined stomach content as well as the contents consisting of fish parts/matter.
 Fish and fish remnants were identified to the lowest taxonomic unit, when possible, or were categorized as unknown.
@@
 
-To estimate the consumption of juvenile Chinook Salmon outmigrants by Northern Pikeminnow in the Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 and applied in the R statistical software [@RcitationFromNick].
-The daily rate of consumption in grams for an individual Northern Pikeminnow was based on predator and prey energy densities, predator start and end weights, and water temperatures.
+Gastric lavage was used in an attempt to document predation of juvenile Chinook Salmon by Northern Pikeminnow.
+However, we were ultimately interested in the potential total number or amount of juveniles consumed during the peaks of DSR (fall) and NRR (spring) emigrations.
+To estimate the consumption of juvenile Chinook Salmon in the Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 and applied in the R statistical software [@RCoreTeam2021].
+The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on predator and prey energy densities, predator start and end weights, and water temperatures.
 Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017].
-Prey energy densities were fixed at 21,500 J/g for juvenile Chinook salmon based on an estimate from @Moss2016.
-We used a weight-length formula [@Parker1995] to calculate the average predator start weight, converted from the average length of Northern Pikeminnow captured during our study within the FSA package [@Ogle2021].
-Continuous water temperature data from 2019 was used in two alternative models, one using the baseline start and end weights and the second with a 10% increase in average end weight of Northern Pikeminnow.
-
-We ran three bioenergetic models for this study.
-For the first model, we chose a 78-day period from September 1 through November 17 where DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration.
-During this time, water temperatures exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon.
-This model assumed no growth for Northern Pikeminnow.
-The second model was run for one full year and assumed no growth for Northern Pikeminnow.
-The assumption of no growth for pikeminnow was included to show what may be occurring if the population of Northern Pikeminnow is stable.
-If growth of the population is occurring, the alternative model shows how an increase of 10% body weight changes the estimates of consumption over a full year period.
-To estimate the number of possible Chinook Salmon that could be consumed, we assumed a weight of 12 g per individual Chinook Salmon.
+Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon based on an estimate from @Moss2016.
+We used a weight-length formula [@Parker1995] to calculate the average predator start weight, converted from the average length of Northern Pikeminnow captured during our study and using the FSA package [@Ogle2021] in R.
+Continuous water temperature data from 2019 was used in two alternative models, one using the baseline start and end weights and the second with a 10% increase in average end weight on Northern Pikeminnow.
+
+We ran three separate bioenergetic models for this study.
+For the first model, we used temperature data from a 78-day period of September 1 through November 17 where DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration.<!-- Did we just use data from Lemhi here?
+All tributaries?
+Was this subjective?
+--> During this time, water temperatures typically exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon<!--citation?-->.
+This first model assumed no growth for Northern Pikeminnow.
+The second model was instead run for one full year, but also assumed no growth for Northern Pikeminnow.
+The assumption of zero growth was included to estimate consumption if either individual Northern Pikeminnow were not growing and/or the population is stable.
+Finally, the alternative third model shows how an increase of 10% body weight or population size changes the estimates of consumption over a full year period.
+
+The model runs provided us with an estimate of the total biomass consumed by an individual Northern Pikeminnow for the given time period.
+We then multipled biomass consumed by an individual Northern Pikeminnow by the estimated population size for that time period to get an estimate of the total biomass consumed.
+The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed by dividing by a eight of 12 g per individual.
 This weight was an average from juvenile Chinook Salmon emigrating past six rotary screw traps operating in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.

### 2021-12-07 analysis/paper/paper.Rmd
cleaned a bit of methods and code

@@
 
-The first objective was to estimate the population size of piscine predators in the Deadwater Slough to begin to their potential impacts on juvenile Chinook Salmon.
-Our initial intent was to estimate predator abundances using mark-recapture methods during fall 2019 and spring 2020 during the emigration of the 2018 brood year.
+Our first objective was to estimate the population size of piscine predators in the Deadwater Slough to begin to understand their potential impacts on juvenile Chinook Salmon.
+The intent was to estimate predator abundances using mark-recapture methods during fall 2019 and spring 2020 during emigration of the 2018 brood year.
 We attempted to sample near the peaks of the juvenile DSR (fall) and NRR (spring) emigrations.
@@
 
-```{r cmr-effort-data}
+```{r capture-data, eval = T}
 raw_df = readxl::read_excel(here("analysis/data/raw_data/Deadwater_predation_mark_recapture_11.09.21.xlsx")) %>%
@@
                 ~ ymd_hms(paste(year(date), month(date), day(date), hour(.), minute(.), second(.))))) %>%
-  # split into sampling periods
+  # break days into "sampling events"
   mutate(event_name = if_else(month(date) > 8,
@@
 
-# focus on pikeminnow
+# focus on pikeminnow, to be used for mark-recap models
 cmr_df = raw_df %>%
@@
 
+```
+
+```{r effort, eval = T}
 effort_df = raw_df %>%
@@
   distinct() %>%
+  filter(method == "Angling") %>%
   mutate(across(total_effort,
                 ~ . * crew_count)) %>%
+  rename(total_effort_min = total_effort) %>%
   group_by(event_name,
            date) %>%
-  summarise(across(c(total_effort,
+  summarise(across(c(total_effort_min,
                      crew_count),
@@
 
+# What size (mm) is considered a large pikeminnow?
+large_size = 228
+
+# catch per unit effort
 cpue_df = effort_df %>%
-  # add how many pikeminnow were caught
+  # add back in the number of pikeminnow caught
   left_join(cmr_df %>%
@@
               summarise(n_fish = n(),
-                        n_large_fish = sum(length>=228,
+                        n_large_fish = sum(length >= large_size,
                                            na.rm = T),
@@
   # calculate CPUE a few different ways
-  mutate(cpue = n_fish / total_effort,
-         cpue_per_day = n_fish / crew_count,
-         cpue_large_per_day = n_large_fish / crew_count)
-
-```
-
-```{r lincoln-petersen-estimator}
+  mutate(catch_per_min = n_fish / total_effort_min,
+         catch_per_hr  = n_fish / (total_effort_min / 60),
+         catch_per_angler_day = n_fish / crew_count,
+         catch_lg_per_angler_day = n_large_fish / crew_count)
+
+```
+
+```{r lincoln-petersen-estimator, eval = T}
 # using Lincoln-Petersen model, combining first 4 and last four days
@@
 
+# add cap_period column, fill w/ NA
 cmr_df$cap_period = NA_character_
+# fill cap_period w/ Mark or Recapture
 for(i in 1:nrow(cap_periods)) {
@@
 
+# Lincoln-Peterson Model
 LP_mod = cmr_df %>%
@@
                         with(x,
-                             mrClosed(M = M,
+                             FSA::mrClosed(M = M,
                                           n = n,
@@
                 ~ str_replace(., "_", " "))) %>%
-  group_by(species, event_name) %>%
-  summarise(across(c(total_effort, 
+  group_by(species,
+           event_name) %>%
+  summarise(across(c(total_effort_min,
                      n_fish,
@@
             .groups = "drop") %>%
-  mutate(across(total_effort,
-                ~ . / 60)) %>%
-  mutate(cpue = n_fish / total_effort,
-         cpue_per_day = n_fish / angler_days,
-         cpue_large_per_day = n_large_fish / angler_days)
+  mutate(catch_per_min = n_fish / total_effort_min,
+         catch_per_hr = n_fish / (total_effort_min / 60),
+         catch_per_angler_day = n_fish / angler_days,
+         catch_lg_per_angler_day = n_large_fish / angler_days)
 
@@
              by = c("event_name", "species")) %>%
-  select(species, model, 
+  select(species,
+         model,
          model_event = event_name,
@@
          Uci_F = Uci,
-         cpue_F = cpue) %>%
+         cpue_F = catch_per_hr) %>%
   full_join(total_cpue %>%
@@
               select(species,
-                     cpue_S = cpue),
+                     cpue_S = catch_per_hr),
             by = "species") %>%
@@
   ggplot(aes(x = event_name,
-             y = cpue_large_per_day,
+             y = catch_lg_per_angler_day,
              fill = event_name)) +

### 2021-12-08 analysis/paper/paper.Rmd
draft abundance results

@@
 
-Abundance estimates for Northern Pikeminnow in the Deadwater Slough during the fall 2019 and 2020 sampling events ranges from `r prettyNum(min(all_est$N), big.mark = ",")` to `r prettyNum(max(all_est$N), big.mark = ",")`.
-The multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied (Figure \@ref(fig:abund-fig)).
-<!-- Need to fill in these fall results more -->
-
-<!-- Add spring abundance results here -->
+We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])` and `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])` Northern Pikeminnow during Fall 2019 and Fall 2020, respectively.
+The catch per unit effort, measured as number of Northern Pikeminnow captured per angler hour, for each was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for Fall 2019 and `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for Fall 2020 (Figure \@ref(fig:cpue-fig)).
+During Fall 2019 we had a total of `r sum(N_mods[[3]][[1]]$m)` recaptures which included `r LP_mod[[3]][[1]]$m` unique individuals i.e.
+`r sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m` was captured more than once.
+There were `r sum(N_mods[[3]][[3]]$m)` recaptures in Fall 2020 which included `r LP_mod[[3]][[3]]$m` unique individuals.
+Summaries of parameters for both the single-census and multiple-census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.
+
+Abundance estimates for Northern Pikeminnow in the Deadwater Slough during Fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for Fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")`.
+In both cases, the multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied (Figure \@ref(fig:abund-fig)).
+Table \@ref(tab:abund-tab) provides point estimates and confidence intervals for each estimator for the fall mark-recapture sampling events.
+
+We captured an additional `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during Spring 2021 and CPUE was `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` fish per angler hour.
+***Finish spring abundance results.*** <!-- Can we assume equal capture probability and perhaps using the ratio of the average fall CPUE and spring CPUE.
+Then multiply the average fall abundance estimate by the ratio.
+Does that seem appropriate?
+-->
 
 The PSD for Northern Pikeminnow in our study and across all three years was `r prettyNum(psd, digits = 1)`%.
-In other words, `r prettyNum(psd, digits = 1)`% of captured fish were over the quality size classification of `r prettyNum(q_size, digits = 0)` mm TL, demonstrating that a high percentage of Northern Pikeminnow within Deadwater Slough are above average size classes defined for the species <!-- Is this useful?
-And if useful, does it make sense to calculate PSD separately for the fall and spring periods?
-I'd expect PSD to vary between seasons.
--->
-
-<!-- Include some results for density e.g., pikeminnow per 100 linear meters or pikeminnow per acre. -->
+In other words, `r prettyNum(psd, digits = 1)`% of captured fish were over the quality size classification of `r prettyNum(q_size, digits = 0)` mm TL, demonstrating that a high percentage of Northern Pikeminnow within Deadwater Slough are above average size classes defined for the species
 
@@
         format.args = list(big.mark = ","),
-        caption = "Data summary for single census estimators.
-M is the number of fish caught in the first sample, marked and returned to the population.
-n is the number of fish caught in the second sample, and m is the number of marked fish caught in the second sample.") %>%
+        caption = "Data summary for single-census mark-recapture estimators.
+M is the total number of fish caught in the first sample event, marked and returned to the population; n is the total number of fish caught in the second sample, and m is the number of previously marked fish caught in the second sample event.") %>%
   kable_styling()
@@
         format.args = list(big.mark = ","),
-        caption = "Data summary for multiple census estimators.
-n is number of fish caught, m is the number of marked fish caught, u is the number of unmarked fish caught, and R is the number of marked fish returned to the population.") %>%
+        caption = "Data summary for multiple-census mark-recapture estimators.
+M is the total number of fish caught in previous events, marked and returned to the population; n is the total number of fish caught in the current event; m is the number of marked fish caught in that event; u is the number of unmarked fish caught, and R is the number of marked fish returned to the population.") %>%
   kable_styling()
@@
         format.args = list(big.mark = ","),
-        caption = "Temporary caption.") %>%
+        caption = "Point estimates, standard errors, and confidence intervals for each of the mark-recapture estimators and for the fall sampling events.
+Standard errors are not available for the multiple-census estimators.") %>%
   kable_styling()
@@
 
-```{r abund-fig, fig.cap = "Estimates of abundance of northern pikeminnow using different estimators."}
+```{r cpue-fig, fig.cap = "The number of Northern Pikeminnow caught per angler effort for each sampling event."}
+total_cpue %>%
+  ggplot(aes(x = event_name,
+             y = catch_per_hr)) +
+  geom_col(fill = "black") +
+  labs(y = "Catch Per Angler Hour") +
+  theme(axis.title.x = element_blank()) +
+  theme_classic()
+
+# total_cpue %>%
+#   ggplot(aes(x = event_name,
+#              y = catch_lg_per_angler_day,
+#              fill = event_name)) +
+#   geom_col() +
+#   scale_fill_brewer(palette = "Set1",
+#                     name = "Event") +
+#   labs(x = "Event",
+#        y = "CPUE (Large Pikeminnow caught per angler day)") +
+#   theme(legend.position = "none")
+
+```
+
+\newpage
+
+```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow using different mark-recapture estimators and for the fall sampling events. Note the y-axis is truncated at 100."}
 all_est %>%
@@
              color = model)) +
+  geom_point(size = 4) +
   geom_errorbar(aes(ymin = Lci,
@@
                 width = 0.1) +
-  geom_point(size = 4) +
   facet_wrap(~ event_name) +
-  scale_color_brewer(palette = "Set1") +
   coord_cartesian(ylim = c(0, 100)) +
   labs(x = "Estimator",
-       y = "Pikeminnow Abundance (x 1,000)") +
+       y = "Abundance (x 1,000)") +
+  theme_bw() +
   theme(legend.position = "none",
         axis.text.x = element_text(angle = 45,
-                                   hjust = 1))
-
-```
-
-\newpage
-
-```{r cpue-fig, fig.cap = "Total catch per unit effort across entire sampling event."}
-total_cpue %>%
-  ggplot(aes(x = event_name,
-             y = catch_lg_per_angler_day,
-             fill = event_name)) +
-  geom_col() +
-  scale_fill_brewer(palette = "Set1",
-                    name = "Event") +
-  labs(x = "Event",
-       y = "CPUE (Large Pikeminnow caught per angler day)") +
-  theme(legend.position = "none")
+                                   hjust = 1)) +
+  scale_color_brewer(palette = "Dark2")
+
+# all_est %>%
+#   mutate(across(c(N, Lci, Uci),
+#                  ~ . / 1000)) %>%
+#   ggplot(aes(x = model,
+#              y = N,
+#              color = model)) +
+#   geom_errorbar(aes(ymin = Lci,
+#                     ymax = Uci),
+#                 width = 0.1) +
+#   geom_point(size = 4) +
+#   facet_wrap(~ event_name) +
+#   scale_color_brewer(palette = "Set1") +
+#   coord_cartesian(ylim = c(0, 100)) +
+#   labs(x = "Estimator",
+#        y = "Abundance (x 1,000)") +
+#   theme(legend.position = "none",
+#         axis.text.x = element_text(angle = 45,
+#                                    hjust = 1))
 

### 2021-12-08 analysis/paper/paper.Rmd
done w draft...going to compile comments for co-authors

@@
 
-The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* each providing important ecological, social, and economic benefits to the region.
+The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations, each providing important ecological, social, and economic benefits to the region.
 The distinct populations include Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017], all of which have become depleted in recent decades.
-Reduction in overall population size, and more specifically, declines in survival of juvenile Chinook Salmon have been associated with hydrosystem passage [@Schaller2014], ocean conditions, regional climate [@Welch2021; @Gosselin2021], and downstream predation by native and non-native piscivorous fish [@Poe1991; @Friesen1999; @Harvey2004].
+Reduction in overall population size, and more specifically, declines in survival of juvenile Chinook Salmon have been associated with hydrosystem passage [@Schaller2014], ocean conditions, regional climate [@Welch2021; @Gosselin2021], and downstream predation by native and non-native piscivorous fishes [@Poe1991; @Friesen1999; @Harvey2004].
 Anthropogenic changes to freshwater habitat including removal of beavers, river channel simplification, water withdrawals, and the proliferation of non-native species (e.g., non-native coastal rainbow trout *O.
-mykiss irideus* and brook trout *Salvelinus fontinalis*) among others [@IdahoOSCTeam2019] have also been attributed to the reductions in the population.
-Moreover, the abundance of returning adults is futher impacted by ocean and mainstem harvests, poor ocean conditions, and changes to the spawning migration corridor [@FCRPSBiOp2020].
+mykiss irideus* and brook trout *Salvelinus fontinalis*) have also been attributed to reductions in population size [@IdahoOSCTeam2019].
+Moreover, the abundance of returning adults is further impacted by ocean and mainstem harvests, poor ocean conditions, and changes to the spawning migration corridor [@FCRPSBiOp2020].
 Each of these factors have contributed, to varying extents, to reduced adult escapement, the primary metric used to assess population viability.
@@
 
-One potentially important, but perhaps under-appreciated source of mortality for Upper Salmon Chinook Salmon is predation on emigrating juveniles by piscine fishes.
-Research in the lower mainstem Snake and Columbia rivers has estimated that during peak emigration, predation on out-migrating salmonids has a significant negative impact (most notably in the presence on non-native piscivores) on the overall population and success of population recovery [@Fresh2003; @Winther2019].
-However, these studies have focused on larger, mainstem and reservoir habitats and direct predation, but very little is known about the interaction between upstream, river dwelling piscine fishes, and their impact on salmonid recovery above reservoir systems of the Columbia and Snake Rivers [@Rubenson2020].
-Adding to this, there is little consensus within the literature related to the magnitude of predation on salmonids in less altered, free flowing systems, with some suggesting lower levels of predation [@Brown1981; @Buchanan1981] and others suggesting increased levels of predation can occur [@Tabor1993; @Ward1995; @Shively1996; @Zimmerman1999].
-
-Dams and reservoirs in the Columbia River basin are the primary locations associated with high rates of predation on salmonids [@Petersen1994; @Ward1995].
+One potentially important, but perhaps under-appreciated source of mortality for Upper Salmon River Chinook Salmon is predation on emigrating juveniles by piscivorous fishes.
+In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recovery, most notably in the presence of non-native piscivores [@Fresh2003; @Winther2019].
+However, these studies have focused on large rivers and reservoirs of the Snake and Columbia rivers whereas very little is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems of the Columbia and Snake river [@Rubenson2020], including the Salmon River.
+Adding to this, there is little consensus within the literature related to the magnitude of predation on salmonids in less altered, free flowing systems, with some suggesting low levels of predation [@Brown1981; @Buchanan1981] while others suggest more impactful levels of predation [@Tabor1993; @Ward1995; @Shively1996; @Zimmerman1999] that may hinder recovery.
+
+Dams and reservoirs in the Columbia River basin are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995].
 There are generally two mechanisms that explain these high predation zones.
 First, migration rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation.
-Second, reservoirs and downstream tailraces resulting from dams on the Columbia and Snake rivers have created favorable slow-water habitat for several species of piscivorous fish known to consume juvenile Chinook Salmon including both non-native (e.g., Smallmouth Bass *Micropterus dolomieu*, Walleye *Sander vitreus*) and native (e.g., Northern Pikeminnow *Ptychocheilus oregonensis*) species.
-In the Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of non-native Smallmouth Bass during some seasons [@Erhardt2018].
-And non-native Walleye are considered a potentially important source for juvenile Chinook Salmon mortality in the Snake River[@Widener2021], particularly as their abundance has significantly increased surrounding Snake River dam facilities between 2008 and 2020 related to their preference for slow water habitats.
-Much effort has been devoted to understanding predation related mortality throughout the hydropower system of the Lower Snake and Columbia rivers, but little is known about native and non-native predation on salmonids throughout free-flowing sections of the Columbia Basin such as the Salmon River [@Rubenson2020].
-
-Along with non-native species, the Salmon River supports native Northern Pikeminnow, a piscine fish tolerating and thriving in relatively warm, slow water habitats.
+Second, reservoirs and downstream tailraces resulting from dams on the Columbia and Snake rivers have created favorable slow-water habitat for several fishes known to consume juvenile Chinook Salmon including both non-native (e.g., Smallmouth Bass *Micropterus dolomieu*, Walleye *Sander vitreus*) and native (e.g., Northern Pikeminnow *Ptychocheilus oregonensis*) species.
+In the Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of Smallmouth Bass during some seasons [@Erhardt2018].
+Similarly, non-native Walleye are considered a potentially important source for juvenile Chinook Salmon mortality in the Snake River [@Widener2021], particularly as their abundance has significantly increased surrounding Snake River dam facilities between 2008 and 2020.
+
+Along with non-native species, the Columbia and Salmon rivers support native Northern Pikeminnow, a piscivorous fish tolerating and thriving in relatively warm, slow water habitats.
 Consequently, they too have benefited from dams on the Columbia River and become abundant predators on salmonid outmigrants [@Knutsen1999].
-Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the estimated 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996].
+Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996].
 Predation by piscivorous fishes, including Northern Pikeminnow, may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@Widener2021].
-However, there is a dearth of current data on piscivorous predator populations in the Columbia River with little known about their movement, life history, or impacts upstream of the mainstem habitats [@Gadomski2001].
+
+While significant effort has been devoted to understanding predation-related mortality throughout the hydropower systems of the lower Snake and Columbia rivers, there is a dearth of contemporary data on piscivorous predator populations and little is known about their movement, life history, or impacts upstream of the mainstem habitats [@Gadomski2001].
+Consequently, the interaction between upstream, river-dwelling piscivorous fishes and emigrating juvenile salmonids, and its impact on salmonid recovery above reservoir systems of the Columbia and Snake rivers is unknown [@Rubenson2020].
 
 Habitats containing slower water velocities and other attributes that support piscivorous predators are not limited to areas near dams and reservoirs.
-One such example is the Deadwater Slough on the mainstem Salmon River, Idaho which is an unnaturally slow and deep reach approximately 1.5 kilometers in length and occupying 29.5 acres (Figure \@ref(fig:map)).
-The slough is located approximately 6 kms downstream from the town of North Fork, Idaho and contains anthropogenically created "reservoir-like" conditions which typically favor piscivorous fish predators like Northern Pikeminnow or Smallmouth Bass [@Watkins2015], and further, lacks hydrological and structural features with little to no fish cover (e.g., large wood or substrate, woody debris, bank undercuts) that can provide essential refuge from predation for juvenile fishes.
-The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
-Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River, forming an alluvial fan and creating an unnaturally slow and deep section of the river resembling a small reservoir.
+One such example is the Deadwater Slough, an approximately 1.5 kilometers long reach of the Salmon River, Idaho (Figure \@ref(fig:map)).
+The slough is located approximately 6 kms downstream from the town of North Fork, Idah with the downstream end located at the confluence with Dump Creek.
+Around 1987, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River, forming an alluvial fan.
+The influx of sediment created an unnaturally slow and deep section of the river, resulting in "reservoir-like" conditions favorable to predators like Northern Pikeminnow and Smallmouth Bass [@Watkins2015].
+Furthermore, the slough lacks hydrological and structureal features that would provide juvenile fishes essential refuge from predation (e.g., large wood or substrate, bank undercuts).
 As a result, predation of juvenile Chinook Salmon and other local ESA-listed salmonids (Sockeye Salmon *O.
@@
 Diversity of migratory tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency.
-Adding to its importance, the slough is part of the migratory pathway for not only juvenile Chinook Salmon (both DSR and NRR emigrants), but also for the endangered Snake River Sockeye Salmon population and several population of threatened Snake River steelhead.
-As such, the main Salmon River has been the target of recent radio telemetry studies to examine the downstream movement, distribution, and (apparent survival) of juvenile salmonid emigrants.
-@Axel2015 demonstrated decreased rates of emigration and survival for Sockeye Salmon from Redfish Lake during the spring emigration.
-Further, recent telemetry studies of Chinook salmon emigration during fall and early winter months indicated a decreased transition probability in the reach spanning Deadwater Slough (approximately 10% less than surrounding reaches).
-
+Because the Deadwater Slough both supports rearing for DSR migrants and provides the migration pathway for NRR migrants, it represents a significant habitat for the Upper Salmon MPG.
+
+The slough is also part of the migratory pathway for the endangered Snake River Sockeye Salmon population and several populations of threatened Snake River steelhead.
+Accordingly, the mainstem Salmon River has been the target of recent radio telemetry studies to examine the downstream movement, distribution, and (apparent survival) of juvenile salmonid emigrants.
+For example, @Axel2015 examined the movement and survival of Sockeye Salmon released in Redfish Lake Creek downstream through the mainstem Salmon River during the spring migration and identified a substantial decrease in estimated survival in the reach spanning the Deadwater Slough.
+They estimated survival among release groups to be approximately 56% whereas survival through the majority of study reaches was greater than 80%.
+Further, recent telemetry studies of Chinook Salmon emigration during the fall and early winter months indicated a decreases transition probability in the reach spanning the Deadwater Slough, approximately 10% less than surrounding reaches [@Ackerman2018; @Porter2019].
+
+In this study, we estimated the abundance of a piscivorous fish population (Northern Pikeminnow) in the Deadwater Slough and its potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
 We hypothesize that increased densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon [@Ackerman2018] and Sockeye Salmon [@Axel2015].
-To test this, we estimated the abundance of a piscivorous fish predator population (Northern Pikeminnow) in the Deadwater Slough and their potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
-Our objectives for the study were three-fold:
+To test this, our objectives for the study were three-fold:
 
@@
 
-This research is an essential first step in characterizing conditions through the Deadwater Reach and providing data necessary to discuss potential mitigation strategies (if needed), where Northern Pikeminnow removal has proven fruitful in increasing survival through the reservoir systems [@Winther2019].
+This research is an essential first step in characterizing conditions through the Deadwater Reach and providing data necessary to discuss potential predation mitigation strategies (if needed).
 We follow with a discussion of the various assumptions that went into the mark-recapture and bioenergetics models and how violations of some assumptions may affect overall results and inferences from the study.
 We also consider how consumption of juvenile Chinook Salmon by predators at Deadwater Slough might impact adult returns, hindering recovery of populations in the Upper Salmon River MPG.
+
+<!-- where Northern Pikeminnow removal has proven fruitful in increasing survival through the reservoir systems [@Winther2019]. -->
 
@@
 
-## Predation On Juvenile Chinook Salmon
+## Gastric Lavage
 
@@
 gastric_df = raw_df %>%
+  select(event_name, date, species, method, lavage_id, length, stomach_contents, comments, stomach_contents_weight,                             fish_content_weight) %>%
+  mutate(bin_50 = lencat(length, w = 50)) %>%
   filter(species == "Northern Pikeminnow") %>%
-  mutate(fish_or_parts = case_when(stri_detect_fixed(stomach_contents, "Fish") ~ "Fish or Fish Parts",
-                                   stri_detect_fixed(stomach_contents, "Empty") ~"Empty",
-                                   stri_detect_fixed(comments, "Whole scuplin") ~ "Fish or Fish Parts",
-                                   stri_detect_fixed(stomach_contents, "Whitefish") ~ "Fish or Fish Parts",
-                                   TRUE ~ "Other")) %>%
-  drop_na(length) %>%
   mutate(non_fish_wt = stomach_contents_weight - fish_content_weight) %>%
-  dplyr::select(date, method, length, lavage_id, stomach_contents, comments, fish_or_parts, stomach_contents_weight, non_fish_wt, fish_content_weight) %>%
-  filter(!stomach_contents %in% c("Lost sample CG27-001", "Did Not Lavage")) %>%
-  drop_na(stomach_contents)
-
-# hungry_fish_df = gastric_df %>%
-#   filter(fish_or_parts == "Fish or Fish Parts") %>%
-#   mutate(fish = case_when(stri_detect_fixed(stomach_contents, "Shiner") ~ "Redside Shiner",
-#                           stri_detect_fixed(stomach_contents, "Whitefish") ~ "Mountain Whitefish",
-#                           stri_detect_fixed(stomach_contents, "scuplin") ~ "Sculpin",
-#                           stri_detect_fixed(stomach_contents, "Chinook") ~ "Chinook",
-#                           stri_detect_fixed(comments, "whitefish") ~ "Mountin Whitefish",
-#                           stri_detect_fixed(comments, "scuplin") ~ "Sculpin",
-#                           stri_detect_fixed(comments, "chinook") ~ "Chinook",
-#                           TRUE ~ "Other"))
-
-```
-
-```{r gastric-summ}
-gastric_df$npm_50 = lencat(gastric_df$length, w = 50)
+  mutate(lavaged = case_when(
+    is.na(lavage_id) & is.na(stomach_contents) ~ FALSE,
+    stomach_contents == "Did Not Lavage" ~ FALSE,
+    stomach_contents == "Lost sample CG27-001" ~ FALSE,
+    TRUE ~ TRUE
+  )) %>%
+  mutate(had_contents = case_when(
+    lavaged == FALSE ~ FALSE,
+    is.na(stomach_contents) ~ FALSE,
+    stomach_contents == "Empty" ~ FALSE,
+    TRUE ~ TRUE
+  )) %>%
+  mutate(had_fish_contents = case_when(
+    fish_content_weight > 0 ~ TRUE,
+    stomach_contents == "MacroInvert-other/Detritus/Fish-Unk" ~ TRUE,
+    stomach_contents == "Fish-Unk (fish bone/cartilage)" ~ TRUE,
+    stomach_contents == "MacroInvert-other/Mountain Whitefish" ~ TRUE,
+    stomach_contents == "MacroInvert-other/Algae/Detritus/Fish-Unk" ~ TRUE,
+    stomach_contents == "Fish -Shiner" ~ TRUE,
+    stomach_contents == "Detritus/Fish-Unk" ~ TRUE,
+    stomach_contents == "Fish Unk" ~ TRUE,
+    stomach_contents == "Detritus/Fish -Shiner" ~ TRUE,
+    stomach_contents == "MacroInvert-other/Fish-Unk" ~ TRUE,
+    stomach_contents == "MacroInvert-other/Detritus/Fish -Shiner" ~ TRUE,
+    stomach_contents == "Fish-Unk" ~ TRUE,
+    stomach_contents == "MacroInvert-Other/Detritus/Fish-Unk" ~ TRUE,
+    stomach_contents == "MacroInvert-Other/Detritus/Fish-Unk/Other"  ~ TRUE,
+    stomach_contents == "Aquatic Veg/Fish-Sucker"  ~ TRUE,
+    stomach_contents == "MacroInvert-Worm/Fish-RedsideShiner"  ~ TRUE,
+    stomach_contents == "Aquatic Veg/Detritus/Fish-RedsideShiner"  ~ TRUE,
+    stomach_contents == "MacroInvert-Other/Fish-Unk" ~ TRUE,
+    stomach_contents == "Fish-Sucker/Other" ~ TRUE,
+    stomach_contents == "Fish-Unk/Other" ~ TRUE,
+    stomach_contents == "MacroInvert-Other/Aquatic Veg/Detritus/Fish-Sucker" ~ TRUE,
+    stomach_contents == "Fish-RedsideShiner" ~ TRUE,
+    stomach_contents == "Fish-Chinook/Other" ~ TRUE,
+    stomach_contents == "Detritus/Fish-RedsideShiner/Other" ~ TRUE,
+    stomach_contents == "MacroInvert-Other/Aquatic Veg/Detritus/Fish-RedsideShiner" ~ TRUE,
+    stomach_contents == "Detritus/Fish-Sucker/Other" ~ TRUE,
+    stomach_contents == "Invertebrates;FishParts"  ~ TRUE,
+    stomach_contents == "Invertebrates;Detritus;FishParts;IntestinalWorms"  ~ TRUE,
+    stomach_contents == "FishParts"  ~ TRUE,
+    stomach_contents == "FishParts;IntestinalWorms"  ~ TRUE,
+    TRUE ~ FALSE
+  )) %>%
+  mutate(had_chinook = case_when(
+    stomach_contents == "Fish-Chinook/Other" ~ TRUE,
+    TRUE ~ FALSE
+  ))
+
+gastric_tbl = gastric_df %>%
+  group_by(event_name) %>%
+  summarise(captured = n(),
+            lavaged = sum(lavaged),
+            stmch_contents = sum(had_contents),
+            fish_contents = sum(had_fish_contents)) %>%
+  adorn_totals("row")
+
+stmch_df = gastric_df %>%
+  filter(stomach_contents_weight > 0) %>%
+  select(event_name,
+         date,
+         length,
+         stomach_contents_weight,
+         fish_content_weight,
+         non_fish_wt) %>%
+  mutate(p_fish = fish_content_weight / stomach_contents_weight)
 
 psd_df = gastric_df %>%
-  group_by(npm_50) %>%
+  group_by(bin_50) %>%
   summarize(n = n())
-
-gastric_df2 = gastric_df %>%
-  drop_na(stomach_contents_weight) %>%
-  group_by(fish_or_parts)
-
-stmch_ctgry_summ = gastric_df2 %>%
-  count()
-
-stmch_wt_summ = gastric_df2 %>%
-  summarise(sum(stomach_contents_weight))
-
-non_fish_wt_summ = gastric_df2 %>%
-  summarise(sum(non_fish_wt))
-
-fish_wt_summ = gastric_df2 %>%
-  summarise(sum(fish_content_weight))
 
@@
 
-## Predation On Juvenile Chinook Salmon
-
-During 2019, 660 Northern Pikeminnow were gastric lavaged.
-Of those fish, we observed 603 individuals with empty stomachs, and 57 individuals with stomach contents.
-Twelve of the samples contained fish remnants including a single Mountain Whitefish and two Redside Shiners.
-During the 2020 season, 805 Northern Pikeminnow were gastric lavaged comprised of 613 individuals with empty stomachs and 188 individuals with stomach contents.
-Twenty-three individuals had fish remnants in their stomach contents including ten with unidentifiable remnants, six Redside Shiners, four suckers, two sculpin, and one juvenile Chinook Salmon.
-In 2021, we collected stomach contents from 105 Northern Pikeminnow with 98 containing non-fish prey items and observed seven containing fish remnants.
-
-<!--- Need to add Nick's gastric lavage table here -->
+## Gastric Lavage
+
+We examined the stomach contents of `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow over the course of the study using gastric lavage methods.
+Of those, we documented some form of stomach contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` of fish of which we were able to confirm fish or fish remnants in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")`.
+Expressed as a percentage, `r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% of lavaged Northern Pikeminnow had stomach contents of which `r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% we identified containing fish or fish remnants.
+Table \@ref(tab:gastric-tab) provides a further breakdown of gastric lavage results for each sampling event.
+
+For Northern Pikeminnow with some form of stomach contents, the average wet weight of total contents was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r prettyNum(min(stmch_df$stomach_contents_weight), digits = 3)`-`r prettyNum(max(stmch_df$stomach_contents_weight), digits = 3)` g).
+Overall, fish or fish remnants were `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall weighted weight among all stomach contents examined.
 
 ## Bioenergetics
+
+We estimated a single average-sized Northern Pikeminnow to consume 61 g of ***fish*** during the modeled fall time period (September 1 - November 17).
+Given an estimated population size of ***xx,xxx***, we'd estimate that ***xxx,xxx*** g of fish are consumed each fall during the emigration of juvenile DSR Chinook Salmon.
+The average weight of a DSR Chinook Salmon is ***xx.x*** taken from ***yyy*** rotary screw traps
+
+If we further assume an average weight of ***xx g*** for DSR Chinook Salmon
 
@@
 Standard errors are not available for the multiple-census estimators.") %>%
+  kable_styling()
+
+```
+
+\newpage
+
+```{r gastric-tab}
+gastric_tbl %>%
+  kable(booktabs = T,
+        format.args = list(big.mark = ","),
+        caption = "Summary of gastric lavage results including the total number of fish captured and gastric lavaged.
+For Northern Pikeminnow that were gastric lavaged, we show the number containing stomach contents, and of those, the number containing fish contents.") %>%
   kable_styling()

### 2021-12-09 analysis/paper/paper.Rmd
some TB edits to Methods

@@
 
-This research is an essential first step in characterizing conditions through the Deadwater Reach and providing data necessary to discuss potential predation mitigation strategies (if needed).
+This research is an essential first step in characterizing conditions through the Deadwater Reach and providing data necessary to discuss potential predation mitigation strategies (if needed), where Northern Pikeminnow removal has proven fruitful in increasing survival through the reservoir systems [@Winther2019].
 We follow with a discussion of the various assumptions that went into the mark-recapture and bioenergetics models and how violations of some assumptions may affect overall results and inferences from the study.
@@
 
-<!-- where Northern Pikeminnow removal has proven fruitful in increasing survival through the reservoir systems [@Winther2019]. -->
-
 # Methods
@@
 
-Our first objective was to estimate the population size of piscine predators in the Deadwater Slough to begin to understand their potential impacts on juvenile Chinook Salmon.
-The intent was to estimate predator abundances using mark-recapture methods during fall 2019 and spring 2020 during emigration of the 2018 brood year.
-We attempted to sample near the peaks of the juvenile DSR (fall) and NRR (spring) emigrations.
-However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 sampling event was delayed until fall 2020.
-Although this provided the benefit of an additional abundance estimate during the DSR emigration, estimates during the NRR emigration were still lacking, and so an additional sampling event was added for spring 2021.
-The study design was such that the fall 2019 and fall 2020 sampling efforts included both a "mark week" and "recapture week" intended for mark-recapture models.
-However, due to budget constraints the spring 2021 effort only included a single "mark week"; instead, we calculated the catch-per-unit-effort (CPUE; fish per hour) for the spring 2021 sampling event for comparison to the fall efforts.
-In the end, sampling dates were as follows: fall 2019 (mark = November 12-14; recapture = November 19-21), fall 2020 (mark = October 20-23; recapture = October 27-30); spring 2021 (mark = May 18-21).
-The recapture events were held just one week following the mark events to reduce opportunity for immigration/emigration of predators between events i.e., to avoid violation of a closed population for mark-recapture models.
-
-We attempted a multiple of capture methods during the first year including raft electrofishing, fyke netting, snorkeling, and angling to reduce selectivity, both for species and size classes.
+Our first objective was to estimate the population size of piscine predators in the Deadwater Slough using mark-recapture methods.
+We attempted to sample predators near the peaks of the juvenile DSR and NRR emigrations for the 2018 brood year (fall 2019 and spring 2020, respectively).
+However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 effort was moved to fall 2020.
+Although this provided the benefit of an additional abundance estimate during the DSR emigration, estimates during the NRR emigration were still lacking, and so an additional effort was added in spring 2021.
+Each of the fall efforts included sampling across a two-week span; however, due to budget constraints the spring 2021 effort only included a single week.
+For the fall efforts, the first week was considered a "mark event" and the second week a "recapture event" for single-census mark-recapture estimators (described below); therefore, the "recapture week" occurred the week following the "mark week" to minimize predator emigration/immigration and constrain the population for the mark-recapture models.
+
+<!-- In the end, sampling dates were as follows: fall 2019 (mark = November 12-14; recapture = November 19-21), fall 2020 (mark = October 20-23; recapture = October 27-30); spring 2021 (mark = May 18-21).
+-->
+
+Multiple capture methods were employed during the fall 2019 effort to reduce selectivity and bias for species and size classes.
+Methods included raft electrofishing, fyke netting, snorkeling, and angling.
 Unfortunately, some methods proved ineffective during initial efforts.
@@
 Snorkeling was ineffective due to the high turbidity of the mainstem Salmon River and Deadwater Slough area.
-And due to debris floating in the water column, there were concerns with fyke nets being pulled from their anchored locations and becoming entangled.
-Additionally, there were concerns with raft electrofishers and fyke nets incidentally capturing ESA-listed adult steelhead that are present in the mainstem Salmon River during fall and spring.
+Significant amounts of debris in the water column also created concerns that fyke nets could become entangled.
+Additionally, there were concerns that raft electrofishers and fyke nets could incidentally capture ESA-listed adult steelhead that are present in the mainstem Salmon River during fall and spring.
 Eventually, angling proved to be the most effective method for capturing piscine predators while minimizing potential impacts to steelhead but recognizing that angling can have both species and size selectivity (bias).
@@
 
-Our study relied heavily on volunteer anglers who were allowed to fish anywhere within Deadwater Slough; periodically or upon filling a livewell anglers would boat or hike their catch to a processing station located at the boat ramp approximately 500 m downstream from the top of the slough.
-For each fish, processors would record the date of capture, species, total length (TL; mm), and whether the fish was previously marked.
-If not previously marked, each fish was provided a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) unique for each marking day.
-We also recorded whether marked individuals were released back to be captured again.
-For release, processors attempted to distribute released individuals throughout Deadwater Slough to facilitate mixing back into the population.
-We also recorded the crew (person or combination of persons) that captured each fish and the start and end times for each crew to estimate total effort.
+Our study relied heavily on volunteer anglers who were permitted to fish anywhere within Deadwater Slough.
+During a sample event, anglers would boat or hike their catch (periodically or upon filling a livewell) to a processing station located at the boat ramp approximately 500 m downstream from the top of the slough.
+For each fish, we recorded the date of capture, species, total length (TL; mm), and whether the fish was previously marked.
+Unmarked fish were given a unique physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) each day and released.
+Released individuals were distributed throughout Deadwater Slough to facilitate mixing back into the population.
+Finally, we recorded the angling start and end times for each crew (person or combinations of persons) to facilitate calculations of catch per unit effort (CPUE).
 
@@
 
-Gastric lavage [@Foster1977] was used to examine the stomach contents of Northern Pikeminnow for the presence of juvenile Chinook Salmon, other fishes (e.g., juvenile steelhead, juvenile Sockeye Salmon, Redside Shiner *Richardsonius balteatus*, etc.), and non-targets (e.g., macroinvertebrates, organic matter, etc.).
-Stomach contents of individuals were preserved with 99% isopropyl alcohol in whirl-paks immediately following lavage for later analysis in a controlled environment.
-For each sample, wet weight (grams) was recorded for the total combined stomach content as well as the contents consisting of fish parts/matter.
-Fish and fish remnants were identified to the lowest taxonomic unit, when possible, or were categorized as unknown.
-A subset of Northern Pikeminnow captures were euthanized for dissection after gastric lavage to validate the efficacy of the methodology.
+Gastric lavage [@Foster1977] was used to examine the stomach contents of Northern Pikeminnow for the presence of juvenile Chinook Salmon and other fishes (e.g., juvenile steelhead, juvenile Sockeye Salmon, Redside Shiner *Richardsonius balteatus*, etc.).
+Immediately following lavage, stomach contents of individuals were preserved with 99% isopropyl alcohol in whirl-paks for later analysis in a controlled environment.
+For each sample, wet weight (grams) was recorded for the total combined stomach content, including all non-fish items (e.g., macroinvertebrates, organic matter), as well as the portion consisting of fish parts/matter.
+Fish and fish remnants were identified to the lowest taxonomic unit, when possible, or were categorized as unknonw.
+A subset of Northern Pikeminnow captures (~ 5%) were euthanized for dissection after gastric lavage to validate the efficacy of the methodology.
 
@@
 
-Gastric lavage was used in an attempt to document predation of juvenile Chinook Salmon by Northern Pikeminnow.
-However, we were ultimately interested in the potential total number or amount of juveniles consumed during the peaks of DSR (fall) and NRR (spring) emigrations.
-To estimate the consumption of juvenile Chinook Salmon in the Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 and applied in the R statistical software [@RCoreTeam2021].
+To estimate the total consumption potential (i.e., number/amount of juvenile consumed during the peaks of DSR [fall] and NRR [spring] emigrations) of Northern Pikeminnow in Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 applied in the R statistical software [@RCoreTeam2021].
 The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on predator and prey energy densities, predator start and end weights, and water temperatures.
@@
 Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon based on an estimate from @Moss2016.
-We used a weight-length formula [@Parker1995] to calculate the average predator start weight, converted from the average length of Northern Pikeminnow captured during our study and using the FSA package [@Ogle2021] in R.
+We used a weight-length formula [@Parker1995] to calculate the average predator start weight from the average length of Northern Pikeminnow captured during our study using the FSA package [@Ogle2021] in R.
 Continuous water temperature data from 2019 was used in two alternative models, one using the baseline start and end weights and the second with a 10% increase in average end weight on Northern Pikeminnow.
@@
 We ran three separate bioenergetic models for this study.
-For the first model, we used temperature data from a 78-day period of September 1 through November 17 where DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration.<!-- Did we just use data from Lemhi here?
+For the first model, we used temperature data from a 78-day period spanning September 1 through November 17 when DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration.<!-- Did we just use data from Lemhi here?
 All tributaries?
@@
 The second model was instead run for one full year, but also assumed no growth for Northern Pikeminnow.
-The assumption of zero growth was included to estimate consumption if either individual Northern Pikeminnow were not growing and/or the population is stable.
-Finally, the alternative third model shows how an increase of 10% body weight or population size changes the estimates of consumption over a full year period.
-
-The model runs provided us with an estimate of the total biomass consumed by an individual Northern Pikeminnow for the given time period.
-We then multipled biomass consumed by an individual Northern Pikeminnow by the estimated population size for that time period to get an estimate of the total biomass consumed.
-The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed by dividing by a eight of 12 g per individual.
+The assumption of zero growth was included to estimate consumption if individual Northern Pikeminnow were not growing and/or the population is stable.
+Finally, the third model shows how an increase of 10% body weight or an increase in population size changes the estimates of consumption over a full year period.
+
+To estimate the total biomass consumed, we then multipled biomass consumed by an individual Northern Pikeminnow for the given time period by the estimated population size for that time period.
+The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed by dividing by a weight of 12 g per individual.
 This weight was an average from juvenile Chinook Salmon emigrating past six rotary screw traps operating in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.

### 2021-12-13 analysis/R scripts/abund_est.R
updated the last modified date

@@ -1,7 +1,7 @@
 # Author: Kevin See
 # Purpose: import data from Deadwater sampling
 # Created: 11/9/21
-# Last Modified: 11/12/21
+# Last Modified: 12/2/21
 # Notes:
 
 #-----------------------------------------------------------------

### 2021-12-13 analysis/paper/paper.Rmd
converted some single digit numbers to english

@@
 library(stringi)
+library(english)
 
@@
 The catch per unit effort, measured as number of Northern Pikeminnow captured per angler hour, for each was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for Fall 2019 and `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for Fall 2020 (Figure \@ref(fig:cpue-fig)).
-During Fall 2019 we had a total of `r sum(N_mods[[3]][[1]]$m)` recaptures which included `r LP_mod[[3]][[1]]$m` unique individuals i.e.
-`r sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m` was captured more than once.
-There were `r sum(N_mods[[3]][[3]]$m)` recaptures in Fall 2020 which included `r LP_mod[[3]][[3]]$m` unique individuals.
+During Fall 2019 we had a total of `r as.english(sum(N_mods[[3]][[1]]$m))` recaptures which included `r as.english(LP_mod[[3]][[1]]$m)` unique individuals i.e.
+`r as.english(sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m)` was captured more than once.
+There were `r as.english(sum(N_mods[[3]][[3]]$m))` recaptures in Fall 2020 which included `r as.english(LP_mod[[3]][[3]]$m)` unique individuals.
 Summaries of parameters for both the single-census and multiple-census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.

### 2021-12-15 analysis/R scripts/Nick Script 11.10.21 cleaning fish data.R
Updating different bioenergetics models for Spring and Fall

@@ -49,6 +49,17 @@ Stomach_content_DF = Clean_NPM %>%
   drop_na(StomachContents)
 
 
+Spring_Avg_length = Clean_NPM %>%
+  filter(Date > "2021-05-17") %>%
+  summarise(avg_TL = mean(Length))
+
+Fall_Avg_length = Clean_NPM %>%
+  filter(Date < "2021-05-17") %>%
+  summarise(avg_TL = mean(Length))
+
+
+
+
 #Read in the Bioenergetics
 
 bioenergetics<- read_csv("analysis/data/raw_data/2020 Bioenergetics 1 year.csv") %>%

### 2021-12-15 analysis/R scripts/Nick Script 12.15.21 bioenergetics.R
Update to Bioenergetics

@@ -1,0 +1,279 @@
+# working to clean up and provide summaries for Deadwater Data
+# Nick Porter
+# 11/2/20201
+rm(list = ls()) # clear the working environment
+
+require(tidyverse)
+require(lubridate)
+require(boot)
+require(bootstrap)
+require(lme4)
+require(Matrix)
+require(numDeriv)
+require(FSA)
+require(FSAdata)
+require(magrittr)
+require(car)
+require(dplyr)
+require(plotrix)
+require(nnet)
+#require(Hmisc) # does not work
+#require(Matching)
+require(Rcapture)
+require(nlstools)
+require(stringi)
+require(MASS)
+require(readr)
+
+#Read in the Deadwater dataset
+deadwater <- read_csv("analysis/data/raw_data/deadwater_cmr_effort_20210614.csv") %>%
+  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
+
+
+NPM = deadwater %>%
+  filter(Species == "Northern Pikeminnow")
+
+Clean_NPM = NPM %>%
+  mutate(Fish_or_Parts = case_when(stri_detect_fixed(StomachContents, "Fish") ~ "Fish or Fish Parts",
+                                   stri_detect_fixed(StomachContents, "Empty") ~ "Empty",
+                                   stri_detect_fixed(Comments, "Whole scuplin") ~ "Fish or Fish Parts",
+                                   stri_detect_fixed(StomachContents, "Whitefish") ~ "Fish or Fish Parts",
+                                   TRUE ~ "Other")) %>%
+  drop_na(Length) %>%
+  mutate(Non_fish_wt = StomachContentsWeight - FishContentWeight)
+
+Stomach_content_DF = Clean_NPM %>%
+  dplyr::select(Date, Length, LavageID, StomachContents, Comments,
+                Fish_or_Parts, StomachContentsWeight, Non_fish_wt, FishContentWeight) %>%
+  filter(!StomachContents %in% c("Lost sample CG27-001","Did Not Lavage")) %>%
+  drop_na(StomachContents)
+
+
+Spring_Avg_length = Clean_NPM %>%
+  filter(Date > "2021-05-17") %>%
+  summarise(avg_TL = mean(Length))
+
+Fall_Avg_length = Clean_NPM %>%
+  filter(Date < "2021-05-17") %>%
[…]

### 2021-12-15 analysis/paper/paper.Rmd
updated author order

@@
 author:
-  - Richard A. Carmichael:
-      email: Richard.Carmichael@merck.com
-      institute: [biomark]
+  - Michael W. Ackerman:
+      email: mike.ackerman@mthoodenvironmental.com
+      institute: [mhe_mccall, biomark]
       correspondence: true
@@
       correspondence: false
-  - Michael W. Ackerman:
-      email: mike.ackerman@mthoodenvironmental.com
-      institute: [biomark, mhe_mccall]
+  - Richard A. Carmichael:
+      email: Richard.Carmichael@merck.com
+      institute: [biomark]
       correspondence: false
 institute:
+  - mhe_mccall: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
   - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
   - wdfw: Washington Department of Fish and Wildife, Fish Program, Science Division, 1111 Washington Street NE, Olympia, Washington, 98501, USA
-  - mhe_mccall: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
   - mhe_sandy: Mount Hood Environmental, 39085 Pioneer Boulevard \#100 Mezzanine, Sandy, Oregon, 97055, USA 

### 2021-12-15 analysis/paper/paper.Rmd
minor edits from Google doc

@@
 ---
-title: "Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations"
+title: "Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations in the Upper Salmon River, Idaho"
 author:
@@
 mykiss irideus* and brook trout *Salvelinus fontinalis*) have also been attributed to reductions in population size [@IdahoOSCTeam2019].
-Moreover, the abundance of returning adults is further impacted by ocean and mainstem harvests, poor ocean conditions, and changes to the spawning migration corridor [@FCRPSBiOp2020].
+Moreover, the abundance of returning adults is impacted by ocean and mainstem harvests, poor ocean conditions, and changes to the spawning migration corridor [@FCRPSBiOp2020].
 Each of these factors have contributed, to varying extents, to reduced adult escapement, the primary metric used to assess population viability.
@@
 However, these studies have focused on large rivers and reservoirs of the Snake and Columbia rivers whereas very little is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems of the Columbia and Snake river [@Rubenson2020], including the Salmon River.
-Adding to this, there is little consensus within the literature related to the magnitude of predation on salmonids in less altered, free flowing systems, with some suggesting low levels of predation [@Brown1981; @Buchanan1981] while others suggest more impactful levels of predation [@Tabor1993; @Ward1995; @Shively1996; @Zimmerman1999] that may hinder recovery.
+Adding to this, there is little consensus within the literature about the magnitude of predation on salmonids in less altered, free flowing systems, with some suggesting lower rates of predation [@Brown1981; @Buchanan1981] while others suggest more impactful rates of predation [@Tabor1993; @Ward1995; @Shively1996; @Zimmerman1999] that may hinder recovery.
 
@@
 There are generally two mechanisms that explain these high predation zones.
-First, migration rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation.
+First, movement (swimming) rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation.
 Second, reservoirs and downstream tailraces resulting from dams on the Columbia and Snake rivers have created favorable slow-water habitat for several fishes known to consume juvenile Chinook Salmon including both non-native (e.g., Smallmouth Bass *Micropterus dolomieu*, Walleye *Sander vitreus*) and native (e.g., Northern Pikeminnow *Ptychocheilus oregonensis*) species.
-In the Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of Smallmouth Bass during some seasons [@Erhardt2018].
+In Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of Smallmouth Bass during some seasons [@Erhardt2018].
 Similarly, non-native Walleye are considered a potentially important source for juvenile Chinook Salmon mortality in the Snake River [@Widener2021], particularly as their abundance has significantly increased surrounding Snake River dam facilities between 2008 and 2020.
@@
 
-The slough is also part of the migratory pathway for the endangered Snake River Sockeye Salmon population and several populations of threatened Snake River steelhead.
+The slough is also part of the migratory pathway for the endangered Snake River Sockeye Salmon population [@Axel2015] and several populations of threatened Snake River steelhead.
 Accordingly, the mainstem Salmon River has been the target of recent radio telemetry studies to examine the downstream movement, distribution, and (apparent survival) of juvenile salmonid emigrants.
@@
 They estimated survival among release groups to be approximately 56% whereas survival through the majority of study reaches was greater than 80%.
-Further, recent telemetry studies of Chinook Salmon emigration during the fall and early winter months indicated a decreases transition probability in the reach spanning the Deadwater Slough, approximately 10% less than surrounding reaches [@Ackerman2018; @Porter2019].
+Further, recent telemetry studies of Chinook Salmon emigration during the fall and early winter months indicated a lower transition probability in the reach spanning the Deadwater Slough, approximately 10% less than surrounding reaches [@Ackerman2018; @Porter2019].
 
@@
 Methods included raft electrofishing, fyke netting, snorkeling, and angling.
-Unfortunately, some methods proved ineffective during initial efforts.
+However, some methods proved ineffective during initial efforts.
 For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) our inability to "chase" fish downstream as they evaded the electricity.
@@
 Additionally, there were concerns that raft electrofishers and fyke nets could incidentally capture ESA-listed adult steelhead that are present in the mainstem Salmon River during fall and spring.
-Eventually, angling proved to be the most effective method for capturing piscine predators while minimizing potential impacts to steelhead but recognizing that angling can have both species and size selectivity (bias).
+After evaluating all methods, angling proved to be the most effective method for capturing piscine predators while minimizing potential impacts to steelhead but recognizing that angling can have both species and size selectivity (bias).
 From here on, we focus on fish captured by angling, unless otherwise noted.
@@
 For each fish, we recorded the date of capture, species, total length (TL; mm), and whether the fish was previously marked.
-Unmarked fish were given a unique physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) each day and released.
+Unmarked fish were given a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) unique to each day and released.
 Released individuals were distributed throughout Deadwater Slough to facilitate mixing back into the population.
@@
 
-We used several mark-recapture abundance estimators that each rely on meeting several assumptions:
-
-1. The population is closed (no immigration, emigration, births or deaths during the sampling period)
-1. Marking a fish does not affect its chances of being caught
-1. No loss of marks
-1. No marks are missed or mistaken
-1. All fish have the same chance of being caught in the second (and subsequent) sample(s)
+All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the sampling period), (2) marking a fish does not affect its chances of recapture, (3) no loss of marks, (4) marks are not missed or mistaken, and (5) all fish have equal chance of being caught in the second (and subsequent) sample(s).
 
@@
 
-The Lincoln-Petersen estimator can be biased with small samples, so we also investigated the Chapman-modified Lincoln-Petersen estimator:
+The Lincoln-Petersen estimator can be biased with small samples, so we also employed the Chapman-modified Lincoln-Petersen estimator:
 
@@
 
-For a multiple census estimator we first considered the Schnabel estimator where the $M$, $n$ and $m$ are indexed by the sampling occassion, $i$.
+For a multiple-census estimator, we first utilized the Schnabel estimator where the $M$, $n$ and $m$ are indexed by the sampling occasion, $i$.
 In our case, the sampling occassions are defined as each day of sampling.
@@
 If that assumption is not true and capture probability did differ, we believe that the capture probability  in the spring would be lower due to higher flows, which would lead to our spring abundance estimate to be biased low.
-Therefore, we presume our spring abundance estimate during the NRR Chinook Salmon emigration to be a conservative estimate of the number of Northern Pikeminnow present in the Deadwater Slough.
+Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 sampling event to be a conservative estimate.
 
@@
 
-Lastly, we calculated the proportional stock density (PSD) for Northern Pikeminnow captured during all field efforts, which is a measure of species size structure.
+Finally, we calculated the proportional stock density (PSD) for Northern Pikeminnow captured during all field efforts, which is a measure of species size structure.
 PSD is a ratio, typically expressed as a percentage, between the number of "quality-sized" individuals or larger individuals and "stock-sized" individuals.
@@
 For each sample, wet weight (grams) was recorded for the total combined stomach content, including all non-fish items (e.g., macroinvertebrates, organic matter), as well as the portion consisting of fish parts/matter.
-Fish and fish remnants were identified to the lowest taxonomic unit, when possible, or were categorized as unknonw.
+Fish and fish remnants were identified to the lowest taxonomic unit, when possible, or were categorized as unknown.
 A subset of Northern Pikeminnow captures (~ 5%) were euthanized for dissection after gastric lavage to validate the efficacy of the methodology.
@@
 Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon based on an estimate from @Moss2016.
-We used a weight-length formula [@Parker1995] to calculate the average predator start weight from the average length of Northern Pikeminnow captured during our study using the FSA package [@Ogle2021] in R.
+We used a weight-length formula [@Parker1995] to calculate the average predator start weight from the average length of Northern Pikeminnow captured during our study using the FSA package in R [@Ogle2021].
+The average length of Northern Pikeminnow captured at the Deadwater Slough during the spring and fall were 352.9 and 393.7 mm, respectively, which calculates to average starting weights of 504.4 and 670.5 g.
 Continuous water temperature data from 2019 was used in two alternative models, one using the baseline start and end weights and the second with a 10% increase in average end weight on Northern Pikeminnow.
@@
 
-To estimate the total biomass consumed, we then multipled biomass consumed by an individual Northern Pikeminnow for the given time period by the estimated population size for that time period.
+To estimate the total biomass consumed, we then multiplied the biomass consumed by an individual Northern Pikeminnow for the given time period by the estimated population size for that time period.
 The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed by dividing by a weight of 12 g per individual.
@@
 
-We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])` and `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])` Northern Pikeminnow during Fall 2019 and Fall 2020, respectively.
-The catch per unit effort, measured as number of Northern Pikeminnow captured per angler hour, for each was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for Fall 2019 and `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for Fall 2020 (Figure \@ref(fig:cpue-fig)).
-During Fall 2019 we had a total of `r as.english(sum(N_mods[[3]][[1]]$m))` recaptures which included `r as.english(LP_mod[[3]][[1]]$m)` unique individuals i.e.
-`r as.english(sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m)` was captured more than once.
+We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during Fall 2019, Fall 2020, and Spring 2021, respectively.
+The catch per unit effort, measured as the number of Northern Pikeminnow captured per angler hour, for each was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for Fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for Fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for Spring 2021 (Figure \@ref(fig:cpue-fig)).
+During Fall 2019, we had a total of `r as.english(sum(N_mods[[3]][[1]]$m))` recaptures which included `r as.english(LP_mod[[3]][[1]]$m)` unique individuals (i.e., `r as.english(sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m)` individual was captured more than once).
 There were `r as.english(sum(N_mods[[3]][[3]]$m))` recaptures in Fall 2020 which included `r as.english(LP_mod[[3]][[3]]$m)` unique individuals.
@@
 
-We captured an additional `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during Spring 2021 and CPUE was `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` fish per angler hour.
 ***Finish spring abundance results.*** <!-- Can we assume equal capture probability and perhaps using the ratio of the average fall CPUE and spring CPUE.

### 2021-12-16 analysis/paper/paper.Rmd
formatted references

@@
 mykiss irideus* and brook trout *Salvelinus fontinalis*) have also been attributed to reductions in population size [@IdahoOSCTeam2019].
-Moreover, the abundance of returning adults is impacted by ocean and mainstem harvests, poor ocean conditions, and changes to the spawning migration corridor [@FCRPSBiOp2020].
+Moreover, the abundance of returning adults is impacted by ocean and mainstem harvests, poor ocean conditions, and changes to the spawning migration corridor [@NMFS2020BiOp].
 Each of these factors have contributed, to varying extents, to reduced adult escapement, the primary metric used to assess population viability.
@@
 The Deadwater Slough is in a reach of the Salmon River believed to be a historically important overwinter rearing and spring emigration area for Chinook Salmon juveniles.
-Spring/summer Chinook Salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014].
+Spring/summer Chinook Salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a].
 The DSR migrants leave the natal spawning area as subyearlings between June and november and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts.

### 2021-12-16 analysis/R scripts/Nick Script 12.15.21 bioenergetics.R
Updated code and graphs.
Graphs need some help but bioenergetics are done.

@@ -1,6 +1,6 @@
 # working to clean up and provide summaries for Deadwater Data
 # Nick Porter
-# 11/2/20201
+# 12/15/20201
 rm(list = ls()) # clear the working environment
 
 require(tidyverse)
@@ -59,106 +59,172 @@ Fall_Avg_length = Clean_NPM %>%
 
 #Read in the Bioenergetics
 
-bioenergetics<- read_csv("analysis/data/raw_data/2020 Bioenergetics 1 year.csv") %>%
-  mutate(Date = as.Date(Date, format = "%m/%d/%y")) %>%
-  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+Fall_I10_F90<- read_csv("analysis/data/raw_data/Bioenergetics_Fall_I10_F90.csv") %>%
+  mutate(Date = as.Date(Date, format = "%m/%d")) %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons.fish.g)) %>%
   mutate(Date = as.Date(Date)) %>%
   mutate(month = month(Date))
 
-bioenergetics_10percet <- read_csv("analysis/data/raw_data/2020 Bioenergetics 1 year 10% growth.csv") %>%
-  mutate(Date = as.Date(Date, format = "%m/%d/%y")) %>%
-  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+Fall_I20_F80<- read_csv("analysis/data/raw_data/Bioenergetics_Fall_I20_F80.csv") %>%
+  mutate(Date = as.Date(Date, format = "%m/%d")) %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons.fish.g)) %>%
   mutate(Date = as.Date(Date)) %>%
   mutate(month = month(Date))
 
-Alternative_bioenergetics <- read_csv("analysis/data/raw_data/Bioenergetics 32 fish 68 inverts.csv") %>%
-  mutate(Date = as.Date(Date, format = "%m/%d/%y")) %>%
-  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+Fall_I30_F70<- read_csv("analysis/data/raw_data/Bioenergetics_Fall_I30_F70.csv") %>%
+  mutate(Date = as.Date(Date, format = "%m/%d")) %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons.fish.g)) %>%
   mutate(Date = as.Date(Date)) %>%
   mutate(month = month(Date))
 
[…]

### 2021-12-16 analysis/R scripts/Nick Script 12.15.21 bioenergetics.R
Updates

@@ -179,7 +179,7 @@ spring_week = c("3/1", "3/8", "3/15", "3/22", "3/29", "4/5", "4/19", "4/26", "5/
 #Spring_bioenergetics %>%
  Spring_Graph = ggplot() +
     geom_line(aes(y = Spring_I10_F90$Cumu_fish_eatten, x =Spring_I10_F90$Date, colour = "blue")) +
-    geom_line(aes(y = Spring_I20_F80$Cumu_fish_eatten, x =Spring_I20_F80$Date, colour = "darkred")) +
+    geom_line(aes(y = Spring_I20_F80$Cumu_fish_eatten, x =Spring_I20_F80$Date, colour = "blue")) +
     geom_line(aes(y = Spring_I30_F70$Cumu_fish_eatten, x =Spring_I30_F70$Date, colour = "blue")) +
     geom_line(aes(y = Spring_I40_F60$Cumu_fish_eatten, x =Spring_I40_F60$Date, colour = "blue")) +
     geom_line(aes(y = Spring_I50_F50$Cumu_fish_eatten, x =Spring_I50_F50$Date, colour = "blue")) +
@@ -190,21 +190,19 @@ spring_week = c("3/1", "3/8", "3/15", "3/22", "3/29", "4/5", "4/19", "4/26", "5/
               geom_line(size = 2.5) +
   #scale_color_distiller(palette = "Spectral", direction = -1) +
   #labs(y = "Cumulative fish eatten (g)") +
-  scale_x_date(breaks = "2 weeks") +
-  #scale_x_continuous(breaks = seq(1,13, by = 1), labels = c("3/1", "3/8", "3/15", "3/22", "3/29", "4/5", "4/19", "4/26", "5/3", "5/10", "5/17", "5/24", "5/31")) +
+  scale_x_date(breaks = "1 weeks") +
   scale_y_continuous(name = "Cumulative fish eatten (g)", breaks = seq(0,600, by = 5),
-                     sec.axis = sec_axis(~. /12, name = "Potential number of Chinook smolts", breaks = seq(0,30, by = 1))) +
+                     sec.axis = sec_axis(~. /9.5, name = "Potential number of Chinook smolts", breaks = seq(0,30, by = 1))) +
   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
   theme_classic() +
-  #theme(legend.position = "top") +
-  labs(x = "weeks")
-
+  theme(legend.position = "none") +
+  labs(x = "Week")
[…]

### 2021-12-16 analysis/R scripts/USR screwtraps script 12.16.21.R

@@ -1,0 +1,86 @@
+rm(list = ls())
+
+require(tidyverse)
+require(lubridate)
+require(esquisse)
+
+tags<- Tagging_Detail_Bioenergetics_2020_avg_weight_ <- read_csv("Merck Sharp & Dohme, Corp/Biomark Field Sampling - Documents/General/Deadwater/Tagging Detail (Bioenergetics).csv")
+glimpse(tags)
+
+Clean_tags = tags %>%
+  rename(Site_Name =`Release Site Name`) %>%
+  rename(Mark_Date = `Mark Date MMDDYYYY`) %>%
+  rename(Tag_code = `Tag Code`) %>%
+  rename(Weight = `Weight g`) %>%
+  rename(Length = `Length mm`) %>%
+  rename(Species = `Species Name`) %>%
+  rename(Count = `Mark Count`) %>%
+  mutate(Mark_Date = as.Date(Mark_Date, format = "%m/%d/%y")) 
+
+Clean_tags_Year_df = Clean_tags %>%
+  select(Site_Name, Mark_Date, Tag_code, Weight, Length, Species, Count, Mark_Date) %>%
+  mutate(Year = year(Mark_Date)) 
+
+unique(Clean_tags_Year_df$Site_Name)
+
+LLR = Clean_tags_Year_df %>%
+  filter(Site_Name == "LLRTP - Lower Lemhi River Rotary Screw Trap")
+
+HYDTRP = Clean_tags_Year_df %>%
+  filter(Site_Name == "HYDTRP - Hayden Creek Rotary Screw Trap" )
+
+PAHTRP = Clean_tags_Year_df %>%
+  filter(Site_Name ==  "PAHTRP - Pahsimeroi River Trap")
+
+SAWTRP = Clean_tags_Year_df %>%
+  filter(Site_Name == "SAWTRP - Sawtooth Trap"  )
+
+NFSTRP = Clean_tags_Year_df %>%
+  filter(Site_Name == "NFSTRP - North Fork Salmon River Rotary Screw Trap")
+
+LEMTRP = Clean_tags_Year_df %>%
+  filter(Site_Name == "LEMTRP - Upper Lemhi River Rotary Screw Trap")
+
+LEMHIW = Clean_tags_Year_df %>%
+  filter(Site_Name == "LEMHIW - Lemhi River Weir")
+
+LEMHIR = Clean_tags_Year_df %>%
+  filter(Site_Name == "LEMHIR - Lemhi River")
+
[…]

### 2021-12-16 analysis/paper/paper.Rmd
minor edits from Google Doc

@@
 Spring/summer Chinook Salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a].
-The DSR migrants leave the natal spawning area as subyearlings between June and november and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts.
+The DSR migrants leave the natal spawning area as subyearlings between June and November and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts.
 Alternatively, NRR migrants remain in their natal spawning areas for approximately one year after emergence until emigration to the ocean as smolts.
@@
 
-## Abundance of Piscivorous Fishes
+## Northern Pikeminnow Abundance
 
@@
 Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017].
-Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon based on an estimate from @Moss2016.
+Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon based on an estimate from @Moss2016 and 3,000 J/g for invertebrates.
+Additional settings included 0% assumed indigestible prety, and a variation in the proportion of diet that was invertebrates or fish ranging from 10% to 90% for both categories.
 We used a weight-length formula [@Parker1995] to calculate the average predator start weight from the average length of Northern Pikeminnow captured during our study using the FSA package in R [@Ogle2021].
@@
 The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed by dividing by a weight of 12 g per individual.
-This weight was an average from juvenile Chinook Salmon emigrating past six rotary screw traps operating in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.
+This weight was an average from juvenile Chinook Salmon captured at six rotary screw traps throughout the operation season March to November in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.
 
@@
 We estimated the population size of Northern Pikeminnow in the Deadwater Slough to be greater than ***xx,xxx*** during the fall emigration period for DSR Chinook Salmon.
-That estimate translates to a density of ***xxx*** Northern Pikeminnow per 100 m or ***xxx*** per 100 m^2^ which is ***similar/more/less*** than estimates from elsewhere in the Columbia River ***(citation)*** where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminno abundance.
-The population size of Northern Pikeminnow was not directly estimated during the spring NRR Chinook salmon emigration period however, the relative abundance measured at CPUE was comparable to the fall sampling periods ***(update statement later)***.
+That estimate translates to a density of ***xxx*** Northern Pikeminnow per 100 m or ***xxx*** per 100 m^2^ which is ***similar/more/less*** than estimates from elsewhere in the Columbia River ***(citation)*** where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminnow abundance.
+The population size of Northern Pikeminnow was not directly estimated during the spring NRR Chinook salmon emigration period; however, the relative abundance measured at CPUE was comparable to the fall sampling periods ***(update statement later)***.
 The population of Northern Pikeminnow in Deadwater Slough was estimated to consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook Salmon during the ***x*** sampling periods and result in an estimated reduction of returning adults between ***xxx*** and ***x,xxx***.
@@
 Additionally, we found that the fish captured in net-traps had a similar proportion of stomach content samples.
-of those, 1,214 (76.2%) were completely empty, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
+Of those, 1,214 (76.2%) were completely empty, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
 We were successfully able to identify Redside Shiners, a Largescale Sucker, sculpin, Mountain Whitefish, and one juvenile Chinook Salmon in the stomach contents.
@@
 
-The authors extent much appreciation to the many volunteers who assisted with field efforts including collaborators from Bureau of Reclamation, Idaho Department of Fish and Game, and Lemhi Regional Land Trust, among others.
+The authors extend much appreciation to the many volunteers who assisted with field efforts including collaborators from Bureau of Reclamation, Idaho Department of Fish and Game, and Lemhi Regional Land Trust, among others.
 This manuscript benefited from reviews and contributions from colleagues at the Idaho Governor's Office of Species Conservation, Rio Applied Science and Engineering, and from Sean Gibbs and Ben Briscoe at Mount Hood Environmental.

### 2021-12-17 analysis/R scripts/Nick Script 12.15.21 bioenergetics.R
Updated Bioenergetics
I added specific weights to each bioenergetics graph. These need help so please use the code in the repo to clean up these graphs.

@@ -192,7 +192,7 @@ spring_week = c("3/1", "3/8", "3/15", "3/22", "3/29", "4/5", "4/19", "4/26", "5/
   #labs(y = "Cumulative fish eatten (g)") +
   scale_x_date(breaks = "1 weeks") +
   scale_y_continuous(name = "Cumulative fish eatten (g)", breaks = seq(0,600, by = 5),
-                     sec.axis = sec_axis(~. /9.5, name = "Potential number of Chinook smolts", breaks = seq(0,30, by = 1))) +
+                     sec.axis = sec_axis(~. /10.9, name = "Potential number of Chinook smolts", breaks = seq(0,30, by = 1))) +
   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
   theme_classic() +
   theme(legend.position = "none") +
@@ -215,7 +215,7 @@ Spring_Graph
     #labs(y = "Cumulative fish eatten (g)") +
     scale_x_date(breaks = "1 weeks") +
     scale_y_continuous(name = "Cumulative fish eatten (g)", breaks = seq(0,600, by = 5),
-                       sec.axis = sec_axis(~. /9.5, name = "Potential number of Chinook smolts", breaks = seq(0,30, by = 1))) +
+                       sec.axis = sec_axis(~. /10.3, name = "Potential number of Chinook smolts", breaks = seq(0,30, by = 1))) +
     theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
     theme_classic() +
     theme(legend.position = "none") +

### 2021-12-17 analysis/R scripts/USR screwtraps script 12.16.21.R

@@ -21,6 +21,22 @@ Clean_tags_Year_df = Clean_tags %>%
   select(Site_Name, Mark_Date, Tag_code, Weight, Length, Species, Count, Mark_Date) %>%
   mutate(Year = year(Mark_Date)) 
 
+Spring_season = Clean_tags_Year_df %>%
+  mutate(Month = month(Mark_Date)) %>%
+  mutate(Day = day(Mark_Date)) %>%
+  filter(Month %in% c(3,4,5))
+
+Fall_season = Clean_tags_Year_df %>%
+  mutate(Month = month(Mark_Date)) %>%
+  mutate(Day = day(Mark_Date)) %>%
+  filter(Month %in% c(9,10,11))
+
+Fall_season$MD = paste(Fall_season$Month, Fall_season$Day)
+
+Fall_season = Fall_season %>%
+  filter(!MD %in% c("9 1","9 2","9 3","9 4","9 5","9 6","9 7","9 8","9 9","9 10","9 11","9 12","9 13","9 14"))
+
+
 unique(Clean_tags_Year_df$Site_Name)
 
 LLR = Clean_tags_Year_df %>%
@@ -57,7 +73,12 @@ Avg_wt_length = Clean_tags_Year_df %>%
   group_by(Site_Name) %>%
   summarise(avg_length = mean(Length), avg_wt = mean(Weight))
 
-Overall_average = Clean_tags_Year_df %>%
+Overall_average_Spring =  Spring_season %>%
+  filter(!Length >300) %>%
+  na.omit(Weight) %>%
+  summarise(avg_length = mean(Length), avg_wt = mean(Weight))
+
+Overall_average_Fall = Fall_season %>%
   filter(!Length >300) %>%
   na.omit(Weight) %>%
   summarise(avg_length = mean(Length), avg_wt = mean(Weight))
@@ -84,3 +105,4 @@ ggplot(PAHTRP) +
   theme_classic() +
   scale_x_date(date_breaks = "1 month") +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+

### 2021-12-17 analysis/R scripts/Nick Script 12.15.21 bioenergetics.R
updates

@@ -228,12 +228,13 @@ Fall_Graph
 
 # Stomach contents and length frequency ############################################################
 
-Clean_NPM$NPM_50=lencat(Clean_NPM$Length, w = 50)
+Clean_NPM$NPM_10=lencat(Clean_NPM$Length, w = 10)
 
 PSD_DF = Clean_NPM %>%
-  group_by(NPM_50) %>%
+  group_by(NPM_10) %>%
   summarise(n = n())
 
+#54.2  S = 250mm P= 380mm
 
 Stomach_catagories_sum = Stomach_content_DF %>%
   drop_na(StomachContentsWeight) %>%

### 2021-12-20 analysis/paper/paper.Rmd
incorporated TBs new Intro

@@
 
+The Snake and Columbia rivers drain a network of inland tributaries that provide essnetial spawning and rearing habitat for anadromous Pacific salmon *Oncorhynchus* spp.
+and steelhead *O.
+mykiss*.
+These fishes were historically abundant throughout the drainage basin and consequently have tremendous ecological, cultural, and economic value.
+Over the last century, salmonid stocks throughout the Snake and Columbia rivers have undergone significant declines related to habitat modification in the freshwater environment (e.g., removal of beavers, river channel simplification, water withdrawals, hydrosystem development) and poor marine conditions, both affecting their survival and recruitment and multiple life stages.
+Consequently, action agencies have made considerable investments in the rehabilitation of tributary ecosystems and evaulation of factors attributed to the species decline [@Roni2018].
+
+One potentially important, but perhaps under-appreciated source of salmonid mortality is predation on emigrating juveniles by piscivorous fishes.
+In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recover [@Fresh2003; @Winther2019].
+Dams and reservoirs in the Columbia River are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995].
+There are generally two mechanisms that explain these high predation zones.
+First, migration (movement) rates of juvenile salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation.
+Second, reservoirs and downstream tailraces associated with dams on the Snake and Columbia rivers create favorable slow-water habitat for predatory fishes known to consume juvenile salmonids, such as the Northern Pikeminnow *Ptychocheilus oregonsis*.
+Northern Pikeminnow tolerate and thrive in relatively warm, slow water habitats [@Wydoski2003] and consequently have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999].
+Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].
+
+Most predation studies have focused on reservoirs [@Murphy2021] and mainstem reaches [@Zimmerman1999; @Tabor1993; @Ward1995; @Shively1996] of the Columbia River basin whereas very little is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems [@Rubenson2020].
+Upstream habitats containing slower water velocities and other attributes that support piscivorous predators may overlap with essential habitat for some salmonid species and life stages.
+One such example is Deadwater Slough, an approximately 1.5 kilometer long reach of unnaturally slow and deep water in the Salmon River, Idaho.
+The Deadwater Slough is located in a river reach containing historically important overwinter rearing and spring emigration habitat for juvenile Chinook Salmon *O.
+tshawytscha* and is inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*.
+Further, the slough is part of the migratory pathway of the endangered Snake River Sockeye Salmon population [@Axel2015] and several populations of threatened Snake River steelhead.
+
+Quantifying mortality associated with piscine predation in upstream habitats of the Salmon River is of particular interest for spring/summer Chinook Salmon in the Upper Salmon major population group (MPG).
+The Salmon River was historically the most productive tributary for spring/summer Chinook Salmon in the Columbia River basin [@Nemeth1999], but those populations have become depleted in recent decades.
+Currently, the Upper salmon MPG supports eight independent, extant spring/summer Chinook Salmon populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River (@NOAA2017).
+The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014].
+Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency.
+Because the Deadwater Slough supports rearing for DSR migrants and is part of the migration corridor for NRR migrants, and juveniles from all Upper Salmon populations must rear in or migrate through it, it represents a significant habitat for the Upper Salmon MPG.
+
+Several studies have attempted to examine the downstream movement, distribution, and (apparent) survival of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach.
+Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have substantially reduced survival in the Deadwater Slough reach compared to adjacent reaches [@Axel2015].
+Similarly, emigrating DSR Chinook Salmon had a 10% reduction in transition probability through the Deadwater Slough compared to surrounding reaches during the fall and early winter months [@Ackerman2018; @Porter2019].
+Low survival in these cases was attributed to low-velocity water delaying movement rates, thereby increasing predation risk.
+
+In this study, we estimated the abundance of Northern Pikeminnow in the Deadwater Slough and its potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
+We hypothesize that increase densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon and Sockeye Salmon.
+To test this, our objectives were three-fold:
+
+1. Estimate the abundance (or relative abundance) of potential predators in the Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations.
+2. Document predation on juvenile Chinook Salmon during the emigration periods using gastric lavage.
+3.
+Use a bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population at Deadwater Slough to assess their potential impacts to DSR and NRR Chinook Salmon emigrants.
+
+<!--
 The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations, each providing important ecological, social, and economic benefits to the region.
@@
 We also consider how consumption of juvenile Chinook Salmon by predators at Deadwater Slough might impact adult returns, hindering recovery of populations in the Upper Salmon River MPG.
+-->
 

### 2021-12-20 analysis/paper/paper.Rmd

@@
 
+## Study Site
+
+The Deadwater Slough is an approximately 1.5-kilometer section of the mainstem Salmon River located roughly 5.8 river kilometers downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
+The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
+Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River [@Emerson1973], thereby creating a slow, deep section in the river, spanning approximately 30 acres and averaging 68 m width.
+
 ## Northern Pikeminnow Abundance

### 2021-12-20 analysis/paper/paper.Rmd
added remaining intro citations

--- 2021-12-20 analysis/paper/paper.Rmd
@@
 
+## Study Site
+
+The Deadwater Slough is an approximately 1.5-kilometer section of the mainstem Salmon River located roughly 5.8 river kilometers downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
+The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
+Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River [@Emerson1973], thereby creating a slow, deep section in the river, spanning approximately 30 acres, and averaging 68 m width.
+
 ## Northern Pikeminnow Abundance

### 2021-12-20 analysis/paper/paper.Rmd
edits to abundance & gastric methods

@@
 One potentially important, but perhaps under-appreciated source of salmonid mortality is predation on emigrating juveniles by piscivorous fishes.
-In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recover [@Fresh2003; @Winther2019].
+In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recover [@Fresh2003; @Winther2020].
 Dams and reservoirs in the Columbia River are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995].
@@
 tshawytscha* and is inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*.
-Further, the slough is part of the migratory pathway of the endangered Snake River Sockeye Salmon population [@Axel2015] and several populations of threatened Snake River steelhead.
+Further, the slough is part of the migratory pathway of the endangered Snake River Sockeye Salmon *O.
+nerka* population [@Axel2015] and several populations of threatened Snake River steelhead.
 
@@
 
-Our first objective was to estimate the population size of piscine predators in the Deadwater Slough using mark-recapture methods.
-We attempted to sample predators near the peaks of the juvenile DSR and NRR emigrations for the 2018 brood year (fall 2019 and spring 2020, respectively).
-However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 effort was moved to fall 2020.
+We estimated the population size of piscine predators in the Deadwater Slough using a mark-recapture survey and a catch per unit effort (CPUE) approach.
+Predators were sampled near the peaks of the juvenile DSR and NRR emigrations.
+Our intent was to sampling during the emigration of the 2018 brood year (fall 2019 and spring 2020).
+However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 effort was delayed until fall 2020.
 Although this provided the benefit of an additional abundance estimate during the DSR emigration, estimates during the NRR emigration were still lacking, and so an additional effort was added in spring 2021.
-Each of the fall efforts included sampling across a two-week span; however, due to budget constraints the spring 2021 effort only included a single week.
-For the fall efforts, the first week was considered a "mark event" and the second week a "recapture event" for single-census mark-recapture estimators (described below); therefore, the "recapture week" occurred the week following the "mark week" to minimize predator emigration/immigration and constrain the population for the mark-recapture models.
+Fall efforts included sampling across a two-week span; however, due to budget constraints the spring 2021 effort was conducted over a single week.
+Sampling was constrained within a two-week window to minimize predator emigration/immigration and to constrain the population for the mark-recapture models.
 
@@
 Methods included raft electrofishing, fyke netting, snorkeling, and angling.
-However, some methods proved ineffective during initial efforts.
-For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) our inability to "chase" fish downstream as they evaded the electricity.
+However, ineffective methods were abandoned in subsequent years.
+After evaluating all methods, angling proved to be the most effective method for capturing piscine predators while minimizing potential impacts to ESA-listed steelhead, but recognizing that angling can have both species and size selectivity (bias).
+From here on, we focus on fish captured by angling, unless otherwise noted.
+Additionally, it quickly became clear during fall 2019 that Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and so is the focus of our study.
+
+<!-- For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) our inability to "chase" fish downstream as they evaded the electricity.
 Snorkeling was ineffective due to the high turbidity of the mainstem Salmon River and Deadwater Slough area.
@@
 Additionally, there were concerns that raft electrofishers and fyke nets could incidentally capture ESA-listed adult steelhead that are present in the mainstem Salmon River during fall and spring.
-After evaluating all methods, angling proved to be the most effective method for capturing piscine predators while minimizing potential impacts to steelhead but recognizing that angling can have both species and size selectivity (bias).
-From here on, we focus on fish captured by angling, unless otherwise noted.
+-->
 
 Our study relied heavily on volunteer anglers who were permitted to fish anywhere within Deadwater Slough.
-During a sample event, anglers would boat or hike their catch (periodically or upon filling a livewell) to a processing station located at the boat ramp approximately 500 m downstream from the top of the slough.
+During a sampling occassion (day), anglers would boat or hike their catch (periodically or upon filling a livewell) to a processing station at the boat ramp approximately 500 m downstream from the top of the slough.
 For each fish, we recorded the date of capture, species, total length (TL; mm), and whether the fish was previously marked.
 Unmarked fish were given a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) unique to each day and released.
+Fish that died prior to release were still counted as captured (or recaptured), but not included in the number of marked fish available for recapture after that date.
 Released individuals were distributed throughout Deadwater Slough to facilitate mixing back into the population.
-Finally, we recorded the angling start and end times for each crew (person or combinations of persons) to facilitate calculations of catch per unit effort (CPUE).
+Finally, we recorded the angling start and end time for each crew (person or combinations of persons) to facilitate calculations of CPUE.
 
@@
 
-All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the sampling period), (2) marking a fish does not affect its chances of recapture, (3) no loss of marks, (4) marks are not missed or mistaken, and (5) all fish have equal chance of being caught in the second (and subsequent) sample(s).
-
-Estimators fell into two broad categories: single census and multiple census.
-For the single census estimators, we treated the first week of sampling as the mark event, and the following (second) week as the recapture event, pooling data within each of those two weeks.
-Alternatively, the multiple census estimators treat each day as a survey, and use information about the total marked fish from all previous surveys to infer the total abundance.
-
-The Lincoln-Peterson estimator, a single-census estimator, is below, where $M$ is the number of fish marked and returned to the population, the $n$ is the number of fish caught in the second/recapture event, and $m$ is the number of marked fish in the second sample.
+Mark-recapture abundance estimators fell into two broad categories: single-census and multiple-census. For the single-census estimators, we used the Lincoln-Peterson estimator:
 
@@
 
-The Lincoln-Petersen estimator can be biased with small samples, so we also employed the Chapman-modified Lincoln-Petersen estimator:
+where $M$ is the number is the total number of fish marked and returned to the population during the first week of sampling (mark event), $n$ is the total number of fish caught in the second week (recapture event), and $m$ is the number of marked fish caught during the recapture event.
+
+Because the Lincoln-Peterson estimator can be biased with small samples, we also employed the Chapman-modified Lincoln-Peterson estimator:
 
@@
 
-For a multiple-census estimator, we first utilized the Schnabel estimator where the $M$, $n$ and $m$ are indexed by the sampling occasion, $i$.
-In our case, the sampling occassions are defined as each day of sampling.
-The Schnabel estimator does not have an associated standard error; however, the 90% confidence intervals can be calculated.
-The Schnabel estimator is essentially a weighted average of a series of Lincoln-Peterson estimators (with a Chapman modfication).
-The Schnabel estimator is shown here:
+For a multiple-census estimator, we first utilized the Schnabel estimator:
 
@@
 
-Finally, we explored one additional estimator for multiple census surveys, called the Schumacher-Eschmeyer estimator, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
+where $M$, $n$, and $m$ are indexed by the sampling occassion, $i$.
+Here, $i$ is defined as each sampling day.
+The Schnabel estimator does not have an associated standard error; however, the 90% confidence intervals were calculated.
+
+Finally, we explored one additional multiple-census estimator, the Schumacher-Eschmeyer estimator, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
 The Schumacher-Eschmeyer estimator is as follows:
@@
 The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and fall 2020 sampling periods.
-However, due to logistical constraints, we were only able to sample during a single week in spring 2021 and released fish were left unmarked.
+However, due to logistical constraints, we were only able to sample during a single week in spring 20201 and released fish were left unmarked.
 Instead, to estimate the abundance of Northern Pikeminnow during spring 2021, we used the ratio of total CPUE in the fall sampling events to the total CPUE in the spring sampling event, and multiplied that ratio by the average abundance of Northern Pikeminnow from the fall sampling events.<!-- Kevin, is this previous statement accurate?
@@
 
+All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the sampling period), (2) marking a fish does not affect its chances of recapture, (3) no loss of marks, (4) marks are not missed or mistaken, and (5) all fish have equal chance of being caught in the second (and subsequent) sample(s).
+
 ```{r psd}
-s_size = 300
-q_size = 400
+s_size = 250
+q_size = 380
 
@@
 
+<!--
 Finally, we calculated the proportional stock density (PSD) for Northern Pikeminnow captured during all field efforts, which is a measure of species size structure.
@@
 where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
-We calculated the PSD for Northern Pikeminnow in our study using `r prettyNum(s_size, digits = 0)` mm TL for stock-size and `r prettyNum(q_size, digits = 0)` mm TL for quality-size.
+We calculated the PSD for Northern Pikeminnow in our study using `r prettyNum(s_size, digits = 0)` mm TL for stock-size and `r prettyNum(q_size, digits = 0)` mm TL for quality-size [@Winther2020].
+-->
+
+<!--
+Juvenile Chinook Salmon abundances were estimated at 5 rotary screw traps in the upper Salmon River basin: Sawtooth, Pahsimeroi, Hayden Creek, Upper Lemhi, and Lower Lemhi.
+Abundances were calculated by Idaho Department of Fish and Game at each trap which made up the production estimate for juvenile Chinook Salmon in the Upper Salmon Basin (Table X) (McClure et al.
+2020).
+-->
 
@@
 For each sample, wet weight (grams) was recorded for the total combined stomach content, including all non-fish items (e.g., macroinvertebrates, organic matter), as well as the portion consisting of fish parts/matter.
-Fish and fish remnants were identified to the lowest taxonomic unit, when possible, or were categorized as unknown.
+Fish and fish remnants were identified to the lowest taxonomic unit using diagnostic bones, or were categorized as unknown [@Hansel1988; @Frost2000].
 A subset of Northern Pikeminnow captures (~ 5%) were euthanized for dissection after gastric lavage to validate the efficacy of the methodology.

### 2021-12-20 analysis/paper/paper.Rmd
minor edits to abundance methods

@@
 Predators were sampled near the peaks of the juvenile DSR and NRR emigrations.
-Our intent was to sampling during the emigration of the 2018 brood year (fall 2019 and spring 2020).
+Our intent was to sample during the emigration of the 2018 brood year (fall 2019 and spring 2020).
 However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 effort was delayed until fall 2020.
@@
 Fall efforts included sampling across a two-week span; however, due to budget constraints the spring 2021 effort was conducted over a single week.
-Sampling was constrained within a two-week window to minimize predator emigration/immigration and to constrain the population for the mark-recapture models.
+Fall efforts were constrained within a two-week window to minimize predator emigration/immigration and to constrain the population for the mark-recapture models.
 
@@
 However, ineffective methods were abandoned in subsequent years.
-After evaluating all methods, angling proved to be the most effective method for capturing piscine predators while minimizing potential impacts to ESA-listed steelhead, but recognizing that angling can have both species and size selectivity (bias).
-From here on, we focus on fish captured by angling, unless otherwise noted.
-Additionally, it quickly became clear during fall 2019 that Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and so is the focus of our study.
-
 <!-- For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) our inability to "chase" fish downstream as they evaded the electricity.
@@
 Additionally, there were concerns that raft electrofishers and fyke nets could incidentally capture ESA-listed adult steelhead that are present in the mainstem Salmon River during fall and spring.
--->
+-->After evaluating all methods, angling proved to be the most effective method for capturing piscine predators while minimizing potential impacts to ESA-listed steelhead.
+<!--, but recognizing that angling can have both species and size selectivity (bias).--> Therefore, the following analyses will focus on fish captured by angling, unless otherwise noted.
+During the fall 2019 effort it additionally became clear that Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and so is the focus of our study.
 
@@
 
-Mark-recapture abundance estimators fell into two broad categories: single-census and multiple-census. For the single-census estimators, we used the Lincoln-Peterson estimator:
+Four different mark-recapture abundance estimators were employed to explore variation in estimates.
+Estimators fell into two broad categories: single-census and multiple-census.
+For the single-census estimators, we used the Lincoln-Peterson estimator:
 
@@
 $$
-
 where $M$ is the number is the total number of fish marked and returned to the population during the first week of sampling (mark event), $n$ is the total number of fish caught in the second week (recapture event), and $m$ is the number of marked fish caught during the recapture event.
-
 Because the Lincoln-Peterson estimator can be biased with small samples, we also employed the Chapman-modified Lincoln-Peterson estimator:
@@
 $$
-
 For a multiple-census estimator, we first utilized the Schnabel estimator:
@@
 The Schnabel estimator does not have an associated standard error; however, the 90% confidence intervals were calculated.
-
-Finally, we explored one additional multiple-census estimator, the Schumacher-Eschmeyer estimator, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
+Lastly, we explored one additional multiple-census estimator, the Schumacher-Eschmeyer, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
 The Schumacher-Eschmeyer estimator is as follows:
@@
 The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and fall 2020 sampling periods.
-However, due to logistical constraints, we were only able to sample during a single week in spring 20201 and released fish were left unmarked.
-Instead, to estimate the abundance of Northern Pikeminnow during spring 2021, we used the ratio of total CPUE in the fall sampling events to the total CPUE in the spring sampling event, and multiplied that ratio by the average abundance of Northern Pikeminnow from the fall sampling events.<!-- Kevin, is this previous statement accurate?
+However, to estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of total CPUE in the fall sampling events to the total CPUE in the spring sampling event, then multiplied that ratio by the average abundance from the fall sampling events.<!-- Kevin, is this previous statement accurate?
 --> This approach assumes equal capture probabilities between the fall and spring capture events.
-If that assumption is not true and capture probability did differ, we believe that the capture probability  in the spring would be lower due to higher flows, which would lead to our spring abundance estimate to be biased low.
-Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 sampling event to be a conservative estimate.
+<!--If that assumption is not true and capture probability did differ, we believe that the capture probability  in the spring would be lower due to higher flows, which would lead to our spring abundance estimate to be biased low.
+Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 sampling event to be a conservative estimate.-->
 

### 2021-12-20 analysis/paper/paper.Rmd
bioenergetic methods edits

@@
 
-To estimate the total consumption potential (i.e., number/amount of juvenile consumed during the peaks of DSR [fall] and NRR [spring] emigrations) of Northern Pikeminnow in Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 applied in the R statistical software [@RCoreTeam2021].
-The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on predator and prey energy densities, predator start and end weights, and water temperatures.
+To estimate the total consumption potential (i.e., the number or amount of juvenile Chinook Salmon consumed during the peaks of DSR [fall] and NRR [spring] emigrations) of Northern Pikeminnow in Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 and applied in the R statistical software [@RCoreTeam2021].
+The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on the following inputs: predator and prey energy densities, predator start and end weights, and water temperatures.
 Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017].
-Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon based on an estimate from @Moss2016 and 3,000 J/g for invertebrates.
-Additional settings included 0% assumed indigestible prety, and a variation in the proportion of diet that was invertebrates or fish ranging from 10% to 90% for both categories.
-We used a weight-length formula [@Parker1995] to calculate the average predator start weight from the average length of Northern Pikeminnow captured during our study using the FSA package in R [@Ogle2021].
-The average length of Northern Pikeminnow captured at the Deadwater Slough during the spring and fall were 352.9 and 393.7 mm, respectively, which calculates to average starting weights of 504.4 and 670.5 g.
-Continuous water temperature data from 2019 was used in two alternative models, one using the baseline start and end weights and the second with a 10% increase in average end weight on Northern Pikeminnow.
-
-We ran three separate bioenergetic models for this study.
-For the first model, we used temperature data from a 78-day period spanning September 1 through November 17 when DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration.<!-- Did we just use data from Lemhi here?
-All tributaries?
-Was this subjective?
---> During this time, water temperatures typically exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon<!--citation?-->.
-This first model assumed no growth for Northern Pikeminnow.
-The second model was instead run for one full year, but also assumed no growth for Northern Pikeminnow.
-The assumption of zero growth was included to estimate consumption if individual Northern Pikeminnow were not growing and/or the population is stable.
-Finally, the third model shows how an increase of 10% body weight or an increase in population size changes the estimates of consumption over a full year period.
-
-To estimate the total biomass consumed, we then multiplied the biomass consumed by an individual Northern Pikeminnow for the given time period by the estimated population size for that time period.
-The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed by dividing by a weight of 12 g per individual.
-This weight was an average from juvenile Chinook Salmon captured at six rotary screw traps throughout the operation season March to November in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.
-
-<!--
-A few things that need to be clarified in this section:
-* Why were the dates Sept 1 - Nov 17 chosen? They make sense, but we should include our rationale
-* Where did the temperature data come from? Hobo? Stream gauge? The methods say temperature data from 2019 was used, but the results seem to use 2020? 
-* Why does it look like we only ran it for dates in 2020? Is that the only year we had temperature data for? Would the results look different if we used temperature data from a different year?
-* The lengths of pikeminnow: did we only use pikeminnow we caught in the fall to get average length? Is there a difference in lengths between 2019 and 2020? 
-* There needs to be more explanation of the inverts vs. fish. What proportion went to each and why? Did it change through time? Based on literature at all? 
-* Where is the data that was used to get a weight of 12 g for Chinook emigrant?
-Was that for Chinook presmolts, or a combination of smolts and presmolts?
-If the latter, can we filter it to look at the average weight for presmolts only?
-And maybe smolts only for a different analysis?
-* Could we run a similar model for the spring emigration period?
-Figure out the dates, look at lengths of pikeminnow from the spring, convert to average weight, use average size of Chinook smolts (not presmolts) from all the Upper Salmon traps?
-Kevin can work on converting the spring catch into some kind of estimate of pikeminnow at Deadwater in the spring
--->
+Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon [@Moss2016].
+We used a weight-length formula [@Parker1995] to calculate the predator start weight using the average length of Northern Pikeminnow captured in Deadwater Slough during our study.
+The average length of Northern Pikeminnow during the fall and spring were 352.9 and 393.7 mm, respectively, which calculated to average starting weights of 504.4 and 670.5 g.
+The assumption of zero growth was included to estimate consumption if individual Northern Pikeminnow were not growing and/or the population is stable, and so start and end weights are equal.
+Mean daily water temperatures were summarized from 15-minute interval temperature readings available March 3, 2013 to June 14, 2021 from a gage station approximately 22 river kilometers downstream of Deadwater Slough near Shoup, Idaho [@USGS2016].
+
+Our largest uncertainty for model parameters was the proportion of the Northern Pikeminnow diet consisting of fishes (e.g., juvenile Chinook Salmon) versus other food items like invertebrates.
+Therefore, we varied the proportion of the diet that was fish or invertebrates ranging from 10-90%, in 10% intervals, for both categories.
+Energy densities for invertebrates was fixed at 3,000 J/g.
+Separate models were also run for a 77-day fall period from September 15 - November 30 and a 92-day period in the spring from March 1 - May 31 to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG.
+This resulted in 18 separate model runs (2 seasons and 9 diet scenarios).
+Each model run provided an estimate of the amount (grams) of fish consumed by an individual Northern Pikeminnow during the fall or spring time periods.
+
+<!-- During this time, water temperatures typically exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon. -->
+
+To estimate the total biomass of fish potentially consumed by Northern Pikeminnow in Deadwater Slough, we then multiplied the biomass consumed by an individual Northern Pikeminnow for the time period by the estimated population size for that time period.
+<!-- from which estimator?
+-->The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed by dividing by a weight of 10.9 g for the spring and 10.3 g for the fall, an average from juvenile Chinook Salmon captured during those date ranges at seven rotary screw traps located throughout the Upper Salmon upstream of Deadwater Slough.
+
+<!-- Sawtooth Hatchery, East Fork Salmon, Pahsimeroi, Upper Lemhi, Hayden Creek, Lower Lemhi, North Fork Salmon -->
 

### 2021-12-21 analysis/R scripts/Nick Script 11.10.21 cleaning fish data.R
updated README, cleaned up folder & file structure

@@ -1,282 +1,0 @@
-# working to clean up and provide summaries for Deadwater Data
-# Nick Porter
-# 11/2/20201
-rm(list = ls()) # clear the working environment
-
-require(tidyverse)
-require(lubridate)
-require(boot)
-require(bootstrap)
-require(lme4)
-require(Matrix)
-require(numDeriv)
-require(FSA)
-require(FSAdata)
-require(magrittr)
-require(car)
-require(dplyr)
-require(plotrix)
-require(nnet)
-#require(Hmisc) # does not work
-#require(Matching)
-require(Rcapture)
-require(nlstools)
-require(stringi)
-require(MASS)
-require(readr)
-
-#Read in the Deadwater dataset
-deadwater <- read_csv("analysis/data/raw_data/deadwater_cmr_effort_20210614.csv") %>%
-  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
-
-
-NPM = deadwater %>%
-  filter(Species == "Northern Pikeminnow")
-
-Clean_NPM = NPM %>%
-  mutate(Fish_or_Parts = case_when(stri_detect_fixed(StomachContents, "Fish") ~ "Fish or Fish Parts",
-                                   stri_detect_fixed(StomachContents, "Empty") ~ "Empty",
-                                   stri_detect_fixed(Comments, "Whole scuplin") ~ "Fish or Fish Parts",
-                                   stri_detect_fixed(StomachContents, "Whitefish") ~ "Fish or Fish Parts",
-                                   TRUE ~ "Other")) %>%
-  drop_na(Length) %>%
-  mutate(Non_fish_wt = StomachContentsWeight - FishContentWeight)
-
-Stomach_content_DF = Clean_NPM %>%
-  dplyr::select(Date, Length, LavageID, StomachContents, Comments,
-                Fish_or_Parts, StomachContentsWeight, Non_fish_wt, FishContentWeight) %>%
-  filter(!StomachContents %in% c("Lost sample CG27-001","Did Not Lavage")) %>%
-  drop_na(StomachContents)
-
-
-Spring_Avg_length = Clean_NPM %>%
-  filter(Date > "2021-05-17") %>%
-  summarise(avg_TL = mean(Length))
-
-Fall_Avg_length = Clean_NPM %>%
-  filter(Date < "2021-05-17") %>%
[…]

### 2021-12-21 analysis/R scripts/Nick Script 12.15.21 bioenergetics.R

@@ -1,341 +1,0 @@
-# working to clean up and provide summaries for Deadwater Data
-# Nick Porter
-# 12/15/20201
-rm(list = ls()) # clear the working environment
-
-require(tidyverse)
-require(lubridate)
-require(boot)
-require(bootstrap)
-require(lme4)
-require(Matrix)
-require(numDeriv)
-require(FSA)
-require(FSAdata)
-require(magrittr)
-require(car)
-require(dplyr)
-require(plotrix)
-require(nnet)
-#require(Hmisc) # does not work
-#require(Matching)
-require(Rcapture)
-require(nlstools)
-require(stringi)
-require(MASS)
-require(readr)
-
-#Read in the Deadwater dataset
-deadwater <- read_csv("analysis/data/raw_data/deadwater_cmr_effort_20210614.csv") %>%
-  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
-
-
-NPM = deadwater %>%
-  filter(Species == "Northern Pikeminnow")
-
-Clean_NPM = NPM %>%
-  mutate(Fish_or_Parts = case_when(stri_detect_fixed(StomachContents, "Fish") ~ "Fish or Fish Parts",
-                                   stri_detect_fixed(StomachContents, "Empty") ~ "Empty",
-                                   stri_detect_fixed(Comments, "Whole scuplin") ~ "Fish or Fish Parts",
-                                   stri_detect_fixed(StomachContents, "Whitefish") ~ "Fish or Fish Parts",
-                                   TRUE ~ "Other")) %>%
-  drop_na(Length) %>%
-  mutate(Non_fish_wt = StomachContentsWeight - FishContentWeight)
-
-Stomach_content_DF = Clean_NPM %>%
-  dplyr::select(Date, Length, LavageID, StomachContents, Comments,
-                Fish_or_Parts, StomachContentsWeight, Non_fish_wt, FishContentWeight) %>%
-  filter(!StomachContents %in% c("Lost sample CG27-001","Did Not Lavage")) %>%
-  drop_na(StomachContents)
-
-
-Spring_Avg_length = Clean_NPM %>%
-  filter(Date > "2021-05-17") %>%
[…]

### 2021-12-21 analysis/R scripts/abund_est.R

@@ -1,311 +1,0 @@
-# Author: Kevin See
-# Purpose: import data from Deadwater sampling
-# Created: 11/9/21
-# Last Modified: 12/2/21
-# Notes:
-
-#-----------------------------------------------------------------
-# load needed libraries
-library(tidyverse)
-library(janitor)
-library(lubridate)
-library(FSA)
-library(here)
-
-theme_set(theme_bw())
-
-#-----------------------------------------------------------------
-# read in data
-#-----------------------------------------------------------------
-raw_df = readxl::read_excel(here("analysis/data/raw_data/Deadwater_predation_mark_recapture_11.09.21.xlsx")) %>%
-  clean_names() %>%
-  mutate(across(date,
-                ymd)) %>%
-  mutate(across(c(start_time, end_time),
-                ~ ymd_hms(paste(year(date), month(date), day(date), hour(.), minute(.), second(.))))) %>%
-  # split into sampling periods
-  mutate(event_name = if_else(month(date) > 8,
-                              paste("Fall", year(date), sep = "_"),
-                              paste("Spring", year(date), sep = "_")))
-
-# focus on pikeminnow
-cmr_df = raw_df %>%
-  filter(count > 0) %>%
-  # pull out only Northern Pikeminnow
-  filter(species == "Northern Pikeminnow") %>%
-  mutate(mark_release = if_else(fish_status == "Live" &
-                                  capture_type == "Mark",
-                                T, F),
-         recap = if_else(capture_type == "Recapture",
-                         T, F))
-
-
-#------------------------------------------------------------
-# CPUE
-#------------------------------------------------------------
-# pull out and summarize effort
-# account for time and number of crew members
-effort_df = raw_df %>%
-  select(event_name,
-         site,
-         date,
-         crew,
-         crew_count,
-         method,
-         equipment,
-         start_time:total_effort,
-         total_day_effort) %>%
-  distinct() %>%
-  mutate(across(total_effort,
-                ~ . * crew_count)) %>%
-  group_by(event_name,
-           date) %>%
-  summarise(across(c(total_effort,
-                     crew_count),
-                   sum),
-            .groups = "drop")
-
-cpue_df = effort_df %>%
-  # add how many pikeminnow were caught
-  left_join(cmr_df %>%
-              group_by(event_name, species, date) %>%
-              summarise(n_fish = n(),
[…]

### 2021-12-21 analysis/supplementary-materials/scripts/Nick Script 11.10.21 cleaning fish data.R

@@ -1,0 +1,282 @@
+# working to clean up and provide summaries for Deadwater Data
+# Nick Porter
+# 11/2/20201
+rm(list = ls()) # clear the working environment
+
+require(tidyverse)
+require(lubridate)
+require(boot)
+require(bootstrap)
+require(lme4)
+require(Matrix)
+require(numDeriv)
+require(FSA)
+require(FSAdata)
+require(magrittr)
+require(car)
+require(dplyr)
+require(plotrix)
+require(nnet)
+#require(Hmisc) # does not work
+#require(Matching)
+require(Rcapture)
+require(nlstools)
+require(stringi)
+require(MASS)
+require(readr)
+
+#Read in the Deadwater dataset
+deadwater <- read_csv("analysis/data/raw_data/deadwater_cmr_effort_20210614.csv") %>%
+  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
+
+
+NPM = deadwater %>%
+  filter(Species == "Northern Pikeminnow")
+
+Clean_NPM = NPM %>%
+  mutate(Fish_or_Parts = case_when(stri_detect_fixed(StomachContents, "Fish") ~ "Fish or Fish Parts",
+                                   stri_detect_fixed(StomachContents, "Empty") ~ "Empty",
+                                   stri_detect_fixed(Comments, "Whole scuplin") ~ "Fish or Fish Parts",
+                                   stri_detect_fixed(StomachContents, "Whitefish") ~ "Fish or Fish Parts",
+                                   TRUE ~ "Other")) %>%
+  drop_na(Length) %>%
+  mutate(Non_fish_wt = StomachContentsWeight - FishContentWeight)
+
+Stomach_content_DF = Clean_NPM %>%
+  dplyr::select(Date, Length, LavageID, StomachContents, Comments,
+                Fish_or_Parts, StomachContentsWeight, Non_fish_wt, FishContentWeight) %>%
+  filter(!StomachContents %in% c("Lost sample CG27-001","Did Not Lavage")) %>%
+  drop_na(StomachContents)
+
+
+Spring_Avg_length = Clean_NPM %>%
+  filter(Date > "2021-05-17") %>%
+  summarise(avg_TL = mean(Length))
+
+Fall_Avg_length = Clean_NPM %>%
+  filter(Date < "2021-05-17") %>%
[…]

### 2021-12-21 analysis/supplementary-materials/scripts/Nick Script 11.2.21 cleaning fish data.R

@@ -1,0 +1,232 @@
+# working to clean up and provide summaries for Deadwater Data
+# Nick Porter
+# 11/2/20201
+rm(list = ls()) # clear the working environment
+
+require(tidyverse)
+require(lubridate)
+require(boot)
+require(bootstrap)
+require(lme4)
+require(Matrix)
+require(numDeriv)
+require(FSA)
+require(FSAdata)
+require(magrittr)
+require(car)
+require(dplyr)
+require(plotrix)
+require(nnet)
+#require(Hmisc) # does not work
+#require(Matching)
+require(Rcapture)
+require(nlstools)
+require(stringi)
+require(MASS)
+require(readr)
+
+#Read in the Deadwater dataset
+deadwater <- read_csv("analysis/data/raw data/deadwater_cmr_effort_20210614.csv") #%>%
+  mutate(Date = as.POSIXct(Date, format = "%m/%d/%y"))
+
+NPM = deadwater %>%
+  filter(Species == "Northern Pikeminnow")
+
+Clean_NPM = NPM %>%
+  mutate(Fish_or_Parts = case_when(stri_detect_fixed(StomachContents, "Fish") ~ "Fish or Fish Parts",
+                                   stri_detect_fixed(StomachContents, "Empty") ~ "Empty",
+                                   stri_detect_fixed(Comments, "Whole scuplin") ~ "Fish or Fish Parts",
+                                   TRUE ~ "Other")) %>%
+  drop_na(Length) %>%
+  mutate(Non_fish_wt = StomachContentsWeight - FishContentWeight)
+
+Stomach_content_DF = Clean_NPM %>%
+  dplyr::select(Date, Length, LavageID, StomachContents, Comments, Fish_or_Parts, StomachContentsWeight, Non_fish_wt, FishContentWeight)
+
+unique(Clean_NPM)
+
+#Read in the Bioenergetics
+
+bioenergetics<- read_csv("analysis/data/raw data/2020 Bioenergetics 1 year.csv") %>%
+  mutate(Date = as.POSIXct(Date, format = "%m/%d/%y")) %>%
+  mutate(Cumu_fish_eatten = cumsum(Cons_fish_g)) %>%
+  mutate(Date = as.Date(Date)) %>%
+  mutate(month = month(Date))
+
[…]

### 2021-12-21 analysis/supplementary-materials/scripts/USR screwtraps script 12.16.21.R

@@ -1,0 +1,108 @@
+rm(list = ls())
+
+require(tidyverse)
+require(lubridate)
+require(esquisse)
+
+tags<- Tagging_Detail_Bioenergetics_2020_avg_weight_ <- read_csv("Merck Sharp & Dohme, Corp/Biomark Field Sampling - Documents/General/Deadwater/Tagging Detail (Bioenergetics).csv")
+glimpse(tags)
+
+Clean_tags = tags %>%
+  rename(Site_Name =`Release Site Name`) %>%
+  rename(Mark_Date = `Mark Date MMDDYYYY`) %>%
+  rename(Tag_code = `Tag Code`) %>%
+  rename(Weight = `Weight g`) %>%
+  rename(Length = `Length mm`) %>%
+  rename(Species = `Species Name`) %>%
+  rename(Count = `Mark Count`) %>%
+  mutate(Mark_Date = as.Date(Mark_Date, format = "%m/%d/%y")) 
+
+Clean_tags_Year_df = Clean_tags %>%
+  select(Site_Name, Mark_Date, Tag_code, Weight, Length, Species, Count, Mark_Date) %>%
+  mutate(Year = year(Mark_Date)) 
+
+Spring_season = Clean_tags_Year_df %>%
+  mutate(Month = month(Mark_Date)) %>%
+  mutate(Day = day(Mark_Date)) %>%
+  filter(Month %in% c(3,4,5))
+
+Fall_season = Clean_tags_Year_df %>%
+  mutate(Month = month(Mark_Date)) %>%
+  mutate(Day = day(Mark_Date)) %>%
+  filter(Month %in% c(9,10,11))
+
+Fall_season$MD = paste(Fall_season$Month, Fall_season$Day)
+
+Fall_season = Fall_season %>%
+  filter(!MD %in% c("9 1","9 2","9 3","9 4","9 5","9 6","9 7","9 8","9 9","9 10","9 11","9 12","9 13","9 14"))
+
+
+unique(Clean_tags_Year_df$Site_Name)
+
+LLR = Clean_tags_Year_df %>%
+  filter(Site_Name == "LLRTP - Lower Lemhi River Rotary Screw Trap")
+
+HYDTRP = Clean_tags_Year_df %>%
+  filter(Site_Name == "HYDTRP - Hayden Creek Rotary Screw Trap" )
+
+PAHTRP = Clean_tags_Year_df %>%
[…]

### 2021-12-21 analysis/supplementary-materials/scripts/abund_est.R

@@ -1,0 +1,311 @@
+# Author: Kevin See
+# Purpose: import data from Deadwater sampling
+# Created: 11/9/21
+# Last Modified: 12/2/21
+# Notes:
+
+#-----------------------------------------------------------------
+# load needed libraries
+library(tidyverse)
+library(janitor)
+library(lubridate)
+library(FSA)
+library(here)
+
+theme_set(theme_bw())
+
+#-----------------------------------------------------------------
+# read in data
+#-----------------------------------------------------------------
+raw_df = readxl::read_excel(here("analysis/data/raw_data/Deadwater_predation_mark_recapture_11.09.21.xlsx")) %>%
+  clean_names() %>%
+  mutate(across(date,
+                ymd)) %>%
+  mutate(across(c(start_time, end_time),
+                ~ ymd_hms(paste(year(date), month(date), day(date), hour(.), minute(.), second(.))))) %>%
+  # split into sampling periods
+  mutate(event_name = if_else(month(date) > 8,
+                              paste("Fall", year(date), sep = "_"),
+                              paste("Spring", year(date), sep = "_")))
+
+# focus on pikeminnow
+cmr_df = raw_df %>%
+  filter(count > 0) %>%
+  # pull out only Northern Pikeminnow
+  filter(species == "Northern Pikeminnow") %>%
+  mutate(mark_release = if_else(fish_status == "Live" &
+                                  capture_type == "Mark",
+                                T, F),
+         recap = if_else(capture_type == "Recapture",
+                         T, F))
+
+
+#------------------------------------------------------------
+# CPUE
+#------------------------------------------------------------
+# pull out and summarize effort
+# account for time and number of crew members
+effort_df = raw_df %>%
+  select(event_name,
+         site,
+         date,
+         crew,
+         crew_count,
+         method,
+         equipment,
+         start_time:total_effort,
+         total_day_effort) %>%
+  distinct() %>%
+  mutate(across(total_effort,
+                ~ . * crew_count)) %>%
+  group_by(event_name,
+           date) %>%
+  summarise(across(c(total_effort,
+                     crew_count),
+                   sum),
+            .groups = "drop")
+
+cpue_df = effort_df %>%
+  # add how many pikeminnow were caught
+  left_join(cmr_df %>%
+              group_by(event_name, species, date) %>%
+              summarise(n_fish = n(),
[…]

### 2021-12-21 analysis/paper/paper.Rmd
compiling bioenergetics results

@@
 
-```{r shoup-temp-data}
+```{r shoup-temp-data, eval = F}
 shoup_temp_df = read_tsv(here("analysis/data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
@@
 ```{r read-bio-outputs}
-# full year, no growth
-full_yr_no_growth = read_csv(here("analysis/data/raw_data/2020 Bioenergetics 1 year.csv")) %>%
+fmt_bio_out = function(x) {
+  x = x %>%
     clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
+    mutate(date = as.Date(date, "%m/%d")) %>%
     mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>% 
     mutate(month = month(date))
-
-# full year, 10% growth
-full_yr_10p_growth = read_csv(here("analysis/data/raw_data/2020 Bioenergetics 1 year 10% growth.csv")) %>%
-  clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
-  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
-  mutate(month = month(date))
-
-# an alternative model
-alternative_mod = read_csv(here("analysis/data/raw_data/Bioenergetics 32 fish 68 inverts.csv")) %>%
-  clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
-  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
-  mutate(month = month(date))
-
-# Fall emigration period (9/1 - 11/17), no growth
-fall_only_no_growth = read_csv(here("analysis/data/raw_data/78_day_bioenergetics.csv")) %>%
-  clean_names() %>%
-  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
-  mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>%
-  mutate(month = month(date))
+  return(x)
+}
+
+# fall, fish 10% - 90%
+fall_f10 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I90_F10.csv")) %>% fmt_bio_out()
+fall_f20 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I80_F20.csv")) %>% fmt_bio_out()
+fall_f30 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I70_F30.csv")) %>% fmt_bio_out()
+fall_f40 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I60_F40.csv")) %>% fmt_bio_out()
+fall_f50 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I50_F50.csv")) %>% fmt_bio_out()
+fall_f60 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I40_F60.csv")) %>% fmt_bio_out()
+fall_f70 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I30_F70.csv")) %>% fmt_bio_out()
+fall_f80 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I20_F80.csv")) %>% fmt_bio_out()
+fall_f90 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I10_F90.csv")) %>% fmt_bio_out()
+
+# spring, fish 10% - 90%
+spring_f10 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I90_F10.csv")) %>% fmt_bio_out()
+spring_f20 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I80_F20.csv")) %>% fmt_bio_out()
+spring_f30 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I70_F30.csv")) %>% fmt_bio_out()
+spring_f40 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I60_F40.csv")) %>% fmt_bio_out()
+spring_f50 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I50_F50.csv")) %>% fmt_bio_out()
+spring_f60 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I40_F60.csv")) %>% fmt_bio_out()
+spring_f70 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I30_F70.csv")) %>% fmt_bio_out()
+spring_f80 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I20_F80.csv")) %>% fmt_bio_out()
+spring_f90 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I10_F90.csv")) %>% fmt_bio_out()
+
+# average size of DSR and NRR juveniles
+avg_DSR_g = 10.3
+avg_NRR_g = 10.9
+
+# compile fall results
+fall_results = fall_f10 %>%
+  select(date, temperature_c, cum_cons_fish_g) %>%
+  rename(f10_g = cum_cons_fish_g) %>%
+  mutate(f20_g = fall_f20$cum_cons_fish_g,
+         f30_g = fall_f30$cum_cons_fish_g,
+         f40_g = fall_f40$cum_cons_fish_g,
+         f50_g = fall_f50$cum_cons_fish_g,
+         f60_g = fall_f60$cum_cons_fish_g,
+         f70_g = fall_f70$cum_cons_fish_g,
+         f80_g = fall_f80$cum_cons_fish_g,
+         f90_g = fall_f90$cum_cons_fish_g) %>%
+  mutate( across(contains('_g'),
+                 .fns = list(n = ~./avg_DSR_g),
+                 .names = "{col}_{fn}") )
+
+rm(fall_f10, fall_f20, fall_f30, fall_f40, fall_f50, fall_f60, fall_f70, fall_f80, fall_f90)
+
+# compile spring results
+spring_results = spring_f10 %>%
+  select(date, temperature_c, cum_cons_fish_g) %>%
+  rename(f10_g = cum_cons_fish_g) %>%
+  mutate(f20_g = spring_f20$cum_cons_fish_g,
+         f30_g = spring_f30$cum_cons_fish_g,
+         f40_g = spring_f40$cum_cons_fish_g,
+         f50_g = spring_f50$cum_cons_fish_g,
+         f60_g = spring_f60$cum_cons_fish_g,
+         f70_g = spring_f70$cum_cons_fish_g,
+         f80_g = spring_f80$cum_cons_fish_g,
+         f90_g = spring_f90$cum_cons_fish_g) %>%
+  mutate( across(contains('_g'),
+                 .fns = list(n = ~./avg_NRR_g),
+                 .names = "{col}_{fn}") )
+
+rm(spring_f10, spring_f20, spring_f30, spring_f40, spring_f50, spring_f60, spring_f70, spring_f80, spring_f90)
 

### 2021-12-21 analysis/paper/paper.Rmd
incorporated edits from Google Doc

@@
 --> This approach assumes equal capture probabilities between the fall and spring capture events.
-<!--If that assumption is not true and capture probability did differ, we believe that the capture probability  in the spring would be lower due to higher flows, which would lead to our spring abundance estimate to be biased low.
-Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 sampling event to be a conservative estimate.-->
 
@@
 We calculated the PSD for Northern Pikeminnow in our study using `r prettyNum(s_size, digits = 0)` mm TL for stock-size and `r prettyNum(q_size, digits = 0)` mm TL for quality-size [@Winther2020].
--->
-
-<!--
-Juvenile Chinook Salmon abundances were estimated at 5 rotary screw traps in the upper Salmon River basin: Sawtooth, Pahsimeroi, Hayden Creek, Upper Lemhi, and Lower Lemhi.
-Abundances were calculated by Idaho Department of Fish and Game at each trap which made up the production estimate for juvenile Chinook Salmon in the Upper Salmon Basin (Table X) (McClure et al.
-2020).
 -->
@@
 
+<!--
+## Juvenile Chinook Salmon Abundance
+
+Juvenile Chinook Salmon abundances were estimated at 5 rotary screw traps in the upper Salmon River basin: Sawtooth, Pahsimeroi, Hayden Creek, Upper Lemhi, and Lower Lemhi.
+Abundances were calculated by Idaho Department of Fish and Game at each trap which made up the production estimate for juvenile Chinook Salmon in the Upper Salmon Basin (Table X) (McClure et al.
+2020).
+-->
+
 # Results
@@
 
-The PSD for Northern Pikeminnow in our study and across all three years was `r prettyNum(psd, digits = 1)`%.
-In other words, `r prettyNum(psd, digits = 1)`% of captured fish were over the quality size classification of `r prettyNum(q_size, digits = 0)` mm TL, demonstrating that a high percentage of Northern Pikeminnow within Deadwater Slough are above average size classes defined for the species
-
 ## Gastric Lavage
@@
 We examined the stomach contents of `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow over the course of the study using gastric lavage methods.
-Of those, we documented some form of stomach contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` of fish of which we were able to confirm fish or fish remnants in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")`.
+Of those, we documented some form of stomach contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` of the fish and we were able to confirm fish or fish remnants in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")`.
 Expressed as a percentage, `r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% of lavaged Northern Pikeminnow had stomach contents of which `r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% we identified containing fish or fish remnants.
@@
 For Northern Pikeminnow with some form of stomach contents, the average wet weight of total contents was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r prettyNum(min(stmch_df$stomach_contents_weight), digits = 3)`-`r prettyNum(max(stmch_df$stomach_contents_weight), digits = 3)` g).
-Overall, fish or fish remnants were `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall weighted weight among all stomach contents examined.
+Overall, fish or fish remnants were `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall weight among all stomach contents examined.
 
@@
 
-We estimated a single average-sized Northern Pikeminnow to consume 61 g of ***fish*** during the modeled fall time period (September 1 - November 17).
-Given an estimated population size of ***xx,xxx***, we'd estimate that ***xxx,xxx*** g of fish are consumed each fall during the emigration of juvenile DSR Chinook Salmon.
-The average weight of a DSR Chinook Salmon is ***xx.x*** taken from ***yyy*** rotary screw traps
-
-If we further assume an average weight of ***xx g*** for DSR Chinook Salmon
-
-Over the fall time period a single average-sized Northern Pikeminnow is modeled to eat 61 g of fish.
-For the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook Salmon.
-Therefore, Northern Pikeminnow in the Deadwater Slough can potentially consume ***xx,xxx*** juvenile Chinook Salmon as a discrete population.
-
-Over an entire year, we estimate that a single Northern Pikeminnow will eat 283 g of fish and 613 g of other food types.
+We estimated a single average-sized Northern Pikeminnow to consume xx.x g of fish during the modeled fall time period (September 15 - November 30) and an additional xx.x g of fish during the modeled spring time period (March 1 - June 31).
+Given estimated population sizes of xx,xxx Northern Pikeminnow during fall and xx,xxx during spring, that would convert to xxx,xxx - xxx,xxx g during fall and xxx,xxx g during spring.
+Given average weights of 10.3 g for DSR Chinook salmon and 10.9 g for NRR Chinook salmon, we would estimate that xx,xxx and xx,xxx juvenile Chinook salmon could be consumed during the fall and spring emigrations in Deadwater Slough, respectively.
 
 <!--
-* Water Temperature measurements.
-    *	Hobo tidbit placed at site. 
-    *	Calculated daily average water temperature for 365 days of the year.
-*	Water temperature plot for the 365 average.
-*	Bioenergetics results. Total mass consumed over a 365 day period. Plot of daily consumption with day of year on x and mass on y.
+For context, throughout the migration periods of DSR and NRR life histories it is estimated that 113,976 individuals are produced and therefore available within the upper Salmon River basin (Table x).
+
+Given an estimated population size of xx,xxx, we’d estimate that xxx,xxx g of NRR fish are consumed each spring during the emigration, and xxxxx.xx g of DSR fish are consumed in the fall during the emigration of juvenile Chinook Salmon.
+Therefore, Northern Pikeminnow in the Deadwater Slough can potentially consume xx,xxx juvenile Chinook Salmon as a discrete population.
+-->
+
+<!--
+## Juvenile Chinook Salmon Abundance
+
+The abundance of juvenile Chinook Salmon for the upper Salmon River basin was estimated at 6 rotary screw traps [@McClure2020].
+Overall, in 2020, there were an estimated 59,289 DSR, 22,264 NRR, and 10,159 parr estimated among all these traps in the upper Salmon River basin (Table X.)
 -->
@@
 
+<!--
+The PSD for Northern Pikeminnow in our study and across all three years was `r prettyNum(psd, digits = 1)`% i.e., `r prettyNum(psd, digits = 1)` of captured fish were over the quality size classification of `r prettyNum(q_size, digits = 0)` mm TL defined by @Winther2020.
+For comparison, @Winther2020 reported a PSD for Northern Pikeminnow captured below Bonneville Dam of 41% and 18% in Bonneville Reservoir, suggesting that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to population reported elsewhere in the Columbia River.
+-->
+
 ## Mark-Recapture Model
@@
 -->
+
+<!--If that assumption is not true and capture probability did differ, we believe that the capture probability  in the spring would be lower due to higher flows, which would lead to our spring abundance estimate to be biased low.
+Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 sampling event to be a conservative estimate.-->
 
@@
 
-```{r}
+```{r, eval = F}
 # pull in SAR numbers
@@
 gastric_tbl %>%
+  rename(Effort = event_name,
+         Captured = captured,
+         Lavaged = lavaged,
+         'Stomach Contents' = stmch_contents,
+         'Fish Contents' = fish_contents) %>%
+  mutate(Effort = recode(Effort,
+                         `Fall_2019` = "Fall 2019",
+                         `Fall_2020` = "Fall 2020",
+                         `Spring_2021` = "Spring 2021")) %>%
   kable(booktabs = T,
@@
 
-```{r fall-bio-fig, fig.cap = "First bioenergetics plot."}
+```{r fall-bio-fig, eval = F, fig.cap = "First bioenergetics plot."}
 fall_bio_p = fall_only_no_growth %>%
@@
 
-```{r full-yr-ng-fig, fig.cap = "Next bioenergetics plot."}
+```{r full-yr-ng-fig, eval = F, fig.cap = "Next bioenergetics plot."}
 full_yr_ng_p = full_yr_no_growth %>%
@@
 
-```{r adult-imp-p, fig.cap = "The adult impacts plot."}
+```{r adult-imp-p, eval = F, fig.cap = "The adult impacts plot."}
 adult_impact %>%

### 2021-12-21 analysis/R scripts/abund_est.R
small additions

@@ -122,7 +122,7 @@ N_mods = cmr_df %>%
 
 # diagnostic plots
 mod = N_mods %>%
-  slice(2) %>%
+  slice(1) %>%
   pull(mr_model) %>%
   magrittr::extract2(1)
 
@@ -142,6 +142,27 @@ tibble(m = mod$m,
               se = F,
               color = "red",
               linetype = 2)
+
+
+N_mods %>%
+  filter(model == "Schnabel") %>%
+  unnest(data) %>%
+  # filter(m > 0) %>%
+  ggplot(aes(x = M,
+             y = m/n)) +
+  geom_point(size = 4) +
+  facet_wrap(~ event_name,
+             scales = "free") +
+  labs(x = "Marked in Population",
+       y = "Prop. Recaptures in Sample") +
+  geom_smooth(method = 'loess',
+              se = F,
+              span = 2) +
+  geom_smooth(method = "lm",
+              se = F,
+              color = "red",
+              linetype = 2)
+
 
 #-----------------------------------------------------------------
 # using Lincoln-Petersen model, combining first 4 and last four days

### 2021-12-21 README.Rmd

@@
 # Please put your title here to include it in the file below.
-Title <- "Northern Pikeminnow Abundance in Deadwater Slough and Potential Impacts to Local Chinook Salmon Populations"
-Authors <- "Carmichael, Richard A., Kevin E. See, Nick Porter, and Michael W. Ackerman"
-yr <- 2021
-Journal <- "TBD"
+Title <- "Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations in the Upper Salmon River, Idaho"
+Authors <- "Ackerman, Michael W., Nicholas J. Porter, Kevin E. See, Michael S. Hall, Tara E. Blackman, and Richard A. Carmichael, and Michael W. Ackerman"
+yr <- 2022
+Journal <- "Transactions of the American Fisheries Society"
 
@@
 
+  - [:file\_folder: data](/analysis/data): Relevant data used in study analyses.
+  - [:file\_folder: figures](/analysis/figures): Figures, plots, and other
+    illustrations.
   - [:file\_folder: paper](/analysis/paper): R Markdown source document
@@
     `paper.docx`, suitable for reading (the code is replaced by figures
-    and tables in this file)
-  - [:file\_folder: data](/analysis/data): Data used in the analysis.
-  - [:file\_folder: figures](/analysis/figures): Plots and other
-    illustrations
-  - [:file\_folder:
-    supplementary-materials](/analysis/supplementary-materials):
-    Supplementary materials including notes and other documents
-    prepared and collected during the analysis.
+    and tables in this file).
+  - [:file\_folder: supplementary-materials](/analysis/supplementary-materials):
+    Supplementary materials including photographs, related R scripts, and files associated with the bioenergetics analysis.
+  - [:file\_folder: templates](/analysis/bioenergetics): Some files used to format the manuscript.
 

### 2021-12-21 analysis/paper/paper_KS.Rmd
started a new Rmd (paper_KS) that includes my edits/revisions to original paper. Knit into an html document.

--- 2021-12-21 analysis/paper/paper.Rmd
@@
 # set default theme for figures
-theme_set(theme_bw())
+# theme_set(theme_bw())
+theme_set(theme_classic())
 
@@
 
-The Snake and Columbia rivers drain a network of inland tributaries that provide essnetial spawning and rearing habitat for anadromous Pacific salmon *Oncorhynchus* spp.
-and steelhead *O.
-mykiss*.
-These fishes were historically abundant throughout the drainage basin and consequently have tremendous ecological, cultural, and economic value.
-Over the last century, salmonid stocks throughout the Snake and Columbia rivers have undergone significant declines related to habitat modification in the freshwater environment (e.g., removal of beavers, river channel simplification, water withdrawals, hydrosystem development) and poor marine conditions, both affecting their survival and recruitment and multiple life stages.
-Consequently, action agencies have made considerable investments in the rehabilitation of tributary ecosystems and evaulation of factors attributed to the species decline [@Roni2018].
-
-One potentially important, but perhaps under-appreciated source of salmonid mortality is predation on emigrating juveniles by piscivorous fishes.
-In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recover [@Fresh2003; @Winther2020].
-Dams and reservoirs in the Columbia River are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995].
-There are generally two mechanisms that explain these high predation zones.
-First, migration (movement) rates of juvenile salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation.
-Second, reservoirs and downstream tailraces associated with dams on the Snake and Columbia rivers create favorable slow-water habitat for predatory fishes known to consume juvenile salmonids, such as the Northern Pikeminnow *Ptychocheilus oregonsis*.
-Northern Pikeminnow tolerate and thrive in relatively warm, slow water habitats [@Wydoski2003] and consequently have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999].
-Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].
-
-Most predation studies have focused on reservoirs [@Murphy2021] and mainstem reaches [@Zimmerman1999; @Tabor1993; @Ward1995; @Shively1996] of the Columbia River basin whereas very little is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems [@Rubenson2020].
-Upstream habitats containing slower water velocities and other attributes that support piscivorous predators may overlap with essential habitat for some salmonid species and life stages.
-One such example is Deadwater Slough, an approximately 1.5 kilometer long reach of unnaturally slow and deep water in the Salmon River, Idaho.
-The Deadwater Slough is located in a river reach containing historically important overwinter rearing and spring emigration habitat for juvenile Chinook Salmon *O.
-tshawytscha* and is inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*.
-Further, the slough is part of the migratory pathway of the endangered Snake River Sockeye Salmon *O.
-nerka* population [@Axel2015] and several populations of threatened Snake River steelhead.
-
-Quantifying mortality associated with piscine predation in upstream habitats of the Salmon River is of particular interest for spring/summer Chinook Salmon in the Upper Salmon major population group (MPG).
-The Salmon River was historically the most productive tributary for spring/summer Chinook Salmon in the Columbia River basin [@Nemeth1999], but those populations have become depleted in recent decades.
-Currently, the Upper salmon MPG supports eight independent, extant spring/summer Chinook Salmon populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River (@NOAA2017).
-The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014].
-Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency.
-Because the Deadwater Slough supports rearing for DSR migrants and is part of the migration corridor for NRR migrants, and juveniles from all Upper Salmon populations must rear in or migrate through it, it represents a significant habitat for the Upper Salmon MPG.
-
-Several studies have attempted to examine the downstream movement, distribution, and (apparent) survival of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach.
-Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have substantially reduced survival in the Deadwater Slough reach compared to adjacent reaches [@Axel2015].
-Similarly, emigrating DSR Chinook Salmon had a 10% reduction in transition probability through the Deadwater Slough compared to surrounding reaches during the fall and early winter months [@Ackerman2018; @Porter2019].
-Low survival in these cases was attributed to low-velocity water delaying movement rates, thereby increasing predation risk.
-
-In this study, we estimated the abundance of Northern Pikeminnow in the Deadwater Slough and its potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
-We hypothesize that increase densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon and Sockeye Salmon.
-To test this, our objectives were three-fold:
-
-1. Estimate the abundance (or relative abundance) of potential predators in the Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations.
-2. Document predation on juvenile Chinook Salmon during the emigration periods using gastric lavage.
-3.
-Use a bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population at Deadwater Slough to assess their potential impacts to DSR and NRR Chinook Salmon emigrants.
-
-<!--
 The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations, each providing important ecological, social, and economic benefits to the region.
@@
 We also consider how consumption of juvenile Chinook Salmon by predators at Deadwater Slough might impact adult returns, hindering recovery of populations in the Upper Salmon River MPG.
--->
 
@@
 
-## Study Site
-
-The Deadwater Slough is an approximately 1.5-kilometer section of the mainstem Salmon River located roughly 5.8 river kilometers downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
-The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
-Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River [@Emerson1973], thereby creating a slow, deep section in the river, spanning approximately 30 acres, and averaging 68 m width.
-
 ## Northern Pikeminnow Abundance
 
-We estimated the population size of piscine predators in the Deadwater Slough using a mark-recapture survey and a catch per unit effort (CPUE) approach.
-Predators were sampled near the peaks of the juvenile DSR and NRR emigrations.
-Our intent was to sample during the emigration of the 2018 brood year (fall 2019 and spring 2020).
-However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 effort was delayed until fall 2020.
-Although this provided the benefit of an additional abundance estimate during the DSR emigration, estimates during the NRR emigration were still lacking, and so an additional effort was added in spring 2021.
-Fall efforts included sampling across a two-week span; however, due to budget constraints the spring 2021 effort was conducted over a single week.
-Fall efforts were constrained within a two-week window to minimize predator emigration/immigration and to constrain the population for the mark-recapture models.
-
-<!-- In the end, sampling dates were as follows: fall 2019 (mark = November 12-14; recapture = November 19-21), fall 2020 (mark = October 20-23; recapture = October 27-30); spring 2021 (mark = May 18-21).
--->
+We estimated the population size of piscine predators in the Deadwater Slough using a mark-recapture study and a CPUE approach.
+Predators were sampled near the peaks of the juvenile DSR and NRR emigrations for brood years 2018 and 2019.
+However, due to logistical concerns during the Covid-19 pandemic, sampling occurred during fall 2019, fall 2020, and spring 2021.
+During the fall sampling, fish caught each day were marked and released, and any previously marked fish were noted as a recapture for that day.
+We limited sampling to a total two week period to minimize predator emigration/immigration and to constrain the population for the mark-recapture models.
+Due to budget constraints the spring 2021 effort was conducted over a single week.
 
@@
 Methods included raft electrofishing, fyke netting, snorkeling, and angling.
-However, ineffective methods were abandoned in subsequent years.
-<!-- For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) our inability to "chase" fish downstream as they evaded the electricity.
-Snorkeling was ineffective due to the high turbidity of the mainstem Salmon River and Deadwater Slough area.
-Significant amounts of debris in the water column also created concerns that fyke nets could become entangled.
-Additionally, there were concerns that raft electrofishers and fyke nets could incidentally capture ESA-listed adult steelhead that are present in the mainstem Salmon River during fall and spring.
--->After evaluating all methods, angling proved to be the most effective method for capturing piscine predators while minimizing potential impacts to ESA-listed steelhead.
-<!--, but recognizing that angling can have both species and size selectivity (bias).--> Therefore, the following analyses will focus on fish captured by angling, unless otherwise noted.
-During the fall 2019 effort it additionally became clear that Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and so is the focus of our study.
+After evaluating all methods, angling proved to be the most effective method for capturing piscine predators while minimizing potential impacts to steelhead.
+Therefore, the following analyses will focus only on fish captured by angling, unless otherwise noted.
 
 Our study relied heavily on volunteer anglers who were permitted to fish anywhere within Deadwater Slough.
-During a sampling occassion (day), anglers would boat or hike their catch (periodically or upon filling a livewell) to a processing station at the boat ramp approximately 500 m downstream from the top of the slough.
+During a sample event, anglers would boat or hike their catch (periodically or upon filling a livewell) to a processing station located at the boat ramp approximately 500 m downstream from the top of the slough.
 For each fish, we recorded the date of capture, species, total length (TL; mm), and whether the fish was previously marked.
 Unmarked fish were given a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) unique to each day and released.
-Fish that died prior to release were still counted as captured (or recaptured), but not included in the number of marked fish available for recapture after that date.
 Released individuals were distributed throughout Deadwater Slough to facilitate mixing back into the population.
-Finally, we recorded the angling start and end time for each crew (person or combinations of persons) to facilitate calculations of CPUE.
+Finally, we recorded the angling start and end times for each crew (person or combinations of persons) to facilitate calculations of catch per unit effort (CPUE).
 
@@
          CI = map(mr_model,
-                  .f = confint),
+                  .f = confint,
+                  conf.level = 0.95),
          Lci = map_dbl(CI,
@@
 spring_abund = all_est %>%
+  filter(model == "Schnabel") %>%
   select(-range, -prop_range) %>%
@@
          cpue_F = catch_per_hr) %>%
+  group_by(species, model) %>%
+  summarise(across(c(N_F:cpue_F),
+                   mean),
+            .groups = "drop") %>%
   full_join(total_cpue %>%
@@
 
-Four different mark-recapture abundance estimators were employed to explore variation in estimates.
-Estimators fell into two broad categories: single-census and multiple-census.
-For the single-census estimators, we used the Lincoln-Peterson estimator:
+Estimators fell into two broad categories: single census and multiple census.
+For the single census estimators, we treated the first week of sampling as the mark event, and the following (second) week as the recapture event, pooling data within each of those weeks.
+Alternatively, the multiple census estimators treat each day as a survey, and use information about the total marked fish from all previous surveys to infer the total abundance.
+
+The Lincoln-Peterson estimator, a single-census estimator, is below, where $M$ is the number of fish marked and returned to the population, the $n$ is the number of fish caught in the second/recapture event, and $m$ is the number of marked fish in the second sample.
 
@@
 $$
-where $M$ is the number is the total number of fish marked and returned to the population during the first week of sampling (mark event), $n$ is the total number of fish caught in the second week (recapture event), and $m$ is the number of marked fish caught during the recapture event.
-Because the Lincoln-Peterson estimator can be biased with small samples, we also employed the Chapman-modified Lincoln-Peterson estimator:
+
+The Lincoln-Petersen estimator can be biased with small samples, so we also employed the Chapman-modified Lincoln-Petersen estimator:
 
@@
 $$
-For a multiple-census estimator, we first utilized the Schnabel estimator:
+
+The Schnabel estimator is a multiple census estimator where the $M$, $n$ and $m$ are indexed by the survey day, $i$.
+The Schnabel estimator does not have an associated standard error; however, the 95% confidence intervals can be calculated using a Poisson approximation.
+The Schnabel estimator is essentially a weighted average of a series of Lincoln-Peterson estimators (with a Chapman modification).
+The Schnabel estimator is shown here:
 
@@
 
-where $M$, $n$, and $m$ are indexed by the sampling occassion, $i$.
-Here, $i$ is defined as each sampling day.
-The Schnabel estimator does not have an associated standard error; however, the 90% confidence intervals were calculated.
-Lastly, we explored one additional multiple-census estimator, the Schumacher-Eschmeyer, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
-The Schumacher-Eschmeyer estimator is as follows:
+Finally, we explored one additional multiple census estimator, the Schumacher-Eschmeyer estimator, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
+The equation for the Schumacher-Eschmeyer estimator is:
 
@@
 The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and fall 2020 sampling periods.
-However, to estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of total CPUE in the fall sampling events to the total CPUE in the spring sampling event, then multiplied that ratio by the average abundance from the fall sampling events.<!-- Kevin, is this previous statement accurate?
---> This approach assumes equal capture probabilities between the fall and spring capture events.
-<!--If that assumption is not true and capture probability did differ, we believe that the capture probability  in the spring would be lower due to higher flows, which would lead to our spring abundance estimate to be biased low.
-Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 sampling event to be a conservative estimate.-->
-
-All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the sampling period), (2) marking a fish does not affect its chances of recapture, (3) no loss of marks, (4) marks are not missed or mistaken, and (5) all fish have equal chance of being caught in the second (and subsequent) sample(s).
+However, due to logistical constraints, we were only able to sample during a single week in spring 2021 and released fish were left unmarked.
+Instead, to estimate the abundance of Northern Pikeminnow during spring 2021, we used the ratio of total CPUE in the spring sampling events to the total CPUE in the fall sampling event, and multiplied that ratio by the average abundance of Northern Pikeminnow from the fall sampling events.
+
+All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the sampling period), (2) all fish have equal chance of being caught in the second (and subsequent) sample(s), (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, (5) marks are not missed or mistaken.
 
 ```{r psd}
-s_size = 250
-q_size = 380
+s_size = 300
+q_size = 400
 
@@
 
-<!--
 Finally, we calculated the proportional stock density (PSD) for Northern Pikeminnow captured during all field efforts, which is a measure of species size structure.
@@
 where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
-We calculated the PSD for Northern Pikeminnow in our study using `r prettyNum(s_size, digits = 0)` mm TL for stock-size and `r prettyNum(q_size, digits = 0)` mm TL for quality-size [@Winther2020].
--->
-
-<!--
-Juvenile Chinook Salmon abundances were estimated at 5 rotary screw traps in the upper Salmon River basin: Sawtooth, Pahsimeroi, Hayden Creek, Upper Lemhi, and Lower Lemhi.
-Abundances were calculated by Idaho Department of Fish and Game at each trap which made up the production estimate for juvenile Chinook Salmon in the Upper Salmon Basin (Table X) (McClure et al.
-2020).
--->
+We calculated the PSD for Northern Pikeminnow in our study using `r prettyNum(s_size, digits = 0)` mm TL for stock-size and `r prettyNum(q_size, digits = 0)` mm TL for quality-size.
 
@@
 For each sample, wet weight (grams) was recorded for the total combined stomach content, including all non-fish items (e.g., macroinvertebrates, organic matter), as well as the portion consisting of fish parts/matter.
-Fish and fish remnants were identified to the lowest taxonomic unit using diagnostic bones, or were categorized as unknown [@Hansel1988; @Frost2000].
+Fish and fish remnants were identified to the lowest taxonomic unit, when possible, or were categorized as unknown.
 A subset of Northern Pikeminnow captures (~ 5%) were euthanized for dissection after gastric lavage to validate the efficacy of the methodology.
@@
 
-```{r shoup-temp-data, eval = F}
+```{r shoup-temp-data}
 shoup_temp_df = read_tsv(here("analysis/data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
@@
 rm(spring_f10, spring_f20, spring_f30, spring_f40, spring_f50, spring_f60, spring_f70, spring_f80, spring_f90)
-
+```
+
+To estimate the total consumption potential (i.e., number/amount of juvenile consumed during the peaks of DSR [fall] and NRR [spring] emigrations) of Northern Pikeminnow in Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 applied in the R statistical software [@RCoreTeam2021].
+The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on predator and prey energy densities, predator start and end weights, and water temperatures.
+Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017].
+Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon based on an estimate from @Moss2016 and 3,000 J/g for invertebrates.
+Additional settings included 0% assumed indigestible prety, and a variation in the proportion of diet that was invertebrates or fish ranging from 10% to 90% for both categories.
+We used a weight-length formula [@Parker1995] to calculate the average predator start weight from the average length of Northern Pikeminnow captured during our study using the FSA package in R [@Ogle2021].
+The average length of Northern Pikeminnow captured at the Deadwater Slough during the spring and fall were 352.9 and 393.7 mm, respectively, which calculates to average starting weights of 504.4 and 670.5 g.
+Continuous water temperature data from 2019 was used in two alternative models, one using the baseline start and end weights and the second with a 10% increase in average end weight on Northern Pikeminnow.
+
+We ran three separate bioenergetic models for this study.
+For the first model, we used temperature data from a 78-day period spanning September 1 through November 17 when DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration.<!-- Did we just use data from Lemhi here?
+All tributaries?
+Was this subjective?
+--> During this time, water temperatures typically exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon<!--citation?-->.
+This first model assumed no growth for Northern Pikeminnow.
+The second model was instead run for one full year, but also assumed no growth for Northern Pikeminnow.
+The assumption of zero growth was included to estimate consumption if individual Northern Pikeminnow were not growing and/or the population is stable.
+Finally, the third model shows how an increase of 10% body weight or an increase in population size changes the estimates of consumption over a full year period.
+
+To estimate the total biomass consumed, we then multiplied the biomass consumed by an individual Northern Pikeminnow for the given time period by the estimated population size for that time period.
+The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed by dividing by a weight of 12 g per individual.
+This weight was an average from juvenile Chinook Salmon captured at six rotary screw traps throughout the operation season March to November in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.
+
+<!--
+A few things that need to be clarified in this section:
+* Why were the dates Sept 1 - Nov 17 chosen? They make sense, but we should include our rationale
+* Where did the temperature data come from? Hobo? Stream gauge? The methods say temperature data from 2019 was used, but the results seem to use 2020? 
+* Why does it look like we only ran it for dates in 2020? Is that the only year we had temperature data for? Would the results look different if we used temperature data from a different year?
+* The lengths of pikeminnow: did we only use pikeminnow we caught in the fall to get average length? Is there a difference in lengths between 2019 and 2020? 
+* There needs to be more explanation of the inverts vs. fish. What proportion went to each and why? Did it change through time? Based on literature at all? 
+* Where is the data that was used to get a weight of 12 g for Chinook emigrant?
+Was that for Chinook presmolts, or a combination of smolts and presmolts?
+If the latter, can we filter it to look at the average weight for presmolts only?
+And maybe smolts only for a different analysis?
+* Could we run a similar model for the spring emigration period?
+Figure out the dates, look at lengths of pikeminnow from the spring, convert to average weight, use average size of Chinook smolts (not presmolts) from all the Upper Salmon traps?
+Kevin can work on converting the spring catch into some kind of estimate of pikeminnow at Deadwater in the spring
+-->
+
+# Results
+
+## Northern Pikeminnow Abundance
+
+We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during Fall 2019, Fall 2020, and Spring 2021, respectively.
+The catch per unit effort, measured as the number of Northern Pikeminnow captured per angler hour, for each was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for Fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for Fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for Spring 2021 (Figure \@ref(fig:cpue-fig)).
+During Fall 2019, we had a total of `r as.english(sum(N_mods[[3]][[1]]$m))` recaptures which included `r as.english(LP_mod[[3]][[1]]$m)` unique individuals (i.e., `r as.english(sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m)` individual was captured more than once).
+There were `r as.english(sum(N_mods[[3]][[3]]$m))` recaptures in Fall 2020 which included `r as.english(LP_mod[[3]][[3]]$m)` unique individuals.
+Summaries of parameters for both the single-census and multiple-census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.
+
+Abundance estimates for Northern Pikeminnow in the Deadwater Slough during Fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for Fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")`.
+In both cases, the multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied (Figure \@ref(fig:abund-fig)).
+Table \@ref(tab:abund-tab) provides point estimates and confidence intervals for each estimator for the fall mark-recapture sampling events.
+
+All the estimators yielded estimates with overlapping confidence intervals.
+Because the sampling design most closely matched a multiple census estimator, we feel the Schnabel estimator is the most appropriate.
+We examined diagnostic plots for that estimator (number of marks vs.
+proportion of recaptures) and they suggested that we were meeting the assumptions of the Schnabel estimator.
+Therefore, further results will be based on the Schnabel estimates of abundance.
+
+Using the mean of the Schnabel estimates of Northern Pikeminnow abundance across two fall sampling events, and the mean of the CPUE from those events as well, we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95%CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring of 2020.
+
+The PSD for Northern Pikeminnow in our study and across all three years was `r prettyNum(psd, digits = 1)`%.
+In other words, `r prettyNum(psd, digits = 1)`% of captured fish were over the quality size classification of `r prettyNum(q_size, digits = 0)` mm TL, demonstrating that a high percentage of Northern Pikeminnow within Deadwater Slough are above average size classes defined for the species
+
+## Gastric Lavage
+
+We examined the stomach contents of `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow over the course of the study using gastric lavage methods.
+Of those, we documented some form of stomach contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` of fish of which we were able to confirm fish or fish remnants in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")`.
+Expressed as a percentage, `r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% of lavaged Northern Pikeminnow had stomach contents of which `r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% we identified containing fish or fish remnants.
+Table \@ref(tab:gastric-tab) provides a further breakdown of gastric lavage results for each sampling event.
+
+For Northern Pikeminnow with some form of stomach contents, the average wet weight of total contents was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r prettyNum(min(stmch_df$stomach_contents_weight), digits = 3)`-`r prettyNum(max(stmch_df$stomach_contents_weight), digits = 3)` g).
+Overall, fish or fish remnants were `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall weighted weight among all stomach contents examined.
+
+## Bioenergetics
+
+We estimated a single average-sized Northern Pikeminnow to consume 61 g of ***fish*** during the modeled fall time period (September 1 - November 17).
+Given an estimated population size of ***xx,xxx***, we'd estimate that ***xxx,xxx*** g of fish are consumed each fall during the emigration of juvenile DSR Chinook Salmon.
+The average weight of a DSR Chinook Salmon is ***xx.x*** taken from ***yyy*** rotary screw traps
+
+If we further assume an average weight of ***xx g*** for DSR Chinook Salmon
+
+Over the fall time period a single average-sized Northern Pikeminnow is modeled to eat 61 g of fish.
+For the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook Salmon.
+Therefore, Northern Pikeminnow in the Deadwater Slough can potentially consume ***xx,xxx*** juvenile Chinook Salmon as a discrete population.
+
+Over an entire year, we estimate that a single Northern Pikeminnow will eat 283 g of fish and 613 g of other food types.
+
+<!--
+* Water Temperature measurements.
+    *	Hobo tidbit placed at site. 
+    *	Calculated daily average water temperature for 365 days of the year.
+*	Water temperature plot for the 365 average.
+*	Bioenergetics results. Total mass consumed over a 365 day period. Plot of daily consumption with day of year on x and mass on y.
+-->
+
+# Discussion
+
+We estimated the population size of Northern Pikeminnow in the Deadwater Slough to be greater than ***xx,xxx*** during the fall emigration period for DSR Chinook Salmon.
+That estimate translates to a density of ***xxx*** Northern Pikeminnow per 100 m or ***xxx*** per 100 m^2^ which is ***similar/more/less*** than estimates from elsewhere in the Columbia River ***(citation)*** where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminnow abundance.
+The population size of Northern Pikeminnow was not directly estimated during the spring NRR Chinook salmon emigration period; however, the relative abundance measured at CPUE was comparable to the fall sampling periods ***(update statement later)***.
+The population of Northern Pikeminnow in Deadwater Slough was estimated to consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook Salmon during the ***x*** sampling periods and result in an estimated reduction of returning adults between ***xxx*** and ***x,xxx***.
+We suggest that the habitat modifications that created the Deadwater Slough have resulted in favorable conditions for Northern Pikeminnow, including improved conditions for predation upon juvenile Chinook Salmon ***(add detail here)***.
+Therefore, predation by Northern Pikeminnow in the Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG.
+
+## Mark-Recapture Model
+
+All estimators suggest a large number of Northern Pikeminnow occupying Deadwater Slough.
+
+<!-- This might be a good place to bring in some comparisons with pikeminnow densities from the Columbia, put our numbers in context. -->
+
+If the first mark-recapture assumption was violated and the population is actually open (potential immigration and emigration during the survey period), as long as the immigration and emigration rates are equal between marked and unmarked fish, our estimators should still provide unbiased estimates.
+Given the size of the sampling area, and the length of the sampling period (two weeks)
+
+We are comfortable with assumptions 4 and 5 (no loss of marks and no mistaken or missed marks).
+Assumptions 2 and 3 indicate whether the second (and subsequent) surveys are random samples of the population.
+Angling methods may have a size selection bias, but that may only affect the interpretation of the abundance estimates (i.e.
+our estimates are the abundance of Northern Pikeminnow above some size threshold).
+Unequal catchability between individual fish is very difficult to assess, especially if that heterogenity arises from being caught once.
+
+<!-- If catching and marking fish makes them want to leave the Deadwater area (more than unmarked fish), that could be making our estimates biased high (due to fewer recaptures).
+I don't imagine that is happening, but does anyone else have concerns about that?
+-->
+
+<!--
+Re-hash out our assumptions?
+What assumptions were most likely violated e.g., the closed population assumption?
+If we assumed an open population with varying immigration/emigration rates how might that affect our estimate?
+In the end, how good do we think our estimate is and is the bottom-line that we belive there still to be a shit-ton of pikeminnow even if our assumptions were violated?
+-->
+
+The spring abundance estimate is smaller than either fall sampling event.
+However, our approach assumes equal capture probabilities between the fall and spring sampling events.
+If it does differ, we believe that the capture probability in the spring could be lower due to higher flows, which would lead to our spring abundance estimate being biased low.
+Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 sampling event to be a conservative estimate.
+
+## Gastric Lavage
+
+The majority of fish collected during this study received gastric lavage, including some non-predatory species.
+To validate the efficacy of this method, we euthanized nine Northern Pikeminnow after gastric lavage was completed and removed the remaining stomach contents via dissection.
+Our results support previous findings that gastric lavage effectively removes stomach contents (@Lott2020).
+Additionally, we found that the fish captured in net-traps had a similar proportion of stomach content samples.
+Of those, 1,214 (76.2%) were completely empty, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
+We were successfully able to identify Redside Shiners, a Largescale Sucker, sculpin, Mountain Whitefish, and one juvenile Chinook Salmon in the stomach contents.
+However, decomposition from digestion rendered most stomach contents unidentifiable.
+
+<!--
+Concerns with gastric lavage.
+We only observed juvenile Chinook salmon (and/or other fishes) in a very small number of stomach contents.
+However, we don't necessarily believe that to mean that pikeminnow in Deadwater Slough rarely consume juvenile Chinook salmon.
+Are there other cases where gastric lavage failed to document predation?
+How soon after consumption do you need to take a sample?
+So instead, we made some assumptions about the proportion of a pikeminnows diet consisting of Chinook salmon.
+-->
+
+## Bioenergetics
+
+What assumptions did we make during the bioenergetics assessment?
+And how might violations of those assumptions change our estimate of the number of juvenile Chinook salmon consumed and resulting impacts to adult returns?
+
+## Impacts to Adult Returns
+
+```{r}
+# pull in SAR numbers
+sar = read_csv(here("analysis/data/raw_data", 
+                    "WSSCh_sar_4See.csv")) %>%
+  summarise(across(c(medsar, sdsar),
+                   median)) %>%
+  summarise(across(c(medsar, sdsar),
+               ~ . / 100))
+
+# summarze potential Chinook eaten for each timeperiod by one pikeminnow
+adult_impact = fall_results %>% 
+  summarise(across(ends_with("g_n"), max)) %>%
+  add_column(season = "Fall",
+             .before = 0) %>%
+  bind_rows(spring_results %>% 
+              summarise(across(ends_with("g_n"), max)) %>%
+              add_column(season = "Spring",
+                         .before = 0)) %>%
+  pivot_longer(cols = ends_with("g_n"),
+               names_to = "perc_diet",
+               values_to = "juv_chnk") %>%
+  mutate(across(perc_diet,
+                str_extract,
+                "[:digit:]+"),
+         across(perc_diet,
+                as.integer)) %>%
+  left_join(spring_abund %>%
+              select(N_F:Uci_F,
+                     N_S:Uci_S) %>%
+              pivot_longer(everything()) %>%
+              mutate(season = if_else(str_detect(name, "_F$"),
+                                                 "Fall",
+                                                 "Spring")) %>%
+              mutate(across(name,
+                            str_remove,
+                            "_F$"),
+                     across(name,
+                            str_remove,
+                            "_S$")) %>%
+              pivot_wider(names_from = name,
+                          values_from = value)) %>%
+  mutate(smolt_N = N * juv_chnk,
+         smolt_Lci = Lci * juv_chnk,
+         smolt_Uci = Uci * juv_chnk) %>%
+  bind_cols(sar) %>%
+  mutate(adult_N = smolt_N * medsar,
+         adult_Lci = smolt_Lci * qnorm(0.025, medsar, sdsar),
+         adult_Uci = smolt_Uci * qnorm(0.975, medsar, sdsar))
+
+# sum fall and spring impacts
+total_impacts = adult_impact %>%
+  group_by(perc_diet) %>%
+  summarise(across(c(ends_with("_N"),
+                     ends_with("_Lci"),
+                     ends_with("_Uci")),
+                   sum))
 ```
@@
 
-To estimate the total biomass of fish potentially consumed by Northern Pikeminnow in Deadwater Slough, we then multiplied the biomass consumed by an individual Northern Pikeminnow for the time period by the estimated population size for that time period.
-<!-- from which estimator?
--->The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed by dividing by a weight of 10.9 g for the spring and 10.3 g for the fall, an average from juvenile Chinook Salmon captured during those date ranges at seven rotary screw traps located throughout the Upper Salmon upstream of Deadwater Slough.
-
-<!-- Sawtooth Hatchery, East Fork Salmon, Pahsimeroi, Upper Lemhi, Hayden Creek, Lower Lemhi, North Fork Salmon -->
-
-# Results
-
-## Northern Pikeminnow Abundance
-
-We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during Fall 2019, Fall 2020, and Spring 2021, respectively.
-The catch per unit effort, measured as the number of Northern Pikeminnow captured per angler hour, for each was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for Fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for Fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for Spring 2021 (Figure \@ref(fig:cpue-fig)).
-During Fall 2019, we had a total of `r as.english(sum(N_mods[[3]][[1]]$m))` recaptures which included `r as.english(LP_mod[[3]][[1]]$m)` unique individuals (i.e., `r as.english(sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m)` individual was captured more than once).
-There were `r as.english(sum(N_mods[[3]][[3]]$m))` recaptures in Fall 2020 which included `r as.english(LP_mod[[3]][[3]]$m)` unique individuals.
-Summaries of parameters for both the single-census and multiple-census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.
-
-Abundance estimates for Northern Pikeminnow in the Deadwater Slough during Fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for Fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")`.
-In both cases, the multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied (Figure \@ref(fig:abund-fig)).
-Table \@ref(tab:abund-tab) provides point estimates and confidence intervals for each estimator for the fall mark-recapture sampling events.
-
-***Finish spring abundance results.*** <!-- Can we assume equal capture probability and perhaps using the ratio of the average fall CPUE and spring CPUE.
-Then multiply the average fall abundance estimate by the ratio.
-Does that seem appropriate?
--->
-
-The PSD for Northern Pikeminnow in our study and across all three years was `r prettyNum(psd, digits = 1)`%.
-In other words, `r prettyNum(psd, digits = 1)`% of captured fish were over the quality size classification of `r prettyNum(q_size, digits = 0)` mm TL, demonstrating that a high percentage of Northern Pikeminnow within Deadwater Slough are above average size classes defined for the species
-
-## Gastric Lavage
-
-We examined the stomach contents of `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow over the course of the study using gastric lavage methods.
-Of those, we documented some form of stomach contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` of fish of which we were able to confirm fish or fish remnants in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")`.
-Expressed as a percentage, `r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% of lavaged Northern Pikeminnow had stomach contents of which `r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% we identified containing fish or fish remnants.
-Table \@ref(tab:gastric-tab) provides a further breakdown of gastric lavage results for each sampling event.
-
-For Northern Pikeminnow with some form of stomach contents, the average wet weight of total contents was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r prettyNum(min(stmch_df$stomach_contents_weight), digits = 3)`-`r prettyNum(max(stmch_df$stomach_contents_weight), digits = 3)` g).
-Overall, fish or fish remnants were `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall weighted weight among all stomach contents examined.
-
-## Bioenergetics
-
-We estimated a single average-sized Northern Pikeminnow to consume 61 g of ***fish*** during the modeled fall time period (September 1 - November 17).
-Given an estimated population size of ***xx,xxx***, we'd estimate that ***xxx,xxx*** g of fish are consumed each fall during the emigration of juvenile DSR Chinook Salmon.
-The average weight of a DSR Chinook Salmon is ***xx.x*** taken from ***yyy*** rotary screw traps
-
-If we further assume an average weight of ***xx g*** for DSR Chinook Salmon
-
-Over the fall time period a single average-sized Northern Pikeminnow is modeled to eat 61 g of fish.
-For the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook Salmon.
-Therefore, Northern Pikeminnow in the Deadwater Slough can potentially consume ***xx,xxx*** juvenile Chinook Salmon as a discrete population.
-
-Over an entire year, we estimate that a single Northern Pikeminnow will eat 283 g of fish and 613 g of other food types.
-
-<!--
-* Water Temperature measurements.
-    *	Hobo tidbit placed at site. 
-    *	Calculated daily average water temperature for 365 days of the year.
-*	Water temperature plot for the 365 average.
-*	Bioenergetics results. Total mass consumed over a 365 day period. Plot of daily consumption with day of year on x and mass on y.
--->
-
-# Discussion
-
-We estimated the population size of Northern Pikeminnow in the Deadwater Slough to be greater than ***xx,xxx*** during the fall emigration period for DSR Chinook Salmon.
-That estimate translates to a density of ***xxx*** Northern Pikeminnow per 100 m or ***xxx*** per 100 m^2^ which is ***similar/more/less*** than estimates from elsewhere in the Columbia River ***(citation)*** where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminnow abundance.
-The population size of Northern Pikeminnow was not directly estimated during the spring NRR Chinook salmon emigration period; however, the relative abundance measured at CPUE was comparable to the fall sampling periods ***(update statement later)***.
-The population of Northern Pikeminnow in Deadwater Slough was estimated to consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook Salmon during the ***x*** sampling periods and result in an estimated reduction of returning adults between ***xxx*** and ***x,xxx***.
-We suggest that the habitat modifications that created the Deadwater Slough have resulted in favorable conditions for Northern Pikeminnow, including improved conditions for predation upon juvenile Chinook Salmon ***(add detail here)***.
-Therefore, predation by Northern Pikeminnow in the Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG.
-
-## Mark-Recapture Model
-
-<!--
-Re-hash out our assumptions?
-What assumptions were most likely violated e.g., the closed population assumption?
-If we assumed an open population with varying immigration/emigration rates how might that affect our estimate?
-In the end, how good do we think our estimate is and is the bottom-line that we belive there still to be a shit-ton of pikeminnow even if our assumptions were violated?
--->
-
-## Gastric Lavage
-
-The majority of fish collected during this study received gastric lavage, including some non-predatory species.
-To validate the efficacy of this method, we euthanized nine Northern Pikeminnow after gastric lavage was completed and removed the remaining stomach contents via dissection.
-Our results support previous findings that gastric lavage effectively removes stomach contents (@Lott2020).
-Additionally, we found that the fish captured in net-traps had a similar proportion of stomach content samples.
-Of those, 1,214 (76.2%) were completely empty, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
-We were successfully able to identify Redside Shiners, a Largescale Sucker, sculpin, Mountain Whitefish, and one juvenile Chinook Salmon in the stomach contents.
-However, decomposition from digestion rendered most stomach contents unidentifiable.
-
-<!--
-Concerns with gastric lavage.
-We only observed juvenile Chinook salmon (and/or other fishes) in a very small number of stomach contents.
-However, we don't necessarily believe that to mean that pikeminnow in Deadwater Slough rarely consume juvenile Chinook salmon.
-Are there other cases where gastric lavage failed to document predation?
-How soon after consumption do you need to take a sample?
-So instead, we made some assumptions about the proportion of a pikeminnows diet consisting of Chinook salmon.
--->
-
-## Bioenergetics
-
-What assumptions did we make during the bioenergetics assessment?
-And how might violations of those assumptions change our estimate of the number of juvenile Chinook salmon consumed and resulting impacts to adult returns?
-
-## Impacts to Adult Returns
-
-```{r}
-# pull in SAR numbers
-sar = read_csv(here("analysis/data/raw_data", 
-                    "WSSCh_sar_4See.csv")) %>%
-  summarise(across(c(medsar, sdsar),
-                   median)) %>%
-  summarise(across(c(medsar, sdsar),
-               ~ . / 100))
-
-adult_impact = fall_only_no_growth %>%
-  filter(cum_cons_fish_g == max(cum_cons_fish_g)) %>%
-  select(cum_cons_fish_g) %>%
-  mutate(n_chnk_smolts = cum_cons_fish_g / 12) %>%
-  bind_cols(all_est %>%
-              select(event_name, 
-                     model,
-                     N,
-                     Lci,
-                     Uci)) %>%
-  mutate(across(c(smolt_N = N, 
-                  smolt_Lci = Lci, 
-                  smolt_Uci = Uci),
-                ~ . * n_chnk_smolts)) %>%
-  bind_cols(sar) %>%
-  mutate(adult_N = smolt_N * medsar,
-         adult_Lci = smolt_Lci * qnorm(0.025, medsar, sdsar),
-         adult_Uci = smolt_Uci * qnorm(0.975, medsar, sdsar))
-
-```
-
-To estimate the number of possible Chinook salmon that could be consumed, we divided the cumulative fish consumption in grams for an individual pikeminnow by the average weight of a Chinook emigrant to come up with a presmolt equivalency and then multiplied this by our abundance estimate of northern pikeminnow in Deadwater Slough, resulting in an estimate of the equivalent total number of Chinook salmon presmolts consumed.
-Finally, we estimated the number of additional adult Chinook salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough reach was reduced or eliminated.
-We multiplied the total estimated consumed juvenile Chinook salmon by an estimate of Granite-to-Granite SARs [@McCann2019].
+To estimate the total biomass of fish potentially consumed by Northern Pikeminnow in Deadwater Slough, we then multiplied the biomass consumed by an individual Northern Pikeminnow for the time period by the estimated population size for that time period (mean Schnabel estimates in the fall, CPUE ratio generated estimate in the spring).
+The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed by dividing by a weight of `r avg_NRR_g` g for the spring and `r avg_DSR_g` g for the fall, an average from juvenile Chinook Salmon captured during those date ranges at seven rotary screw traps located throughout the Upper Salmon upstream of Deadwater Slough.
+Finally, we estimated the number of additional adult Chinook salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough reach was reduced or eliminated by multiplying the total estimated consumed juvenile Chinook salmon by an estimate of Granite-to-Granite SARs (`r round(sar$medsar, 5)`, SE: `r round(sar$sdsar, 5)`) [@McCann2019].
+
+The estimated number of Chinook Salmon consumed by Northern Pikeminnow begins to taper off between 30-90% of Pikeminnow diet.
+So although we are uncertain of that diet proportion, the impact of that uncertainty is somewhat constrained.
+We estimated the total juvenile outmigrants consumed (fall and spring) to be between `r paste(prettyNum(round(range(total_impacts$smolt_N)), big.mark = ","), collapse = " and ")`.
+This translates to adult Chinook equivalents between `r paste(prettyNum(round(range(total_impacts$adult_N)), big.mark = ","), collapse = " and ")` (Figure \@ref(fig:adult-imp-p)).
 
@@
 
-*	SAR results. Maybe a plot of multiple scenarios depending on gastric lavage results? And pop estimates if they vary in time.
+*	SAR results. Those results are GRA - GRA SARs, estimated from multiple years of PIT tag data. Seem pretty good to me.
+* Maybe a plot of multiple scenarios depending on gastric lavage results? 
+* Maybe pop estimates if they vary in time.
 
@@
 
+```{r spring-N}
+spring_abund %>%
+  rowwise() %>%
+  mutate(`Avg. Fall N (CI)` = paste0(prettyNum(round(N_F), big.mark = ","),
+                                     " (",
+                                     paste0(prettyNum(round(c(Lci_F, Uci_F)), big.mark = ","),
+                                            collapse = ", "),
+                                     ")"),
+         `Spring N (CI)` = paste0(prettyNum(round(N_S), big.mark = ","),
+                                     " (",
+                                     paste0(prettyNum(round(c(Lci_S, Uci_S)), big.mark = ","),
+                                            collapse = ", "),
+                                     ")")) %>%
+  ungroup() %>%
+  select(Species = species,
+         `Avg Fall CPUE` = cpue_F,
+         `Spring CPUE` = cpue_S,
+         ends_with("(CI)")) %>%
+  kable(booktabs = T,
+        caption = "Average CPUE and abundance in the fall (95% confidence interval), and CPUE and estimates of spring abundance (95% confidence interval) for Northern Pikeminnow in Deadwater Slough.") %>%
+  kable_styling()
+```
+
 ```{r gastric-tab}
@@
 
+```{r adult-imp}
+adult_impact %>%
+  mutate(stage = "Smolt") %>%
+  select(#event_name, 
+         #model,
+         stage,
+         starts_with("smolt")) %>%
+  rename_with(.fn = ~ str_remove(., "smolt_"),
+              .cols = starts_with("smolt")) %>%
+  bind_rows(adult_impact %>%
+              mutate(stage = "Adult Equivalent") %>%
+              select(#event_name, model,
+                     stage,
+                     starts_with("adult")) %>%
+              rename_with(.fn = ~ str_remove(., "adult_"),
+                          .cols = starts_with("adult"))) %>%
+  kable(booktabs = T,
+        format.args = list(big.mark = ","),
+        col.names = c("Stage",
+                      "N",
+                      "Lower 95% CI",
+                      "Upper 95% CI"),
+        digits = 0,
+        caption = "Estimates of smolts consumed during fall migration by Northern Pikeminnow, and the adult equivalent of those smolts based on SARs.") %>%
+  kable_styling()
+```
+
 # Figures
@@
              y = catch_per_hr)) +
-  geom_col(fill = "black") +
+  geom_col(fill = "gray30") +
   labs(y = "Catch Per Angler Hour") +
-  theme(axis.title.x = element_blank()) +
-  theme_classic()
-
-# total_cpue %>%
-#   ggplot(aes(x = event_name,
-#              y = catch_lg_per_angler_day,
-#              fill = event_name)) +
-#   geom_col() +
-#   scale_fill_brewer(palette = "Set1",
-#                     name = "Event") +
-#   labs(x = "Event",
-#        y = "CPUE (Large Pikeminnow caught per angler day)") +
-#   theme(legend.position = "none")
+  theme(axis.title.x = element_blank())
 
@@
        y = "Abundance (x 1,000)") +
-  theme_bw() +
   theme(legend.position = "none",
@@
 
-# all_est %>%
-#   mutate(across(c(N, Lci, Uci),
-#                  ~ . / 1000)) %>%
-#   ggplot(aes(x = model,
-#              y = N,
-#              color = model)) +
-#   geom_errorbar(aes(ymin = Lci,
-#                     ymax = Uci),
-#                 width = 0.1) +
-#   geom_point(size = 4) +
-#   facet_wrap(~ event_name) +
-#   scale_color_brewer(palette = "Set1") +
-#   coord_cartesian(ylim = c(0, 100)) +
-#   labs(x = "Estimator",
-#        y = "Abundance (x 1,000)") +
-#   theme(legend.position = "none",
-#         axis.text.x = element_text(angle = 45,
-#                                    hjust = 1))
-
-```
-
-\newpage
-
-```{r fall-bio-fig, fig.cap = "First bioenergetics plot."}
-fall_bio_p = fall_only_no_growth %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Scumulative fish eaten (g)",
-                     breaks = seq(0, 200, by = 10),
-                     sec.axis = sec_axis(~. / 12, 
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 20, by = 1))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)")
-
-fall_bio_p
-
-```
-
-\newpage
-
-```{r full-yr-ng-fig, fig.cap = "Next bioenergetics plot."}
-full_yr_ng_p = full_yr_no_growth %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  #scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Cumulative fish eaten (g)",
-                     breaks = seq(0, 600, by = 50),
-                     sec.axis = sec_axis(~. / 12,
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 30, by = 2))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)",
-       x = "Month")
-
-full_yr_ng_p
-
-```
-
-```{r full-yr-10p-fig, eval = F, fig.cap = "Another bioenergetics plot."}
-full_yr_10p_p = full_yr_10p_growth %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  #scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Cumulative fish eaten (g)",
-                     breaks = seq(0, 600, by = 50),
-                     sec.axis = sec_axis(~. / 12,
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 40, by = 2))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)",
-       x = "Month")
-
-full_yr_10p_p
-
-```
-
-```{r alt-mod-fig, eval = F, fig.cap = "Another bioenergetics plot, again."}
-alt_mod_p = alternative_mod %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  #scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Cumulative fish eaten (g)",
-                     breaks = seq(0, 600, by = 50),
-                     sec.axis = sec_axis(~. / 12,
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 30, by = 2))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)",
-       x = "Month")
-
-alt_mod_p
-
-```
+```
+
+\newpage
 
@@
 
-```{r adult-imp-p, fig.cap = "The adult impacts plot."}
+```{r adult-imp-p, fig.cap = "How many Chinook Salmon are consumed by Northern Pikeminnow, depending on what percent of diet is Chinook. Faceted by migration season and life-stage."}
 adult_impact %>%
-  mutate(stage = "Smolt") %>%
-  select(event_name, model,
-         stage,
+  mutate(stage = "Juv. Emigrant") %>%
+  select(stage,
+         season,
+         perc_diet,
          starts_with("smolt")) %>%
@@
               mutate(stage = "Adult Equivalent") %>%
-              select(event_name, model,
-                     stage,
+              select(stage,
+                     season,
+                     perc_diet,
                      starts_with("adult")) %>%
@@
                           .cols = starts_with("adult"))) %>%
-  filter(str_detect(model, "Schuma", negate = T)) %>%
-  ggplot(aes(x = model,
+  ggplot(aes(x = perc_diet,
              y = N,
-             color = model)) +
+             color = as.factor(perc_diet))) +
   geom_errorbar(aes(ymin = Lci,
                     ymax = Uci),
-                width = 0.5) +
+                width = 0.1) +
   geom_point(size = 4) +
-  facet_grid(stage ~ event_name,
+  facet_grid(fct_rev(stage) ~ season,
              scales = "free_y") +
-  scale_color_brewer(palette = "Set1",
-                     name = "Model") +
+  scale_color_viridis_d(name = "Percent Chinook\nin Diet") +
   theme(axis.text.x = element_blank(),
-        legend.position = "bottom") +
-  labs(x = "N. Pikeminnow Abundance Model")
+        legend.position = "bottom",
+        axis.title.x = element_blank(),
+        axis.ticks.x = element_blank()) +
+  labs(y = "Number Chinook Consumed")
 ```

### 2021-12-21 analysis/paper/paper.Rmd
added KS content

@@
 # set default theme for figures
-theme_set(theme_bw())
+#theme_set(theme_bw())
+theme_set(theme_classic())
 
@@
 
-Four different mark-recapture abundance estimators were employed to explore variation in estimates.
+Four different mark-recapture abundance estimators were employed to explore variation in estimates among them.
 Estimators fell into two broad categories: single-census and multiple-census.
-For the single-census estimators, we used the Lincoln-Peterson estimator:
+For the single census estimators, we treated the first week of sampling as the mark event, and the following (second) week as the recapture week, pooling data within each of those weeks.
+Alternatively, the multiple census estimators treat each day as a sampling occassion, and use information about the total marked fish from all previous occassions to infer the total abundance.
+
+For the single-census estimators, we first used the Lincoln-Peterson estimator:
 
@@
 $$
-For a multiple-census estimator, we first utilized the Schnabel estimator:
+For a multiple-census estimator, we first utilized the Schnabel method:
 
@@
 
-All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the sampling period), (2) marking a fish does not affect its chances of recapture, (3) no loss of marks, (4) marks are not missed or mistaken, and (5) all fish have equal chance of being caught in the second (and subsequent) sample(s).
+All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the sampling period), (2) all fish have equal chance of being caught in the second (and subsequent) sample(s), (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, (5) marks are not missed or mistaken.
 
@@
 
-```{r read-bio-outputs}
+```{r read-bio-outputs, message=FALSE, warning=FALSE}
 fmt_bio_out = function(x) {
@@
   return(x)
+
 }
@@
 
-To estimate the total biomass of fish potentially consumed by Northern Pikeminnow in Deadwater Slough, we then multiplied the biomass consumed by an individual Northern Pikeminnow for the time period by the estimated population size for that time period.
-<!-- from which estimator?
--->The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed by dividing by a weight of 10.9 g for the spring and 10.3 g for the fall, an average from juvenile Chinook Salmon captured during those date ranges at seven rotary screw traps located throughout the Upper Salmon upstream of Deadwater Slough.
+To estimate the total biomass of fish potentially consumed by Northern Pikeminnow in Deadwater Slough, we then multiplied the biomass consumed by an individual Northern Pikeminnow for the timer period by the estimated population size for that time period (mean Schnabel estimates in the fall, CPUE ratio generated estimate in the spring).
+The total biomass was further converted to an estimate of the total number of Chinook salmon consumed by dividing by a weight of `r avg_DSR_g` for the fall and `r avg_NRR_g` for the spring, an average from juvenile Chinook Salmon captured during those date ranges at seven rotary screw traps located throughout the Upper Salmon upstream of Deadwater Slough.
 
@@
 -->
+
+```{r adult-impacts}
+# pull in SAR numbers
+sar = read_csv(here("analysis/data/raw_data", 
+                    "WSSCh_sar_4See.csv")) %>%
+  summarise(across(c(medsar, sdsar),
+                   median)) %>%
+  summarise(across(c(medsar, sdsar),
+               ~ . / 100))
+
+# summarize potential Chinook eaten for each time period by one pikeminnow
+adult_impact = fall_results %>%
+  summarise(across(ends_with("g_n"), max)) %>%
+  add_column(season = "Fall",
+             .before = 0) %>%
+  bind_rows(spring_results %>%
+              summarise(across(ends_with("g_n"), max)) %>%
+              add_column(season = "Spring",
+                         .before = 0)) %>%
+  pivot_longer(cols = ends_with("g_n"),
+               names_to = "perc_diet",
+               values_to = "juv_chnk") %>%
+  mutate(across(perc_diet,
+                str_extract,
+                "[:digit:]+"),
+         across(perc_diet,
+                as.integer)) %>%
+  left_join(spring_abund %>%
+              select(N_F:Uci_F,
+                     N_S:Uci_S) %>%
+              pivot_longer(everything()) %>%
+              mutate(season = if_else(str_detect(name, "_F$"),
+                                      "Fall",
+                                      "Spring")) %>%
+              mutate(across(name,
+                            str_remove,
+                            "_F$"),
+                     across(name,
+                            str_remove,
+                            "_S$")) %>%
+              pivot_wider(names_from = name,
+                          values_from = value)) %>%
+  mutate(smolt_N = N * juv_chnk,
+         smolt_Lci = Lci * juv_chnk,
+         smolt_Uci = Uci * juv_chnk) %>%
+  bind_cols(sar) %>%
+  mutate(adult_N = smolt_N * medsar,
+         adult_Lci = smolt_Lci * qnorm(0.025, medsar, sdsar),
+         adult_Uci = smolt_Uci * qnorm(0.975, medsar, sdsar))
+
+# sum fall and spring impacts
+total_impacts = adult_impact %>%
+  group_by(perc_diet) %>%
+  summarise(across(c(ends_with("_N"),
+                     ends_with("_Lci"),
+                     ends_with("_Uci")),
+                   sum))
+
+```
 
@@
 
-***Finish spring abundance results.*** <!-- Can we assume equal capture probability and perhaps using the ratio of the average fall CPUE and spring CPUE.
-Then multiply the average fall abundance estimate by the ratio.
-Does that seem appropriate?
--->
+All the estimators yielded estimates with overlapping confidence intervals.
+Because the sampling design most closely matched a multiple census estimator, we feel the Schnabel estimator is the most appropriate.
+We examined diagnostic plots for that estimator (number of marks vs.
+proportion of recaptures) and they suggested that we were meeting the assumptions of the Schnabel estimator.
+Therefore, further results will be based on the Schnabel estimates of abundance.
+
+Using the mean of the Schnabel estimates of Northern Pikeminnow abundance across two fall sampling events, and the mean of the CPUE from those events as well, we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95%CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring of 2020.
 
@@
 Given average weights of 10.3 g for DSR Chinook salmon and 10.9 g for NRR Chinook salmon, we would estimate that xx,xxx and xx,xxx juvenile Chinook salmon could be consumed during the fall and spring emigrations in Deadwater Slough, respectively.
+
+The estimated number of Chinook Salmon consumed by Northern Pikeminnow begins to taper off between 30-90% of Pikeminnow diet.
+So although we are uncertain of that diet proportion, the impact of that uncertainty is somewhat constrained.
+We estimated the total juvenile outmigrants consumed (fall and spring) to be between `r paste(prettyNum(round(range(total_impacts$smolt_N)), big.mark = ","), collapse = " and ")`.
+This translates to adult Chinook equivalents between `r paste(prettyNum(round(range(total_impacts$adult_N)), big.mark = ","), collapse = " and ")` (Figure \@ref(fig:adult-imp-p)).
 
@@
 
-<!--If that assumption is not true and capture probability did differ, we believe that the capture probability  in the spring would be lower due to higher flows, which would lead to our spring abundance estimate to be biased low.
-Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 sampling event to be a conservative estimate.-->
+All estimators suggest a large number of Northern Pikeminnow occupying Deadwater Slough.
+<!-- This might be a good place to bring in some comparisons with Northern Pikeminnow densities from elsewhere to put our numbers in context.-->
+
+If the first mark-recapture assumption was violated and the population is actually open (potential immigration and emigration during the survey period), as long as the immigration and emigration rates are equal between marked and unmarked fsih, our estimators should still provide unbiased estimates of abundance.
+Given the size of the sampling area, and the length of the sampling period (two weeks).
+
+We are comfortable with assumptions 4 and 5 (no loss of marks and no mistaken or missed marks).
+Assumptions 2 and 3 indicate whether the second (and subsequent) surveys are random samples of the population.
+Angling methods may provide a size selection bias, but that may only affect the interpretation of the abundance estimates (i.e., our estimate are the abundance of Northern Pikeminnow above some size threshold).
+Unequal catchability between individual fish is very difficult to assess, especially if that heterogeneity arises from being caught once.
+
+<!-- If catching and marking fish makes them want to leve the Deadwater area (more than unmarked fish), that could be making our estimates biased high (due to fewer recaptures).
+I don't imagine this is happening, but does anyone else have concerns about that?
+-->
+
+The spring abundance estimate is smaller than either fall sampling event.
+However, our approach assumes equal capture probabilities betwee the fall and spring sampling events.
+If it does differ, we believe that the capture probability in the spring could be lower due to higher flows, which would lead to our spring abundance estimate being biased low.
+Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 sampling event to be a conservative estimate.
 
@@
 
-```{r, eval = F}
-# pull in SAR numbers
-sar = read_csv(here("analysis/data/raw_data", 
-                    "WSSCh_sar_4See.csv")) %>%
-  summarise(across(c(medsar, sdsar),
-                   median)) %>%
-  summarise(across(c(medsar, sdsar),
-               ~ . / 100))
-
-adult_impact = fall_only_no_growth %>%
-  filter(cum_cons_fish_g == max(cum_cons_fish_g)) %>%
-  select(cum_cons_fish_g) %>%
-  mutate(n_chnk_smolts = cum_cons_fish_g / 12) %>%
-  bind_cols(all_est %>%
-              select(event_name, 
-                     model,
-                     N,
-                     Lci,
-                     Uci)) %>%
-  mutate(across(c(smolt_N = N, 
-                  smolt_Lci = Lci, 
-                  smolt_Uci = Uci),
-                ~ . * n_chnk_smolts)) %>%
-  bind_cols(sar) %>%
-  mutate(adult_N = smolt_N * medsar,
-         adult_Lci = smolt_Lci * qnorm(0.025, medsar, sdsar),
-         adult_Uci = smolt_Uci * qnorm(0.975, medsar, sdsar))
-
-```
-
 To estimate the number of possible Chinook salmon that could be consumed, we divided the cumulative fish consumption in grams for an individual pikeminnow by the average weight of a Chinook emigrant to come up with a presmolt equivalency and then multiplied this by our abundance estimate of northern pikeminnow in Deadwater Slough, resulting in an estimate of the equivalent total number of Chinook salmon presmolts consumed.
@@
 We multiplied the total estimated consumed juvenile Chinook salmon by an estimate of Granite-to-Granite SARs [@McCann2019].
+
+Finally, we estimated the number of additional adult Chinook Salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough reach was reduced or eliminated by multiplying the total estimated consumed Chinook Salmon by an estimate of Granite-to-Granite smolt-to-adult survival rates (SARs) (`r round(sar$medsar, 5)`, SE: `r round(sar$sdsar, 5)`) [@McCann2019].
 
@@
 
-*	SAR results. Maybe a plot of multiple scenarios depending on gastric lavage results? And pop estimates if they vary in time.
+*	SAR results. Those results are GRA - GRA SARs, estimated from multiple years of PIT tag data. Seem pretty good to me.
+* Maybe a plot of multiple scenarios depending on gastric lavage results? And pop estimates if they vary in time.
+* Maybe pop estimates if they vary in time.
 
@@
 Standard errors are not available for the multiple-census estimators.") %>%
+  kable_styling()
+
+```
+
+\newpage
+
+```{r spring-N}
+spring_abund %>%
+  rowwise() %>%
+  mutate(`Avg. Fall N (CI)` = paste0(prettyNum(round(N_F), big.mark = ","),
+                                     " (",
+                                     paste0(prettyNum(round(c(Lci_F, Uci_F)), big.mark = ","),
+                                            collapse = ", "),
+                                     ")"),
+         `Spring N (CI)` = paste0(prettyNum(round(N_S), big.mark = ","),
+                                     " (",
+                                     paste0(prettyNum(round(c(Lci_S, Uci_S)), big.mark = ","),
+                                            collapse = ", "),
+                                     ")")) %>%
+  ungroup() %>%
+  select(Species = species,
+         `Avg Fall CPUE` = cpue_F,
+         `Spring CPUE` = cpue_S,
+         ends_with("(CI)")) %>%
+  kable(booktabs = T,
+        caption = "Average CPUE and abundance in the fall (95% confidence interval), and CPUE and estimates of spring abundance (95% confidence interval) for Northern Pikeminnow in Deadwater Slough.") %>%
   kable_styling()
@@
 
+```{r adult-imp}
+adult_impact %>%
+  mutate(stage = "Smolt") %>%
+  select(#event_name, 
+         #model,
+         stage,
+         starts_with("smolt")) %>%
+  rename_with(.fn = ~ str_remove(., "smolt_"),
+              .cols = starts_with("smolt")) %>%
+  bind_rows(adult_impact %>%
+              mutate(stage = "Adult Equivalent") %>%
+              select(#event_name, model,
+                     stage,
+                     starts_with("adult")) %>%
+              rename_with(.fn = ~ str_remove(., "adult_"),
+                          .cols = starts_with("adult"))) %>%
+  kable(booktabs = T,
+        format.args = list(big.mark = ","),
+        col.names = c("Stage",
+                      "N",
+                      "Lower 95% CI",
+                      "Upper 95% CI"),
+        digits = 0,
+        caption = "Estimates of smolts consumed during fall migration by Northern Pikeminnow, and the adult equivalent of those smolts based on SARs.") %>%
+  kable_styling()
+```
+
+\newpage
+
 # Figures
@@
        y = "Abundance (x 1,000)") +
-  theme_bw() +
+  #theme_bw() +
   theme(legend.position = "none",
@@
 
-# all_est %>%
-#   mutate(across(c(N, Lci, Uci),
-#                  ~ . / 1000)) %>%
-#   ggplot(aes(x = model,
-#              y = N,
-#              color = model)) +
-#   geom_errorbar(aes(ymin = Lci,
-#                     ymax = Uci),
-#                 width = 0.1) +
-#   geom_point(size = 4) +
-#   facet_wrap(~ event_name) +
-#   scale_color_brewer(palette = "Set1") +
-#   coord_cartesian(ylim = c(0, 100)) +
-#   labs(x = "Estimator",
-#        y = "Abundance (x 1,000)") +
-#   theme(legend.position = "none",
-#         axis.text.x = element_text(angle = 45,
-#                                    hjust = 1))
-
-```
-
-\newpage
-
-```{r fall-bio-fig, eval = F, fig.cap = "First bioenergetics plot."}
-fall_bio_p = fall_only_no_growth %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Scumulative fish eaten (g)",
-                     breaks = seq(0, 200, by = 10),
-                     sec.axis = sec_axis(~. / 12, 
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 20, by = 1))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)")
-
-fall_bio_p
-
-```
-
-\newpage
-
-```{r full-yr-ng-fig, eval = F, fig.cap = "Next bioenergetics plot."}
-full_yr_ng_p = full_yr_no_growth %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  #scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Cumulative fish eaten (g)",
-                     breaks = seq(0, 600, by = 50),
-                     sec.axis = sec_axis(~. / 12,
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 30, by = 2))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)",
-       x = "Month")
-
-full_yr_ng_p
-
-```
-
-```{r full-yr-10p-fig, eval = F, fig.cap = "Another bioenergetics plot."}
-full_yr_10p_p = full_yr_10p_growth %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  #scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Cumulative fish eaten (g)",
-                     breaks = seq(0, 600, by = 50),
-                     sec.axis = sec_axis(~. / 12,
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 40, by = 2))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)",
-       x = "Month")
-
-full_yr_10p_p
-
-```
-
-```{r alt-mod-fig, eval = F, fig.cap = "Another bioenergetics plot, again."}
-alt_mod_p = alternative_mod %>%
-  ggplot() +
-  aes(x = date,
-      y = cum_cons_fish_g,
-      colour = temperature_c) +
-  geom_line(size = 2.5) +
-  scale_color_distiller(palette = "Spectral",
-                        direction = -1) +
-  #scale_x_date(breaks = "2 week") +
-  scale_y_continuous(name = "Cumulative fish eaten (g)",
-                     breaks = seq(0, 600, by = 50),
-                     sec.axis = sec_axis(~. / 12,
-                                         name = "Potential number of Chinook Salmon smolts",
-                                         breaks = seq(0, 30, by = 2))) +
-  theme_classic() +
-  theme(legend.position = "top") +
-  labs(color = "Temperature (C)",
-       x = "Month")
-
-alt_mod_p
-
-```
+```
+
+\newpage
 
@@
 
-```{r adult-imp-p, eval = F, fig.cap = "The adult impacts plot."}
+```{r adult-imp-p, fig.cap = "How many Chinook Salmon are consumed by Northern Pikeminnow, depending on what percent of diet is Chinook. Faceted by migration season and life-stage."}
 adult_impact %>%
-  mutate(stage = "Smolt") %>%
-  select(event_name, model,
-         stage,
+  mutate(stage = "Juv. Emigrant") %>%
+  select(stage,
+         season,
+         perc_diet,
          starts_with("smolt")) %>%
@@
                           .cols = starts_with("adult"))) %>%
-  filter(str_detect(model, "Schuma", negate = T)) %>%
-  ggplot(aes(x = model,
+  ggplot(aes(x = perc_diet,
              y = N,
-             color = model)) +
+             color = as.factor(perc_diet))) +
   geom_errorbar(aes(ymin = Lci,
                     ymax = Uci),
-                width = 0.5) +
+                width = 0.1) +
   geom_point(size = 4) +
-  facet_grid(stage ~ event_name,
+  facet_grid(fct_rev(stage) ~ season,
              scales = "free_y") +
-  scale_color_brewer(palette = "Set1",
-                     name = "Model") +
+  scale_color_viridis_d(name = "Percent Chinook\nin Diet") +
   theme(axis.text.x = element_blank(),
-        legend.position = "bottom") +
-  labs(x = "N. Pikeminnow Abundance Model")
+        legend.position = "bottom",
+        axis.title.x = element_blank(),
+        axis.ticks.x = element_blank()) +
+  labs(y = "Number Chinook Consumed")
+
 ```

### 2021-12-22 analysis/paper/paper.Rmd
added TBs bio plot

@@
 library(english)
+library(lemon)
+library(gridExtra)
 
@@
 
-\newpage
-
 ```{r length-freq-p, eval = T, fig.cap = "A length frequency plot."}
@@
 length_freq_p
+
+```
+
+\newpage
+
+```{r bio-p, fig.cap = "Temp."}
+spring_results$season = "Spring"
+fall_results$season = "Fall"
+bio_results = rbind(spring_results, fall_results)
+bio_results$ymd = (ymd(bio_results$date))
+
+pFish = bio_results %>%
+  ggplot(aes(ymd, f50_g)) +
+  geom_line(size = 1.15) +
+  geom_ribbon(data = bio_results,
+              aes(ymd,
+                  ymin = f10_g,
+                  ymax = f90_g),
+              color = "grey65",
+              alpha = 0.4) +
+  theme_bw() +
+  labs(y = "Cumulative fish consumed (g)",
+       x = "",
+       title = "") +
+  facet_rep_wrap(~season,
+                 scales = "free_x",
+                 repeat.tick.labels = F) +
+  theme(strip.background = element_rect(color = "white",
+                                        fill = "white"),
+        strip.text.x = element_text(size = 12),
+        axis.text.x = element_blank(),
+        axis.text.y = element_text(size = 11,
+                                   color = "black"),
+        axis.title.y = element_text(vjust = 1,
+                                    size = 12),
+        panel.grid.major.x = element_blank(),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        plot.margin = unit(c(-.1, .3, -.3, .2), "cm"))
+
+pTemp = bio_results %>%
+  ggplot(aes(ymd, temperature_c)) +
+  geom_line(color = "navyblue",
+            size = 1.15) +
+  theme_bw() +
+  labs(y = expression("Water temperature"*~degree*C),
+       x = "Month",
+       title = "") +
+  facet_rep_wrap(~season,
+                 scales = "free_x",
+                 repeat.tick.labels = F) +
+  theme(strip.background = element_rect(color = "white",
+                                        fill = "white"),
+        strip.text.x = element_blank(),
+        axis.text.x = element_text(size = 11,
+                                   color = "black"),
+        axis.title.x = element_text(size = 12),
+        axis.text.y = element_text(size = 11,
+                                   color = "black"),
+        axis.title.y = element_text(vjust = 0.1,
+                                    size = 12),
+        panel.grid.major.x = element_blank(),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        plot.margin = unit(c(-.1, .3, .1, .1), "cm"))
+
+p_list = list(pFish, pTemp)
+n_col = floor(sqrt(length(p_list)))
+
+bio_p = do.call("grid.arrange",
+                c(p_list, ncol = n_col))
 

### 2021-12-22 analysis/paper/paper.Rmd
draft Intro complete

@@
 ---
-title: "Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations in the Upper Salmon River, Idaho"
+title: "Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations"
 author:
@@
 keywords: |
-  northern pikeminnow; Chinook salmon; predation; mark-recapture; bioenergetics
+  Northern Pikeminnow; Chinook Salmon; predation; mark-recapture; bioenergetics
 highlights: |
@@
 
-The Snake and Columbia rivers drain a network of inland tributaries that provide essnetial spawning and rearing habitat for anadromous Pacific salmon *Oncorhynchus* spp.
+The Snake and Columbia rivers drain a network of inland tributaries that provide essential spawning and rearing habitat for anadromous Pacific salmon *Oncorhynchus* spp.
 and steelhead *O.
@@
 These fishes were historically abundant throughout the drainage basin and consequently have tremendous ecological, cultural, and economic value.
-Over the last century, salmonid stocks throughout the Snake and Columbia rivers have undergone significant declines related to habitat modification in the freshwater environment (e.g., removal of beavers, river channel simplification, water withdrawals, hydrosystem development) and poor marine conditions, both affecting their survival and recruitment and multiple life stages.
+Over the last century, salmonid stocks throughout the Snake and Columbia rivers have undergone significant declines related to habitat modification in the freshwater environment (e.g., removal of beavers, river channel simplification, water withdrawals, hydrosystem development) and poor marine conditions, both affecting their survival and recruitment at multiple life stages.
 Consequently, action agencies have made considerable investments in the rehabilitation of tributary ecosystems and evaulation of factors attributed to the species decline [@Roni2018].
@@
 One potentially important, but perhaps under-appreciated source of salmonid mortality is predation on emigrating juveniles by piscivorous fishes.
-In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recover [@Fresh2003; @Winther2020].
+In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recovery [@Fresh2003; @Winther2020].
 Dams and reservoirs in the Columbia River are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995].
@@
 
-Most predation studies have focused on reservoirs [@Murphy2021] and mainstem reaches [@Zimmerman1999; @Tabor1993; @Ward1995; @Shively1996] of the Columbia River basin whereas very little is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems [@Rubenson2020].
+Most predation studies have focused on reservoirs [@Murphy2021] and mainstem reaches [@Zimmerman1999; @Tabor1993; @Ward1995; @Shively1996] of the Columbia River basin whereas considerably less is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems [@Rubenson2020].
 Upstream habitats containing slower water velocities and other attributes that support piscivorous predators may overlap with essential habitat for some salmonid species and life stages.
@@
 The Salmon River was historically the most productive tributary for spring/summer Chinook Salmon in the Columbia River basin [@Nemeth1999], but those populations have become depleted in recent decades.
-Currently, the Upper salmon MPG supports eight independent, extant spring/summer Chinook Salmon populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River (@NOAA2017).
+Currently, the Upper Salmon MPG supports eight independent, extant spring/summer Chinook Salmon populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
 The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014].
@@
 
-Several studies have attempted to examine the downstream movement, distribution, and (apparent) survival of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach.
+Recent studies have attempted to examine the downstream movement, distribution, and (apparent) survival of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach.
 Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have substantially reduced survival in the Deadwater Slough reach compared to adjacent reaches [@Axel2015].
@@
 
-In this study, we estimated the abundance of Northern Pikeminnow in the Deadwater Slough and its potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
-We hypothesize that increase densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon and Sockeye Salmon.
+In this study, we estimated the abundance of Northern Pikeminnow in the Deadwater Slough and evaluated potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
+We hypothesize that high densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon and Sockeye Salmon.
 To test this, our objectives were three-fold:
@@
 
-<!--
-The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations, each providing important ecological, social, and economic benefits to the region.
-The distinct populations include Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017], all of which have become depleted in recent decades.
-Reduction in overall population size, and more specifically, declines in survival of juvenile Chinook Salmon have been associated with hydrosystem passage [@Schaller2014], ocean conditions, regional climate [@Welch2021; @Gosselin2021], and downstream predation by native and non-native piscivorous fishes [@Poe1991; @Friesen1999; @Harvey2004].
-Anthropogenic changes to freshwater habitat including removal of beavers, river channel simplification, water withdrawals, and the proliferation of non-native species (e.g., non-native coastal rainbow trout *O.
-mykiss irideus* and brook trout *Salvelinus fontinalis*) have also been attributed to reductions in population size [@IdahoOSCTeam2019].
-Moreover, the abundance of returning adults is impacted by ocean and mainstem harvests, poor ocean conditions, and changes to the spawning migration corridor [@NMFS2020BiOp].
-Each of these factors have contributed, to varying extents, to reduced adult escapement, the primary metric used to assess population viability.
-In response to the decline in Upper Salmon Chinook Salmon abundance from the myriad of human activities and associated habitat degradation, action agencies have attempted to improve juvenile survival and adult spawning conditions by investing in rehabilitation of tributary ecosystems as well as investing in the study and evaluation of many factors attributed to the species decline [@Roni2018].
-
-One potentially important, but perhaps under-appreciated source of mortality for Upper Salmon River Chinook Salmon is predation on emigrating juveniles by piscivorous fishes.
-In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recovery, most notably in the presence of non-native piscivores [@Fresh2003; @Winther2019].
-However, these studies have focused on large rivers and reservoirs of the Snake and Columbia rivers whereas very little is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems of the Columbia and Snake river [@Rubenson2020], including the Salmon River.
-Adding to this, there is little consensus within the literature about the magnitude of predation on salmonids in less altered, free flowing systems, with some suggesting lower rates of predation [@Brown1981; @Buchanan1981] while others suggest more impactful rates of predation [@Tabor1993; @Ward1995; @Shively1996; @Zimmerman1999] that may hinder recovery.
-
-Dams and reservoirs in the Columbia River basin are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995].
-There are generally two mechanisms that explain these high predation zones.
-First, movement (swimming) rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation.
-Second, reservoirs and downstream tailraces resulting from dams on the Columbia and Snake rivers have created favorable slow-water habitat for several fishes known to consume juvenile Chinook Salmon including both non-native (e.g., Smallmouth Bass *Micropterus dolomieu*, Walleye *Sander vitreus*) and native (e.g., Northern Pikeminnow *Ptychocheilus oregonensis*) species.
-In Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of Smallmouth Bass during some seasons [@Erhardt2018].
-Similarly, non-native Walleye are considered a potentially important source for juvenile Chinook Salmon mortality in the Snake River [@Widener2021], particularly as their abundance has significantly increased surrounding Snake River dam facilities between 2008 and 2020.
-
-Along with non-native species, the Columbia and Salmon rivers support native Northern Pikeminnow, a piscivorous fish tolerating and thriving in relatively warm, slow water habitats.
-Consequently, they too have benefited from dams on the Columbia River and become abundant predators on salmonid outmigrants [@Knutsen1999].
-Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996].
-Predation by piscivorous fishes, including Northern Pikeminnow, may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@Widener2021].
-
-While significant effort has been devoted to understanding predation-related mortality throughout the hydropower systems of the lower Snake and Columbia rivers, there is a dearth of contemporary data on piscivorous predator populations and little is known about their movement, life history, or impacts upstream of the mainstem habitats [@Gadomski2001].
-Consequently, the interaction between upstream, river-dwelling piscivorous fishes and emigrating juvenile salmonids, and its impact on salmonid recovery above reservoir systems of the Columbia and Snake rivers is unknown [@Rubenson2020].
-
-Habitats containing slower water velocities and other attributes that support piscivorous predators are not limited to areas near dams and reservoirs.
-One such example is the Deadwater Slough, an approximately 1.5 kilometers long reach of the Salmon River, Idaho (Figure \@ref(fig:map)).
-The slough is located approximately 6 kms downstream from the town of North Fork, Idah with the downstream end located at the confluence with Dump Creek.
-Around 1987, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River, forming an alluvial fan.
-The influx of sediment created an unnaturally slow and deep section of the river, resulting in "reservoir-like" conditions favorable to predators like Northern Pikeminnow and Smallmouth Bass [@Watkins2015].
-Furthermore, the slough lacks hydrological and structureal features that would provide juvenile fishes essential refuge from predation (e.g., large wood or substrate, bank undercuts).
-As a result, predation of juvenile Chinook Salmon and other local ESA-listed salmonids (Sockeye Salmon *O.
-nerka*, redband steelhead *O.
-mykiss gairdneri*) has been cited as a concern for populations in the Upper Salmon River.
-
-The Deadwater Slough is in a reach of the Salmon River believed to be a historically important overwinter rearing and spring emigration area for Chinook Salmon juveniles.
-Spring/summer Chinook Salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a].
-The DSR migrants leave the natal spawning area as subyearlings between June and November and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts.
-Alternatively, NRR migrants remain in their natal spawning areas for approximately one year after emergence until emigration to the ocean as smolts.
-Diversity of migratory tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency.
-Because the Deadwater Slough both supports rearing for DSR migrants and provides the migration pathway for NRR migrants, it represents a significant habitat for the Upper Salmon MPG.
-
-The slough is also part of the migratory pathway for the endangered Snake River Sockeye Salmon population [@Axel2015] and several populations of threatened Snake River steelhead.
-Accordingly, the mainstem Salmon River has been the target of recent radio telemetry studies to examine the downstream movement, distribution, and (apparent survival) of juvenile salmonid emigrants.
-For example, @Axel2015 examined the movement and survival of Sockeye Salmon released in Redfish Lake Creek downstream through the mainstem Salmon River during the spring migration and identified a substantial decrease in estimated survival in the reach spanning the Deadwater Slough.
-They estimated survival among release groups to be approximately 56% whereas survival through the majority of study reaches was greater than 80%.
-Further, recent telemetry studies of Chinook Salmon emigration during the fall and early winter months indicated a lower transition probability in the reach spanning the Deadwater Slough, approximately 10% less than surrounding reaches [@Ackerman2018; @Porter2019].
-
-In this study, we estimated the abundance of a piscivorous fish population (Northern Pikeminnow) in the Deadwater Slough and its potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
-We hypothesize that increased densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon [@Ackerman2018] and Sockeye Salmon [@Axel2015].
-To test this, our objectives for the study were three-fold:
-
-1. Estimate the abundance (or relative abundance) of potential predators in the Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations.
-2. Document predation on juvenile Chinook Salmon during the emigration periods using gastric lavage.
-3.
-Use a bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population at Deadwater Slough and the percent of consumption consisting of juvenile DSR and NRR Chinook salmon during defined emigration periods.
-
-This research is an essential first step in characterizing conditions through the Deadwater Reach and providing data necessary to discuss potential predation mitigation strategies (if needed), where Northern Pikeminnow removal has proven fruitful in increasing survival through the reservoir systems [@Winther2019].
-We follow with a discussion of the various assumptions that went into the mark-recapture and bioenergetics models and how violations of some assumptions may affect overall results and inferences from the study.
-We also consider how consumption of juvenile Chinook Salmon by predators at Deadwater Slough might impact adult returns, hindering recovery of populations in the Upper Salmon River MPG.
--->
+We also consider how consumption of juvenile Chinook Salmo emigrants at Deadwater Slough by Northern Pikeminnow may impact adult returns to the Upper Salmon MPG.
 

### 2021-12-22 analysis/paper/paper.Rmd
draft Methods complete

@@
 
-The Deadwater Slough is an approximately 1.5-kilometer section of the mainstem Salmon River located roughly 5.8 river kilometers downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
+The Deadwater Slough is an approximately 1.5 km section of the mainstem Salmon River located approximately 6 river km downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
 The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
@@
 Predators were sampled near the peaks of the juvenile DSR and NRR emigrations.
-Our intent was to sample during the emigration of the 2018 brood year (fall 2019 and spring 2020).
+Our initial intent was to sample during the emigration of the 2018 brood year (fall 2019 and spring 2020).
 However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 effort was delayed until fall 2020.
 Although this provided the benefit of an additional abundance estimate during the DSR emigration, estimates during the NRR emigration were still lacking, and so an additional effort was added in spring 2021.
-Fall efforts included sampling across a two-week span; however, due to budget constraints the spring 2021 effort was conducted over a single week.
+Fall efforts included sampling across a two-week span, but due to budget constraints the spring 2021 effort was limited to a single week.
 Fall efforts were constrained within a two-week window to minimize predator emigration/immigration and to constrain the population for the mark-recapture models.
@@
 
-Multiple capture methods were employed during the fall 2019 effort to reduce selectivity and bias for species and size classes.
+Multiple capture methods were employed during the initial fall 2019 effort to reduce selectivity and bias for species and size classes.
 Methods included raft electrofishing, fyke netting, snorkeling, and angling.
-However, ineffective methods were abandoned in subsequent years.
+Ineffective methods were then abandoned in subsequent years.
 <!-- For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) our inability to "chase" fish downstream as they evaded the electricity.
@@
 Additionally, there were concerns that raft electrofishers and fyke nets could incidentally capture ESA-listed adult steelhead that are present in the mainstem Salmon River during fall and spring.
--->After evaluating all methods, angling proved to be the most effective method for capturing piscine predators while minimizing potential impacts to ESA-listed steelhead.
-<!--, but recognizing that angling can have both species and size selectivity (bias).--> Therefore, the following analyses will focus on fish captured by angling, unless otherwise noted.
+-->After evaluating all methods, angling proved to be the most effective method for capturing piscine predators while also minimizing potential impacts to ESA-listed adult steelhead present in the Salmon River during our sampling efforts.<!--, but recognizing that angling can have both species and size selectivity (bias).--> Therefore, the following analyses will focus on fish captured by angling, unless otherwise noted.
 During the fall 2019 effort it additionally became clear that Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and so is the focus of our study.
@@
 
-Four different mark-recapture abundance estimators were employed to explore variation in estimates among them.
-Estimators fell into two broad categories: single-census and multiple-census.
-For the single census estimators, we treated the first week of sampling as the mark event, and the following (second) week as the recapture week, pooling data within each of those weeks.
-Alternatively, the multiple census estimators treat each day as a sampling occassion, and use information about the total marked fish from all previous occassions to infer the total abundance.
-
-For the single-census estimators, we first used the Lincoln-Peterson estimator:
+Four different mark-recapture estimators were employed to explore variation in abundance estimates among them.
+Estimators fell into two broad categories: single census and multiple census.
+For the single census estimators, we treated the first week as the first (mark) sample, and the following (second) week as the second (recapture) sample, pooling data within each of those weeks.
+Alternatively, the multiple census estimators treat each day as a sample and use information about the total marked fish from all previous samples to infer the total abundance.
+
+For the single census estimators, we first used the Lincoln-Peterson estimator:
 
@@
 $$
-where $M$ is the number is the total number of fish marked and returned to the population during the first week of sampling (mark event), $n$ is the total number of fish caught in the second week (recapture event), and $m$ is the number of marked fish caught during the recapture event.
+where $M$ is the total number of fish marked and returned to the population during the first sample (week), $n$ is the total number of fish caught in the second sample (week), and $m$ is the number of marked fish caught during the second week.
 Because the Lincoln-Peterson estimator can be biased with small samples, we also employed the Chapman-modified Lincoln-Peterson estimator:
@@
 $$
-For a multiple-census estimator, we first utilized the Schnabel method:
+For a multiple census estimator, we first utilized the Schnabel method:
 
@@
 
-where $M$, $n$, and $m$ are indexed by the sampling occassion, $i$.
-Here, $i$ is defined as each sampling day.
-The Schnabel estimator does not have an associated standard error; however, the 90% confidence intervals were calculated.
+where $M$, $n$, and $m$ are indexed by each sample, $i$.
+Here, sample $i$ is defined as each day.
+The Schnabel estimator does not have an associated standard error; however, the 95% confidence intervals were calculated.
 Lastly, we explored one additional multiple-census estimator, the Schumacher-Eschmeyer, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
@@
 
-The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and fall 2020 sampling periods.
-However, to estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of total CPUE in the fall sampling events to the total CPUE in the spring sampling event, then multiplied that ratio by the average abundance from the fall sampling events.<!-- Kevin, is this previous statement accurate?
---> This approach assumes equal capture probabilities between the fall and spring capture events.
-
-All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the sampling period), (2) all fish have equal chance of being caught in the second (and subsequent) sample(s), (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, (5) marks are not missed or mistaken.
+The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and 2020 sample efforts (the estimate is of the initial population size).
+However, to estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of total CPUE in the fall efforts to the total CPUE in the spring efforts, then multiplied that ratio by the average abundance from the fall efforts.
+This approach assumes equal capture probabilities between fall and spring efforts.
+
+All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the sampling period), (2) all fish have equal chance of being caught in the second (and subsequent) sample(s), (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or mistaken.
 
@@
 Immediately following lavage, stomach contents of individuals were preserved with 99% isopropyl alcohol in whirl-paks for later analysis in a controlled environment.
-For each sample, wet weight (grams) was recorded for the total combined stomach content, including all non-fish items (e.g., macroinvertebrates, organic matter), as well as the portion consisting of fish parts/matter.
+For each sample, wet weight (grams) was recorded for the total combined stomach content, including all non-fish items (e.g., macroinvertebrates, organic matter), as well as that portion consisting of fish parts or matter.
 Fish and fish remnants were identified to the lowest taxonomic unit using diagnostic bones, or were categorized as unknown [@Hansel1988; @Frost2000].
@@
 Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon [@Moss2016].
-We used a weight-length formula [@Parker1995] to calculate the predator start weight using the average length of Northern Pikeminnow captured in Deadwater Slough during our study.
-The average length of Northern Pikeminnow during the fall and spring were 352.9 and 393.7 mm, respectively, which calculated to average starting weights of 504.4 and 670.5 g.
-The assumption of zero growth was included to estimate consumption if individual Northern Pikeminnow were not growing and/or the population is stable, and so start and end weights are equal.
+The predator start weight was calculated using a weight-length formula [@Parker1995] and the average length of Northern Pikeminnow caught in Deadwater Slough during our study.
+The average length of Northern Pikeminnow caught during the fall and spring efforts were 352.9 and 393.7 mm, respectively, which calculated to average starting weights of 504.4 and 670.5 g.
+We assumed no growth in individual Northern Pikeminnow, and so predator start and end weights are equal, i.e., we assumed the population biomass was stable.
 Mean daily water temperatures were summarized from 15-minute interval temperature readings available March 3, 2013 to June 14, 2021 from a gage station approximately 22 river kilometers downstream of Deadwater Slough near Shoup, Idaho [@USGS2016].
 
-Our largest uncertainty for model parameters was the proportion of the Northern Pikeminnow diet consisting of fishes (e.g., juvenile Chinook Salmon) versus other food items like invertebrates.
-Therefore, we varied the proportion of the diet that was fish or invertebrates ranging from 10-90%, in 10% intervals, for both categories.
+The largest uncertainty among model parameters was the proportion of the Northern Pikeminnow diet consisting of fishes (e.g., juvenile Chinook Salmon) versus other food items (e.g., invertebrates).
+Therefore, we varied the proportion of the diet consisting of fish versus invertebrates to range from 10-90%, in 10% intervals, for both categories.
 Energy densities for invertebrates was fixed at 3,000 J/g.
 Separate models were also run for a 77-day fall period from September 15 - November 30 and a 92-day period in the spring from March 1 - May 31 to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG.
-This resulted in 18 separate model runs (2 seasons and 9 diet scenarios).
+This resulted in 18 separate model runs (9 diet scenarios and 2 seasons).
 Each model run provided an estimate of the amount (grams) of fish consumed by an individual Northern Pikeminnow during the fall or spring time periods.
@@
 
-To estimate the total biomass of fish potentially consumed by Northern Pikeminnow in Deadwater Slough, we then multiplied the biomass consumed by an individual Northern Pikeminnow for the timer period by the estimated population size for that time period (mean Schnabel estimates in the fall, CPUE ratio generated estimate in the spring).
-The total biomass was further converted to an estimate of the total number of Chinook salmon consumed by dividing by a weight of `r avg_DSR_g` for the fall and `r avg_NRR_g` for the spring, an average from juvenile Chinook Salmon captured during those date ranges at seven rotary screw traps located throughout the Upper Salmon upstream of Deadwater Slough.
+To estimate the total biomass of DSR and NRR emigrants potentially consumed by all Northern Pikeminnow in Deadwater Slough, we then multiplied the grams of fish consumed by an individual Northern Pikeminnow  during each period (fall and spring) by the estimated population size for that period.
+The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
+The weight of an average DSR and NRR emigrants was taken from fish caught at seven rotary screw traps located throughout the Upper Salmon MPG upstream of Deadwater Slough during the fall and spring time periods.
 

### 2021-12-22 analysis/paper/paper.Rmd
draft abundance and gastric results complete

@@
 
-We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during Fall 2019, Fall 2020, and Spring 2021, respectively.
-The catch per unit effort, measured as the number of Northern Pikeminnow captured per angler hour, for each was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for Fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for Fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for Spring 2021 (Figure \@ref(fig:cpue-fig)).
-During Fall 2019, we had a total of `r as.english(sum(N_mods[[3]][[1]]$m))` recaptures which included `r as.english(LP_mod[[3]][[1]]$m)` unique individuals (i.e., `r as.english(sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m)` individual was captured more than once).
-There were `r as.english(sum(N_mods[[3]][[3]]$m))` recaptures in Fall 2020 which included `r as.english(LP_mod[[3]][[3]]$m)` unique individuals.
-Summaries of parameters for both the single-census and multiple-census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.
-
-Abundance estimates for Northern Pikeminnow in the Deadwater Slough during Fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for Fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")`.
+We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 efforts, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish))` fish caught using hook-and-line angling.
+The CPUE, measured as number of Northern Pikeminnow caught per angler hour, was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021 (Figure \@ref(fig:cpue-fig)).
+During the mark-recapture efforts, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019 which included `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals (i.e., `r as.character(english(sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m))` individual was captured more than once).
+There were an additional `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020 which included `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals.
+Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.
+
+Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` (Table \@ref(tab:abund-tab)).
 In both cases, the multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied (Figure \@ref(fig:abund-fig)).
-Table \@ref(tab:abund-tab) provides point estimates and confidence intervals for each estimator for the fall mark-recapture sampling events.
-
-All the estimators yielded estimates with overlapping confidence intervals.
-Because the sampling design most closely matched a multiple census estimator, we feel the Schnabel estimator is the most appropriate.
-We examined diagnostic plots for that estimator (number of marks vs.
-proportion of recaptures) and they suggested that we were meeting the assumptions of the Schnabel estimator.
-Therefore, further results will be based on the Schnabel estimates of abundance.
-
-Using the mean of the Schnabel estimates of Northern Pikeminnow abundance across two fall sampling events, and the mean of the CPUE from those events as well, we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95%CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring of 2020.
+Table \@ref(tab:abund-tab) provides point estimates and confidence intervals for each estimator for the fall mark-recapture efforts.
+
+All mark-recapture estimators yielded estimates with overlapping confidence intervals.
+However, because our sampling design most closely matched a multiple census estimator (i.e., we released marked fish each day and recaptures from all previous days were recorded for each day), we fell those estimators most appropriate.
+Additionally, because the Schnabel estimator includes a small sample size modification, as in our case (i.e., few recaptures), we feel it is most appropriate.
+The Schnabel estimator has the added benefit of more precise confidence intervals and is more straightforward than the Schumacher-Eschmeyer estimate as it is essentially the weighted average among multiple Chapman-modified Lincoln-Peterson estimates.
+Therefore, further analyses and results will be based on the Schnabel estimates of abundance.
+
+Using the mean of the Schnabel estimates of Northern Pikeminnow abundance across the two fall sampling events (`r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")`) and multiplying the ratio (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`) between the spring CPUE (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"], 2)`) and mean of the fall CPUEs (`r round(mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring 2021.
 
@@
 
-We examined the stomach contents of `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow over the course of the study using gastric lavage methods.
-Of those, we documented some form of stomach contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` of the fish and we were able to confirm fish or fish remnants in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")`.
-Expressed as a percentage, `r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% of lavaged Northern Pikeminnow had stomach contents of which `r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% we identified containing fish or fish remnants.
-Table \@ref(tab:gastric-tab) provides a further breakdown of gastric lavage results for each sampling event.
-
-For Northern Pikeminnow with some form of stomach contents, the average wet weight of total contents was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r prettyNum(min(stmch_df$stomach_contents_weight), digits = 3)`-`r prettyNum(max(stmch_df$stomach_contents_weight), digits = 3)` g).
-Overall, fish or fish remnants were `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall weight among all stomach contents examined.
+We examined the stomach contents of `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow from Deadwater Slough using gastric lavage methods.
+Of those, we found contents of some form in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) of those.
+Table \@ref(tab:gastric-tab) summarizes gastric lavage results for each sampling effort.
+For those individuals with stomach contents, the average wet weight of total contents was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r prettyNum(min(stmch_df$stomach_contents_weight), digits = 3)`-`r prettyNum(max(stmch_df$stomach_contents_weight), digits = 3)` g).
+Overall, fish or fish remnants consisted of `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall of all stomach contents examined.
 

### 2021-12-22 analysis/paper/paper.Rmd
draft Results complete

@@
 
+# average size of fall and spring pikeminnow
+avg_fall_NPM   = 352.9
+avg_spring_NPM = 393.7
+
 # average size of DSR and NRR juveniles
@@
 
+fall_g_summ = fall_results %>%
+  select(f10_g:f90_g) %>%
+  pivot_longer(f10_g:f90_g) %>%
+  group_by(name) %>%
+  summarize(cumm_g = max(value))
+
 rm(fall_f10, fall_f20, fall_f30, fall_f40, fall_f50, fall_f60, fall_f70, fall_f80, fall_f90)
@@
 
+spring_g_summ = spring_results %>%
+  select(f10_g:f90_g) %>%
+  pivot_longer(f10_g:f90_g) %>%
+  group_by(name) %>%
+  summarize(cumm_g = max(value))
+
 rm(spring_f10, spring_f20, spring_f30, spring_f40, spring_f50, spring_f60, spring_f70, spring_f80, spring_f90)
@@
 The predator start weight was calculated using a weight-length formula [@Parker1995] and the average length of Northern Pikeminnow caught in Deadwater Slough during our study.
-The average length of Northern Pikeminnow caught during the fall and spring efforts were 352.9 and 393.7 mm, respectively, which calculated to average starting weights of 504.4 and 670.5 g.
+The average length of Northern Pikeminnow caught during the fall and spring efforts were `r avg_fall_NPM` and `r avg_spring_NPM` mm, respectively, which calculated to average starting weights of 504.4 and 670.5 g.
 We assumed no growth in individual Northern Pikeminnow, and so predator start and end weights are equal, i.e., we assumed the population biomass was stable.
@@
 
-We estimated a single average-sized Northern Pikeminnow to consume xx.x g of fish during the modeled fall time period (September 15 - November 30) and an additional xx.x g of fish during the modeled spring time period (March 1 - June 31).
-Given estimated population sizes of xx,xxx Northern Pikeminnow during fall and xx,xxx during spring, that would convert to xxx,xxx - xxx,xxx g during fall and xxx,xxx g during spring.
-Given average weights of 10.3 g for DSR Chinook salmon and 10.9 g for NRR Chinook salmon, we would estimate that xx,xxx and xx,xxx juvenile Chinook salmon could be consumed during the fall and spring emigrations in Deadwater Slough, respectively.
-
-The estimated number of Chinook Salmon consumed by Northern Pikeminnow begins to taper off between 30-90% of Pikeminnow diet.
-So although we are uncertain of that diet proportion, the impact of that uncertainty is somewhat constrained.
-We estimated the total juvenile outmigrants consumed (fall and spring) to be between `r paste(prettyNum(round(range(total_impacts$smolt_N)), big.mark = ","), collapse = " and ")`.
-This translates to adult Chinook equivalents between `r paste(prettyNum(round(range(total_impacts$adult_N)), big.mark = ","), collapse = " and ")` (Figure \@ref(fig:adult-imp-p)).
+We estimated an individual, average-size (`r avg_fall_NPM` mm) Northern Pikeminnow to consume `r median(fall_g_summ$cumm_g)` g (range `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the DSR emigration, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
+Similarly, we estimated the average-size (`r avg_spring_NPM` mm) Northern Pikeminnow to consume `r median(spring_g_summ$cumm_g)` g (range `r min(spring_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
+Given an average weight of `r avg_DSR_g` g for DSR emigrants, that would equate to `r round(median(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` (range `r round(min(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)`) fish consumed by an individual during fall and an additional `r round(median(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` (range `r round(min(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)`) fish consumed using an average weight of `r avg_NRR_g` g for NRR emigrants.
+
+Given a mean population size of `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` from fall estimates and an estimate of `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` for spring 2021, and assuming the median value from diet scenarios (i.e, 50% of their diet consists of fish), we'd estimate that `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")`) DSR and `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")`) NRR juveniles may be consumed in Deadwater Slough during peak emigrations.
+Table \@ref(tab:adult-imp) shows the estimated number of juvenile Chinook Salmon consumed by and individual Northern Pikeminnow and by the Deadwater Slough population across all diet scenarios and for both the fall (DSR) and spring (NRR) peak emigration seasons.
+
+Across both seasons and all diet scenarios, we estimated the total juvenile outmigrants consumed to be between `r prettyNum(round(min(total_impacts$smolt_N), 0), big.mark = ",")` and `r prettyNum(round(max(total_impacts$smolt_N), 0), big.mark = ",")`.
+The estimated number of Chinook Salmon consumed by a Northern Pikeminnow begins to taper off betwen the 30-90% diet scenarios.
+So although we are uncertain of the diet proportions, the impact of that uncertainty is somewhat constrained.
 
@@
 We multiplied the total estimated consumed juvenile Chinook salmon by an estimate of Granite-to-Granite SARs [@McCann2019].
+
+This translates to adult Chinook equivalents between `r paste(prettyNum(round(range(total_impacts$adult_N)), big.mark = ","), collapse = " and ")` (Figure \@ref(fig:adult-imp-p)).
 

### 2021-12-23 analysis/paper/paper.Rmd
cleaned up tables and fig 1

@@
 
-Using the mean of the Schnabel estimates of Northern Pikeminnow abundance across the two fall sampling events (`r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")`) and multiplying the ratio (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`) between the spring CPUE (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"], 2)`) and mean of the fall CPUEs (`r round(mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring 2021.
+Using the mean of the Schnabel estimates of Northern Pikeminnow abundance across the two fall sampling events, `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`), and multiplying the ratio (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`) between the spring CPUE (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"], 2)`) and mean of the fall CPUEs (`r round(mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring 2021.
 
@@
 <!--
+## Juvenile Chinook Salmon Abundance
 For context, throughout the migration periods of DSR and NRR life histories it is estimated that 113,976 individuals are produced and therefore available within the upper Salmon River basin (Table x).
-
-Given an estimated population size of xx,xxx, we’d estimate that xxx,xxx g of NRR fish are consumed each spring during the emigration, and xxxxx.xx g of DSR fish are consumed in the fall during the emigration of juvenile Chinook Salmon.
-Therefore, Northern Pikeminnow in the Deadwater Slough can potentially consume xx,xxx juvenile Chinook Salmon as a discrete population.
--->
-
-<!--
-## Juvenile Chinook Salmon Abundance
 
@@
   distinct() %>%
+  select(-species) %>%
   unnest(data) %>%
@@
                 ~ str_replace(., "_", " "))) %>%
-  rename(`Sampling Event` = event_name,
-         Species = species) %>%
+  rename(`Effort` = event_name) %>%
   kable(booktabs = T,
         format.args = list(big.mark = ","),
-        caption = "Data summary for single-census mark-recapture estimators.
-M is the total number of fish caught in the first sample event, marked and returned to the population; n is the total number of fish caught in the second sample, and m is the number of previously marked fish caught in the second sample event.") %>%
+        caption = "Summary of Northern Pikeminnow angling survey data for single-census mark-recapture estimators.
+M is the total number of fish caught in the first sample (week), marked and returned to the population; n is the total number of fish caught in the second sample (week), and m is the number of previously marked fish caught in the second week.") %>%
   kable_styling()
@@
                 ~ str_replace(., "_", " "))) %>%
-  rename(`Sampling Event` = event_name,
-         Species = species,
+  select(event_name, date, M, n, m, u, R) %>%
+  rename(`Effort` = event_name,
          Date = date) %>%
@@
         format.args = list(big.mark = ","),
-        caption = "Data summary for multiple-census mark-recapture estimators.
-M is the total number of fish caught in previous events, marked and returned to the population; n is the total number of fish caught in the current event; m is the number of marked fish caught in that event; u is the number of unmarked fish caught, and R is the number of marked fish returned to the population.") %>%
+        caption = "Summary of Northern Pikeminnow angling survey data for multiple-census mark-recapture estimators.
+M is the total number of fish caught in previous samples (days), marked and returned to the population; n is the total number of fish caught in the current sample (day); m is the number of marked fish caught in that sample; u is the number of unmarked Northern Pikeminnow caught, and R is the number of marked fish returned to the population.
+Dates of sampling are also provided.") %>%
   kable_styling()
@@
   select(-c(species, range:prop_range)) %>%
-  rename(`Sampling Event` = event_name,
-         Estimator = model) %>%
+  rename(`Effort` = event_name,
+         Estimator = model,
+         `L CI` = Lci,
+         `U CI` = Uci) %>%
   kable(booktabs = T,
         format.args = list(big.mark = ","),
-        caption = "Point estimates, standard errors, and confidence intervals for each of the mark-recapture estimators and for the fall sampling events.
+        caption = "Point estimates, standard errors, and 95% confidence intervals for each of the mark-recapture estimators and for the fall sampling efforts.
 Standard errors are not available for the multiple-census estimators.") %>%
@@
 
-\newpage
-
-```{r spring-N}
+```{r spring-N, eval = F}
 spring_abund %>%
@@
 adult_impact %>%
-  mutate(stage = "Smolt") %>%
-  select(#event_name, 
-         #model,
-         stage,
-         starts_with("smolt")) %>%
-  rename_with(.fn = ~ str_remove(., "smolt_"),
-              .cols = starts_with("smolt")) %>%
-  bind_rows(adult_impact %>%
-              mutate(stage = "Adult Equivalent") %>%
-              select(#event_name, model,
-                     stage,
-                     starts_with("adult")) %>%
-              rename_with(.fn = ~ str_remove(., "adult_"),
-                          .cols = starts_with("adult"))) %>%
+  select(season,
+         perc_diet,
+         juv_chnk,
+         smolt_N,
+         smolt_Lci,
+         smolt_Uci,
+         medsar,
+         sdsar,
+         adult_N,
+         adult_Lci,
+         adult_Uci) %>%
+  mutate(juv_chnk = round(juv_chnk, 2),
+         smolt_N = round(smolt_N, 0),
+         smolt_Lci = round(smolt_Lci, 0),
+         smolt_Uci = round(smolt_Uci, 0),
+         medsar = round(medsar, 5),
+         sdsar = round(sdsar, 5),
+         adult_N = round(adult_N, 0),
+         adult_Lci = round(adult_Lci, 0),
+         adult_Uci = round(adult_Uci, 0)) %>%
+  rename(Effort = season,
+         `p Fish` = perc_diet,
+         `Juv Chinook` = juv_chnk,
+         `Juv N` = smolt_N,
+         `Juv L CI` = smolt_Lci,
+         `Juv U CI` = smolt_Uci,
+         `Median SAR` = medsar,
+         `SD SAR` = sdsar,
+         `Adult N` = adult_N,
+         `Adult L CI` = adult_Lci,
+         `Adult U CI` = adult_Uci) %>%
   kable(booktabs = T,
         format.args = list(big.mark = ","),
-        col.names = c("Stage",
-                      "N",
-                      "Lower 95% CI",
-                      "Upper 95% CI"),
-        digits = 0,
-        caption = "Estimates of smolts consumed during fall migration by Northern Pikeminnow, and the adult equivalent of those smolts based on SARs.") %>%
-  kable_styling()
+        caption = "Estimates of the number of DSR (fall) and NRR (spring) emigrants consumed for each season and diet scenario.
+The diet scenario indicates the proportion of the Northern Pikeminnow diet consisting of fish.
+The 'Juv Chinook' column indicates the number of juvenile Chinook consumed by an individual Northern Pikeminnow during the fall or spring seasons and is taken from each of the 18 bioenergetics model runs.
+The number of smolts consumed (and 95% confidence intervals) are calculated using Northern Pikeminnow abundance estimates of 28,144 (95% CI: 14,378 - 59,978) for fall and 14,897 (95% CI: 7,610 - 31,747) for spring.
+The number of 'adult equivalents' is also estimated using a Lower Granite Dam smolt-to-adult rate (SAR) from McCann et al.
+2019.") %>%
+  kable_styling(font_size = 8)
+
+# adult_impact %>%
+#   mutate(stage = "Smolt") %>%
+#   select(#event_name, 
+#          #model,
+#          stage,
+#          starts_with("smolt")) %>%
+#   rename_with(.fn = ~ str_remove(., "smolt_"),
+#               .cols = starts_with("smolt")) %>%
+#   bind_rows(adult_impact %>%
+#               mutate(stage = "Adult Equivalent") %>%
+#               select(#event_name, model,
+#                      stage,
+#                      starts_with("adult")) %>%
+#               rename_with(.fn = ~ str_remove(., "adult_"),
+#                           .cols = starts_with("adult"))) %>%
+#   kable(booktabs = T,
+#         format.args = list(big.mark = ","),
+#         col.names = c("Stage",
+#                       "N",
+#                       "Lower 95% CI",
+#                       "Upper 95% CI"),
+#         digits = 0,
+#         caption = "Estimates of smolts consumed during fall migration by Northern Pikeminnow, and the adult equivalent of those smolts based on SARs.") %>%
+#   kable_styling()
+
 ```
@@
 ```{r map, out.width = "85%", fig.cap = "Map of the Deadwater Slough study area.
-The high-resolution orthoimage portion directly surrounding the Deadwater Slough was generated from aerial images taken by a drone.
+The high-resolution orthoimage portion surrounding the Deadwater Slough was generated from aerial images taken by an unmanned aircraft system (drone).
 The red polygon indicates the reach characterized by unnaturally slow water velocities and a deepened channel.

### 2021-12-23 analysis/paper/paper.Rmd
cleaned up figures

@@
 
-```{r cpue-fig, fig.cap = "The number of Northern Pikeminnow caught per angler effort for each sampling event."}
+```{r cpue-fig, fig.cap = "The number of Northern Pikeminnow caught per angler effort for each sample effort."}
 total_cpue %>%
@@
   labs(y = "Catch Per Angler Hour") +
-  theme(axis.title.x = element_blank())
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_text(size = 11,
+                                   color = "black"),
+        axis.title.y = element_text(vjust = 3),
+        axis.text.y = element_text(size = 11,
+                                   color = "black"))
 
@@
   theme(legend.position = "none",
+        axis.title.x = element_blank(),
         axis.text.x = element_text(angle = 45,
-                                   hjust = 1)) +
+                                   hjust = 1,
+                                   size = 11,
+                                   color = "black"),
+        axis.text.y = element_text(size = 11,
+                                   color = "black"),
+        strip.text.x = element_text(size = 11,
+                                    color = "black")) +
   scale_color_brewer(palette = "Dark2")
@@
 
-```{r length-freq-p, eval = T, fig.cap = "A length frequency plot."}
+```{r length-freq-p, fig.cap = "Length frequency histogram of Northern Pikeminnow caught using hook-and-line angling during the study."}
 length_freq_p = gastric_df %>%
@@
   geom_histogram(breaks = seq(150, 650, by = 50),
-                 col = "#DB740E",
-                 fill = "#DB740E",
+                 col = "black",
+                 fill = "gray30",
                  alpha = .2) +
-  labs(x = "TL (mm)",
+  labs(x = "Total Length (mm)",
        y = "Frequency") +
@@
   scale_y_continuous(breaks = seq(0, 550, by = 50)) +
+  theme(axis.text.x = element_text(size = 11,
+                                   color = "black"),
+        axis.text.y = element_text(size = 11,
+                                   color = "black",
+                                   vjust = 3)) +
   theme_classic()
@@
 
-```{r bio-p, fig.cap = "Temp."}
+```{r bio-p, fig.cap = "The cumulative amount of fish (g) consumed by an individual Northern Pikeminnow during the peak DSR (fall) and NRR (spring) Chinook Salmon emigrations (top panel) and corresponding daily mean water temperatures (bottom panel).
+The black line is the median among diet scenarios (i.e., where 50% of the Northern Pikeminnow diet consists of fish) and the gray area shows the range among model scenarios (i.e., 10 - 90% of Northern Pikeminnow diet consists of fish).
+Daily mean water temperatures were summarized from six years of data available from the Shoup gage station.
+"}
 spring_results$season = "Spring"
@@
   theme_bw() +
-  labs(y = "Cumulative fish consumed (g)",
+  labs(y = "Fish consumed (g)",
        x = "",
@@
   theme_bw() +
-  labs(y = expression("Water temperature"*~degree*C),
+  labs(y = expression("Water temp"*~degree*C),
        x = "Month",
@@
                                    color = "black"),
-        axis.title.x = element_text(size = 12),
+        axis.title.x = element_blank(),
         axis.text.y = element_text(size = 11,
@@
 
-```{r adult-imp-p, fig.cap = "How many Chinook Salmon are consumed by Northern Pikeminnow, depending on what percent of diet is Chinook. Faceted by migration season and life-stage."}
+```{r adult-imp-p, fig.cap = "The estimated number of DSR (fall) and NRR (spring) emigrants consumed by Northern Pikeminnow in Deadwater Slough (top panel) across a range of diet scenarios where 10-90% of the diet consists of juvenile Chinook salmon.
+The bottom panel provides the adult equivalents assuming a Lower Granite Dam smolt-to-adult rate (SAR) from McCann et al.
+2019."}
 adult_impact %>%
@@
              scales = "free_y") +
-  scale_color_viridis_d(name = "Percent Chinook\nin Diet") +
+  scale_color_viridis_d(name = "Percent\nFish in Diet") +
   theme(axis.text.x = element_blank(),
@@
         axis.title.x = element_blank(),
-        axis.ticks.x = element_blank()) +
+        axis.ticks.x = element_blank(),
+        axis.text.y = element_text(size = 11,
+                                   color = "black"),
+        strip.text.x = element_text(size = 11,
+                                    color = "black"),
+        strip.text.y = element_text(size = 11,
+                                    color = "black")) +
   labs(y = "Number Chinook Consumed")

### 2021-12-23 analysis/paper/paper.Rmd
added clarification of fish versus invertebrates

@@
 
-The largest uncertainty among model parameters was the proportion of the Northern Pikeminnow diet consisting of fishes (e.g., juvenile Chinook Salmon) versus other food items (e.g., invertebrates).
+The largest uncertainty among model parameters was the proportion of the Northern Pikeminnow diet consisting of fish (e.g., juvenile Chinook Salmon) versus other food items (e.g., invertebrates).
 Therefore, we varied the proportion of the diet consisting of fish versus invertebrates to range from 10-90%, in 10% intervals, for both categories.
 Energy densities for invertebrates was fixed at 3,000 J/g.
+We estimate the g of (any) fish consumed for each model run.
+For simplicity, we assume that any fish prey have the same energy densities as juvenile Chinook Salmon and we do not differentiate juvenile Chinook Salmon from other fish prey than may be present in Deadwater slough (e.g., juvenile steelhead, juvenile Sockeye Salmon, Redside Shiner).
 Separate models were also run for a 77-day fall period from September 15 - November 30 and a 92-day period in the spring from March 1 - May 31 to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG.
@@
 
-To estimate the total biomass of DSR and NRR emigrants potentially consumed by all Northern Pikeminnow in Deadwater Slough, we then multiplied the grams of fish consumed by an individual Northern Pikeminnow  during each period (fall and spring) by the estimated population size for that period.
-The total biomass was further converted to an estimate of the total number of Chinook Salmon consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
+To estimate the total biomass of fish potentially consumed by all Northern Pikeminnow in Deadwater Slough, we then multiplied the grams of fish consumed by an individual Northern Pikeminnow  during each period (fall and spring) by the estimated predator population size for that period.
+The total biomass was further converted to an estimate of the total number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
 The weight of an average DSR and NRR emigrants was taken from fish caught at seven rotary screw traps located throughout the Upper Salmon MPG upstream of Deadwater Slough during the fall and spring time periods.
+This calculation assumes that all fish prey items consumed by Northern Pikeminnow are, on average, the same size as DSR and NRR emigrants.
 
@@
 
-We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 efforts, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish))` fish caught using hook-and-line angling.
+We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 efforts, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish caught using hook-and-line angling.
 The CPUE, measured as number of Northern Pikeminnow caught per angler hour, was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021 (Figure \@ref(fig:cpue-fig)).
@@
 
-Given a mean population size of `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` from fall estimates and an estimate of `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` for spring 2021, and assuming the median value from diet scenarios (i.e, 50% of their diet consists of fish), we'd estimate that `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")`) DSR and `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")`) NRR juveniles may be consumed in Deadwater Slough during peak emigrations.
-Table \@ref(tab:adult-imp) shows the estimated number of juvenile Chinook Salmon consumed by and individual Northern Pikeminnow and by the Deadwater Slough population across all diet scenarios and for both the fall (DSR) and spring (NRR) peak emigration seasons.
-
-Across both seasons and all diet scenarios, we estimated the total juvenile outmigrants consumed to be between `r prettyNum(round(min(total_impacts$smolt_N), 0), big.mark = ",")` and `r prettyNum(round(max(total_impacts$smolt_N), 0), big.mark = ",")`.
-The estimated number of Chinook Salmon consumed by a Northern Pikeminnow begins to taper off betwen the 30-90% diet scenarios.
-So although we are uncertain of the diet proportions, the impact of that uncertainty is somewhat constrained.
+Given a mean population size of `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` from fall estimates and an estimate of `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` for spring 2021, and assuming the median value from diet scenarios (i.e, 50% of their diet consists of fish), we'd estimate that Northern Pikeminnow can consume `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")`) fish during the DSR emigration and `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")`) fish during the NRR emigration in Deadwater Slough.
+Table \@ref(tab:adult-imp) shows the estimated number of fish consumed by an individual Northern Pikeminnow and by the population across all diet scenarios during both the fall (DSR) and spring (NRR) peak emigration seasons.
+
+Across both seasons and all diet scenarios, we estimated the total number of fish consumed to be between `r prettyNum(round(min(total_impacts$smolt_N), 0), big.mark = ",")` and `r prettyNum(round(max(total_impacts$smolt_N), 0), big.mark = ",")` (Table \@ref(tab:adult-imp)).
+The estimated number of fish consumed by a Northern Pikeminnow begins to taper off between the 30-90% diet scenarios.
+So although we are uncertain of the diet proportions consisting of fish versus invertebrates, the impact of that uncertainty is somewhat constrained.
 

### 2021-12-23 analysis/paper/paper.Rmd
updates to acknowledgements

@@
 
-<!-- Please feel free to add others that I missed or improve -->
-
-The authors extend much appreciation to the many volunteers who assisted with field efforts including collaborators from Bureau of Reclamation, Idaho Department of Fish and Game, and Lemhi Regional Land Trust, among others.
-This manuscript benefited from reviews and contributions from colleagues at the Idaho Governor's Office of Species Conservation, Rio Applied Science and Engineering, and from Sean Gibbs and Ben Briscoe at Mount Hood Environmental.
+The authors extend much appreciation to the many volunteers who assisted with field efforts including collaborators from the Bureau of Reclamation, Idaho Office of Species Conservation, Idaho Deparment of Fish and Game, and the Lemhi Regional Land Trust, among others.
+Special thanks to Chelsea Welke, Tulley Mackey, Jared Barker, and Brian Hamilton for their help in the field.
+Special thanks to Caitlin Alcott and Inter-Fluve for their administrative support and guidance.
 Funding for this study was provided by the Bureau of Reclamation, Pacific Northwest Regional Office (contract No.
-140R1021F0018).
-Special thanks to Caitlin Alcott and Inter-Fluve for their administrative support and guidance.
+140R1021F0018) with assistance from the Idaho Office of Species Conservation.
 

### 2021-12-28 analysis/paper/paper.Rmd
start Discussion edits

@@
 
+```{r deadwater-attr}
+# set some deadwater attributes
+ddw_rkms = 1.871
+ddw_m2 = 119431
+ddw_avg_width_m = 68
+ddw_acres = ddw_m2 / 4047
+ddw_hectares = ddw_m2 / 10000
+
+```
+
 # Introduction
@@
 Upstream habitats containing slower water velocities and other attributes that support piscivorous predators may overlap with essential habitat for some salmonid species and life stages.
-One such example is Deadwater Slough, an approximately 1.5 kilometer long reach of unnaturally slow and deep water in the Salmon River, Idaho.
+One such example is Deadwater Slough, an approximately `r round(ddw_rkms, 1)` kilometer long reach of unnaturally slow and deep water in the Salmon River, Idaho.
 The Deadwater Slough is located in a river reach containing historically important overwinter rearing and spring emigration habitat for juvenile Chinook Salmon *O.
@@
 The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
-Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River [@Emerson1973], thereby creating a slow, deep section in the river, spanning approximately 30 acres, and averaging 68 m width.
+Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River [@Emerson1973], thereby creating a slow, deep section in the river, spanning approximately `r round(ddw_acres)` acres (`r round(ddw_hectares)` hectares), and averaging `r round(ddw_avg_width_m)` m width.
 
@@
 All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the sampling period), (2) all fish have equal chance of being caught in the second (and subsequent) sample(s), (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or mistaken.
-
-```{r psd}
-s_size = 250
-q_size = 380
-
-psd_df = cmr_df %>%
-  select(species, count, length, recap, event_name) %>%
-  filter(recap  != TRUE)
-
-fs = psd_df %>%
-  filter(length >= s_size) %>%
-  nrow()
-
-fq = psd_df %>%
-  filter(length >= q_size) %>%
-  nrow()
-
-psd = 100 * (fq/fs)
-
-```
-
-<!--
-Finally, we calculated the proportional stock density (PSD) for Northern Pikeminnow captured during all field efforts, which is a measure of species size structure.
-PSD is a ratio, typically expressed as a percentage, between the number of "quality-sized" individuals or larger individuals and "stock-sized" individuals.
-Stock and quality size definitions vary by species.
-PSD is expressed as:
-
-$$
-PSD_{i} = 100 * \frac{FQ_{i}}{FS_{i}}
-$$
-
-where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
-We calculated the PSD for Northern Pikeminnow in our study using `r prettyNum(s_size, digits = 0)` mm TL for stock-size and `r prettyNum(q_size, digits = 0)` mm TL for quality-size [@Winther2020].
--->
 
@@
 
-We estimated an individual, average-size (`r avg_fall_NPM` mm) Northern Pikeminnow to consume `r median(fall_g_summ$cumm_g)` g (range `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the DSR emigration, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
-Similarly, we estimated the average-size (`r avg_spring_NPM` mm) Northern Pikeminnow to consume `r median(spring_g_summ$cumm_g)` g (range `r min(spring_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
+We estimated an individual, average-size (`r avg_fall_NPM` mm) Northern Pikeminnow to consume `r median(fall_g_summ$cumm_g)` g (range `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the DSR emigration to maintain their body size, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
+Similarly, we estimated an average-size (`r avg_spring_NPM` mm) Northern Pikeminnow to consume `r median(spring_g_summ$cumm_g)` g (range `r min(spring_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
 Given an average weight of `r avg_DSR_g` g for DSR emigrants, that would equate to `r round(median(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` (range `r round(min(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)`) fish consumed by an individual during fall and an additional `r round(median(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` (range `r round(min(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)`) fish consumed using an average weight of `r avg_NRR_g` g for NRR emigrants.
 
-Given a mean population size of `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` from fall estimates and an estimate of `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` for spring 2021, and assuming the median value from diet scenarios (i.e, 50% of their diet consists of fish), we'd estimate that Northern Pikeminnow can consume `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")`) fish during the DSR emigration and `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")`) fish during the NRR emigration in Deadwater Slough.
+Given a mean Northern Pikeminnow population size of `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` from fall estimates and an estimate of `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` for spring 2021 and using the median value from diet scenarios (i.e, 50% of their diet consists of fish), we'd estimate that Northern Pikeminnow can consume `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")`) fish during the DSR emigration and `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")`) fish during the NRR emigration in Deadwater Slough.
 Table \@ref(tab:adult-imp) shows the estimated number of fish consumed by an individual Northern Pikeminnow and by the population across all diet scenarios during both the fall (DSR) and spring (NRR) peak emigration seasons.
@@
 
-We estimated the population size of Northern Pikeminnow in the Deadwater Slough to be greater than ***xx,xxx*** during the fall emigration period for DSR Chinook Salmon.
-That estimate translates to a density of ***xxx*** Northern Pikeminnow per 100 m or ***xxx*** per 100 m^2^ which is ***similar/more/less*** than estimates from elsewhere in the Columbia River ***(citation)*** where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminnow abundance.
-The population size of Northern Pikeminnow was not directly estimated during the spring NRR Chinook salmon emigration period; however, the relative abundance measured at CPUE was comparable to the fall sampling periods ***(update statement later)***.
-The population of Northern Pikeminnow in Deadwater Slough was estimated to consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook Salmon during the ***x*** sampling periods and result in an estimated reduction of returning adults between ***xxx*** and ***x,xxx***.
-We suggest that the habitat modifications that created the Deadwater Slough have resulted in favorable conditions for Northern Pikeminnow, including improved conditions for predation upon juvenile Chinook Salmon ***(add detail here)***.
-Therefore, predation by Northern Pikeminnow in the Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG.
-
-<!--
-The PSD for Northern Pikeminnow in our study and across all three years was `r prettyNum(psd, digits = 1)`% i.e., `r prettyNum(psd, digits = 1)` of captured fish were over the quality size classification of `r prettyNum(q_size, digits = 0)` mm TL defined by @Winther2020.
-For comparison, @Winther2020 reported a PSD for Northern Pikeminnow captured below Bonneville Dam of 41% and 18% in Bonneville Reservoir, suggesting that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to population reported elsewhere in the Columbia River.
--->
+```{r john-day}
+john_day_res_abund = 85316
+john_day_rkms = 123
+john_day_dens = john_day_res_abund / john_day_rkms
+
+low_columbia_fish_km = 2814
+
+```
+
+The mean population size of Northern Pikeminnow in the Deadwater Slough during the fall season was estimated to be `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` and during the spring 2021 to be `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")`.
+Those estimates translate to linear densities of `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms), big.mark = ",")` which is `r round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms / john_day_dens)` and `r round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms / john_day_dens)` times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r as.character(english(round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms / low_columbia_fish_km)))` and `r as.character(english(round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms / low_columbia_fish_km)))` times greater than projections for the lower Columbia River [@Beamesderfer1996].
+These densities for Deadwater Slough initially seem unrealistic especially given the size (volume) of the Salmon River at the slough relative to the Columbia River.
+However, we did catch `r prettyNum(nrow(cmr_df), big.mark = ",")` Northern Pikeminnow over the course of the study, of which only `r sum(cmr_df$capture_type == "Recapture", na.rm = T)` fish (one in every `r round(nrow(cmr_df) / sum(cmr_df$capture_type == "Recapture", na.rm = T))`) were recaptures.
+Further, @Beamesderfer1991 did acknowledge that they were unsuccessful in finding a suitable method for sampling offshore; they noted that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective and so their abundance estimates only apply to nearshore areas.
+This was not the case at Deadwater Slough where maximum depth was approximately 10-15 m with a relatively homogeneous bottom contour i.e., angling was effective for the entire area.
+Our estimates of Northern Pikeminnow areal density are also higher than those reported elsewhere.
+We estimate the density of Northern Pikeminnow in Deadwater Slough, as fish per hectare, to be `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_hectares), big.mark = ",")` for fall and `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_hectares), big.mark = ",")` for spring.
+For comparison, the only other available estimates of areal density, that the authors are aware of, are 4.4 fish per hectare for John Day Reservoir [@Beamesderfer1991] and 15 fish per hectare in Lake Washington [@Bartoo1977].
+
+```{r psd}
+s_size = 250
+q_size = 380
+
+psd_df = cmr_df %>%
+  select(species, count, length, recap, event_name) %>%
+  filter(recap  != TRUE)
+
+fs = psd_df %>%
+  filter(length >= s_size) %>%
+  nrow()
+
+fq = psd_df %>%
+  filter(length >= q_size) %>%
+  nrow()
+
+psd = 100 * (fq/fs)
+
+```
+
+Northern Pikeminnow in Deadwater Slough were not only abundant, but anglers noticed that many were also large-bodied.
+For that reason, we calculated the proportional stock density (PSD) for Northern Pikeminnow in Deadwater Slough, which is a measure of species size structure.
+PSD is a ratio, typically expressed as a percentage, between the number of "quality_sized" (larger) individuals and "stock-sized" individuals.
+
+$$
+PSD_{i} = 100 * \frac{FQ_{i}}{FS_{i}}
+$$
+
+where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
+Stock and quality size definitions vary by species.
+We calculated the PSD for Northern Pikeminnow in Deadwater Slough using `r prettyNum(s_size, digits = 0)` mm TL for stock-size and `r prettyNum(q_size, digits = 0)` mm TL for quality-size [@Winther2020].
+The PSD for Northern Pikeminnow in Deadwater Slough across all three efforts was `r prettyNum(psd, digits = 1)`% i.e., `r prettyNum(psd, digits = 1)`% of captured fish were over the quality size classification defined by @Winther2020.
+For comparison, @Winther2020 reported a PSD for Northern Pikeminnow captured below Bonneville Dam of 41% and 18% in Bonneville Reservoir, suggesting that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to populations reported elsewhere in the Columbia River.
+In fact, the state record Northern Pikeminnow, which measured `r max(cmr_df$length, na.rm = T)` mm (`r round(max(cmr_df$length, na.rm = T) / 25.4, 1)` in) TL, was caught in Deadwater Slough during the fall 2020 effort.
 
@@
 ## Bioenergetics
+
+The population of Northern Pikeminnow in Deadwater Slough was estimated to consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook Salmon during the ***x*** sampling periods and result in an estimated reduction of returning adults between ***xxx*** and ***x,xxx***.
 
@@
 
+projections for the lower Columbia River (Beamesderfer et al.
+1996) where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminnow abundance.
+
 ## Conclusions
-
+* We suggest that the habitat modifications that created the Deadwater Slough have resulted in favorable conditions for Northern Pikeminnow, including improved conditions for predation upon juvenile Chinook Salmon ***(add detail here)***.
+Therefore, predation by Northern Pikeminnow in the Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG.
 *	We have also presented a novel modeling framework for estimating predation on native, critically endangered anadromous species which can be applied to other areas of interest. John day? Others?
@@
          adult_Uci = round(adult_Uci, 0)) %>%
-  rename(Effort = season,
+  rename(Season = season,
          `p Fish` = perc_diet,
-         `Juv Chinook` = juv_chnk,
-         `Juv N` = smolt_N,
-         `Juv L CI` = smolt_Lci,
-         `Juv U CI` = smolt_Uci,
+         `Fish Per NPM` = juv_chnk,
+         `Fish N` = smolt_N,
+         `Fish L CI` = smolt_Lci,
+         `Fish U CI` = smolt_Uci,
          `Median SAR` = medsar,
@@
         format.args = list(big.mark = ","),
-        caption = "Estimates of the number of DSR (fall) and NRR (spring) emigrants consumed for each season and diet scenario.
+        caption = "Estimates of the number of fish consumed for each season and diet scenario.
 The diet scenario indicates the proportion of the Northern Pikeminnow diet consisting of fish.
-The 'Juv Chinook' column indicates the number of juvenile Chinook consumed by an individual Northern Pikeminnow during the fall or spring seasons and is taken from each of the 18 bioenergetics model runs.
-The number of smolts consumed (and 95% confidence intervals) are calculated using Northern Pikeminnow abundance estimates of 28,144 (95% CI: 14,378 - 59,978) for fall and 14,897 (95% CI: 7,610 - 31,747) for spring.
-The number of 'adult equivalents' is also estimated using a Lower Granite Dam smolt-to-adult rate (SAR) from McCann et al.
+The 'Fish Per NPM' column indicates the number of fish an individual Northern Pikeminnow to consume during the fall or spring seasons assuming an average weight of 10.3 (DSR) and 10.9 (NRR) grams for juvenile Chinook Salmon and is taken from each of the 18 bioenergetics model runs.
+The number of fish consumed (and 95% confidence intervals) are calculated using Northern Pikeminnow abundance estimates of 28,144 (95% CI: 14,378 - 59,978) for fall and 14,897 (95% CI: 7,610 - 31,747) for spring.
+The number of 'adult equivalents' is also estimated using a Lower Granite Dam smolt-to-adult rate (SAR) for Chinook Salmon from McCann et al.
 2019.") %>%

### 2021-12-28 analysis/paper/paper.Rmd
edits to m-r and gastric discussion

@@
 
-<!--
-Re-hash out our assumptions?
-What assumptions were most likely violated e.g., the closed population assumption?
-If we assumed an open population with varying immigration/emigration rates how might that affect our estimate?
-In the end, how good do we think our estimate is and is the bottom-line that we belive there still to be a shit-ton of pikeminnow even if our assumptions were violated?
--->
-
-All estimators suggest a large number of Northern Pikeminnow occupying Deadwater Slough.
-<!-- This might be a good place to bring in some comparisons with Northern Pikeminnow densities from elsewhere to put our numbers in context.-->
-
-If the first mark-recapture assumption was violated and the population is actually open (potential immigration and emigration during the survey period), as long as the immigration and emigration rates are equal between marked and unmarked fsih, our estimators should still provide unbiased estimates of abundance.
-Given the size of the sampling area, and the length of the sampling period (two weeks).
-
-We are comfortable with assumptions 4 and 5 (no loss of marks and no mistaken or missed marks).
-Assumptions 2 and 3 indicate whether the second (and subsequent) surveys are random samples of the population.
-Angling methods may provide a size selection bias, but that may only affect the interpretation of the abundance estimates (i.e., our estimate are the abundance of Northern Pikeminnow above some size threshold).
-Unequal catchability between individual fish is very difficult to assess, especially if that heterogeneity arises from being caught once.
-
-<!-- If catching and marking fish makes them want to leve the Deadwater area (more than unmarked fish), that could be making our estimates biased high (due to fewer recaptures).
-I don't imagine this is happening, but does anyone else have concerns about that?
--->
-
-The spring abundance estimate is smaller than either fall sampling event.
-However, our approach assumes equal capture probabilities betwee the fall and spring sampling events.
-If it does differ, we believe that the capture probability in the spring could be lower due to higher flows, which would lead to our spring abundance estimate being biased low.
-Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 sampling event to be a conservative estimate.
+All mark-recapture abundance estimators suggest a large number of Northern Pikeminnow occupying Deadwater Slough.
+Estimators assumed: (1) the population is closed (no immigration, emigration, births, or deaths during the sampling period), (2) all fish have equal chance of being caught in the second (and subsequent) samples(s), (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or mistaken.
+If the population is actually open (potential immigration and/or emigration during the survey period), our estimators should still provide unbiased estimates of abundance as long as the immigration and emigration rates are equal between marked and unmarked fish.
+Given the size of the sampling area, and the limited length of the sampling period (two weeks), we are comfortable with the assumption of no immigration or emigration.
+Assumption (3) indicates whether the second (and subsequent) surveys are random samples of the population.
+Angling methods may provide a size selection bias, but that only affects the interpretation of the abundance estimates (i.e., our estimates are the abundance of Northern Pikeminnow susceptible to angling above some size threshold).
+Unequal catchability between individual fish is very difficult to assess, especially if that heterogeneity arises from being caught once; however, we did recapture two individual fish more than once suggesting that marked fish were still susceptible to angling.
+We are also comfortable with assumptions 4 and 5 (no loss of marks and no mistaken or missed marks) given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.
+
+We do consider two potential violations of assumption, both which reduces the probability that a marked fish is susceptible to capture in the second (and subsequent) samples.
+First, the assumption of a closed population could be violated if some marked fish that are released back to the population die.
+We don't believe this case to be likely, as fish that were dead or nearly so after processing were not released back to the population (nor recorded as such), and those that were released were hand-released only if they were able to swim away on their own volition; however, the potential for marked and released fish to have died at a higher rate than unmarked fish should be acknowledged.
+The second, and more likely, reason is that marked and released fish did not mix thoroughly back into the population (at least not at a fast enough rate), reducing their probability of capture in subsequent captures, thus violating assumption (2).
+We attempted to release marked fish throughout Deadwater Slough to facilitate mixing back into the population; however, anecdotally volunteer anglers would congregate to particular areas of the slough either to socialize with other anglers, because the fishing was good, etc.
+and if marked and released fish were not readily available in those areas that would make marked fish less susceptible to capture.
+This violation would, in effect, reduce the number of marked fish available for recapture or reduce the marking fraction in some areas of the slough.
+Violations of assumption (2) would lead to an overestimate of abundance of Northern Pikeminnow which is a real possibility in our study.
+
+The spring abundance estimate was smaller than either fall abundance estimate.
+However, our approach assumed equal capture probabilities between the fall and spring sample efforts.
+If it does differ, we suspect that the capture probability in the spring would be lower due to higher flows (i.e., reduced densities), which would lead to our spring abundance estimate being biased low.
+Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 effort to be a conservative estimate.
 
@@
 
-The majority of fish collected during this study received gastric lavage, including some non-predatory species.
-To validate the efficacy of this method, we euthanized nine Northern Pikeminnow after gastric lavage was completed and removed the remaining stomach contents via dissection.
-Our results support previous findings that gastric lavage effectively removes stomach contents (@Lott2020).
-Additionally, we found that the fish captured in net-traps had a similar proportion of stomach content samples.
-Of those, 1,214 (76.2%) were completely empty, 345 (21.6%) had stomach contents, and 35 (2.2%) contained fish remnants.
-We were successfully able to identify Redside Shiners, a Largescale Sucker, sculpin, Mountain Whitefish, and one juvenile Chinook Salmon in the stomach contents.
-However, decomposition from digestion rendered most stomach contents unidentifiable.
+The majority of fish caught during the study received gastric lavage.
+To evaluate the efficacy of the method, we euthanized nine Northern Pikeminnow after gastric lavage was completed and removed the remaining stomach contents via dissection.
+In each case, no additional stomach contents were identified, suggesting that gastric lavage was effective in removing stomach contents, although our sample size was low.
+However, in total, we only identified fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r round(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"] * 100, 1)`%) of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow examined using gastric lavage.
+Aside from juvenile Chinook Salmon, identified species included Redshine Shiner, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*; however, decomposition from digestion rendered most stomach contents unidentifiable.
+
+Although we only identified fish or fish parts within `r round(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"] * 100, 1)`% of Northern Pikeminnow examined and that fish remnants were only `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall stomach contents, we believe those numbers to be misleading.
 

### 2021-12-28 analysis/paper/paper.Rmd
bio & adult impacts discussion additions

@@
 
-```{r read-bio-outputs, message=FALSE, warning=FALSE}
+```{r read-bio-outputs, message=FALSE, warning=FALSE, include=FALSE}
 fmt_bio_out = function(x) {
@@
 We estimated an individual, average-size (`r avg_fall_NPM` mm) Northern Pikeminnow to consume `r median(fall_g_summ$cumm_g)` g (range `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the DSR emigration to maintain their body size, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
-Similarly, we estimated an average-size (`r avg_spring_NPM` mm) Northern Pikeminnow to consume `r median(spring_g_summ$cumm_g)` g (range `r min(spring_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
+Similarly, we estimated an average-size (`r avg_spring_NPM` mm) Northern Pikeminnow to consume `r median(spring_g_summ$cumm_g)` g (range `r min(spring_g_summ$cumm_g)` - `r max(spring_g_summ$cumm_g)` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
 Given an average weight of `r avg_DSR_g` g for DSR emigrants, that would equate to `r round(median(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` (range `r round(min(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)`) fish consumed by an individual during fall and an additional `r round(median(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` (range `r round(min(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)`) fish consumed using an average weight of `r avg_NRR_g` g for NRR emigrants.
@@
 
-Although we only identified fish or fish parts within `r round(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"] * 100, 1)`% of Northern Pikeminnow examined and that fish remnants were only `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall stomach contents, we believe those numbers to be misleading.
+Although we only identified fish or fish parts within `r round(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"] * 100, 1)`% of Northern Pikeminnow examined and that fish remnants were only `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall stomach contents, we believe those numbers to be misleading...
 
@@
 ## Bioenergetics
-
-The population of Northern Pikeminnow in Deadwater Slough was estimated to consume between ***xx,xxx*** and ***xx,xxx*** juvenile Chinook Salmon during the ***x*** sampling periods and result in an estimated reduction of returning adults between ***xxx*** and ***x,xxx***.
-
-What assumptions did we make during the bioenergetics assessment?
-And how might violations of those assumptions change our estimate of the number of juvenile Chinook salmon consumed and resulting impacts to adult returns?
+<!--
+OUR APPROACH:
+1. g of fish consumed during DSR and NRR peak emigrations
+2. Combine results from #1 for total g consumed by an individual NPM
+3. Parse out the percentage of fish prey that is juvenile Chinook salmon
+  - start w/ conservative scenarios of 10% and 25%
+4. Divide total g consumed by 10.6 g, the average juvenile Chinook Salmon from Upper Salmon MPG, to estimate the # of juvenile Chinook Salmon consumed
+5. Multiply results from #4 by a "stable" NPM population size estimate
+6. Multiply results from #5 by SAR from McCann et al. 2019 for Upper Salmon MPG Chinook
+-->
+
+An average-size Northern Pikeminnow from Deadwater Slough is estimated to consume `r median(fall_g_summ$cumm_g)` g (range `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) and `r median(spring_g_summ$cumm_g)` g (range `r min(spring_g_summ$cumm_g)` - `r max(spring_g_summ$cumm_g)` g) of fish during spring and fall, respectively, to maintain its body weight using mean daily water temperatures recorded at a nearby stream gage [@USGS2016].
+This equates to `r median(fall_g_summ$cumm_g) + median(spring_g_summ$cumm_g)` g of fish consumed during the DSR and NRR peak emigrations assuming the median value among diet scenarios (i.e., if 50% of their diet consists of fish); values ranged from a minimum of `r min(fall_g_summ$cumm_g) + min(spring_g_summ$cumm_g)` g if only 10% of the diet is fish to a maximum of `r max(fall_g_summ$cumm_g) + max(spring_g_summ$cumm_g)` g if 90% of the diet is fish.
+
+<!-- The following needs to be updated after we finalize some results -->
+This approach currently assumes that juvenile Chinook Salmon are the only fish prey available to or preyed upon by Northern Pikeminnow.
+It further assumes that all prey available have energy densities equal to juvenile Chinook Salmon, which may not be far from the truth, given that other salmonids including juvenile steelhead and Sockeye Salmon are likely also prevalent as available prey in Deadwater Slough, especially during the spring migration.
+Additionally, hatchery smolt releases (Chinook Salmon, steelhead, and Sockeye Salmon) are also prevalent in the mainstem Salmon River during the spring migration.
+However, our study is focused on potential impacts to natural-origin DSR and NRR Chinook Salmon, in which case, let’s consider a “thought exercise” where DSR and NRR Chinook Salmon represent only 10% and 25% of total fish prey consumed by Northern Pikeminnow during these periods.
+Under these scenarios, the average Northern Pikeminnow would consume 8.95 g (range 4.52 – 10.05 g) if 10% of prey were natural-origin Chinook Salmon and 22.38 g (range 11.31 – 25.12) if 25%.
+
+To estimate the total number of possible juvenile Chinook Salmon that could be consumed, we divided the cumulative consumption in grams for an individual Northern Pikeminnow by the average weight of a Chinook Salmon emigrant (taken from all rotary screw traps in the Upper Salmon MPG) to come up with a juvenile equivalency, and then multiplied this by a "stable" estimate of abundance for Northern Pikeminnow in Deadwater Slough, resulting in an estimate of the equivalent toal number of Chinook Salmon juveniles consumed.
+If we assume an average juvenile Chinook Salmon of 10.6 g, the mean from DSR and NRR Chinook Salmon emigrants in the Upper Salmon MPG, this equates to 0.84 juvenile Chinook Salmon (range 0.43 – 0.94) under the 10% scenario and 2.11 juvenile Chinook Salmon (range 1.07 – 2.37) under the 25% scenario.
+
+Let’s now consider a stable Northern Pikeminnow population size of 10,000 (roughly the mean of fall and spring abundance estimates, divided by two to account for half of marked individuals not being available for recapture).
+In this case, the Northern Pikeminnow population in Deadwater Slough would be estimated to consume 8,400 (range 4,300 – 9,400) or 21,100 (range 10,700 – 23,700) under the 10% and 25% scenarios, respectively.
 
@@
 
-To estimate the number of possible Chinook salmon that could be consumed, we divided the cumulative fish consumption in grams for an individual pikeminnow by the average weight of a Chinook emigrant to come up with a presmolt equivalency and then multiplied this by our abundance estimate of northern pikeminnow in Deadwater Slough, resulting in an estimate of the equivalent total number of Chinook salmon presmolts consumed.
-Finally, we estimated the number of additional adult Chinook salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough reach was reduced or eliminated.
-We multiplied the total estimated consumed juvenile Chinook salmon by an estimate of Granite-to-Granite SARs [@McCann2019].
-
-This translates to adult Chinook equivalents between `r paste(prettyNum(round(range(total_impacts$adult_N)), big.mark = ","), collapse = " and ")` (Figure \@ref(fig:adult-imp-p)).
-
-Finally, we estimated the number of additional adult Chinook Salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough reach was reduced or eliminated by multiplying the total estimated consumed Chinook Salmon by an estimate of Granite-to-Granite smolt-to-adult survival rates (SARs) (`r round(sar$medsar, 5)`, SE: `r round(sar$sdsar, 5)`) [@McCann2019].
+<!-- Also update this after finalizing results -->
+Finally, let's take our thought exercise one additional step and estimate the number of additional adult Chinook Salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough was reduced or eliminated.
+Our approach here was to multiply the total estimated juvenile Chinook Salmon consumed by an estimate of Granite-to-Granite smolt-to-adult survival rate (SAR) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.
+Assuming an SAR of `r round(sar$medsar, 5)` (SE: `r prettyNum(round(sar$sdsar, 5))`) we'll consider the "adult equivalents" to our estimates of juveniles consumed.
+Under this scenario, the consumed juveniles would be equivalent to 51.6 adults (range 26.4 – 57.7) and 129.6 adults (range 65.7 – 145.6).
+However, we believe these 10% and 25% scenarios to be conservative, especially during the fall DSR emigration period.
+During spring, natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, and Sockeye Salmon) are present in the river, and so 10% or 25% may be reasonable.
+However, during fall, DSR Chinook Salmon are likely the dominant prey species available to Chinook Salmon in Deadwater Slough as 1) no (or at least minimal) hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent than in Chinook Salmon.
+Note that this thought exercise only includes the Northern Pikeminnow population in Deadwater Slough even though they are present throughout the upper Salmon River, although likely at lower densities than in Deadwater Slough, and also ignores potential impacts to ESA-listed steelhead and Sockeye Salmon that are also a portion of the Northern Pikeminnow diet.
 
 <!--
-After making some assumptions (e.g., on SARs) attempt to estimate how results from objective #3 could translate into impacts on adult returns (via a thought experiment or simulation approach)
-
-Currently only using one bioenergetics model run.
-Could incorporate more bioenergetics models, but if so then maybe consider focusing on a single abundance estimator (e.g.
-Schnabel) to avoid too many combinations in the results
+Additional thoughts here:
+*	SAR results. Those results are GRA - GRA SARs, estimated from multiple years of PIT tag data. Seem pretty good to me.
+* Maybe a plot of multiple scenarios
 -->
-
-Again, what assumptions did we make here and how might violations of those assumptions change our estimate of impacts to adult returns.
-
-*	SAR results. Those results are GRA - GRA SARs, estimated from multiple years of PIT tag data. Seem pretty good to me.
-* Maybe a plot of multiple scenarios depending on gastric lavage results? And pop estimates if they vary in time.
-* Maybe pop estimates if they vary in time.
 
@@
 ## Conclusions
+
 * We suggest that the habitat modifications that created the Deadwater Slough have resulted in favorable conditions for Northern Pikeminnow, including improved conditions for predation upon juvenile Chinook Salmon ***(add detail here)***.

### 2021-12-29 analysis/paper/paper.Rmd
updates to avian predation discussion

@@
 
-Although not formally assessed in this study, avian predators including Great Blue Herons *Ardea herodias* and Bald Eagles *Haliaeetus leucocephalus* are another potential source of mortality for juvenile salmon in the Deadwater Slough.
-The Deadwater Slough is recognized as an important bird watching and nesting area due to the associated riparian and backwater habitats [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
-Several piscivorous bird species have been documented using the Deadwater Slough that include the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
-During the initial sampling event in 2019, a two-person crew walked the entire reach and surrounding and upstream areas scanning for passive integrated transponder (PIT) tags.
-During that informal survey, nine PIT tags were recovered near active bird nests or in an upstream anastomizing reach where herons and eagles are prevalent, suggesting that mortality may have been a result of avian predation.
-The PIT tag histories in [PTAGIS](https://www.ptagis.org/) indicate these tags were implanted into a combination of juvenile Chinook Salmon (3), Sockeye Salmon (3), and steelhead (3).
-Avian predation contributes a major component of the total mortality for yearling Chinook Salmon in some locations in the lower Snake River and Columbia River, particularly at hydroelectric dams and within reservoirs [@Evans2012; -@Evans2016]; however, we did not observe large colonies of piscivorous birds within the study area.
-Although there is documentation of individual Double-crested Cormorants [@eBird2021] at the Deadwater Slough, the site is not within their breeding range, rather, it is part of a migration corridor.
+Although not formally assessed in this study, avian predators including Great Blue Herons *Ardea herodias* and Bald Eagles *Haliaeetus leucocephalus* are another potential source of mortality for juvenile Chinook Salmon and other ESA-listed salmonids in the Deadwater Slough.
+the Deadwater Slough is recognized as an important bird watching and nesting area due to the associated upstream braided channels and riparian and backwater habitats [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
+Several piscivorous bird species have been documented using Deadwater Slough including the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
+During the initial sampling effort in fall 2019, a two-person crew walked the entire reach including surrounding and upstream areas scanning for passive integrated transponder (PIT) tags that are implanted into thousands of natural- and hatchery-origin salmonids in the Upper Salmon River, annually, to monitor movement and survival.
+During that informal survey, 10 PIT tags were recovered near active bird nests or in an upstream anastomizing reach where herons and eagles are prevalent, suggesting that mortality may have been a result of avian predation.
+The PIT tag histories in [PTAGIS](https://www.ptagis.org/) indicate these tags were implanted into a combination of juvenile Chinook Salmon (3), Sockeye Salmon (4), and steelhead (3), all juveniles.
+Two of the Chinook Salmon and one steelhead were natural-origin, the remaining were of hatchery-origin.
+Avian predation contributes a major component of the total mortality for yearling Chinook Salmon in some locations in the lower Snake and Columbia rivers, particularly at hydroelectric dams and within reservoirs [@Evans2012; -@Evans2016]; however, we did not observe large colonies of piscivorous birds within the study area.
+Although we did observe Belted Kingfishers, mergansers, and Great Blue Herons at Deadwater Slough during field efforts; Great Blue Herons were especially prevalent in the anastomizing reach immediately upriver of the slackwater reach.
+Additionally, there is documentation of individual Double-crested Cormorants [@eBird2021] at the Deadwater Slough, although the site is not within their breeding range; rather, it is part of a migration corridor.
 Given the current avian species known to occupy Deadwater Slough, it is unlikely that avian predation on juvenile salmonids is comparable to elsewhere in the Columbia River basin with large piscivorous bird colonies.
-Nevertheless, we hypothesize that the reservoir-like conditions at the Deadwater Slough may increase the probability of avian predation on juvenile Chinook Salmon from the many piscivorous birds known to use the site.
+Nevertheless, we hypothesize that the reservoir-like conditions and associated habitat features at the Deadwater Slough may increase the probability of avian predation on juvenile Chinook Salmon and other ESA-listed salmonids from the many piscivorous birds known to use the site.
 Future estimates of predation would benefit from consideration of the contribution of piscivorous avian predators.

### 2021-12-29 analysis/paper/paper.Rmd
updates to management implications

@@
 
-At least five of the eight populations in the Upper Salmon MPG must meet criteria set forth by @McElhany2000 and @ICTRT2007 for the MPG to be considered viable and for the recovery of the Snake River Evolutionary Significant Unit.
-Populations with the ESU have substantial cultural value, support downriver mainstem Snake and Columbia River commercial and subsistence fisheries, and support local fisheries and economies in years with sufficient abundance.
-
-We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces annual adult returns by ***xxx - x,xxx*** (perhaps express as percent, too) to upriver populations.
-Presumably, that reduction in adult returns impacts both the ESA-listed natural populations in the Upper Salmon River MPG and two hatchery populations in the Upper Salmon, Pahsimeroi and Sawtooth hatcheries, which provide for recrational fishing opportunities.
-Consequently, reducing juvenile Chinook Salmon predation mortalities at Deadwater Slough can potentially benefit multiple upriver natural and hatchery populations upriver.
-Because the Deadwater Slough is part of the migratory pathway for emigrating salmonids, including for multiple species and populations, the impact of that benefit would likely be higher than tributary habitat rehabilitation actions, which typically benefit a single population.
+<!-- Again, some of the following numbers need to be updated after results are formalized -->
+Chinook Salmon populations in the Upper Salmon River MPG support local fisheries and economies in years with sufficient abundance and have substantial cultural value in the region.
+These populations additionally support downriver commercial and subsistence fisheries in the Snake and Columbia rivers.
+Recovery of the MPG is desired which requires that at least five of the eight populations must meet criteria set forth by @McElhany2000 and @ICTRT2007 for the MPG to be considered viable and for recovery of the Snake River Evolutionary Significant Unit.
+Spawner abundance is perhaps the most important metric considered in determining a population's viability.
+We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces annual adult natural-origin returns by 26.4 to 145.6 to upriver populations.
+Although this does not sound like a substantial amount, it represents 4.7% - 26.1% of the total mean adult escapement of 558 reported for the Upper Salmon River MPG in 2017 - 2019 [@Kinzer2020].
+Presumably, reductions in adult returns also affect natural-origin Sockeye Salmon and steelhead as well as hatchery populations in the Upper Salmon which provide for recreational fishing opportunities.
+Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations for multiple ESA-listed species.
+Because the Deadwater Slough is part of the migratory pathway for emigrating salmonids, including for multiple species and populations, the impact of that benefit could be higher than tributary rehabilitation actions, which typically benefit a single population.
 Moreover, the deepened, slack water conditions that favor northern pikeminnow at Deadwater Slough are indirectly the result of manmade activities i.e., the failure of a manmade mining reservoir dam.
@@
 We see two potential management actions: 1) removing the Dump Creek delta to restore flow and 2) a local northern pikeminnow bounty program to encourage harvest of northern pikeminnow in Deadwater Slough.
+
+projections for the lower Columbia River (Beamesderfer et al.
+1996) where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminnow abundance.
 
@@
 -->
-
-projections for the lower Columbia River (Beamesderfer et al.
-1996) where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminnow abundance.
 

### 2021-12-29 analysis/paper/paper.Rmd
more edits to management implications

@@
 
-We see two potential management actions: 1) removing the Dump Creek delta to restore flow and 2) a local northern pikeminnow bounty program to encourage harvest of northern pikeminnow in Deadwater Slough.
-
-projections for the lower Columbia River (Beamesderfer et al.
-1996) where substantial Northern Pikeminnow predation impacts on salmonids have led to bounty programs aimed at reducing Northern Pikeminnow abundance.
-
+We envision two potential management actions at reducing predation at Deadwater Slough: 1) removing the Dump Creek alluvial fan and 2) a local Northern Pikeminnow bounty program to encourage harvest in Deadwater Slough aimed at reducing the predator population size.
+Both actions have evident pros and cons.
+Removing the Dump Creek alluvial fan has the benefit of restoring the Salmon River to a more natural flow (e.g., increased water velocity, decreased depth) and bathymetry (e.g., heterogeneous substrate) that likely existed in the reach prior to the failure of the mining reservoir dam in Dump Creek.
+Restoring flow would presumably reduce Northern Pikeminnow densities to levels similar to upstream and downstream reaches where recent studies identified higher survival and transition probabilities than those at Deadwater Slough (@Axel2015; @Ackerman2018; @Porter2019).
+However, we do not know the feasibility of removing or alleviating the Dump Creek alluvial fan, and if it could be removed or altered, future disturbance events could potentially move additional loose sediment in the Dump Creek drainage, thereby creating a new alluvial fan.
+Additionally, restoring the flow would likely diminish the value of the reach as an important bird watching and nesting area.
+
+Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere (citation).
+A local bounty program in the Salmon River has the potential to boost the local economy (to an unknown degree) which has been negatively impacted by low returns of Chinook Salmon, steelhead, and Sockeye Salmon in recent decades.
+Many local business rely on license and gear sales for local sport fisheries, but low salmon and steelhead returns have eliminated or reduced the duration of fisheries in some years, and a local Northern Pikeminnow bounty program could allow the local economy to recoup some of those lost revenues.
+On the downside, bounty programs require continual effort and harvest to suppress the population size and so would require year of continued funding and local buy-in to remain effective.
+If funding or effort was alleviated, Northern Pikeminnow would likely recover to a higher population size without additional alterations to the habitat.
 <!--
+What about adding cover? It could potentially provide refugia for juvenile salmonids, but could make recreational angling more difficult
+
 Is there anything in the literature that discusses habitat preferences for pike minnow?

### 2021-12-29 analysis/paper/paper.Rmd
updated some citations

@@
 Currently, the Upper Salmon MPG supports eight independent, extant spring/summer Chinook Salmon populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
-The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014].
+The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a].
 Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency.

### 2021-12-29 analysis/paper/paper.Rmd
added draft conclusion

@@
 
-* We suggest that the habitat modifications that created the Deadwater Slough have resulted in favorable conditions for Northern Pikeminnow, including improved conditions for predation upon juvenile Chinook Salmon ***(add detail here)***.
-Therefore, predation by Northern Pikeminnow in the Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG.
-*	We have also presented a novel modeling framework for estimating predation on native, critically endangered anadromous species which can be applied to other areas of interest. John day? Others?
-*	The end.
+The failure of a mining reservoir dam in the late 19th century created an alluvial fan at the confluence of the Salmon River and Dump Creek which has backed up the Salmon River and created an almost 2-kilometer section of unnaturally slow and deep river, resulting in favorable conditions for Northern Pikeminnow.
+We estimate that greater than 10,000 Northern Pikeminnow occupy Deadwater Slough with densities greater than reported elsewhere.
+Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids emigrating through Deadwater Slough including Chinook Salmon i.e., juvenile Chinook Salmon have slower movement rates and little refugia to avoid predation.
+We estimate that predation from Northern Pikeminnow in Deadwater Slough may reduce adult returns by as much as 146 or 26% or recent returns to the Upper Salmon River MPG.
+Therefore, predation in Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
 

### 2021-12-29 analysis/paper/paper.Rmd
minor edits to acknowledgements

@@
 
-The authors extend much appreciation to the many volunteers who assisted with field efforts including collaborators from the Bureau of Reclamation, Idaho Office of Species Conservation, Idaho Deparment of Fish and Game, and the Lemhi Regional Land Trust, among others.
-Special thanks to Chelsea Welke, Tulley Mackey, Jared Barker, and Brian Hamilton for their help in the field.
-Special thanks to Caitlin Alcott and Inter-Fluve for their administrative support and guidance.
+The authors extend much appreciation to the many volunteers that assisted with field efforts including collaborators from the Bureau of Reclamation, Idaho Governor’s Office of Species Conservation, Idaho Department of Fish and Game, and the Lemhi Regional Land Trust, among others.
+Special thanks to Jared Barker, Brian Hamilton, Tulley Mackey, and Chelsea Welke for their help in the field.
+We further appreciate the administrative support and guidance from Caitlin Alcott and Inter-Fluve.
 Funding for this study was provided by the Bureau of Reclamation, Pacific Northwest Regional Office (contract No.
-140R1021F0018) with assistance from the Idaho Office of Species Conservation.
+140R1021F0018) with assistance from the Idaho Governor’s Office of Species Conservation.
 

### 2021-12-29 analysis/paper/paper.Rmd
added very draft abstract

@@
 abstract: |
-  Abstract text here. 
 keywords: |
@@
 ```
+
+\newpage
+
+## Abstract
+
+Predation of emigrating juvenile salmonids by piscivorous fishes is perhaps an under-appreciated source of mortality hindering population recovery in the Columbia River basin.
+Reservoirs and tailraces associated with dams on the Snake and Columbia rivers have created favorable conditions for fishes known to consume juvenile salmonids, including the Northern Pikeminnow *Ptychocheilus oregonsis*.
+Indeed, Northern Pikeminnow have been estimated to consume 8% of the approximately 200 million juvenile salmonids emigration through those rivers, annually.
+Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above the reservoir systems.
+In this study, we examine predation by Northern Pikeminnow on juvenile salmonids, focusing on Chinook Salmon *Oncorhynchus tshawytscha*, in a slow-water reach of the Salmon River, Idaho known as Deadwater Slough.
+The reach contains historically important overwinter rearing and spring emigration habitat for Chinook Salmon and other ESA-listed species, and was formed by the failure of a small mining dam in a reservoir in the late 19th century.
+We estimate the abundance of Northern Pikeminnow in Deadwater Slough during peak emigrations of Chinook salmon, and further, use a bioenergetics approach to estimate the consumption potential of the Northern Pikeminnow population to assess potential impacts to Chinook Salmon emigrants.
+Finally, we evaulate potential impacts to adult returns of local populations.
+***We found that...
+(insert concluding findings here.*** It is likely that predation of juvenile Chinook Salmon in Deadwater Slough has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG hindering local recovery efforts.
 

### 2021-12-30 analysis/paper/paper.Rmd
added SG gastric lavage edits

@@
 
-We examined the stomach contents of `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow from Deadwater Slough using gastric lavage methods.
-Of those, we found contents of some form in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) of those.
-Table \@ref(tab:gastric-tab) summarizes gastric lavage results for each sampling effort.
-For those individuals with stomach contents, the average wet weight of total contents was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r prettyNum(min(stmch_df$stomach_contents_weight), digits = 3)`-`r prettyNum(max(stmch_df$stomach_contents_weight), digits = 3)` g).
-Overall, fish or fish remnants consisted of `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall of all stomach contents examined.
+We examined the stomach contents of Northern Pikeminnow from Deadwater Slough using gastric lavage.
+To validate the efficacy of the method, we euthanized 75 Northern Pikeminnow after gastric lavage was completed and removed the remaining food items via dissection.
+We found that gastric lavage successfully removed food items in all dissected individual, indicating the method was effective.
+Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in some form in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) of those.
+Northern Pikeminnow captured during the spring 2021 effort had a higher proportion of diets with fish contents (`r round(100 * gastric_tbl$fish_contents[gastric_tbl$event_name == "Spring_2021"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Spring_2021"], 1)`%) than fish captured in the fall 2019 (`r round(100 * gastric_tbl$fish_contents[gastric_tbl$event_name == "Fall_2019"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Fall_2019"], 1)`%) and fall 2020 (`r round(100 * gastric_tbl$fish_contents[gastric_tbl$event_name == "Fall_2020"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Fall_2020"], 1)`%) (Table \@ref(tab:gastric-tab)).
+For those individuals with stomach contents, the average wet weight of total contents was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r round(min(stmch_df$stomach_contents_weight), 2)`-`r round(max(stmch_df$stomach_contents_weight), 2)` g).
+Overall, fish or fish remnants were `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall stomach contents examined.
 
@@
 So although we are uncertain of the diet proportions consisting of fish versus invertebrates, the impact of that uncertainty is somewhat constrained.
-
 <!--
@@
 
-The majority of fish caught during the study received gastric lavage.
-To evaluate the efficacy of the method, we euthanized nine Northern Pikeminnow after gastric lavage was completed and removed the remaining stomach contents via dissection.
-In each case, no additional stomach contents were identified, suggesting that gastric lavage was effective in removing stomach contents, although our sample size was low.
-However, in total, we only identified fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r round(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"] * 100, 1)`%) of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow examined using gastric lavage.
-Aside from juvenile Chinook Salmon, identified species included Redshine Shiner, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*; however, decomposition from digestion rendered most stomach contents unidentifiable.
-
+We performed gastric lavage on nearly all Northern Pikeminnow collected during this study.
+Aside from juvenile Chinook Salmon, identified species included Redside Shiner, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*; although decomposition from digestion rendered many fish contents unidentifiable.
+While this approach provided useful information on prey species, we have concerns about using angling and gastric lavage to quantify diet composition in Northern Pikeminnow.
+First, previous studies have indicated that diets of angled fish differ from those collected by other methods (e.g., electrofishing, [@Hodgson1998]), and further, angled fish are much more likely to have empty stomachs [@Jurajda2016].
+Second, unlike other predators such as Smallmouth Bass and Walleye that are commonly sampled using gastric lavage [@Kamler2001], Northern Pikeminnow are cyprinids and therefore lack a true stomach.
+Previous research indicates that performing gastric lavage on cyprinid species can be ineffective at quantifying diet composition [@Hartleb1995], or alternately, requires a modification in methodology to flush food items out the vent of the fish [@Wasowicz1994].
+Our intent, rather, was to simply document predation on juvenile Chinook Salmon in Deadwater Slough, and given the potential ineffectiveness of gastric lavage using angling and for Northern Pikeminnow, we consider our diet composition estimates to be conservative estimates i.e., we believe that a larger proportion of the Northern Pikeminnow diet is fish and juvenile Chinook Salmon than our results suggest.
+<!--
 Although we only identified fish or fish parts within `r round(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"] * 100, 1)`% of Northern Pikeminnow examined and that fish remnants were only `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall stomach contents, we believe those numbers to be misleading...
-
-<!--
-Concerns with gastric lavage.
-We only observed juvenile Chinook salmon (and/or other fishes) in a very small number of stomach contents.
-However, we don't necessarily believe that to mean that pikeminnow in Deadwater Slough rarely consume juvenile Chinook salmon.
-Are there other cases where gastric lavage failed to document predation?
-How soon after consumption do you need to take a sample?
-So instead, we made some assumptions about the proportion of a pikeminnows diet consisting of Chinook salmon.
 -->

### 2021-12-30 analysis/paper/paper.Rmd
TB methods edits

@@
 The average length of Northern Pikeminnow caught during the fall and spring efforts were `r avg_fall_NPM` and `r avg_spring_NPM` mm, respectively, which calculated to average starting weights of 504.4 and 670.5 g.
-We assumed no growth in individual Northern Pikeminnow, and so predator start and end weights are equal, i.e., we assumed the population biomass was stable.
+We assumed no growth in individual Northern Pikeminnow, and so predator start and end weights are equal (i.e., we assumed the population biomass was stable).
 Mean daily water temperatures were summarized from 15-minute interval temperature readings available March 3, 2013 to June 14, 2021 from a gage station approximately 22 river kilometers downstream of Deadwater Slough near Shoup, Idaho [@USGS2016].
 
-The largest uncertainty among model parameters was the proportion of the Northern Pikeminnow diet consisting of fish (e.g., juvenile Chinook Salmon) versus other food items (e.g., invertebrates).
-Therefore, we varied the proportion of the diet consisting of fish versus invertebrates to range from 10-90%, in 10% intervals, for both categories.
+The largest uncertainty among model parameters was the proportion of the Northern Pikeminnow diet consisting of fish versus non-fish prey items (e.g., invertebrates).
+Therefore, we conducted nine model runs with varying proportions of fish in the diet ranging from 10-90%, in 10% increments.
 Energy densities for invertebrates was fixed at 3,000 J/g.
-We estimate the g of (any) fish consumed for each model run.
-For simplicity, we assume that any fish prey have the same energy densities as juvenile Chinook Salmon and we do not differentiate juvenile Chinook Salmon from other fish prey than may be present in Deadwater slough (e.g., juvenile steelhead, juvenile Sockeye Salmon, Redside Shiner).
+Because we were unable to differentiate juvenile Chinook Salmon from other fish prey consumed, we assume all fish prey have the same energy densities as juvenile Chinook Salmon.
 Separate models were also run for a 77-day fall period from September 15 - November 30 and a 92-day period in the spring from March 1 - May 31 to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG.
@@
 
-<!-- During this time, water temperatures typically exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon. -->
-
-To estimate the total biomass of fish potentially consumed by all Northern Pikeminnow in Deadwater Slough, we then multiplied the grams of fish consumed by an individual Northern Pikeminnow  during each period (fall and spring) by the estimated predator population size for that period.
-The total biomass was further converted to an estimate of the total number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
-The weight of an average DSR and NRR emigrants was taken from fish caught at seven rotary screw traps located throughout the Upper Salmon MPG upstream of Deadwater Slough during the fall and spring time periods.
+To estimate the total biomass of Chinook Salmon potentially consumed by all Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated predator population size during each period (fall and spring).
+The estimate of total biomass was converted to the total number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
+The average weights of DSR and NRR emigrants were taken from fish caught at seven rotary screw traps located throughout the Upper Salmon MPG upstream of Deadwater Slough during the fall and spring time periods.
 This calculation assumes that all fish prey items consumed by Northern Pikeminnow are, on average, the same size as DSR and NRR emigrants.

### 2021-12-30 analysis/paper/paper.Rmd
TB results edits

@@
 
-We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 efforts, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish caught using hook-and-line angling.
+Using hook-and-line angling, we caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 efforts, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish.
 The CPUE, measured as number of Northern Pikeminnow caught per angler hour, was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021 (Figure \@ref(fig:cpue-fig)).
-During the mark-recapture efforts, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019 which included `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals (i.e., `r as.character(english(sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m))` individual was captured more than once).
-There were an additional `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020 which included `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals.
+During the mark-recapture efforts, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019 which included `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals.
+There were an additional `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020 consisting of `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals.
 Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.
@@
 Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` (Table \@ref(tab:abund-tab)).
-In both cases, the multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied (Figure \@ref(fig:abund-fig)).
-Table \@ref(tab:abund-tab) provides point estimates and confidence intervals for each estimator for the fall mark-recapture efforts.
+In both cases, the multiple census estimators were larger than the single census estimators.
+Confidence intervals for the fall 2020 effort were more varied, likely resulting from fewer recaptures in that year.
 
 All mark-recapture estimators yielded estimates with overlapping confidence intervals.
-However, because our sampling design most closely matched a multiple census estimator (i.e., we released marked fish each day and recaptures from all previous days were recorded for each day), we fell those estimators most appropriate.
+However, because our sampling design most closely matched a multiple census estimator (i.e., we released marked fish each day and recaptures from all previous days were recorded for each day) and required a small sample size modification, we used the Schnabel estimates of abundance for further analyses and results.
+Accordingly, the mean of Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`).
+By multiplying the ratio (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`) between the spring CPUE (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"], 2)`) and mean of the fall CPUEs (`r round(mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring 2021.
+
+<!--However, because our sampling design most closely matched a multiple census estimator (i.e., we released marked fish each day and recaptures from all previous days were recorded for each day), we fell those estimators most appropriate.
 Additionally, because the Schnabel estimator includes a small sample size modification, as in our case (i.e., few recaptures), we feel it is most appropriate.
 The Schnabel estimator has the added benefit of more precise confidence intervals and is more straightforward than the Schumacher-Eschmeyer estimate as it is essentially the weighted average among multiple Chapman-modified Lincoln-Peterson estimates.
-Therefore, further analyses and results will be based on the Schnabel estimates of abundance.
-
-Using the mean of the Schnabel estimates of Northern Pikeminnow abundance across the two fall sampling events, `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`), and multiplying the ratio (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`) between the spring CPUE (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"], 2)`) and mean of the fall CPUEs (`r round(mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring 2021.
+Therefore, further analyses and results will be based on the Schnabel estimates of abundance.-->
 
@@
 
-```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow using different mark-recapture estimators and for the fall sampling events. Note the y-axis is truncated at 100."}
+```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow using different mark-recapture estimators and for the fall sampling events.
+Single census estimators are shown as circles; multiple census estimators are shown as triangles.
+Note the y-axis is truncated at 100 and the upper confidence interval for the Schumacker-Eschmeyer estimator in 2020 extends beyond that."}
 all_est %>%
@@
                 ~ . / 1000)) %>%
+  mutate(census = case_when(grepl("Chapman", model) ~ "single",
+                            grepl("Petersen", model) ~ "single",
+                            grepl("Schnabel", model) ~ "multiple",
+                            grepl("Schumacher-Eschmeyer", model) ~ "multiple")) %>%
   ggplot(aes(x = model,
              y = N,
-             color = model)) +
-  geom_point(size = 4) +
+             #color = model
+             )) +
+  geom_point(aes(size = 4,
+                 shape = census)) +
+  scale_shape_manual(values = c(17, 19)) +
   geom_errorbar(aes(ymin = Lci,

### 2021-12-30 analysis/paper/paper.Rmd
TB Discussion edits

@@
 
+```{r psd}
+s_size = 250
+q_size = 380
+
+psd_df = cmr_df %>%
+  select(species, count, length, recap, event_name) %>%
+  filter(recap  != TRUE)
+
+fs = psd_df %>%
+  filter(length >= s_size) %>%
+  nrow()
+
+fq = psd_df %>%
+  filter(length >= q_size) %>%
+  nrow()
+
+psd = 100 * (fq/fs)
+
+```
+
+Northern Pikeminnow in Deadwater Slough were not only abundant, but anglers reported that many were large-bodied.
+For that reason, we calculated the proportional stock density (PSD) for Northern Pikeminnow in Deadwater Slough, which is a measure of species size structure.
+PSD is a ratio, typically expressed as a percentage, between the number of "quality_sized" (larger) individuals and "stock-sized" individuals.
+
+$$
+PSD_{i} = 100 * \frac{FQ_{i}}{FS_{i}}
+$$
+
+where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
+Stock and quality size definitions vary by species.
+We calculated the PSD for Northern Pikeminnow in Deadwater Slough using `r prettyNum(s_size, digits = 0)` mm TL for stock-size and `r prettyNum(q_size, digits = 0)` mm TL for quality-size [@Winther2020].
+
 ## Gastric Lavage
@@
 
+The PSD for Northern Pikeminnow in Deadwater Slough across all three efforts was `r prettyNum(psd, digits = 1)`% i.e., `r prettyNum(psd, digits = 1)`% of captured fish were over the quality size classification defined by @Winther2020.
+For comparison, @Winther2020 reported a PSD for Northern Pikeminnow captured below Bonneville Dam of 41% and 18% in Bonneville Reservoir, suggesting that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to populations reported elsewhere in the Columbia River.
+In fact, the state record Northern Pikeminnow, which measured `r max(cmr_df$length, na.rm = T)` mm (`r round(max(cmr_df$length, na.rm = T) / 25.4, 1)` in) TL, was caught in Deadwater Slough during the fall 2020 effort.
+
 ## Gastric Lavage
@@
 The mean population size of Northern Pikeminnow in the Deadwater Slough during the fall season was estimated to be `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` and during the spring 2021 to be `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")`.
-Those estimates translate to linear densities of `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms), big.mark = ",")` which is `r round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms / john_day_dens)` and `r round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms / john_day_dens)` times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r as.character(english(round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms / low_columbia_fish_km)))` and `r as.character(english(round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms / low_columbia_fish_km)))` times greater than projections for the lower Columbia River [@Beamesderfer1996].
-These densities for Deadwater Slough initially seem unrealistic especially given the size (volume) of the Salmon River at the slough relative to the Columbia River.
-However, we did catch `r prettyNum(nrow(cmr_df), big.mark = ",")` Northern Pikeminnow over the course of the study, of which only `r sum(cmr_df$capture_type == "Recapture", na.rm = T)` fish (one in every `r round(nrow(cmr_df) / sum(cmr_df$capture_type == "Recapture", na.rm = T))`) were recaptures.
-Further, @Beamesderfer1991 did acknowledge that they were unsuccessful in finding a suitable method for sampling offshore; they noted that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective and so their abundance estimates only apply to nearshore areas.
-This was not the case at Deadwater Slough where maximum depth was approximately 10-15 m with a relatively homogeneous bottom contour i.e., angling was effective for the entire area.
-Our estimates of Northern Pikeminnow areal density are also higher than those reported elsewhere.
+Those estimates translate to linear densities of `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms), big.mark = ",")` which is `r round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms / john_day_dens)` and `r round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms / john_day_dens)` times greater, respectively, than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r as.character(english(round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms / low_columbia_fish_km)))` and `r as.character(english(round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms / low_columbia_fish_km)))` times greater than projections for the lower Columbia River [@Beamesderfer1996].
+Although estimated densities for Deadwater Slough are high given the relative size (volume) of the Salmon River at the slough relative to the Columbia River, high CPUE combined with low recapture rates (one in every `r round(nrow(cmr_df) / sum(cmr_df$capture_type == "Recapture", na.rm = T))` fish were recaptures) suggest this area may support exceptionally high predator densities.
+Further, @Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective.
+Conversely, at Deadwater Slough the maximum depth was approsimately 10-15 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach.
+Our estimates of Northern Pikeminnow areal density are also higher than those reported elsewhere in the Columbia River basin.
 We estimate the density of Northern Pikeminnow in Deadwater Slough, as fish per hectare, to be `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_hectares), big.mark = ",")` for fall and `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_hectares), big.mark = ",")` for spring.
 For comparison, the only other available estimates of areal density, that the authors are aware of, are 4.4 fish per hectare for John Day Reservoir [@Beamesderfer1991] and 15 fish per hectare in Lake Washington [@Bartoo1977].
-
-```{r psd}
-s_size = 250
-q_size = 380
-
-psd_df = cmr_df %>%
-  select(species, count, length, recap, event_name) %>%
-  filter(recap  != TRUE)
-
-fs = psd_df %>%
-  filter(length >= s_size) %>%
-  nrow()
-
-fq = psd_df %>%
-  filter(length >= q_size) %>%
-  nrow()
-
-psd = 100 * (fq/fs)
-
-```
-
-Northern Pikeminnow in Deadwater Slough were not only abundant, but anglers noticed that many were also large-bodied.
-For that reason, we calculated the proportional stock density (PSD) for Northern Pikeminnow in Deadwater Slough, which is a measure of species size structure.
-PSD is a ratio, typically expressed as a percentage, between the number of "quality_sized" (larger) individuals and "stock-sized" individuals.
-
-$$
-PSD_{i} = 100 * \frac{FQ_{i}}{FS_{i}}
-$$
-
-where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
-Stock and quality size definitions vary by species.
-We calculated the PSD for Northern Pikeminnow in Deadwater Slough using `r prettyNum(s_size, digits = 0)` mm TL for stock-size and `r prettyNum(q_size, digits = 0)` mm TL for quality-size [@Winther2020].
-The PSD for Northern Pikeminnow in Deadwater Slough across all three efforts was `r prettyNum(psd, digits = 1)`% i.e., `r prettyNum(psd, digits = 1)`% of captured fish were over the quality size classification defined by @Winther2020.
-For comparison, @Winther2020 reported a PSD for Northern Pikeminnow captured below Bonneville Dam of 41% and 18% in Bonneville Reservoir, suggesting that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to populations reported elsewhere in the Columbia River.
-In fact, the state record Northern Pikeminnow, which measured `r max(cmr_df$length, na.rm = T)` mm (`r round(max(cmr_df$length, na.rm = T) / 25.4, 1)` in) TL, was caught in Deadwater Slough during the fall 2020 effort.
+Given the dearth of current data available in the Columbia River basin, it is unclear how Northern Pikeminnow abundance estimates in this study compare to elsewhere in the Columbia River Basin.
+@Widener2021 has identified a need for more current information on piscine predator populations in the Columbia River.
 
@@
 
+All mark-recapture abundance estimators suggest a large number of Northern Pikeminnow occupying Deadwater Slough.
+Several assumptions in our estimators may have influenced the magnitude of our results.
+First, all models assumed a closed population, however, it is possible that immigration and emigration occurred.
+Nevertheless, our estimators still provide unbiased estimates of abundance assuming the immigration and emigration rates are equal between marked and unmarked fish.
+Given the size of the sampling area, and the limited length of the sampling period (two weeks), that assumption is likely met.
+The assumption of a closed population can also affect results if mortality occurs for some marked fish released back to the population.
+During our field processing, we did not observe mortalities or injury post release; however, the potential for mortality amongst marked and released fish cannot be discounted.
+Next, our estimates assumed that the second (and subsequent) sample(s) are random samples of the population.
+Yet, angling methods may have a size selection bias, thereby limiting our abundance estimates to a size threshold susceptible to angling.
+We also consider the validity of our assumption of equal catchability of individual fish between sampling events to have introduced potential bias in our results.
+Unequal catchability for individual fish is difficult to assess, especially if being caught once influences the likelihood of recapture.
+During our study, we recaptured two individual fish during multiple surveys indicting that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.--> We also recognize that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
+Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit the "good" fishing locales.
+If marked fish were not recently released or mixed back into those areas that would make marked fish less susceptible to capture.
+Violations of the assumption of equal catchability between marked and unmarked fish may lead to an overestimate of abundance of Northern Pikeminnow in our study if marked fish have lower catchability.
+
+<!--
 All mark-recapture abundance estimators suggest a large number of Northern Pikeminnow occupying Deadwater Slough.
@@
 Violations of assumption (2) would lead to an overestimate of abundance of Northern Pikeminnow which is a real possibility in our study.
-
-The spring abundance estimate was smaller than either fall abundance estimate.
-However, our approach assumed equal capture probabilities between the fall and spring sample efforts.
-If it does differ, we suspect that the capture probability in the spring would be lower due to higher flows (i.e., reduced densities), which would lead to our spring abundance estimate being biased low.
-Therefore, we presume our abundance estimate of Northern Pikeminnow for the spring 2021 effort to be a conservative estimate.
+-->
+
+The spring abundance estimate was smaller than fall abundance estimates.
+Our approach assumed equal capture probabilities between the fall and spring sample efforts.
+Yet, we suspect that the capture probability in the spring could be lower as higher spring flows reduce Northern Pikeminnow densities biasing our spring abundance estimate low.
+Therefore, we believe our spring abundance estimate to be conservative.
 
@@
 We performed gastric lavage on nearly all Northern Pikeminnow collected during this study.
-Aside from juvenile Chinook Salmon, identified species included Redside Shiner, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*; although decomposition from digestion rendered many fish contents unidentifiable.
+In addition to juvenile Chinook Salmon, species included Redside Shiner, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*; although decomposition from digestion rendered many fish contents unidentifiable.
 While this approach provided useful information on prey species, we have concerns about using angling and gastric lavage to quantify diet composition in Northern Pikeminnow.
@@
 
-An average-size Northern Pikeminnow from Deadwater Slough is estimated to consume `r median(fall_g_summ$cumm_g)` g (range `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) and `r median(spring_g_summ$cumm_g)` g (range `r min(spring_g_summ$cumm_g)` - `r max(spring_g_summ$cumm_g)` g) of fish during spring and fall, respectively, to maintain its body weight using mean daily water temperatures recorded at a nearby stream gage [@USGS2016].
-This equates to `r median(fall_g_summ$cumm_g) + median(spring_g_summ$cumm_g)` g of fish consumed during the DSR and NRR peak emigrations assuming the median value among diet scenarios (i.e., if 50% of their diet consists of fish); values ranged from a minimum of `r min(fall_g_summ$cumm_g) + min(spring_g_summ$cumm_g)` g if only 10% of the diet is fish to a maximum of `r max(fall_g_summ$cumm_g) + max(spring_g_summ$cumm_g)` g if 90% of the diet is fish.
-
-<!-- The following needs to be updated after we finalize some results -->
-This approach currently assumes that juvenile Chinook Salmon are the only fish prey available to or preyed upon by Northern Pikeminnow.
-It further assumes that all prey available have energy densities equal to juvenile Chinook Salmon, which may not be far from the truth, given that other salmonids including juvenile steelhead and Sockeye Salmon are likely also prevalent as available prey in Deadwater Slough, especially during the spring migration.
+An average-size Northern Pikeminnow from Deadwater Slough is estimated to consume `r median(fall_g_summ$cumm_g)` g (range `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) and `r median(spring_g_summ$cumm_g)` g (range `r min(spring_g_summ$cumm_g)` - `r max(spring_g_summ$cumm_g)` g) of fish during spring and fall, respectively, to maintain its body weight.
+This equates to `r median(fall_g_summ$cumm_g) + median(spring_g_summ$cumm_g)` g of fish consumed during the DSR and NRR peak emigrations, assuming the median value among diet scenarios (i.e., if 50% of their diet consists of fish).
+Consumption ranged from `r min(fall_g_summ$cumm_g) + min(spring_g_summ$cumm_g)` for a diet comprised of 10% fish to `r max(fall_g_summ$cumm_g) + max(spring_g_summ$cumm_g)` for a diet comprised of 90% fish.
+
+Our approach assumed that all available prey consumed by Northern Pikeminnow have energy densities equal to juvenile Chinook Salmon.
+Other salmonid including juvenile steelhead and Sockeye Salmon are likely also available prey in Deadwater Slough, especially during the spring migration.
 Additionally, hatchery smolt releases (Chinook Salmon, steelhead, and Sockeye Salmon) are also prevalent in the mainstem Salmon River during the spring migration.
-However, our study is focused on potential impacts to natural-origin DSR and NRR Chinook Salmon, in which case, let’s consider a “thought exercise” where DSR and NRR Chinook Salmon represent only 10% and 25% of total fish prey consumed by Northern Pikeminnow during these periods.
+Other studies have used a similar approach where the energy densities of different salmonids were generalized using an estimate from a single species [@Peterson1999].<!-- The following needs to be updated after we finalize some results --> However, our study is focused on potential impacts to natural-origin DSR and NRR Chinook Salmon, in which case, let's consider a thought exercise where DSR and NRR Chinook Salmon represent only 10% and 25% of total fish prey consumed by Northern Pikemnnow during these periods.
 Under these scenarios, the average Northern Pikeminnow would consume 8.95 g (range 4.52 – 10.05 g) if 10% of prey were natural-origin Chinook Salmon and 22.38 g (range 11.31 – 25.12) if 25%.
 
-To estimate the total number of possible juvenile Chinook Salmon that could be consumed, we divided the cumulative consumption in grams for an individual Northern Pikeminnow by the average weight of a Chinook Salmon emigrant (taken from all rotary screw traps in the Upper Salmon MPG) to come up with a juvenile equivalency, and then multiplied this by a "stable" estimate of abundance for Northern Pikeminnow in Deadwater Slough, resulting in an estimate of the equivalent toal number of Chinook Salmon juveniles consumed.
-If we assume an average juvenile Chinook Salmon of 10.6 g, the mean from DSR and NRR Chinook Salmon emigrants in the Upper Salmon MPG, this equates to 0.84 juvenile Chinook Salmon (range 0.43 – 0.94) under the 10% scenario and 2.11 juvenile Chinook Salmon (range 1.07 – 2.37) under the 25% scenario.
-
+Our estimated consumption of Chinook Salmon juveniles by Northern Pikeminnow was calculated using a static predator population.
+If we assume an average juvenile Upper Salmon MPG Chinook Salmon of 10.6 g, this equates to 0.84 juvenile Chinook Salmon (range 0.43 – 0.94) under the 10% scenario and 2.11 juvenile Chinook Salmon (range 1.07 – 2.37) under the 25% scenario.
 Let’s now consider a stable Northern Pikeminnow population size of 10,000 (roughly the mean of fall and spring abundance estimates, divided by two to account for half of marked individuals not being available for recapture).
@@
 <!-- Also update this after finalizing results -->
-Finally, let's take our thought exercise one additional step and estimate the number of additional adult Chinook Salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough was reduced or eliminated.
+Our primary objective was to assess the impact of predation on juvenile Chinook Salmon in the Upper Salmon MPG.
+However, there is also interest in estimating the number of additional adult Chinook Salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough was reduced or eliminated.
 Our approach here was to multiply the total estimated juvenile Chinook Salmon consumed by an estimate of Granite-to-Granite smolt-to-adult survival rate (SAR) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.
@@
 During the initial sampling effort in fall 2019, a two-person crew walked the entire reach including surrounding and upstream areas scanning for passive integrated transponder (PIT) tags that are implanted into thousands of natural- and hatchery-origin salmonids in the Upper Salmon River, annually, to monitor movement and survival.
-During that informal survey, 10 PIT tags were recovered near active bird nests or in an upstream anastomizing reach where herons and eagles are prevalent, suggesting that mortality may have been a result of avian predation.
-The PIT tag histories in [PTAGIS](https://www.ptagis.org/) indicate these tags were implanted into a combination of juvenile Chinook Salmon (3), Sockeye Salmon (4), and steelhead (3), all juveniles.
+During that informal survey, 10 PIT tags were recovered near active bird nests and in an upstream anastomizing reach where herons and eagles are prevalent, suggesting that mortality may have been a result of avian predation.
+The PIT tag histories in [PTAGIS](https://www.ptagis.org/) indicate these tags were implanted into a combination of juvenile Chinook Salmon (3), Sockeye Salmon (4), and steelhead (3).
 Two of the Chinook Salmon and one steelhead were natural-origin, the remaining were of hatchery-origin.
@@
 Although we did observe Belted Kingfishers, mergansers, and Great Blue Herons at Deadwater Slough during field efforts; Great Blue Herons were especially prevalent in the anastomizing reach immediately upriver of the slackwater reach.
-Additionally, there is documentation of individual Double-crested Cormorants [@eBird2021] at the Deadwater Slough, although the site is not within their breeding range; rather, it is part of a migration corridor.
+Additionally, there is documentation of individual Double-crested Cormorants [@eBird2021] at the Deadwater Slough; however, the site is not within their breeding range, rather, it is part of a migration corridor.
 Given the current avian species known to occupy Deadwater Slough, it is unlikely that avian predation on juvenile salmonids is comparable to elsewhere in the Columbia River basin with large piscivorous bird colonies.
 Nevertheless, we hypothesize that the reservoir-like conditions and associated habitat features at the Deadwater Slough may increase the probability of avian predation on juvenile Chinook Salmon and other ESA-listed salmonids from the many piscivorous birds known to use the site.
-Future estimates of predation would benefit from consideration of the contribution of piscivorous avian predators.
+Future studies of predation would benefit from estimates from piscivorous avian predators.
 
@@
 <!-- Again, some of the following numbers need to be updated after results are formalized -->
-Chinook Salmon populations in the Upper Salmon River MPG support local fisheries and economies in years with sufficient abundance and have substantial cultural value in the region.
-These populations additionally support downriver commercial and subsistence fisheries in the Snake and Columbia rivers.
-Recovery of the MPG is desired which requires that at least five of the eight populations must meet criteria set forth by @McElhany2000 and @ICTRT2007 for the MPG to be considered viable and for recovery of the Snake River Evolutionary Significant Unit.
+Recovery of the Upper Salmon MPG is desired to support local fisheries and economies.
+At least five of the eight populations must meet criteria set forth by @McElhany2000 and @ICTRT2007 for the MPG to be considered viable and for recovery of the Snake River Evolutionary Significant Unit.
 Spawner abundance is perhaps the most important metric considered in determining a population's viability.
-We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces annual adult natural-origin returns by 26.4 to 145.6 to upriver populations.
-Although this does not sound like a substantial amount, it represents 4.7% - 26.1% of the total mean adult escapement of 558 reported for the Upper Salmon River MPG in 2017 - 2019 [@Kinzer2020].
-Presumably, reductions in adult returns also affect natural-origin Sockeye Salmon and steelhead as well as hatchery populations in the Upper Salmon which provide for recreational fishing opportunities.
-Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations for multiple ESA-listed species.
+We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by 26.4 to 145.6 adults, annually.
+This is between 4.7% and 26.1% of the total mean adult escapement<!--of 558--> reported for the Upper Salmon River MPG in 2017 - 2019 [@Kinzer2020].
+These estimates only take into account predation on juvenile Chinook Salmon by Northern Pikeminnow, but presumably, juvenile steelhead and Sockeye Salmon are also prey items in Deadwater Slough, including hatchery releases.
+Therefore, adult returns of steelhead and Sockeye Salmon are affected as well, including for hathchery populations which provide for recreational fishing opportunities.
+Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including multiple ESA-listed species.
 Because the Deadwater Slough is part of the migratory pathway for emigrating salmonids, including for multiple species and populations, the impact of that benefit could be higher than tributary rehabilitation actions, which typically benefit a single population.
-Moreover, the deepened, slack water conditions that favor northern pikeminnow at Deadwater Slough are indirectly the result of manmade activities i.e., the failure of a manmade mining reservoir dam.
-Given these reasons, it seems that Deadwater Slough could be a candidate for management or restoration actions to benefit local Chinook Salmon populations.
+Moreover, the deepened, slack water conditions that favor Northern Pikeminnow at Deadwater Slough are a result of the failure of a manmade mining reservoir dam.
+Taken together, the Deadwater Slough is a favorable candidate for management or restoration actions to benefit local Chinook Salmon populations.
 
@@
 A local bounty program in the Salmon River has the potential to boost the local economy (to an unknown degree) which has been negatively impacted by low returns of Chinook Salmon, steelhead, and Sockeye Salmon in recent decades.
-Many local business rely on license and gear sales for local sport fisheries, but low salmon and steelhead returns have eliminated or reduced the duration of fisheries in some years, and a local Northern Pikeminnow bounty program could allow the local economy to recoup some of those lost revenues.
+Many local businesses rely on license and gear sales for local sport fisheries, but low salmon and steelhead returns have eliminated or reduced the duration of fisheries in some years, and a local Northern Pikeminnow bounty program could allow the local economy to recoup some of those lost revenues.
 On the downside, bounty programs require continual effort and harvest to suppress the population size and so would require year of continued funding and local buy-in to remain effective.

### 2021-12-30 analysis/paper/paper.Rmd
knit complete draft and saved to repo

@@
 While this approach provided useful information on prey species, we have concerns about using angling and gastric lavage to quantify diet composition in Northern Pikeminnow.
-First, previous studies have indicated that diets of angled fish differ from those collected by other methods (e.g., electrofishing, [@Hodgson1998]), and further, angled fish are much more likely to have empty stomachs [@Jurajda2016].
+First, previous studies have indicated that diets of angled fish differ from those collected by other methods (e.g., electrofishing, [@Hodgson1998]), and further, angled fish are more likely to have empty stomachs [@Jurajda2016], perhaps because they are hungry and actively searching for food.
 Second, unlike other predators such as Smallmouth Bass and Walleye that are commonly sampled using gastric lavage [@Kamler2001], Northern Pikeminnow are cyprinids and therefore lack a true stomach.

### 2021-12-30 analysis/paper/paper.Rmd
added last references

--- 2021-12-30 analysis/paper/paper.Rmd
@@
 While this approach provided useful information on prey species, we have concerns about using angling and gastric lavage to quantify diet composition in Northern Pikeminnow.
-First, previous studies have indicated that diets of angled fish differ from those collected by other methods (e.g., electrofishing, [@Hodgson1998]), and further, angled fish are much more likely to have empty stomachs [@Jurajda2016].
+First, previous studies have indicated that diets of angled fish differ from those collected by other methods (e.g., electrofishing, [@Hodgson1988]), and further, angled fish are more likely to have empty stomachs [@Jurajda2016], perhaps because they are hungry and actively searching for food.
 Second, unlike other predators such as Smallmouth Bass and Walleye that are commonly sampled using gastric lavage [@Kamler2001], Northern Pikeminnow are cyprinids and therefore lack a true stomach.
@@
 Additionally, hatchery smolt releases (Chinook Salmon, steelhead, and Sockeye Salmon) are also prevalent in the mainstem Salmon River during the spring migration.
-Other studies have used a similar approach where the energy densities of different salmonids were generalized using an estimate from a single species [@Peterson1999].<!-- The following needs to be updated after we finalize some results --> However, our study is focused on potential impacts to natural-origin DSR and NRR Chinook Salmon, in which case, let's consider a thought exercise where DSR and NRR Chinook Salmon represent only 10% and 25% of total fish prey consumed by Northern Pikemnnow during these periods.
+Other studies have used a similar approach where the energy densities of different salmonids were generalized using an estimate from a single species [@Petersen1999].<!-- The following needs to be updated after we finalize some results --> However, our study is focused on potential impacts to natural-origin DSR and NRR Chinook Salmon, in which case, let's consider a thought exercise where DSR and NRR Chinook Salmon represent only 10% and 25% of total fish prey consumed by Northern Pikemnnow during these periods.
 Under these scenarios, the average Northern Pikeminnow would consume 8.95 g (range 4.52 – 10.05 g) if 10% of prey were natural-origin Chinook Salmon and 22.38 g (range 11.31 – 25.12) if 25%.

### 2022-01-03 analysis/paper/paper.Rmd
complete draft

@@
 Indeed, Northern Pikeminnow have been estimated to consume 8% of the approximately 200 million juvenile salmonids emigration through those rivers, annually.
-Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above the reservoir systems.
+Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above these reservoir systems.
 In this study, we examine predation by Northern Pikeminnow on juvenile salmonids, focusing on Chinook Salmon *Oncorhynchus tshawytscha*, in a slow-water reach of the Salmon River, Idaho known as Deadwater Slough.
@@
 mykiss*.
-These fishes were historically abundant throughout the drainage basin and consequently have tremendous ecological, cultural, and economic value.
+These fishes were historically abundant throughout the basin and consequently have tremendous ecological, cultural, and economic value.
 Over the last century, salmonid stocks throughout the Snake and Columbia rivers have undergone significant declines related to habitat modification in the freshwater environment (e.g., removal of beavers, river channel simplification, water withdrawals, hydrosystem development) and poor marine conditions, both affecting their survival and recruitment at multiple life stages.
@@
 Upstream habitats containing slower water velocities and other attributes that support piscivorous predators may overlap with essential habitat for some salmonid species and life stages.
-One such example is Deadwater Slough, an approximately `r round(ddw_rkms, 1)` kilometer long reach of unnaturally slow and deep water in the Salmon River, Idaho.
+One such example is Deadwater Slough, an approximately `r round(ddw_rkms, 1)` km long reach of unnaturally slow and deep water in the Salmon River, Idaho.
 The Deadwater Slough is located in a river reach containing historically important overwinter rearing and spring emigration habitat for juvenile Chinook Salmon *O.
@@
 
-Quantifying mortality associated with piscine predation in upstream habitats of the Salmon River is of particular interest for spring/summer Chinook Salmon in the Upper Salmon major population group (MPG).
-The Salmon River was historically the most productive tributary for spring/summer Chinook Salmon in the Columbia River basin [@Nemeth1999], but those populations have become depleted in recent decades.
-Currently, the Upper Salmon MPG supports eight independent, extant spring/summer Chinook Salmon populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
+Quantifying mortality associated with piscine predation in upstream habitats of the Salmon River is of particular interest for spring/summer-run Chinook Salmon (hereafter Chinook Salmon) in the Upper Salmon major population group (MPG).
+The Salmon River was historically the most productive tributary for Chinook Salmon in the Columbia River basin [@Nemeth1999], but those populations have become depleted in recent decades.
+Currently, the Upper Salmon MPG supports eight independent, extant populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
 The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a].
+The DSR migrants leave their natal area as subyearlings between June and November and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts.
+Alternatively, NRR migrants remain in their natal areas for approximately one year after emergence until emigration to the ocean as smolts.
 Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency.
@@
 
-Recent studies have attempted to examine the downstream movement, distribution, and (apparent) survival of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach.
+Recent studies have examined the downstream movement, distribution, and (apparent) survival of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach.
 Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have substantially reduced survival in the Deadwater Slough reach compared to adjacent reaches [@Axel2015].
@@
 1. Estimate the abundance (or relative abundance) of potential predators in the Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations.
-2. Document predation on juvenile Chinook Salmon during the emigration periods using gastric lavage.
+2. Document predation on juvenile Chinook Salmon during the DSR and NRR emigration periods using gastric lavage.
 3.
@@
 
-We also consider how consumption of juvenile Chinook Salmo emigrants at Deadwater Slough by Northern Pikeminnow may impact adult returns to the Upper Salmon MPG.
+We also consider how consumption of juvenile Chinook Salmon emigrants at Deadwater Slough by Northern Pikeminnow may impact adult returns to the Upper Salmon MPG.
 
@@
 
-The Deadwater Slough is an approximately 1.5 km section of the mainstem Salmon River located approximately 6 river km downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
+The Deadwater Slough is an approximately `r round(ddw_rkms, 1)` km section of the mainstem Salmon River located approximately 6 river km downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
 The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
-Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River [@Emerson1973], thereby creating a slow, deep section in the river, spanning approximately `r round(ddw_acres)` acres (`r round(ddw_hectares)` hectares), and averaging `r round(ddw_avg_width_m)` m width.
-
-## Northern Pikeminnow Abundance
+Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River [@Emerson1973] forming an alluvial fan.
+The fan created a slow, deep section in the river, spanning approximately `r round(ddw_acres)` acres (`r round(ddw_hectares)` hectares), averaging `r round(ddw_avg_width_m)` m width, and resembling a small, narrow pond or reservoir.
+
+## Northern Pikeminnow Population Size
 
@@
 -->After evaluating all methods, angling proved to be the most effective method for capturing piscine predators while also minimizing potential impacts to ESA-listed adult steelhead present in the Salmon River during our sampling efforts.<!--, but recognizing that angling can have both species and size selectivity (bias).--> Therefore, the following analyses will focus on fish captured by angling, unless otherwise noted.
-During the fall 2019 effort it additionally became clear that Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and so is the focus of our study.
+During the fall 2019 effort it additionally became clear that Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and so is our focus.
 
@@
 The Schnabel estimator does not have an associated standard error; however, the 95% confidence intervals were calculated.
+The Schnabel estimator is essentially a weighted average of a series of Lincoln-Peterson estimators (with a Champman modification).
 Lastly, we explored one additional multiple-census estimator, the Schumacher-Eschmeyer, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
@@
 For that reason, we calculated the proportional stock density (PSD) for Northern Pikeminnow in Deadwater Slough, which is a measure of species size structure.
-PSD is a ratio, typically expressed as a percentage, between the number of "quality_sized" (larger) individuals and "stock-sized" individuals.
+PSD is a ratio, typically expressed as a percentage, between the number of "quality-sized" (larger) individuals and "stock-sized" individuals:
 
@@
 where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
-Stock and quality size definitions vary by species.
-We calculated the PSD for Northern Pikeminnow in Deadwater Slough using `r prettyNum(s_size, digits = 0)` mm TL for stock-size and `r prettyNum(q_size, digits = 0)` mm TL for quality-size [@Winther2020].
-
-## Gastric Lavage
+Quality and stock size definitions vary by species.
+We calculated the PSD for Northern Pikeminnow in Deadwater Slough using `r prettyNum(q_size, digits = 0)` mm TL for quality-size and `r prettyNum(s_size, digits = 0)` mm TL for stock-size [@Winther2020].
+
+## Stomach Contents
 
@@
 
-## Bioenergetics
+## Fish Consumption Potential
 
@@
 Energy densities for invertebrates was fixed at 3,000 J/g.
-Because we were unable to differentiate juvenile Chinook Salmon from other fish prey consumed, we assume all fish prey have the same energy densities as juvenile Chinook Salmon.
+Because we were unable to differentiate juvenile Chinook Salmon from other fish prey consumed during model runs, we assume all fish prey have the same energy densities as juvenile Chinook Salmon i.e., juvenile Chinook Salmon were used as surrogates for all fish prey items.
 Separate models were also run for a 77-day fall period from September 15 - November 30 and a 92-day period in the spring from March 1 - May 31 to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG.
 This resulted in 18 separate model runs (9 diet scenarios and 2 seasons).
-Each model run provided an estimate of the amount (grams) of fish consumed by an individual Northern Pikeminnow during the fall or spring time periods.
-
-To estimate the total biomass of Chinook Salmon potentially consumed by all Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated predator population size during each period (fall and spring).
-The estimate of total biomass was converted to the total number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
+Each model run provided an estimate of the total amount (grams) of fish consumed by an individual Northern Pikeminnow during the fall or spring time periods.
+To estimate the total biomass of fish potentially consumed by all Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated predator population size during each period (fall and spring).
+
+## Potential Impacts to Chinook Salmon Populations
+
+Multiplying the grams of fish consumed by an individual Northern Pikeminnow by the Northern Pikeminnow population size provides an estimate of total fish consumption potential for the population.
+However, the bioenergetics model runs do not differentiate juvenile Chinook Salmon from other fish prey items.
+Therefore, to understand potential impacts to local Chinook Salmon populations, we need to parse out the portion of total fish consumed that is juvenile Chinook Salmon.
+Unfortunately, estimates of the proportion of the Northern Pikeminnow fish prey that are juvenile Chinook Salmon are sparse, as well as information on the prevalence of juvenile Chinook Salmon relative to other fish prey items.
+Therefore, we instead consider hypothetical scenarios where Chinook Salmon are 1%, 10%, 30%, and 50% of fish prey consumed, by weight, which provide estimates of the total biomass (grams) of juvenile Chinook Salmon consumed by Northern Pikeminnow during the DSR and NRR emigration periods.
+
+The estimate of total juvenile Chinook Salmon biomass consumed was converted to the total number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
 The average weights of DSR and NRR emigrants were taken from fish caught at seven rotary screw traps located throughout the Upper Salmon MPG upstream of Deadwater Slough during the fall and spring time periods.
-This calculation assumes that all fish prey items consumed by Northern Pikeminnow are, on average, the same size as DSR and NRR emigrants.
 
@@
 
-## Northern Pikeminnow Abundance
+## Northern Pikeminnow Population Size
 
@@
 In both cases, the multiple census estimators were larger than the single census estimators.
-Confidence intervals for the fall 2020 effort were more varied, likely resulting from fewer recaptures in that year.
+Confidence intervals for the fall 2020 effort were more varied, likely due to fewer recaptures during that year.
 
@@
 
-## Gastric Lavage
+## Stomach Contents
 
@@
 To validate the efficacy of the method, we euthanized 75 Northern Pikeminnow after gastric lavage was completed and removed the remaining food items via dissection.
-We found that gastric lavage successfully removed food items in all dissected individual, indicating the method was effective.
+We found that gastric lavage successfully removed food items in all dissected individuals, indicating the method was effective.
 Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in some form in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) of those.
@@
 
-## Bioenergetics
-
-We estimated an individual, average-size (`r avg_fall_NPM` mm) Northern Pikeminnow to consume `r median(fall_g_summ$cumm_g)` g (range `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the DSR emigration to maintain their body size, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
-Similarly, we estimated an average-size (`r avg_spring_NPM` mm) Northern Pikeminnow to consume `r median(spring_g_summ$cumm_g)` g (range `r min(spring_g_summ$cumm_g)` - `r max(spring_g_summ$cumm_g)` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
-Given an average weight of `r avg_DSR_g` g for DSR emigrants, that would equate to `r round(median(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` (range `r round(min(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)`) fish consumed by an individual during fall and an additional `r round(median(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` (range `r round(min(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)`) fish consumed using an average weight of `r avg_NRR_g` g for NRR emigrants.
+## Fish Consumption Potential
+
+We estimated an individual, average-size (`r avg_fall_NPM` mm) Northern Pikeminnow to consume `r median(fall_g_summ$cumm_g)` g (range: `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the DSR emigration to maintain their body size, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
+Similarly, we estimated an average-size (`r avg_spring_NPM` mm) Northern Pikeminnow to consume `r median(spring_g_summ$cumm_g)` g (range: `r min(spring_g_summ$cumm_g)` - `r max(spring_g_summ$cumm_g)` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
+Given a mean Northern Pikeminnow population size of `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` from fall estimates and `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` for spring 2021 and using the median value from diet scenarios (i.e., 50% of their diet consists of fish), we'd estimate that the Northern Pikeminnow population can consume `r prettyNum(median(fall_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` g (range: `r prettyNum(min(fall_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` - `r prettyNum(max(fall_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` g) fish during the DSR emigration and an additional `r prettyNum(round(median(spring_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` g (range: `r prettyNum(round(min(spring_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` - `r prettyNum(round(max(spring_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` g) during the NRR emigration in Deadwater Slough.
+In total, the Northern Pikeminnow population can consume `r prettyNum(median(fall_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Fall"]) + median(spring_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Spring"]), big.mark = ",")` g of fish during the DSR and NRR emigrations to maintain their population size if we assume 50% of their diet consists of fish.
+
+## Potential Impacts to Chinook Salmon Populations
+
+1. Start with total g of fish consumed by Northern Pikemnnow population, summing DSR and NRR time periods.
+2. Parse out portion of fish total g consumed that is juvenile Chinook Salmon (1%, 10%, 30%, 50%).
+3. Divide by average weight of juvenile Chinook Salmon (10.6 g), the average of DSR (10.3 g) and NRR (10.9 g) emigrants.
+4. Estimate of total juvenile Chinook Salmon consumed.
+5. Multiply by SAR to estimate "adult equivalents".
+
+<!--
+Given an average weight of `r avg_DSR_g` g for DSR emigrants, that would equate to `r round(median(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` (range: `r round(min(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)`) fish consumed by an individual during fall and an additional `r round(median(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` (range: `r round(min(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)`) fish consumed using an average weight of `r avg_NRR_g` g for NRR emigrants.
 
@@
 So although we are uncertain of the diet proportions consisting of fish versus invertebrates, the impact of that uncertainty is somewhat constrained.
+-->
+
 <!--
@@
 Further, @Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective.
-Conversely, at Deadwater Slough the maximum depth was approsimately 10-15 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach.
+Conversely, at Deadwater Slough the maximum depth was approximately 10-15 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach.
 Our estimates of Northern Pikeminnow areal density are also higher than those reported elsewhere in the Columbia River basin.
@@
 
-## Mark-Recapture Model
+## Mark-Recapture Models
 
@@
 Next, our estimates assumed that the second (and subsequent) sample(s) are random samples of the population.
-Yet, angling methods may have a size selection bias, thereby limiting our abundance estimates to a size threshold susceptible to angling.
+Angling methods may have a size selection bias, thereby limiting our abundance estimates to a size threshold susceptible to angling.
 We also consider the validity of our assumption of equal catchability of individual fish between sampling events to have introduced potential bias in our results.
 Unequal catchability for individual fish is difficult to assess, especially if being caught once influences the likelihood of recapture.
-During our study, we recaptured two individual fish during multiple surveys indicting that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.--> We also recognize that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
+During our study, we recaptured two individual fish during multiple surveys indicating that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.--> We further recognize that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
 Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit the "good" fishing locales.
-If marked fish were not recently released or mixed back into those areas that would make marked fish less susceptible to capture.
+If marked fish were not recently released or mixed back into those areas that would result in marked fish being less susceptible to capture.
 Violations of the assumption of equal catchability between marked and unmarked fish may lead to an overestimate of abundance of Northern Pikeminnow in our study if marked fish have lower catchability.
@@
 While this approach provided useful information on prey species, we have concerns about using angling and gastric lavage to quantify diet composition in Northern Pikeminnow.
-First, previous studies have indicated that diets of angled fish differ from those collected by other methods (e.g., electrofishing, [@Hodgson1988]), and further, angled fish are more likely to have empty stomachs [@Jurajda2016], perhaps because they are hungry and actively searching for food.
+First, previous studies have indicated that diets of angled fish differ from those collected by other methods (e.g., electrofishing) [@Hodgson1988], and further, angled fish are more likely to have empty stomachs [@Jurajda2016], perhaps because they are hungry and actively searching for food.
 Second, unlike other predators such as Smallmouth Bass and Walleye that are commonly sampled using gastric lavage [@Kamler2001], Northern Pikeminnow are cyprinids and therefore lack a true stomach.
 Previous research indicates that performing gastric lavage on cyprinid species can be ineffective at quantifying diet composition [@Hartleb1995], or alternately, requires a modification in methodology to flush food items out the vent of the fish [@Wasowicz1994].
-Our intent, rather, was to simply document predation on juvenile Chinook Salmon in Deadwater Slough, and given the potential ineffectiveness of gastric lavage using angling and for Northern Pikeminnow, we consider our diet composition estimates to be conservative estimates i.e., we believe that a larger proportion of the Northern Pikeminnow diet is fish and juvenile Chinook Salmon than our results suggest.
+Our intent, rather, was to simply document predation on juvenile Chinook Salmon in Deadwater Slough, and given the potential ineffectiveness of gastric lavage using angling and for Northern Pikeminnow, we consider our diet composition estimates to be conservative estimates i.e., we believe that a larger proportion of the Northern Pikeminnow diet is fish and juvenile Chinook Salmon than our gastric results suggest.
 <!--
@@
 Although not formally assessed in this study, avian predators including Great Blue Herons *Ardea herodias* and Bald Eagles *Haliaeetus leucocephalus* are another potential source of mortality for juvenile Chinook Salmon and other ESA-listed salmonids in the Deadwater Slough.
-the Deadwater Slough is recognized as an important bird watching and nesting area due to the associated upstream braided channels and riparian and backwater habitats [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
+The Deadwater Slough is recognized as an important bird watching and nesting area due to the associated upstream braided channels and riparian and backwater habitats [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
 Several piscivorous bird species have been documented using Deadwater Slough including the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
-During the initial sampling effort in fall 2019, a two-person crew walked the entire reach including surrounding and upstream areas scanning for passive integrated transponder (PIT) tags that are implanted into thousands of natural- and hatchery-origin salmonids in the Upper Salmon River, annually, to monitor movement and survival.
+During the initial sampling effort in fall 2019, a two-person crew walked the entire reach including surrounding and upstream areas scanning for passive integrated transponder (PIT) tags that are implanted into thousands of natural- and hatchery-origin salmonids in the Upper Salmon River, annually.
 During that informal survey, 10 PIT tags were recovered near active bird nests and in an upstream anastomizing reach where herons and eagles are prevalent, suggesting that mortality may have been a result of avian predation.
@@
 We further appreciate the administrative support and guidance from Caitlin Alcott and Inter-Fluve.
+This manuscript benefitted from a review by Sean Gibbs.
 Funding for this study was provided by the Bureau of Reclamation, Pacific Northwest Regional Office (contract No.

### 2022-01-05 analysis/supplementary-materials/scripts/abund_est.R
minor tweaks to plotting code.

@@ -122,7 +122,8 @@ N_mods = cmr_df %>%
 
 # diagnostic plots
 mod = N_mods %>%
-  slice(1) %>%
+  # slice(1) %>%
+  slice(3) %>%
   pull(mr_model) %>%
   magrittr::extract2(1)
 
@@ -147,7 +148,9 @@ tibble(m = mod$m,
 N_mods %>%
   filter(model == "Schnabel") %>%
   unnest(data) %>%
-  # filter(m > 0) %>%
+  group_by(event_name) %>%
+  filter(m > 0 |
+           date == min(date)) %>%
   ggplot(aes(x = M,
              y = m/n)) +
   geom_point(size = 4) +
@@ -164,7 +167,7 @@ N_mods %>%
               linetype = 2)
 
 
-#-----------------------------------------------------------------
+# -----------------------------------------------------------------
 # using Lincoln-Petersen model, combining first 4 and last four days
 # define mark and recapture periods
 cap_periods = tibble(event_name = "Fall_2019",

### 2022-01-06 analysis/supplementary-materials/scripts/temp_comparison.R
added temp comparison script

@@ -1,0 +1,45 @@
+library(magrittr)
+library(tidyverse)
+library(here)
+library(lubridate)
+
+# shoup gage data
+shoup_df = read_tsv(here("analysis/data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
+                         comment = "#") %>%
+  rename(discharge_cfs = `47768_00060`,
+         discharge_cfs_cd = `47768_00060_cd`,
+         gage_ht_ft = `47769_00065`,
+         gage_ht_ft_cd = `47769_00065_cd`,
+         water_temp_c = `47773_00010`,
+         water_temp_c_cd = `47773_00010_cd`)
+
+# deadwater boat ramp tidbit data
+tidbit_df = read_csv(here("analysis/data/raw_data/deadwater_tidbit_compiled.csv")) %>%
+  mutate(datetime = paste(date, time)) %>%
+  mutate(across('datetime', ~ as.POSIXct(.x, format = "%m/%d/%Y %H:%M:%S"))) %>%
+  select(datetime, temp_c) %>%
+  mutate(source = "Tibdit")
+
+temp_df = shoup_df %>%
+  select(datetime, water_temp_c) %>%
+  mutate(source = "Shoup") %>%
+  rename(temp_c = water_temp_c) %>%
+  rbind(tidbit_df) %>%
+  mutate(year = year(datetime),
+         week = week(datetime)) %>%
+  group_by(source, year, week) %>%
+  summarize(temp_c = mean(temp_c, na.rm = T)) %>%
+  drop_na(week) %>%
+  mutate(source_yr = paste(source, year))
+
+temp_p = temp_df %>%
+  ggplot() +
+  geom_line(aes(x = week,
+                 y = temp_c,
+                 color = source_yr)) +
+  labs(x = "Week",
+       y = "Temperature (C)",
+       color = "Source and Year") +
+  scale_color_brewer(palette = "Set1") +
+  theme_bw()
+temp_p

### 2022-01-06 analysis/paper/paper.Rmd
cleaning paper code up; updated impact scenarios to include p_chnk; more work to do

@@
 library(magrittr)
+library(janitor)
+library(here)
+library(lubridate)
+library(FSA)
+library(english)
 library(lubridate)
 library(tidytable)
-library(janitor)
-library(purrr)
-library(here)
-library(FSA)
-library(knitr)
-library(kableExtra)
-library(stringi)
-library(english)
-library(lemon)
-library(gridExtra)
+# library(purrr)
+# library(knitr)
+# library(kableExtra)
+# library(stringi)
+
+# library(lemon)
+# library(gridExtra)
 
@@
 
-```{r deadwater-attr}
-# set some deadwater attributes
-ddw_rkms = 1.871
-ddw_m2 = 119431
-ddw_avg_width_m = 68
-ddw_acres = ddw_m2 / 4047
-ddw_hectares = ddw_m2 / 10000
-
-```
-
 # Introduction
@@
 
+```{r deadwater-attr}
+# set some deadwater attributes
+ddw_rkms = 1.871
+ddw_m2 = 119431
+ddw_avg_width_m = 68
+ddw_acres = ddw_m2 / 4047
+ddw_hectares = ddw_m2 / 10000
+
+```
+
 Most predation studies have focused on reservoirs [@Murphy2021] and mainstem reaches [@Zimmerman1999; @Tabor1993; @Ward1995; @Shively1996] of the Columbia River basin whereas considerably less is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems [@Rubenson2020].
@@
 Fall efforts were constrained within a two-week window to minimize predator emigration/immigration and to constrain the population for the mark-recapture models.
-
-<!-- In the end, sampling dates were as follows: fall 2019 (mark = November 12-14; recapture = November 19-21), fall 2020 (mark = October 20-23; recapture = October 27-30); spring 2021 (mark = May 18-21).
--->
 
@@
 
-# focus on pikeminnow, to be used for mark-recap models
+# capture/mark/recpature data frame; focus on pikeminnow, to be used for mark-recap models
 cmr_df = raw_df %>%
-  filter(count > 0) %>%
-  # pull out only Northern Pikeminnow
-  filter(species == "Northern Pikeminnow") %>%
+  filter(count > 0) %>%                                     # filter out effort-only records
+  filter(species == "Northern Pikeminnow") %>%              # focus on NPM
+  filter(method == "Angling") %>%
+  select(event_name, 
+         date,
+         method, 
+         species, 
+         count, 
+         length, 
+         fish_status, 
+         capture_type) %>%
+  # if fish_status = Live & capture_type = Mark, set as mark_release as TRUE
+  # if capture_type = Recapture, set recap to TRUE
   mutate(mark_release = if_else(fish_status == "Live" &     
@@
 ```{r effort, eval = T}
-effort_df = raw_df %>%
+# What size (mm) is considered a large pikeminnow?
+large_size = 380
+
+# catch per unit effort
+cpue_df = raw_df %>%
   select(event_name,
@@
                    sum),
-            .groups = "drop")
-
-# What size (mm) is considered a large pikeminnow?
-large_size = 228
-
-# catch per unit effort
-cpue_df = effort_df %>%
-  # add back in the number of pikeminnow caught
+            .groups = "drop") %>%
     left_join(cmr_df %>%
@@
                         .groups = "drop")) %>%
-  # calculate CPUE a few different ways
-  mutate(catch_per_min = n_fish / total_effort_min,
-         catch_per_hr  = n_fish / (total_effort_min / 60),
+  mutate(catch_per_hr = n_fish / (total_effort_min / 60),
          catch_per_angler_day = n_fish / crew_count,
@@
 LP_mod = cmr_df %>%
-  # filter out spring 2021 (no recapture)
   filter(event_name != "Spring_2021") %>%
@@
             .groups = "drop") %>%
-  mutate(catch_per_min = n_fish / total_effort_min,
-         catch_per_hr = n_fish / (total_effort_min / 60),
+  mutate(catch_per_hr = n_fish / (total_effort_min / 60),
          catch_per_angler_day = n_fish / angler_days,
@@
             by = "species") %>%
+  # calculate spring abundance estimate
   mutate(N_S = N_F * (cpue_S / cpue_F),
@@
 ```{r psd}
-s_size = 250
-q_size = 380
+s_size = 250        # stock size
+q_size = large_size # quality size
 
 psd_df = cmr_df %>%
-  select(species, count, length, recap, event_name) %>%
-  filter(recap  != TRUE)
+  select(event_name,
+         species,
+         count,
+         length,
+         recap) %>%
+  filter(recap != T)
+
+fq = psd_df %>%
+  filter(length >= q_size) %>%
+  nrow()
 
@@
   filter(length >= s_size) %>%
-  nrow()
-
-fq = psd_df %>%
-  filter(length >= q_size) %>%
   nrow()
@@
 gastric_df = raw_df %>%
-  select(event_name, date, species, method, lavage_id, length, stomach_contents, comments, stomach_contents_weight,                             fish_content_weight) %>%
-  mutate(bin_50 = lencat(length, w = 50)) %>%
-  filter(species == "Northern Pikeminnow") %>%
+  select(event_name, 
+         date, 
+         species,
+         method,
+         length,
+         lavage_id,
+         stomach_contents,
+         stomach_contents_weight,
+         fish_content_weight,
+         comments) %>%
+  #mutate(bin_50 = lencat(length, w = 50)) %>%
+  filter(species == "Northern Pikeminnow",
+         method == "Angling") %>%
   mutate(non_fish_wt = stomach_contents_weight - fish_content_weight) %>%
@@
     TRUE ~ FALSE
-  ))
+  )) %>%
+  mutate(across(event_name,
+                ~ str_replace(., "_", " ")))
 
@@
 stmch_df = gastric_df %>%
-  filter(stomach_contents_weight > 0) %>%
   select(event_name,
+         species,
          date,
@@
          non_fish_wt) %>%
+  filter(stomach_contents_weight > 0) %>%
   mutate(p_fish = fish_content_weight / stomach_contents_weight)
 
-psd_df = gastric_df %>%
-  group_by(bin_50) %>%
-  summarize(n = n())
+# psd_df = gastric_df %>%
+#   group_by(bin_50) %>%
+#   summarize(n = n())
 
@@
          water_temp_c = `47773_00010`,
-         water_temp_c_cd = `47773_00010_cd`)
+         water_temp_c_cd = `47773_00010_cd`) %>%
+  select(datetime, water_temp_c) %>%
+  mutate(month = month(datetime),
+         day = day(datetime)) %>%
+  group_by(month, 
+           day) %>%
+  summarize(mean_water_temp = mean(water_temp_c, na.rm = T))
 
@@
          f90_g = fall_f90$cum_cons_fish_g) %>%
-  mutate( across(contains('_g'),
-                 .fns = list(n = ~./avg_DSR_g),
-                 .names = "{col}_{fn}") )
-
-fall_g_summ = fall_results %>%
-  select(f10_g:f90_g) %>%
-  pivot_longer(f10_g:f90_g) %>%
-  group_by(name) %>%
-  summarize(cumm_g = max(value))
-
-rm(fall_f10, fall_f20, fall_f30, fall_f40, fall_f50, fall_f60, fall_f70, fall_f80, fall_f90)
+  mutate(season = "Fall") #%>%
+  # mutate( across(contains('_g'),
+  #                .fns = list(n = ~./avg_DSR_g),
+  #                .names = "{col}_{fn}") )
 
@@
          f90_g = spring_f90$cum_cons_fish_g) %>%
-  mutate( across(contains('_g'),
-                 .fns = list(n = ~./avg_NRR_g),
-                 .names = "{col}_{fn}") )
-
-spring_g_summ = spring_results %>%
-  select(f10_g:f90_g) %>%
+  mutate(season = "Spring") # %>%
+  # mutate( across(contains('_g'),
+  #                .fns = list(n = ~./avg_NRR_g),
+  #                .names = "{col}_{fn}") )
+
+bio_results = rbind(fall_results, spring_results)
+bio_g_summ = bio_results %>%
+  select(season, f10_g:f90_g) %>%
   pivot_longer(f10_g:f90_g) %>%
-  group_by(name) %>%
+  group_by(season, name) %>%
   summarize(cumm_g = max(value))
 
-rm(spring_f10, spring_f20, spring_f30, spring_f40, spring_f50, spring_f60, spring_f70, spring_f80, spring_f90)
+rm(fall_results,
+   fall_f10, 
+   fall_f20, 
+   fall_f30, 
+   fall_f40, 
+   fall_f50, 
+   fall_f60, 
+   fall_f70, 
+   fall_f80, 
+   fall_f90,
+   spring_results,
+   spring_f10, 
+   spring_f20, 
+   spring_f30, 
+   spring_f40, 
+   spring_f50, 
+   spring_f60, 
+   spring_f70, 
+   spring_f80, 
+   spring_f90)
 
@@
 # summarize potential Chinook eaten for each time period by one pikeminnow
-adult_impact = fall_results %>%
-  summarise(across(ends_with("g_n"), max)) %>%
-  add_column(season = "Fall",
-             .before = 0) %>%
-  bind_rows(spring_results %>%
-              summarise(across(ends_with("g_n"), max)) %>%
-              add_column(season = "Spring",
-                         .before = 0)) %>%
-  pivot_longer(cols = ends_with("g_n"),
+impact_df = bio_results %>%
+  group_by(season) %>%
+  summarize(across(ends_with("_g"), max)) %>%
+  pivot_longer(cols = ends_with("_g"),
                names_to = "perc_diet",
-               values_to = "juv_chnk") %>%
+               values_to = "g_fish_cons") %>%
   mutate(across(perc_diet,
@@
                 as.integer)) %>%
+  crossing(p_chnk = c(.01, .10, .30, .50)) %>%
+  mutate(g_chnk = g_fish_cons * p_chnk) %>%
+  mutate(n_chnk = ifelse(season == "Fall",
+                         g_chnk / avg_DSR_g,
+                         g_chnk / avg_NRR_g)) %>%
   left_join(spring_abund %>%
@@
                         values_from = value)) %>%
-  mutate(smolt_N = N * juv_chnk,
-         smolt_Lci = Lci * juv_chnk,
-         smolt_Uci = Uci * juv_chnk) %>%
+  mutate(juv_N = N * n_chnk,
+         juv_Lci = Lci * n_chnk,
+         juv_Uci = Uci * n_chnk) %>%
   bind_cols(sar) %>%
-  mutate(adult_N = smolt_N * medsar,
-         adult_Lci = smolt_Lci * qnorm(0.025, medsar, sdsar),
-         adult_Uci = smolt_Uci * qnorm(0.975, medsar, sdsar))
+  mutate(adult_N = juv_N * medsar,
+         adult_Lci = juv_Lci * qnorm(0.025, medsar, sdsar),
+         adult_Uci = juv_Uci * qnorm(0.975, medsar, sdsar))
 
 # sum fall and spring impacts
-total_impacts = adult_impact %>%
-  group_by(perc_diet) %>%
+total_impacts = impact_df %>%
+  group_by(perc_diet, p_chnk) %>%
   summarise(across(c(ends_with("_N"),
@@
                      ends_with("_Uci")),
-                   sum))
+                   sum),
+            .groups = "drop")
 

### 2022-01-07 analysis/paper/paper.Rmd
cleaned up some paper code

@@
 library(tidytable)
-# library(purrr)
-# library(knitr)
-# library(kableExtra)
-# library(stringi)
-
-# library(lemon)
-# library(gridExtra)
+library(kableExtra)
+library(lemon)
+library(gridExtra)
 
@@
 
-# psd_df = gastric_df %>%
-#   group_by(bin_50) %>%
-#   summarize(n = n())
-
 ```
@@
                 as.integer)) %>%
-  crossing(p_chnk = c(.01, .10, .30, .50)) %>%
+  crossing(p_chnk = c(.01, .10, .30, .50, .70, .90, 1)) %>%
   mutate(g_chnk = g_fish_cons * p_chnk) %>%
@@
 
-We estimated an individual, average-size (`r avg_fall_NPM` mm) Northern Pikeminnow to consume `r median(fall_g_summ$cumm_g)` g (range: `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the DSR emigration to maintain their body size, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
-Similarly, we estimated an average-size (`r avg_spring_NPM` mm) Northern Pikeminnow to consume `r median(spring_g_summ$cumm_g)` g (range: `r min(spring_g_summ$cumm_g)` - `r max(spring_g_summ$cumm_g)` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
-Given a mean Northern Pikeminnow population size of `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` from fall estimates and `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` for spring 2021 and using the median value from diet scenarios (i.e., 50% of their diet consists of fish), we'd estimate that the Northern Pikeminnow population can consume `r prettyNum(median(fall_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` g (range: `r prettyNum(min(fall_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` - `r prettyNum(max(fall_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` g) fish during the DSR emigration and an additional `r prettyNum(round(median(spring_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` g (range: `r prettyNum(round(min(spring_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` - `r prettyNum(round(max(spring_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` g) during the NRR emigration in Deadwater Slough.
-In total, the Northern Pikeminnow population can consume `r prettyNum(median(fall_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Fall"]) + median(spring_g_summ$cumm_g) * unique(adult_impact$N[adult_impact$season == "Spring"]), big.mark = ",")` g of fish during the DSR and NRR emigrations to maintain their population size if we assume 50% of their diet consists of fish.
+We estimated an individual, average-size (`r avg_fall_NPM` mm) Northern Pikeminnow to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g) of fish during the DSR emigration to maintain their body size, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
+Similarly, we estimated an average-size (`r avg_spring_NPM` mm) Northern Pikeminnow to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
+Given a mean Northern Pikeminnow population size of `r prettyNum(unique(impact_df$N[impact_df$season == "Fall"]), big.mark = ",")` from fall estimates and `r prettyNum(unique(impact_df$N[impact_df$season == "Spring"]), big.mark = ",")` for spring 2021 and using the median value from diet scenarios (i.e., 50% of their diet consists of fish), we'd estimate that the Northern Pikeminnow population can consume `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 50]), big.mark = ",")` g (range: `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 10]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 10]), big.mark = ",")` - `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 90]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 90]), big.mark = ",")` g) fish during the DSR emigration and an additional `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 50]), big.mark = ",")` g (range: `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 10]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 10]), big.mark = ",")` - `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 90]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 90]), big.mark = ",")` g) during the NRR emigration in Deadwater Slough.
+In total, the Northern Pikeminnow population can consume `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 50]) + unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 50]), big.mark = ",")` g of fish during the DSR and NRR emigrations to maintain their population size if we assume 50% of their diet consists of fish.
 
@@
 <!--
-Given an average weight of `r avg_DSR_g` g for DSR emigrants, that would equate to `r round(median(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` (range: `r round(min(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)`) fish consumed by an individual during fall and an additional `r round(median(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` (range: `r round(min(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)`) fish consumed using an average weight of `r avg_NRR_g` g for NRR emigrants.
-
-Given a mean Northern Pikeminnow population size of `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` from fall estimates and an estimate of `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` for spring 2021 and using the median value from diet scenarios (i.e, 50% of their diet consists of fish), we'd estimate that Northern Pikeminnow can consume `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")`) fish during the DSR emigration and `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")`) fish during the NRR emigration in Deadwater Slough.
-Table \@ref(tab:adult-imp) shows the estimated number of fish consumed by an individual Northern Pikeminnow and by the population across all diet scenarios during both the fall (DSR) and spring (NRR) peak emigration seasons.
-
-Across both seasons and all diet scenarios, we estimated the total number of fish consumed to be between `r prettyNum(round(min(total_impacts$smolt_N), 0), big.mark = ",")` and `r prettyNum(round(max(total_impacts$smolt_N), 0), big.mark = ",")` (Table \@ref(tab:adult-imp)).
-The estimated number of fish consumed by a Northern Pikeminnow begins to taper off between the 30-90% diet scenarios.
-So although we are uncertain of the diet proportions consisting of fish versus invertebrates, the impact of that uncertainty is somewhat constrained.
--->
-
-<!--
 ## Juvenile Chinook Salmon Abundance
@@
 
-The mean population size of Northern Pikeminnow in the Deadwater Slough during the fall season was estimated to be `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` and during the spring 2021 to be `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")`.
-Those estimates translate to linear densities of `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms), big.mark = ",")` which is `r round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms / john_day_dens)` and `r round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms / john_day_dens)` times greater, respectively, than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r as.character(english(round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_rkms / low_columbia_fish_km)))` and `r as.character(english(round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_rkms / low_columbia_fish_km)))` times greater than projections for the lower Columbia River [@Beamesderfer1996].
+The mean population size of Northern Pikeminnow in the Deadwater Slough during the fall season was estimated to be `r prettyNum(unique(impact_df$N[impact_df$season == "Fall"]), big.mark = ",")` and during the spring 2021 to be `r prettyNum(unique(impact_df$N[impact_df$season == "Spring"]), big.mark = ",")`.
+Those estimates translate to linear densities of `r prettyNum(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms), big.mark = ",")` which is `r round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / john_day_dens)` and `r round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / john_day_dens)` times greater, respectively, than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r as.character(english(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / low_columbia_fish_km)))` and `r as.character(english(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / low_columbia_fish_km)))` times greater than projections for the lower Columbia River [@Beamesderfer1996].
 Although estimated densities for Deadwater Slough are high given the relative size (volume) of the Salmon River at the slough relative to the Columbia River, high CPUE combined with low recapture rates (one in every `r round(nrow(cmr_df) / sum(cmr_df$capture_type == "Recapture", na.rm = T))` fish were recaptures) suggest this area may support exceptionally high predator densities.
@@
 Our estimates of Northern Pikeminnow areal density are also higher than those reported elsewhere in the Columbia River basin.
-We estimate the density of Northern Pikeminnow in Deadwater Slough, as fish per hectare, to be `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Fall"]) / ddw_hectares), big.mark = ",")` for fall and `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]) / ddw_hectares), big.mark = ",")` for spring.
+We estimate the density of Northern Pikeminnow in Deadwater Slough, as fish per hectare, to be `r prettyNum(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_hectares), big.mark = ",")` for fall and `r prettyNum(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_hectares), big.mark = ",")` for spring.
 For comparison, the only other available estimates of areal density, that the authors are aware of, are 4.4 fish per hectare for John Day Reservoir [@Beamesderfer1991] and 15 fish per hectare in Lake Washington [@Bartoo1977].
@@
 Our intent, rather, was to simply document predation on juvenile Chinook Salmon in Deadwater Slough, and given the potential ineffectiveness of gastric lavage using angling and for Northern Pikeminnow, we consider our diet composition estimates to be conservative estimates i.e., we believe that a larger proportion of the Northern Pikeminnow diet is fish and juvenile Chinook Salmon than our gastric results suggest.
-<!--
-Although we only identified fish or fish parts within `r round(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"] * 100, 1)`% of Northern Pikeminnow examined and that fish remnants were only `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall stomach contents, we believe those numbers to be misleading...
--->
 
@@
 
-An average-size Northern Pikeminnow from Deadwater Slough is estimated to consume `r median(fall_g_summ$cumm_g)` g (range `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) and `r median(spring_g_summ$cumm_g)` g (range `r min(spring_g_summ$cumm_g)` - `r max(spring_g_summ$cumm_g)` g) of fish during spring and fall, respectively, to maintain its body weight.
-This equates to `r median(fall_g_summ$cumm_g) + median(spring_g_summ$cumm_g)` g of fish consumed during the DSR and NRR peak emigrations, assuming the median value among diet scenarios (i.e., if 50% of their diet consists of fish).
-Consumption ranged from `r min(fall_g_summ$cumm_g) + min(spring_g_summ$cumm_g)` for a diet comprised of 10% fish to `r max(fall_g_summ$cumm_g) + max(spring_g_summ$cumm_g)` for a diet comprised of 90% fish.
+An average-size Northern Pikeminnow from Deadwater Slough is estimated to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g) and `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g) of fish during fall and spring, respectively, to maintain its body weight.
+This equates to `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"]) + median(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g of fish consumed during the DSR and NRR peak emigrations, assuming the median value among diet scenarios (i.e., if 50% of their diet consists of fish).
+Consumption ranged from `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"]) + min(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` for a diet comprised of 10% fish to `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"]) + max(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` for a diet comprised of 90% fish.
 
@@
 
-```{r adult-imp}
-adult_impact %>%
+```{r adult-imp, eval = F}
+impact_df %>%
   select(season,
@@
 
-# adult_impact %>%
-#   mutate(stage = "Smolt") %>%
-#   select(#event_name, 
-#          #model,
-#          stage,
-#          starts_with("smolt")) %>%
-#   rename_with(.fn = ~ str_remove(., "smolt_"),
-#               .cols = starts_with("smolt")) %>%
-#   bind_rows(adult_impact %>%
-#               mutate(stage = "Adult Equivalent") %>%
-#               select(#event_name, model,
-#                      stage,
-#                      starts_with("adult")) %>%
-#               rename_with(.fn = ~ str_remove(., "adult_"),
-#                           .cols = starts_with("adult"))) %>%
-#   kable(booktabs = T,
-#         format.args = list(big.mark = ","),
-#         col.names = c("Stage",
-#                       "N",
-#                       "Lower 95% CI",
-#                       "Upper 95% CI"),
-#         digits = 0,
-#         caption = "Estimates of smolts consumed during fall migration by Northern Pikeminnow, and the adult equivalent of those smolts based on SARs.") %>%
-#   kable_styling()
-
 ```
@@
 "}
-spring_results$season = "Spring"
-fall_results$season = "Fall"
-bio_results = rbind(spring_results, fall_results)
+# spring_results$season = "Spring"
+# fall_results$season = "Fall"
+# bio_results = rbind(spring_results, fall_results)
 bio_results$ymd = (ymd(bio_results$date))
@@
 2019."}
-adult_impact %>%
+impact_df %>%
   mutate(stage = "Juv. Emigrant") %>%
@@
               .cols = starts_with("smolt")) %>%
-  bind_rows(adult_impact %>%
+  bind_rows(impact_df %>%
               mutate(stage = "Adult Equivalent") %>%

### 2022-01-07 README.Rmd
updated README

@@
 # Please put your title here to include it in the file below.
-Title <- "Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations in the Upper Salmon River, Idaho"
+Title <- "Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations"
 Authors <- "Ackerman, Michael W., Nicholas J. Porter, Kevin E. See, Michael S. Hall, Tara E. Blackman, and Richard A. Carmichael, and Michael W. Ackerman"
@@
 
+<!--
 Our pre-print is online here:
@@
 > `r Authors` (`r yr`). _`r Title`_. `r Journal`, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+-->
 

### 2022-01-07 analysis/paper/paper.Rmd
updates to impacts plot

@@
 Note that this thought exercise only includes the Northern Pikeminnow population in Deadwater Slough even though they are present throughout the upper Salmon River, although likely at lower densities than in Deadwater Slough, and also ignores potential impacts to ESA-listed steelhead and Sockeye Salmon that are also a portion of the Northern Pikeminnow diet.
-
 <!--
@@
          'Fish Contents' = fish_contents) %>%
-  mutate(Effort = recode(Effort,
-                         `Fall_2019` = "Fall 2019",
-                         `Fall_2020` = "Fall 2020",
-                         `Spring_2021` = "Spring 2021")) %>%
   kable(booktabs = T,
@@
 2019."}
-impact_df %>%
-  mutate(stage = "Juv. Emigrant") %>%
-  select(stage,
-         season,
+impact_p = impact_df %>%
+  select(season,
          perc_diet,
-         starts_with("smolt")) %>%
-  rename_with(.fn = ~ str_remove(., "smolt_"),
-              .cols = starts_with("smolt")) %>%
+         p_chnk,
+         N = juv_N,
+         Lci = juv_Lci,
+         Uci = juv_Uci) %>%
+  group_by(perc_diet, p_chnk) %>%
+  summarise(across(c("N", "Lci", "Uci"),
+                   sum),
+            .groups = "drop") %>%
+  mutate(Stage = "Juveniles") %>%
   bind_rows(impact_df %>%
-              mutate(stage = "Adult Equivalent") %>%
-              select(stage,
-                     season,
+              select(season,
                      perc_diet,
-                     starts_with("adult")) %>%
-              rename_with(.fn = ~ str_remove(., "adult_"),
-                          .cols = starts_with("adult"))) %>%
+                     p_chnk,
+                     N = adult_N,
+                     Lci = adult_Lci,
+                     Uci = adult_Uci) %>%
+              group_by(perc_diet, p_chnk) %>%
+              summarise(across(c("N", "Lci", "Uci"),
+                   sum),
+             .groups = "drop") %>%
+             mutate(Stage = "Adult Equivalents")) %>%
+  mutate(p_chnk = p_chnk * 100) %>%
+  rename(perc_chnk = p_chnk) %>%
+  filter(perc_chnk %in% c(10, 30, 50)) %>%
   ggplot(aes(x = perc_diet,
@@
                 width = 0.1) +
-  geom_point(size = 4) +
-  facet_grid(fct_rev(stage) ~ season,
+  geom_point(size = 3) +
+  facet_grid(vars(Stage),
+             vars(perc_chnk),
              scales = "free_y") +
   scale_color_viridis_d(name = "Percent\nFish in Diet") +
+  scale_y_continuous(labels = scales::comma) +
   theme(axis.text.x = element_blank(),
@@
 
+impact_p
+
+# impact_df %>%
+#   mutate(stage = "Juv. Emigrant") %>%
+#   select(stage,
+#          season,
+#          perc_diet,
+#          starts_with("smolt")) %>%
+#   rename_with(.fn = ~ str_remove(., "smolt_"),
+#               .cols = starts_with("smolt")) %>%
+#   bind_rows(impact_df %>%
+#               mutate(stage = "Adult Equivalent") %>%
+#               select(stage,
+#                      season,
+#                      perc_diet,
+#                      starts_with("adult")) %>%
+#               rename_with(.fn = ~ str_remove(., "adult_"),
+#                           .cols = starts_with("adult"))) %>%
+# ggplot(aes(x = perc_diet,
+#            y = N,
+#            color = as.factor(perc_diet))) +
+# geom_errorbar(aes(ymin = Lci,
+#                   ymax = Uci),
+#               width = 0.1) +
+# geom_point(size = 4) +
+# facet_grid(fct_rev(stage) ~ season,
+#            scales = "free_y") +
+# scale_color_viridis_d(name = "Percent\nFish in Diet") +
+# theme(axis.text.x = element_blank(),
+#       legend.position = "bottom",
+#       axis.title.x = element_blank(),
+#       axis.ticks.x = element_blank(),
+#       axis.text.y = element_text(size = 11,
+#                                  color = "black"),
+#       strip.text.x = element_text(size = 11,
+#                                   color = "black"),
+#       strip.text.y = element_text(size = 11,
+#                                   color = "black")) +
+# labs(y = "Number Chinook Consumed")
+
 ```

### 2022-01-08 analysis/paper/paper.Rmd
made a few writing edits

@@
 Fall efforts included sampling across a two-week span, but due to budget constraints the spring 2021 effort was limited to a single week.
-Fall efforts were constrained within a two-week window to minimize predator emigration/immigration and to constrain the population for the mark-recapture models.
+Fall efforts were constrained within a two-week window to minimize predator emigration/immigration and to constrain the population to meet assumptions of the mark-recapture models.
 
@@
   mutate(model = fct_recode(model,
-                            "Schumacher-Eschmeyer" = "SchumacherEschmeyer")) %>%
+                            "Schumacher-Eschmeyer" = "SchumacherEschmeyer"),
+         model = fct_relevel(model,
+                             "Petersen",
+                             after = 0)) %>%
   arrange(event_name, model)
@@
 
-where $M$, $n$, and $m$ are indexed by each sample, $i$.
-Here, sample $i$ is defined as each day.
-The Schnabel estimator does not have an associated standard error; however, the 95% confidence intervals were calculated.
+where $M$, $n$, and $m$ are indexed by each sample (day), $i$.
+With small numbers of recaptures, it may not be appropriate to calculate the standard error of the Schnabel estimator but the 95% confidence intervals can be calculated using a Poisson approximation [@Krebs1999].
 The Schnabel estimator is essentially a weighted average of a series of Lincoln-Peterson estimators (with a Champman modification).
@@
 Next, our estimates assumed that the second (and subsequent) sample(s) are random samples of the population.
-Angling methods may have a size selection bias, thereby limiting our abundance estimates to a size threshold susceptible to angling.
+Angling methods may have a size selection bias, thereby limiting our abundance estimates to a size threshold susceptible to angling, and possibly biasing them low.
 We also consider the validity of our assumption of equal catchability of individual fish between sampling events to have introduced potential bias in our results.

### 2022-01-10 analysis/supplementary-materials/scripts/abund_est.R
updated script to examine the impacts if not all marked fish were available for recapture.

@@ -333,3 +333,74 @@ cpue_df %>%
        y = "CPUE (Pikeminnow caught per hour)") +
   theme(legend.position = "none")
 
+
+#-----------------------------------------------------------------
+# what if marked fish were not as susceptible to recapture?
+
+N_mods2 = N_mods %>%
+  crossing(perc_unavail = c(0, 0.1, 0.3, 0.5, 0.7, 0.9)) %>%
+  mutate(data = map2(data,
+                     perc_unavail,
+                    .f = function(x, y) {
+                      x %>%
+                        mutate(across(R,
+                                      ~ . * (1 - y))) %>%
+                        mutate(M = lag(R - m),
+                               M = replace_na(M, 0),
+                               M = cumsum(M))
+                    })) %>%
+  mutate(mr_model = map2(data,
+                         model,
+                         .f = function(x, y) {
+                           with(x,
+                                mrClosed(n = n,
+                                         m = m,
+                                         R = R,
+                                         method = y,
+                                         chapman.mod = TRUE))
+                         }),
+         N = map_dbl(mr_model,
+                     .f = summary),
+         CI = map(mr_model,
+                  .f = confint),
+         Lci = map_dbl(CI,
+                       .f = function(x) x[1]),
+         Uci = map_dbl(CI,
+                       .f = function(x) x[2])) %>%
+  ungroup() %>%
+  select(-CI) %>%
+  group_by(event_name,
+           model) %>%
+  mutate(perc_N = N / N[perc_unavail == 0],
+         perc_avail = 1 - perc_unavail)
+
+all_equal(N_mods %>%
+            select(event_name, model,
+                   N:Uci),
+          N_mods2 %>%
+            filter(perc_unavail == 0) %>%
+            select(event_name, model,
+                   N:Uci))
+
+N_mods2 %>%
+  ggplot(aes(x = perc_unavail,
+             y = N,
+             color = model)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~ event_name,
+             scales = "free_y") +
+  theme(legend.position = "bottom")
+
+N_mods2 %>%
+  ggplot(aes(x = perc_avail,
+             y = perc_N,
+             color = model)) +
+  geom_point(aes(shape = event_name)) +
+  geom_abline(linetype = 2) +
+  geom_smooth()
+
+N_mods2 %>%
+  filter(model == "Schnabel")
+
+# I don't know how to do something similar for single census models

### 2022-01-10 analysis/paper/paper.Rmd
made a few editorial changes, and re-knit to html

@@
   filter(model == "Schnabel") %>%
+  # filter(model == "Chapman") %>%
   select(-range, -prop_range) %>%
@@
 all_est %>%
-  select(-c(species, range:prop_range)) %>%
+  mutate(across(c(N, SE, Lci, Uci),
+                ~ prettyNum(round(.), big.mark = ","))) %>%
+  rowwise() %>%
+  mutate(CI = paste(Lci, Uci,
+                    sep = " - ")) %>%
+  select(-c(species, range:prop_range),
+         -Lci, -Uci) %>%
   rename(`Effort` = event_name,
          Estimator = model,
-         `L CI` = Lci,
-         `U CI` = Uci) %>%
+         `95% Confidence Interval` = CI) %>%
+         # `L CI` = Lci,
+         # `U CI` = Uci) %>%
   kable(booktabs = T,

### 2022-01-10 analysis/paper/paper.Rmd
Added a few sentences about the degree of bias that could be introduced by difference in catchability between marked and unmarked fish

@@
 Unequal catchability for individual fish is difficult to assess, especially if being caught once influences the likelihood of recapture.
-During our study, we recaptured two individual fish during multiple surveys indicating that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.--> We further recognize that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
+During our study, we recaptured two individual fish during multiple surveys indicating that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.-->
+
+We further recognize that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
 Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit the "good" fishing locales.
@@
 Violations of the assumption of equal catchability between marked and unmarked fish may lead to an overestimate of abundance of Northern Pikeminnow in our study if marked fish have lower catchability.
+This bias will be proportional to the difference in catchability of marked and unmarked fish, so if marked fish are only 70% as likely to be caught as unmarked fish (due to behavior or spatial distribution) the true abundance is closer to 70% of our abundance estimate.
+We calculated this for the Schnabel estimator by assuming that only 70% of the fish marked each sample were returned to the populuation and available for recapture.
 

### 2022-01-10 analysis/paper/paper.Rmd
one editorial comment

@@
 
-The mean population size of Northern Pikeminnow in the Deadwater Slough during the fall season was estimated to be `r prettyNum(unique(impact_df$N[impact_df$season == "Fall"]), big.mark = ",")` and during the spring 2021 to be `r prettyNum(unique(impact_df$N[impact_df$season == "Spring"]), big.mark = ",")`.
-Those estimates translate to linear densities of `r prettyNum(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms), big.mark = ",")` which is `r round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / john_day_dens)` and `r round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / john_day_dens)` times greater, respectively, than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r as.character(english(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / low_columbia_fish_km)))` and `r as.character(english(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / low_columbia_fish_km)))` times greater than projections for the lower Columbia River [@Beamesderfer1996].
-Although estimated densities for Deadwater Slough are high given the relative size (volume) of the Salmon River at the slough relative to the Columbia River, high CPUE combined with low recapture rates (one in every `r round(nrow(cmr_df) / sum(cmr_df$capture_type == "Recapture", na.rm = T))` fish were recaptures) suggest this area may support exceptionally high predator densities.
+The mean population size of Northern Pikeminnow in the Deadwater Slough during the fall season was estimated to be `r prettyNum(unique(round(impact_df$N[impact_df$season == "Fall"])), big.mark = ",")` and during the spring 2021 to be `r prettyNum(unique(round(impact_df$N[impact_df$season == "Spring"])), big.mark = ",")`.
+Those estimates translate to linear densities of `r prettyNum(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms), big.mark = ",")`
+<!-- need units  -->
+which is `r round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / john_day_dens)` and `r round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / john_day_dens)` times greater, respectively, than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r as.character(english(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / low_columbia_fish_km)))` and `r as.character(english(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / low_columbia_fish_km)))` times greater than projections for the lower Columbia River [@Beamesderfer1996].
+Although estimated densities for Deadwater Slough are high given the relative size (volume) of the Salmon River at the slough relative to the Columbia River, high CPUE combined with low recapture rates (one in every `r round(nrow(cmr_df) / sum(cmr_df$capture_type == "Recapture", na.rm = T))` captured fish were recaptures) suggest this area may support exceptionally high predator densities.
 Further, @Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective.

### 2022-01-10 analysis/paper/paper.Rmd
added a few comments and made a few editorial revisions.

@@
 Therefore, we believe our spring abundance estimate to be conservative.
+<!-- how do higher spring flows reduce pikeminnow densities?
+If pikeminnow densities are lower, then the total abundance SHOULD be lower.
+Makes more sense to me to say that higher flows directly lower capture probabilities.
+-->
 
@@
 <!-- Again, some of the following numbers need to be updated after results are formalized -->
-Recovery of the Upper Salmon MPG is desired to support local fisheries and economies.
+Recovery of Chinook Salmon in the Upper Salmon MPG is desired to support local fisheries and economies.
 At least five of the eight populations must meet criteria set forth by @McElhany2000 and @ICTRT2007 for the MPG to be considered viable and for recovery of the Snake River Evolutionary Significant Unit.
@@
 These estimates only take into account predation on juvenile Chinook Salmon by Northern Pikeminnow, but presumably, juvenile steelhead and Sockeye Salmon are also prey items in Deadwater Slough, including hatchery releases.
-Therefore, adult returns of steelhead and Sockeye Salmon are affected as well, including for hathchery populations which provide for recreational fishing opportunities.
+Therefore, adult returns of steelhead and Sockeye Salmon are affected as well, including for hatchery populations which provide recreational fishing opportunities.
 Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including multiple ESA-listed species.
-Because the Deadwater Slough is part of the migratory pathway for emigrating salmonids, including for multiple species and populations, the impact of that benefit could be higher than tributary rehabilitation actions, which typically benefit a single population.
+Because the Deadwater Slough is part of the migratory pathway for multiple species and populations of emigrating salmonids, the impact of that benefit could be higher than tributary rehabilitation actions which typically benefit a single population.
 Moreover, the deepened, slack water conditions that favor Northern Pikeminnow at Deadwater Slough are a result of the failure of a manmade mining reservoir dam.
@@
 
-Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere (citation).
+Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere (*citation*).
 A local bounty program in the Salmon River has the potential to boost the local economy (to an unknown degree) which has been negatively impacted by low returns of Chinook Salmon, steelhead, and Sockeye Salmon in recent decades.
 Many local businesses rely on license and gear sales for local sport fisheries, but low salmon and steelhead returns have eliminated or reduced the duration of fisheries in some years, and a local Northern Pikeminnow bounty program could allow the local economy to recoup some of those lost revenues.
-On the downside, bounty programs require continual effort and harvest to suppress the population size and so would require year of continued funding and local buy-in to remain effective.
+On the downside, bounty programs require continual effort and harvest to suppress the population size and so would require years of continued funding and local buy-in to remain effective.
 If funding or effort was alleviated, Northern Pikeminnow would likely recover to a higher population size without additional alterations to the habitat.
@@
 Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids emigrating through Deadwater Slough including Chinook Salmon i.e., juvenile Chinook Salmon have slower movement rates and little refugia to avoid predation.
-We estimate that predation from Northern Pikeminnow in Deadwater Slough may reduce adult returns by as much as 146 or 26% or recent returns to the Upper Salmon River MPG.
+We estimate that predation from Northern Pikeminnow in Deadwater Slough may reduce adult returns by as much as 146 or 26% of recent returns to the Upper Salmon River MPG.
 Therefore, predation in Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon MPG.

### 2022-01-13 analysis/paper/paper.Rmd
made a mess

@@
 # Introduction
+
+Here's a mess!
 

### 2022-01-14 analysis/supplementary-materials/scripts/abund_est.R
one more quick analysis of relationship between percent of marks available for recapture and estimate of abundance.

@@ -383,6 +383,16 @@ all_equal(N_mods %>%
                    N:Uci))
 
 N_mods2 %>%
+  group_by(event_name,
+           model) %>%
+  mutate(org_N = N[perc_avail == 1]) %>%
+  ungroup() %>%
+  lm(N / org_N ~ perc_avail*model - 1,
+     data = .) %>%
+  summary() %>%
+  broom::tidy()
+
+N_mods2 %>%
   ggplot(aes(x = perc_unavail,
              y = N,
              color = model)) +

### 2022-01-14 analysis/paper/paper.Rmd
cleaned up the mess
Modified plot

@@
 
-Here's a mess!
-
 The Snake and Columbia rivers drain a network of inland tributaries that provide essential spawning and rearing habitat for anadromous Pacific salmon *Oncorhynchus* spp.
@@
 Note the y-axis is truncated at 100 and the upper confidence interval for the Schumacker-Eschmeyer estimator in 2020 extends beyond that."}
+
+pos = position_dodge(width=0.3)
 all_est %>%
+
   mutate(across(c(N, Lci, Uci),
@@
                             grepl("Schumacher-Eschmeyer", model) ~ "multiple")) %>%
+
   ggplot(aes(x = model,
@@
              )) +
-  geom_point(aes(size = 4,
-                 shape = census)) +
-  scale_shape_manual(values = c(17, 19)) +
-  geom_errorbar(aes(ymin = Lci,
-                    ymax = Uci),
-                width = 0.1) +
-  facet_wrap(~ event_name) +
+  geom_errorbar(aes(ymin = Lci,ymax = Uci, group=event_name), position = pos, width=0) + 
+  geom_point(aes(shape = event_name, color = census), position = pos, size = 4) +
+    #facet_wrap(~ event_name) +
   coord_cartesian(ylim = c(0, 100)) +
+  scale_shape_manual(values = c(21, 19), name="Event") +
+  scale_color_manual(values = c("grey35", "darkred"), name="Census type") +
   labs(x = "Estimator",
        y = "Abundance (x 1,000)") +
-  #theme_bw() +
-  theme(legend.position = "none",
-        axis.title.x = element_blank(),
-        axis.text.x = element_text(angle = 45,
-                                   hjust = 1,
-                                   size = 11,
-                                   color = "black"),
-        axis.text.y = element_text(size = 11,
-                                   color = "black"),
-        strip.text.x = element_text(size = 11,
-                                    color = "black")) +
-  scale_color_brewer(palette = "Dark2")
+  #scale_size(guide = "none")+
+  theme(legend.position="top",legend.key = element_blank(),
+        legend.background = element_blank(),
+        legend.key.size = unit(0.50, "cm"),
+        #axis.title.x = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1, size = 11, color = "black"),
+        axis.text.y = element_text(size = 11, color = "black"),
+        strip.text.x = element_text(size = 11,color = "black"))
+  #scale_color_brewer(palette = "Dark2")
 

### 2022-01-14 analysis/paper/paper.Rmd
added code to examine if marked fish weren't available for recapture until 2 days after capture.

@@
 
+```{r assump-violation}
+# assume that marked fish aren't available for recapture until 2 days after capture
+# using Schnabel estimator
+N_mods_viol = N_mods %>%
+  filter(model == "Schnabel") %>%
+  select(event_name:model) %>%
+  mutate(data = map(data,
+                    .f = function(x) {
+                      x %>%
+                        mutate(R_lag = lag(R),
+                               R_lag = replace_na(R_lag,
+                                                  0))
+                    })) %>%
+  mutate(mr_model = map2(data,
+                         model,
+                         .f = function(x, y) {
+                           with(x,
+                                mrClosed(n = n,
+                                         m = m,
+                                         R = R_lag,
+                                         method = y,
+                                         chapman.mod = TRUE))
+                         }),
+         N = map_dbl(mr_model,
+                     .f = summary),
+         CI = map(mr_model,
+                  .f = confint),
+         Lci = map_dbl(CI,
+                       .f = function(x) x[1]),
+         Uci = map_dbl(CI,
+                       .f = function(x) x[2])) %>%
+  ungroup() %>%
+  select(-CI) %>%
+  mutate(model = "Schnabel - Delayed Mixing")
+```
+
 ```{r abund-est}
@@
   bind_rows(N_mods %>%
+             select(event_name, species, model, N:Uci)) %>%
+  bind_rows(N_mods_viol %>%
              select(event_name, species, model, N:Uci)) %>%
@@
                              after = 0)) %>%
-  arrange(event_name, model)
+  arrange(event_name, model) %>%
+  filter(! model %in% c("Petersen",
+                        "Schumacher-Eschmeyer"))
 

### 2022-01-14 analysis/supplementary-materials/scripts/abund_est.R

@@ -414,3 +414,42 @@ N_mods2 %>%
   filter(model == "Schnabel")
 
 # I don't know how to do something similar for single census models
+
+
+
+# assume that marked fish aren't available for recapture until 2 days after capture
+# using Schnabel estimator
+
+N_mods_viol <- N_mods %>%
+  filter(model == "Schnabel") %>%
+  select(event_name:model) %>%
+  mutate(data = map(data,
+                    .f = function(x) {
+                      x %>%
+                        mutate(R_lag = lag(R),
+                               R_lag = replace_na(R_lag,
+                                                  0))
+                    })) %>%
+  mutate(mr_model = map2(data,
+                         model,
+                         .f = function(x, y) {
+                           with(x,
+                                mrClosed(n = n,
+                                         m = m,
+                                         R = R_lag,
+                                         method = y,
+                                         chapman.mod = TRUE))
+                         }),
+         N = map_dbl(mr_model,
+                     .f = summary),
+         CI = map(mr_model,
+                  .f = confint),
+         Lci = map_dbl(CI,
+                       .f = function(x) x[1]),
+         Uci = map_dbl(CI,
+                       .f = function(x) x[2])) %>%
+  ungroup() %>%
+  select(-CI)
+
+N_mods %>%
+  filter(model == 'Schnabel')

### 2022-01-18 analysis/paper/paper.Rmd
Figure edits

@@
 pos = position_dodge(width=0.3)
+pm_est_labels <- c("Chapman","Schnabel","Schnabel \n Delayed-mixing")
 all_est %>%
@@
   geom_errorbar(aes(ymin = Lci,ymax = Uci, group=event_name), position = pos, width=0) + 
-  geom_point(aes(shape = event_name, color = census), position = pos, size = 4) +
+  geom_point(aes(shape = event_name), position = pos, size = 4) +
     #facet_wrap(~ event_name) +
@@
   scale_shape_manual(values = c(21, 19), name="Event") +
-  scale_color_manual(values = c("grey35", "darkred"), name="Census type") +
+  #scale_color_manual(values = c("grey35", "darkred"), name="Census type") +
+  scale_x_discrete(labels=c(pm_est_labels))+
   labs(x = "Estimator",
@@
   #scale_size(guide = "none")+
-  theme(legend.position="top",legend.key = element_blank(),
-        legend.background = element_blank(),
-        legend.key.size = unit(0.50, "cm"),
+  theme(legend.position = c(.15, .85),legend.background = element_blank(),
+        legend.key.size = unit(.5, "cm"),legend.box = 'horizontal',
         #axis.title.x = element_blank(),
-        axis.text.x = element_text(angle = 45, hjust = 1, size = 11, color = "black"),
+        axis.text.x = element_text(size = 11, color = "black"),
         axis.text.y = element_text(size = 11, color = "black"),
@@
   labs(x = "Total Length (mm)",
-       y = "Frequency") +
+       y = "Number of Northern Pikeminnow") +
   xlim(c(100, 700)) +
@@
         axis.text.y = element_text(size = 11,
-                                   color = "black",
-                                   vjust = 3)) +
-  theme_classic()
+                                   color = "black"))
 
@@
   ggplot(aes(ymd, f50_g)) +
-  geom_line(size = 1.15) +
+  geom_line(size = 1.1) +
   geom_ribbon(data = bio_results,
-              aes(ymd,
-                  ymin = f10_g,
+              aes(ymd,ymin = f10_g,
                   ymax = f90_g),
-              color = "grey65",
-              alpha = 0.4) +
-  theme_bw() +
+              color = "grey55",
+              alpha = 0.4) +  theme_bw() +
   labs(y = "Fish consumed (g)",
-       x = "",
-       title = "") +
-  facet_rep_wrap(~season,
-                 scales = "free_x",
+       x = "", title = "") +
+  facet_rep_wrap(~season,scales = "free_x",
                  repeat.tick.labels = F) +
@@
                                    color = "black"),
-        axis.title.x = element_blank(),
         axis.text.y = element_text(size = 11,
@@
         panel.grid.minor.y = element_blank(),
-        plot.margin = unit(c(-.1, .3, .1, .1), "cm"))
+        plot.margin = unit(c(-.2, .3, .1, .1), "cm"))
 
@@
 2019."}
+#impact_p = impact_df %>%
+#  select(season,
+#         perc_diet,
+#         p_chnk,
+#         N = juv_N,
+#         Lci = juv_Lci,
+#         Uci = juv_Uci) %>%
+#  group_by(perc_diet, p_chnk) %>%
+#  summarise(across(c("N", "Lci", "Uci"),sum),
+#            .groups = "drop") %>%
+#  mutate(Stage = "Juveniles") %>%
+#  bind_rows(impact_df %>%
+#              select(season,
+#                     perc_diet,
+#                     p_chnk,
+#                     N = adult_N,
+#                     Lci = adult_Lci,
+#                     Uci = adult_Uci) %>%
+#              group_by(perc_diet, p_chnk) %>%
+#              summarise(across(c("N", "Lci", "Uci"),
+#                   sum),
+#             .groups = "drop") %>%
+#             mutate(Stage = "Adult Equivalents")) %>%
+#  mutate(p_chnk = p_chnk * 100) %>%
+#  rename(perc_chnk = p_chnk) %>%
+#  filter(perc_chnk %in% c(10, 30, 50)) %>%
+#  ggplot(aes(x = perc_diet,y = N)) +
+#  geom_errorbar(aes(ymin = Lci,
+#                    ymax = Uci),
+#                width = 0.1) +
+#  geom_point(size = 3) +
+#  facet_grid(vars(Stage),
+#             vars(perc_chnk),
+#             scales = "free_y") +
+#  scale_color_viridis_d(name = "Percent\nFish in Diet") +
+#  scale_y_continuous(labels = scales::comma) +
+#  scale_x_continuous(breaks = seq(0, 100, by = 10)) +
+#  theme(#axis.text.x = element_blank(),
+        #legend.position = "bottom",
+        #axis.title.x = element_blank(),
+        #axis.ticks.x = element_blank(),
+#        axis.text.y = element_text(size = 11,
+#                                   color = "black"),
+#        strip.text.x = element_text(size = 11,
+#                                    color = "black"),
+#        strip.text.y = element_text(size = 11,
+#                                    color = "black")) +
+#  labs(y = "Number Chinook Consumed")
+
+#impact_p
+
 impact_p = impact_df %>%
@@
   group_by(perc_diet, p_chnk) %>%
-  summarise(across(c("N", "Lci", "Uci"),
-                   sum),
+  summarise(across(c("N", "Lci", "Uci"),sum),
             .groups = "drop") %>%
   mutate(Stage = "Juveniles") %>%
-  bind_rows(impact_df %>%
-              select(season,
-                     perc_diet,
-                     p_chnk,
-                     N = adult_N,
-                     Lci = adult_Lci,
-                     Uci = adult_Uci) %>%
-              group_by(perc_diet, p_chnk) %>%
-              summarise(across(c("N", "Lci", "Uci"),
-                   sum),
-             .groups = "drop") %>%
-             mutate(Stage = "Adult Equivalents")) %>%
   mutate(p_chnk = p_chnk * 100) %>%
@@
   filter(perc_chnk %in% c(10, 30, 50)) %>%
-  ggplot(aes(x = perc_diet,
-             y = N,
-             color = as.factor(perc_diet))) +
-  geom_errorbar(aes(ymin = Lci,
-                    ymax = Uci),
-                width = 0.1) +
-  geom_point(size = 3) +
-  facet_grid(vars(Stage),
-             vars(perc_chnk),
-             scales = "free_y") +
-  scale_color_viridis_d(name = "Percent\nFish in Diet") +
+  ggplot(aes(x = perc_diet,y = N, group=as.factor(perc_chnk), color=as.factor(perc_chnk))) +
+  geom_point(size = 2) + geom_line(size = 1.05)+
   scale_y_continuous(labels = scales::comma) +
-  theme(axis.text.x = element_blank(),
-        legend.position = "bottom",
-        axis.title.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        axis.text.y = element_text(size = 11,
+  scale_x_continuous(breaks = seq(0, 100, by = 10)) +
+  scale_color_manual(name="Proportion Chinook", values=c("lightsteelblue","steelblue3","navyblue"))+
+  theme(axis.text.y = element_text(size = 11,
                                    color = "black"),
-        strip.text.x = element_text(size = 11,
+        axis.text.x = element_text(size = 11,
                                     color = "black"),
         strip.text.y = element_text(size = 11,
-                                    color = "black")) +
-  labs(y = "Number Chinook Consumed")
+                                    color = "black"),
+        legend.position = c(.15, .85),legend.background = element_blank(),
+        legend.key.size = unit(.5, "cm"),legend.box = 'horizontal') +
+  labs(y = "Estimated Chinook Consumed", x="Percent Fish in Pikeminnow Diet")
 

### 2022-01-22 analysis/paper/paper.Rmd
Updated sensitivity plot with values from Zimmerman et. al (1999)

@@
                 as.integer)) %>%
-  crossing(p_chnk = c(.01, .10, .30, .50, .70, .90, 1)) %>%
+  crossing(p_chnk = c(.01, .10, .30, .50, .65, .90, 1)) %>%
   mutate(g_chnk = g_fish_cons * p_chnk) %>%
@@
 pos = position_dodge(width=0.3)
-pm_est_labels <- c("Chapman","Schnabel","Schnabel \n Delayed-mixing")
+pm_est_labels <- c("Chapman","Schnabel","Schnabel \n [Delayed-mixing]")
 all_est %>%
@@
   rename(perc_chnk = p_chnk) %>%
-  filter(perc_chnk %in% c(10, 30, 50)) %>%
-  ggplot(aes(x = perc_diet,y = N, group=as.factor(perc_chnk), color=as.factor(perc_chnk))) +
-  geom_point(size = 2) + geom_line(size = 1.05)+
+  filter(perc_chnk %in% c(30, 50, 65)) %>%
+  ggplot(aes(x = perc_diet,y = N, group=as.factor(perc_chnk), linetype=as.factor(perc_chnk))) +
+  geom_line(size = 1.05)+
   scale_y_continuous(labels = scales::comma) +
   scale_x_continuous(breaks = seq(0, 100, by = 10)) +
-  scale_color_manual(name="Proportion Chinook", values=c("lightsteelblue","steelblue3","navyblue"))+
+  scale_linetype_manual(name="Percent juvenile chinook \nof total fish consumed", values=c("dotted","solid","dashed"))+
   theme(axis.text.y = element_text(size = 11,
@@
                                     color = "black"),
-        legend.position = c(.15, .85),legend.background = element_blank(),
-        legend.key.size = unit(.5, "cm"),legend.box = 'horizontal') +
+        legend.position = c(.19, .82),legend.background = element_blank(),
+        legend.key.size = unit(.56, "cm"),legend.box = 'horizontal') +
   labs(y = "Estimated Chinook Consumed", x="Percent Fish in Pikeminnow Diet")

### 2022-01-25 analysis/paper/paper.Rmd
Added adult impact table & minor caption edits

@@
 
-```{r adult-imp, eval = F}
+```{r adult-imp}
 impact_df %>%
@@
 
+```{r adult-imp-tab, echo =TRUE}
+
+impact_tbl = impact_df %>%
+  select(season,
+         perc_diet,
+         p_chnk,
+         N = juv_N,
+         Lci = juv_Lci,
+         Uci = juv_Uci) %>%
+  group_by(perc_diet, p_chnk) %>%
+  summarise(across(c("N", "Lci", "Uci"),sum),
+            .groups = "drop") %>%
+  mutate(Stage = "Juveniles") %>%
+  mutate(p_chnk = p_chnk * 100) %>%
+  rename(perc_chnk = p_chnk) %>%
+  filter(perc_chnk %in% c(30, 50, 65))%>%
+
+  bind_rows(impact_df %>%
+              select(season,
+                     perc_diet,
+                     p_chnk,
+                     N = adult_N,
+                     Lci = adult_Lci,
+                     Uci = adult_Uci) %>%
+              group_by(perc_diet, p_chnk) %>%
+              summarise(across(c("N", "Lci", "Uci"),
+                   sum),
+             .groups = "drop") %>%
+             mutate(Stage = "Adult Equivalents") %>%
+  mutate(p_chnk = p_chnk * 100) %>%
+  rename(perc_chnk = p_chnk) %>%
+  filter(perc_chnk %in% c(30, 50, 65))) %>%
+  kable(booktabs = T,
+        format.args = list(big.mark = ","),
+        caption = "Estimates of juvenile fish consumed and adult equivalents for diet scenarios comprised of 10-90% fish.
+Estimates and 95% confidence intervals are calculated using Northern Pikeminnow abundance estimates of 28,144 (95% CI: 14,378 - 59,978) for fall and 14,897 (95% CI: 7,610 - 31,747) for spring.
+'Adult equivalents' are estimated using a Lower Granite Dam smolt-to-adult rate (SAR) for Chinook Salmon from McCann et al.
+2019.") %>%
+  kable_styling(font_size = 10)
+
+impact_tbl
+
+```
+\newpage
+
 # Figures
@@
 
-```{r cpue-fig, fig.cap = "The number of Northern Pikeminnow caught per angler effort for each sample effort."}
+```{r cpue-fig, fig.cap = "The number of Northern Pikeminnow caught per angler effort for each sample effort.", eval = FALSE}
 total_cpue %>%
@@
 
-```{r bio-p, fig.cap = "The cumulative amount of fish (g) consumed by an individual Northern Pikeminnow during the peak DSR (fall) and NRR (spring) Chinook Salmon emigrations (top panel) and corresponding daily mean water temperatures (bottom panel).
-The black line is the median among diet scenarios (i.e., where 50% of the Northern Pikeminnow diet consists of fish) and the gray area shows the range among model scenarios (i.e., 10 - 90% of Northern Pikeminnow diet consists of fish).
+```{r bio-p, fig.cap = "Cumulative amount of fish (g) consumed by an individual Northern Pikeminnow during the peak DSR (fall) and NRR (spring) Chinook Salmon emigrations (top panel) and corresponding daily mean water temperatures (bottom panel).
+The black line indicates the diet scenario comprised of 50% fish prey and the gray area indicates the range among diet scenarios (10 - 90% of Northern Pikeminnow diet consists of fish).
 Daily mean water temperatures were summarized from six years of data available from the Shoup gage station.
@@
   filter(perc_chnk %in% c(30, 50, 65)) %>%
-  ggplot(aes(x = perc_diet,y = N, group=as.factor(perc_chnk), linetype=as.factor(perc_chnk))) +
-  geom_line(size = 1.05)+
+  ggplot(aes(x = perc_diet,y = N, group=as.factor(perc_chnk), color=as.factor(perc_chnk))) +
+  geom_line(size = 1.1)+
   scale_y_continuous(labels = scales::comma) +
   scale_x_continuous(breaks = seq(0, 100, by = 10)) +
-  scale_linetype_manual(name="Percent juvenile chinook \nof total fish consumed", values=c("dotted","solid","dashed"))+
+  scale_color_manual(name="% Juvenile Chinook \nof total fish consumed", values=c("lightsteelblue","steelblue3","navyblue"))+
   theme(axis.text.y = element_text(size = 11,
@@
         legend.key.size = unit(.56, "cm"),legend.box = 'horizontal') +
-  labs(y = "Estimated Chinook Consumed", x="Percent Fish in Pikeminnow Diet")
+  labs(y = "Estimated Juvenile Chinook Consumed", x="Percent Fish in Pikeminnow Diet")
 

### 2022-01-29 analysis/paper/paper.Rmd
started incorporating TBs awesome edits

@@
 ---
-title: "Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations"
+title: "Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Chinook Salmon Populations"
 author:
@@
 Predation of emigrating juvenile salmonids by piscivorous fishes is perhaps an under-appreciated source of mortality hindering population recovery in the Columbia River basin.
-Reservoirs and tailraces associated with dams on the Snake and Columbia rivers have created favorable conditions for fishes known to consume juvenile salmonids, including the Northern Pikeminnow *Ptychocheilus oregonsis*.
-Indeed, Northern Pikeminnow have been estimated to consume 8% of the approximately 200 million juvenile salmonids emigration through those rivers, annually.
+Reservoirs and tailraces associated with dams on the Snake and Columbia rivers have created favorable conditions for fishes known to consume juvenile salmonids.
+The native Northern Pikeminnow are one of these fishes and have been estimated to consume ~8% of the approximately 200 million juvenile salmonids emigrating throu the lower Snake and Columbia rivers, annually.
 Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above these reservoir systems.
-In this study, we examine predation by Northern Pikeminnow on juvenile salmonids, focusing on Chinook Salmon *Oncorhynchus tshawytscha*, in a slow-water reach of the Salmon River, Idaho known as Deadwater Slough.
-The reach contains historically important overwinter rearing and spring emigration habitat for Chinook Salmon and other ESA-listed species, and was formed by the failure of a small mining dam in a reservoir in the late 19th century.
-We estimate the abundance of Northern Pikeminnow in Deadwater Slough during peak emigrations of Chinook salmon, and further, use a bioenergetics approach to estimate the consumption potential of the Northern Pikeminnow population to assess potential impacts to Chinook Salmon emigrants.
-Finally, we evaulate potential impacts to adult returns of local populations.
-***We found that...
-(insert concluding findings here.*** It is likely that predation of juvenile Chinook Salmon in Deadwater Slough has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG hindering local recovery efforts.
+In this study, we examine predation by Northern Pikeminnow on Chinook Salmon in a slow-water reach of the Salmon River, Idaho known as Deadwater Slough.
+The reach was (at least partially) formed by the failure of a small mining dam in a nearby drainage during the late 19th century and contains historically important overwinter rearing habitat is a migration corridor for Chinook Salmon and other Endangered Species Act-listed species.
+We estimate the abundance of Northern Pikeminnow during peak emigrations to be ***insert numbers here***.
+Using a bioenergetics approach we estimate the consumption potential of the Northern Pikeminnow population to be ***key results here***.
+This resulted in the potential consumption of ***numbers*** Chinook Salmon emigrants.
+Using a sensitivity analysis, we evaluated the potential impact to adult returns of local Chinook Salmon populations to be between ***number with CIs***.
+It is likely that predation of juvenile Chinook Salmon in Deadwater Slough has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River major population group, thereby hindering local recovery efforts.
 
@@
 mykiss*.
-These fishes were historically abundant throughout the basin and consequently have tremendous ecological, cultural, and economic value.
-Over the last century, salmonid stocks throughout the Snake and Columbia rivers have undergone significant declines related to habitat modification in the freshwater environment (e.g., removal of beavers, river channel simplification, water withdrawals, hydrosystem development) and poor marine conditions, both affecting their survival and recruitment at multiple life stages.
-Consequently, action agencies have made considerable investments in the rehabilitation of tributary ecosystems and evaulation of factors attributed to the species decline [@Roni2018].
+These fishes were historically abundant throughout the basin and consequently have tremendous ecological, cultural, and economic value [@Atlas2021; @Lewis2019].
+Over the last century, salmonid stocks throughout the Snake and Columbia rivers have undergone significant declines related to habitat modification in the freshwater environment (e.g., removal of beavers, river channel simplification, water withdrawals, hydrosystem development) and poor marine conditions, affecting their survival and recruitment at multiple life stages [@Crozier2020; @Clark2020; @Justice2017].
+Consequently, action agencies have made considerable investments in the rehabilitation of tributary ecosystems and evaluation of factors attributed to the species decline [@White2021; @Roni2018].
 
@@
 Northern Pikeminnow tolerate and thrive in relatively warm, slow water habitats [@Wydoski2003] and consequently have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999].
-Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].
+Indeed, Northern Pikeminnow are estimated to consume ~8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].
 
@@
 One such example is Deadwater Slough, an approximately `r round(ddw_rkms, 1)` km long reach of unnaturally slow and deep water in the Salmon River, Idaho.
-The Deadwater Slough is located in a river reach containing historically important overwinter rearing and spring emigration habitat for juvenile Chinook Salmon *O.
-tshawytscha* and is inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*.
+The Deadwater Slough is in a river reach containing historically important overwinter rearing habitat, includes the emigration corridor for juvenile Chinook Salmon *O.
+tshawytscha*, and is inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*.
 Further, the slough is part of the migratory pathway of the endangered Snake River Sockeye Salmon *O.
@@
 Currently, the Upper Salmon MPG supports eight independent, extant populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
+The Lemhi River historically contained the largest Chinook Salmon population within the MPG and is an intensively monitored watershed to determine responses to restoration actions [@Zimmerman2012].
 The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a].
@@
 Alternatively, NRR migrants remain in their natal areas for approximately one year after emergence until emigration to the ocean as smolts.
-Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency.
-Because the Deadwater Slough supports rearing for DSR migrants and is part of the migration corridor for NRR migrants, and juveniles from all Upper Salmon populations must rear in or migrate through it, it represents a significant habitat for the Upper Salmon MPG.
-
-Recent studies have examined the downstream movement, distribution, and (apparent) survival of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach.
+Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency [@Dodson2013].
+Deadwater Slough represents an important habitat for the Upper Salmon MPG as is supports rearing and migration of juveniles from all populations including fall DSR and spring NRR migrants.
+
+Recent studies have examined the downstream movement, distribution, and apparent (hereafter 'survival') of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach.
 Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have substantially reduced survival in the Deadwater Slough reach compared to adjacent reaches [@Axel2015].
 Similarly, emigrating DSR Chinook Salmon had a 10% reduction in transition probability through the Deadwater Slough compared to surrounding reaches during the fall and early winter months [@Ackerman2018; @Porter2019].
-Low survival in these cases was attributed to low-velocity water delaying movement rates, thereby increasing predation risk.
+Low survival was attributed to low-velocity water delaying movement rates, thereby increasing predation risk.
 
 In this study, we estimated the abundance of Northern Pikeminnow in the Deadwater Slough and evaluated potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
-We hypothesize that high densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon and Sockeye Salmon.
-To test this, our objectives were three-fold:
-
-1. Estimate the abundance (or relative abundance) of potential predators in the Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations.
-2. Document predation on juvenile Chinook Salmon during the DSR and NRR emigration periods using gastric lavage.
-3.
-Use a bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population at Deadwater Slough to assess their potential impacts to DSR and NRR Chinook Salmon emigrants.
-
-We also consider how consumption of juvenile Chinook Salmon emigrants at Deadwater Slough by Northern Pikeminnow may impact adult returns to the Upper Salmon MPG.
+We hypothesize that high densities of piscivorous predators in the Deadwater Slough may explain the reduced survival observed for juvenile Chinook Salmon.
+To test this, our objectives were:
+
+1. Estimate the abundance of potential predators in the Deadwater Slough during the peaks of fall DSR and spring NRR juvenile emigrations;
+2. Document predation on juvenile Chinook Salmon during the DSR and NRR emigration periods using gastric lavage;
+3. Use a bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population on DSR and NRR Chinook Salmon emigrants at Deadwater Slough;
+4. Estimate how consumption of juvenile Chinook Salmon emigrants at Deadwater Slough by Northern Pikeminnow impacts adult returns to the Upper Salmon MPG.
 
@@
 
-The Deadwater Slough is an approximately `r round(ddw_rkms, 1)` km section of the mainstem Salmon River located approximately 6 river km downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
+The Deadwater Slough is an approximately `r round(ddw_rkms, 1)` km section of the mainstem Salmon River located roughly 6 river kilometers (RKM) downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
 The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
-Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River [@Emerson1973] forming an alluvial fan.
-The fan created a slow, deep section in the river, spanning approximately `r round(ddw_acres)` acres (`r round(ddw_hectares)` hectares), averaging `r round(ddw_avg_width_m)` m width, and resembling a small, narrow pond or reservoir.
-
-## Northern Pikeminnow Population Size
-
-We estimated the population size of piscine predators in the Deadwater Slough using a mark-recapture survey and a catch per unit effort (CPUE) approach.
+Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event and upslope instability that deposited substantial amounts of sediment at the confluence of the Salmon River [@Emerson1973].
+That sediment formed an alluvial fan which created a slow, deep section in the river, spanning approximately `r round(ddw_hectares)` hectares, averaging `r round(ddw_avg_width_m)` m wide, and resembling a small reservoir.
+
+## Northern Pikeminnow Demographics
+
+We estimated the population size of piscine predators in the Deadwater Slough using a mark-recapture survey design and a catch per unit effort (CPUE) approach.
 Predators were sampled near the peaks of the juvenile DSR and NRR emigrations.
-Our initial intent was to sample during the emigration of the 2018 brood year (fall 2019 and spring 2020).
-However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 effort was delayed until fall 2020.
-Although this provided the benefit of an additional abundance estimate during the DSR emigration, estimates during the NRR emigration were still lacking, and so an additional effort was added in spring 2021.
-Fall efforts included sampling across a two-week span, but due to budget constraints the spring 2021 effort was limited to a single week.
-Fall efforts were constrained within a two-week window to minimize predator emigration/immigration and to constrain the population to meet assumptions of the mark-recapture models.
-
-Multiple capture methods were employed during the initial fall 2019 effort to reduce selectivity and bias for species and size classes.
+During our initial survey in 2019, Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and consequently became our focal taxa.
+Northern Pikeminnow were surveyed near the peaks of the fall DSR and spring NRR emigrations.
+The initial intent was to survey during the emigration of the 2018 brood year (fall 2019 and spring 2020), however logistical constraints during the onset of the Covid-19 pandemic delayed the spring 2020 survey until fall 2020.
+Although this allowed an additional survey during the fall DSR emigration, surveys during the spring NRR emigration were still lacking.
+Consequently, an additional week of surveying was added in spring 2021.
+Fall surveys occurred over two weeks to minimize Northern Pikeminnow immigration/emigration.
+
+Multiple capture methods were employed during the initial fall 2019 survey to reduce selectivity and bias for species and size classes.
 Methods included raft electrofishing, fyke netting, snorkeling, and angling.
-Ineffective methods were then abandoned in subsequent years.
+Ineffective methods were abandoned in subsequent years.
 <!-- For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) our inability to "chase" fish downstream as they evaded the electricity.
@@
 Additionally, there were concerns that raft electrofishers and fyke nets could incidentally capture ESA-listed adult steelhead that are present in the mainstem Salmon River during fall and spring.
--->After evaluating all methods, angling proved to be the most effective method for capturing piscine predators while also minimizing potential impacts to ESA-listed adult steelhead present in the Salmon River during our sampling efforts.<!--, but recognizing that angling can have both species and size selectivity (bias).--> Therefore, the following analyses will focus on fish captured by angling, unless otherwise noted.
-During the fall 2019 effort it additionally became clear that Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and so is our focus.
+-->After evaluating all methods, angling was the most effective method for capturing Northern Pikeminnow while also minimizing potential impacts to ESA-listed adult steelhead that were present in the Salmon River during our surveys.<!--, but recognizing that angling can have both species and size selectivity (bias).--> Therefore, the following analyses will focus on fish captured by angling, unless otherwise noted.
 
 Our study relied heavily on volunteer anglers who were permitted to fish anywhere within Deadwater Slough.
-During a sampling occassion (day), anglers would boat or hike their catch (periodically or upon filling a livewell) to a processing station at the boat ramp approximately 500 m downstream from the top of the slough.
+Each survey day, anglers boated or hiked their catch (periodically or upon filling a livewell) to a processing station at the boat ramp, approximately 500 m downstream from the top of the slough.
 For each fish, we recorded the date of capture, species, total length (TL; mm), and whether the fish was previously marked.
-Unmarked fish were given a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) unique to each day and released.
-Fish that died prior to release were still counted as captured (or recaptured), but not included in the number of marked fish available for recapture after that date.
-Released individuals were distributed throughout Deadwater Slough to facilitate mixing back into the population.
-Finally, we recorded the angling start and end time for each crew (person or combinations of persons) to facilitate calculations of CPUE.
+Unmarked fish were given a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) unique to each day and then were released.
+Released fish were distributed throughout Deadwater Slough to facilitate mixing back into the population.
+Fish that died prior to release were included in the total capture (or recapture) count, but not included in the number of marked fish available for recapture after that date.
+Finally, we recorded the angling start and end time for each crew (person or combinations of persons) to calculate CPUE as number of Northern Pikeminnow caught per angler hour.
 
@@
 
-Four different mark-recapture estimators were employed to explore variation in abundance estimates among them.
+Two types of mark-recapture estimators were used to explore variation in abundance estimates: a single census and a multiple census estimator.
 Estimators fell into two broad categories: single census and multiple census.
-For the single census estimators, we treated the first week as the first (mark) sample, and the following (second) week as the second (recapture) sample, pooling data within each of those weeks.
-Alternatively, the multiple census estimators treat each day as a sample and use information about the total marked fish from all previous samples to infer the total abundance.
-
-For the single census estimators, we first used the Lincoln-Peterson estimator:
-
+For the single census estimator, we treated the first week of the survey as the mark event and the following week as the recapture event, pooling data within each event week.
+Alternatively, the multiple census estimator treated each day as an event and use information about the total number of marked fish from all previous events to calculate the total abundance.
+
+For the single census estimator, we used the Chapman-modified Lincoln-Peterson estimator:
+
+<!--
 $$
@@
 $$
-where $M$ is the total number of fish marked and returned to the population during the first sample (week), $n$ is the total number of fish caught in the second sample (week), and $m$ is the number of marked fish caught during the second week.
-Because the Lincoln-Peterson estimator can be biased with small samples, we also employed the Chapman-modified Lincoln-Peterson estimator:
-
+-->
 $$
@@
 $$
-For a multiple census estimator, we first utilized the Schnabel method:
+
+where $M$ is the total number of fish marked and returned to the population during the first event (week), $n$ is the total number of fish caught in the second event (week), and $m$ is the number of marked fish caught during the second week.
+For a multiple census estimator, we used the Schnabel method:
 
@@
 
-where $M$, $n$, and $m$ are indexed by each sample (day), $i$.
+where $M$, $n$, and $m$ are indexed by each survey (day), $i$.
 With small numbers of recaptures, it may not be appropriate to calculate the standard error of the Schnabel estimator but the 95% confidence intervals can be calculated using a Poisson approximation [@Krebs1999].
-The Schnabel estimator is essentially a weighted average of a series of Lincoln-Peterson estimators (with a Champman modification).
+The Schnabel estimator is essentially a weighted average of a series of Chapman-modified Lincoln-Peterson estimators.
+
+Because captures of new and marked fish were dependent on volunteer anglers, there was concern that sampling was biased to locations preferred by anglers.
+If marked fish were not recently released or mixed back into those locales, marked fish may have been less susceptible to capture.
+To account for this potential bias, we adjusted the Schnabel estimator with an assumption that marked and released fish during each survey day were not available for recapture for 48 hours.
+
+<!-- delayed mixing Schnabel equation here; include statement about fish being "hook shy" -->
+
+<!--
 Lastly, we explored one additional multiple-census estimator, the Schumacher-Eschmeyer, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
@@
 $$
-
-The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and 2020 sample efforts (the estimate is of the initial population size).
-However, to estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of total CPUE in the fall efforts to the total CPUE in the spring efforts, then multiplied that ratio by the average abundance from the fall efforts.
-This approach assumes equal capture probabilities between fall and spring efforts.
-
-All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the sampling period), (2) all fish have equal chance of being caught in the second (and subsequent) sample(s), (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or mistaken.
+-->
+
+The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and 2020 surveys.
+However, to estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of total CPUE in the fall survey to the total CPUE in the spring survey, then multiplied that ratio by the average abundance from the fall surveys.
+
+$$
+\hat{N}_{spring} = \overline{N}_{fall} * \frac{CPUE_{spring}}{CPUE_{fall}}
+$$
+
+This approach assumes equal capture probabilities for fall and spring surveys.
+Fall and spring population estimates were converted into a linear density by dividing by the length of Deadwater Slough (km) measured in GIS software and using drone-generated orthography.
+
+All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the survey period), (2) all fish have equal chance of being caught in subsequent survey days, (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or mistaken.
 
@@
 
-Northern Pikeminnow in Deadwater Slough were not only abundant, but anglers reported that many were large-bodied.
-For that reason, we calculated the proportional stock density (PSD) for Northern Pikeminnow in Deadwater Slough, which is a measure of species size structure.
-PSD is a ratio, typically expressed as a percentage, between the number of "quality-sized" (larger) individuals and "stock-sized" individuals:
+Anglers reported that Northern Pikeminnow in Deadwater Slough were large-bodied.
+To describe the population size structure, we calculated the proportional stock density (PSD) for Northern Pikeminnow in Deadwater Slough.
+PSD is a ratio, expressed as a percentage, between the number of "quality-sized" (larger) individuals and "stock-sized" individuals:
 
@@
 where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
-Quality and stock size definitions vary by species.
-We calculated the PSD for Northern Pikeminnow in Deadwater Slough using `r prettyNum(q_size, digits = 0)` mm TL for quality-size and `r prettyNum(s_size, digits = 0)` mm TL for stock-size [@Winther2020].
+For Northern Pikeminnow in Deadwater Slough, we used `r prettyNum(q_size, digits = 0)` mm TL for quality-length and `r prettyNum(s_size, digits = 0)` mm TL for stock-length [@Winther2020].
 
@@
 
-Gastric lavage [@Foster1977] was used to examine the stomach contents of Northern Pikeminnow for the presence of juvenile Chinook Salmon and other fishes (e.g., juvenile steelhead, juvenile Sockeye Salmon, Redside Shiner *Richardsonius balteatus*, etc.).
-Immediately following lavage, stomach contents of individuals were preserved with 99% isopropyl alcohol in whirl-paks for later analysis in a controlled environment.
-For each sample, wet weight (grams) was recorded for the total combined stomach content, including all non-fish items (e.g., macroinvertebrates, organic matter), as well as that portion consisting of fish parts or matter.
+Gastric lavage [@Foster1977] was used to examine the stomach contents of Northern Pikeminnow for the presence of juvenile Chinook Salmon and other fishes occupying Deadwater Slough.
+Immediately following lavage, stomach contents of individuals were preserved with 99% isopropyl alcohol in whirl-paks  to be analyzed in a controlled environment.
+For each stomach sample, wet weight (grams) was recorded for the total stomach contents, including all non-fish items (e.g., macroinvertebrates, organic matter), as well as the portion consisting of fish.
 Fish and fish remnants were identified to the lowest taxonomic unit using diagnostic bones, or were categorized as unknown [@Hansel1988; @Frost2000].
-A subset of Northern Pikeminnow captures (~ 5%) were euthanized for dissection after gastric lavage to validate the efficacy of the methodology.
+Approximately 5% of Northern Pikeminnow were euthanized for dissection (n = 75) after gastric lavage to validate the efficacy of the methodology.
 
@@
 
-## Northern Pikeminnow Population Size
+## Northern Pikeminnow Demographics
 

### 2022-01-29 analysis/paper/paper.Rmd
knit word doc

@@
 
-```{r adult-imp}
+```{r adult-imp, eval = F}
 impact_df %>%

### 2022-01-31 analysis/paper/paper.Rmd
methods edits

@@
 
-To estimate the total consumption potential (i.e., the number or amount of juvenile Chinook Salmon consumed during the peaks of DSR [fall] and NRR [spring] emigrations) of Northern Pikeminnow in Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 and applied in the R statistical software [@RCoreTeam2021].
+To estimate the total consumption potential of Northern Pikeminnow in Deadwater Slough during the peaks of fall DSR and spring NRR emigrations, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 applied in R statistical software [@RCoreTeam2021].
 The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on the following inputs: predator and prey energy densities, predator start and end weights, and water temperatures.
+Separate models were run for the fall period (September 15 - November 30) and the spring period (March 1 - May 31) to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG.
+
 Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017].
-Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon [@Moss2016].
-The predator start weight was calculated using a weight-length formula [@Parker1995] and the average length of Northern Pikeminnow caught in Deadwater Slough during our study.
-The average length of Northern Pikeminnow caught during the fall and spring efforts were `r avg_fall_NPM` and `r avg_spring_NPM` mm, respectively, which calculated to average starting weights of 504.4 and 670.5 g.
-We assumed no growth in individual Northern Pikeminnow, and so predator start and end weights are equal (i.e., we assumed the population biomass was stable).
-Mean daily water temperatures were summarized from 15-minute interval temperature readings available March 3, 2013 to June 14, 2021 from a gage station approximately 22 river kilometers downstream of Deadwater Slough near Shoup, Idaho [@USGS2016].
-
-The largest uncertainty among model parameters was the proportion of the Northern Pikeminnow diet consisting of fish versus non-fish prey items (e.g., invertebrates).
-Therefore, we conducted nine model runs with varying proportions of fish in the diet ranging from 10-90%, in 10% increments.
-Energy densities for invertebrates was fixed at 3,000 J/g.
-Because we were unable to differentiate juvenile Chinook Salmon from other fish prey consumed during model runs, we assume all fish prey have the same energy densities as juvenile Chinook Salmon i.e., juvenile Chinook Salmon were used as surrogates for all fish prey items.
-Separate models were also run for a 77-day fall period from September 15 - November 30 and a 92-day period in the spring from March 1 - May 31 to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG.
-This resulted in 18 separate model runs (9 diet scenarios and 2 seasons).
-Each model run provided an estimate of the total amount (grams) of fish consumed by an individual Northern Pikeminnow during the fall or spring time periods.
-To estimate the total biomass of fish potentially consumed by all Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated predator population size during each period (fall and spring).
+Prey energy densities were fixed at 3,000 J/g for invertebrates and 21,500 J/g for juvenile Chinook Salmon [@Moss2016].
+Because were were unable to differentiate juvenile Chinook Salmon from other fish prey, we assume all fish prey have the same energy densities as juvenile Chinook Salmon.
+The average length of Northern Pikeminnow caught in Deadwater Slough during our study was converted to weight using a weight-length formula from @Parker1995.
+This resulted in average spring and fall starting eights of 504.4 and 670.5 g for Northern Pikeminnow with average lengths of `r avg_fall_NPM` and `r avg_spring_NPM` mm, respectively.
+We assumed no growth in individual Northern Pikeminnow, resulting in equal start and end weights.
+Mean daily water temperatures were summarized from 15-minute interval temperature readings between March 3, 2013 and June 14, 2021 from USGS gage station 13307000, approximately 22 RKM downstream of Deadwater Slough.
+
+The proportion of fish in the Northern Pikeminnow's diet relative to non-fish prey items is unknown in Deadwater Slough.
+Therefore, we conducted separate model runs that assume different proportions of fish in the diet ranging from 30 - 90%, in 10% increments.
+These values were supported by observations elsewhere in the Columbia River basin (54-86%: @Shively1996; 48-86%: @Zimmerman1999; 37%: @Gray2001) that suggest fish are the majority of Northern Pikeminnow prey.
+14 models were run (7 diet scenarios and 2 seasons), each providing an estimate of the total grams of fish consumed by an individual Northern Pikeminnow (***table reference***).
+To estimate the total biomass of fish consumed by Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated Northern Pikeminnow population size during fall and spring.
 
@@
 
-Multiplying the grams of fish consumed by an individual Northern Pikeminnow by the Northern Pikeminnow population size provides an estimate of total fish consumption potential for the population.
-However, the bioenergetics model runs do not differentiate juvenile Chinook Salmon from other fish prey items.
-Therefore, to understand potential impacts to local Chinook Salmon populations, we need to parse out the portion of total fish consumed that is juvenile Chinook Salmon.
-Unfortunately, estimates of the proportion of the Northern Pikeminnow fish prey that are juvenile Chinook Salmon are sparse, as well as information on the prevalence of juvenile Chinook Salmon relative to other fish prey items.
-Therefore, we instead consider hypothetical scenarios where Chinook Salmon are 1%, 10%, 30%, and 50% of fish prey consumed, by weight, which provide estimates of the total biomass (grams) of juvenile Chinook Salmon consumed by Northern Pikeminnow during the DSR and NRR emigration periods.
-
-The estimate of total juvenile Chinook Salmon biomass consumed was converted to the total number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
-The average weights of DSR and NRR emigrants were taken from fish caught at seven rotary screw traps located throughout the Upper Salmon MPG upstream of Deadwater Slough during the fall and spring time periods.
-
-<!-- Sawtooth Hatchery, East Fork Salmon, Pahsimeroi, Upper Lemhi, Hayden Creek, Lower Lemhi, North Fork Salmon -->
-
-<!--
-## Juvenile Chinook Salmon Abundance
-
-Juvenile Chinook Salmon abundances were estimated at 5 rotary screw traps in the upper Salmon River basin: Sawtooth, Pahsimeroi, Hayden Creek, Upper Lemhi, and Lower Lemhi.
-Abundances were calculated by Idaho Department of Fish and Game at each trap which made up the production estimate for juvenile Chinook Salmon in the Upper Salmon Basin (Table X) (McClure et al.
-2020).
--->
+The proportion of Chinook Salmon prey relative to other fish prey in the Northern Pikeminnow diet is also unknown.
+However, there is some evidence that juvenile salmonids are by far the most consumed fish prey [@Zimmerman1999; @Shively1996].
+Moreover, it is estimated that Chinook Salmon make up 64.2%, 29.3%, and 49.3% of the fish prey consumed by Northern Pikeminnow in the Columbia River below Bonneville Dam, Columbia River reservoirs, and in the lower Snake River, respectively [@Zimmerman1999].
+Because the bioenergetics model does not differentiate juvenile Chinook Salmn from other fish prey items, we performed a sensitivity analysis to assess potential impacts of Northern Pikeminnow predation on local Chinook Salmon populations.
+Using values similar to @Zimmerman1999 we modeled three diet scenarios where Chinook Salmon comprised 30%, 50%, and 65% of the fish prey, by weight, in the Northern Pikeminnow diet during the fall DSR and spring NRR emigration periods.
+In this analysis, we assume 60% of the Northern Pikeminnow's total diet is comprised of fish, representing the median in our range of modeled diets.
+
+The estimate of total juvenile Chinook Salmon biomass consumed was then converted to the total estimated number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
+The average weights of DSR and NRR emigrants were calculated from fish captured at seven rotary screw traps upstream of Deadwater Slough during the fall and spring periods.
+
+Although the primary impact to Chinook Salmon in Deadwater Slough by Northern Pikeminnow is juvenile predation, the common metric to evaluate salmon recovery is often adult returns.
+We assessed the impacts on adult returns by estimating the number of additional adult Chinook Salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough was reduced or eliminated.
+This calculation was made by multiplying the total estimated juvenile Chinook Salmon consumed by an estimate of Granite-to-Granite smolt-to-adult survival rate (SAR) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.
 
@@
 
-Using hook-and-line angling, we caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 efforts, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish.
-The CPUE, measured as number of Northern Pikeminnow caught per angler hour, was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021 (Figure \@ref(fig:cpue-fig)).
-During the mark-recapture efforts, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019 which included `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals.
-There were an additional `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020 consisting of `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals.
+Using hook-and-line angling, we caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 surveys, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish.
+The CPUE was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021 (Figure \@ref(fig:cpue-fig)).
+During the mark-recapture events, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019, including `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals and `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020, including `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals.
 Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.
 
-Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` (Table \@ref(tab:abund-tab)).
-In both cases, the multiple census estimators were larger than the single census estimators.
-Confidence intervals for the fall 2020 effort were more varied, likely due to fewer recaptures during that year.
+Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` (Table \@ref(tab:abund-tab)).For both estimates, the multiple census estimates were larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (***figure reference***).
 

### 2022-01-31 analysis/paper/paper.Rmd
incorporated TB results edits

@@
 
+$$
+\hat{N} = \frac{\sum\limits_{i = 1}^k n_i M_{i-1}}{\left(\sum\limits_{i = 1}^k m_i \right) + 1}
+$$
+
 <!-- delayed mixing Schnabel equation here; include statement about fish being "hook shy" -->
-
-<!--
-Lastly, we explored one additional multiple-census estimator, the Schumacher-Eschmeyer, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population.
-The Schumacher-Eschmeyer estimator is as follows:
-
-$$
-\hat{N} = \frac{\sum\limits_{i = 1}^k n_i M^2_i}{\sum\limits_{i = 1}^k m_i M_i}
-$$
--->
 
@@
 The average length of Northern Pikeminnow caught in Deadwater Slough during our study was converted to weight using a weight-length formula from @Parker1995.
+
+$$
+log_{10} W_s = -4.886 + 2.986[log_{10}(FL)]
+$$
+
 This resulted in average spring and fall starting eights of 504.4 and 670.5 g for Northern Pikeminnow with average lengths of `r avg_fall_NPM` and `r avg_spring_NPM` mm, respectively.
@@
 Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` (Table \@ref(tab:abund-tab)).For both estimates, the multiple census estimates were larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (***figure reference***).
-
-All mark-recapture estimators yielded estimates with overlapping confidence intervals.
-However, because our sampling design most closely matched a multiple census estimator (i.e., we released marked fish each day and recaptures from all previous days were recorded for each day) and required a small sample size modification, we used the Schnabel estimates of abundance for further analyses and results.
-Accordingly, the mean of Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`).
-By multiplying the ratio (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`) between the spring CPUE (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"], 2)`) and mean of the fall CPUEs (`r round(mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring 2021.
-
-<!--However, because our sampling design most closely matched a multiple census estimator (i.e., we released marked fish each day and recaptures from all previous days were recorded for each day), we fell those estimators most appropriate.
-Additionally, because the Schnabel estimator includes a small sample size modification, as in our case (i.e., few recaptures), we feel it is most appropriate.
-The Schnabel estimator has the added benefit of more precise confidence intervals and is more straightforward than the Schumacher-Eschmeyer estimate as it is essentially the weighted average among multiple Chapman-modified Lincoln-Peterson estimates.
-Therefore, further analyses and results will be based on the Schnabel estimates of abundance.-->
-
-The PSD for Northern Pikeminnow in Deadwater Slough across all three efforts was `r prettyNum(psd, digits = 1)`% i.e., `r prettyNum(psd, digits = 1)`% of captured fish were over the quality size classification defined by @Winther2020.
-For comparison, @Winther2020 reported a PSD for Northern Pikeminnow captured below Bonneville Dam of 41% and 18% in Bonneville Reservoir, suggesting that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to populations reported elsewhere in the Columbia River.
-In fact, the state record Northern Pikeminnow, which measured `r max(cmr_df$length, na.rm = T)` mm (`r round(max(cmr_df$length, na.rm = T) / 25.4, 1)` in) TL, was caught in Deadwater Slough during the fall 2020 effort.
+Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, the Schnabel estimator was determined to provide the best estimates of abundance.
+Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`).<!--Need to add the delayed-mixing Schnabel estimate results--> For spring 2021, we estimated abundance by multiplying the mean fall abundance by the ratio between the spring CPUE and mean of the fall CPUEs (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), resulting in an estimate of `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough.
+Those estimates translate to linear densities of ***value*** and ***value*** Northern Pikeminnow per RKM.
+
+<!-- need to adjust the spring estimate -->
+
+The lengths of Northern Pikeminnow captured ranged from ***value*** to ***value*** mm with an average length of ***value*** mm (\@ref(fig:length-freq-p)).
+The PSD for Northern Pikeminnow in Deadwater Slough across all three efforts was `r prettyNum(psd, digits = 1)`%.
 
@@
 
-We examined the stomach contents of Northern Pikeminnow from Deadwater Slough using gastric lavage.
-To validate the efficacy of the method, we euthanized 75 Northern Pikeminnow after gastric lavage was completed and removed the remaining food items via dissection.
 We found that gastric lavage successfully removed food items in all dissected individuals, indicating the method was effective.
-Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in some form in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) of those.
-Northern Pikeminnow captured during the spring 2021 effort had a higher proportion of diets with fish contents (`r round(100 * gastric_tbl$fish_contents[gastric_tbl$event_name == "Spring_2021"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Spring_2021"], 1)`%) than fish captured in the fall 2019 (`r round(100 * gastric_tbl$fish_contents[gastric_tbl$event_name == "Fall_2019"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Fall_2019"], 1)`%) and fall 2020 (`r round(100 * gastric_tbl$fish_contents[gastric_tbl$event_name == "Fall_2020"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Fall_2020"], 1)`%) (Table \@ref(tab:gastric-tab)).
-For those individuals with stomach contents, the average wet weight of total contents was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r round(min(stmch_df$stomach_contents_weight), 2)`-`r round(max(stmch_df$stomach_contents_weight), 2)` g).
-Overall, fish or fish remnants were `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall stomach contents examined.
+Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs.
+Northern Pikeminnow captured during spring 2021 had a higher proportion of fish content in their diet compared to individuals captured in the fall surveys (Table \@ref(tab:gastric-tab)).
+The average wet weight of total contents for an individual was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r round(min(stmch_df$stomach_contents_weight), 2)`-`r round(max(stmch_df$stomach_contents_weight), 2)` g).
+Overall, fish or fish remnants made up `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of all stomach contents examined.
 
@@
 
+During the fall DSR emigration, we estimated an average-size Northern Pikeminnow (***value*** mm) to consume ***value*** g of fish to maintain their body size, assuming 60% of their diet consists of fish prey (Figure \@ref(fig:bio-p)).
+Fall consumption ranged from ***value*** g for a diet of 30% fish to ***value*** g for a diet of 90% fish.
+During the spring NRR emigration, we estimated an average-size Northern Pikeminnow (***value***) to consume ***value*** 44.29 g of fish, assuming 60% of their diet is fish prey (Figure \@ref(fig:bio-p)).
+Spring consumption ranged from ***value*** g for a diet of 30% fish to ***value*** g for a diet of 90% fish.
+Given estimated mean Northern Pikeminnow population sizes of ***value*** from fall estimates and ***value*** for spring 2021, and assuming 60% of their diet is fish, we estimate that Northern Pikeminnow can consume ***value*** metric tons of fish during the fall DSR emigration and an additional ***value*** metric tons during the spring NRR emigration in Deadwater Slough.
+In total, the Northern Pikeminnow population can consume ***value*** metric tons of fish during the DSR and NRR emigrations to maintain their population size, assuming 60% of their diet consists of fish.
+
+<!--
 We estimated an individual, average-size (`r avg_fall_NPM` mm) Northern Pikeminnow to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g) of fish during the DSR emigration to maintain their body size, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
@@
 In total, the Northern Pikeminnow population can consume `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 50]) + unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 50]), big.mark = ",")` g of fish during the DSR and NRR emigrations to maintain their population size if we assume 50% of their diet consists of fish.
+-->
 
@@
 
-1. Start with total g of fish consumed by Northern Pikemnnow population, summing DSR and NRR time periods.
-2. Parse out portion of fish total g consumed that is juvenile Chinook Salmon (1%, 10%, 30%, 50%).
-3. Divide by average weight of juvenile Chinook Salmon (10.6 g), the average of DSR (10.3 g) and NRR (10.9 g) emigrants.
-4. Estimate of total juvenile Chinook Salmon consumed.
-5. Multiply by SAR to estimate "adult equivalents".
-
-<!--
-## Juvenile Chinook Salmon Abundance
-For context, throughout the migration periods of DSR and NRR life histories it is estimated that 113,976 individuals are produced and therefore available within the upper Salmon River basin (Table x).
-
-The abundance of juvenile Chinook Salmon for the upper Salmon River basin was estimated at 6 rotary screw traps [@McClure2020].
-Overall, in 2020, there were an estimated 59,289 DSR, 22,264 NRR, and 10,159 parr estimated among all these traps in the upper Salmon River basin (Table X.)
--->
+Using the total g of fish consumed by the Northern Pikeminnow population (summing fall and spring surveys) we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where juvenile Chinook Salmn make up 30%, 50%, and 65% of total fish prey consumed (Figure ***reference***).
+Assuming that 50% of the total fish consumed are juvenile Chinook Salmon, we estimate between (30% fish scenario with CIs) and (90% fish scenario with CIs) juvenile Chinook Salmon will be consumed.
+We then used our estimates of total juvenile Chinook Salmon consumed to estimate "adult equivalents" using the Granite-to-Granite smolt-to-adult survival rate of ***value*** (SE: ***value***; Table ***reference***).
+Assuming predation in Deadwater Slough was reduced or eliminated, the juvenile consumption would be equivalent to ***value*** adults in the spring (range ***value*** - ***value***) and ***value*** adults (range ***value*** - ***value***) in the fall, totaling ***value*** adult Chinook Salmon.
 

### 2022-01-31 analysis/paper/paper.Rmd
incorporated TB discussion edits

@@
 
+## Mark-Recapture Models
+
 ```{r john-day}
@@
 
+All mark-recapture abundance estimators suggest a large population of Northern Pikeminnow occupy Deadwater Slough.
+The mean population size of Northern Pikeminnow in the Deadwater Slough was estimated to be ***value*** during the fall DSR emigration and ***value*** during the spring NRR emigration.
+Those estimates translate to densities (fish/km) of Northern Pikeminnow that are ***value*** and ***value*** times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately ***value*** and ***value*** times greater than projections for the lower Columbia River [@Beamesderfer1996].
+@Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective; therefore it is possible that their estimates may have been low.
+Conversely, at Deadwater Slough the maximum depth was approximately 6 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach.
+Given the size (volume) of the Salmon River at the slough relative to the Columbia River, the estimated Northern Pikeminnow densities for Deadwater Slough are remarkable.
+Further, our estimate densities of Northern Pikeminnow suggest that slow water reaches outside of the reservoir complexes on the Snake and Columbia rivers may support exceptionally high predator densities, consistent with findings of @Harnish2014, @Gray2001, and @Zimmerman1999.
+Given the dearth of current data available in the Columbia River basin, it is unclear how Northern Pikeminnow abundance estimates in this study compare to elsewhere in the Columbia River Basin; contemporary information on piscine predator abundances may be needed [@Widener2021].
+
+<!--
 The mean population size of Northern Pikeminnow in the Deadwater Slough during the fall season was estimated to be `r prettyNum(unique(round(impact_df$N[impact_df$season == "Fall"])), big.mark = ",")` and during the spring 2021 to be `r prettyNum(unique(round(impact_df$N[impact_df$season == "Spring"])), big.mark = ",")`.
 Those estimates translate to linear densities of `r prettyNum(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms), big.mark = ",")`
-<!-- need units  -->
 which is `r round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / john_day_dens)` and `r round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / john_day_dens)` times greater, respectively, than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r as.character(english(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / low_columbia_fish_km)))` and `r as.character(english(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / low_columbia_fish_km)))` times greater than projections for the lower Columbia River [@Beamesderfer1996].
@@
 @Widener2021 has identified a need for more current information on piscine predator populations in the Columbia River.
-
-## Mark-Recapture Models
-
-All mark-recapture abundance estimators suggest a large number of Northern Pikeminnow occupying Deadwater Slough.
-Several assumptions in our estimators may have influenced the magnitude of our results.
+-->
+
+Several assumptions in our estimators may have influenced the magnitude of our abundance results.
 First, all models assumed a closed population, however, it is possible that immigration and emigration occurred.
 Nevertheless, our estimators still provide unbiased estimates of abundance assuming the immigration and emigration rates are equal between marked and unmarked fish.
-Given the size of the sampling area, and the limited length of the sampling period (two weeks), that assumption is likely met.
+Given the size of the sampling area, and the short duration of our surveys, that assumption is likely met.
 The assumption of a closed population can also affect results if mortality occurs for some marked fish released back to the population.
-During our field processing, we did not observe mortalities or injury post release; however, the potential for mortality amongst marked and released fish cannot be discounted.
-Next, our estimates assumed that the second (and subsequent) sample(s) are random samples of the population.
-Angling methods may have a size selection bias, thereby limiting our abundance estimates to a size threshold susceptible to angling, and possibly biasing them low.
-We also consider the validity of our assumption of equal catchability of individual fish between sampling events to have introduced potential bias in our results.
-Unequal catchability for individual fish is difficult to assess, especially if being caught once influences the likelihood of recapture.
-During our study, we recaptured two individual fish during multiple surveys indicating that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.-->
-
-We further recognize that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
+During our field processing, no mortalities or injury post release was observed; however, the potential for mortality amongst marked and released fish cannot be discounted.
+Next, our estimates assumed that capture and recapture events are random samples of the population.
+Angling methods may have a size selection bias, thereby limiting our abundance estimates to a size range susceptible to angling, resulting in a conservative estimate of the total population.
+We also question the validity of our assumption of equal catchability of individual fish between sampling events.
+During our study, we recaptured two individual fish during multiple surveys indicating that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.--> However, we suspect that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
 Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit the "good" fishing locales.
-If marked fish were not recently released or mixed back into those areas that would result in marked fish being less susceptible to capture.
 Violations of the assumption of equal catchability between marked and unmarked fish may lead to an overestimate of abundance of Northern Pikeminnow in our study if marked fish have lower catchability.
 This bias will be proportional to the difference in catchability of marked and unmarked fish, so if marked fish are only 70% as likely to be caught as unmarked fish (due to behavior or spatial distribution) the true abundance is closer to 70% of our abundance estimate.
-We calculated this for the Schnabel estimator by assuming that only 70% of the fish marked each sample were returned to the populuation and available for recapture.
 
@@
 The spring abundance estimate was smaller than fall abundance estimates.
-Our approach assumed equal capture probabilities between the fall and spring sample efforts.
-Yet, we suspect that the capture probability in the spring could be lower as higher spring flows reduce Northern Pikeminnow densities biasing our spring abundance estimate low.
+Although our approach assumed equal capture probabilities between the fall and spring surveys, we believe higher spring flows may result in lower capture probability.
 Therefore, we believe our spring abundance estimate to be conservative.
-<!-- how do higher spring flows reduce pikeminnow densities?
-If pikeminnow densities are lower, then the total abundance SHOULD be lower.
-Makes more sense to me to say that higher flows directly lower capture probabilities.
--->
+
+In addition to the large population, the observed PSD (***value***%) in this study was large compared to observations of 41% below Bonneville Dam and 18% in Bonneville Reservoir [@Winther2020].
+This suggests that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to populations reported elsewhere in the Columbia River.
+Notably, the state catch-and-release record Northern Pikeminnow, measuring ***value*** mm TL, was caught in Deadwater Slough during the fall 2020 survey.
 
@@
 We performed gastric lavage on nearly all Northern Pikeminnow collected during this study.
-In addition to juvenile Chinook Salmon, species included Redside Shiner, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*; although decomposition from digestion rendered many fish contents unidentifiable.
-While this approach provided useful information on prey species, we have concerns about using angling and gastric lavage to quantify diet composition in Northern Pikeminnow.
-First, previous studies have indicated that diets of angled fish differ from those collected by other methods (e.g., electrofishing) [@Hodgson1988], and further, angled fish are more likely to have empty stomachs [@Jurajda2016], perhaps because they are hungry and actively searching for food.
-Second, unlike other predators such as Smallmouth Bass and Walleye that are commonly sampled using gastric lavage [@Kamler2001], Northern Pikeminnow are cyprinids and therefore lack a true stomach.
-Previous research indicates that performing gastric lavage on cyprinid species can be ineffective at quantifying diet composition [@Hartleb1995], or alternately, requires a modification in methodology to flush food items out the vent of the fish [@Wasowicz1994].
-Our intent, rather, was to simply document predation on juvenile Chinook Salmon in Deadwater Slough, and given the potential ineffectiveness of gastric lavage using angling and for Northern Pikeminnow, we consider our diet composition estimates to be conservative estimates i.e., we believe that a larger proportion of the Northern Pikeminnow diet is fish and juvenile Chinook Salmon than our gastric results suggest.
+Observed prey species included juvenile Chinook Salmon, species included Redside Shiner *Richardsonius balteatus*, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*, though decomposition from digestion rendered most fish contents unidentifiable.
+Although our approach provided useful information on prey species, we have concerns about using angling and gastric lavage to quantify diet composition in Northern Pikeminnow.
+First, diets of angled fish can differ from those collected by other methods such as electrofishing [@Hodgson1988].
+Further, angled fish are more likely to have empty stomachs [@Jurajda2016], perhaps because they are hungry and actively searching for food.
+Second, unlike other predators such as Smallmouth Bass and Walleye, that are commonly sampled using gastric lavage [@Kamler2001], Northern Pikeminnow are cyprinids and therefore lack a true stomach.
+Consequently, performing gastric lavage on cyprinid species can be ineffective for accurately quantifying diet composition [@Hartleb1995], or alternately, requires a modification in methodology to flush food items out the vent of the fish [@Wasowicz1994].
+Therefore, we consider our diet composition estimates and subsequent consumption estimates to be conservative.
 
 ## Bioenergetics
-<!--
-OUR APPROACH:
-1. g of fish consumed during DSR and NRR peak emigrations
-2. Combine results from #1 for total g consumed by an individual NPM
-3. Parse out the percentage of fish prey that is juvenile Chinook salmon
-  - start w/ conservative scenarios of 10% and 25%
-4. Divide total g consumed by 10.6 g, the average juvenile Chinook Salmon from Upper Salmon MPG, to estimate the # of juvenile Chinook Salmon consumed
-5. Multiply results from #4 by a "stable" NPM population size estimate
-6. Multiply results from #5 by SAR from McCann et al. 2019 for Upper Salmon MPG Chinook
--->
-
-An average-size Northern Pikeminnow from Deadwater Slough is estimated to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g) and `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g) of fish during fall and spring, respectively, to maintain its body weight.
-This equates to `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"]) + median(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g of fish consumed during the DSR and NRR peak emigrations, assuming the median value among diet scenarios (i.e., if 50% of their diet consists of fish).
-Consumption ranged from `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"]) + min(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` for a diet comprised of 10% fish to `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"]) + max(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` for a diet comprised of 90% fish.
 
 Our approach assumed that all available prey consumed by Northern Pikeminnow have energy densities equal to juvenile Chinook Salmon.
-Other salmonid including juvenile steelhead and Sockeye Salmon are likely also available prey in Deadwater Slough, especially during the spring migration.
+Generalizing energy densities using a single species is a common approach [@Petersen1999], however, we recognize that other salmonids including juvenile steelhead and Sockeye Salmon are likely also available prey in Deadwater Slough, especially during the spring migration.
 Additionally, hatchery smolt releases (Chinook Salmon, steelhead, and Sockeye Salmon) are also prevalent in the mainstem Salmon River during the spring migration.
-Other studies have used a similar approach where the energy densities of different salmonids were generalized using an estimate from a single species [@Petersen1999].<!-- The following needs to be updated after we finalize some results --> However, our study is focused on potential impacts to natural-origin DSR and NRR Chinook Salmon, in which case, let's consider a thought exercise where DSR and NRR Chinook Salmon represent only 10% and 25% of total fish prey consumed by Northern Pikemnnow during these periods.
-Under these scenarios, the average Northern Pikeminnow would consume 8.95 g (range 4.52 – 10.05 g) if 10% of prey were natural-origin Chinook Salmon and 22.38 g (range 11.31 – 25.12) if 25%.
-
-Our estimated consumption of Chinook Salmon juveniles by Northern Pikeminnow was calculated using a static predator population.
-If we assume an average juvenile Upper Salmon MPG Chinook Salmon of 10.6 g, this equates to 0.84 juvenile Chinook Salmon (range 0.43 – 0.94) under the 10% scenario and 2.11 juvenile Chinook Salmon (range 1.07 – 2.37) under the 25% scenario.
-Let’s now consider a stable Northern Pikeminnow population size of 10,000 (roughly the mean of fall and spring abundance estimates, divided by two to account for half of marked individuals not being available for recapture).
-In this case, the Northern Pikeminnow population in Deadwater Slough would be estimated to consume 8,400 (range 4,300 – 9,400) or 21,100 (range 10,700 – 23,700) under the 10% and 25% scenarios, respectively.
-
-## Impacts to Adult Returns
-
-<!-- Also update this after finalizing results -->
-Our primary objective was to assess the impact of predation on juvenile Chinook Salmon in the Upper Salmon MPG.
-However, there is also interest in estimating the number of additional adult Chinook Salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough was reduced or eliminated.
-Our approach here was to multiply the total estimated juvenile Chinook Salmon consumed by an estimate of Granite-to-Granite smolt-to-adult survival rate (SAR) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.
-Assuming an SAR of `r round(sar$medsar, 5)` (SE: `r prettyNum(round(sar$sdsar, 5))`) we'll consider the "adult equivalents" to our estimates of juveniles consumed.
-Under this scenario, the consumed juveniles would be equivalent to 51.6 adults (range 26.4 – 57.7) and 129.6 adults (range 65.7 – 145.6).
-However, we believe these 10% and 25% scenarios to be conservative, especially during the fall DSR emigration period.
-During spring, natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, and Sockeye Salmon) are present in the river, and so 10% or 25% may be reasonable.
-However, during fall, DSR Chinook Salmon are likely the dominant prey species available to Chinook Salmon in Deadwater Slough as 1) no (or at least minimal) hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent than in Chinook Salmon.
-Note that this thought exercise only includes the Northern Pikeminnow population in Deadwater Slough even though they are present throughout the upper Salmon River, although likely at lower densities than in Deadwater Slough, and also ignores potential impacts to ESA-listed steelhead and Sockeye Salmon that are also a portion of the Northern Pikeminnow diet.
-<!--
-Additional thoughts here:
-*	SAR results. Those results are GRA - GRA SARs, estimated from multiple years of PIT tag data. Seem pretty good to me.
-* Maybe a plot of multiple scenarios
--->
+To address the prevalence of other fish in the Northern Pikeminnow's diet, we considered scenarios where DSR and NRR Chinook Salmon represent only 30%, 50%, and 65% of total fish prey consumed.
+These values may be considered conservative during the emigration periods when Chinook Salmon are abundant and the Northern Pikeminnow diet shifts almost entirely to piscivory, presumably on juvenile Chinook Salmon [@Shively1996; @Poe1991].
+This is likely the case during the fall when DSR Chinook Salmon are the dominant prey species available in Deadwater Slough as 1) no (or at least minimal) hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent that in Chinook Salmon.
+During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river <50% Chinook Salmon in their diet may be more likely.
+Future work to quantify and identify the juvenile salmonid in Deadwater Slough during their seasonal migration would be useful to validate the diet composition assumptions used in our model.
 
@@
 
-Although not formally assessed in this study, avian predators including Great Blue Herons *Ardea herodias* and Bald Eagles *Haliaeetus leucocephalus* are another potential source of mortality for juvenile Chinook Salmon and other ESA-listed salmonids in the Deadwater Slough.
+Although not formally assessed in this study, avian predators are another potential source of mortality for juvenile Chinook Salmon and other ESA-listed salmonids in the Deadwater Slough.
 The Deadwater Slough is recognized as an important bird watching and nesting area due to the associated upstream braided channels and riparian and backwater habitats [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
-Several piscivorous bird species have been documented using Deadwater Slough including the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
+Several piscivorous bird species have been documented using Deadwater Slough including the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron *Ardea herodias*, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
 During the initial sampling effort in fall 2019, a two-person crew walked the entire reach including surrounding and upstream areas scanning for passive integrated transponder (PIT) tags that are implanted into thousands of natural- and hatchery-origin salmonids in the Upper Salmon River, annually.
@@
 Avian predation contributes a major component of the total mortality for yearling Chinook Salmon in some locations in the lower Snake and Columbia rivers, particularly at hydroelectric dams and within reservoirs [@Evans2012; -@Evans2016]; however, we did not observe large colonies of piscivorous birds within the study area.
-Although we did observe Belted Kingfishers, mergansers, and Great Blue Herons at Deadwater Slough during field efforts; Great Blue Herons were especially prevalent in the anastomizing reach immediately upriver of the slackwater reach.
+Although we did observe Belted Kingfishers, mergansers, and Great Blue Herons at Deadwater Slough; Great Blue Herons were especially prevalent in the anastomizing reach immediately upriver of the slack water reach.
 Additionally, there is documentation of individual Double-crested Cormorants [@eBird2021] at the Deadwater Slough; however, the site is not within their breeding range, rather, it is part of a migration corridor.
@@
 
+## Potential Impacts to Salmon Populations
+
+Recovery of Chinook Salmon in the Upper Salmon MPG is desired to support local fisheries and economies.
+At least five of the eight populations must meet criteria set forth by @McElhany2000 and the @ICTRT2007 for the MPG to be considered viable and for recovery of the Snake River Evolutionary Significant Unit.
+We estimate that ***value*** juvenile Chinook Salmon will be consumed by Northern Pikeminnow, which translates to ***value*** "adult equivalents" which is approximately ***value***% of contemporary returns to the MPG.
+Considering the diet scenarios used in our study are likely conservative assumptions during the fall DSR emigration, the estimated number of Chinook Salmon consumed is substantial.
+
+Spawner abundance is perhaps the most important metric considered in determining a population's viability.
+We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by ***value*** to ***value*** adults, annually.
+This is between ***value***% and ***value***% of the total mean adult escapement<!--of 558--> reported for the Upper Salmon MPG in 2017 - 2019 [@Kinzer2020].
+In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.
+Accordingly, we surmise that adult returns of steelhead and Sockeye Salmon are affected by predation, including for hatchery populations which provide recreational fishing opportunities.
+Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including ESA-listed species.
+Because the Deadwater Slough is part of the migratory pathway for multiple species and populations of emigrating salmonids, the positive impact could be greater than individual tributary rehabilitation actions which typically benefit a single population.
+
 ## Management Implications
 
-<!-- Again, some of the following numbers need to be updated after results are formalized -->
-Recovery of Chinook Salmon in the Upper Salmon MPG is desired to support local fisheries and economies.
-At least five of the eight populations must meet criteria set forth by @McElhany2000 and @ICTRT2007 for the MPG to be considered viable and for recovery of the Snake River Evolutionary Significant Unit.
-Spawner abundance is perhaps the most important metric considered in determining a population's viability.
-We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by 26.4 to 145.6 adults, annually.
-This is between 4.7% and 26.1% of the total mean adult escapement<!--of 558--> reported for the Upper Salmon River MPG in 2017 - 2019 [@Kinzer2020].
-These estimates only take into account predation on juvenile Chinook Salmon by Northern Pikeminnow, but presumably, juvenile steelhead and Sockeye Salmon are also prey items in Deadwater Slough, including hatchery releases.
-Therefore, adult returns of steelhead and Sockeye Salmon are affected as well, including for hatchery populations which provide recreational fishing opportunities.
-Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including multiple ESA-listed species.
-Because the Deadwater Slough is part of the migratory pathway for multiple species and populations of emigrating salmonids, the impact of that benefit could be higher than tributary rehabilitation actions which typically benefit a single population.
-Moreover, the deepened, slack water conditions that favor Northern Pikeminnow at Deadwater Slough are a result of the failure of a manmade mining reservoir dam.
-Taken together, the Deadwater Slough is a favorable candidate for management or restoration actions to benefit local Chinook Salmon populations.
-
-We envision two potential management actions at reducing predation at Deadwater Slough: 1) removing the Dump Creek alluvial fan and 2) a local Northern Pikeminnow bounty program to encourage harvest in Deadwater Slough aimed at reducing the predator population size.
-Both actions have evident pros and cons.
-Removing the Dump Creek alluvial fan has the benefit of restoring the Salmon River to a more natural flow (e.g., increased water velocity, decreased depth) and bathymetry (e.g., heterogeneous substrate) that likely existed in the reach prior to the failure of the mining reservoir dam in Dump Creek.
-Restoring flow would presumably reduce Northern Pikeminnow densities to levels similar to upstream and downstream reaches where recent studies identified higher survival and transition probabilities than those at Deadwater Slough (@Axel2015; @Ackerman2018; @Porter2019).
-However, we do not know the feasibility of removing or alleviating the Dump Creek alluvial fan, and if it could be removed or altered, future disturbance events could potentially move additional loose sediment in the Dump Creek drainage, thereby creating a new alluvial fan.
-Additionally, restoring the flow would likely diminish the value of the reach as an important bird watching and nesting area.
-
-Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere (*citation*).
-A local bounty program in the Salmon River has the potential to boost the local economy (to an unknown degree) which has been negatively impacted by low returns of Chinook Salmon, steelhead, and Sockeye Salmon in recent decades.
-Many local businesses rely on license and gear sales for local sport fisheries, but low salmon and steelhead returns have eliminated or reduced the duration of fisheries in some years, and a local Northern Pikeminnow bounty program could allow the local economy to recoup some of those lost revenues.
-On the downside, bounty programs require continual effort and harvest to suppress the population size and so would require years of continued funding and local buy-in to remain effective.
-If funding or effort was alleviated, Northern Pikeminnow would likely recover to a higher population size without additional alterations to the habitat.
+The Deadwater Slough is a favorable candidate for management or restoration actions to benefit the local Chinook Salmon populations.
+Two potential management actions could reduce predation at Deadwater Slough: 1) removing the Dump Creek alluvial fan and 2) a local Northern Pikeminnow bounty program to encourage harvest in Deadwater Slough aimed at reducing the predator population size.
+Removing the Dump Creek alluvial fan has the benefit of restoring the natural fluvial processes in the Salmon River that likely existed in the reach prior to the failure of the mining reservoir dam in Dump Creek.
+Restoring natural processes could reduce Northern Pikeminnow densities to levels similar to upstream and downstream reaches where higher survival and transition probabilities have been observed compared to Deadwater Slough [@Axel2015; @Ackerman2018; @Porter2019].
+Managers ought also to consider the feasibility and net benefit of restoring fluvial processes relative to potential losses for recreational fishing and bird watching opportunities.
+
+Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere (@Winther2019).
+A local Northern Pikeminnow bounty program could provide monetary incentive for sportfisherman, in addition to boosting the local economy from lost revenues.
+Bounty programs could be conducted year-round or seasonally to coincide with peak juvenile outmigration(s).
+Each approach would require continual/annual effort and harvest to suppress the population size and as such, appropriate funding would be needed to support the program.
+
 <!--
 What about adding cover? It could potentially provide refugia for juvenile salmonids, but could make recreational angling more difficult
-
-Is there anything in the literature that discusses habitat preferences for pike minnow?
-If we speed up velocities and add some cover potentially pike minnow predation success will be lowered.
-Will these fish just move elsewhere?
 -->
@@
 
-The failure of a mining reservoir dam in the late 19th century created an alluvial fan at the confluence of the Salmon River and Dump Creek which has backed up the Salmon River and created an almost 2-kilometer section of unnaturally slow and deep river, resulting in favorable conditions for Northern Pikeminnow.
-We estimate that greater than 10,000 Northern Pikeminnow occupy Deadwater Slough with densities greater than reported elsewhere.
-Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids emigrating through Deadwater Slough including Chinook Salmon i.e., juvenile Chinook Salmon have slower movement rates and little refugia to avoid predation.
-We estimate that predation from Northern Pikeminnow in Deadwater Slough may reduce adult returns by as much as 146 or 26% of recent returns to the Upper Salmon River MPG.
-Therefore, predation in Deadwater Slough likely has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
+The failure of a mining reservoir dam in the late 19th century created an alluvial fan at the confluence of the Salmon River and Dump Creek which has backed up the Salmon River and created favorable conditions for Northern Pikeminnow.
+We estimate that more than ***value*** Northern Pikeminnow occupy Deadwater Slough, resulting in a remarkable density given the size of Deadwater Slough.
+Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids that are rearing in or emigrating through Deadwater Slough, including Chinook Salmon.
+We estimate that predation from Northern Pikeminnow in Deadwater Slough may reduce adult returns by as much as 146 or 26%, and therefore, has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
 

### 2022-01-31 analysis/paper/paper.Rmd
fixed some table captions

@@
 
-```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow using different mark-recapture estimators and for the fall sampling events.
-Single census estimators are shown as circles; multiple census estimators are shown as triangles.
-Note the y-axis is truncated at 100 and the upper confidence interval for the Schumacker-Eschmeyer estimator in 2020 extends beyond that."}
+```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow using different mark-recapture estimators and for the fall sampling events."}
 
@@
   coord_cartesian(ylim = c(0, 100)) +
-  scale_shape_manual(values = c(21, 19), name="Event") +
+  scale_shape_manual(values = c(21, 19), name="Survey") +
   #scale_color_manual(values = c("grey35", "darkred"), name="Census type") +
@@
 
-```{r bio-p, fig.cap = "Cumulative amount of fish (g) consumed by an individual Northern Pikeminnow during the peak DSR (fall) and NRR (spring) Chinook Salmon emigrations (top panel) and corresponding daily mean water temperatures (bottom panel).
-The black line indicates the diet scenario comprised of 50% fish prey and the gray area indicates the range among diet scenarios (10 - 90% of Northern Pikeminnow diet consists of fish).
+```{r bio-p, fig.cap = "The cumulative fish (g) pre consumed by an individual Northern Pikeminnow during the peak DSR (fall) and NRR (spring) Chinook Salmon emigrations (top panel) and corresponding daily mean water temperatures (bottom panel).
+The black line is the median among diet scenarios (60% of the Northern Pikeminnow diet consists of fish) and the gray area shows the range between diet scenarios of 30 and 90% fish prey.
 Daily mean water temperatures were summarized from six years of data available from the Shoup gage station.

### 2022-02-01 analysis/paper/paper.Rmd
added some citations
(hopefully) final edits to intro

@@
 One potentially important, but perhaps under-appreciated source of salmonid mortality is predation on emigrating juveniles by piscivorous fishes.
-In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recovery [@Fresh2003; @Winther2020].
+In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of recovery [@Fresh2003; @Winther2020].
 Dams and reservoirs in the Columbia River are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995].
@@
 Currently, the Upper Salmon MPG supports eight independent, extant populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
-The Lemhi River historically contained the largest Chinook Salmon population within the MPG and is an intensively monitored watershed to determine responses to restoration actions [@Zimmerman2012].
+The Lemhi River historically contained the largest Chinook Salmon population within the MPG and is an intensively monitored watershed to measure responses to restoration actions [@Zimmerman2012].
 The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a].
@@
 
-Recent studies have examined the downstream movement, distribution, and apparent (hereafter 'survival') of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach.
-Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have substantially reduced survival in the Deadwater Slough reach compared to adjacent reaches [@Axel2015].
-Similarly, emigrating DSR Chinook Salmon had a 10% reduction in transition probability through the Deadwater Slough compared to surrounding reaches during the fall and early winter months [@Ackerman2018; @Porter2019].
+Recent studies have examined the downstream movement, distribution, and apparent survival (hereafter 'survival') of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach.
+Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have substantially reduced survival in the Deadwater Slough reach [@Axel2015].
+Similarly, emigrating DSR Chinook Salmon had an approximate 10% reduction in transition probability through the Deadwater Slough compared to surrounding reaches during fall and early winter months [@Ackerman2018; @Porter2019].
 Low survival was attributed to low-velocity water delaying movement rates, thereby increasing predation risk.
@@
 We hypothesize that high densities of piscivorous predators in the Deadwater Slough may explain the reduced survival observed for juvenile Chinook Salmon.
-To test this, our objectives were:
-
-1. Estimate the abundance of potential predators in the Deadwater Slough during the peaks of fall DSR and spring NRR juvenile emigrations;
+To test this, our objectives were four-fold:
+
+1. Estimate the abundance of Northern Pikeminnow in the Deadwater Slough during the peaks of fall DSR and spring NRR juvenile emigrations;
 2. Document predation on juvenile Chinook Salmon during the DSR and NRR emigration periods using gastric lavage;
 3. Use a bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population on DSR and NRR Chinook Salmon emigrants at Deadwater Slough;
-4. Estimate how consumption of juvenile Chinook Salmon emigrants at Deadwater Slough by Northern Pikeminnow impacts adult returns to the Upper Salmon MPG.
+4. Estimate how consumption of juvenile Chinook Salmon emigrants at Deadwater Slough by Northern Pikeminnow may impact adult returns to the Upper Salmon MPG.
 
@@
 
-Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere (@Winther2019).
+Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere (@Winther2020).
 A local Northern Pikeminnow bounty program could provide monetary incentive for sportfisherman, in addition to boosting the local economy from lost revenues.

### 2022-02-01 analysis/paper/paper.Rmd
near final methods edits

@@
 
-The Deadwater Slough is an approximately `r round(ddw_rkms, 1)` km section of the mainstem Salmon River located roughly 6 river kilometers (RKM) downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
+The Deadwater Slough is an approximately `r round(ddw_rkms, 1)` km section of the mainstem Salmon River located roughly 6 river kilometers (rkm) downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)).
 The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
@@
 We estimated the population size of piscine predators in the Deadwater Slough using a mark-recapture survey design and a catch per unit effort (CPUE) approach.
-Predators were sampled near the peaks of the juvenile DSR and NRR emigrations.
+Predators were sampled near the peaks of the fall DSR and spring NRR emigrations.
 During our initial survey in 2019, Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and consequently became our focal taxa.
-Northern Pikeminnow were surveyed near the peaks of the fall DSR and spring NRR emigrations.
 The initial intent was to survey during the emigration of the 2018 brood year (fall 2019 and spring 2020), however logistical constraints during the onset of the Covid-19 pandemic delayed the spring 2020 survey until fall 2020.
@@
 Consequently, an additional week of surveying was added in spring 2021.
-Fall surveys occurred over two weeks to minimize Northern Pikeminnow immigration/emigration.
+The fall surveys were constrained to two weeks to minimize Northern Pikeminnow immigration/emigration.
 
@@
 Ineffective methods were abandoned in subsequent years.
-<!-- For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) our inability to "chase" fish downstream as they evaded the electricity.
-Snorkeling was ineffective due to the high turbidity of the mainstem Salmon River and Deadwater Slough area.
-Significant amounts of debris in the water column also created concerns that fyke nets could become entangled.
-Additionally, there were concerns that raft electrofishers and fyke nets could incidentally capture ESA-listed adult steelhead that are present in the mainstem Salmon River during fall and spring.
--->After evaluating all methods, angling was the most effective method for capturing Northern Pikeminnow while also minimizing potential impacts to ESA-listed adult steelhead that were present in the Salmon River during our surveys.<!--, but recognizing that angling can have both species and size selectivity (bias).--> Therefore, the following analyses will focus on fish captured by angling, unless otherwise noted.
+After evaluating all methods, angling was the most effective method for capturing Northern Pikeminnow while also minimizing potential impacts to ESA-listed adult steelhead that were present in the Salmon River during our surveys.
+Therefore, the following analyses will focus on fish captured by angling, unless otherwise noted.
 
@@
 Fish that died prior to release were included in the total capture (or recapture) count, but not included in the number of marked fish available for recapture after that date.
-Finally, we recorded the angling start and end time for each crew (person or combinations of persons) to calculate CPUE as number of Northern Pikeminnow caught per angler hour.
+Finally, we recorded the angling start and end time for each crew (person or combinations of persons) to calculate CPUE as the number of Northern Pikeminnow caught per angler hour.
 
@@
 Two types of mark-recapture estimators were used to explore variation in abundance estimates: a single census and a multiple census estimator.
-Estimators fell into two broad categories: single census and multiple census.
-For the single census estimator, we treated the first week of the survey as the mark event and the following week as the recapture event, pooling data within each event week.
-Alternatively, the multiple census estimator treated each day as an event and use information about the total number of marked fish from all previous events to calculate the total abundance.
+For the single census estimator, we treated the first week of the survey as the mark event and the following week as the recapture event, pooling data within each week.
+Alternatively, the multiple census estimator treated each day as an event and used information about the total number of marked fish from all previous events to calculate the total abundance.
 
@@
 
-<!--
-$$
-\hat{N} = \frac{(M)(n)}{(m)}
-$$
--->
 $$
@@
 where $M$, $n$, and $m$ are indexed by each survey (day), $i$.
-With small numbers of recaptures, it may not be appropriate to calculate the standard error of the Schnabel estimator but the 95% confidence intervals can be calculated using a Poisson approximation [@Krebs1999].
-The Schnabel estimator is essentially a weighted average of a series of Chapman-modified Lincoln-Peterson estimators.
+With small numbers of recaptures, it may not be appropriate to calculate the standard error of the Schnabel estimator, but the 95% confidence intervals can be calculated using a Poisson approximation [@Krebs1999].
+The Schnabel estimator is essentially the weighted average of a series of Chapman-modified Lincoln-Peterson estimators.
 
 Because captures of new and marked fish were dependent on volunteer anglers, there was concern that sampling was biased to locations preferred by anglers.
-If marked fish were not recently released or mixed back into those locales, marked fish may have been less susceptible to capture.
-To account for this potential bias, we adjusted the Schnabel estimator with an assumption that marked and released fish during each survey day were not available for recapture for 48 hours.
+If marked fish were not recently released or mixed back into those preferred locales, marked fish may have been less susceptible to capture than unmarked fish.
+To account for this potential bias, we also used an adjusted the Schnabel estimator which assumed that marked and released fish during each survey day were not available for recapture for 48 hours.
+The adjusted Schnabel estimator, which we refer to as the "delayed mixing" Schnabel estimator, would also account for marked fish being "hook shy" for 48 hours following release.
 
@@
 
-<!-- delayed mixing Schnabel equation here; include statement about fish being "hook shy" -->
-
 The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and 2020 surveys.
-However, to estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of total CPUE in the fall survey to the total CPUE in the spring survey, then multiplied that ratio by the average abundance from the fall surveys.
+However, to estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of total CPUE in the fall survey to the total CPUE in the spring survey, then multiplied that ratio by the mean abundance from fall surveys.
 
@@
 This approach assumes equal capture probabilities for fall and spring surveys.
-Fall and spring population estimates were converted into a linear density by dividing by the length of Deadwater Slough (km) measured in GIS software and using drone-generated orthography.
+Fall and spring population estimates were converted into a linear density by dividing by the length of Deadwater Slough (rkm) measured in GIS software and using drone-generated orthography.
 
@@
 For each stomach sample, wet weight (grams) was recorded for the total stomach contents, including all non-fish items (e.g., macroinvertebrates, organic matter), as well as the portion consisting of fish.
-Fish and fish remnants were identified to the lowest taxonomic unit using diagnostic bones, or were categorized as unknown [@Hansel1988; @Frost2000].
+Fish and fish remnants were identified to the lowest taxonomic unit using diagnostic bones [@Hansel1988; @Frost2000] or were categorized as unknown.
 Approximately 5% of Northern Pikeminnow were euthanized for dissection (n = 75) after gastric lavage to validate the efficacy of the methodology.
@@
 Prey energy densities were fixed at 3,000 J/g for invertebrates and 21,500 J/g for juvenile Chinook Salmon [@Moss2016].
-Because were were unable to differentiate juvenile Chinook Salmon from other fish prey, we assume all fish prey have the same energy densities as juvenile Chinook Salmon.
+Because we were unable to differentiate juvenile Chinook Salmon from other fish prey, we assume all fish prey have the same energy densities as juvenile Chinook Salmon.
 The average length of Northern Pikeminnow caught in Deadwater Slough during our study was converted to weight using a weight-length formula from @Parker1995.
@@
 
-This resulted in average spring and fall starting eights of 504.4 and 670.5 g for Northern Pikeminnow with average lengths of `r avg_fall_NPM` and `r avg_spring_NPM` mm, respectively.
+This resulted in average spring and fall starting weights of 504.4 and 670.5 g for Northern Pikeminnow with average lengths of `r avg_fall_NPM` and `r avg_spring_NPM` mm, respectively.
 We assumed no growth in individual Northern Pikeminnow, resulting in equal start and end weights.
-Mean daily water temperatures were summarized from 15-minute interval temperature readings between March 3, 2013 and June 14, 2021 from USGS gage station 13307000, approximately 22 RKM downstream of Deadwater Slough.
+Mean daily water temperatures were summarized from 15-minute interval temperature readings between March 3, 2013 and June 14, 2021 from USGS gage station 13307000, approximately 22 rkm downstream of Deadwater Slough.
 
@@
 These values were supported by observations elsewhere in the Columbia River basin (54-86%: @Shively1996; 48-86%: @Zimmerman1999; 37%: @Gray2001) that suggest fish are the majority of Northern Pikeminnow prey.
-14 models were run (7 diet scenarios and 2 seasons), each providing an estimate of the total grams of fish consumed by an individual Northern Pikeminnow (***table reference***).
-To estimate the total biomass of fish consumed by Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated Northern Pikeminnow population size during fall and spring.
+Fourteen model runs were completed (7 diet scenarios and 2 seasons), each providing an estimate of the total grams of fish consumed by an individual Northern Pikeminnow (***table reference***).
+To estimate the total biomass of fish consumed by Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated Northern Pikeminnow population sizes during fall and spring.
 
@@
 However, there is some evidence that juvenile salmonids are by far the most consumed fish prey [@Zimmerman1999; @Shively1996].
-Moreover, it is estimated that Chinook Salmon make up 64.2%, 29.3%, and 49.3% of the fish prey consumed by Northern Pikeminnow in the Columbia River below Bonneville Dam, Columbia River reservoirs, and in the lower Snake River, respectively [@Zimmerman1999].
-Because the bioenergetics model does not differentiate juvenile Chinook Salmn from other fish prey items, we performed a sensitivity analysis to assess potential impacts of Northern Pikeminnow predation on local Chinook Salmon populations.
-Using values similar to @Zimmerman1999 we modeled three diet scenarios where Chinook Salmon comprised 30%, 50%, and 65% of the fish prey, by weight, in the Northern Pikeminnow diet during the fall DSR and spring NRR emigration periods.
-In this analysis, we assume 60% of the Northern Pikeminnow's total diet is comprised of fish, representing the median in our range of modeled diets.
+Moreover, it is estimated that Chinook Salmon make up 64.2%, 29.3%, and 49.3% of the fish prey consumed by Northern Pikeminnow in the Columbia River below Bonneville Dam, in Columbia River reservoirs, and in the lower Snake River, respectively [@Zimmerman1999].
+Because the bioenergetics model does not differentiate juvenile Chinook Salmon from other fish prey items, we performed a sensitivity analysis to assess potential impacts of Northern Pikeminnow predation on local Chinook Salmon populations.
+Using values similar to @Zimmerman1999, we modeled three diet scenarios where Chinook Salmon comprised 30%, 50%, and 65% of the fish prey, by weight, in the Northern Pikeminnow diet during the fall DSR and spring NRR emigration periods.
+In this analysis, we assume 60% of the Northern Pikeminnow's total diet is comprised of fish, representing the median scenario in our range of modeled diets.
 
@@
 Although the primary impact to Chinook Salmon in Deadwater Slough by Northern Pikeminnow is juvenile predation, the common metric to evaluate salmon recovery is often adult returns.
-We assessed the impacts on adult returns by estimating the number of additional adult Chinook Salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough was reduced or eliminated.
+We assessed the potential impacts on adult returns by estimating the number of additional adult Chinook Salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough was reduced or eliminated.
 This calculation was made by multiplying the total estimated juvenile Chinook Salmon consumed by an estimate of Granite-to-Granite smolt-to-adult survival rate (SAR) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.

### 2022-02-01 analysis/paper/paper.Rmd
adund and stomach results edits

@@
 spring_abund = all_est %>%
-  filter(model == "Schnabel") %>%
+  filter(model == "Schnabel - Delayed Mixing") %>%
   # filter(model == "Chapman") %>%
@@
 
-## Potential Impacts to Chinook Salmon Populations
+## Impacts to Chinook Salmon Populations
 
@@
 Using hook-and-line angling, we caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 surveys, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish.
-The CPUE was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021 (Figure \@ref(fig:cpue-fig)).
+The CPUE was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021.
 During the mark-recapture events, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019, including `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals and `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020, including `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals.
@@
 
-Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` (Table \@ref(tab:abund-tab)).For both estimates, the multiple census estimates were larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (***figure reference***).
-Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, the Schnabel estimator was determined to provide the best estimates of abundance.
-Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`).<!--Need to add the delayed-mixing Schnabel estimate results--> For spring 2021, we estimated abundance by multiplying the mean fall abundance by the ratio between the spring CPUE and mean of the fall CPUEs (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), resulting in an estimate of `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough.
-Those estimates translate to linear densities of ***value*** and ***value*** Northern Pikeminnow per RKM.
-
-<!-- need to adjust the spring estimate -->
-
-The lengths of Northern Pikeminnow captured ranged from ***value*** to ***value*** mm with an average length of ***value*** mm (\@ref(fig:length-freq-p)).
+Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` depending on the estimator used (Table \@ref(tab:abund-tab)).
+For both surveys, the multiple census estimates were generally larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (Figure \@ref(fig:abund-fig)).
+Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, the Schnabel estimates were determined to provide the best estimates of abundance.
+Additionally, due to concerns about marked and released fish not immediately mixing back into preferred angling locales and/or being "hook shy", we used the conservative delayed-mixing Schnabel estimates.
+Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`).
+For spring 2021, we estimated abundance by multiplying the mean fall abundance by the ratio between the spring CPUE and mean fall CPUEs (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), resulting in an estimate of `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough.
+Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm.
+
+The lengths of Northern Pikeminnow captured ranged from `r min(cmr_df$length, na.rm = T)` to `r max(cmr_df$length, na.rm = T)` mm with an average TL of `r round(mean(cmr_df$length, na.rm = T))` mm (Figure \@ref(fig:length-freq-p)).
 The PSD for Northern Pikeminnow in Deadwater Slough across all three efforts was `r prettyNum(psd, digits = 1)`%.
@@
 Northern Pikeminnow captured during spring 2021 had a higher proportion of fish content in their diet compared to individuals captured in the fall surveys (Table \@ref(tab:gastric-tab)).
-The average wet weight of total contents for an individual was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r round(min(stmch_df$stomach_contents_weight), 2)`-`r round(max(stmch_df$stomach_contents_weight), 2)` g).
+The mean wet weight of total contents for an individual was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; sd = `r round(sd(stmch_df$stomach_contents_weight), 2)` g).
 Overall, fish or fish remnants made up `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of all stomach contents examined.
@@
 ```{r adult-imp-tab, echo =TRUE}
-
 impact_tbl = impact_df %>%

### 2022-02-01 analysis/paper/paper.Rmd
cleaned YAML header

@@
 csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
-abstract: |
 keywords: |
   Northern Pikeminnow; Chinook Salmon; predation; mark-recapture; bioenergetics
-highlights: |
-  These are the highlights. 
 ---

### 2022-02-01 analysis/paper/paper.Rmd
added length-weight section to code

@@
 # set some deadwater attributes
-ddw_rkms = 1.871
-ddw_m2 = 119431
-ddw_avg_width_m = 68
+ddw_rkms = 1.871               # length km
+ddw_m2 = 119431                # area square m
+ddw_avg_width_m = 68           # measured avg width
 ddw_acres = ddw_m2 / 4047
@@
 
-```{r capture-data, eval = T}
+```{r capture-data}
 raw_df = readxl::read_excel(here("analysis/data/raw_data/Deadwater_predation_mark_recapture_11.09.21.xlsx")) %>%
@@
 
-```{r effort, eval = T}
+```{r effort}
 # What size (mm) is considered a large pikeminnow?
@@
 
-```{r lincoln-petersen-estimator, eval = T}
+```{r lincoln-petersen-estimator}
 # using Lincoln-Petersen model, combining first 4 and last four days
@@
 ```{r assump-violation}
-# assume that marked fish aren't available for recapture until 2 days after capture
+# assume that marked fish aren't available for recapture until 48 days after release
 # using Schnabel estimator
@@
   arrange(event_name, model) %>%
+  # remove the Petersen and Schumacher-Eschmeyer estimators
   filter(! model %in% c("Petersen",
@@
 
+```{r length-weight}
+# average size of fall and spring pikeminnow
+avg_fall_NPM_TL = cmr_df %>%
+  filter(grepl('Fall', event_name)) %>%
+  select(length) %>%
+  pull() %>%
+  mean(na.rm = T)
+
+avg_spring_NPM_TL = cmr_df %>%
+  filter(grepl('Spring', event_name)) %>%
+  select(length) %>%
+  pull() %>%
+  mean(na.rm = T)
+
+# conversions from Fishbase.org
+#TL = 1.052 * FL
+#TL = 1.087 * FL (12-30cm)
+#TL = 1.114 * FL (5-12cm) don't use
+
+# use mean of coefficients
+# TL = 1.0695 * FL or
+# FL = 0.9350 * TL
+
+# convert to FL
+avg_fall_NPM_FL = avg_fall_NPM_TL * 0.9350
+avg_spring_NPM_FL = avg_spring_NPM_TL * 0.9350
+
+# convert FL to g using Parker (1995) equation
+avg_fall_NPM_g = 10^(-4.886 + (2.986*log10(avg_fall_NPM_FL)))
+avg_spring_NPM_g = 10^(-4.886 + (2.986*log10(avg_spring_NPM_FL)))
+
+# average size of DSR and NRR juveniles
+avg_DSR_g = 10.3
+avg_NRR_g = 10.9
+
+```
+
 ```{r read-bio-outputs, message=FALSE, warning=FALSE, include=FALSE}
@@
 spring_f90 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I10_F90.csv")) %>% fmt_bio_out()
-
-# average size of fall and spring pikeminnow
-avg_fall_NPM   = 352.9
-avg_spring_NPM = 393.7
-
-# average size of DSR and NRR juveniles
-avg_DSR_g = 10.3
-avg_NRR_g = 10.9
 
@@
 Because we were unable to differentiate juvenile Chinook Salmon from other fish prey, we assume all fish prey have the same energy densities as juvenile Chinook Salmon.
-The average length of Northern Pikeminnow caught in Deadwater Slough during our study was converted to weight using a weight-length formula from @Parker1995.
+The average TL of Northern Pikeminnow caught in Deadwater Slough during our study was converted to fork length (FL) and then to weight (grams) using a weight-length formula from @Parker1995.
 
@@
 
-This resulted in average spring and fall starting weights of 504.4 and 670.5 g for Northern Pikeminnow with average lengths of `r avg_fall_NPM` and `r avg_spring_NPM` mm, respectively.
+This resulted in average fall and spring starting weights of `r round(avg_fall_NPM_g, 1)` and `r round(avg_spring_NPM_g, 1)` g for Northern Pikeminnow with average TLs of `r round(avg_fall_NPM_TL, 1)` and `r round(avg_spring_NPM_TL, 1)` mm, respectively.
 We assumed no growth in individual Northern Pikeminnow, resulting in equal start and end weights.
@@
 <!--
-We estimated an individual, average-size (`r avg_fall_NPM` mm) Northern Pikeminnow to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g) of fish during the DSR emigration to maintain their body size, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
-Similarly, we estimated an average-size (`r avg_spring_NPM` mm) Northern Pikeminnow to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
+We estimated an individual, average-size (`r avg_fall_NPM_TL` mm) Northern Pikeminnow to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g) of fish during the DSR emigration to maintain their body size, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
+Similarly, we estimated an average-size (`r avg_spring_NPM_TL` mm) Northern Pikeminnow to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
 Given a mean Northern Pikeminnow population size of `r prettyNum(unique(impact_df$N[impact_df$season == "Fall"]), big.mark = ",")` from fall estimates and `r prettyNum(unique(impact_df$N[impact_df$season == "Spring"]), big.mark = ",")` for spring 2021 and using the median value from diet scenarios (i.e., 50% of their diet consists of fish), we'd estimate that the Northern Pikeminnow population can consume `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 50]), big.mark = ",")` g (range: `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 10]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 10]), big.mark = ",")` - `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 90]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 90]), big.mark = ",")` g) fish during the DSR emigration and an additional `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 50]), big.mark = ",")` g (range: `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 10]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 10]), big.mark = ",")` - `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 90]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 90]), big.mark = ",")` g) during the NRR emigration in Deadwater Slough.

### 2022-02-01 analysis/paper/paper.Rmd
added bioenergetics table

@@
 
-# fall, fish 10% - 90%
-fall_f10 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I90_F10.csv")) %>% fmt_bio_out()
-fall_f20 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I80_F20.csv")) %>% fmt_bio_out()
+# fall, fish 30% - 90%
 fall_f30 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I70_F30.csv")) %>% fmt_bio_out()
@@
 
-# spring, fish 10% - 90%
-spring_f10 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I90_F10.csv")) %>% fmt_bio_out()
-spring_f20 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I80_F20.csv")) %>% fmt_bio_out()
+# spring, fish 30% - 90%
 spring_f30 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I70_F30.csv")) %>% fmt_bio_out()
@@
 # compile fall results
-fall_results = fall_f10 %>%
+fall_results = fall_f30 %>%
   select(date, temperature_c, cum_cons_fish_g) %>%
-  rename(f10_g = cum_cons_fish_g) %>%
-  mutate(f20_g = fall_f20$cum_cons_fish_g,
-         f30_g = fall_f30$cum_cons_fish_g,
-         f40_g = fall_f40$cum_cons_fish_g,
+  rename(f30_g = cum_cons_fish_g) %>%
+  mutate(f40_g = fall_f40$cum_cons_fish_g,
          f50_g = fall_f50$cum_cons_fish_g,
@@
 # compile spring results
-spring_results = spring_f10 %>%
+spring_results = spring_f30 %>%
   select(date, temperature_c, cum_cons_fish_g) %>%
-  rename(f10_g = cum_cons_fish_g) %>%
-  mutate(f20_g = spring_f20$cum_cons_fish_g,
-         f30_g = spring_f30$cum_cons_fish_g,
-         f40_g = spring_f40$cum_cons_fish_g,
+  rename(f30_g = cum_cons_fish_g) %>%
+  mutate(f40_g = spring_f40$cum_cons_fish_g,
          f50_g = spring_f50$cum_cons_fish_g,
@@
 bio_g_summ = bio_results %>%
-  select(season, f10_g:f90_g) %>%
-  pivot_longer(f10_g:f90_g) %>%
+  select(season, f30_g:f90_g) %>%
+  pivot_longer(f30_g:f90_g) %>%
   group_by(season, name) %>%
@@
 rm(fall_results,
-   fall_f10, 
-   fall_f20, 
    fall_f30, 
@@
    spring_results,
-   spring_f10, 
-   spring_f20, 
    spring_f30, 
@@
 These values were supported by observations elsewhere in the Columbia River basin (54-86%: @Shively1996; 48-86%: @Zimmerman1999; 37%: @Gray2001) that suggest fish are the majority of Northern Pikeminnow prey.
-Fourteen model runs were completed (7 diet scenarios and 2 seasons), each providing an estimate of the total grams of fish consumed by an individual Northern Pikeminnow (***table reference***).
+Fourteen model runs were completed (7 diet scenarios and 2 seasons), each providing an estimate of the total grams of fish consumed by an individual Northern Pikeminnow (Table \@ref(tab:bio-tbl)).
 To estimate the total biomass of fish consumed by Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated Northern Pikeminnow population sizes during fall and spring.
@@
 <div id="tables"></div>
+
+```{r bio-tbl}
+bio_tbl = bio_g_summ %>%
+  mutate(`Date Range` = season) %>%
+  select(season,
+         `Date Range`,
+         name,
+         cumm_g) %>%
+  mutate(name = str_sub(name, 2, 3)) %>%
+  rename(Season = season,
+         `% Fish in Diet` = name,
+         `g Consumed` = cumm_g) %>%
+  mutate(`Date Range` = recode(`Date Range`,
+                               Fall = "Sep 15 - Nov 30",
+                               Spring = "Mar 1 - May 31")) %>%
+  kable(booktabs = T,
+        caption = "Bioenergetics model runs to estimate the grams of fish consumed during the DSR fall and NRR spring Chinook Salmon emigrations.
+The date ranges for each model run, the assumed proportion of fish in the diet, and the total grams of fish consumed by an individual Northern Pikeminnow to maintain its body weight for a given season are shown.
+The Northern Pikeminnow start/end weights were 598.7 and 430.7 g for the fall and spring seasons, respectively.") %>%
+  kable_styling()
+
+bio_tbl
+
+```
+
+\newpage
 
@@
 pFish = bio_results %>%
-  ggplot(aes(ymd, f50_g)) +
+  ggplot(aes(ymd, f60_g)) +
   geom_line(size = 1.1) +
   geom_ribbon(data = bio_results,
-              aes(ymd,ymin = f10_g,
+              aes(ymd,ymin = f30_g,
                   ymax = f90_g),

### 2022-02-01 analysis/paper/paper.Rmd
cleaned up demographics results

@@
 spring_abund = all_est %>%
-  filter(model == "Schnabel - Delayed Mixing") %>%
-  # filter(model == "Chapman") %>%
+  filter(model == "Schnabel") %>%
+  #filter(model == "Schnabel - Delayed Mixing") %>%
   select(-range, -prop_range) %>%
@@
 The average weights of DSR and NRR emigrants were calculated from fish captured at seven rotary screw traps upstream of Deadwater Slough during the fall and spring periods.
-
 Although the primary impact to Chinook Salmon in Deadwater Slough by Northern Pikeminnow is juvenile predation, the common metric to evaluate salmon recovery is often adult returns.
@@
                 as.integer)) %>%
-  crossing(p_chnk = c(.01, .10, .30, .50, .65, .90, 1)) %>%
+  filter(perc_diet == 60) %>%
+  crossing(p_chnk = c(.30, .50, .65)) %>%
   mutate(g_chnk = g_fish_cons * p_chnk) %>%
@@
 For both surveys, the multiple census estimates were generally larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (Figure \@ref(fig:abund-fig)).
-Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, the Schnabel estimates were determined to provide the best estimates of abundance.
-Additionally, due to concerns about marked and released fish not immediately mixing back into preferred angling locales and/or being "hook shy", we used the conservative delayed-mixing Schnabel estimates.
-Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`).
-For spring 2021, we estimated abundance by multiplying the mean fall abundance by the ratio between the spring CPUE and mean fall CPUEs (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), resulting in an estimate of `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough.
+Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, we deemed the Schnabel estimates most appropriate.
+Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`).
+If we assume that marked and released fish are unavailable for recapture for 48 hours following release to address concerns about delayed mixing into preferred angling locales and/or released fish being "hook shy", the mean fall abundance estimate was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r prettyNum(mean(all_est$Lci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` - `r prettyNum(mean(all_est$Uci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")`).
+
+For spring 2021, we estimated abundance by multiplying the mean fall abundance (standard Schnabel estimates) by the ratio between the spring CPUE and mean fall CPUEs (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), resulting in an estimate of `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough.
 Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm.
 
-The lengths of Northern Pikeminnow captured ranged from `r min(cmr_df$length, na.rm = T)` to `r max(cmr_df$length, na.rm = T)` mm with an average TL of `r round(mean(cmr_df$length, na.rm = T))` mm (Figure \@ref(fig:length-freq-p)).
+The lengths of Northern Pikeminnow captured ranged from `r min(cmr_df$length, na.rm = T)` to `r max(cmr_df$length, na.rm = T)` mm TL with an average of `r round(mean(cmr_df$length, na.rm = T))` mm (Figure \@ref(fig:length-freq-p)).
 The PSD for Northern Pikeminnow in Deadwater Slough across all three efforts was `r prettyNum(psd, digits = 1)`%.

### 2022-02-01 analysis/paper/paper.Rmd
cleaned up some references

@@
                    sum),
-            .groups = "drop")
+            .groups = "drop") %>%
+  select(perc_diet,
+         p_chnk,
+         juv_N,
+         juv_Lci,
+         juv_Uci,
+         adult_N,
+         adult_Lci,
+         adult_Uci)
 
@@
 
-During the fall DSR emigration, we estimated an average-size Northern Pikeminnow (***value*** mm) to consume ***value*** g of fish to maintain their body size, assuming 60% of their diet consists of fish prey (Figure \@ref(fig:bio-p)).
-Fall consumption ranged from ***value*** g for a diet of 30% fish to ***value*** g for a diet of 90% fish.
-During the spring NRR emigration, we estimated an average-size Northern Pikeminnow (***value***) to consume ***value*** 44.29 g of fish, assuming 60% of their diet is fish prey (Figure \@ref(fig:bio-p)).
-Spring consumption ranged from ***value*** g for a diet of 30% fish to ***value*** g for a diet of 90% fish.
-Given estimated mean Northern Pikeminnow population sizes of ***value*** from fall estimates and ***value*** for spring 2021, and assuming 60% of their diet is fish, we estimate that Northern Pikeminnow can consume ***value*** metric tons of fish during the fall DSR emigration and an additional ***value*** metric tons during the spring NRR emigration in Deadwater Slough.
-In total, the Northern Pikeminnow population can consume ***value*** metric tons of fish during the DSR and NRR emigrations to maintain their population size, assuming 60% of their diet consists of fish.
-
-<!--
-We estimated an individual, average-size (`r avg_fall_NPM_TL` mm) Northern Pikeminnow to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g) of fish during the DSR emigration to maintain their body size, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)).
-Similarly, we estimated an average-size (`r avg_spring_NPM_TL` mm) Northern Pikeminnow to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)).
-Given a mean Northern Pikeminnow population size of `r prettyNum(unique(impact_df$N[impact_df$season == "Fall"]), big.mark = ",")` from fall estimates and `r prettyNum(unique(impact_df$N[impact_df$season == "Spring"]), big.mark = ",")` for spring 2021 and using the median value from diet scenarios (i.e., 50% of their diet consists of fish), we'd estimate that the Northern Pikeminnow population can consume `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 50]), big.mark = ",")` g (range: `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 10]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 10]), big.mark = ",")` - `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 90]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 90]), big.mark = ",")` g) fish during the DSR emigration and an additional `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 50]), big.mark = ",")` g (range: `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 10]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 10]), big.mark = ",")` - `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 90]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 90]), big.mark = ",")` g) during the NRR emigration in Deadwater Slough.
-In total, the Northern Pikeminnow population can consume `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 50]) + unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 50]), big.mark = ",")` g of fish during the DSR and NRR emigrations to maintain their population size if we assume 50% of their diet consists of fish.
--->
+```{r tons-consumption}
+tot_g_DSR = (bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"] * mean(all_est$N[all_est$model == "Schnabel"])) / 1000000
+tot_g_NRR = (bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f60_g"] * round(spring_abund$N_S)) / 1000000
+
+```
+
+During the fall DSR emigration, we estimate an average-size Northern Pikeminnow (`r round(avg_fall_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"]` g of fish to maintain their body size, assuming 60% of their diet consists of fish prey (Figure \@ref(fig:bio-p)).
+Fall consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f30_g"]` g for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f90_g"]` g for a diet of 90% fish.
+During the spring NRR emigration, we estimated an average-size Northern Pikeminnow (`r round(avg_spring_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f60_g"]` g of fish, assuming 60% of their diet is fish prey (Figure \@ref(fig:bio-p)).
+Spring consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f30_g"]` for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f90_g"]` for a diet of 90% fish.
+Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` from fall estimates and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` for spring 2021, and assuming 60% of their diet is fish, we estimate that Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and an additional `r round(tot_g_NRR, 1)` metric tons during the NRR emigration in Deadwater Slough.
 
@@
 
-Using the total g of fish consumed by the Northern Pikeminnow population (summing fall and spring surveys) we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where juvenile Chinook Salmn make up 30%, 50%, and 65% of total fish prey consumed (Figure ***reference***).
-Assuming that 50% of the total fish consumed are juvenile Chinook Salmon, we estimate between (30% fish scenario with CIs) and (90% fish scenario with CIs) juvenile Chinook Salmon will be consumed.
-We then used our estimates of total juvenile Chinook Salmon consumed to estimate "adult equivalents" using the Granite-to-Granite smolt-to-adult survival rate of ***value*** (SE: ***value***; Table ***reference***).
-Assuming predation in Deadwater Slough was reduced or eliminated, the juvenile consumption would be equivalent to ***value*** adults in the spring (range ***value*** - ***value***) and ***value*** adults (range ***value*** - ***value***) in the fall, totaling ***value*** adult Chinook Salmon.
+Using the total fish consumed by the Northern Pikeminnow population (summing fall and spring surveys), we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where juvenile Chinook Salmon make up 30%, 50%, and 65% of total fish prey consumed (Figure \@ref(fig:impact-p)).
+Assuming that 50% of the total fish consumed are juvenile Chinook salmon, we estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon will be consumed.
+We then used our estimate of total juvenile Chinook Salmon consumed to estimate an "adult equivalents" using the median Granite-to-Granite SAR of `r round(sar$medsar, 5)` (sd = `r format(round(sar$sdsar, 5), scientific = F)`) from @McCann2019.
+Assuming predation was non-existent in Deadwater Slough, the juvenile consumption would be equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$p_chnk == 0.50]), big.mark = ",")`).
 
@@
 We estimate that predation from Northern Pikeminnow in Deadwater Slough may reduce adult returns by as much as 146 or 26%, and therefore, has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
+
+\newpage
 

### 2022-02-01 analysis/paper/paper.Rmd
update length freq plot

@@
   kable(booktabs = T,
-        caption = "Bioenergetics model runs to estimate the grams of fish consumed during the DSR fall and NRR spring Chinook Salmon emigrations.
+        caption = "Bioenergetics model runs to estimate the grams of fish consumed during the fall DSR and spring NRR Chinook Salmon emigrations.
 The date ranges for each model run, the assumed proportion of fish in the diet, and the total grams of fish consumed by an individual Northern Pikeminnow to maintain its body weight for a given season are shown.
-The Northern Pikeminnow start/end weights were 598.7 and 430.7 g for the fall and spring seasons, respectively.") %>%
+Northern Pikeminnow start and end weights were 598.7 and 430.7 g for the fall and spring seasons, respectively.") %>%
   kable_styling()
@@
                 ~ str_replace(., "_", " "))) %>%
-  rename(`Effort` = event_name) %>%
+  rename(`Survey` = event_name) %>%
   kable(booktabs = T,
         format.args = list(big.mark = ","),
-        caption = "Summary of Northern Pikeminnow angling survey data for single-census mark-recapture estimators.
+        caption = "Summary of Northern Pikeminnow angling survey data for the single-census mark-recapture estimator.
 M is the total number of fish caught in the first sample (week), marked and returned to the population; n is the total number of fish caught in the second sample (week), and m is the number of previously marked fish caught in the second week.") %>%
@@
   select(event_name, date, M, n, m, u, R) %>%
-  rename(`Effort` = event_name,
+  rename(`Survey` = event_name,
          Date = date) %>%
@@
         format.args = list(big.mark = ","),
-        caption = "Summary of Northern Pikeminnow angling survey data for multiple-census mark-recapture estimators.
-M is the total number of fish caught in previous samples (days), marked and returned to the population; n is the total number of fish caught in the current sample (day); m is the number of marked fish caught in that sample; u is the number of unmarked Northern Pikeminnow caught, and R is the number of marked fish returned to the population.
+        caption = "Summary of Northern Pikeminnow angling survey data for standard Schnabel estimator.
+M is the total number of fish caught in previous samples (days), marked and returned to the population; n is the total number of fish caught in the current sample (day); m is the number of marked fish caught in that sample; u is the number of unmarked fish caught, and R is the number of marked fish returned to the population.
 Dates of sampling are also provided.") %>%
@@
          -Lci, -Uci) %>%
-  rename(`Effort` = event_name,
+  rename(`Survey` = event_name,
          Estimator = model,
-         `95% Confidence Interval` = CI) %>%
+         `95% CI` = CI) %>%
          # `L CI` = Lci,
@@
 
-```{r spring-N, eval = F}
-spring_abund %>%
-  rowwise() %>%
-  mutate(`Avg. Fall N (CI)` = paste0(prettyNum(round(N_F), big.mark = ","),
-                                     " (",
-                                     paste0(prettyNum(round(c(Lci_F, Uci_F)), big.mark = ","),
-                                            collapse = ", "),
-                                     ")"),
-         `Spring N (CI)` = paste0(prettyNum(round(N_S), big.mark = ","),
-                                     " (",
-                                     paste0(prettyNum(round(c(Lci_S, Uci_S)), big.mark = ","),
-                                            collapse = ", "),
-                                     ")")) %>%
-  ungroup() %>%
-  select(Species = species,
-         `Avg Fall CPUE` = cpue_F,
-         `Spring CPUE` = cpue_S,
-         ends_with("(CI)")) %>%
-  kable(booktabs = T,
-        caption = "Average CPUE and abundance in the fall (95% confidence interval), and CPUE and estimates of spring abundance (95% confidence interval) for Northern Pikeminnow in Deadwater Slough.") %>%
-  kable_styling()
-
-```
-
 \newpage
@@
 gastric_tbl %>%
-  rename(Effort = event_name,
+  rename(Survey = event_name,
          Captured = captured,
@@
 
-\newpage
-
-```{r adult-imp-tab, echo =TRUE}
-impact_tbl = impact_df %>%
-  select(season,
-         perc_diet,
+```{r adult-imp-tab}
+impact_tbl = total_impacts %>%
+  select(perc_diet,
          p_chnk,
@@
          Uci = juv_Uci) %>%
-  group_by(perc_diet, p_chnk) %>%
-  summarise(across(c("N", "Lci", "Uci"),sum),
-            .groups = "drop") %>%
-  mutate(Stage = "Juveniles") %>%
   mutate(p_chnk = p_chnk * 100) %>%
   rename(perc_chnk = p_chnk) %>%
-  filter(perc_chnk %in% c(30, 50, 65))%>%
-
-  bind_rows(impact_df %>%
-              select(season,
-                     perc_diet,
+  mutate(Stage = "Juveniles") %>%
+  bind_rows(total_impacts %>%
+              select(perc_diet,
                      p_chnk,
@@
                      Uci = adult_Uci) %>%
-              group_by(perc_diet, p_chnk) %>%
-              summarise(across(c("N", "Lci", "Uci"),
-                   sum),
-             .groups = "drop") %>%
-             mutate(Stage = "Adult Equivalents") %>%
               mutate(p_chnk = p_chnk * 100) %>%
               rename(perc_chnk = p_chnk) %>%
-  filter(perc_chnk %in% c(30, 50, 65))) %>%
+              mutate(Stage = "Adult Equivalents")) %>%
+  select(Stage,
+         `% Chinook` = perc_chnk,
+         N,
+         `95% L CI` = Lci,
+         `95% U CI` = Uci) %>% 
+  mutate(across(where(is.numeric), round)) %>%
   kable(booktabs = T,
         format.args = list(big.mark = ","),
-        caption = "Estimates of juvenile fish consumed and adult equivalents for diet scenarios comprised of 10-90% fish.
-Estimates and 95% confidence intervals are calculated using Northern Pikeminnow abundance estimates of 28,144 (95% CI: 14,378 - 59,978) for fall and 14,897 (95% CI: 7,610 - 31,747) for spring.
-'Adult equivalents' are estimated using a Lower Granite Dam smolt-to-adult rate (SAR) for Chinook Salmon from McCann et al.
-2019.") %>%
-  kable_styling(font_size = 10)
+        caption = "Estimates of juvenile Chinook Salmon consumed and adult equivalents for scenario where 60% of Northern Pikeminnow diet is fish, the median of scenarios evaluated.
+Table shows scenarios where 30%, 50%, and 65% of total fish consumed is Chinook Salmon.
+Estimates and 95% confidence intervals are calculated using Northern Pikeminnow abundance estimates of 27,874 (95% CI: 14,244 - 59,388) for fall and 14,799 (95% CI: 7,562 - 31,530) for spring.
+'Adult equivalents' are estimated using a Lower Granite Dam smolt-to-adult rate (SAR) for upper Salmon River Chinook Salmon from McCann et al.
+(2019).") %>%
+  kable_styling()
 
@@
 ```
+
 \newpage
@@
 
-```{r cpue-fig, fig.cap = "The number of Northern Pikeminnow caught per angler effort for each sample effort.", eval = FALSE}
+```{r cpue-fig, fig.cap = "The number of Northern Pikeminnow caught per angler effort for each sample effort.", eval = F}
 total_cpue %>%
@@
 
-\newpage
-
-```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow using different mark-recapture estimators and for the fall sampling events."}
+```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow in Deadwater Slough using different mark-recapture estimators and for the fall sampling events."}
 
@@
     #facet_wrap(~ event_name) +
-  coord_cartesian(ylim = c(0, 100)) +
+  coord_cartesian(ylim = c(0, 80)) +
   scale_shape_manual(values = c(21, 19), name="Survey") +
@@
   scale_x_discrete(labels=c(pm_est_labels))+
-  labs(x = "Estimator",
-       y = "Abundance (x 1,000)") +
+  labs(y = "Abundance (x 1,000)") +
   #scale_size(guide = "none")+
@@
         legend.key.size = unit(.5, "cm"),legend.box = 'horizontal',
-        #axis.title.x = element_blank(),
+        axis.title.x = element_blank(),
         axis.text.x = element_text(size = 11, color = "black"),
@@
 length_freq_p = gastric_df %>%
-  ggplot(aes(length)) +
-  geom_histogram(breaks = seq(150, 650, by = 50),
+  select(event_name,
+         length) %>%
+  mutate(season = event_name) %>%
+  mutate(season = recode(season,
+                         `Fall 2019` = "Fall",
+                         `Fall 2020` = "Fall",
+                         `Spring 2021` = "Spring"))
+
+ggplot(length_freq_p, 
+       aes(x = length, 
+           fill = season)) +
+  geom_histogram(breaks = seq(150, 650, by = 25),
                  col = "black",
-                 fill = "gray30",
-                 alpha = .2) +
+                 alpha = 0.5) +
+  scale_fill_manual(name = "Season",
+                    values = c("gray", "black"),
+                    labels = c("Fall", "Spring")) +
   labs(x = "Total Length (mm)",
-       y = "Number of Northern Pikeminnow") +
-  xlim(c(100, 700)) +
-  ylim(c(0, 800)) +
-  scale_x_continuous(breaks = seq(0, 650, by = 50)) +
-  scale_y_continuous(breaks = seq(0, 550, by = 50)) +
+       y = "Frequency") +
   theme(axis.text.x = element_text(size = 11,
@@
         axis.text.y = element_text(size = 11,
-                                   color = "black"))
-
-length_freq_p
+                                   color = "black"),
+        legend.position = c(.15, .85))
 

### 2022-02-01 analysis/paper/paper.Rmd
error in impacts results

@@
                 as.integer)) %>%
-  filter(perc_diet == 60) %>%
+  #filter(perc_diet == 60) %>%
   crossing(p_chnk = c(.30, .50, .65)) %>%
@@
 Using the total fish consumed by the Northern Pikeminnow population (summing fall and spring surveys), we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where juvenile Chinook Salmon make up 30%, 50%, and 65% of total fish prey consumed (Figure \@ref(fig:impact-p)).
-Assuming that 50% of the total fish consumed are juvenile Chinook salmon, we estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon will be consumed.
+Assuming that 50% of the total fish consumed are juvenile Chinook salmon, we estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon will be consumed.
 We then used our estimate of total juvenile Chinook Salmon consumed to estimate an "adult equivalents" using the median Granite-to-Granite SAR of `r round(sar$medsar, 5)` (sd = `r format(round(sar$sdsar, 5), scientific = F)`) from @McCann2019.
-Assuming predation was non-existent in Deadwater Slough, the juvenile consumption would be equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$p_chnk == 0.50]), big.mark = ",")`).
+Assuming predation was non-existent in Deadwater Slough, the juvenile consumption would be equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`).
 
@@
 
-\newpage
-
-```{r adult-imp, eval = F}
-impact_df %>%
-  select(season,
-         perc_diet,
-         juv_chnk,
-         smolt_N,
-         smolt_Lci,
-         smolt_Uci,
-         medsar,
-         sdsar,
-         adult_N,
-         adult_Lci,
-         adult_Uci) %>%
-  mutate(juv_chnk = round(juv_chnk, 2),
-         smolt_N = round(smolt_N, 0),
-         smolt_Lci = round(smolt_Lci, 0),
-         smolt_Uci = round(smolt_Uci, 0),
-         medsar = round(medsar, 5),
-         sdsar = round(sdsar, 5),
-         adult_N = round(adult_N, 0),
-         adult_Lci = round(adult_Lci, 0),
-         adult_Uci = round(adult_Uci, 0)) %>%
-  rename(Season = season,
-         `p Fish` = perc_diet,
-         `Fish Per NPM` = juv_chnk,
-         `Fish N` = smolt_N,
-         `Fish L CI` = smolt_Lci,
-         `Fish U CI` = smolt_Uci,
-         `Median SAR` = medsar,
-         `SD SAR` = sdsar,
-         `Adult N` = adult_N,
-         `Adult L CI` = adult_Lci,
-         `Adult U CI` = adult_Uci) %>%
-  kable(booktabs = T,
-        format.args = list(big.mark = ","),
-        caption = "Estimates of the number of fish consumed for each season and diet scenario.
-The diet scenario indicates the proportion of the Northern Pikeminnow diet consisting of fish.
-The 'Fish Per NPM' column indicates the number of fish an individual Northern Pikeminnow to consume during the fall or spring seasons assuming an average weight of 10.3 (DSR) and 10.9 (NRR) grams for juvenile Chinook Salmon and is taken from each of the 18 bioenergetics model runs.
-The number of fish consumed (and 95% confidence intervals) are calculated using Northern Pikeminnow abundance estimates of 28,144 (95% CI: 14,378 - 59,978) for fall and 14,897 (95% CI: 7,610 - 31,747) for spring.
-The number of 'adult equivalents' is also estimated using a Lower Granite Dam smolt-to-adult rate (SAR) for Chinook Salmon from McCann et al.
-2019.") %>%
-  kable_styling(font_size = 8)
-
-```
-
-```{r adult-imp-tab}
+```{r adult-imp-tab, eval = F}
 impact_tbl = total_impacts %>%
@@
 all_est %>%
-
   mutate(across(c(N, Lci, Uci),
@@
 
-```{r adult-imp-p, fig.cap = "The estimated number of DSR (fall) and NRR (spring) emigrants consumed by Northern Pikeminnow in Deadwater Slough (top panel) across a range of diet scenarios where 10-90% of the diet consists of juvenile Chinook salmon.
-The bottom panel provides the adult equivalents assuming a Lower Granite Dam smolt-to-adult rate (SAR) from McCann et al.
-2019."}
-#impact_p = impact_df %>%
-#  select(season,
-#         perc_diet,
-#         p_chnk,
-#         N = juv_N,
-#         Lci = juv_Lci,
-#         Uci = juv_Uci) %>%
-#  group_by(perc_diet, p_chnk) %>%
-#  summarise(across(c("N", "Lci", "Uci"),sum),
-#            .groups = "drop") %>%
-#  mutate(Stage = "Juveniles") %>%
-#  bind_rows(impact_df %>%
-#              select(season,
-#                     perc_diet,
-#                     p_chnk,
-#                     N = adult_N,
-#                     Lci = adult_Lci,
-#                     Uci = adult_Uci) %>%
-#              group_by(perc_diet, p_chnk) %>%
-#              summarise(across(c("N", "Lci", "Uci"),
-#                   sum),
-#             .groups = "drop") %>%
-#             mutate(Stage = "Adult Equivalents")) %>%
-#  mutate(p_chnk = p_chnk * 100) %>%
-#  rename(perc_chnk = p_chnk) %>%
-#  filter(perc_chnk %in% c(10, 30, 50)) %>%
-#  ggplot(aes(x = perc_diet,y = N)) +
-#  geom_errorbar(aes(ymin = Lci,
-#                    ymax = Uci),
-#                width = 0.1) +
-#  geom_point(size = 3) +
-#  facet_grid(vars(Stage),
-#             vars(perc_chnk),
-#             scales = "free_y") +
-#  scale_color_viridis_d(name = "Percent\nFish in Diet") +
-#  scale_y_continuous(labels = scales::comma) +
-#  scale_x_continuous(breaks = seq(0, 100, by = 10)) +
-#  theme(#axis.text.x = element_blank(),
-        #legend.position = "bottom",
-        #axis.title.x = element_blank(),
-        #axis.ticks.x = element_blank(),
-#        axis.text.y = element_text(size = 11,
-#                                   color = "black"),
-#        strip.text.x = element_text(size = 11,
-#                                    color = "black"),
-#        strip.text.y = element_text(size = 11,
-#                                    color = "black")) +
-#  labs(y = "Number Chinook Consumed")
-
-#impact_p
-
-impact_p = impact_df %>%
+```{r juv-con-p, fig.cap = "test"}
+juv_cons_p = impact_df %>%
   select(season,
@@
          Uci = juv_Uci) %>%
-  group_by(perc_diet, p_chnk) %>%
+  group_by(perc_diet,
+           p_chnk) %>%
   summarise(across(c("N", "Lci", "Uci"),sum),
@@
   rename(perc_chnk = p_chnk) %>%
-  filter(perc_chnk %in% c(30, 50, 65)) %>%
-  ggplot(aes(x = perc_diet,y = N, group=as.factor(perc_chnk), color=as.factor(perc_chnk))) +
-  geom_line(size = 1.1)+
+  ggplot(aes(x = perc_diet,
+             y = N,
+             group = as.factor(perc_chnk),
+             color = as.factor(perc_chnk))) +
+  geom_point(size = 2) +
+  geom_line(size = 1.1) +
   scale_y_continuous(labels = scales::comma) +
   scale_x_continuous(breaks = seq(0, 100, by = 10)) +
-  scale_color_manual(name="% Juvenile Chinook \nof total fish consumed", values=c("lightsteelblue","steelblue3","navyblue"))+
+  scale_color_manual(name = "Percent Chinook", 
+                     values=c("lightsteelblue","steelblue3","navyblue")) +
   theme(axis.text.y = element_text(size = 11,
@@
                                     color = "black"),
-        legend.position = c(.19, .82),legend.background = element_blank(),
-        legend.key.size = unit(.56, "cm"),legend.box = 'horizontal') +
-  labs(y = "Estimated Juvenile Chinook Consumed", x="Percent Fish in Pikeminnow Diet")
-
-impact_p
-
-# impact_df %>%
-#   mutate(stage = "Juv. Emigrant") %>%
-#   select(stage,
-#          season,
-#          perc_diet,
-#          starts_with("smolt")) %>%
-#   rename_with(.fn = ~ str_remove(., "smolt_"),
-#               .cols = starts_with("smolt")) %>%
-#   bind_rows(impact_df %>%
-#               mutate(stage = "Adult Equivalent") %>%
-#               select(stage,
-#                      season,
-#                      perc_diet,
-#                      starts_with("adult")) %>%
-#               rename_with(.fn = ~ str_remove(., "adult_"),
-#                           .cols = starts_with("adult"))) %>%
-# ggplot(aes(x = perc_diet,
-#            y = N,
-#            color = as.factor(perc_diet))) +
-# geom_errorbar(aes(ymin = Lci,
-#                   ymax = Uci),
-#               width = 0.1) +
-# geom_point(size = 4) +
-# facet_grid(fct_rev(stage) ~ season,
-#            scales = "free_y") +
-# scale_color_viridis_d(name = "Percent\nFish in Diet") +
-# theme(axis.text.x = element_blank(),
-#       legend.position = "bottom",
-#       axis.title.x = element_blank(),
-#       axis.ticks.x = element_blank(),
-#       axis.text.y = element_text(size = 11,
-#                                  color = "black"),
-#       strip.text.x = element_text(size = 11,
-#                                   color = "black"),
-#       strip.text.y = element_text(size = 11,
-#                                   color = "black")) +
-# labs(y = "Number Chinook Consumed")
+        legend.position = c(.19, .89),
+        legend.background = element_blank(),
+        legend.key.size = unit(.56, "cm"),
+        legend.box = 'horizontal') +
+  labs(y = "Estimated Juvenile Chinook Consumed", 
+       x = "Percent Fish in Pikeminnow Diet")
+
+juv_cons_p
+
+```
+
+\newpage
+
+```{r adult-imp-p, fig.cap = "temp"}
+#adult impacts
+adult_imp_p = total_impacts %>%
+  filter(perc_diet == 60) %>%
+  select(p_chnk,
+         adult_N,
+         adult_Lci,
+         adult_Uci) %>%
+  mutate(perc_chnk = p_chnk * 100) %>%
+  select(-p_chnk) %>%
+  ggplot(aes(x = as.factor(perc_chnk),
+             y = adult_N)) +
+  geom_point(size = 3,
+             aes(shape = as.factor(perc_chnk))) +
+  geom_errorbar(aes(ymin = adult_Lci,
+                    ymax = adult_Uci),
+                width = 0.1) +
+  labs(y = "Adult Equivalents",
+       shape = "Percent\nChinook") +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank(),
+        axis.text.y = element_text(size = 11,
+                                   color = "black"),
+        legend.position = c(.2, .8),
+        legend.box = "horizontal")
+
+adult_imp_p
 

### 2022-02-02 analysis/paper/paper.Rmd
some discussion edits

@@
 All mark-recapture abundance estimators suggest a large population of Northern Pikeminnow occupy Deadwater Slough.
-The mean population size of Northern Pikeminnow in the Deadwater Slough was estimated to be ***value*** during the fall DSR emigration and ***value*** during the spring NRR emigration.
-Those estimates translate to densities (fish/km) of Northern Pikeminnow that are ***value*** and ***value*** times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately ***value*** and ***value*** times greater than projections for the lower Columbia River [@Beamesderfer1996].
+The mean population size of Northern Pikeminnow was estimated to be `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` during the fall DSR emigration and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` during the spring NRR emigration.
+Those estimates translate to linear densities (fish/rkm) that are `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_rkms) / john_day_dens)` and `r round((spring_abund$N_S / ddw_rkms) / john_day_dens)` times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_rkms) / low_columbia_fish_km)` and `r round((spring_abund$N_S / ddw_rkms) / low_columbia_fish_km)` times greater than projections for the lower Columbia River [@Beamesderfer1996].
+Our areal densities (fish/100 $m^2$) for Deadwater Slough are `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_m2) * 100, 1)` and `r round((spring_abund$N_S / ddw_m2) * 100, 1)` for fall and spring, respectively.
 @Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective; therefore it is possible that their estimates may have been low.
@@
 
-<!--
-The mean population size of Northern Pikeminnow in the Deadwater Slough during the fall season was estimated to be `r prettyNum(unique(round(impact_df$N[impact_df$season == "Fall"])), big.mark = ",")` and during the spring 2021 to be `r prettyNum(unique(round(impact_df$N[impact_df$season == "Spring"])), big.mark = ",")`.
-Those estimates translate to linear densities of `r prettyNum(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms), big.mark = ",")`
-which is `r round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / john_day_dens)` and `r round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / john_day_dens)` times greater, respectively, than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r as.character(english(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / low_columbia_fish_km)))` and `r as.character(english(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / low_columbia_fish_km)))` times greater than projections for the lower Columbia River [@Beamesderfer1996].
-Although estimated densities for Deadwater Slough are high given the relative size (volume) of the Salmon River at the slough relative to the Columbia River, high CPUE combined with low recapture rates (one in every `r round(nrow(cmr_df) / sum(cmr_df$capture_type == "Recapture", na.rm = T))` captured fish were recaptures) suggest this area may support exceptionally high predator densities.
-Further, @Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective.
-Conversely, at Deadwater Slough the maximum depth was approximately 10-15 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach.
-Our estimates of Northern Pikeminnow areal density are also higher than those reported elsewhere in the Columbia River basin.
-We estimate the density of Northern Pikeminnow in Deadwater Slough, as fish per hectare, to be `r prettyNum(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_hectares), big.mark = ",")` for fall and `r prettyNum(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_hectares), big.mark = ",")` for spring.
-For comparison, the only other available estimates of areal density, that the authors are aware of, are 4.4 fish per hectare for John Day Reservoir [@Beamesderfer1991] and 15 fish per hectare in Lake Washington [@Bartoo1977].
-Given the dearth of current data available in the Columbia River basin, it is unclear how Northern Pikeminnow abundance estimates in this study compare to elsewhere in the Columbia River Basin.
-@Widener2021 has identified a need for more current information on piscine predator populations in the Columbia River.
--->
-
 Several assumptions in our estimators may have influenced the magnitude of our abundance results.
@@
 Given the size of the sampling area, and the short duration of our surveys, that assumption is likely met.
+We also caught one individual that ws confirmed tagged in Deadwater Slough in a previous year suggesting that at least some individuals remain in the reach or at least migrate to Deadwater Slough multiple times, annually.
 The assumption of a closed population can also affect results if mortality occurs for some marked fish released back to the population.
-During our field processing, no mortalities or injury post release was observed; however, the potential for mortality amongst marked and released fish cannot be discounted.
+During field processing, no mortalities or injury post release was observed; however, the potential for mortality amongst marked and released fish cannot be discounted.
 Next, our estimates assumed that capture and recapture events are random samples of the population.
@@
 We also question the validity of our assumption of equal catchability of individual fish between sampling events.
-During our study, we recaptured two individual fish during multiple surveys indicating that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.--> However, we suspect that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
+During our study, we recaptured two individual fish during multiple days within a survey indicating that marked fish were still susceptible to angling.
+We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.
+However, we suspect that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
 Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit the "good" fishing locales.
 Violations of the assumption of equal catchability between marked and unmarked fish may lead to an overestimate of abundance of Northern Pikeminnow in our study if marked fish have lower catchability.
-This bias will be proportional to the difference in catchability of marked and unmarked fish, so if marked fish are only 70% as likely to be caught as unmarked fish (due to behavior or spatial distribution) the true abundance is closer to 70% of our abundance estimate.
-
-<!--
-All mark-recapture abundance estimators suggest a large number of Northern Pikeminnow occupying Deadwater Slough.
-Estimators assumed: (1) the population is closed (no immigration, emigration, births, or deaths during the sampling period), (2) all fish have equal chance of being caught in the second (and subsequent) samples(s), (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or mistaken.
-If the population is actually open (potential immigration and/or emigration during the survey period), our estimators should still provide unbiased estimates of abundance as long as the immigration and emigration rates are equal between marked and unmarked fish.
-Given the size of the sampling area, and the limited length of the sampling period (two weeks), we are comfortable with the assumption of no immigration or emigration.
-Assumption (3) indicates whether the second (and subsequent) surveys are random samples of the population.
-Angling methods may provide a size selection bias, but that only affects the interpretation of the abundance estimates (i.e., our estimates are the abundance of Northern Pikeminnow susceptible to angling above some size threshold).
-Unequal catchability between individual fish is very difficult to assess, especially if that heterogeneity arises from being caught once; however, we did recapture two individual fish more than once suggesting that marked fish were still susceptible to angling.
-We are also comfortable with assumptions 4 and 5 (no loss of marks and no mistaken or missed marks) given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.
-
-We do consider two potential violations of assumption, both which reduces the probability that a marked fish is susceptible to capture in the second (and subsequent) samples.
-First, the assumption of a closed population could be violated if some marked fish that are released back to the population die.
-We don't believe this case to be likely, as fish that were dead or nearly so after processing were not released back to the population (nor recorded as such), and those that were released were hand-released only if they were able to swim away on their own volition; however, the potential for marked and released fish to have died at a higher rate than unmarked fish should be acknowledged.
-The second, and more likely, reason is that marked and released fish did not mix thoroughly back into the population (at least not at a fast enough rate), reducing their probability of capture in subsequent captures, thus violating assumption (2).
-We attempted to release marked fish throughout Deadwater Slough to facilitate mixing back into the population; however, anecdotally volunteer anglers would congregate to particular areas of the slough either to socialize with other anglers, because the fishing was good, etc.
-and if marked and released fish were not readily available in those areas that would make marked fish less susceptible to capture.
-This violation would, in effect, reduce the number of marked fish available for recapture or reduce the marking fraction in some areas of the slough.
-Violations of assumption (2) would lead to an overestimate of abundance of Northern Pikeminnow which is a real possibility in our study.
--->
+This bias will be proportional to the difference in catchability of marked and unmarked fish.
+As an example, if marked fish are only 70% as likely to be caught as unmarked fish (due to behavior or spatial distribution) the true abundance is closer to 70% of our abundance estimate.
 
@@
 
-In addition to the large population, the observed PSD (***value***%) in this study was large compared to observations of 41% below Bonneville Dam and 18% in Bonneville Reservoir [@Winther2020].
+In addition to the large population, the observed PSD (`r round(psd)`%) in this study was large compared to observations of 41% below Bonneville Dam and 18% in Bonneville Reservoir [@Winther2020].
 This suggests that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to populations reported elsewhere in the Columbia River.
-Notably, the state catch-and-release record Northern Pikeminnow, measuring ***value*** mm TL, was caught in Deadwater Slough during the fall 2020 survey.
+Notably, the state catch-and-release record Northern Pikeminnow, measuring `r max(gastric_df$length, na.rm = T)` mm TL, was caught in Deadwater Slough during the fall 2020 survey.
 
@@
 At least five of the eight populations must meet criteria set forth by @McElhany2000 and the @ICTRT2007 for the MPG to be considered viable and for recovery of the Snake River Evolutionary Significant Unit.
-We estimate that ***value*** juvenile Chinook Salmon will be consumed by Northern Pikeminnow, which translates to ***value*** "adult equivalents" which is approximately ***value***% of contemporary returns to the MPG.
-Considering the diet scenarios used in our study are likely conservative assumptions during the fall DSR emigration, the estimated number of Chinook Salmon consumed is substantial.
+We estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` juvenile Chinook Salmon may be consumed by Northern Pikeminnow assuming that 60% of their diet is fish and 50% of that is Chinook Salmon.
+Considering that that diet scenarios used in our study are likely conservative assumptions, especially during the fall DSR emigration, the estimated number of Chinook Salmon consumed is substantial.
 
 Spawner abundance is perhaps the most important metric considered in determining a population's viability.
-We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by ***value*** to ***value*** adults, annually.
-This is between ***value***% and ***value***% of the total mean adult escapement<!--of 558--> reported for the Upper Salmon MPG in 2017 - 2019 [@Kinzer2020].
+We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by `r round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` (95% CI: `r round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` - `r round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])`) adults, annually.
+This is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean adult escapement estimated<!--of 558--> for the Upper Salmon MPG in 2017 - 2019 [@Kinzer2020].
+Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of that is Chinook Salmon we estimate the "adult equivalents" of the juveniles consumed to be `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` (95% CI: `r round(total_impacts$adult_Lci[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` - `r round(total_impacts$adult_Uci[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])`) which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% or recent natural-origin adult escapements.
 In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.
@@
 Each approach would require continual/annual effort and harvest to suppress the population size and as such, appropriate funding would be needed to support the program.
-
-<!--
-What about adding cover? It could potentially provide refugia for juvenile salmonids, but could make recreational angling more difficult
--->
 

### 2022-02-02 analysis/paper/paper.Rmd
discussion edits

@@
 This is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean adult escapement estimated<!--of 558--> for the Upper Salmon MPG in 2017 - 2019 [@Kinzer2020].
-Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of that is Chinook Salmon we estimate the "adult equivalents" of the juveniles consumed to be `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` (95% CI: `r round(total_impacts$adult_Lci[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` - `r round(total_impacts$adult_Uci[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])`) which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% or recent natural-origin adult escapements.
+Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of that is Chinook Salmon we estimate the "adult equivalents" of the juveniles consumed to be `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` (95% CI: `r round(total_impacts$adult_Lci[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` - `r round(total_impacts$adult_Uci[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])`) which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% of recent natural-origin adult escapements.
 In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.

### 2022-02-02 analysis/paper/paper.Rmd
complete draft

@@
 
-Predation of emigrating juvenile salmonids by piscivorous fishes is perhaps an under-appreciated source of mortality hindering population recovery in the Columbia River basin.
-Reservoirs and tailraces associated with dams on the Snake and Columbia rivers have created favorable conditions for fishes known to consume juvenile salmonids.
-The native Northern Pikeminnow are one of these fishes and have been estimated to consume ~8% of the approximately 200 million juvenile salmonids emigrating throu the lower Snake and Columbia rivers, annually.
+Predation of emigrating juvenile salmonids by piscivorous fishes is perhaps an under-appreciated source of mortality hindinger population recovery efforts in the Columbia River basin.
+Reservoirs and tailraces associated with dams on the Snake and Columbia river have created favorable conditions for fishes known to consume juvenile salmonids.
+The native Northern Pikeminnow are one of these fished and have been estimate to consume ~8% of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually.
 Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above these reservoir systems.
-In this study, we examine predation by Northern Pikeminnow on Chinook Salmon in a slow-water reach of the Salmon River, Idaho known as Deadwater Slough.
-The reach was (at least partially) formed by the failure of a small mining dam in a nearby drainage during the late 19th century and contains historically important overwinter rearing habitat is a migration corridor for Chinook Salmon and other Endangered Species Act-listed species.
-We estimate the abundance of Northern Pikeminnow during peak emigrations to be ***insert numbers here***.
-Using a bioenergetics approach we estimate the consumption potential of the Northern Pikeminnow population to be ***key results here***.
-This resulted in the potential consumption of ***numbers*** Chinook Salmon emigrants.
-Using a sensitivity analysis, we evaluated the potential impact to adult returns of local Chinook Salmon populations to be between ***number with CIs***.
-It is likely that predation of juvenile Chinook Salmon in Deadwater Slough has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River major population group, thereby hindering local recovery efforts.
+In this study, we examine the abundance of Northern Pikeminnow in a slow-water reach of the Salmon River, Idaho, known as Deadwater Slough, and consider potential impacts of them on local Chinook Salmon populations.
+The reach was (at least partially) formed by the failure of a small mining dam in a nearby drainage during the late 19th century and contains historically important overwinter rearing habitat and is a migration corridor for Chinook Salmon and other Endangered Species Act-listed species including steelhead and Sockeye Salmon.
+We estimated the abundance of Northern Pikeminnow in Deadwater Slough to be 27,874 in the fall and 19,499 in the spring during the peak emigrations of juvenile Chinook Salmon.
+Assuming a scenario where 60% of the Northern Pikeminnow diet is fish, of which 50% is Chinook Salmon, we estimated the Northern Pikeminnow population to consume 95,120 juvenile Chinook Salmon, annually.
+In the most conservative scenario where 30% of diet is fish and 30% of fish are juvenile Chinook Salmon 47,079 juveniles are consumed.
+Using a sensitivity analysis, we further considered potential impacts to adult returns to local populations.
+Given the size of the Salmon River, the estimated Northern Pikeminnow densities for Deadwater Slough are remarkable and predation from the population likely has consequential impacts on Chinook Salmon in the Upper Salmon River major population group, thereby hindering local recovery efforts.
 
@@
 The failure of a mining reservoir dam in the late 19th century created an alluvial fan at the confluence of the Salmon River and Dump Creek which has backed up the Salmon River and created favorable conditions for Northern Pikeminnow.
-We estimate that more than ***value*** Northern Pikeminnow occupy Deadwater Slough, resulting in a remarkable density given the size of Deadwater Slough.
+Even under the most conservative scenarios, we estimate more than 10,000 Northern Pikeminnow occupy Deadwater Slough, resulting in a remarkable density given the size of Deadwater Slough.
 Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids that are rearing in or emigrating through Deadwater Slough, including Chinook Salmon.
-We estimate that predation from Northern Pikeminnow in Deadwater Slough may reduce adult returns by as much as 146 or 26%, and therefore, has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
+We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than 24,000 juvenile Chinook salmon (the lower 95% CI of the most conservative scenario evaluated).
+It is more likely that this value is closer to 100,000, annually, creating a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
 
@@
 We further appreciate the administrative support and guidance from Caitlin Alcott and Inter-Fluve.
-This manuscript benefitted from a review by Sean Gibbs.
+This manuscript benefited from a review by Sean Gibbs.
 Funding for this study was provided by the Bureau of Reclamation, Pacific Northwest Regional Office (contract No.

### 2022-02-02 analysis/paper/paper.Rmd
added statement for GitHub repo

@@
 
-Predation of emigrating juvenile salmonids by piscivorous fishes is perhaps an under-appreciated source of mortality hindinger population recovery efforts in the Columbia River basin.
+Predation of emigrating juvenile salmonids by piscivorous fishes is perhaps an under-appreciated source of mortality hindering population recovery efforts in the Columbia River basin.
 Reservoirs and tailraces associated with dams on the Snake and Columbia river have created favorable conditions for fishes known to consume juvenile salmonids.
-The native Northern Pikeminnow are one of these fished and have been estimate to consume ~8% of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually.
+The native Northern Pikeminnow are one of these fishes and have been estimated to consume ~8% of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually.
 Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above these reservoir systems.
 In this study, we examine the abundance of Northern Pikeminnow in a slow-water reach of the Salmon River, Idaho, known as Deadwater Slough, and consider potential impacts of them on local Chinook Salmon populations.
-The reach was (at least partially) formed by the failure of a small mining dam in a nearby drainage during the late 19th century and contains historically important overwinter rearing habitat and is a migration corridor for Chinook Salmon and other Endangered Species Act-listed species including steelhead and Sockeye Salmon.
+The reach was (at least partially) formed by the failure of a small mining dam in a nearby drainage during the late 19th century and contains historically important overwinter rearing habitat and is a migration corridor for Chinook Salmon as well as other Endangered Species Act-listed species including steelhead and Sockeye Salmon.
 We estimated the abundance of Northern Pikeminnow in Deadwater Slough to be 27,874 in the fall and 19,499 in the spring during the peak emigrations of juvenile Chinook Salmon.
 Assuming a scenario where 60% of the Northern Pikeminnow diet is fish, of which 50% is Chinook Salmon, we estimated the Northern Pikeminnow population to consume 95,120 juvenile Chinook Salmon, annually.
-In the most conservative scenario where 30% of diet is fish and 30% of fish are juvenile Chinook Salmon 47,079 juveniles are consumed.
+In the most conservative scenario where 30% of diet is fish and 30% of fish are juvenile Chinook Salmon, 47,079 juveniles are consumed.
 Using a sensitivity analysis, we further considered potential impacts to adult returns to local populations.
@@
 If marked fish were not recently released or mixed back into those preferred locales, marked fish may have been less susceptible to capture than unmarked fish.
-To account for this potential bias, we also used an adjusted the Schnabel estimator which assumed that marked and released fish during each survey day were not available for recapture for 48 hours.
+To account for this potential bias, we also used an adjusted Schnabel estimator, which assumed that marked and released fish during each survey day were not available for recapture for 48 hours.
 The adjusted Schnabel estimator, which we refer to as the "delayed mixing" Schnabel estimator, would also account for marked fish being "hook shy" for 48 hours following release.
@@
 
+All data and code for the analyses presented here can be found in a GitHub repository at ***zenodo DOI*** here.
+
 # Results
@@
 We then used our estimate of total juvenile Chinook Salmon consumed to estimate an "adult equivalents" using the median Granite-to-Granite SAR of `r round(sar$medsar, 5)` (sd = `r format(round(sar$sdsar, 5), scientific = F)`) from @McCann2019.
-Assuming predation was non-existent in Deadwater Slough, the juvenile consumption would be equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`).
+Assuming predation was non-existent in Deadwater Slough, the juvenile consumption would be equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) adults.
 
@@
 Given the size of the sampling area, and the short duration of our surveys, that assumption is likely met.
-We also caught one individual that ws confirmed tagged in Deadwater Slough in a previous year suggesting that at least some individuals remain in the reach or at least migrate to Deadwater Slough multiple times, annually.
+We also caught one individual that was confirmed tagged in Deadwater Slough in a previous year suggesting that at least some individuals remain in the reach or at least migrate to Deadwater Slough multiple times, annually.
 The assumption of a closed population can also affect results if mortality occurs for some marked fish released back to the population.
@@
 We performed gastric lavage on nearly all Northern Pikeminnow collected during this study.
-Observed prey species included juvenile Chinook Salmon, species included Redside Shiner *Richardsonius balteatus*, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*, though decomposition from digestion rendered most fish contents unidentifiable.
+Observed prey species included juvenile Chinook Salmon as well as Redside Shiner *Richardsonius balteatus*, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*, though decomposition from digestion rendered most fish contents unidentifiable.
 Although our approach provided useful information on prey species, we have concerns about using angling and gastric lavage to quantify diet composition in Northern Pikeminnow.
@@
 This is likely the case during the fall when DSR Chinook Salmon are the dominant prey species available in Deadwater Slough as 1) no (or at least minimal) hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent that in Chinook Salmon.
-During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river <50% Chinook Salmon in their diet may be more likely.
+During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river, <50% Chinook Salmon in their diet may be more likely.
 Future work to quantify and identify the juvenile salmonid in Deadwater Slough during their seasonal migration would be useful to validate the diet composition assumptions used in our model.
@@
 We estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` juvenile Chinook Salmon may be consumed by Northern Pikeminnow assuming that 60% of their diet is fish and 50% of that is Chinook Salmon.
-Considering that that diet scenarios used in our study are likely conservative assumptions, especially during the fall DSR emigration, the estimated number of Chinook Salmon consumed is substantial.
+Considering that diet scenarios used in our study are likely conservative assumptions, especially during the fall DSR emigration, the estimated number of Chinook Salmon consumed is substantial.
 
@@
 In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.
-Accordingly, we surmise that adult returns of steelhead and Sockeye Salmon are affected by predation, including for hatchery populations which provide recreational fishing opportunities.
-Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including ESA-listed species.
+Accordingly, we surmise that adult returns of steelhead and Sockeye Salmon are also affected by predation, including hatchery populations, which provide recreational fishing opportunities.
+Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including other ESA-listed species.
 Because the Deadwater Slough is part of the migratory pathway for multiple species and populations of emigrating salmonids, the positive impact could be greater than individual tributary rehabilitation actions which typically benefit a single population.
@@
 Restoring natural processes could reduce Northern Pikeminnow densities to levels similar to upstream and downstream reaches where higher survival and transition probabilities have been observed compared to Deadwater Slough [@Axel2015; @Ackerman2018; @Porter2019].
-Managers ought also to consider the feasibility and net benefit of restoring fluvial processes relative to potential losses for recreational fishing and bird watching opportunities.
+Managers ought also consider the feasibility and net benefit of restoring fluvial processes relative to potential losses for recreational fishing and bird watching opportunities.
 
@@
 The failure of a mining reservoir dam in the late 19th century created an alluvial fan at the confluence of the Salmon River and Dump Creek which has backed up the Salmon River and created favorable conditions for Northern Pikeminnow.
-Even under the most conservative scenarios, we estimate more than 10,000 Northern Pikeminnow occupy Deadwater Slough, resulting in a remarkable density given the size of Deadwater Slough.
+Even under the most conservative scenarios, we estimate more than 10,000 Northern Pikeminnow occupy Deadwater Slough, resulting in a remarkable density given the size of the area.
 Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids that are rearing in or emigrating through Deadwater Slough, including Chinook Salmon.

### 2022-02-03 analysis/paper/paper.Rmd
latest edits from TB

@@
 The Salmon River was historically the most productive tributary for Chinook Salmon in the Columbia River basin [@Nemeth1999], but those populations have become depleted in recent decades.
-Currently, the Upper Salmon MPG supports eight independent, extant populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
-The Lemhi River historically contained the largest Chinook Salmon population within the MPG and is an intensively monitored watershed to measure responses to restoration actions [@Zimmerman2012].
+Currently, the Upper Salmon MPG supports eight independent, extant populations; this includes Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River populations [@NOAA2017].
+The Lemhi River historically contained the largest Chinook Salmon population within the MPG and is therefore of particular interest for restoration actions [@Zimmerman2012].
 The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a].
@@
 Methods included raft electrofishing, fyke netting, snorkeling, and angling.
-Ineffective methods were abandoned in subsequent years.
 After evaluating all methods, angling was the most effective method for capturing Northern Pikeminnow while also minimizing potential impacts to ESA-listed adult steelhead that were present in the Salmon River during our surveys.
-Therefore, the following analyses will focus on fish captured by angling, unless otherwise noted.
+Ineffective methods were abandoned in subsequent years and the following analyses will focus on fish captured by angling, unless otherwise noted.
 
@@
 where $M$, $n$, and $m$ are indexed by each survey (day), $i$.
-With small numbers of recaptures, it may not be appropriate to calculate the standard error of the Schnabel estimator, but the 95% confidence intervals can be calculated using a Poisson approximation [@Krebs1999].
-The Schnabel estimator is essentially the weighted average of a series of Chapman-modified Lincoln-Peterson estimators.
+With small numbers of recaptures, it may not be appropriate to calculate the standard error of the Schnabel estimator, but the 95% confidence intervals can be calculated using a Poisson approximation [@Krebs1999].<!--The Schnabel estimator is essentially the weighted average of a series of Chapman-modified Lincoln-Peterson estimators.-->
 
@@
 To account for this potential bias, we also used an adjusted Schnabel estimator, which assumed that marked and released fish during each survey day were not available for recapture for 48 hours.
-The adjusted Schnabel estimator, which we refer to as the "delayed mixing" Schnabel estimator, would also account for marked fish being "hook shy" for 48 hours following release.
+The adjusted Schnabel estimator, referred to as the "delayed mixing" Schnabel estimator, is given by:
 
@@
 This approach assumes equal capture probabilities for fall and spring surveys.
-Fall and spring population estimates were converted into a linear density by dividing by the length of Deadwater Slough (rkm) measured in GIS software and using drone-generated orthography.
+Fall and spring population estimates were converted into linear and areal densities by dividing by the length and area, respectively, of Deadwater Slough measured in GIS software and using drone-generated orthography.
 
@@
 Although the primary impact to Chinook Salmon in Deadwater Slough by Northern Pikeminnow is juvenile predation, the common metric to evaluate salmon recovery is often adult returns.
-We assessed the potential impacts on adult returns by estimating the number of additional adult Chinook Salmon that might be expected to return to Lower Granite Dam if predation in the Deadwater Slough was reduced or eliminated.
-This calculation was made by multiplying the total estimated juvenile Chinook Salmon consumed by an estimate of Granite-to-Granite smolt-to-adult survival rate (SAR) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.
+We assessed the potential impacts on adult returns by estimating the number of additional adult Chinook Salmon expected to return to Lower Granite Dam if predation was reduced or eliminated in the Deadwater Slough.
+This calculation was made by multiplying the total estimated juvenile Chinook Salmon consumed by the median Granite-to-Granite SAR of 0.00614 (sd = 0.00051) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.
 
@@
 Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`).
-If we assume that marked and released fish are unavailable for recapture for 48 hours following release to address concerns about delayed mixing into preferred angling locales and/or released fish being "hook shy", the mean fall abundance estimate was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r prettyNum(mean(all_est$Lci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` - `r prettyNum(mean(all_est$Uci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")`).
-
-For spring 2021, we estimated abundance by multiplying the mean fall abundance (standard Schnabel estimates) by the ratio between the spring CPUE and mean fall CPUEs (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), resulting in an estimate of `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough.
-Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm.
+Using the delayed-mixing Schnabel estimator, the mean fall abundance estimate was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r prettyNum(mean(all_est$Lci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` - `r prettyNum(mean(all_est$Uci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")`).
+For spring 2021, we estimated `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough.
+Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm and areal densities (fish/100 m$^2$) of `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_m2) * 100, 1)` and `r round((spring_abund$N_S / ddw_m2) * 100, 1)` for fall and spring, respectively.
 
@@
 Spring consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f30_g"]` for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f90_g"]` for a diet of 90% fish.
-Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` from fall estimates and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` for spring 2021, and assuming 60% of their diet is fish, we estimate that Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and an additional `r round(tot_g_NRR, 1)` metric tons during the NRR emigration in Deadwater Slough.
+Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` from fall estimates and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` for spring 2021, and assuming 60% of their diet is fish, we estimate that Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and an additional `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
 
@@
 
-Using the total fish consumed by the Northern Pikeminnow population (summing fall and spring surveys), we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where juvenile Chinook Salmon make up 30%, 50%, and 65% of total fish prey consumed (Figure \@ref(fig:impact-p)).
+Using the total fish consumed by the Northern Pikeminnow population (summing fall and spring surveys), we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where juvenile Chinook Salmon make up 30%, 50%, and 65% of total fish prey consumed (Figure \@ref(fig:juv-con-p)).
 Assuming that 50% of the total fish consumed are juvenile Chinook salmon, we estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon will be consumed.
-We then used our estimate of total juvenile Chinook Salmon consumed to estimate an "adult equivalents" using the median Granite-to-Granite SAR of `r round(sar$medsar, 5)` (sd = `r format(round(sar$sdsar, 5), scientific = F)`) from @McCann2019.
-Assuming predation was non-existent in Deadwater Slough, the juvenile consumption would be equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) adults.
+Using the median Granite-to-Granite SAR, we estimated the juvenile consumption would is equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) adults (Figure \@ref(fig:adult-imp-p)).
 
@@
 Those estimates translate to linear densities (fish/rkm) that are `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_rkms) / john_day_dens)` and `r round((spring_abund$N_S / ddw_rkms) / john_day_dens)` times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_rkms) / low_columbia_fish_km)` and `r round((spring_abund$N_S / ddw_rkms) / low_columbia_fish_km)` times greater than projections for the lower Columbia River [@Beamesderfer1996].
-Our areal densities (fish/100 $m^2$) for Deadwater Slough are `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_m2) * 100, 1)` and `r round((spring_abund$N_S / ddw_m2) * 100, 1)` for fall and spring, respectively.
 @Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective; therefore it is possible that their estimates may have been low.
@@
 Given the size of the sampling area, and the short duration of our surveys, that assumption is likely met.
-We also caught one individual that was confirmed tagged in Deadwater Slough in a previous year suggesting that at least some individuals remain in the reach or at least migrate to Deadwater Slough multiple times, annually.
-The assumption of a closed population can also affect results if mortality occurs for some marked fish released back to the population.
+<!--We also caught one individual that was confirmed tagged in Deadwater Slough in a previous year suggesting that at least some individuals remain in the reach or at least migrate to Deadwater Slough multiple times, annually.-->The assumption of a closed population can also affect results if mortality occurs for some marked fish released back to the population.
 During field processing, no mortalities or injury post release was observed; however, the potential for mortality amongst marked and released fish cannot be discounted.
@@
 We also question the validity of our assumption of equal catchability of individual fish between sampling events.
-During our study, we recaptured two individual fish during multiple days within a survey indicating that marked fish were still susceptible to angling.
-We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.
-However, we suspect that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
+During our study, we recaptured two individual fish during multiple days within a survey indicating that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.--> However, we suspect that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
 Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit the "good" fishing locales.
@@
 This bias will be proportional to the difference in catchability of marked and unmarked fish.
-As an example, if marked fish are only 70% as likely to be caught as unmarked fish (due to behavior or spatial distribution) the true abundance is closer to 70% of our abundance estimate.
+For example, if marked fish were only 70% as likely to be caught as unmarked fish, the true abundance is closer to 70% of our abundance estimate.
 
@@
 This is likely the case during the fall when DSR Chinook Salmon are the dominant prey species available in Deadwater Slough as 1) no (or at least minimal) hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent that in Chinook Salmon.
-During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river, <50% Chinook Salmon in their diet may be more likely.
+During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river, <50% Chinook Salmon in the Pikeminnow's diet may be more likely.
 Future work to quantify and identify the juvenile salmonid in Deadwater Slough during their seasonal migration would be useful to validate the diet composition assumptions used in our model.
@@
 Spawner abundance is perhaps the most important metric considered in determining a population's viability.
-We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by `r round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` (95% CI: `r round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` - `r round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])`) adults, annually.
-This is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean adult escapement estimated<!--of 558--> for the Upper Salmon MPG in 2017 - 2019 [@Kinzer2020].
-Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of that is Chinook Salmon we estimate the "adult equivalents" of the juveniles consumed to be `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` (95% CI: `r round(total_impacts$adult_Lci[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` - `r round(total_impacts$adult_Uci[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])`) which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% of recent natural-origin adult escapements.
+We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by `r round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` adults, annually, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean adult escapement estimated<!--of 558--> for the Upper Salmon MPG in 2017 - 2019 [@Kinzer2020].
+Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of that is Chinook Salmon we estimate the "adult equivalents" of the juveniles consumed to be `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])`, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% of recent natural-origin adult escapements.
 In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.
@@
 
-```{r cpue-fig, fig.cap = "The number of Northern Pikeminnow caught per angler effort for each sample effort.", eval = F}
-total_cpue %>%
-  ggplot(aes(x = event_name,
-             y = catch_per_hr)) +
-  geom_col(fill = "gray30") +
-  labs(y = "Catch Per Angler Hour") +
-  theme(axis.title.x = element_blank(),
-        axis.text.x = element_text(size = 11,
-                                   color = "black"),
-        axis.title.y = element_text(vjust = 3),
-        axis.text.y = element_text(size = 11,
-                                   color = "black"))
-
-```
-
-```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow in Deadwater Slough using different mark-recapture estimators and for the fall sampling events."}
+```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow in Deadwater Slough using different mark-recapture estimators and for the fall sampling events.
+Error bars indicate 95% confidence intervals."}
 
@@
   scale_x_discrete(labels=c(pm_est_labels))+
-  labs(y = "Abundance (x 1,000)") +
+  labs(y = "Estimated Abundance (x 1,000)",
+       x = "Estimator") +
   #scale_size(guide = "none")+
@@
 \newpage
-
-```{r stmch-cont-p, eval = F, fig.cap = "A stomach contents plot."}
-stmch_cont_per_sz_p = gastric_df %>%
-  filter(method %in% "Angling") %>%
-  ggplot() +
-  aes(x = fish_or_parts) +
-  geom_bar(fill = "#DB740E") +
-  labs(x = "Stomach Contents",
-       y = "Frequency") +
-  theme_classic() +
-  theme(axis.text.x = element_text(angle = 90,
-                                   vjust = 0.5,
-                                   hjust = 1)) +
-  facet_wrap(vars(npm_50),
-             nrow = 2L)
-
-stmch_cont_per_sz_p
-
-```
 
@@
   labs(x = "Total Length (mm)",
-       y = "Frequency") +
+       y = "Number of Northern Pikeminnow") +
   theme(axis.text.x = element_text(size = 11,
@@
              y = adult_N)) +
-  geom_point(size = 3,
-             aes(shape = as.factor(perc_chnk))) +
+  geom_point(size = 3) +
   geom_errorbar(aes(ymin = adult_Lci,
                     ymax = adult_Uci),
-                width = 0.1) +
-  labs(y = "Adult Equivalents",
-       shape = "Percent\nChinook") +
-  theme(axis.title.x = element_blank(),
-        axis.text.x = element_blank(),
+                width = 0) +
+  labs(y = "Estimated Adult Chinook",
+       x = "Percent Chinook in Diet") +
+  theme(axis.text.x = element_text(size = 11,
+                                   color = "black"),
         axis.text.y = element_text(size = 11,

### 2022-02-23 analysis/paper/paper.Rmd
changed map file name
Incorporated edits from coauthors

@@
 ---
-title: "Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Chinook Salmon Populations"
+title: "Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho, and Potential Impacts to Local Chinook Salmon Populations"
 author:
@@
 
-Predation of emigrating juvenile salmonids by piscivorous fishes is perhaps an under-appreciated source of mortality hindering population recovery efforts in the Columbia River basin.
-Reservoirs and tailraces associated with dams on the Snake and Columbia river have created favorable conditions for fishes known to consume juvenile salmonids.
-The native Northern Pikeminnow are one of these fishes and have been estimated to consume ~8% of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually.
-Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above these reservoir systems.
+Predation on emigrating juvenile salmonids by piscivorous fishes is a widely-studied source of mortality within reservoirs and dam tailraces in the Columbia River basin.
+The native Northern Pikeminnow are one of these predators and have been estimated to consume ~8% of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually.
+Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above the reservoir systems.
 In this study, we examine the abundance of Northern Pikeminnow in a slow-water reach of the Salmon River, Idaho, known as Deadwater Slough, and consider potential impacts of them on local Chinook Salmon populations.
-The reach was (at least partially) formed by the failure of a small mining dam in a nearby drainage during the late 19th century and contains historically important overwinter rearing habitat and is a migration corridor for Chinook Salmon as well as other Endangered Species Act-listed species including steelhead and Sockeye Salmon.
-We estimated the abundance of Northern Pikeminnow in Deadwater Slough to be 27,874 in the fall and 19,499 in the spring during the peak emigrations of juvenile Chinook Salmon.
-Assuming a scenario where 60% of the Northern Pikeminnow diet is fish, of which 50% is Chinook Salmon, we estimated the Northern Pikeminnow population to consume 95,120 juvenile Chinook Salmon, annually.
-In the most conservative scenario where 30% of diet is fish and 30% of fish are juvenile Chinook Salmon, 47,079 juveniles are consumed.
-Using a sensitivity analysis, we further considered potential impacts to adult returns to local populations.
-Given the size of the Salmon River, the estimated Northern Pikeminnow densities for Deadwater Slough are remarkable and predation from the population likely has consequential impacts on Chinook Salmon in the Upper Salmon River major population group, thereby hindering local recovery efforts.
+The Deadwater Slough contains historically important overwinter rearing habitat and is a migration corridor for Chinook Salmon as well as other Endangered Species Act-listings including steelhead and Sockeye Salmon.
+We estimated the abundance of Northern Pikeminnow in Deadwater Slough during the peak emigrations of juvenile Chinook Salmonto be 27,874 in the fall and 19,499 in the spring.
+Using these abundance values, we estimated Northern Pikeminnow consumption of juvenile Chinook Salmon for varying proportions of fish in their diet, ranging from 30% to 90%.
+Assuming a median scenario where 60% of the Northern Pikeminnow diet is fish, of which 50% is juvenile Chinook Salmon, we estimated the Northern Pikeminnow population can consume 95,120 juvenile Chinook Salmon, annually.
+Using a sensitivity analysis, the potential impact of predation on juvenile Chinook Salmon under the median scenario is a reduction in adult returns by 539 (95% CI: 231 - 1,335).
+Given the relatively small size of the Salmon River, the large densities and potential consumption rates of Northern Pikeminnow in Deadwater Slough suggest predation likely has consequential impacts on Chinook Salmon in the Upper Salmon River major population group.
 
@@
 
-One potentially important, but perhaps under-appreciated source of salmonid mortality is predation on emigrating juveniles by piscivorous fishes.
+Predation on emigrating juveniles by piscivorous fishes is another important, and potentially under-estimated, source of salmonid mortality.
 In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of recovery [@Fresh2003; @Winther2020].
@@
 Second, reservoirs and downstream tailraces associated with dams on the Snake and Columbia rivers create favorable slow-water habitat for predatory fishes known to consume juvenile salmonids, such as the Northern Pikeminnow *Ptychocheilus oregonsis*.
-Northern Pikeminnow tolerate and thrive in relatively warm, slow water habitats [@Wydoski2003] and consequently have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999].
-Indeed, Northern Pikeminnow are estimated to consume ~8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].
+Northern Pikeminnow tolerate and thrive in relatively warm, slow-water habitats [@Wydoski2003] and consequently have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999].
+Northern Pikeminnow are estimated to consume ~8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].
 
@@
 One such example is Deadwater Slough, an approximately `r round(ddw_rkms, 1)` km long reach of unnaturally slow and deep water in the Salmon River, Idaho.
-The Deadwater Slough is in a river reach containing historically important overwinter rearing habitat, includes the emigration corridor for juvenile Chinook Salmon *O.
-tshawytscha*, and is inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*.
-Further, the slough is part of the migratory pathway of the endangered Snake River Sockeye Salmon *O.
+The Deadwater Slough contains historically important overwinter rearing habitat for juvenile Chinook Salmon *O.
+tshawytscha* and is part of the migratory pathway for upstream Chinook Salmon, the endangered Snake River Sockeye Salmon *O.
 nerka* population [@Axel2015] and several populations of threatened Snake River steelhead.
+The slough is also inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu* that are potential sources of mortality for these rearing and migrating salmonids.
 
@@
 Currently, the Upper Salmon MPG supports eight independent, extant populations; this includes Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River populations [@NOAA2017].
-The Lemhi River historically contained the largest Chinook Salmon population within the MPG and is therefore of particular interest for restoration actions [@Zimmerman2012].
+The Lemhi River historically contained the largest Chinook Salmon population within the MPG and is therefore a prioritized candidate for restoration of natural processes [@Zimmerman2012].
+
 The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a].
@@
 Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency [@Dodson2013].
-Deadwater Slough represents an important habitat for the Upper Salmon MPG as is supports rearing and migration of juveniles from all populations including fall DSR and spring NRR migrants.
+Deadwater Slough represents an important habitat for the Upper Salmon MPG as it supports rearing and migration of juveniles from all populations including fall DSR and spring NRR migrants.
 
@@
 2. Document predation on juvenile Chinook Salmon during the DSR and NRR emigration periods using gastric lavage;
-3. Use a bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population on DSR and NRR Chinook Salmon emigrants at Deadwater Slough;
+3. Use an established bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population on DSR and NRR Chinook Salmon emigrants at Deadwater Slough;
 4. Estimate how consumption of juvenile Chinook Salmon emigrants at Deadwater Slough by Northern Pikeminnow may impact adult returns to the Upper Salmon MPG.
@@
 The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
-Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event and upslope instability that deposited substantial amounts of sediment at the confluence of the Salmon River [@Emerson1973].
+Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event and upslope instability that deposited substantial amounts of sediment at its confluence with the Salmon River [@Emerson1973].
 That sediment formed an alluvial fan which created a slow, deep section in the river, spanning approximately `r round(ddw_hectares)` hectares, averaging `r round(ddw_avg_width_m)` m wide, and resembling a small reservoir.
@@
 During our initial survey in 2019, Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and consequently became our focal taxa.
-The initial intent was to survey during the emigration of the 2018 brood year (fall 2019 and spring 2020), however logistical constraints during the onset of the Covid-19 pandemic delayed the spring 2020 survey until fall 2020.
-Although this allowed an additional survey during the fall DSR emigration, surveys during the spring NRR emigration were still lacking.
-Consequently, an additional week of surveying was added in spring 2021.
+The initial intent was to survey during the emigration of the 2018 Chinook Salmon brood year (fall 2019 and spring 2020), however logistical constraints during the onset of the Covid-19 pandemic delayed the spring 2020 survey until 2021.
+Over this period, an additional fall survey occurred during the 2020 DSR emigration.
 The fall surveys were constrained to two weeks to minimize Northern Pikeminnow immigration/emigration.
@@
 Methods included raft electrofishing, fyke netting, snorkeling, and angling.
-After evaluating all methods, angling was the most effective method for capturing Northern Pikeminnow while also minimizing potential impacts to ESA-listed adult steelhead that were present in the Salmon River during our surveys.
+After evaluating all methods, angling was the most effective method for capturing Northern Pikeminnow while also minimizing potential impacts to ESA-listed adult steelhead that were present during our surveys.
 Ineffective methods were abandoned in subsequent years and the following analyses will focus on fish captured by angling, unless otherwise noted.
@@
 For the single census estimator, we treated the first week of the survey as the mark event and the following week as the recapture event, pooling data within each week.
-Alternatively, the multiple census estimator treated each day as an event and used information about the total number of marked fish from all previous events to calculate the total abundance.
-
+Alternatively, the multiple census estimator treated each day as a sampling event and used information about the total number of marked fish from all previous events to calculate the total abundance.
 For the single census estimator, we used the Chapman-modified Lincoln-Peterson estimator:
@@
 where $M$ is the total number of fish marked and returned to the population during the first event (week), $n$ is the total number of fish caught in the second event (week), and $m$ is the number of marked fish caught during the second week.
-For a multiple census estimator, we used the Schnabel method:
+For a multiple census estimator, we used the Chapman-modified Schnabel estimator:
 
@@
 To account for this potential bias, we also used an adjusted Schnabel estimator, which assumed that marked and released fish during each survey day were not available for recapture for 48 hours.
-The adjusted Schnabel estimator, referred to as the "delayed mixing" Schnabel estimator, is given by:
+The adjusted Schnabel estimator, referred to as the "Delayed-mixing" Schnabel estimator, is given by:
 
@@
 $$
-
-The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and 2020 surveys.
-However, to estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of total CPUE in the fall survey to the total CPUE in the spring survey, then multiplied that ratio by the mean abundance from fall surveys.
+All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the survey period), (2) all fish have equal chance of being caught in subsequent survey days, (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or mistaken.
+
+To estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of mean total CPUE in the fall surveys to the total CPUE in the spring survey, then multiplied that ratio by the mean abundance from fall surveys.
 
@@
 This approach assumes equal capture probabilities for fall and spring surveys.
-Fall and spring population estimates were converted into linear and areal densities by dividing by the length and area, respectively, of Deadwater Slough measured in GIS software and using drone-generated orthography.
-
-All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the survey period), (2) all fish have equal chance of being caught in subsequent survey days, (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or mistaken.
+Fall and spring population estimates were converted into linear and areal densities by dividing by the length and area of Deadwater Slough.
+Length and area were measured in GIS software and using drone-generated orthoimagery.
 
@@
 During the mark-recapture events, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019, including `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals and `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020, including `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals.
-Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.
-
-Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` depending on the estimator used (Table \@ref(tab:abund-tab)).
-For both surveys, the multiple census estimates were generally larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (Figure \@ref(fig:abund-fig)).
+Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn).
+
+Mark-recapture abundance estimates of Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` depending on the estimator used (Table \@ref(tab:abund-tab)).
+For both surveys, the unadjusted multiple-census estimates were larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (Figure \@ref(fig:abund-fig)).
 Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, we deemed the Schnabel estimates most appropriate.
@@
 
-The lengths of Northern Pikeminnow captured ranged from `r min(cmr_df$length, na.rm = T)` to `r max(cmr_df$length, na.rm = T)` mm TL with an average of `r round(mean(cmr_df$length, na.rm = T))` mm (Figure \@ref(fig:length-freq-p)).
-The PSD for Northern Pikeminnow in Deadwater Slough across all three efforts was `r prettyNum(psd, digits = 1)`%.
+The lengths of Northern Pikeminnow ranged from `r min(cmr_df$length, na.rm = T)` to `r max(cmr_df$length, na.rm = T)` mm TL with an average of `r round(mean(cmr_df$length, na.rm = T))` mm (Figure \@ref(fig:length-freq-p)).
+The PSD for Northern Pikeminnow in Deadwater Slough across all three surveys was `r prettyNum(psd, digits = 1)`%.
 
@@
 Spring consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f30_g"]` for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f90_g"]` for a diet of 90% fish.
-Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` from fall estimates and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` for spring 2021, and assuming 60% of their diet is fish, we estimate that Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and an additional `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
+Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` from fall estimates and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` for spring 2021, and assuming 60% of their diet is fish, we estimate that the Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and an additional `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
 
@@
 
-Using the total fish consumed by the Northern Pikeminnow population (summing fall and spring surveys), we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where juvenile Chinook Salmon make up 30%, 50%, and 65% of total fish prey consumed (Figure \@ref(fig:juv-con-p)).
+Using the total fish consumed by the Northern Pikeminnow population (summing fall and spring surveys), we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where 60% of the diet is fish, and juvenile Chinook Salmon make up 30%, 50%, and 65% of total fish prey consumed (Figure \@ref(fig:juv-con-p)).
 Assuming that 50% of the total fish consumed are juvenile Chinook salmon, we estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon will be consumed.
-Using the median Granite-to-Granite SAR, we estimated the juvenile consumption would is equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) adults (Figure \@ref(fig:adult-imp-p)).
+Using the median Granite-to-Granite SAR, we estimated that juvenile consumption is equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) adults (Figure \@ref(fig:adult-imp-p)).
 
@@
 
-All mark-recapture abundance estimators suggest a large population of Northern Pikeminnow occupy Deadwater Slough.
+All mark-recapture abundance estimators suggest a large population of Northern Pikeminnow occupies Deadwater Slough.
 The mean population size of Northern Pikeminnow was estimated to be `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` during the fall DSR emigration and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` during the spring NRR emigration.
@@
 Conversely, at Deadwater Slough the maximum depth was approximately 6 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach.
-Given the size (volume) of the Salmon River at the slough relative to the Columbia River, the estimated Northern Pikeminnow densities for Deadwater Slough are remarkable.
-Further, our estimate densities of Northern Pikeminnow suggest that slow water reaches outside of the reservoir complexes on the Snake and Columbia rivers may support exceptionally high predator densities, consistent with findings of @Harnish2014, @Gray2001, and @Zimmerman1999.
+Our estimated densities of Northern Pikeminnow suggest that slow-water reaches outside of the reservoir complexes on the Snake and Columbia rivers may support exceptionally high predator densities, consistent with findings of @Harnish2014, @Gray2001, and @Zimmerman1999.
 Given the dearth of current data available in the Columbia River basin, it is unclear how Northern Pikeminnow abundance estimates in this study compare to elsewhere in the Columbia River Basin; contemporary information on piscine predator abundances may be needed [@Widener2021].
@@
 Nevertheless, our estimators still provide unbiased estimates of abundance assuming the immigration and emigration rates are equal between marked and unmarked fish.
-Given the size of the sampling area, and the short duration of our surveys, that assumption is likely met.
-<!--We also caught one individual that was confirmed tagged in Deadwater Slough in a previous year suggesting that at least some individuals remain in the reach or at least migrate to Deadwater Slough multiple times, annually.-->The assumption of a closed population can also affect results if mortality occurs for some marked fish released back to the population.
+Given the size of the sampling area, and the short duration of our surveys, the closed population assumption is likely met.
+<!--We also caught one individual that was confirmed tagged in Deadwater Slough in a previous year suggesting that at least some individuals remain in the reach or at least migrate to Deadwater Slough multiple times, annually.-->This assumption can also affect results if mortality occurs for some marked fish released back to the population.
 During field processing, no mortalities or injury post release was observed; however, the potential for mortality amongst marked and released fish cannot be discounted.
@@
 During our study, we recaptured two individual fish during multiple days within a survey indicating that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.--> However, we suspect that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
-Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit the "good" fishing locales.
+Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit "good" fishing locales.
 Violations of the assumption of equal catchability between marked and unmarked fish may lead to an overestimate of abundance of Northern Pikeminnow in our study if marked fish have lower catchability.
@@
 
-In addition to the large population, the observed PSD (`r round(psd)`%) in this study was large compared to observations of 41% below Bonneville Dam and 18% in Bonneville Reservoir [@Winther2020].
+In addition to the large population, the observed (`r round(psd)`%) PSD in this study was large compared to observations of 41% below Bonneville Dam and 18% in Bonneville Reservoir [@Winther2020].
 This suggests that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to populations reported elsewhere in the Columbia River.
-Notably, the state catch-and-release record Northern Pikeminnow, measuring `r max(gastric_df$length, na.rm = T)` mm TL, was caught in Deadwater Slough during the fall 2020 survey.
+Notably, the Idaho state catch-and-release record Northern Pikeminnow, measuring `r max(gastric_df$length, na.rm = T)` mm TL, was caught in Deadwater Slough during the fall 2020 survey.
 
@@
 Observed prey species included juvenile Chinook Salmon as well as Redside Shiner *Richardsonius balteatus*, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*, though decomposition from digestion rendered most fish contents unidentifiable.
-Although our approach provided useful information on prey species, we have concerns about using angling and gastric lavage to quantify diet composition in Northern Pikeminnow.
+Although our approach provided useful information on prey species, using angling and gastric lavage to quantify diet composition in Northern Pikeminnow can be problematic.
 First, diets of angled fish can differ from those collected by other methods such as electrofishing [@Hodgson1988].
-Further, angled fish are more likely to have empty stomachs [@Jurajda2016], perhaps because they are hungry and actively searching for food.
+For instance, angled fish are more likely to have empty stomachs [@Jurajda2016]; possibly because they are captured while actively searching for food, suggeting they are hungry.
 Second, unlike other predators such as Smallmouth Bass and Walleye, that are commonly sampled using gastric lavage [@Kamler2001], Northern Pikeminnow are cyprinids and therefore lack a true stomach.
@@
 To address the prevalence of other fish in the Northern Pikeminnow's diet, we considered scenarios where DSR and NRR Chinook Salmon represent only 30%, 50%, and 65% of total fish prey consumed.
-These values may be considered conservative during the emigration periods when Chinook Salmon are abundant and the Northern Pikeminnow diet shifts almost entirely to piscivory, presumably on juvenile Chinook Salmon [@Shively1996; @Poe1991].
-This is likely the case during the fall when DSR Chinook Salmon are the dominant prey species available in Deadwater Slough as 1) no (or at least minimal) hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent that in Chinook Salmon.
+These values may be considered conservative during the peak emigration periods when Chinook Salmon are abundant and the Northern Pikeminnow diet shifts almost entirely to piscivory, presumably on juvenile Chinook Salmon [@Shively1996; @Poe1991].
+This is likely the case during the fall when DSR Chinook Salmon are the dominant prey species available in Deadwater Slough as 1) no (or at least minimal) hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent than Chinook Salmon emigration.
 During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river, <50% Chinook Salmon in the Pikeminnow's diet may be more likely.
 Future work to quantify and identify the juvenile salmonid in Deadwater Slough during their seasonal migration would be useful to validate the diet composition assumptions used in our model.
+
+## Potential Impacts to Salmon Populations
+
+Recovery of Chinook Salmon in the Upper Salmon MPG is desired to support local fisheries and economies.
+At least five of the eight populations must meet criteria set forth by @McElhany2000 and the @ICTRT2007 for the MPG to be considered viable and for recovery of the Snake River Evolutionary Significant Unit.
+We estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` juvenile Chinook Salmon may be consumed by Northern Pikeminnow assuming that 60% of their diet is fish and 50% of fish prey are Chinook Salmon.
+Considering that diet scenarios used in our study are likely conservative assumptions, especially during the fall DSR emigration, the estimated number of Chinook Salmon consumed is substantial.
+
+Spawner abundance is perhaps the most important metric considered in determining a population's viability and productivity.
+We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by `r round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` adults, annually, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean adult escapement estimated<!--of 558--> for the Upper Salmon MPG in 2017 - 2019 [@Kinzer2020].
+Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of that is Chinook Salmon we estimate the "adult equivalents" of the juveniles consumed to be `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])`, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% of recent natural-origin adult escapements.
+In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.
+Accordingly, we surmise that adult returns of steelhead and Sockeye Salmon are also affected by predation, including hatchery populations, which provide recreational fishing opportunities.
+Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including other ESA-listed species.
+Because the Deadwater Slough is part of the migratory pathway for multiple species and populations of emigrating salmonids, the positive impact could be greater than individual tributary rehabilitation actions which typically benefit a single population.
 
@@
 
-## Potential Impacts to Salmon Populations
-
-Recovery of Chinook Salmon in the Upper Salmon MPG is desired to support local fisheries and economies.
-At least five of the eight populations must meet criteria set forth by @McElhany2000 and the @ICTRT2007 for the MPG to be considered viable and for recovery of the Snake River Evolutionary Significant Unit.
-We estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` juvenile Chinook Salmon may be consumed by Northern Pikeminnow assuming that 60% of their diet is fish and 50% of that is Chinook Salmon.
-Considering that diet scenarios used in our study are likely conservative assumptions, especially during the fall DSR emigration, the estimated number of Chinook Salmon consumed is substantial.
-
-Spawner abundance is perhaps the most important metric considered in determining a population's viability.
-We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by `r round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` adults, annually, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean adult escapement estimated<!--of 558--> for the Upper Salmon MPG in 2017 - 2019 [@Kinzer2020].
-Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of that is Chinook Salmon we estimate the "adult equivalents" of the juveniles consumed to be `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])`, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% of recent natural-origin adult escapements.
-In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.
-Accordingly, we surmise that adult returns of steelhead and Sockeye Salmon are also affected by predation, including hatchery populations, which provide recreational fishing opportunities.
-Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including other ESA-listed species.
-Because the Deadwater Slough is part of the migratory pathway for multiple species and populations of emigrating salmonids, the positive impact could be greater than individual tributary rehabilitation actions which typically benefit a single population.
-
 ## Management Implications
@@
 Removing the Dump Creek alluvial fan has the benefit of restoring the natural fluvial processes in the Salmon River that likely existed in the reach prior to the failure of the mining reservoir dam in Dump Creek.
-Restoring natural processes could reduce Northern Pikeminnow densities to levels similar to upstream and downstream reaches where higher survival and transition probabilities have been observed compared to Deadwater Slough [@Axel2015; @Ackerman2018; @Porter2019].
-Managers ought also consider the feasibility and net benefit of restoring fluvial processes relative to potential losses for recreational fishing and bird watching opportunities.
+Restoring natural processes could reduce Northern Pikeminnow densities to levels similar to upstream and downstream reaches where higher survival and transition probabilities for juvenile salmon have been observed compared to Deadwater Slough [@Axel2015; @Ackerman2018; @Porter2019].
+Managers ought also to consider the feasibility and net benefit of restoring fluvial processes relative to potential losses for recreational fishing and bird watching opportunities.
 
@@
 The failure of a mining reservoir dam in the late 19th century created an alluvial fan at the confluence of the Salmon River and Dump Creek which has backed up the Salmon River and created favorable conditions for Northern Pikeminnow.
-Even under the most conservative scenarios, we estimate more than 10,000 Northern Pikeminnow occupy Deadwater Slough, resulting in a remarkable density given the size of the area.
+Even under the most conservative scenarios, we estimate more than 10,000 Northern Pikeminnow occupy Deadwater Slough, resulting in a remarkably high density given the size of the area.
 Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids that are rearing in or emigrating through Deadwater Slough, including Chinook Salmon.
-We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than 24,000 juvenile Chinook salmon (the lower 95% CI of the most conservative scenario evaluated).
-It is more likely that this value is closer to 100,000, annually, creating a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
+We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than 24,000 juvenile Chinook salmon, however, it is likely that this value is closer to 100,000, annually.
+The magnitude of predation by Northern Pikeminnow can reduce adult escapement between 47% and 97% and should be considered a major factor influencing the recovery ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
 
@@
         format.args = list(big.mark = ","),
-        caption = "Summary of gastric lavage results including the total number of fish captured and gastric lavaged.
-For Northern Pikeminnow that were gastric lavaged, we show the number containing stomach contents, and of those, the number containing fish contents.") %>%
+        caption = "Summary of gastric lavage results including counts of the total number of Northern Pikeminnow captured, gastric lavaged, individuals with stomach contents, and individuals with fish prey contents") %>%
   kable_styling()
@@
 
-```{r map, out.width = "85%", fig.cap = "Map of the Deadwater Slough study area.
-The high-resolution orthoimage portion surrounding the Deadwater Slough was generated from aerial images taken by an unmanned aircraft system (drone).
-The red polygon indicates the reach characterized by unnaturally slow water velocities and a deepened channel.
-The location of the Dump Creek delta is indicated."}
-knitr::include_graphics("../figures/deadwater_map_low_res.png")
-
-```
-
-\newpage
-
-```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow in Deadwater Slough using different mark-recapture estimators and for the fall sampling events.
-Error bars indicate 95% confidence intervals."}
+```{r map, out.width = "85%", fig.cap = "Map of the Deadwater Slough study area within the Upper Salmon River MPG. The Dump Creek alluvial fan is loacted at the downstream end of the study reach."}
+knitr::include_graphics("../figures/Study Area Map 5.png")
+
+```
+
+\newpage
+
+```{r abund-fig, fig.cap = "Estimated abundance of Northern Pikeminnow in Deadwater Slough from three mark-recapture estimators for the fall surveys. Error bars indicate 95% confidence intervals."}
 
@@
   scale_x_discrete(labels=c(pm_est_labels))+
-  labs(y = "Estimated Abundance (x 1,000)",
+  labs(y = "Estimated Pikeminnow Abundance (x 1,000)",
        x = "Estimator") +
@@
         legend.key.size = unit(.5, "cm"),legend.box = 'horizontal',
-        axis.title.x = element_blank(),
+        axis.title.x = element_text(size = 11, color = "black"),
         axis.text.x = element_text(size = 11, color = "black"),
@@
                  alpha = 0.5) +
-  scale_fill_manual(name = "Season",
+  scale_fill_manual(name = "Survey",
                     values = c("gray", "black"),
@@
 
-```{r bio-p, fig.cap = "The cumulative fish (g) pre consumed by an individual Northern Pikeminnow during the peak DSR (fall) and NRR (spring) Chinook Salmon emigrations (top panel) and corresponding daily mean water temperatures (bottom panel).
-The black line is the median among diet scenarios (60% of the Northern Pikeminnow diet consists of fish) and the gray area shows the range between diet scenarios of 30 and 90% fish prey.
+```{r bio-p, fig.cap = "The cumulative fish prey consumed (g) by an individual Northern Pikeminnow during the peak Chinook Salmon emigrations for fall DSR (top left) and spring NRR  (top right) and corresponding daily mean water temperatures (bottom panels).
+The black line is the median Northern Pikeminnow diet scenario, consisting of 60% fish and the gray area shows the range between diet scenarios of 30% and 90% fish prey.
 Daily mean water temperatures were summarized from six years of data available from the Shoup gage station.
@@
 
-```{r juv-con-p, fig.cap = "test"}
+```{r juv-con-p, fig.cap = "Estimated number of juvenile Chinook Salmon consumed by Northern Pikeminnow from a sensitivity analysis of variable diet scenarios.
+The proportion of Chinook Salmon in the total fish prey consumed by Northern Pikminnow were evaluated for 30%, 50%, and 65% scenarios, similar to those reported by Zimmerman and Ward 1999 "}
+
 juv_cons_p = impact_df %>%
@@
 
-```{r adult-imp-p, fig.cap = "temp"}
+```{r adult-imp-p, fig.cap = "Estimated reduction in the annual number of returning adult Chinook Salmon as a result of Northern Pikeminnow predation. Error bars indicate 95% confidence intervals."}
 #adult impacts
@@
   labs(y = "Estimated Adult Chinook",
-       x = "Percent Chinook in Diet") +
+       x = "Percent Chinook Salmon in Total Fish Consumed") +
   theme(axis.text.x = element_text(size = 11,

### 2022-02-23 analysis/paper/paper.Rmd
Incorporated edits from coauthors

@@
 
-Most predation studies have focused on reservoirs [@Murphy2021] and mainstem reaches [@Zimmerman1999; @Tabor1993; @Ward1995; @Shively1996] of the Columbia River basin whereas considerably less is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems [@Rubenson2020].
+Most predation studies focus on reservoirs [@Murphy2021] and mainstem reaches [@Zimmerman1999; @Tabor1993; @Ward1995; @Shively1996] of the Columbia River basin whereas considerably less is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems [@Rubenson2020].
 Upstream habitats containing slower water velocities and other attributes that support piscivorous predators may overlap with essential habitat for some salmonid species and life stages.

### 2022-02-23 analysis/paper/paper.Rmd
some very minor typos

--- 2022-02-23 analysis/paper/paper.Rmd
@@
 The Deadwater Slough contains historically important overwinter rearing habitat and is a migration corridor for Chinook Salmon as well as other Endangered Species Act-listings including steelhead and Sockeye Salmon.
-We estimated the abundance of Northern Pikeminnow in Deadwater Slough during the peak emigrations of juvenile Chinook Salmonto be 27,874 in the fall and 19,499 in the spring.
+We estimated the abundance of Northern Pikeminnow in Deadwater Slough during the peak emigrations of juvenile Chinook Salmon to be 27,874 in the fall and 19,499 in the spring.
 Using these abundance values, we estimated Northern Pikeminnow consumption of juvenile Chinook Salmon for varying proportions of fish in their diet, ranging from 30% to 90%.
@@
 Given the relatively small size of the Salmon River, the large densities and potential consumption rates of Northern Pikeminnow in Deadwater Slough suggest predation likely has consequential impacts on Chinook Salmon in the Upper Salmon River major population group.
-
-\newpage
 
@@
 Similarly, emigrating DSR Chinook Salmon had an approximate 10% reduction in transition probability through the Deadwater Slough compared to surrounding reaches during fall and early winter months [@Ackerman2018; @Porter2019].
-Low survival was attributed to low-velocity water delaying movement rates, thereby increasing predation risk.
+Low survival was attributed to lack of fish cover and low-velocity water delaying movement rates, thereby increasing predation risk.
 
@@
 Fall and spring population estimates were converted into linear and areal densities by dividing by the length and area of Deadwater Slough.
-Length and area were measured in GIS software and using drone-generated orthoimagery.
+Length and area were measured in QGIS software [@QGISDevelopmentTeam2022] and using drone-generated orthoimagery.
 

### 2022-03-11 analysis/paper/paper.Rmd
typo on title page. Added Mark and Ian to acknowledgements.

@@
   - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
-  - wdfw: Washington Department of Fish and Wildife, Fish Program, Science Division, 1111 Washington Street NE, Olympia, Washington, 98501, USA
+  - wdfw: Washington Department of Fish and Wildlife, Fish Program, Science Division, 1111 Washington Street NE, Olympia, Washington, 98501, USA
   - mhe_sandy: Mount Hood Environmental, 39085 Pioneer Boulevard \#100 Mezzanine, Sandy, Oregon, 97055, USA 
@@
 We further appreciate the administrative support and guidance from Caitlin Alcott and Inter-Fluve.
-This manuscript benefited from a review by Sean Gibbs.
+This manuscript benefited from reviews by Sean Gibbs, Mark Roes, and Ian Courter.
 Funding for this study was provided by the Bureau of Reclamation, Pacific Northwest Regional Office (contract No.

### 2022-03-20 analysis/paper/paper.Rmd
Incorporated edits from reviewers IC and MR

@@
 Predation on emigrating juvenile salmonids by piscivorous fishes is a widely-studied source of mortality within reservoirs and dam tailraces in the Columbia River basin.
-The native Northern Pikeminnow are one of these predators and have been estimated to consume ~8% of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually.
+Native Northern Pikeminnow have been estimated to consume ~8% of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually.
 Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above the reservoir systems.
-In this study, we examine the abundance of Northern Pikeminnow in a slow-water reach of the Salmon River, Idaho, known as Deadwater Slough, and consider potential impacts of them on local Chinook Salmon populations.
+In this study, we examine the abundance of Northern Pikeminnow in a slow-water reach of the Salmon River, Idaho, known as Deadwater Slough, and consider their potential impacts on local Chinook Salmon populations.
 The Deadwater Slough contains historically important overwinter rearing habitat and is a migration corridor for Chinook Salmon as well as other Endangered Species Act-listings including steelhead and Sockeye Salmon.
-We estimated the abundance of Northern Pikeminnow in Deadwater Slough during the peak emigrations of juvenile Chinook Salmon to be 27,874 in the fall and 19,499 in the spring.
-Using these abundance values, we estimated Northern Pikeminnow consumption of juvenile Chinook Salmon for varying proportions of fish in their diet, ranging from 30% to 90%.
-Assuming a median scenario where 60% of the Northern Pikeminnow diet is fish, of which 50% is juvenile Chinook Salmon, we estimated the Northern Pikeminnow population can consume 95,120 juvenile Chinook Salmon, annually.
-Using a sensitivity analysis, the potential impact of predation on juvenile Chinook Salmon under the median scenario is a reduction in adult returns by 539 (95% CI: 231 - 1,335).
+We estimated the abundance of Northern Pikeminnow in Deadwater Slough to be 27,874 in the fall and 19,499 in the spring, corresponding with the peak emigrations of juvenile Chinook Salmon.
+Using these abundance values, we estimated Northern Pikeminnow consumption of juvenile Chinook Salmon.
+Assuming 60% of the Northern Pikeminnow diet is fish, of which 50% is juvenile Chinook Salmon, we estimated the Northern Pikeminnow population can consume 95,120 juvenile Chinook Salmon, annually.
+Using a sensitivity analysis, Northern Pikeminnow predation on juvenile Chinook Salmon was predicted to reduce adult returns by 539 (95% CI: 231 - 1,335).
 Given the relatively small size of the Salmon River, the large densities and potential consumption rates of Northern Pikeminnow in Deadwater Slough suggest predation likely has consequential impacts on Chinook Salmon in the Upper Salmon River major population group.
@@
 nerka* population [@Axel2015] and several populations of threatened Snake River steelhead.
-The slough is also inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu* that are potential sources of mortality for these rearing and migrating salmonids.
-
-Quantifying mortality associated with piscine predation in upstream habitats of the Salmon River is of particular interest for spring/summer-run Chinook Salmon (hereafter Chinook Salmon) in the Upper Salmon major population group (MPG).
+The slough is also inhabited by piscivorous predators, Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*, that are potential sources of mortality for rearing and migrating salmonids.
+
 The Salmon River was historically the most productive tributary for Chinook Salmon in the Columbia River basin [@Nemeth1999], but those populations have become depleted in recent decades.
+Therefore, quantifying the mortality of spring/summer-run Chinook Salmon (hereafter Chinook Salmon) in the Upper Salmon major population group (MPG) associated with piscine predation is of particular interest.
 Currently, the Upper Salmon MPG supports eight independent, extant populations; this includes Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River populations [@NOAA2017].
@@
 Predators were sampled near the peaks of the fall DSR and spring NRR emigrations.
-During our initial survey in 2019, Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and consequently became our focal taxa.
-The initial intent was to survey during the emigration of the 2018 Chinook Salmon brood year (fall 2019 and spring 2020), however logistical constraints during the onset of the Covid-19 pandemic delayed the spring 2020 survey until 2021.
+During our initial survey in 2019, Northern Pikeminnow were the most prevalent piscine predator in Deadwater Slough and consequently became our focal taxa.
+Our intent was to survey during the emigration of the 2018 Chinook Salmon brood year that occurred fall 2019 and spring 2020, however logistical constraints during the onset of the Covid-19 pandemic delayed the spring 2020 survey until 2021.
 Over this period, an additional fall survey occurred during the 2020 DSR emigration.
-The fall surveys were constrained to two weeks to minimize Northern Pikeminnow immigration/emigration.
-
-Multiple capture methods were employed during the initial fall 2019 survey to reduce selectivity and bias for species and size classes.
+Fall surveys were constrained to two weeks to minimize Northern Pikeminnow immigration/emigration.
+
+Multiple capture methods were employed during the fall 2019 survey to reduce selectivity and bias for species and size classes.
 Methods included raft electrofishing, fyke netting, snorkeling, and angling.
 After evaluating all methods, angling was the most effective method for capturing Northern Pikeminnow while also minimizing potential impacts to ESA-listed adult steelhead that were present during our surveys.
-Ineffective methods were abandoned in subsequent years and the following analyses will focus on fish captured by angling, unless otherwise noted.
+Less effective methods were abandoned in subsequent years and the following analyses will focus on fish captured by angling, unless otherwise noted.
 
@@
 Unmarked fish were given a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) unique to each day and then were released.
-Released fish were distributed throughout Deadwater Slough to facilitate mixing back into the population.
-Fish that died prior to release were included in the total capture (or recapture) count, but not included in the number of marked fish available for recapture after that date.
+Fish release sites were distributed throughout Deadwater Slough to facilitate mixing back into the population.
+Fish that died prior to release were included in that day's total capture (or recapture) count, but not included in the number of marked fish available for recapture after that date.
 Finally, we recorded the angling start and end time for each crew (person or combinations of persons) to calculate CPUE as the number of Northern Pikeminnow caught per angler hour.
@@
 where $M$, $n$, and $m$ are indexed by each survey (day), $i$.
-With small numbers of recaptures, it may not be appropriate to calculate the standard error of the Schnabel estimator, but the 95% confidence intervals can be calculated using a Poisson approximation [@Krebs1999].<!--The Schnabel estimator is essentially the weighted average of a series of Chapman-modified Lincoln-Peterson estimators.-->
-
-Because captures of new and marked fish were dependent on volunteer anglers, there was concern that sampling was biased to locations preferred by anglers.
-If marked fish were not recently released or mixed back into those preferred locales, marked fish may have been less susceptible to capture than unmarked fish.
-To account for this potential bias, we also used an adjusted Schnabel estimator, which assumed that marked and released fish during each survey day were not available for recapture for 48 hours.
-The adjusted Schnabel estimator, referred to as the "Delayed-mixing" Schnabel estimator, is given by:
+The 95% confidence intervals were calculated using a Poisson approximation [@Krebs1999].<!--The Schnabel estimator is essentially the weighted average of a series of Chapman-modified Lincoln-Peterson estimators.-->
+
+Because fish capture was dependent on volunteer anglers, there was concern that sampling effort was biased to locations preferred by anglers.
+If release sites for marked fish were not proximal to preferred angling locales, recently released fish may have been less susceptible to immediate capture than unmarked fish.
+To account for this potential bias, we included a "Delayed-mixing" Schnabel estimator that assumed fish marked and released on a given survey day were not available for recapture for 48 hours.
+The "Delayed-mixing" Schnabel estimator, is given by:
 
@@
 $$
-All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the survey period), (2) all fish have equal chance of being caught in subsequent survey days, (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or mistaken.
-
-To estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of mean total CPUE in the fall surveys to the total CPUE in the spring survey, then multiplied that ratio by the mean abundance from fall surveys.
+All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the survey period), (2) all fish have equal chance of being caught in subsequent survey days, (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or misidentified.
+
+Surveys conducted in spring 2021 did not follow a mark-recapture study design so Northern Pikeminnow abundance was estimated as the ratio of mean total CPUE in the fall surveys to the total CPUE in the spring survey, then multiplied that ratio by the mean abundance from fall surveys.
 
@@
 This approach assumes equal capture probabilities for fall and spring surveys.
-Fall and spring population estimates were converted into linear and areal densities by dividing by the length and area of Deadwater Slough.
+Fall and spring population estimates were converted into linear and areal densities by dividing by the length and area of Deadwater Slough, respectively.
 Length and area were measured in QGIS software [@QGISDevelopmentTeam2022] and using drone-generated orthoimagery.
@@
 To describe the population size structure, we calculated the proportional stock density (PSD) for Northern Pikeminnow in Deadwater Slough.
-PSD is a ratio, expressed as a percentage, between the number of "quality-sized" (larger) individuals and "stock-sized" individuals:
+PSD is the percentage of all fish that meet criteria for "quality-length" (larger) individuals and "stock-length" individuals:
 
 $$
-PSD_{i} = 100 * \frac{FQ_{i}}{FS_{i}}
+PSD = 100 * \frac{FQ}{FS}
 $$
 
-where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$.
+where $FQ$ is the number of fish $\ge$ quality-length, and $FS$ is the number of fish $\ge$ stock-length.
 For Northern Pikeminnow in Deadwater Slough, we used `r prettyNum(q_size, digits = 0)` mm TL for quality-length and `r prettyNum(s_size, digits = 0)` mm TL for stock-length [@Winther2020].
@@
 Immediately following lavage, stomach contents of individuals were preserved with 99% isopropyl alcohol in whirl-paks  to be analyzed in a controlled environment.
-For each stomach sample, wet weight (grams) was recorded for the total stomach contents, including all non-fish items (e.g., macroinvertebrates, organic matter), as well as the portion consisting of fish.
+For each stomach sample, total wet weight (grams) was recorded for all stomach contents, including all fish and non-fish items (e.g., macroinvertebrates, organic matter).
 Fish and fish remnants were identified to the lowest taxonomic unit using diagnostic bones [@Hansel1988; @Frost2000] or were categorized as unknown.
@@
 To estimate the total consumption potential of Northern Pikeminnow in Deadwater Slough during the peaks of fall DSR and spring NRR emigrations, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 applied in R statistical software [@RCoreTeam2021].
-The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on the following inputs: predator and prey energy densities, predator start and end weights, and water temperatures.
+The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on predator and prey energy densities, predator start and end weights, and water temperatures.
 Separate models were run for the fall period (September 15 - November 30) and the spring period (March 1 - May 31) to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG.
 
-Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017].
+Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules (J)/g [@Deslauriers2017].
 Prey energy densities were fixed at 3,000 J/g for invertebrates and 21,500 J/g for juvenile Chinook Salmon [@Moss2016].
@@
 The proportion of fish in the Northern Pikeminnow's diet relative to non-fish prey items is unknown in Deadwater Slough.
-Therefore, we conducted separate model runs that assume different proportions of fish in the diet ranging from 30 - 90%, in 10% increments.
-These values were supported by observations elsewhere in the Columbia River basin (54-86%: @Shively1996; 48-86%: @Zimmerman1999; 37%: @Gray2001) that suggest fish are the majority of Northern Pikeminnow prey.
-Fourteen model runs were completed (7 diet scenarios and 2 seasons), each providing an estimate of the total grams of fish consumed by an individual Northern Pikeminnow (Table \@ref(tab:bio-tbl)).
+Therefore, we conducted a series of model runs with varying proportions of fish in the diet ranging from 30 - 90%, in 10% increments.
+These values were supported by observations elsewhere in the Columbia River basin (54-86%: @Shively1996; 48-86%: @Zimmerman1999; 37%: @Gray2001) that suggest fish are the majority of Northern Pikeminnow prey by volume.
+A model run was conducted for each combination of diet scenario and season, resulting in fourteen estimates of the total grams of fish consumed by an individual Northern Pikeminnow (Table \@ref(tab:bio-tbl)).
 To estimate the total biomass of fish consumed by Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated Northern Pikeminnow population sizes during fall and spring.
@@
 
-The estimate of total juvenile Chinook Salmon biomass consumed was then converted to the total estimated number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
+We obtained an estimate of the number of juvenile Chinook Salmon consumed by dividing total biomass consumed by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
 The average weights of DSR and NRR emigrants were calculated from fish captured at seven rotary screw traps upstream of Deadwater Slough during the fall and spring periods.
-Although the primary impact to Chinook Salmon in Deadwater Slough by Northern Pikeminnow is juvenile predation, the common metric to evaluate salmon recovery is often adult returns.
-We assessed the potential impacts on adult returns by estimating the number of additional adult Chinook Salmon expected to return to Lower Granite Dam if predation was reduced or eliminated in the Deadwater Slough.
-This calculation was made by multiplying the total estimated juvenile Chinook Salmon consumed by the median Granite-to-Granite SAR of 0.00614 (sd = 0.00051) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.
+We also quantified the potential impact of Northern Pikeminnow predation on juvenile Chinook Salmon to adult returns by estimating the number of additional adult Chinook Salmon expected to return to Lower Granite Dam if predation by Pikeminnow was reduced or eliminated in the Deadwater Slough.
+We multiplied the estimated total juvenile Chinook Salmon consumed by the median Granite-to-Granite smolt-to-adult return rate (SAR) of 0.00614 (SD = 0.00051) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.
 
@@
 
-Mark-recapture abundance estimates of Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` depending on the estimator used (Table \@ref(tab:abund-tab)).
+Mark-recapture abundance estimates of Northern Pikeminnow in Deadwater Slough ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` in fall 2019 and from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` in fall 2020 (Table \@ref(tab:abund-tab)).
 For both surveys, the unadjusted multiple-census estimates were larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (Figure \@ref(fig:abund-fig)).

### 2022-03-20 analysis/paper/paper.Rmd
Incorporated edits from reviewers IC and MR

@@
 
-We obtained an estimate of the number of juvenile Chinook Salmon consumed by dividing total biomass consumed by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
+The estimate of total juvenile Chinook Salmon biomass consumed was then converted to the total estimated number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
 The average weights of DSR and NRR emigrants were calculated from fish captured at seven rotary screw traps upstream of Deadwater Slough during the fall and spring periods.
-We also quantified the potential impact of Northern Pikeminnow predation on juvenile Chinook Salmon to adult returns by estimating the number of additional adult Chinook Salmon expected to return to Lower Granite Dam if predation by Pikeminnow was reduced or eliminated in the Deadwater Slough.
-We multiplied the estimated total juvenile Chinook Salmon consumed by the median Granite-to-Granite smolt-to-adult return rate (SAR) of 0.00614 (SD = 0.00051) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.
+Although the primary impact to Chinook Salmon in Deadwater Slough by Northern Pikeminnow is juvenile predation, the common metric to evaluate salmon recovery is often adult returns.
+We assessed the potential impacts on adult returns by estimating the number of additional adult Chinook Salmon expected to return to Lower Granite Dam if predation was reduced or eliminated in the Deadwater Slough.
+This calculation was made by multiplying the total estimated juvenile Chinook Salmon consumed by the median Granite-to-Granite SAR of 0.00614 (sd = 0.00051) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.
 
@@
 
-Mark-recapture abundance estimates of Northern Pikeminnow in Deadwater Slough ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` in fall 2019 and from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` in fall 2020 (Table \@ref(tab:abund-tab)).
+Mark-recapture abundance estimates of Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` depending on the estimator used (Table \@ref(tab:abund-tab)).
 For both surveys, the unadjusted multiple-census estimates were larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (Figure \@ref(fig:abund-fig)).

### 2022-03-23 analysis/paper/paper.Rmd
Updated with edits from IC and MR

--- 2022-03-20 analysis/paper/paper.Rmd
@@
 Mark-recapture abundance estimates of Northern Pikeminnow in Deadwater Slough ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` in fall 2019 and from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` in fall 2020 (Table \@ref(tab:abund-tab)).
-For both surveys, the unadjusted multiple-census estimates were larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (Figure \@ref(fig:abund-fig)).
+We estimated larger populations of Northern Pikeminnow using the unadjusted multiple-census estimator compared to adjusted and single census estimators (Figure \@ref(fig:abund-fig)).
 Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, we deemed the Schnabel estimates most appropriate.
@@
 Northern Pikeminnow captured during spring 2021 had a higher proportion of fish content in their diet compared to individuals captured in the fall surveys (Table \@ref(tab:gastric-tab)).
-The mean wet weight of total contents for an individual was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; sd = `r round(sd(stmch_df$stomach_contents_weight), 2)` g).
+The mean wet weight of total contents for an individual was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; SD = `r round(sd(stmch_df$stomach_contents_weight), 2)` g).
 Overall, fish or fish remnants made up `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of all stomach contents examined.
@@
 Spring consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f30_g"]` for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f90_g"]` for a diet of 90% fish.
-Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` from fall estimates and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` for spring 2021, and assuming 60% of their diet is fish, we estimate that the Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and an additional `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
+Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` in the fall, `r prettyNum(round(spring_abund$N_S), big.mark = ",")` in the spring 2021, and a diet consisting of 60% fish, the Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
 
@@
 
-Using the total fish consumed by the Northern Pikeminnow population (summing fall and spring surveys), we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where 60% of the diet is fish, and juvenile Chinook Salmon make up 30%, 50%, and 65% of total fish prey consumed (Figure \@ref(fig:juv-con-p)).
-Assuming that 50% of the total fish consumed are juvenile Chinook salmon, we estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon will be consumed.
+Given yearly fish consumption estimates during the fall DSR and spring NRR periods, we produced a sensitivity analysis to calculated consumption of juvenile Chinook Salmon where 30%, 50%, and 65% of total fish prey consumed are juvenile Chinook Salmon (Figure \@ref(fig:juv-con-p)).
+At the median value of 50%, we estimated that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon will be consumed.
 Using the median Granite-to-Granite SAR, we estimated that juvenile consumption is equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) adults (Figure \@ref(fig:adult-imp-p)).
@@
 We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by `r round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` adults, annually, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean adult escapement estimated<!--of 558--> for the Upper Salmon MPG in 2017 - 2019 [@Kinzer2020].
-Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of that is Chinook Salmon we estimate the "adult equivalents" of the juveniles consumed to be `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])`, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% of recent natural-origin adult escapements.
+Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of that is juvenile Chinook Salmon, we estimated `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` "adult equivalents" would be consumed, which is approximately `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% of recent natural-origin adult Chinook Salmon escapements.
 In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.

### 2022-03-30 analysis/paper/paper.Rmd
made sure all numbers in text jived

@@
 The Deadwater Slough contains historically important overwinter rearing habitat and is a migration corridor for Chinook Salmon as well as other Endangered Species Act-listings including steelhead and Sockeye Salmon.
-We estimated the abundance of Northern Pikeminnow in Deadwater Slough to be 27,874 in the fall and 19,499 in the spring, corresponding with the peak emigrations of juvenile Chinook Salmon.
+We estimated the abundance of Northern Pikeminnow in Deadwater Slough to be 19,499 in the fall and 10,352 in the spring, corresponding with the peak emigrations of juvenile Chinook Salmon.
 Using these abundance values, we estimated Northern Pikeminnow consumption of juvenile Chinook Salmon.
-Assuming 60% of the Northern Pikeminnow diet is fish, of which 50% is juvenile Chinook Salmon, we estimated the Northern Pikeminnow population can consume 95,120 juvenile Chinook Salmon, annually.
-Using a sensitivity analysis, Northern Pikeminnow predation on juvenile Chinook Salmon was predicted to reduce adult returns by 539 (95% CI: 231 - 1,335).
+Assuming 60% of the Northern Pikeminnow diet is fish, of which 50% is juvenile Chinook Salmon, we estimated the Northern Pikeminnow population can consume 61,409 juvenile Chinook Salmon, annually.
+Using a sensitivity analysis, Northern Pikeminnow predation on juvenile Chinook Salmon was predicted to reduce adult returns by 377 (95% CI: 161 - 935).
 Given the relatively small size of the Salmon River, the large densities and potential consumption rates of Northern Pikeminnow in Deadwater Slough suggest predation likely has consequential impacts on Chinook Salmon in the Upper Salmon River major population group.
@@
 spring_abund = all_est %>%
-  filter(model == "Schnabel") %>%
-  #filter(model == "Schnabel - Delayed Mixing") %>%
+  filter(model == "Schnabel - Delayed Mixing") %>%
+  # filter(model == "Schnabel") %>%
   select(-range, -prop_range) %>%
@@
 Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, we deemed the Schnabel estimates most appropriate.
-Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`).
+Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r prettyNum(round(mean(all_est$Lci[all_est$model == "Schnabel"])), big.mark = ",")` - `r prettyNum(round(mean(all_est$Uci[all_est$model == "Schnabel"])), big.mark = ",")`).
 Using the delayed-mixing Schnabel estimator, the mean fall abundance estimate was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r prettyNum(mean(all_est$Lci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` - `r prettyNum(mean(all_est$Uci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")`).
 For spring 2021, we estimated `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough.
-Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm and areal densities (fish/100 m$^2$) of `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_m2) * 100, 1)` and `r round((spring_abund$N_S / ddw_m2) * 100, 1)` for fall and spring, respectively.
+Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm and areal densities (fish/100 m$^2$) of `r round((mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_m2) * 100, 1)` and `r round((spring_abund$N_S / ddw_m2) * 100, 1)` for fall and spring, respectively.
 
@@
 We found that gastric lavage successfully removed food items in all dissected individuals, indicating the method was effective.
-Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs.
+Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` stomachs.
 Northern Pikeminnow captured during spring 2021 had a higher proportion of fish content in their diet compared to individuals captured in the fall surveys (Table \@ref(tab:gastric-tab)).
@@
 ```{r tons-consumption}
-tot_g_DSR = (bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"] * mean(all_est$N[all_est$model == "Schnabel"])) / 1000000
+tot_g_DSR = (bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"] * mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"])) / 1000000
 tot_g_NRR = (bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f60_g"] * round(spring_abund$N_S)) / 1000000
@@
 Spring consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f30_g"]` for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f90_g"]` for a diet of 90% fish.
-Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` in the fall, `r prettyNum(round(spring_abund$N_S), big.mark = ",")` in the spring 2021, and a diet consisting of 60% fish, the Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
+Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` in the fall, `r prettyNum(round(spring_abund$N_S), big.mark = ",")` in the spring 2021, and a diet consisting of 60% fish, the Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
 
@@
 All mark-recapture abundance estimators suggest a large population of Northern Pikeminnow occupies Deadwater Slough.
-The mean population size of Northern Pikeminnow was estimated to be `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` during the fall DSR emigration and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` during the spring NRR emigration.
-Those estimates translate to linear densities (fish/rkm) that are `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_rkms) / john_day_dens)` and `r round((spring_abund$N_S / ddw_rkms) / john_day_dens)` times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_rkms) / low_columbia_fish_km)` and `r round((spring_abund$N_S / ddw_rkms) / low_columbia_fish_km)` times greater than projections for the lower Columbia River [@Beamesderfer1996].
+The mean population size of Northern Pikeminnow was estimated to be `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` during the fall DSR emigration and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` during the spring NRR emigration.
+Those estimates translate to linear densities (fish/rkm) that are `r round((mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms) / john_day_dens)` and `r round((spring_abund$N_S / ddw_rkms) / john_day_dens)` times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r round((mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms) / low_columbia_fish_km)` and `r round((spring_abund$N_S / ddw_rkms) / low_columbia_fish_km)` times greater than projections for the lower Columbia River [@Beamesderfer1996].
 @Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective; therefore it is possible that their estimates may have been low.

### 2022-04-01 analysis/paper/paper.Rmd
moved Avian Predation to Supplementary Materials
minor edits in Discussion prior to near final editing

@@
 
-## Avian Predation
-
-Although not formally assessed in this study, avian predators are another potential source of mortality for juvenile Chinook Salmon and other ESA-listed salmonids in the Deadwater Slough.
-The Deadwater Slough is recognized as an important bird watching and nesting area due to the associated upstream braided channels and riparian and backwater habitats [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
-Several piscivorous bird species have been documented using Deadwater Slough including the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron *Ardea herodias*, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
-During the initial sampling effort in fall 2019, a two-person crew walked the entire reach including surrounding and upstream areas scanning for passive integrated transponder (PIT) tags that are implanted into thousands of natural- and hatchery-origin salmonids in the Upper Salmon River, annually.
-During that informal survey, 10 PIT tags were recovered near active bird nests and in an upstream anastomizing reach where herons and eagles are prevalent, suggesting that mortality may have been a result of avian predation.
-The PIT tag histories in [PTAGIS](https://www.ptagis.org/) indicate these tags were implanted into a combination of juvenile Chinook Salmon (3), Sockeye Salmon (4), and steelhead (3).
-Two of the Chinook Salmon and one steelhead were natural-origin, the remaining were of hatchery-origin.
-Avian predation contributes a major component of the total mortality for yearling Chinook Salmon in some locations in the lower Snake and Columbia rivers, particularly at hydroelectric dams and within reservoirs [@Evans2012; -@Evans2016]; however, we did not observe large colonies of piscivorous birds within the study area.
-Although we did observe Belted Kingfishers, mergansers, and Great Blue Herons at Deadwater Slough; Great Blue Herons were especially prevalent in the anastomizing reach immediately upriver of the slack water reach.
-Additionally, there is documentation of individual Double-crested Cormorants [@eBird2021] at the Deadwater Slough; however, the site is not within their breeding range, rather, it is part of a migration corridor.
-Given the current avian species known to occupy Deadwater Slough, it is unlikely that avian predation on juvenile salmonids is comparable to elsewhere in the Columbia River basin with large piscivorous bird colonies.
-Nevertheless, we hypothesize that the reservoir-like conditions and associated habitat features at the Deadwater Slough may increase the probability of avian predation on juvenile Chinook Salmon and other ESA-listed salmonids from the many piscivorous birds known to use the site.
-Future studies of predation would benefit from estimates from piscivorous avian predators.
-
 ## Management Implications
@@
 
-Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere (@Winther2020).
+Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere [@Winther2020].
 A local Northern Pikeminnow bounty program could provide monetary incentive for sportfisherman, in addition to boosting the local economy from lost revenues.
@@
 Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids that are rearing in or emigrating through Deadwater Slough, including Chinook Salmon.
-We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than 24,000 juvenile Chinook salmon, however, it is likely that this value is closer to 100,000, annually.
-The magnitude of predation by Northern Pikeminnow can reduce adult escapement between 47% and 97% and should be considered a major factor influencing the recovery ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
+We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than `r prettyNum(round(min(total_impacts$juv_Lci), -3), big.mark = ",")` juvenile Chinook salmon, however, it is likely that this value is closer to `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50], -3), big.mark = ",")`, annually.
+The magnitude of predation by Northern Pikeminnow can reduce adult escapement between `r round((min(total_impacts$adult_N) / 558) * 100)`% and `r round((max(total_impacts$adult_N) / 558) * 100)`% and should be considered a major factor influencing the recovery ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
 
@@
 ```
+
+\newpage
+
+# Supplementary Materials
+
+Although not formally assessed in this study, avian predators are another potential source of mortality for juvenile Chinook Salmon and other ESA-listed salmonids in the Deadwater Slough.
+The Deadwater Slough is recognized as an important bird watching and nesting area due to the associated upstream braided channels and riparian and backwater habitats [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
+Several piscivorous bird species have been documented using Deadwater Slough including the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron *Ardea herodias*, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
+During the initial sampling effort in fall 2019, a two-person crew walked the entire reach including surrounding and upstream areas scanning for passive integrated transponder (PIT) tags that are implanted into thousands of natural- and hatchery-origin salmonids in the Upper Salmon River, annually.
+During that informal survey, 10 PIT tags were recovered near active bird nests and in an upstream anastomizing reach where herons and eagles are prevalent, suggesting that mortality may have been a result of avian predation.
+The PIT tag histories in [PTAGIS](https://www.ptagis.org/) indicate these tags were implanted into a combination of juvenile Chinook Salmon (3), Sockeye Salmon (4), and steelhead (3).
+Two of the Chinook Salmon and one steelhead were natural-origin, the remaining were of hatchery-origin.
+Avian predation contributes a major component of the total mortality for yearling Chinook Salmon in some locations in the lower Snake and Columbia rivers, particularly at hydroelectric dams and within reservoirs [@Evans2012; -@Evans2016]; however, we did not observe large colonies of piscivorous birds within the study area.
+Although we did observe Belted Kingfishers, mergansers, and Great Blue Herons at Deadwater Slough; Great Blue Herons were especially prevalent in the anastomizing reach immediately upriver of the slack water reach.
+Additionally, there is documentation of individual Double-crested Cormorants [@eBird2021] at the Deadwater Slough; however, the site is not within their breeding range, rather, it is part of a migration corridor.
+Given the current avian species known to occupy Deadwater Slough, it is unlikely that avian predation on juvenile salmonids is comparable to elsewhere in the Columbia River basin with large piscivorous bird colonies.
+Nevertheless, we hypothesize that the reservoir-like conditions and associated habitat features at the Deadwater Slough may increase the probability of avian predation on juvenile Chinook Salmon and other ESA-listed salmonids from the many piscivorous birds known to use the site.
+Future studies of predation would benefit from estimates from piscivorous avian predators.

### 2022-04-04 analysis/paper/paper.Rmd
draft edits complete

@@
   ggplot(aes(x = perc_diet,
-             y = N,
+             y = N / 1000,
              group = as.factor(perc_chnk),
@@
         legend.box = 'horizontal') +
-  labs(y = "Estimated Juvenile Chinook Consumed", 
+  labs(y = "Juvenile Chinook Consumed (x 1,000)", 
        x = "Percent Fish in Pikeminnow Diet")

### 2022-04-07 analysis/paper/paper.Rmd
some final edits to Intro

@@
 In this study, we examine the abundance of Northern Pikeminnow in a slow-water reach of the Salmon River, Idaho, known as Deadwater Slough, and consider their potential impacts on local Chinook Salmon populations.
-The Deadwater Slough contains historically important overwinter rearing habitat and is a migration corridor for Chinook Salmon as well as other Endangered Species Act-listings including steelhead and Sockeye Salmon.
-We estimated the abundance of Northern Pikeminnow in Deadwater Slough to be 19,499 in the fall and 10,352 in the spring, corresponding with the peak emigrations of juvenile Chinook Salmon.
+The Deadwater Slough is within a migration corridor for several Chinook Salmon populations as well as other Endangered Species Act-listings including steelhead and Sockeye Salmon.
+We estimated the abundance of Northern Pikeminnow in Deadwater Slough, an approximately 1.9 kilometer reach of the Salmon River, to be 19,499 in the fall and 10,352 in the spring, corresponding with the peak emigrations of juvenile Chinook Salmon.
 Using these abundance values, we estimated Northern Pikeminnow consumption of juvenile Chinook Salmon.
 Assuming 60% of the Northern Pikeminnow diet is fish, of which 50% is juvenile Chinook Salmon, we estimated the Northern Pikeminnow population can consume 61,409 juvenile Chinook Salmon, annually.
-Using a sensitivity analysis, Northern Pikeminnow predation on juvenile Chinook Salmon was predicted to reduce adult returns by 377 (95% CI: 161 - 935).
-Given the relatively small size of the Salmon River, the large densities and potential consumption rates of Northern Pikeminnow in Deadwater Slough suggest predation likely has consequential impacts on Chinook Salmon in the Upper Salmon River major population group.
+Using a sensitivity analysis, estimated Northern Pikeminnow predation on juvenile Chinook Salmon was equivalent to 377 (95% CI: 161 - 935) returning adults.
+Given the relatively small size of the Salmon River, the large densities and potential consumption rates of Northern Pikeminnow in Deadwater Slough suggest predation likely has consequential impact on Chinook Salmon recovery in the Upper Salmon River.
 
@@
 Predation on emigrating juveniles by piscivorous fishes is another important, and potentially under-estimated, source of salmonid mortality.
-In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of recovery [@Fresh2003; @Winther2020].
+In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of recovery efforts [@Friesen1999; @Tabor1993; @Beamesderfer1996].
 Dams and reservoirs in the Columbia River are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995].
@@
 One such example is Deadwater Slough, an approximately `r round(ddw_rkms, 1)` km long reach of unnaturally slow and deep water in the Salmon River, Idaho.
-The Deadwater Slough contains historically important overwinter rearing habitat for juvenile Chinook Salmon *O.
+The Deadwater Slough is within a section of the Salmon River containing historically important overwinter rearing habitat for juvenile Chinook Salmon *O.
 tshawytscha* and is part of the migratory pathway for upstream Chinook Salmon, the endangered Snake River Sockeye Salmon *O.
@@
 The Salmon River was historically the most productive tributary for Chinook Salmon in the Columbia River basin [@Nemeth1999], but those populations have become depleted in recent decades.
-Therefore, quantifying the mortality of spring/summer-run Chinook Salmon (hereafter Chinook Salmon) in the Upper Salmon major population group (MPG) associated with piscine predation is of particular interest.
-Currently, the Upper Salmon MPG supports eight independent, extant populations; this includes Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River populations [@NOAA2017].
-The Lemhi River historically contained the largest Chinook Salmon population within the MPG and is therefore a prioritized candidate for restoration of natural processes [@Zimmerman2012].
-
-The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a].
+Therefore, quantifying the mortality of spring/summer-run Chinook Salmon (hereafter Chinook Salmon) in the Upper Salmon River associated with piscine predation is of particular interest.
+Importantly, the Salmon River supports eight extant populations of Chinook Salmon upstream of Deadwater Slough [@NOAA2017].
+This includes the Lemhi River, which was historically the largest population, and is a prioritized candidate for restoration of natural processes to increase production of juvenile Chinook Salmon [@Zimmerman2012].
+
+Chinook Salmon in the Upper Salmon River are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a].
 The DSR migrants leave their natal area as subyearlings between June and November and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts.
@@
 Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency [@Dodson2013].
-Deadwater Slough represents an important habitat for the Upper Salmon MPG as it supports rearing and migration of juveniles from all populations including fall DSR and spring NRR migrants.
+Deadwater Slough represents an important habitat for Chinook Salmon in the Upper Salmon River as it supports rearing and migration of juveniles from all populations including fall DSR and spring NRR migrants.
 
 Recent studies have examined the downstream movement, distribution, and apparent survival (hereafter 'survival') of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach.
-Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have substantially reduced survival in the Deadwater Slough reach [@Axel2015].
+Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have 32% lower survival in the Deadwater Slough reach compared to directly adjacent reaches [@Axel2015].
 Similarly, emigrating DSR Chinook Salmon had an approximate 10% reduction in transition probability through the Deadwater Slough compared to surrounding reaches during fall and early winter months [@Ackerman2018; @Porter2019].
@@
 3. Use an established bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population on DSR and NRR Chinook Salmon emigrants at Deadwater Slough;
-4. Estimate how consumption of juvenile Chinook Salmon emigrants at Deadwater Slough by Northern Pikeminnow may impact adult returns to the Upper Salmon MPG.
+4. Estimate how consumption of juvenile Chinook Salmon emigrants at Deadwater Slough by Northern Pikeminnow may impact adult returns to the Upper Salmon River.
 

### 2022-04-07 analysis/paper/paper.Rmd
near final edits to Methods

@@
 The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River.
-Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event and upslope instability that deposited substantial amounts of sediment at its confluence with the Salmon River [@Emerson1973].
-That sediment formed an alluvial fan which created a slow, deep section in the river, spanning approximately `r round(ddw_hectares)` hectares, averaging `r round(ddw_avg_width_m)` m wide, and resembling a small reservoir.
+A large alluvial fan at the mouth of Dump Creek has, at least partially, created a hydraulic control in the Salmon River resulting in the formation of Deadwater Slough.
+Although the origin and timing of the alluvial fan and Deadwater Slough is somewhat ambiguous [@Reichmuth1958; @USACE1986], they likely coincided with the failure of a small mining reservoir in the Dump Creek drainage around 1897 that deposited substantial amounts of sediment at its confluence with the Salmon River [@Emerson1973].
+The slough is currently a slow, deep section in the river, spanning approximately `r round(ddw_hectares)` hectares, averaging `r round(ddw_avg_width_m)` m wide, and has characteristics resembling a small reservoir.
 
@@
 Over this period, an additional fall survey occurred during the 2020 DSR emigration.
-Fall surveys were constrained to two weeks to minimize Northern Pikeminnow immigration/emigration.
+Fall surveys were constrained to two weeks to minimize Northern Pikeminnow immigration and emigration.
 
@@
 where $M$, $n$, and $m$ are indexed by each survey (day), $i$.
-The 95% confidence intervals were calculated using a Poisson approximation [@Krebs1999].<!--The Schnabel estimator is essentially the weighted average of a series of Chapman-modified Lincoln-Peterson estimators.-->
+The 95% confidence intervals of the Schnabel estimator were calculated using a Poisson approximation [@Krebs1999].<!--The Schnabel estimator is essentially the weighted average of a series of Chapman-modified Lincoln-Peterson estimators.-->
 
@@
 If release sites for marked fish were not proximal to preferred angling locales, recently released fish may have been less susceptible to immediate capture than unmarked fish.
-To account for this potential bias, we included a "Delayed-mixing" Schnabel estimator that assumed fish marked and released on a given survey day were not available for recapture for 48 hours.
-The "Delayed-mixing" Schnabel estimator, is given by:
+To account for this potential bias, we included an adjusted "delayed-mixing" Schnabel estimator that assumed fish marked and released on a given survey day were not available for recapture for 48 hours.
+The delayed-mixing Schnabel estimator, is given by:
 
@@
 $$
-All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the survey period), (2) all fish have equal chance of being caught in subsequent survey days, (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or misidentified.
+All mark-recapture abundance estimators include a Chapman correction for small sample sizes [@Chapman1951] and assumed: (1) the population is closed (no immigration, emigration, births or deaths during the survey period), (2) all fish have equal chance of being caught in subsequent survey days, (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or misidentified.
 
@@
 The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on predator and prey energy densities, predator start and end weights, and water temperatures.
-Separate models were run for the fall period (September 15 - November 30) and the spring period (March 1 - May 31) to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG.
+Separate models were run for the fall period (September 15 - November 30) and the spring period (March 1 - May 31) to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon River.
 
@@
 
-This resulted in average fall and spring starting weights of `r round(avg_fall_NPM_g, 1)` and `r round(avg_spring_NPM_g, 1)` g for Northern Pikeminnow with average TLs of `r round(avg_fall_NPM_TL, 1)` and `r round(avg_spring_NPM_TL, 1)` mm, respectively.
+This resulted in average fall DSR and spring NRR starting weights of `r round(avg_fall_NPM_g, 1)` and `r round(avg_spring_NPM_g, 1)` g for Northern Pikeminnow with average TLs of `r round(avg_fall_NPM_TL, 1)` and `r round(avg_spring_NPM_TL, 1)` mm, respectively.
 We assumed no growth in individual Northern Pikeminnow, resulting in equal start and end weights.
@@
 These values were supported by observations elsewhere in the Columbia River basin (54-86%: @Shively1996; 48-86%: @Zimmerman1999; 37%: @Gray2001) that suggest fish are the majority of Northern Pikeminnow prey by volume.
-A model run was conducted for each combination of diet scenario and season, resulting in fourteen estimates of the total grams of fish consumed by an individual Northern Pikeminnow (Table \@ref(tab:bio-tbl)).
+A model run was conducted for each combination of diet scenario and season, resulting in fourteen estimates of the total grams of fish consumed by an individual Northern Pikeminnow.
 To estimate the total biomass of fish consumed by Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated Northern Pikeminnow population sizes during fall and spring.
@@
 The average weights of DSR and NRR emigrants were calculated from fish captured at seven rotary screw traps upstream of Deadwater Slough during the fall and spring periods.
-We also quantified the potential impact of Northern Pikeminnow predation on juvenile Chinook Salmon to adult returns by estimating the number of additional adult Chinook Salmon expected to return to Lower Granite Dam if predation by Pikeminnow was reduced or eliminated in the Deadwater Slough.
-We multiplied the estimated total juvenile Chinook Salmon consumed by the median Granite-to-Granite smolt-to-adult return rate (SAR) of 0.00614 (SD = 0.00051) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.
+Although the primary impact to Chinook Salmon in Deadwater Slough by Northern Pikeminnow is juvenile predation, the common metric to evaluate salmon recovery is adult returns.
+We quantified the potential impact of Northern Pikeminnow predation on adult returns by estimating the number of equivalent adults expected to return to Lower Granite Dam if predation by Pikeminnow was eliminated in the Deadwater Slough.
+To accomplish this, we multiplied the estimated total juvenile Chinook Salmon consumed by the median Granite-to-Granite smolt-to-adult return rate (SAR) of 0.00614 (SD = 0.00051) from @McCann2019 for Chinook Salmon in the Upper Salmon River.
 
@@
 
-All data and code for the analyses presented here can be found in a GitHub repository at ***zenodo DOI*** here.
+All data and code for the analyses presented here can be found in a GitHub repository at ***zenodo DOI***.
 

### 2022-04-07 analysis/paper/paper.Rmd
near final edits of Discussion

@@
 
-Using hook-and-line angling, we caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 surveys, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish.
-The CPUE was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021.
-During the mark-recapture events, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019, including `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals and `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020, including `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals.
-Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn).
-
+Using hook-and-line angling, we caught a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` Northern Pikeminnow over the course of the study including `r sum(N_mods[[3]][[1]]$m) + sum(N_mods[[3]][[3]]$m)` recaptures; the overall CPUE was `r round(sum(cpue_df$n_fish) / (sum(cpue_df$total_effort_min) / 60), 2)` Northern Pikeminnow per angler hour (table reference needed).
 Mark-recapture abundance estimates of Northern Pikeminnow in Deadwater Slough ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` in fall 2019 and from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` in fall 2020 (Table \@ref(tab:abund-tab)).
-We estimated larger populations of Northern Pikeminnow using the unadjusted multiple-census estimator compared to adjusted and single census estimators (Figure \@ref(fig:abund-fig)).
-Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, we deemed the Schnabel estimates most appropriate.
-Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r prettyNum(round(mean(all_est$Lci[all_est$model == "Schnabel"])), big.mark = ",")` - `r prettyNum(round(mean(all_est$Uci[all_est$model == "Schnabel"])), big.mark = ",")`).
-Using the delayed-mixing Schnabel estimator, the mean fall abundance estimate was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r prettyNum(mean(all_est$Lci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` - `r prettyNum(mean(all_est$Uci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")`).
+We estimated larger populations of Northern Pikeminnow using the unadjusted multiple-census estimator compared to the single census and adjusted delayed-mixing multiple census estimators (Figure \@ref(fig:abund-fig)).
+Our sampling design most closely matched a multiple-census estimator; therefore, the Schnabel estimates were considered most appropriate.
+Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r prettyNum(round(mean(all_est$Lci[all_est$model == "Schnabel"])), big.mark = ",")` - `r prettyNum(round(mean(all_est$Uci[all_est$model == "Schnabel"])), big.mark = ",")`) using the unadjusted Schabel estimator.
+Using the delayed-mixing Schnabel estimator the mean fall abundance estimate was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r prettyNum(mean(all_est$Lci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` - `r prettyNum(mean(all_est$Uci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")`).
+All subsequent analyses use results from the adjusted delayed-mixing Schnabel estimator.
 For spring 2021, we estimated `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough.
@@
 We found that gastric lavage successfully removed food items in all dissected individuals, indicating the method was effective.
-Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` stomachs.
+Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` of those.
 Northern Pikeminnow captured during spring 2021 had a higher proportion of fish content in their diet compared to individuals captured in the fall surveys (Table \@ref(tab:gastric-tab)).
@@
 
-During the fall DSR emigration, we estimate an average-size Northern Pikeminnow (`r round(avg_fall_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"]` g of fish to maintain their body size, assuming 60% of their diet consists of fish prey (Figure \@ref(fig:bio-p)).
+During the fall DSR emigration, we estimated an average-size Northern Pikeminnow (`r round(avg_fall_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"]` g of fish to maintain their body size, assuming 60% of their diet consists of fish prey (Figure \@ref(fig:bio-p)).
 Fall consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f30_g"]` g for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f90_g"]` g for a diet of 90% fish.
@@
 Spring consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f30_g"]` for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f90_g"]` for a diet of 90% fish.
-Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` in the fall, `r prettyNum(round(spring_abund$N_S), big.mark = ",")` in the spring 2021, and a diet consisting of 60% fish, the Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
+Daily consumption rates were higher early in the fall and late in the spring and corresponded with water temperatures above 8 degrees Celsius (Figure \@ref(fig:bio-p)).
+Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` in the fall and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` in the spring 2021, and a diet consisting of 60% fish, the Northern Pikeminnow population is estimated to consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
 
@@
 
-Given yearly fish consumption estimates during the fall DSR and spring NRR periods, we produced a sensitivity analysis to calculated consumption of juvenile Chinook Salmon where 30%, 50%, and 65% of total fish prey consumed are juvenile Chinook Salmon (Figure \@ref(fig:juv-con-p)).
+Given yearly fish consumption estimates during the fall DSR and spring NRR periods, we produced a sensitivity analysis to calculate the consumption of juvenile Chinook Salmon where 30%, 50%, and 65% of total fish prey consumed are juvenile Chinook Salmon (Figure \@ref(fig:juv-con-p)).
 At the median value of 50%, we estimated that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon will be consumed.
-Using the median Granite-to-Granite SAR, we estimated that juvenile consumption is equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) adults (Figure \@ref(fig:adult-imp-p)).
-
-# Discussion
-
-## Mark-Recapture Models
+Using the median Granite-to-Granite SAR for Chinook Salmon in the Upper Salmon River, we estimated the "adult equivalents" of juveniles consumed to be `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) adults (Figure \@ref(fig:adult-imp-p)).
 
@@
 
+# Discussion
+
 All mark-recapture abundance estimators suggest a large population of Northern Pikeminnow occupies Deadwater Slough.
-The mean population size of Northern Pikeminnow was estimated to be `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` during the fall DSR emigration and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` during the spring NRR emigration.
+The mean population size of Northern Pikeminnow was estimated to be `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` during the fall DSR emigration and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` during the spring NRR emigration even after accounting for the potential of delayed mixing of marked fish.
 Those estimates translate to linear densities (fish/rkm) that are `r round((mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms) / john_day_dens)` and `r round((spring_abund$N_S / ddw_rkms) / john_day_dens)` times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r round((mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms) / low_columbia_fish_km)` and `r round((spring_abund$N_S / ddw_rkms) / low_columbia_fish_km)` times greater than projections for the lower Columbia River [@Beamesderfer1996].
-@Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective; therefore it is possible that their estimates may have been low.
+@Beamesderfer1991 acknowledged that their methods were unsuitable for sampling offshore, noting that water velocity, depth, and irregular bottom contours, and barge traffic made sampling offshore ineffective; therefore it is possible that their estimates may have been low.
 Conversely, at Deadwater Slough the maximum depth was approximately 6 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach.
@@
 Several assumptions in our estimators may have influenced the magnitude of our abundance results.
-First, all models assumed a closed population, however, it is possible that immigration and emigration occurred.
-Nevertheless, our estimators still provide unbiased estimates of abundance assuming the immigration and emigration rates are equal between marked and unmarked fish.
-Given the size of the sampling area, and the short duration of our surveys, the closed population assumption is likely met.
-<!--We also caught one individual that was confirmed tagged in Deadwater Slough in a previous year suggesting that at least some individuals remain in the reach or at least migrate to Deadwater Slough multiple times, annually.-->This assumption can also affect results if mortality occurs for some marked fish released back to the population.
-During field processing, no mortalities or injury post release was observed; however, the potential for mortality amongst marked and released fish cannot be discounted.
-Next, our estimates assumed that capture and recapture events are random samples of the population.
-Angling methods may have a size selection bias, thereby limiting our abundance estimates to a size range susceptible to angling, resulting in a conservative estimate of the total population.
+First, all models assumed a closed population; however, it is possible that immigration and emigration occurred.
+Nevertheless, our estimators still provide unbiased estimates of abundance assuming the immigration and emigration rates are equal between marked and marked fish.
+Emigration of marked individuals would reduce the marking fraction in the population resulting in an upward bias of abundance estimate, but emigration rates would need to be substantial.
+Further, the multiple census estimators are more robust to this assumption because the marking fraction is estimated daily.
+We additionally have no reason to believe that emigration occurs during the two-week window as Deadwater Slough provides favorable habitat to Northern Pikeminnow relative to adjacent reaches.
+Given the size of the sampling area and the short duration of our surveys, the closed population assumption is likely met.
+This assumption can also affect results if mortality for some marked fish released back to the population.
+During field processing, no mortalities or injury post release was observed and fish injured prior to release were retained.
+Nevertheless, the potential for mortality amongst marked and released fish cannot be discounted.
+
+Next, our abundance estimates assumed that capture and recapture events are random samples of the population.
+Angling methods often have a size selection bias, thereby limiting abundance estimates to a size range susceptible to angling and resulting in a conservative estimate of the total population.
 We also question the validity of our assumption of equal catchability of individual fish between sampling events.
-During our study, we recaptured two individual fish during multiple days within a survey indicating that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.--> However, we suspect that marked and released fish may not have mixed thoroughly back into the population prior to recapture events.
-Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit "good" fishing locales.
+Although we recaptured two individual fish during multiple days within a survey (indicating that marked fish were still susceptible to angling), we suspect that marked and released fish may not have immediately mixed thoroughly back into the population.
+Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit “good” fishing locales.
 Violations of the assumption of equal catchability between marked and unmarked fish may lead to an overestimate of abundance of Northern Pikeminnow in our study if marked fish have lower catchability.
@@
 For example, if marked fish were only 70% as likely to be caught as unmarked fish, the true abundance is closer to 70% of our abundance estimate.
-
-The spring abundance estimate was smaller than fall abundance estimates.
-Although our approach assumed equal capture probabilities between the fall and spring surveys, we believe higher spring flows may result in lower capture probability.
+This would additionally account for marked and released fish being “hook shy” for a period of time.
+The adjusted delayed-mixing Schnabel estimator may account for these biases; however the magnitude of bias is unknown.
+
+The spring abundance estimate was smaller than fall abundance estimates due to the spring CPUE being approximately half of the mean fall CPUE.
+Although our approach assumed equal capture probabilities between the fall and spring surveys, we believe that higher spring flows may result in lower capture probability of Northern Pikeminnow.
 Therefore, we believe our spring abundance estimate to be conservative.
 
-In addition to the large population, the observed (`r round(psd)`%) PSD in this study was large compared to observations of 41% below Bonneville Dam and 18% in Bonneville Reservoir [@Winther2020].
+In addition to the large population abundance, the observed (`r round(psd)`%) PSD in this study was greater than observations of 41% below Bonneville Dam and 18% in Bonneville Reservoir [@Winther2020].
 This suggests that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to populations reported elsewhere in the Columbia River.
@@
 
-## Stomach Contents
-
 We performed gastric lavage on nearly all Northern Pikeminnow collected during this study.
 Observed prey species included juvenile Chinook Salmon as well as Redside Shiner *Richardsonius balteatus*, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*, though decomposition from digestion rendered most fish contents unidentifiable.
-Although our approach provided useful information on prey species, using angling and gastric lavage to quantify diet composition in Northern Pikeminnow can be problematic.
+Although our approach provided useful information on prey species, the use of angling and gastric lavage to quantify diet composition in Northern Pikeminnow can be problematic.
 First, diets of angled fish can differ from those collected by other methods such as electrofishing [@Hodgson1988].
-For instance, angled fish are more likely to have empty stomachs [@Jurajda2016]; possibly because they are captured while actively searching for food, suggeting they are hungry.
+For instance, angled fish are more likely to have empty stomachs [@Jurajda2016]; possibly because they are captured while actively searching for food, suggesting they are hungry.
 Second, unlike other predators such as Smallmouth Bass and Walleye, that are commonly sampled using gastric lavage [@Kamler2001], Northern Pikeminnow are cyprinids and therefore lack a true stomach.
 Consequently, performing gastric lavage on cyprinid species can be ineffective for accurately quantifying diet composition [@Hartleb1995], or alternately, requires a modification in methodology to flush food items out the vent of the fish [@Wasowicz1994].
-Therefore, we consider our diet composition estimates and subsequent consumption estimates to be conservative.
-
-## Bioenergetics
-
-Our approach assumed that all available prey consumed by Northern Pikeminnow have energy densities equal to juvenile Chinook Salmon.
-Generalizing energy densities using a single species is a common approach [@Petersen1999], however, we recognize that other salmonids including juvenile steelhead and Sockeye Salmon are likely also available prey in Deadwater Slough, especially during the spring migration.
+Therefore, we consider our diet composition estimates from gastric lavage to be conservative.
+
+Our bioenergetics approach assumed that all available prey consumed by Northern Pikeminnow have energy densities equal to juvenile Chinook Salmon.
+Generalizing energy densities using a single species is a common approach [@Petersen1999].
+Other salmonids including juvenile steelhead and Sockeye Salmon are likely also available prey in Deadwater Slough, especially during the spring migration, and so generalizing energy densities among salmonids may not be problematic.
 Additionally, hatchery smolt releases (Chinook Salmon, steelhead, and Sockeye Salmon) are also prevalent in the mainstem Salmon River during the spring migration.
-To address the prevalence of other fish in the Northern Pikeminnow's diet, we considered scenarios where DSR and NRR Chinook Salmon represent only 30%, 50%, and 65% of total fish prey consumed.
+To address the prevalence of other fish in the Northern Pikeminnow’s diet, we considered scenarios where DSR and NRR Chinook Salmon represent only 30%, 50%, and 65% of total fish prey consumed.
 These values may be considered conservative during the peak emigration periods when Chinook Salmon are abundant and the Northern Pikeminnow diet shifts almost entirely to piscivory, presumably on juvenile Chinook Salmon [@Shively1996; @Poe1991].
-This is likely the case during the fall when DSR Chinook Salmon are the dominant prey species available in Deadwater Slough as 1) no (or at least minimal) hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent than Chinook Salmon emigration.
-During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river, <50% Chinook Salmon in the Pikeminnow's diet may be more likely.
-Future work to quantify and identify the juvenile salmonid in Deadwater Slough during their seasonal migration would be useful to validate the diet composition assumptions used in our model.
-
-## Potential Impacts to Salmon Populations
-
-Recovery of Chinook Salmon in the Upper Salmon MPG is desired to support local fisheries and economies.
+This is likely the case during fall months when DSR Chinook Salmon are the dominant prey species available in Deadwater Slough as 1) few to no hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent than Chinook Salmon.
+During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river, less than 50% Chinook Salmon in the Northern Pikeminnow’s diet may be more likely.
+Predation on juveniles from any of the ESA-listed salmonid species is detrimental to their recovery.
+Future work to quantify and identify juvenile salmonids in Deadwater Slough during their seasonal migration would be useful to validate the diet composition assumptions used in our model and to understand impacts to local populations.
+
+Chinook Salmon populations above Deadwater Slough are within the Upper Salmon major population group (MPG) which supports eight independent, extant populations including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017].
+Recovery of the MPG is desired to support local fisheries and economies.
 At least five of the eight populations must meet criteria set forth by @McElhany2000 and the @ICTRT2007 for the MPG to be considered viable and for recovery of the Snake River Evolutionary Significant Unit.
-We estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` juvenile Chinook Salmon may be consumed by Northern Pikeminnow assuming that 60% of their diet is fish and 50% of fish prey are Chinook Salmon.
+We estimated that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` juvenile Chinook Salmon may be consumed by Northern Pikeminnow assuming that 60% of their diet is fish and 50% of fish prey are Chinook Salmon.
+For context, this is approximately `r round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 45416, 2)` times the average annual combined DSR and NRR emigration estimates for brood years 2011 - 2018 at a rotary screw trap located in the lower Lemhi River [@McClure2021; @Feeken2020; @Poole2019], the largest population in the Upper Salmon MPG.
+Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of fish prey are Chinook Salmon, we estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30]), big.mark = ",")` juvenile Chinook Salmon may be consumed which is `r round((total_impacts$juv_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 45416) * 100)`% of the total DSR and NRR emigration for brood years 2011 - 2018 from the Lemhi River.
 Considering that diet scenarios used in our study are likely conservative assumptions, especially during the fall DSR emigration, the estimated number of Chinook Salmon consumed is substantial.
 
-Spawner abundance is perhaps the most important metric considered in determining a population's viability and productivity.
-We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by `r round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` adults, annually, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean adult escapement estimated<!--of 558--> for the Upper Salmon MPG in 2017 - 2019 [@Kinzer2020].
-Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of that is juvenile Chinook Salmon, we estimated `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` "adult equivalents" would be consumed, which is approximately `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% of recent natural-origin adult Chinook Salmon escapements.
-In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.
-Accordingly, we surmise that adult returns of steelhead and Sockeye Salmon are also affected by predation, including hatchery populations, which provide recreational fishing opportunities.
-Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including other ESA-listed species.
-Because the Deadwater Slough is part of the migratory pathway for multiple species and populations of emigrating salmonids, the positive impact could be greater than individual tributary rehabilitation actions which typically benefit a single population.
-
-## Management Implications
-
-The Deadwater Slough is a favorable candidate for management or restoration actions to benefit the local Chinook Salmon populations.
-Two potential management actions could reduce predation at Deadwater Slough: 1) removing the Dump Creek alluvial fan and 2) a local Northern Pikeminnow bounty program to encourage harvest in Deadwater Slough aimed at reducing the predator population size.
-Removing the Dump Creek alluvial fan has the benefit of restoring the natural fluvial processes in the Salmon River that likely existed in the reach prior to the failure of the mining reservoir dam in Dump Creek.
+Spawner abundance is perhaps the most important metric considered in determining a population’s viability and productivity.
+We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough is equivalent to `r round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` adults, annually, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean spawner escapement for the Upper Salmon MPG in 2017 – 2019 [@Kinzer2020].
+In the most conservative scenario, we estimated juvenile consumption equivalent to `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` adults, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% of recent natural-origin adult Chinook Salmon returns.
+In the least conservative scenarios, adult equivalents approached recent escapements for the entire Upper Salmon MPG.
+The sensitivity analysis used some simplifying assumptions.
+For example, it assumed no juvenile mortality between Deadwater Slough and Lower Granite Dam which is an unrealistic assumption.
+In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including both natural-origin juveniles and hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.
+Accordingly, we surmise that adult returns of steelhead and Sockeye Salmon are also affected by predation, including hatchery populations, which provide for recreational fishing opportunities in the Upper Salmon River.
+Consequently, reducing predation mortality in Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including other ESA-listed species.
+Because Deadwater Slough is part of the migratory pathway for multiple species and populations of emigrating salmonids, the positive impact could be greater than individual tributary rehabilitation actions which typically benefit a single population.
+
+Deadwater Slough is a favorable candidate for management or restoration actions to benefit local Chinook Salmon populations.
+Three potential management actions could reduce predation at Deadwater Slough: 1) removing or reducing the Dump Creek alluvial fan, 2) a local Northern Pikeminnow bounty program to encourage harvest in Deadwater Slough aimed at reducing the predator population size, and 3) adding structure or cover within Deadwater Slough to provide refuge for juvenile salmonids with the goal or reducing predation rates.
+Reducing or removing the Dump Creek alluvial fan has the benefit of restoring natural fluvial processes in the Salmon River that likely existed in the reach prior to the formation or increase in size of the alluvial fan related to human activities in upper Dump Creek.
 Restoring natural processes could reduce Northern Pikeminnow densities to levels similar to upstream and downstream reaches where higher survival and transition probabilities for juvenile salmon have been observed compared to Deadwater Slough [@Axel2015; @Ackerman2018; @Porter2019].
-Managers ought also to consider the feasibility and net benefit of restoring fluvial processes relative to potential losses for recreational fishing and bird watching opportunities.
-
-Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere [@Winther2020].
+Managers ought also to consider the feasibility and net benefit of restoring fluvial processes relative to potential losses for recreational fishing and bird watching opportunities (https://www.audobon.org/important-bird-areas/deadwater-slough).
+
+Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids in the Columbia River [@Winther2020].
 A local Northern Pikeminnow bounty program could provide monetary incentive for sportfisherman, in addition to boosting the local economy from lost revenues.
@@
 Each approach would require continual/annual effort and harvest to suppress the population size and as such, appropriate funding would be needed to support the program.
-
-## Conclusions
-
-The failure of a mining reservoir dam in the late 19th century created an alluvial fan at the confluence of the Salmon River and Dump Creek which has backed up the Salmon River and created favorable conditions for Northern Pikeminnow.
-Even under the most conservative scenarios, we estimate more than 10,000 Northern Pikeminnow occupy Deadwater Slough, resulting in a remarkably high density given the size of the area.
-Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids that are rearing in or emigrating through Deadwater Slough, including Chinook Salmon.
-We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than `r prettyNum(round(min(total_impacts$juv_Lci), -3), big.mark = ",")` juvenile Chinook salmon, however, it is likely that this value is closer to `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50], -3), big.mark = ",")`, annually.
-The magnitude of predation by Northern Pikeminnow can reduce adult escapement between `r round((min(total_impacts$adult_N) / 558) * 100)`% and `r round((max(total_impacts$adult_N) / 558) * 100)`% and should be considered a major factor influencing the recovery ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
+Finally, a more passive approach could be to provide cover or structure within Deadwater Slough that provides refugia for juvenile salmonids during their emigration.
+Cover should be appropriately sized to provide concealment to juvenile fishes while reducing access by larger fishes like Northern Pikeminnow.
+Adding cover could be a cost-effective approach; however, its potential effectiveness is unclear.
+
+Although the origins of Deadwater Slough are somewhat ambiguous [@Reichmuth1985; @USACE1986], it is likely that human activities in Dump Creek appear to have either exacerbated or contributed to its formation and in turn created favorable conditions for Northern Pikeminnow.
+Among the three surveys, our lowest population abundance estimate was greater than 10,000 Northern Pikeminnow that occupy Deadwater Slough, suggesting a remarkably high density given the size of the area.
+The slow water velocity and lack of cover for fish also create conditions where juvenile salmonids that are rearing in or emigrating through Deadwater Slough, including Chinook Salmon, are susceptible to predation.
+We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than `r prettyNum(round(min(total_impacts$juv_Lci), -3), big.mark = ",")` juvenile Chinook Salmon (the lower endpoint of the confidence interval for the most conservative scenario); however, it is likely that this value is closer to `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50], -3), big.mark = ",")`, annually.
+The magnitude of predation by Northern Pikeminnow is equivalent to `r round((min(total_impacts$adult_N) / 558) * 100)`% and `r round((max(total_impacts$adult_N) / 558) * 100)`% of recent adult escapement.
+Deadwater Slough predates the ESA-listing of Chinook Salmon populations in the Upper Salmon MPG in the 1990s and is therefore unlikely to be the primary cause for the population’s decline.
+However, predation by Northern Pikeminnow in Deadwater Slough and elsewhere should be considered among limiting factors hindering recovery efforts of ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
 
@@
 Special thanks to Jared Barker, Brian Hamilton, Tulley Mackey, and Chelsea Welke for their help in the field.
-We further appreciate the administrative support and guidance from Caitlin Alcott and Inter-Fluve.
+We further appreciate the administrative support and guidance from staff at Inter-Fluve.
 This manuscript benefited from reviews by Sean Gibbs, Mark Roes, and Ian Courter.

### 2022-04-07 analysis/paper/paper.Rmd
merged all of TB 20220406 edits

@@
 
-Using hook-and-line angling, we caught a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` Northern Pikeminnow over the course of the study including `r sum(N_mods[[3]][[1]]$m) + sum(N_mods[[3]][[3]]$m)` recaptures; the overall CPUE was `r round(sum(cpue_df$n_fish) / (sum(cpue_df$total_effort_min) / 60), 2)` Northern Pikeminnow per angler hour (table reference needed).
+Using hook-and-line angling, we caught a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` Northern Pikeminnow over the course of the study including `r sum(N_mods[[3]][[1]]$m) + sum(N_mods[[3]][[3]]$m)` recaptures; the overall CPUE was `r round(sum(cpue_df$n_fish) / (sum(cpue_df$total_effort_min) / 60), 2)` Northern Pikeminnow per angler hour (Table \@ref(tab:summary-tbl)).
 Mark-recapture abundance estimates of Northern Pikeminnow in Deadwater Slough ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` in fall 2019 and from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` in fall 2020 (Table \@ref(tab:abund-tab)).
@@
 <div id="tables"></div>
+
+```{r summary-tbl}
+summary_tbl = NULL %>%
+  kable(booktabs = T,
+        caption = "Summary of Northern Pikeminnow angling survey data.
+Fall surveys included mark and recapture weeks for the single census estimate.
+Spring surveys included only a single capture week to estimate catch-per-unit-effort.")
+
+```
+
+\newpage
 
@@
 
-\newpage
-
-```{r abund-data-lp}
+```{r abund-data-lp, eval = F}
 LP_mod %>%
@@
 
-\newpage
-
-```{r abund-data-schn}
+```{r abund-data-schn, eval = F}
 N_mods %>%
@@
         format.args = list(big.mark = ","),
-        caption = "Summary of gastric lavage results including counts of the total number of Northern Pikeminnow captured, gastric lavaged, individuals with stomach contents, and individuals with fish prey contents") %>%
+        caption = "Summary of gastric lavage results including counts of the total number of Northern Pikeminnow captured, gastric lavaged, individuals with stomach contents, and individuals with fish prey contents.") %>%
   kable_styling()

### 2022-04-07 analysis/paper/paper.Rmd
updated BibTex library

@@
 A large alluvial fan at the mouth of Dump Creek has, at least partially, created a hydraulic control in the Salmon River resulting in the formation of Deadwater Slough.
-Although the origin and timing of the alluvial fan and Deadwater Slough is somewhat ambiguous [@Reichmuth1958; @USACE1986], they likely coincided with the failure of a small mining reservoir in the Dump Creek drainage around 1897 that deposited substantial amounts of sediment at its confluence with the Salmon River [@Emerson1973].
+Although the origin and timing of the alluvial fan and Deadwater Slough is somewhat ambiguous [@Reichmuth1985; @USACE1986], they likely coincided with the failure of a small mining reservoir in the Dump Creek drainage around 1897 that deposited substantial amounts of sediment at its confluence with the Salmon River [@Emerson1973].
 The slough is currently a slow, deep section in the river, spanning approximately `r round(ddw_hectares)` hectares, averaging `r round(ddw_avg_width_m)` m wide, and has characteristics resembling a small reservoir.

### 2022-04-07 analysis/paper/paper.Rmd
updated draft in repo

@@
 A large alluvial fan at the mouth of Dump Creek has, at least partially, created a hydraulic control in the Salmon River resulting in the formation of Deadwater Slough.
-Although the origin and timing of the alluvial fan and Deadwater Slough is somewhat ambiguous [@Reichmuth1985; @USACE1986], they likely coincided with the failure of a small mining reservoir in the Dump Creek drainage around 1897 that deposited substantial amounts of sediment at its confluence with the Salmon River [@Emerson1973].
+Although the origin and timing of the alluvial fan and Deadwater Slough is somewhat ambiguous [@Reichmuth1985; @USACE1986], some believe they coincided with the failure of a small mining reservoir in the Dump Creek drainage around 1897 that deposited substantial amounts of sediment at its confluence with the Salmon River [@Emerson1973].
 The slough is currently a slow, deep section in the river, spanning approximately `r round(ddw_hectares)` hectares, averaging `r round(ddw_avg_width_m)` m wide, and has characteristics resembling a small reservoir.
@@
 To account for this potential bias, we included an adjusted "delayed-mixing" Schnabel estimator that assumed fish marked and released on a given survey day were not available for recapture for 48 hours.
-The delayed-mixing Schnabel estimator, is given by:
+The delayed-mixing Schnabel estimator is given by:
 
@@
 
-We obtained an estimate of the number of juvenile Chinook Salmon consumed by dividing total biomass consumed by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG.
+We obtained an estimate of the number of juvenile Chinook Salmon consumed by dividing total biomass consumed by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon River.
 The average weights of DSR and NRR emigrants were calculated from fish captured at seven rotary screw traps upstream of Deadwater Slough during the fall and spring periods.
@@
 ```{r summary-tbl}
-summary_tbl = NULL %>%
+summary_tbl = tibble(dummy = 1:4, bogus = 5:8) %>%
   kable(booktabs = T,
@@
 Fall surveys included mark and recapture weeks for the single census estimate.
-Spring surveys included only a single capture week to estimate catch-per-unit-effort.")
+Spring surveys included only a single capture week to estimate catch-per-unit-effort.") %>%
+  kable_styling()
+
+summary_tbl
 

### 2022-04-07 analysis/paper/paper.Rmd
updated author order

@@
       correspondence: false
+  - Tara E. Blackman:
+      email: tara.blackman@mthoodenvironmental.com
+      institute: [mhe_sandy]
+      correspondence: false
   - Michael S. Hall:
@@
       institute: [biomark]
-      correspondence: false
-  - Tara E. Blackman:
-      email: tara.blackman@mthoodenvironmental.com
-      institute: [mhe_sandy]
       correspondence: false

### 2022-04-07 analysis/paper/paper.Rmd
incorporated some final edits and saved 20220407 draft

@@
 Predation on emigrating juvenile salmonids by piscivorous fishes is a widely-studied source of mortality within reservoirs and dam tailraces in the Columbia River basin.
-Native Northern Pikeminnow have been estimated to consume ~8% of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually.
+Native Northern Pikeminnow have been estimated to consume \~8% of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually.
 Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above the reservoir systems.
@@
 Northern Pikeminnow tolerate and thrive in relatively warm, slow-water habitats [@Wydoski2003] and consequently have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999].
-Northern Pikeminnow are estimated to consume ~8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].
+Northern Pikeminnow are estimated to consume \~8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].
 
@@
 $$
+
 All mark-recapture abundance estimators include a Chapman correction for small sample sizes [@Chapman1951] and assumed: (1) the population is closed (no immigration, emigration, births or deaths during the survey period), (2) all fish have equal chance of being caught in subsequent survey days, (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or misidentified.
@@
    spring_f90)
-
 ```
@@
 @Beamesderfer1991 acknowledged that their methods were unsuitable for sampling offshore, noting that water velocity, depth, and irregular bottom contours, and barge traffic made sampling offshore ineffective; therefore it is possible that their estimates may have been low.
-Conversely, at Deadwater Slough the maximum depth was approximately 6 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach.
+Conversely, the maximum depth at Deadwater Slough was approximately 6 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach.
 Our estimated densities of Northern Pikeminnow suggest that slow-water reaches outside of the reservoir complexes on the Snake and Columbia rivers may support exceptionally high predator densities, consistent with findings of @Harnish2014, @Gray2001, and @Zimmerman1999.
-Given the dearth of current data available in the Columbia River basin, it is unclear how Northern Pikeminnow abundance estimates in this study compare to elsewhere in the Columbia River Basin; contemporary information on piscine predator abundances may be needed [@Widener2021].
+Given the dearth of current data available throughout the Columbia River basin, it is unclear how Northern Pikeminnow abundance estimates in this study compare to elsewhere in the basin; additional information on contemporary piscine predator abundances may be needed [@Widener2021].
 
@@
 First, all models assumed a closed population; however, it is possible that immigration and emigration occurred.
-Nevertheless, our estimators still provide unbiased estimates of abundance assuming the immigration and emigration rates are equal between marked and marked fish.
-Emigration of marked individuals would reduce the marking fraction in the population resulting in an upward bias of abundance estimate, but emigration rates would need to be substantial.
-Further, the multiple census estimators are more robust to this assumption because the marking fraction is estimated daily.
-We additionally have no reason to believe that emigration occurs during the two-week window as Deadwater Slough provides favorable habitat to Northern Pikeminnow relative to adjacent reaches.
+Nevertheless, our estimators still provide unbiased estimates of abundance assuming the immigration and emigration rates are equal between marked and unmarked fish.
+Emigration of marked individuals could reduce the marking fraction in the population leading to an upward bias in abundance estimates, but emigration rates would need to be substantial.
+The multiple census estimators, which we report, are more robust to this assumption because the marking fraction is estimated daily.
+We additionally have little reason to believe that emigration occurs during the two-week survey window owing to favorable habitat for Northern Pikeminnow in Deadwater Slough relative to adjacent reaches.
 Given the size of the sampling area and the short duration of our surveys, the closed population assumption is likely met.
-This assumption can also affect results if mortality for some marked fish released back to the population.
+This assumption can likewise affect results if mortality occurs for some marked fish released back to the population.
 During field processing, no mortalities or injury post release was observed and fish injured prior to release were retained.
@@
 Next, our abundance estimates assumed that capture and recapture events are random samples of the population.
-Angling methods often have a size selection bias, thereby limiting abundance estimates to a size range susceptible to angling and resulting in a conservative estimate of the total population.
+Angling methods often have a size selection bias, thereby limiting abundance estimates to a size range susceptible to angling, resulting in a conservative estimate of the total population.
 We also question the validity of our assumption of equal catchability of individual fish between sampling events.
-Although we recaptured two individual fish during multiple days within a survey (indicating that marked fish were still susceptible to angling), we suspect that marked and released fish may not have immediately mixed thoroughly back into the population.
-Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit “good” fishing locales.
+Two individual fish were recaptured during multiple days within a survey, indicating that marked fish were continually susceptible to angling.
+We additionally suspect that marked and released fish may not have immediately mixed thoroughly back into the population.
+Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit "good" fishing locales.
 Violations of the assumption of equal catchability between marked and unmarked fish may lead to an overestimate of abundance of Northern Pikeminnow in our study if marked fish have lower catchability.
 This bias will be proportional to the difference in catchability of marked and unmarked fish.
-For example, if marked fish were only 70% as likely to be caught as unmarked fish, the true abundance is closer to 70% of our abundance estimate.
-This would additionally account for marked and released fish being “hook shy” for a period of time.
+As an example, if marked fish were only 70% as likely to be caught as unmarked fish, the true abundance is closer to 70% of our abundance estimate.
+This would additionally account for marked and released fish being "hook shy" for a period of time.
 The adjusted delayed-mixing Schnabel estimator may account for these biases; however the magnitude of bias is unknown.
 
-The spring abundance estimate was smaller than fall abundance estimates due to the spring CPUE being approximately half of the mean fall CPUE.
-Although our approach assumed equal capture probabilities between the fall and spring surveys, we believe that higher spring flows may result in lower capture probability of Northern Pikeminnow.
+The spring abundance estimate was smaller than fall abundance estimates due to the spring CPUE being approximately half of the mean fall CPUE (Table \@ref(tab:summary-tbl)).
+Although our approach assumed equal capture probabilities between the fall and spring surveys, we believe that higher spring flows may result in lower capture probabilities of Northern Pikeminnow.
 Therefore, we believe our spring abundance estimate to be conservative.
@@
 Additionally, hatchery smolt releases (Chinook Salmon, steelhead, and Sockeye Salmon) are also prevalent in the mainstem Salmon River during the spring migration.
-To address the prevalence of other fish in the Northern Pikeminnow’s diet, we considered scenarios where DSR and NRR Chinook Salmon represent only 30%, 50%, and 65% of total fish prey consumed.
+To address the prevalence of other fish in the Northern Pikeminnow's diet, we considered scenarios where DSR and NRR Chinook Salmon represent only 30%, 50%, and 65% of total fish prey consumed.
 These values may be considered conservative during the peak emigration periods when Chinook Salmon are abundant and the Northern Pikeminnow diet shifts almost entirely to piscivory, presumably on juvenile Chinook Salmon [@Shively1996; @Poe1991].
-This is likely the case during fall months when DSR Chinook Salmon are the dominant prey species available in Deadwater Slough as 1) few to no hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent than Chinook Salmon.
-During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river, less than 50% Chinook Salmon in the Northern Pikeminnow’s diet may be more likely.
-Predation on juveniles from any of the ESA-listed salmonid species is detrimental to their recovery.
+This is likely the case during fall months when DSR Chinook Salmon are the dominant prey species available the Salmon River and Deadwater Slough as 1) few to no hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent than Chinook Salmon.
+During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river, less than 50% Chinook Salmon in the Northern Pikeminnow's diet may be more likely.
 Future work to quantify and identify juvenile salmonids in Deadwater Slough during their seasonal migration would be useful to validate the diet composition assumptions used in our model and to understand impacts to local populations.
@@
 For context, this is approximately `r round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 45416, 2)` times the average annual combined DSR and NRR emigration estimates for brood years 2011 - 2018 at a rotary screw trap located in the lower Lemhi River [@McClure2021; @Feeken2020; @Poole2019], the largest population in the Upper Salmon MPG.
-Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of fish prey are Chinook Salmon, we estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30]), big.mark = ",")` juvenile Chinook Salmon may be consumed which is `r round((total_impacts$juv_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 45416) * 100)`% of the total DSR and NRR emigration for brood years 2011 - 2018 from the Lemhi River.
+Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of fish prey are Chinook Salmon, we estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30]), big.mark = ",")` juvenile Chinook Salmon may be consumed which is `r round((total_impacts$juv_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 45416) * 100)`% of the total DSR and NRR emigration for brood years 2011 - 2018.
 Considering that diet scenarios used in our study are likely conservative assumptions, especially during the fall DSR emigration, the estimated number of Chinook Salmon consumed is substantial.
 
-Spawner abundance is perhaps the most important metric considered in determining a population’s viability and productivity.
-We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough is equivalent to `r round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` adults, annually, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean spawner escapement for the Upper Salmon MPG in 2017 – 2019 [@Kinzer2020].
+Spawner abundance is perhaps the most important metric considered in determining a population's viability and productivity.
+We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough is equivalent to `r round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` adults, annually, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean spawner escapement for the Upper Salmon MPG in 2017 -- 2019 [@Kinzer2020].
 In the most conservative scenario, we estimated juvenile consumption equivalent to `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])` adults, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% of recent natural-origin adult Chinook Salmon returns.
 In the least conservative scenarios, adult equivalents approached recent escapements for the entire Upper Salmon MPG.
-The sensitivity analysis used some simplifying assumptions.
-For example, it assumed no juvenile mortality between Deadwater Slough and Lower Granite Dam which is an unrealistic assumption.
-In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including both natural-origin juveniles and hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.
-Accordingly, we surmise that adult returns of steelhead and Sockeye Salmon are also affected by predation, including hatchery populations, which provide for recreational fishing opportunities in the Upper Salmon River.
+The sensitivity analysis used some simplifying assumptions; for example, it assumed no juvenile mortality between Deadwater Slough and Lower Granite Dam which is an unrealistic assumption.
+Nevertheless, we found converting juveniles consumed to “adult equivalents” a useful exercise to place results in context with a metric commonly used for recovery.
+
+In additional to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including both natural-origin juveniles and hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.
+Accordingly, we surmise that juvenile emigration and adult returns of steelhead and Sockeye Salmon are also affected by predation, including hatchery populations, which provide for recreational fishing opportunities in the Upper Salmon River.
 Consequently, reducing predation mortality in Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including other ESA-listed species.
 Because Deadwater Slough is part of the migratory pathway for multiple species and populations of emigrating salmonids, the positive impact could be greater than individual tributary rehabilitation actions which typically benefit a single population.
+Predation on juveniles from any of the ESA-listed salmonid species is likely detrimental to their recovery.
 
 Deadwater Slough is a favorable candidate for management or restoration actions to benefit local Chinook Salmon populations.
-Three potential management actions could reduce predation at Deadwater Slough: 1) removing or reducing the Dump Creek alluvial fan, 2) a local Northern Pikeminnow bounty program to encourage harvest in Deadwater Slough aimed at reducing the predator population size, and 3) adding structure or cover within Deadwater Slough to provide refuge for juvenile salmonids with the goal or reducing predation rates.
-Reducing or removing the Dump Creek alluvial fan has the benefit of restoring natural fluvial processes in the Salmon River that likely existed in the reach prior to the formation or increase in size of the alluvial fan related to human activities in upper Dump Creek.
-Restoring natural processes could reduce Northern Pikeminnow densities to levels similar to upstream and downstream reaches where higher survival and transition probabilities for juvenile salmon have been observed compared to Deadwater Slough [@Axel2015; @Ackerman2018; @Porter2019].
-Managers ought also to consider the feasibility and net benefit of restoring fluvial processes relative to potential losses for recreational fishing and bird watching opportunities (https://www.audobon.org/important-bird-areas/deadwater-slough).
+Three potential management actions could reduce predation at Deadwater Slough: 1) removing or reducing the Dump Creek alluvial fan, 2) a local Northern Pikeminnow bounty program to encourage harvest in Deadwater Slough aimed at reducing the predator population size, and 3) adding structure or cover within Deadwater Slough to provide refuge for juvenile salmonids to reduce predation rates.
+Reducing or removing the Dump Creek alluvial fan has the benefit of restoring natural fluvial processes in the Salmon River that likely existed in the reach prior to the formation or increase in size of the alluvial fan.
+Restoring natural processes could reduce Northern Pikeminnow densities to levels similar to upstream and downstream reaches where higher survival and transition probabilities for juvenile salmon have been observed relative to Deadwater Slough [@Axel2015; @Ackerman2018; @Porter2019].
+Managers ought also to consider the feasibility and net benefit of restoring fluvial processes relative to potential losses for recreational fishing and bird watching opportunities (<https://www.audobon.org/important-bird-areas/deadwater-slough>).
 
@@
 Each approach would require continual/annual effort and harvest to suppress the population size and as such, appropriate funding would be needed to support the program.
-Finally, a more passive approach could be to provide cover or structure within Deadwater Slough that provides refugia for juvenile salmonids during their emigration.
+Lastly, a more passive approach could be to provide cover or structure within Deadwater Slough that provides refugia for juvenile salmonids during their emigration.
 Cover should be appropriately sized to provide concealment to juvenile fishes while reducing access by larger fishes like Northern Pikeminnow.
@@
 
-Although the origins of Deadwater Slough are somewhat ambiguous [@Reichmuth1985; @USACE1986], it is likely that human activities in Dump Creek appear to have either exacerbated or contributed to its formation and in turn created favorable conditions for Northern Pikeminnow.
+Although the origins of Deadwater Slough are somewhat ambiguous [@Reichmuth1985; @USACE1986], it is likely that human activities in Dump Creek appear to have either exacerbated or contributed to its formation, and in turn, created favorable conditions for Northern Pikeminnow.
 Among the three surveys, our lowest population abundance estimate was greater than 10,000 Northern Pikeminnow that occupy Deadwater Slough, suggesting a remarkably high density given the size of the area.
@@
 We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than `r prettyNum(round(min(total_impacts$juv_Lci), -3), big.mark = ",")` juvenile Chinook Salmon (the lower endpoint of the confidence interval for the most conservative scenario); however, it is likely that this value is closer to `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50], -3), big.mark = ",")`, annually.
-The magnitude of predation by Northern Pikeminnow is equivalent to `r round((min(total_impacts$adult_N) / 558) * 100)`% and `r round((max(total_impacts$adult_N) / 558) * 100)`% of recent adult escapement.
-Deadwater Slough predates the ESA-listing of Chinook Salmon populations in the Upper Salmon MPG in the 1990s and is therefore unlikely to be the primary cause for the population’s decline.
+The magnitude of predation by Northern Pikeminnow is equivalent to `r round((min(total_impacts$adult_N) / 558) * 100)`% and `r round((max(total_impacts$adult_N) / 558) * 100)`% of recent adult escapements.
+Deadwater Slough predates the ESA-listing of Chinook Salmon populations in the Upper Salmon MPG in the 1990s and is therefore unlikely to be the primary cause for the population's decline.
 However, predation by Northern Pikeminnow in Deadwater Slough and elsewhere should be considered among limiting factors hindering recovery efforts of ESA-listed Chinook Salmon populations in the Upper Salmon MPG.
@@
 
-The authors extend much appreciation to the many volunteers that assisted with field efforts including collaborators from the Bureau of Reclamation, Idaho Governor’s Office of Species Conservation, Idaho Department of Fish and Game, and the Lemhi Regional Land Trust, among others.
+The authors extend much appreciation to the many volunteers that assisted with field efforts including collaborators from the Bureau of Reclamation, Idaho Governor's Office of Species Conservation, Idaho Department of Fish and Game, and the Lemhi Regional Land Trust, among others.
 Special thanks to Jared Barker, Brian Hamilton, Tulley Mackey, and Chelsea Welke for their help in the field.
@@
 Funding for this study was provided by the Bureau of Reclamation, Pacific Northwest Regional Office (contract No.
-140R1021F0018) with assistance from the Idaho Governor’s Office of Species Conservation.
+140R1021F0018) with assistance from the Idaho Governor's Office of Species Conservation.
 
@@
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 
-<div id="tables"></div>
+::: {#tables}
+:::
 
@@
 
-<div id="figures"></div>
+::: {#figures}
+:::
 

### 2022-04-07 README.Rmd
updating README

--- 2021-12-21 README.Rmd
@@
 # Please put your title here to include it in the file below.
-Title <- "Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations in the Upper Salmon River, Idaho"
-Authors <- "Ackerman, Michael W., Nicholas J. Porter, Kevin E. See, Michael S. Hall, Tara E. Blackman, and Richard A. Carmichael, and Michael W. Ackerman"
+Title <- "Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho, and Potential Impacts to Local Chinook Salmon Populations"
+Authors <- "Ackerman, Michael W., Nicholas J. Porter, Kevin E. See, Tara E. Blackman, Michael S. Hall, and Richard A. Carmichael"
 yr <- 2022
@@
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/BiomarkABS/DeadwaterPaper/main?urlpath=rstudio)
+[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/mackerman44/DeadwaterPaper/main?urlpath=rstudio)
 
@@
 > `r Authors` (`r yr`). _`r Title`_. `r Journal` <https://doi.org/xxx/xxx>
-
-Our pre-print is online here:
-
-> `r Authors` (`r yr`). _`r Title`_. `r Journal`, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
 

### 2022-04-07 README.Rmd
tiny updates to README

@@
 The simplest way to explore the text, code and data is to click on
-[binder](https://mybinder.org/v2/gh/BiomarkABS/DeadwaterPaper/main?urlpath=rstudio)
+[binder](https://mybinder.org/v2/gh/mackerman44/DeadwaterPaper/main?urlpath=rstudio)
 to open an instance of RStudio in your browser, which will have the

### 2022-04-07 analysis/paper/paper.Rmd
removed reference to zenodo DOI... for now
minor edits from NP

@@
 During our initial survey in 2019, Northern Pikeminnow were the most prevalent piscine predator in Deadwater Slough and consequently became our focal taxa.
-Our intent was to survey during the emigration of the 2018 Chinook Salmon brood year that occurred fall 2019 and spring 2020, however logistical constraints during the onset of the Covid-19 pandemic delayed the spring 2020 survey until 2021.
+Our intent was to survey during the emigration of the 2018 Chinook Salmon brood year that occurred fall 2019 and spring 2020, however logistical constraints during the onset of the COVID-19 pandemic delayed the spring 2020 survey until 2021.
 Over this period, an additional fall survey occurred during the 2020 DSR emigration.
@@
 Immediately following lavage, stomach contents of individuals were preserved with 99% isopropyl alcohol in whirl-paks to be analyzed in a controlled environment.
-For each stomach sample, total wet weight (grams) was recorded for all stomach contents, including all fish and non-fish items (e.g., macroinvertebrates, organic matter).
+For each stomach sample, total wet weight (grams) was recorded, including all fish and non-fish items (e.g., macroinvertebrates, organic matter).
 Fish and fish remnants were identified to the lowest taxonomic unit using diagnostic bones [@Hansel1988; @Frost2000] or were categorized as unknown.
 Approximately 5% of Northern Pikeminnow were euthanized for dissection (n = 75) after gastric lavage to validate the efficacy of the methodology.
+Additionally, contents of livewells were examined during fish processing to ensure food items were not regurgitated.
 
@@
 
-All data and code for the analyses presented here can be found in a GitHub repository at ***zenodo DOI***.
-
 # Results
@@
 
-The lengths of Northern Pikeminnow ranged from `r min(cmr_df$length, na.rm = T)` to `r max(cmr_df$length, na.rm = T)` mm TL with an average of `r round(mean(cmr_df$length, na.rm = T))` mm (Figure \@ref(fig:length-freq-p)).
+Lengths of Northern Pikeminnow ranged from `r min(cmr_df$length, na.rm = T)` to `r max(cmr_df$length, na.rm = T)` mm TL with an average of `r round(mean(cmr_df$length, na.rm = T))` mm (Figure \@ref(fig:length-freq-p)).
 The PSD for Northern Pikeminnow in Deadwater Slough across all three surveys was `r prettyNum(psd, digits = 1)`%.
@@
 Given yearly fish consumption estimates during the fall DSR and spring NRR periods, we produced a sensitivity analysis to calculate the consumption of juvenile Chinook Salmon where 30%, 50%, and 65% of total fish prey consumed are juvenile Chinook Salmon (Figure \@ref(fig:juv-con-p)).
-At the median value of 50%, we estimated that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon will be consumed.
+At the median value of 50%, we estimated that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon would be consumed.
 Using the median Granite-to-Granite SAR for Chinook Salmon in the Upper Salmon River, we estimated the "adult equivalents" of juveniles consumed to be `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) adults (Figure \@ref(fig:adult-imp-p)).
@@
 Conversely, the maximum depth at Deadwater Slough was approximately 6 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach.
-Our estimated densities of Northern Pikeminnow suggest that slow-water reaches outside of the reservoir complexes on the Snake and Columbia rivers may support exceptionally high predator densities, consistent with findings of @Harnish2014, @Gray2001, and @Zimmerman1999.
+Our estimated densities of Northern Pikeminnow suggest that slow-water reaches outside of reservoir complexes on the Snake and Columbia rivers may support exceptionally high predator densities, consistent with findings of @Harnish2014, @Gray2001, and @Zimmerman1999.
 Given the dearth of current data available throughout the Columbia River basin, it is unclear how Northern Pikeminnow abundance estimates in this study compare to elsewhere in the basin; additional information on contemporary piscine predator abundances may be needed [@Widener2021].
@@
 Emigration of marked individuals could reduce the marking fraction in the population leading to an upward bias in abundance estimates, but emigration rates would need to be substantial.
+We also captured two fish with marks from previous years, indicating seasonal residency of some fish.
 The multiple census estimators, which we report, are more robust to this assumption because the marking fraction is estimated daily.
@@
 This suggests that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to populations reported elsewhere in the Columbia River.
-Notably, the Idaho state catch-and-release record Northern Pikeminnow, measuring `r max(gastric_df$length, na.rm = T)` mm TL, was caught in Deadwater Slough during the fall 2020 survey.
+Notably, the Idaho state catch-and-release record Northern Pikeminnow, measuring `r max(gastric_df$length, na.rm = T)` mm TL, was caught in Deadwater Slough during the fall 2020 survey alluding to the favorable conditions for Northern Pikeminnow.
 
@@
 Lastly, a more passive approach could be to provide cover or structure within Deadwater Slough that provides refugia for juvenile salmonids during their emigration.
-Cover should be appropriately sized to provide concealment to juvenile fishes while reducing access by larger fishes like Northern Pikeminnow.
+Cover would need to be appropriately sized to provide concealment to juvenile fishes while reducing access by larger fishes like Northern Pikeminnow.
 Adding cover could be a cost-effective approach; however, its potential effectiveness is unclear.
@@
 ```{r juv-con-p, fig.cap = "Estimated number of juvenile Chinook Salmon consumed by Northern Pikeminnow from a sensitivity analysis of variable diet scenarios.
-The proportion of Chinook Salmon in the total fish prey consumed by Northern Pikminnow were evaluated for 30%, 50%, and 65% scenarios, similar to those reported by Zimmerman and Ward 1999 "}
+The proportion of Chinook Salmon in the total fish prey consumed by Northern Pikminnow were evaluated for 30%, 50%, and 65% scenarios, similar to those reported by Zimmerman and Ward 1999."}
 

### 2022-04-11 analysis/paper/paper.Rmd
incorporated MH edits

@@
 nerka* population [@Axel2015] and several populations of threatened Snake River steelhead.
-The slough is also inhabited by piscivorous predators, Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*, that are potential sources of mortality for rearing and migrating salmonids.
+The slough is also inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*, that are potential sources of mortality for rearing and migrating salmonids.
 
@@
 
-In additional to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including both natural-origin juveniles and hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough.
+In additional to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon (both natural-origin and hatchery releases) are also likely prey items for Northern Pikeminnow in Deadwater Slough.
 Accordingly, we surmise that juvenile emigration and adult returns of steelhead and Sockeye Salmon are also affected by predation, including hatchery populations, which provide for recreational fishing opportunities in the Upper Salmon River.
@@
 
-```{r map, out.width = "85%", fig.cap = "Map of the Deadwater Slough study area within the Upper Salmon River MPG. The Dump Creek alluvial fan is loacted at the downstream end of the study reach."}
+```{r map, out.width = "85%", fig.cap = "Map of the Deadwater Slough study area within the Upper Salmon River MPG. The Dump Creek alluvial fan is located at the downstream end of the study reach."}
 knitr::include_graphics("../figures/Study Area Map 5.png")
@@
 ```{r juv-con-p, fig.cap = "Estimated number of juvenile Chinook Salmon consumed by Northern Pikeminnow from a sensitivity analysis of variable diet scenarios.
-The proportion of Chinook Salmon in the total fish prey consumed by Northern Pikminnow were evaluated for 30%, 50%, and 65% scenarios, similar to those reported by Zimmerman and Ward 1999."}
+The proportion of Chinook Salmon in the total fish prey consumed by Northern Pikeminnow were evaluated for 30%, 50%, and 65% scenarios, similar to those reported by Zimmerman and Ward 1999."}
 

### 2022-04-11 analysis/paper/paper.Rmd
edits from KS

@@
 The Deadwater Slough is within a migration corridor for several Chinook Salmon populations as well as other Endangered Species Act-listings including steelhead and Sockeye Salmon.
-We estimated the abundance of Northern Pikeminnow in Deadwater Slough, an approximately 1.9 kilometer reach of the Salmon River, to be 19,499 in the fall and 10,352 in the spring, corresponding with the peak emigrations of juvenile Chinook Salmon.
+We estimated the abundance of Northern Pikeminnow in Deadwater Slough, an approximately 1.9 kilometer reach of the Salmon River, to be 19,499 in the fall and 10,352 in the spring, corresponding with the peak emigration windows of juvenile Chinook Salmon.
 Using these abundance values, we estimated Northern Pikeminnow consumption of juvenile Chinook Salmon.
 Assuming 60% of the Northern Pikeminnow diet is fish, of which 50% is juvenile Chinook Salmon, we estimated the Northern Pikeminnow population can consume 61,409 juvenile Chinook Salmon, annually.
-Using a sensitivity analysis, estimated Northern Pikeminnow predation on juvenile Chinook Salmon was equivalent to 377 (95% CI: 161 - 935) returning adults.
-Given the relatively small size of the Salmon River, the large densities and potential consumption rates of Northern Pikeminnow in Deadwater Slough suggest predation likely has consequential impact on Chinook Salmon recovery in the Upper Salmon River.
+After performing a sensitivity analysis, we estimated Northern Pikeminnow predation on juvenile Chinook Salmon was equivalent to 377 (95% CI: 161 - 935) returning adults.
+Given the relatively small size of the Salmon River, the large densities and potential consumption rates of Northern Pikeminnow in Deadwater Slough suggest predation likely has a consequential impact on Chinook Salmon recovery in the Upper Salmon River.
 
@@
 In this study, we estimated the abundance of Northern Pikeminnow in the Deadwater Slough and evaluated potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
-We hypothesize that high densities of piscivorous predators in the Deadwater Slough may explain the reduced survival observed for juvenile Chinook Salmon.
+We hypothesized that high densities of piscivorous predators in the Deadwater Slough might explain the reduced survival observed for juvenile Chinook Salmon.
 To test this, our objectives were four-fold:
@@
 
-During the fall DSR emigration, we estimated an average-size Northern Pikeminnow (`r round(avg_fall_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"]` g of fish to maintain their body size, assuming 60% of their diet consists of fish prey (Figure \@ref(fig:bio-p)).
+During the fall DSR emigration, we estimated an average-size Northern Pikeminnow (`r round(avg_fall_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"]` g of fish to maintain their body size, assuming 60% of their diet consisted of fish prey (Figure \@ref(fig:bio-p)).
 Fall consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f30_g"]` g for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f90_g"]` g for a diet of 90% fish.
-During the spring NRR emigration, we estimated an average-size Northern Pikeminnow (`r round(avg_spring_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f60_g"]` g of fish, assuming 60% of their diet is fish prey (Figure \@ref(fig:bio-p)).
+During the spring NRR emigration, we estimated an average-size Northern Pikeminnow (`r round(avg_spring_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f60_g"]` g of fish, assuming 60% of their diet was fish prey (Figure \@ref(fig:bio-p)).
 Spring consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f30_g"]` for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f90_g"]` for a diet of 90% fish.
 Daily consumption rates were higher early in the fall and late in the spring and corresponded with water temperatures above 8 degrees Celsius (Figure \@ref(fig:bio-p)).
-Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` in the fall and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` in the spring 2021, and a diet consisting of 60% fish, the Northern Pikeminnow population is estimated to consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
+Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` in the fall and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` in the spring 2021, and a diet consisting of 60% fish, the Northern Pikeminnow population was estimated to consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
 
@@
 
-Given yearly fish consumption estimates during the fall DSR and spring NRR periods, we produced a sensitivity analysis to calculate the consumption of juvenile Chinook Salmon where 30%, 50%, and 65% of total fish prey consumed are juvenile Chinook Salmon (Figure \@ref(fig:juv-con-p)).
+Given yearly fish consumption estimates during the fall DSR and spring NRR periods, we produced a sensitivity analysis to calculate the consumption of juvenile Chinook Salmon where 30%, 50%, and 65% of total fish prey consumed were juvenile Chinook Salmon (Figure \@ref(fig:juv-con-p)).
 At the median value of 50%, we estimated that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon would be consumed.
@@
 First, all models assumed a closed population; however, it is possible that immigration and emigration occurred.
-Nevertheless, our estimators still provide unbiased estimates of abundance assuming the immigration and emigration rates are equal between marked and unmarked fish.
-Emigration of marked individuals could reduce the marking fraction in the population leading to an upward bias in abundance estimates, but emigration rates would need to be substantial.
+Nevertheless, our estimators still provide unbiased estimates of abundance assuming the immigration and emigration rates were equal between marked and unmarked fish.
+Emigration of marked individuals could reduce the marking fraction in the population leading to an upward bias in abundance estimates, but emigration rates would have needed to be substantial.
 We also captured two fish with marks from previous years, indicating seasonal residency of some fish.
-The multiple census estimators, which we report, are more robust to this assumption because the marking fraction is estimated daily.
+The multiple census estimators, which we report, are more robust to this assumption because the marking fraction was estimated daily.
 We additionally have little reason to believe that emigration occurs during the two-week survey window owing to favorable habitat for Northern Pikeminnow in Deadwater Slough relative to adjacent reaches.
-Given the size of the sampling area and the short duration of our surveys, the closed population assumption is likely met.
-This assumption can likewise affect results if mortality occurs for some marked fish released back to the population.
-During field processing, no mortalities or injury post release was observed and fish injured prior to release were retained.
+Given the size of the sampling area and the short duration of our surveys, the closed population assumption was likely met.
+This assumption could likewise have affected results if mortality occurred for some marked fish released back to the population.
+During field processing, no mortalities or injuries post release were observed and fish injured prior to release were retained.
 Nevertheless, the potential for mortality amongst marked and released fish cannot be discounted.
@@
 Angling methods often have a size selection bias, thereby limiting abundance estimates to a size range susceptible to angling, resulting in a conservative estimate of the total population.
-We also question the validity of our assumption of equal catchability of individual fish between sampling events.
-Two individual fish were recaptured during multiple days within a survey, indicating that marked fish were continually susceptible to angling.
-We additionally suspect that marked and released fish may not have immediately mixed thoroughly back into the population.
+The estimators also assumed equal catchability of individual fish between sampling events; two fish were recaptured during multiple days within a survey, indicating that marked fish were continually susceptible to angling.
+It is also possible that marked and released fish may not have immediately mixed thoroughly back into the population.
 Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit "good" fishing locales.
-Violations of the assumption of equal catchability between marked and unmarked fish may lead to an overestimate of abundance of Northern Pikeminnow in our study if marked fish have lower catchability.
+Violations of the assumption of equal catchability between marked and unmarked fish may have led to an overestimate of abundance of Northern Pikeminnow in our study if marked fish have lower catchability.
 This bias will be proportional to the difference in catchability of marked and unmarked fish.
@@
 Although the origins of Deadwater Slough are somewhat ambiguous [@Reichmuth1985; @USACE1986], it is likely that human activities in Dump Creek appear to have either exacerbated or contributed to its formation, and in turn, created favorable conditions for Northern Pikeminnow.
-Among the three surveys, our lowest population abundance estimate was greater than 10,000 Northern Pikeminnow that occupy Deadwater Slough, suggesting a remarkably high density given the size of the area.
+Among the three surveys, our lowest population abundance estimate was greater than 10,000 Northern Pikeminnow occupying Deadwater Slough, suggesting a remarkably high density given the size of the area.
 The slow water velocity and lack of cover for fish also create conditions where juvenile salmonids that are rearing in or emigrating through Deadwater Slough, including Chinook Salmon, are susceptible to predation.
@@
   rename(Season = season,
-         `% Fish in Diet` = name,
-         `g Consumed` = cumm_g) %>%
+         `Percent Fish in Diet` = name,
+         `g Fish Consumed` = cumm_g) %>%
   mutate(`Date Range` = recode(`Date Range`,
@@
         format.args = list(big.mark = ","),
-        caption = "Summary of gastric lavage results including counts of the total number of Northern Pikeminnow captured, gastric lavaged, individuals with stomach contents, and individuals with fish prey contents.") %>%
+        caption = "Counts of the total number of Northern Pikeminnow captured, gastric lavaged, individuals with stomach contents, and individuals with fish prey contents.") %>%
   kable_styling()
@@
   ggplot(aes(ymd, temperature_c)) +
-  geom_line(color = "navyblue",
+  geom_line(color = "black",
             size = 1.15) +
@@
 
-```{r juv-con-p, fig.cap = "Estimated number of juvenile Chinook Salmon consumed by Northern Pikeminnow from a sensitivity analysis of variable diet scenarios.
+```{r juv-con-p, fig.cap = "Estimated number of juvenile Chinook Salmon consumed by Northern Pikeminnow from a sensitivity analysis of diet scenarios made up of variable percentages of fish.
 The proportion of Chinook Salmon in the total fish prey consumed by Northern Pikeminnow were evaluated for 30%, 50%, and 65% scenarios, similar to those reported by Zimmerman and Ward 1999."}
@@
   scale_color_manual(name = "Percent Chinook", 
-                     values=c("lightsteelblue","steelblue3","navyblue")) +
+                     values=c("gray70", "gray35","gray0")) +
   theme(axis.text.y = element_text(size = 11,

### 2022-04-12 analysis/paper/paper.Rmd
incorporated KS and RC edits and re-knit paper.docx

@@
 Second, reservoirs and downstream tailraces associated with dams on the Snake and Columbia rivers create favorable slow-water habitat for predatory fishes known to consume juvenile salmonids, such as the Northern Pikeminnow *Ptychocheilus oregonsis*.
-Northern Pikeminnow tolerate and thrive in relatively warm, slow-water habitats [@Wydoski2003] and consequently have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999].
+Northern Pikeminnow tolerate and thrive in relatively warm, slow-water habitats [@Wydoski2003] and thus have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999].
 Northern Pikeminnow are estimated to consume \~8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].
@@
 Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency [@Dodson2013].
-Deadwater Slough represents an important habitat for Chinook Salmon in the Upper Salmon River as it supports rearing and migration of juveniles from all populations including fall DSR and spring NRR migrants.
+Deadwater Slough represents an important habitat for Chinook Salmon in the Upper Salmon River as it supports rearing and migration of juveniles from all upstream populations including fall DSR and spring NRR migrants.
 
@@
 This would additionally account for marked and released fish being "hook shy" for a period of time.
-The adjusted delayed-mixing Schnabel estimator may account for these biases; however the magnitude of bias is unknown.
+The adjusted delayed-mixing Schnabel estimator may account for these biases; however, the magnitude of bias is unknown.
 
@@
 These values may be considered conservative during the peak emigration periods when Chinook Salmon are abundant and the Northern Pikeminnow diet shifts almost entirely to piscivory, presumably on juvenile Chinook Salmon [@Shively1996; @Poe1991].
-This is likely the case during fall months when DSR Chinook Salmon are the dominant prey species available the Salmon River and Deadwater Slough as 1) few to no hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent than Chinook Salmon.
+This is likely the case during fall months when DSR Chinook Salmon are the dominant prey species available within the Salmon River and Deadwater Slough as 1) few to no hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent than Chinook Salmon.
 During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river, less than 50% Chinook Salmon in the Northern Pikeminnow's diet may be more likely.
@@
 
-In additional to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon (both natural-origin and hatchery releases) are also likely prey items for Northern Pikeminnow in Deadwater Slough.
+In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon (both natural-origin and hatchery releases) are also likely prey items for Northern Pikeminnow in Deadwater Slough.
 Accordingly, we surmise that juvenile emigration and adult returns of steelhead and Sockeye Salmon are also affected by predation, including hatchery populations, which provide for recreational fishing opportunities in the Upper Salmon River.
@@
 Each approach would require continual/annual effort and harvest to suppress the population size and as such, appropriate funding would be needed to support the program.
-Lastly, a more passive approach could be to provide cover or structure within Deadwater Slough that provides refugia for juvenile salmonids during their emigration.
+Lastly, a more passive approach could be to improve cover or structure within Deadwater Slough that provides refugia for juvenile salmonids during their emigration.
 Cover would need to be appropriately sized to provide concealment to juvenile fishes while reducing access by larger fishes like Northern Pikeminnow.
@@
 The authors extend much appreciation to the many volunteers that assisted with field efforts including collaborators from the Bureau of Reclamation, Idaho Governor's Office of Species Conservation, Idaho Department of Fish and Game, and the Lemhi Regional Land Trust, among others.
-Special thanks to Jared Barker, Brian Hamilton, Tulley Mackey, and Chelsea Welke for their help in the field.
+Special thanks to Braden Lott, Jared Barker, Brian Hamilton, Tulley Mackey, and Chelsea Welke for their help in the field.
 We further appreciate the administrative support and guidance from staff at Inter-Fluve.

### 2022-04-12 analysis/paper/paper.Rmd
crossing fingers, final edits before preparing for submission

@@
 After performing a sensitivity analysis, we estimated Northern Pikeminnow predation on juvenile Chinook Salmon was equivalent to 377 (95% CI: 161 - 935) returning adults.
-Given the relatively small size of the Salmon River, the large densities and potential consumption rates of Northern Pikeminnow in Deadwater Slough suggest predation likely has a consequential impact on Chinook Salmon recovery in the Upper Salmon River.
+Given the relatively small size of the Salmon River, the high densities and potential consumption rates of Northern Pikeminnow in Deadwater Slough suggest predation likely has a consequential impact on Chinook Salmon recovery in the Upper Salmon River.
 
@@
 Northern Pikeminnow tolerate and thrive in relatively warm, slow-water habitats [@Wydoski2003] and thus have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999].
-Northern Pikeminnow are estimated to consume \~8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].
+Northern Pikeminnow are estimated to consume \~8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996].
 
@@
 The Deadwater Slough is within a section of the Salmon River containing historically important overwinter rearing habitat for juvenile Chinook Salmon *O.
-tshawytscha* and is part of the migratory pathway for upstream Chinook Salmon, the endangered Snake River Sockeye Salmon *O.
-nerka* population [@Axel2015] and several populations of threatened Snake River steelhead.
+tshawytscha* and is part of the migratory pathway for upstream Chinook Salmon populations, the endangered Snake River Sockeye Salmon *O.
+nerka* population [@Axel2015], and several populations of threatened Snake River steelhead.
 The slough is also inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*, that are potential sources of mortality for rearing and migrating salmonids.
@@
 
-In this study, we estimated the abundance of Northern Pikeminnow in the Deadwater Slough and evaluated potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
+In this study, we estimated the abundance of Northern Pikeminnow in Deadwater Slough and evaluated potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon.
 We hypothesized that high densities of piscivorous predators in the Deadwater Slough might explain the reduced survival observed for juvenile Chinook Salmon.
@@
 
-1.  Estimate the abundance of Northern Pikeminnow in the Deadwater Slough during the peaks of fall DSR and spring NRR juvenile emigrations;
+1.  Estimate the abundance of Northern Pikeminnow in Deadwater Slough during the peaks of fall DSR and spring NRR juvenile emigrations;
 2.  Document predation on juvenile Chinook Salmon during the DSR and NRR emigration periods using gastric lavage;
 3.  Use an established bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population on DSR and NRR Chinook Salmon emigrants at Deadwater Slough;
-4.  Estimate how consumption of juvenile Chinook Salmon emigrants at Deadwater Slough by Northern Pikeminnow may impact adult returns to the Upper Salmon River.
+4.  Estimate how consumption of juvenile Chinook Salmon emigrants at Deadwater Slough by Northern Pikeminnow might impact adult returns to the Upper Salmon River.
 
@@
 Methods included raft electrofishing, fyke netting, snorkeling, and angling.
-After evaluating all methods, angling was the most effective method for capturing Northern Pikeminnow while also minimizing potential impacts to ESA-listed adult steelhead that were present during our surveys.
+After evaluating all methods, angling was the most effective method for capturing Northern Pikeminnow while also minimizing potential impacts to Endangered Species Act (ESA)-listed adult steelhead that were present during our surveys.
 Less effective methods were abandoned in subsequent years and the following analyses will focus on fish captured by angling, unless otherwise noted.
@@
 
-Surveys conducted in spring 2021 did not follow a mark-recapture study design so Northern Pikeminnow abundance was estimated as the ratio of mean total CPUE in the fall surveys to the total CPUE in the spring survey, then multiplied that ratio by the mean abundance from fall surveys.
+Surveys conducted in spring 2021 did not follow a mark-recapture study design so Northern Pikeminnow abundance was estimated as the ratio of mean total CPUE in the fall surveys to the total CPUE in the spring survey, then multiplied that ratio by the mean abundance from fall surveys:
 
@@
 Approximately 5% of Northern Pikeminnow were euthanized for dissection (n = 75) after gastric lavage to validate the efficacy of the methodology.
-Additionally, contents of livewells were examined during fish processing to ensure food items were not regurgitated.
+Contents of livewells were examined during fish processing to ensure food items were not regurgitated.
 
@@
 Prey energy densities were fixed at 3,000 J/g for invertebrates and 21,500 J/g for juvenile Chinook Salmon [@Moss2016].
-Because we were unable to differentiate juvenile Chinook Salmon from other fish prey, we assume all fish prey have the same energy densities as juvenile Chinook Salmon.
+Because we were unable to differentiate juvenile Chinook Salmon from other fish prey in model runs, we assume all fish prey have the same energy densities as juvenile Chinook Salmon.
 The average TL of Northern Pikeminnow caught in Deadwater Slough during our study was converted to fork length (FL) and then to weight (grams) using a weight-length formula from @Parker1995.
@@
 A model run was conducted for each combination of diet scenario and season, resulting in fourteen estimates of the total grams of fish consumed by an individual Northern Pikeminnow.
-To estimate the total biomass of fish consumed by Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated Northern Pikeminnow population sizes during fall and spring.
+To estimate the total biomass of fish consumed by Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated Northern Pikeminnow population sizes during fall DSR and spring NRR emigrations.
 
@@
 Although the primary impact to Chinook Salmon in Deadwater Slough by Northern Pikeminnow is juvenile predation, the common metric to evaluate salmon recovery is adult returns.
-We quantified the potential impact of Northern Pikeminnow predation on adult returns by estimating the number of equivalent adults expected to return to Lower Granite Dam if predation by Pikeminnow was eliminated in the Deadwater Slough.
+We quantified the potential impact of Northern Pikeminnow predation on adult returns by estimating the number of equivalent adults expected to return to Lower Granite Dam if predation by Pikeminnow was eliminated in Deadwater Slough.
 To accomplish this, we multiplied the estimated total juvenile Chinook Salmon consumed by the median Granite-to-Granite smolt-to-adult return rate (SAR) of 0.00614 (SD = 0.00051) from @McCann2019 for Chinook Salmon in the Upper Salmon River.
@@
 Mark-recapture abundance estimates of Northern Pikeminnow in Deadwater Slough ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` in fall 2019 and from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` in fall 2020 (Table \@ref(tab:abund-tab)).
-We estimated larger populations of Northern Pikeminnow using the unadjusted multiple-census estimator compared to the single census and adjusted delayed-mixing multiple census estimators (Figure \@ref(fig:abund-fig)).
-Our sampling design most closely matched a multiple-census estimator; therefore, the Schnabel estimates were considered most appropriate.
+We estimated larger populations of Northern Pikeminnow using the unadjusted multiple census estimator compared to the single census and adjusted delayed-mixing multiple census estimators (Figure \@ref(fig:abund-fig)).
+Our sampling design most closely matched a multiple census estimator; therefore, the Schnabel estimates were considered most appropriate.
 Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r prettyNum(round(mean(all_est$Lci[all_est$model == "Schnabel"])), big.mark = ",")` - `r prettyNum(round(mean(all_est$Uci[all_est$model == "Schnabel"])), big.mark = ",")`) using the unadjusted Schabel estimator.
@@
 For spring 2021, we estimated `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough.
-Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm and areal densities (fish/100 m$^2$) of `r round((mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_m2) * 100, 1)` and `r round((spring_abund$N_S / ddw_m2) * 100, 1)` for fall and spring, respectively.
+Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm and areal densities (fish per 100 m$^2$) of `r round((mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_m2) * 100, 1)` and `r round((spring_abund$N_S / ddw_m2) * 100, 1)` for fall and spring, respectively.
 
@@
 We found that gastric lavage successfully removed food items in all dissected individuals, indicating the method was effective.
-Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` of those.
+Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")`<!--(`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%)--> stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` of those.
 Northern Pikeminnow captured during spring 2021 had a higher proportion of fish content in their diet compared to individuals captured in the fall surveys (Table \@ref(tab:gastric-tab)).
@@
 Spring consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f30_g"]` for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f90_g"]` for a diet of 90% fish.
-Daily consumption rates were higher early in the fall and late in the spring and corresponded with water temperatures above 8 degrees Celsius (Figure \@ref(fig:bio-p)).
+Daily consumption rates were higher early in the fall and late in the spring and corresponded with higher water temperatures, especially above 8 degrees Celsius (Figure \@ref(fig:bio-p)).
 Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` in the fall and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` in the spring 2021, and a diet consisting of 60% fish, the Northern Pikeminnow population was estimated to consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.
@@
 All mark-recapture abundance estimators suggest a large population of Northern Pikeminnow occupies Deadwater Slough.
-The mean population size of Northern Pikeminnow was estimated to be `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` during the fall DSR emigration and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` during the spring NRR emigration even after accounting for the potential of delayed mixing of marked fish.
-Those estimates translate to linear densities (fish/rkm) that are `r round((mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms) / john_day_dens)` and `r round((spring_abund$N_S / ddw_rkms) / john_day_dens)` times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r round((mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms) / low_columbia_fish_km)` and `r round((spring_abund$N_S / ddw_rkms) / low_columbia_fish_km)` times greater than projections for the lower Columbia River [@Beamesderfer1996].
-@Beamesderfer1991 acknowledged that their methods were unsuitable for sampling offshore, noting that water velocity, depth, and irregular bottom contours, and barge traffic made sampling offshore ineffective; therefore it is possible that their estimates may have been low.
+The mean population size of Northern Pikeminnow was estimated to be `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` during the fall DSR emigration and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` during the spring NRR emigration even after accounting for the potential of delayed mixing of marked and released fish.
+Those estimates translate to linear densities (fish per rkm) that are `r round((mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms) / john_day_dens)` and `r round((spring_abund$N_S / ddw_rkms) / john_day_dens)` times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r round((mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms) / low_columbia_fish_km)` and `r round((spring_abund$N_S / ddw_rkms) / low_columbia_fish_km)` times greater than projections for the lower Columbia River [@Beamesderfer1996].
+@Beamesderfer1991 acknowledged that their methods were unsuitable for sampling offshore, noting that water velocity, depth, and irregular bottom contours, and barge traffic made sampling offshore ineffective; therefore, it is possible that their estimates may have been low.
 Conversely, the maximum depth at Deadwater Slough was approximately 6 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach.
@@
 Bounty programs could be conducted year-round or seasonally to coincide with peak juvenile outmigration(s).
-Each approach would require continual/annual effort and harvest to suppress the population size and as such, appropriate funding would be needed to support the program.
+Each approach would require continual/annual effort and harvest to suppress the population size and as such, appropriate funding would be needed to support such a program.
 Lastly, a more passive approach could be to improve cover or structure within Deadwater Slough that provides refugia for juvenile salmonids during their emigration.
@@
 
-Although the origins of Deadwater Slough are somewhat ambiguous [@Reichmuth1985; @USACE1986], it is likely that human activities in Dump Creek appear to have either exacerbated or contributed to its formation, and in turn, created favorable conditions for Northern Pikeminnow.
+Although the origins of Deadwater Slough are somewhat ambiguous [@Reichmuth1985; @USACE1986], it appears likely that human activities in Dump Creek have either exacerbated or contributed to its formation, and in turn, created favorable conditions for Northern Pikeminnow.
 Among the three surveys, our lowest population abundance estimate was greater than 10,000 Northern Pikeminnow occupying Deadwater Slough, suggesting a remarkably high density given the size of the area.
@@
         format.args = list(big.mark = ","),
-        caption = "Summary of Northern Pikeminnow angling survey data for the single-census mark-recapture estimator.
+        caption = "Summary of Northern Pikeminnow angling survey data for the single census mark-recapture estimator.
 M is the total number of fish caught in the first sample (week), marked and returned to the population; n is the total number of fish caught in the second sample (week), and m is the number of previously marked fish caught in the second week.") %>%
@@
         caption = "Point estimates, standard errors, and 95% confidence intervals for each of the mark-recapture estimators and for the fall sampling efforts.
-Standard errors are not available for the multiple-census estimators.") %>%
+Standard errors are not available for the multiple census estimators.") %>%
   kable_styling()

### 2022-04-13 analysis/paper/paper.Rmd
minor edits in .Rmd that are in submission

@@
 ---
-title: "Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho, and Potential Impacts to Local Chinook Salmon Populations"
+title: "Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho, and Potential Impacts to Local Chinook Salmon"
 author:
@@
   Northern Pikeminnow; Chinook Salmon; predation; mark-recapture; bioenergetics
+runninghead: |
+  Northern Pikeminnow in Deadwater Slough
 ---
+
+Running Head: `r rmarkdown::metadata$runninghead`
 
@@
 
+## Impact Statement
+
+A large population of Northern Pikeminnow in a slow-water reach of the Salmon River, Idaho can potentially consume large numbers of juvenile Chinook Salmon, hindering ongoing recovery efforts.
+
 ## Abstract
@@
 Given the relatively small size of the Salmon River, the high densities and potential consumption rates of Northern Pikeminnow in Deadwater Slough suggest predation likely has a consequential impact on Chinook Salmon recovery in the Upper Salmon River.
+
+\newpage
 
@@
   rename(Season = season,
-         `Percent Fish in Diet` = name,
-         `g Fish Consumed` = cumm_g) %>%
+         `Fish in Diet (%)` = name,
+         `Fish Consumed (g)` = cumm_g) %>%
   mutate(`Date Range` = recode(`Date Range`,
@@
         format.args = list(big.mark = ","),
-        caption = "Counts of the total number of Northern Pikeminnow captured, gastric lavaged, individuals with stomach contents, and individuals with fish prey contents.") %>%
+        caption = "Counts of the total number of Northern Pikeminnow captured and gastric lavaged.
+The number of individuals with stomach contents, including those with confirmed fish contents are shown.") %>%
   kable_styling()
@@
 
-# Figures
+# Figure Captions
 
@@
 
-```{r adult-imp-p, fig.cap = "Estimated reduction in the annual number of returning adult Chinook Salmon as a result of Northern Pikeminnow predation. Error bars indicate 95% confidence intervals."}
+```{r adult-imp-p, fig.cap = "Estimated “adult equivalents” of the total juvenile Chinook Salmon consumed by Northern Pikeminnow in Deadwater Slough. Error bars indicate 95% confidence intervals."}
 #adult impacts
@@
 
-# Supplementary Materials
+## Supplement - Avian Predation in Deadwater Slough
 

### 2022-04-13 analysis/paper/paper.Rmd
added zenodo DOI
merging final edits into .Rmd

@@
 
-A large population of Northern Pikeminnow in a slow-water reach of the Salmon River, Idaho can potentially consume large numbers of juvenile Chinook Salmon, hindering ongoing recovery efforts.
+A large population of Northern Pikeminnow occupying a slow-water reach of the Salmon River, Idaho can potentially consume large numbers of juvenile Chinook Salmon, hindering ongoing recovery efforts.
 
@@
 
+All data and code for the analyses presented here can be found in a GitHub repository release (https://doi.org/10.5281/zenodo.6458124).
+
 # Results
@@
   kable(booktabs = T,
-        caption = "Summary of Northern Pikeminnow angling survey data.
+        caption = "Summary of Northern Pikeminnow captures from angling surveys in Deadwater Slough.
 Fall surveys included mark and recapture weeks for the single census estimate.
-Spring surveys included only a single capture week to estimate catch-per-unit-effort.") %>%
+Spring surveys included only a single capture week to estimate CPUE.") %>%
   kable_styling()
@@
   kable(booktabs = T,
-        caption = "Bioenergetics model runs to estimate the grams of fish consumed during the fall DSR and spring NRR Chinook Salmon emigrations.
-The date ranges for each model run, the assumed proportion of fish in the diet, and the total grams of fish consumed by an individual Northern Pikeminnow to maintain its body weight for a given season are shown.
+        caption = "Estimated fish consumed by an individual Northern Pikeminnow to maintain its body weight during the fall DSR and spring NRR Chinook Salmon emigrations.
+Bioenergetic model results for each season include seven model runs that assume different percentages of fish in the diet.
 Northern Pikeminnow start and end weights were 598.7 and 430.7 g for the fall and spring seasons, respectively.") %>%
@@
         format.args = list(big.mark = ","),
-        caption = "Point estimates, standard errors, and 95% confidence intervals for each of the mark-recapture estimators and for the fall sampling efforts.
+        caption = "Estimated Northern Pikeminnow in Deadwater Slough and the associated standard errors and 95% confidence intervals from mark-recapture estimators calculated for the fall surveys.
 Standard errors are not available for the multiple census estimators.") %>%
@@
         format.args = list(big.mark = ","),
-        caption = "Counts of the total number of Northern Pikeminnow captured and gastric lavaged.
-The number of individuals with stomach contents, including those with confirmed fish contents are shown.") %>%
+        caption = "Total number of Northern Pikeminnow gastric lavaged and the number of those individuals with contents (including fish) and with contents that include fish. ") %>%
   kable_styling()
@@
 ```{r bio-p, fig.cap = "The cumulative fish prey consumed (g) by an individual Northern Pikeminnow during the peak Chinook Salmon emigrations for fall DSR (top left) and spring NRR (top right) and corresponding daily mean water temperatures (bottom panels).
-The black line is the median Northern Pikeminnow diet scenario, consisting of 60% fish and the gray area shows the range between diet scenarios of 30% and 90% fish prey.
-Daily mean water temperatures were summarized from six years of data available from the Shoup gage station.
-"}
+The black line shows the consumption by Northern Pikeminnow when the diet consists of 60% fish and the gray area shows the range between all model runs in the series, from 30% to 90% fish prey.
+Daily mean water temperatures were summarized from six years of data available from the Shoup gage station."}
 # spring_results$season = "Spring"
@@
 ```{r juv-con-p, fig.cap = "Estimated number of juvenile Chinook Salmon consumed by Northern Pikeminnow from a sensitivity analysis of diet scenarios made up of variable percentages of fish.
-The proportion of Chinook Salmon in the total fish prey consumed by Northern Pikeminnow were evaluated for 30%, 50%, and 65% scenarios, similar to those reported by Zimmerman and Ward 1999."}
+Diet scenarios include 30%, 50%, and 65% Chinook Salmon in the total fish prey consumed by Northern Pikeminnow, similar to those reported by Zimmerman and Ward 1999."}
 
@@
 The Deadwater Slough is recognized as an important bird watching and nesting area due to the associated upstream braided channels and riparian and backwater habitats [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#).
-Several piscivorous bird species have been documented using Deadwater Slough including the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron *Ardea herodias*, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021].
+Several piscivorous bird species have been documented using Deadwater Slough including the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron *Ardea herodias*, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* (eBird 2021).
 During the initial sampling effort in fall 2019, a two-person crew walked the entire reach including surrounding and upstream areas scanning for passive integrated transponder (PIT) tags that are implanted into thousands of natural- and hatchery-origin salmonids in the Upper Salmon River, annually.
@@
 Two of the Chinook Salmon and one steelhead were natural-origin, the remaining were of hatchery-origin.
-Avian predation contributes a major component of the total mortality for yearling Chinook Salmon in some locations in the lower Snake and Columbia rivers, particularly at hydroelectric dams and within reservoirs [@Evans2012; -@Evans2016]; however, we did not observe large colonies of piscivorous birds within the study area.
+Avian predation contributes a major component of the total mortality for yearling Chinook Salmon in some locations in the lower Snake and Columbia rivers, particularly at hydroelectric dams and within reservoirs (Evans et al.
+2012; 2016); however, we did not observe large colonies of piscivorous birds within the study area.
 Although we did observe Belted Kingfishers, mergansers, and Great Blue Herons at Deadwater Slough; Great Blue Herons were especially prevalent in the anastomizing reach immediately upriver of the slack water reach.
-Additionally, there is documentation of individual Double-crested Cormorants [@eBird2021] at the Deadwater Slough; however, the site is not within their breeding range, rather, it is part of a migration corridor.
+Additionally, there is documentation of individual Double-crested Cormorants (eBird 2021) at the Deadwater Slough; however, the site is not within their breeding range, rather, it is part of a migration corridor.
 Given the current avian species known to occupy Deadwater Slough, it is unlikely that avian predation on juvenile salmonids is comparable to elsewhere in the Columbia River basin with large piscivorous bird colonies.

### 2022-06-12 analysis/paper/paper.Rmd
minor edits from TAFS review

@@
 Upstream habitats containing slower water velocities and other attributes that support piscivorous predators may overlap with essential habitat for some salmonid species and life stages.
-One such example is Deadwater Slough, an approximately `r round(ddw_rkms, 1)` km long reach of unnaturally slow and deep water in the Salmon River, Idaho.
+One such example is Deadwater Slough, an approximately `r round(ddw_rkms, 1)` km long reach of uncharacteristically slow and deep water in the Salmon River, Idaho.
 The Deadwater Slough is within a section of the Salmon River containing historically important overwinter rearing habitat for juvenile Chinook Salmon *O.
@@
 Although the origin and timing of the alluvial fan and Deadwater Slough is somewhat ambiguous [@Reichmuth1985; @USACE1986], some believe they coincided with the failure of a small mining reservoir in the Dump Creek drainage around 1897 that deposited substantial amounts of sediment at its confluence with the Salmon River [@Emerson1973].
-The slough is currently a slow, deep section in the river, spanning approximately `r round(ddw_hectares)` hectares, averaging `r round(ddw_avg_width_m)` m wide, and has characteristics resembling a small reservoir.
+The slough is currently a slow, deep (approximately 6 m maximum depth) section in the river, spanning `r round(ddw_hectares)` hectares, averaging `r round(ddw_avg_width_m)` m wide, and has characteristics resembling a small reservoir.
 
@@
 Over this period, an additional fall survey occurred during the 2020 DSR emigration.
-Fall surveys were constrained to two weeks to minimize Northern Pikeminnow immigration and emigration.
+Fall surveys were constrained to two weeks to minimize the effect of Northern Pikeminnow immigration and emigration on the population size estimate.
 
@@
 Our sampling design most closely matched a multiple census estimator; therefore, the Schnabel estimates were considered most appropriate.
-Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r prettyNum(round(mean(all_est$Lci[all_est$model == "Schnabel"])), big.mark = ",")` - `r prettyNum(round(mean(all_est$Uci[all_est$model == "Schnabel"])), big.mark = ",")`) using the unadjusted Schabel estimator.
+Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events (2019 and 2020) was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r prettyNum(round(mean(all_est$Lci[all_est$model == "Schnabel"])), big.mark = ",")` - `r prettyNum(round(mean(all_est$Uci[all_est$model == "Schnabel"])), big.mark = ",")`) using the unadjusted Schabel estimator.
 Using the delayed-mixing Schnabel estimator the mean fall abundance estimate was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r prettyNum(mean(all_est$Lci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` - `r prettyNum(mean(all_est$Uci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")`).
@@
 
-During the fall DSR emigration, we estimated an average-size Northern Pikeminnow (`r round(avg_fall_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"]` g of fish to maintain their body size, assuming 60% of their diet consisted of fish prey (Figure \@ref(fig:bio-p)).
+During the fall DSR emigration, we estimated an average-size Northern Pikeminnow (`r round(avg_fall_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"]` g of fish across the season to maintain their body size, assuming 60% of their diet consisted of fish prey (Figure \@ref(fig:bio-p)).
 Fall consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f30_g"]` g for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f90_g"]` g for a diet of 90% fish.
@@
 We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than `r prettyNum(round(min(total_impacts$juv_Lci), -3), big.mark = ",")` juvenile Chinook Salmon (the lower endpoint of the confidence interval for the most conservative scenario); however, it is likely that this value is closer to `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50], -3), big.mark = ",")`, annually.
-The magnitude of predation by Northern Pikeminnow is equivalent to `r round((min(total_impacts$adult_N) / 558) * 100)`% and `r round((max(total_impacts$adult_N) / 558) * 100)`% of recent adult escapements.
+The magnitude of predation by Northern Pikeminnow on adult returns is equivalent to `r round((min(total_impacts$adult_N) / 558) * 100)`% and `r round((max(total_impacts$adult_N) / 558) * 100)`% of recent adult escapements.
 Deadwater Slough predates the ESA-listing of Chinook Salmon populations in the Upper Salmon MPG in the 1990s and is therefore unlikely to be the primary cause for the population's decline.
@@
 Bioenergetic model results for each season include seven model runs that assume different percentages of fish in the diet.
-Northern Pikeminnow start and end weights were 598.7 and 430.7 g for the fall and spring seasons, respectively.") %>%
+Northern Pikeminnow start and end weights were 598.7 g for the fall season and 430.7 g for the spring season.") %>%
   kable_styling()
